### 2016-11-25 analyze_monthly_data.Rmd
initial analysis of maragra data

---
title: "Maragra-CISM status report"
date: ""
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setOptions, message=FALSE, echo = FALSE}
library(googleVis)
op <- options(gvis.plot.tag='chart')
```

```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(brew)
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

**Joe Brew, Laia Cirera, Kizito Gondo, Elton Dorkin, Elisa Sicuri**

`r format(Sys.Date(), '%B %d, %Y')`

```{r}
# Packages
library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)
library(ggthemes)
library(RColorBrewer)
library(pander)

# Read data
df <- read_csv(file = 'data/maragra_monthly_data.csv')

# Create date column
df$date <- as.Date(paste0(df$year, '-', df$month, '-', 15))

# Create a percent postiive column
df$percent_positive <- df$positive / df$tested * 100

# Arrange by date
df <- df %>% arrange(date)

# Get rid of any unknowns
df <- df %>% filter(!is.na(tested))
```

# Status

The formalities and agreements pertaining to the transfer of data from Ilovo Maragra to CISM are more or less finished. With Dr. Gondo's help, Joe Brew has retrieved aggregated monthly summarical data pertaining to malaria cases among workers. More time will be needed to record person-specific data, since it is all handwritten and will need to be digitized.

# Next steps

Beginning next week (November 28, 2016), the digitization of all positive cases will be carried out. For this, the following needs to take place.

- Joe Brew will request RRHH named from Ilovo-Maragra.
- Joe Brew will build a name-searcher app, so as to identify the worker ID number at the moment of data entry.
- Laia Cirera will request the enlistment of several data entry technicians.

Full data collection will be carried out over the coming weeks. All digitized data, in addition to being stored by the CISM for this project, will also be given to Ilovo-Maragra as an "in-kind" service for their willingness to have their data examined.

# Data overview

## Cases over time

### All workers

If we combine all workers, we can examine the total incidence of malaria since 2011.

```{r}
x <- 
  df %>%
  group_by(date) %>%
  summarise(p = sum(positive),
            pp = sum(positive) / sum(tested) * 100,
            tested = sum(tested))

# Gather into long format
y <- 
  rbind(x %>% dplyr::select(date, p) %>%
          rename(val = p) %>%
          mutate(key = 'positive'),
        x %>% dplyr::select(date, tested) %>%
          rename(val = tested) %>%
          mutate(key = 'tested')) %>%
  mutate(key = factor(key, levels = c('tested', 'positive')))

ggplot(data = x,
       aes(x = date,
           y = p)) +
  geom_bar(stat = 'identity') +
  xlab('Date') +
  ylab('Number of positive cases') +
    theme_wsj() +
  ggtitle('Malaria miscroscopy positive results',
          'All workers') +
  theme(title = element_text(size = 12))
```

### By nationality

The below charts show the total number of tests and postive cases, by month, for Mozambican and foreign workers, respectively, at Ilovo-Maragra.

```{r, results = 'asis'}
gg <- gvisLineChart(data = df %>% filter(group == 'nacionais'),
                    xvar = 'date',
                    yvar = c('tested', 'positive'),
                    options = list(curveType='function',
                                  title="Malaria among Mozambican employees",
                                  titleTextStyle="{color:'black',fontName:'Courier',fontSize:16}"))
plot(gg)

gg <- gvisLineChart(data = df %>% filter(group == 'expatriados'),
                    xvar = 'date',
                    yvar = c('tested', 'positive'),
                    options = list(curveType='function',
                                  title="Malaria among expatriated employees",
                                  titleTextStyle="{color:'black',fontName:'Courier',fontSize:16}"))
plot(gg)
# 
# # Heat calendar chart
# gvisCalendar

# # # Motion chart
# x <- gvisMotionChart(df,
#    idvar="group",
#    timevar="date")
```

## Cases and tests

### Absolute numbers

The below chart shows both the number of positive cases (all workers, in red) and the number of tests, by month.

```{r}
ggplot(data = y,
       aes(x = date,
           y = val,
           group = key,
           fill = key)) +
  geom_area(stat = 'identity', position = 'stack', alpha = 0.5) +
  xlab('Date') +
  ylab('Number of positive cases') +
  theme_wsj() +
  scale_fill_manual(name = '',
                    values = c('darkgreen', 'darkred')) +
  ggtitle('Malaria miscroscopy tests and positive results',
          'All workers, absolute') +
    theme(title = element_text(size = 12))

```

### Relative numbers

The below chart shows the same data as above, but convers the number of positive cases to a percentage of all tests, rather than an absolute number.

```{r}
ggplot(data = x,
       aes(x = date,
           y = pp)) +
  geom_area(stat = 'identity', fill = 'darkblue', alpha = 0.5) +
  xlab('Date') +
  ylab('Number of positive cases') +
  theme_wsj() +
  ggtitle('Malaria miscroscopy tests and positive results',
          'All workers, relative') +
    theme(title = element_text(size = 12))

```

# Seasonality analysis

We can examine the annual seasonality of positive cases by overlaying all years' data onto one axis.

```{r}
make_month_name <- function(number){
  format(as.Date(paste0('2015-', number, '-01')), '%B')
}
x <- df %>%
  # mutate(day_number = as.numeric(format(date, '%j'))) %>%
  # mutate(day_number = round(day_number, -1) + 5)
  mutate(month_number = as.numeric(format(date, '%m'))) %>%
  group_by(month_number, year) %>%
  summarise(p = sum(positive),
            pp = sum(positive) / sum(tested) * 100) %>%
  mutate(year = factor(year,
                       levels = as.character(2011:2016))) %>%
  ungroup %>%
  mutate(month_number = factor(make_month_name(month_number),
                               levels = make_month_name(1:12)))

cols <- colorRampPalette(brewer.pal(n = 9,  'Spectral'))(length(unique(x$year)))

ggplot(data = x,
       aes(x = month_number,
           y = p,
           group = year,
           color = year)) +
  geom_line() +
  scale_color_manual(name = 'Year',
                     values = cols) +
  theme_wsj() +
  ggtitle('Malaria miscroscopy positive results',
          'Absolute cases, year overlapping') +
    theme(title = element_text(size = 12)) +
  xlab('Month') +
  ylab('Positive results') +
  theme(axis.text.x = element_text(angle = 90))
```

The below chart uses the same data as above, but instead of positive cases, it shows the percent of tests which were positive.

```{r}
ggplot(data = x,
       aes(x = month_number,
           y = p,
           group = year,
           color = year)) +
  geom_line() +
  scale_color_manual(name = 'Year',
                     values = cols) +
  theme_wsj() +
  ggtitle('Malaria miscroscopy positive results',
          'Percent positive, year overlapping') +
    theme(title = element_text(size = 12)) +
  xlab('Month') +
  ylab('Positive results') +
  theme(axis.text.x = element_text(angle = 90))
```

If we aggregate and view distributions (via "violin" charts) at the level of the month, seasonality is more apparent.

```{r}
ggplot(data = x,
       aes(x = month_number,
           y = p)) +
  geom_violin(alpha = 0.6, fill = 'darkorange') +
    theme_wsj() +
  ggtitle('Malaria miscroscopy positive results',
          'Percent positive, year overlapping') +
    theme(title = element_text(size = 12)) +
  xlab('Month') +
  ylab('Positive results') +
  theme(axis.text.x = element_text(angle = 90))
```

# Data

The raw data collected so far appear as such.

```{r}
df %>%
  dplyr::select(year, month, group, tested, positive, negative) %>%
  pander
```

### 2017-01-03 analyze_monthly_data.Rmd
monitoring or data entry

@@
   ylab('Number of positive cases') +
-    theme_wsj() +
+    theme_fivethirtyeight() +
   ggtitle('Malaria miscroscopy positive results',

### 2017-01-03 monitor.R

@@ -1,0 +1,34 @@
+library(cism)
+library(dplyr)
+library(ggplot2)
+
+# Get data
+df <- get_data(tab = 'ECONOMICS_MAL_CORE',
+               dbname = 'sapodk',
+               port = 4706)
+
+# Clean up
+df$date_time <- as.POSIXct(df$`_CREATION_DATE`)
+
+# Plot by time
+by_time <-
+  df %>%
+  mutate(dummy = 1) %>%
+  arrange(date_time) %>%
+  group_by(person = DEVICEID) %>%
+  mutate(cs = cumsum(dummy)) %>%
+  ungroup
+
+ggplot(data = by_time,
+       aes(x = date_time,
+           y = cs,
+           group = DEVICEID,
+           color = DEVICEID)) +
+  geom_line() +
+  theme_cism() +
+  labs(x = 'Date-time',
+       y = 'Cumulative entries',
+       title = 'Maragra data entry progress',
+       subtitle = 'Device IDs are unique for each worker') +
+  scale_color_manual(name = 'Device ID',
+                     values = c('purple', 'green'))

### 2017-08-21 README.Rmd
major structuring

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Maragra and malaria

_Joe Brew, Elisa Sicuri, Laia Cirera, Kizito Gondo, Elton Dorkin_

# Details

- This paper-as-a-package concept was inspired by https://github.com/Pakillo/template

### 2017-08-21 analyses/analyze_monthly_data.Rmd
major structuring

--- 2017-01-03 analyze_monthly_data.Rmd
@@
 library(brew)
+library(maragra)
 ## Global options
@@
 library(pander)
+library(cism)
 
 # Read data
-df <- read_csv(file = 'data/maragra_monthly_data.csv')
+clinic_agg <- maragra::clinic_agg
 
 # Create date column
-df$date <- as.Date(paste0(df$year, '-', df$month, '-', 15))
+clinic_agg$date <- as.Date(paste0(clinic_agg$year, '-', clinic_agg$month, '-', 15))
 
 # Create a percent postiive column
-df$percent_positive <- df$positive / df$tested * 100
+clinic_agg$percent_positive <- clinic_agg$positive / clinic_agg$tested * 100
 
 # Arrange by date
-df <- df %>% arrange(date)
+clinic_agg <- clinic_agg %>% arrange(date)
 
 # Get rid of any unknowns
-df <- df %>% filter(!is.na(tested))
+clinic_agg <- clinic_agg %>% filter(!is.na(tested))
 ```
@@
 x <- 
-  df %>%
+  clinic_agg %>%
   group_by(date) %>%
@@
 ```{r, results = 'asis'}
-gg <- gvisLineChart(data = df %>% filter(group == 'nacionais'),
+gg <- gvisLineChart(data = clinic_agg %>% filter(group == 'nacionais'),
                     xvar = 'date',
@@
 
-gg <- gvisLineChart(data = df %>% filter(group == 'expatriados'),
+gg <- gvisLineChart(data = clinic_agg %>% filter(group == 'expatriados'),
                     xvar = 'date',
@@
 # # # Motion chart
-# x <- gvisMotionChart(df,
+# x <- gvisMotionChart(clinic_agg,
 #    idvar="group",
@@
   ylab('Number of positive cases') +
-  theme_wsj() +
+  theme_cism() +
   scale_fill_manual(name = '',
@@
   ylab('Number of positive cases') +
-  theme_wsj() +
+  theme_cism() +
   ggtitle('Malaria miscroscopy tests and positive results',
@@
 }
-x <- df %>%
+x <- clinic_agg %>%
   # mutate(day_number = as.numeric(format(date, '%j'))) %>%
@@
 ```{r}
-df %>%
+clinic_agg %>%
   dplyr::select(year, month, group, tested, positive, negative) %>%
-  pander
+  kable
 ```

### 2017-08-21 manuscript/manuscript.Rmd
major structuring

---
title: "Manuscript"
author: "Joe Brew"
date: "August 21, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2017-08-21 R/maragra-package.r

@@ -1,0 +1,5 @@
+#' maragra.
+#'
+#' @name maragra
+#' @docType package
+NULL

### 2017-08-21 build_package.R

@@ -1,0 +1,8 @@
+library(devtools)
+library(roxygen2)
+library(rmarkdown)
+document('.')
+install('.')
+render('README.Rmd')
+setwd('vignettes')
+render('vignette.Rmd')

### 2017-08-21 monitor.R

@@ -1,84 +1,0 @@
-library(cism)
-library(dplyr)
-library(ggplot2)
-
-# Get data for clinic
-df <- get_data(tab = 'ECONOMICS_MAL_CORE',
-               dbname = 'sapodk',
-               port = 3306)
-
-# Clean up
-df$date_time <- as.POSIXct(df$`_CREATION_DATE`)
-
-# Plot by time
-by_time <-
-  df %>%
-  mutate(dummy = 1) %>%
-  arrange(date_time) %>%
-  group_by(person = DEVICEID) %>%
-  mutate(cs = cumsum(dummy)) %>%
-  ungroup
-
-ggplot(data = by_time,
-       aes(x = date_time,
-           y = cs,
-           group = DEVICEID,
-           color = DEVICEID)) +
-  geom_line() +
-  theme_cism() +
-  labs(x = 'Date-time',
-       y = 'Cumulative entries',
-       title = 'Maragra data entry progress',
-       subtitle = 'Device IDs are unique for each worker') +
-  scale_color_manual(name = 'Device ID',
-                     values = c('purple', 'green'))
-
-
-# Get data for malaria control
-mc <- get_data(tab = 'MARAGRA_VECTOR_CONTROL_CORE',
-               dbname = 'sapodk',
-               port = 3306)
-
-# Clean up
-mc$date_time <- as.POSIXct(mc$`_CREATION_DATE`)
-mc$date <- as.POSIXct(mc$DATA)
-
-# Plot by time
-by_time <-
-  mc %>%
-  mutate(dummy = 1) %>%
-  arrange(date_time) %>%
-  mutate(cs = cumsum(dummy)) %>%
-  ungroup
-
-ggplot(data = by_time,
-       aes(x = date_time,
-           y = cs)) +
-  geom_line() +
-  theme_cism() +
-  labs(x = 'Date-time',
-       y = 'Cumulative entries',
-       title = 'Maragra data entry progress',
-       subtitle = 'Malaria control')
-
-
-# Also explore by year
-by_year <- 
-  mc %>%
-  mutate(year = as.numeric(format(date, '%Y'))) %>%
-  group_by(year) %>%
-  tally
-
-ggplot(data = by_year,
-       aes(x = year,
-           y = n)) +
-  geom_bar(stat = 'identity',
-           fill = 'darkgreen',
-           alpha = 0.6) +
-  theme_cism() +
-  labs(x = 'Year',
-       y = 'Records entered',
[…]

### 2017-08-21 tests/testthat.R

@@ -1,0 +1,4 @@
+library(testthat)
+library(maragra)
+
+test_check("maragra")

### 2017-08-22 analyses/analyze_monthly_data.Rmd
update to raw data files

@@
 clinic_agg <- maragra::clinic_agg
-
-# Create date column
-clinic_agg$date <- as.Date(paste0(clinic_agg$year, '-', clinic_agg$month, '-', 15))
-
-# Create a percent postiive column
-clinic_agg$percent_positive <- clinic_agg$positive / clinic_agg$tested * 100
-
-# Arrange by date
-clinic_agg <- clinic_agg %>% arrange(date)
-
-# Get rid of any unknowns
-clinic_agg <- clinic_agg %>% filter(!is.na(tested))
 ```

### 2017-08-22 data-raw/create_data_files.R

@@ -1,0 +1,155 @@
+library(devtools)
+library(readr)
+library(tidyr)
+library(dplyr)
+library(cism)
+library(dplyr)
+library(ggplot2)
+library(readxl)
+
+# Initiate the raw-data into package session
+devtools::use_data_raw()
+
+# READ DATA ###################################
+
+# Read in absenteeism data
+# ab <- readxl::read_excel('HRS - Leave Applications.xls') # too large
+ab <- readr::read_csv('HRS - Leave Applications.csv')
+ab2 <- readr::read_csv('HRS - Leave Applications - Sheet 2.csv')
+ab <- bind_rows(ab, ab2)
+rm(ab2)
+
+# Read in clinic data
+clinic <- get_data(tab = 'ECONOMICS_MAL_CORE',
+               dbname = 'sapodk')
+
+# Read in aggregate level clinic data
+clinic_agg <- read_csv('maragra_monthly_data.csv')
+
+# Read in malaria control data
+mc <- get_data(tab = 'MARAGRA_VECTOR_CONTROL_CORE',
+               dbname = 'sapodk')
+
+# Read in workers data
+workers <- readxl::read_excel('All PERMANENT & NOM PERMANENT EE TO USE IN COMPLEMENT REPORT.xls',
+                              skip = 2)
+
+# Save a backup
+save.image(paste0('../backups/', Sys.Date(), '.RData'))
+
+# CLEAN DATA ###################################
+
+# ab ########
+# Get a date column
+ab$date <- as.Date(ab$`Date Captured`, format = '%d-%b-%Y')
+
+# Clean up column names
+names(ab) <- tolower(gsub(' ', '_', names(ab)))
+
+# clinic ######
+# Clean up column names
+names(clinic) <- tolower(gsub(' ', '_', names(clinic)))
+
+# Create a name column
+person_name <- iconv(clinic$b_group_full_name,"WINDOWS-1252","UTF-8")
+person_name <- tolower(person_name)
+person_name <- gsub(' _pf| -pf| -po| _po|_pf|pf|-pf|4g| _pg| _f| _|_po|3,pm| _f| _pm|_po|_pt|_pg|_f|_f,g| _f| _g| _pm|_pm', '', person_name)
+clinic$name <- toupper(person_name)
+
+# Clean up dates
+clinic$creation_date <- as.Date(clinic$`_creation_date`)
+clinic$date <- as.Date(clinic$b_group_date)
+
+# Get severity
[…]

### 2017-08-22 README.Rmd
making folders visible through gitignores

@@
 
-# Maragra and malaria
+# Malaria and sugar
+_An in-depth examination of the effect of malaria control activities on the health and productivity of Maragra sugarcane factory workers_
 
-_Joe Brew, Elisa Sicuri, Laia Cirera, Kizito Gondo, Elton Dorkin_
+# Purpose
+
+This code repository serves as the "research compendium" for the analysis of Maragra malaria incidence, malaria control, and worker absenteeism and productivity data.
+The purpose of this repository - which takes the form of a fully documented R package - is "to integrate the computations and code used in data analyses, methodological descriptions, simulations, etc.
+with the documents that describe and rely on them" ([Gentleman and Lang, 2004](http://biostats.bepress.com/bioconductor/paper2/)).
+By using this package, anyone (with access to the original data in its original form) will be able to reproduce every step of the analysis and mansucript writing.
+
+# Instructions
+
+1. Clone this repository (using the following code in bash).
+
+```
+$ git clone https://github.com/joebrew/maragra
+```
+
+2. Install the code in the repository as an R package.
+
+```
+# !/usr/bin/R
+devtools::install('maragra')
+devtools::document('maragra')
+```
+
+3.
+
+```
+
+```
 

### 2017-08-22 README.Rmd
major update

@@
 with the documents that describe and rely on them" ([Gentleman and Lang, 2004](http://biostats.bepress.com/bioconductor/paper2/)).
-By using this package, anyone (with access to the original data in its original form) will be able to reproduce every step of the analysis and mansucript writing.
+
+In research, reproducibility is key to transparency and reliability.
+By using this package, any authorized collaborator will be able to reproduce every step of the analysis and mansucript writing by following the below instructions.
 
@@
 
-1. Clone this repository (using the following code in bash).
+#### 1. Clone this repository (using the following code in bash).
 
@@
 
-2. Install the code in the repository as an R package.
+#### 2. Prepare data
+
+First, youĺl need to populate your `data-raw` folder. This is where you should place those files which are not publicly available. These are:
+
+```
+├── HRS - Leave Applications.csv
+├── HRS - Leave Applications - Sheet 2.csv
+├── HRS - Leave Applications.xls
+└── maragra_monthly_data.csv
+```
+
+If you do not have these files but are an authorized collaborator, email joebrew@gmail.com.
+
+#### 3. Populate your `credentials.yaml` folder
+
+You'll need to send some credentials to the CISM database when accessing it. You should create a `credentials.yaml` file in the `credentials` folder with the following values:
+
+```
+dbname: openhds
+host: sap.manhica.net
+port: 4706 # (use 3306 instead if based in the CISM)
+user: xxx
+password: xxx
+```
+
+Note that the `user` and `password` fields should be changed to database credentials provided by the CISM.
+If you do not have these credentials but are an authorized collaborator, email joebrew@gmail.com.
+
+#### 4. Process the data
+
+The raw data is not in suitable form for analysis.
+By running the below script, you'll create [tidy datasets](http://vita.had.co.nz/papers/tidy-data.html) from the raw data, which can then be called by simply naming them after attaching the `maragra` package.
+
+To process data, run the following in an R session from within the `maragra` directory:
+
+```
+setwd('data-raw')
+source('create_data_files.R')
+```
+
+Alternatively, you can run this directly from bash:
+
+```
+$ cd data-raw
+$ Rscript create_data_files.R
+```
+
+#### 5. Install the code in the repository and build the R package.
 
@@
 
-3.
+If the above fails, it will most likely be due to R package dependencies which you do not have.
+Examine the error messages, and install R packages as necessary.
+Note that the `cism` and `brew` packages are available from https://github.com/joebrew, whereas the `databrew` package is available from https://github.com/databrew.
+
+As an alternative to the above, or to update documentation (including this README), run the following in an R session from within the `maragra` directory:
 
 ```
+source('build_package.R')
+```
+
+Alternatively, you can run the script from directly within bash:
 
 ```
+$ Rscript build_package.R
+```
 
-# Details
+#### 6. Run the tests
 
-- This paper-as-a-package concept was inspired by https://github.com/Pakillo/template
+Unit tests have been written to ensure that the processed data and packages are behaving correctly. Run these tests by running the following in an R session from within the `maragra` directory:
+
+```
+setwd('tests')
+source('testthat.R')
+```
+
+#### 7. Generate outputs
+
+All analysis is written in the `.Rmd` format.
+Analysis files are in the `inst/rmd` folder.
+These can be knitted directly using `rmarkdown::render`, or called via specific functions.
+For example, to compile the "Maragra clinical malaria incidence" report, simply run the following from within an R session:
+
+```
+maragra::generate_maragra_clinical_malaria_incidence()
+```
+
+All functions which begin with the term `generate` will produce an output (either pdf or html).
+These functions also all share arguments for `date`, `output_dir`, and `output_file`.
+So, one could specify that the output an analysis should go to the `outputs` folder and be named "incidence_report.hml" as follows:
+
+```
+maragra::generate_maragra_clinical_malaria_incidence(date = Sys.Date(),
+                                                     output_dir = 'outputs',
+                                                     output_file = 'incidence_report.html')
+```
+
+To run _all_ analyses, and produce all reports and papers to the `outputs` folder, once can simply run the following from an R session within the `maragra` folder:
+
+```
+maragra::run_all_analyses(output_dir = 'outputs')
+```
+
+## Details
+
+### Funding
+
+This research is for Joe Brew's PhD.
+Full details at [www.economicsofmalaria.com](http://economicsofmalaria.com) He is generously funded by the Erasmus Mundus Joint Doctorate Fellowship, Specific Grant Agreement 2016-1346.
+His program of study is the [Transdisciplinary Global Health programme](http://www.transglobalhealth.org/).
+
+### Collaboration
+
+A debt of gratitude is owed to the following people and institutions, without whom this research would not be possible:
+
+#### People
+
+- Elisa Sicuri, for her guidance and expertise   
+- Laia Cirera, for her on-the-ground knowledge and willingness to work hard    
+- Jacqueline Broerse, for her global persepctive and critical contributions    
+- Menno Pradhan, for his technical knowledge and experience    
+- Kizito Gondo, for helping to make this collaboration go from idea to reality  
+- Elton Dorkin, for his subject matter insight and administrative familiarity  
+- The entire team at the Centro de Investigação em Saude de Manhiça
+
+#### Institutions
+
+- Maragra Açucar and Illovo Sugar (Mozambique)  
+- Institut de Salut Global de Barcelona (Spain)   
+- Centro de Investigação de Saude de Manhiça (Mozambique)
+- Vrije Universiteit, Amsterdam (Netherlands)
+- The European Commission, Erasmus Mundus fellowship, and Transdisciplinary Global Health programme (Europe)

### 2017-08-22 R/generate_maragra_clinical_malaria_incidence.R

@@ -1,0 +1,42 @@
+#' Generate Maragra clinical malaria incidence report
+#' 
+#' Generate an html report of Maragra clinical malaria incidence, based
+#' on aggregate numbers manually collected by Joe in December 2016 at the 
+#' Maragra clinic and laboratory
+#' @param date A date to be printed on the report. If \code{NULL} (the 
+#' default), the current date will be used
+#' @param output_dir The directory to which the file should be written. If 
+#' \code{NULL} (the default), the current working directory will be used.
+#' @param output_file The name of the file to be written. 
+#' @return An html will be written
+#' @importFrom rmarkdown render
+#' @export
+
+generate_maragra_clinical_malaria_incidence <- function(date = NULL,
+                                                     output_dir = NULL,
+                                                     output_file = 'malaria_clinical_malaria_incidence.html'){
+  
+  # If no output directory, make current wd
+  if(is.null(output_dir)){
+    output_dir <- getwd()
+  }
+  
+  # If not date, use today's
+  if(is.null(date)){
+    date <- Sys.Date()
+  }
+  
+  # Combine parameters into a list, so as to pass to Rmd
+  parameters <- list(date = date)
+  
+  # Find location the rmd to knit
+  file_to_knit <- 
+    system.file('rmd/maragra_clinical_malaria_incidence.Rmd', 
+                package='maragra')
+  
+  # Knit file
+  rmarkdown::render(file_to_knit,
+                    output_dir = output_dir,
+                    output_file = output_file,
+                    params = parameters)
+}

### 2017-08-22 R/run_all_analyses.R

@@ -1,0 +1,31 @@
+#' Run all analyses
+#' 
+#' Run all analyses and produce outputs in the \code{outputs_dir} folder
+#' @param date A date to be printed on the report. If \code{NULL} (the 
+#' default), the current date will be used
+#' @param output_dir The directory to which the analyses should be written. If 
+#' \code{NULL} (the default), the current working directory will be used.
+#' @return html and pdf files will be written
+#' @importFrom rmarkdown render
+#' @export
+
+run_all_analyses <- function(date = NULL,
+                               output_dir = NULL){
+  
+  # If no output directory, make current wd
+  if(is.null(output_dir)){
+    output_dir <- getwd()
+  }
+  
+  # If not date, use today's
+  if(is.null(date)){
+    date <- Sys.Date()
+  }
+  
+  # Clinical malaria incidence report
+  message(paste0('Generating the clinical malaria incidence report in ',
+          output_dir))
+  generate_maragra_clinical_malaria_incidence(date = date,
+                                              output_dir = output_dir)
+  
+}

### 2017-08-22 data-raw/create_data_files.R

@@ -10,7 +10,7 @@ library(readxl)
 # Initiate the raw-data into package session
 devtools::use_data_raw()
 
-# READ DATA ###################################
+# ab ####################################################
 
 # Read in absenteeism data
 # ab <- readxl::read_excel('HRS - Leave Applications.xls') # too large
@@ -19,37 +19,29 @@ ab2 <- readr::read_csv('HRS - Leave Applications - Sheet 2.csv')
 ab <- bind_rows(ab, ab2)
 rm(ab2)
 
-# Read in clinic data
-clinic <- get_data(tab = 'ECONOMICS_MAL_CORE',
-               dbname = 'sapodk')
-
-# Read in aggregate level clinic data
-clinic_agg <- read_csv('maragra_monthly_data.csv')
-
-# Read in malaria control data
-mc <- get_data(tab = 'MARAGRA_VECTOR_CONTROL_CORE',
-               dbname = 'sapodk')
-
-# Read in workers data
-workers <- readxl::read_excel('All PERMANENT & NOM PERMANENT EE TO USE IN COMPLEMENT REPORT.xls',
-                              skip = 2)
-
-# Save a backup
-save.image(paste0('../backups/', Sys.Date(), '.RData'))
-
-# CLEAN DATA ###################################
-
-# ab ########
 # Get a date column
 ab$date <- as.Date(ab$`Date Captured`, format = '%d-%b-%Y')
 
 # Clean up column names
 names(ab) <- tolower(gsub(' ', '_', names(ab)))
 
-# clinic ######
+# Save for use in package
+devtools::use_data(ab,
+                   overwrite = TRUE)
+
+# clinic ##################################################
+
+# Read in clinic data
+if('clinic.RData' %in% dir()){
+  load('clinic.RData')
+} else {
+  clinic <- get_data(tab = 'ECONOMICS_MAL_CORE',
+                     dbname = 'sapodk')
+  save(clinic,
+       file = 'clinic.RData')
+}
 # Clean up column names
 names(clinic) <- tolower(gsub(' ', '_', names(clinic)))
-
 # Create a name column
 person_name <- iconv(clinic$b_group_full_name,"WINDOWS-1252","UTF-8")
 person_name <- tolower(person_name)
@@ -68,7 +60,14 @@ clinic <-
   clinic %>%
   dplyr::select(date, name, severity, creation_date)
 
-# clinic_agg ######
+# Save for use in package
+devtools::use_data(clinic,
[…]

### 2017-08-22 build_package.R
minor change to make creation of data simpler

@@ -1,3 +1,5 @@
+#! /usr/bin/R
+reconstruct_data <- FALSE
 library(devtools)
 library(roxygen2)
 library(rmarkdown)
@@ -6,3 +8,9 @@ install('.')
 render('README.Rmd')
 setwd('vignettes')
 render('vignette.Rmd')
+setwd('..')
+if(reconstruct_data){
+  setwd('data-raw')
+  source('create_data_files.R')
+  setwd('..')
+}

### 2017-08-22 data-raw/create_data_files.R
creating panel data from expanded absences and presences

@@ -10,6 +10,32 @@ library(readxl)
 # Initiate the raw-data into package session
 devtools::use_data_raw()
 
+# Define some helper functions
+
+# Generate date columns
+make_date_columns <- function(df){
+  if(!'date' %in% names(df)){
+    stop('Must have a column named date')
+  }
+  if(class(df$date) != 'Date'){
+    stop('date column must be of class Date')
+  }
+  if('month' %in% names(df) | 'year' %in% names(df) |
+     'dow' %in% names(df) | 'day' %in% names(df) |
+     'day_number' %in% names(df) | 'year_month' %in% names(df)){
+    stop('Cannot have already have columns named month, year, dow, day, day_number, or year_month')
+  }
+  df <- df %>%
+    mutate(month = as.numeric(format(date, '%m')),
+           year = as.numeric(format(date, '%Y')),
+           dow = weekdays(date),
+           day = as.numeric(format(date, '%d')),
+           day_number = as.numeric(format(date, '%j'))) %>%
+    mutate(year_month = paste0(year, '-', month))
+  return(df)
+}
+
+
 # ab ####################################################
 
 # Read in absenteeism data
@@ -19,11 +45,39 @@ ab2 <- readr::read_csv('HRS - Leave Applications - Sheet 2.csv')
 ab <- bind_rows(ab, ab2)
 rm(ab2)
 
-# Get a date column
-ab$date <- as.Date(ab$`Date Captured`, format = '%d-%b-%Y')
-
 # Clean up column names
 names(ab) <- tolower(gsub(' ', '_', names(ab)))
+
+# Format dates
+format_date <- function(x){as.Date(x, '%d-%b-%Y')}
+ab <- ab %>%
+  mutate(date_captured = format_date(date_captured),
+         leave_from_date = format_date(leave_from_date),
+         leave_to_date = format_date(leave_to_date),
+         date_approved = format_date(date_approved),
+         date_captured = format_date(date_captured),
+         date_signed_leave_form_rec = format_date(date_signed_leave_form_rec))
+
+# Keep only relevant columns
+ab <-
+  ab %>%
+  dplyr::select(-update_indicator,
+                -leave_sequence_number,
+                -leave_reason,
+                -date_captured,
+                -leave_form_number,
+                -paid_unpaid_indicator,
+                -oracle_number_approved_by,
+                -oracle_number_denied_by,
+                -date_denied,
+                -date_approved,
[…]

### 2017-08-24 data-raw/create_data_files.R
better snapshotting

@@ -235,53 +235,72 @@ expandify <- function(oracle_number = 'abc',
 }
 
 # Go through each worker and get the eligible working days
-eligible_working_days <- list()
-nrw <- nrow(workers)
-for(i in 1:nrw){
-  message(i, ' of ', nrw)
-  worker_type <- workers$employee_indicator_type[i]
-  if(worker_type == 'C'){
-    start_date <- workers$contract_start_date[i]
-    end_date <- workers$contract_end_date[i]
-  } else {
-    start_date <- workers$company_entry_date[i]
-    end_date <- as.Date('2017-03-31')
+if('eligible_working_days.RData' %in% dir()){
+  load('eligible_working_days.RData')
+} else {
+  eligible_working_days <- list()
+  nrw <- nrow(workers)
+  for(i in 1:nrw){
+    message(i, ' of ', nrw)
+    worker_type <- workers$employee_indicator_type[i]
+    if(worker_type == 'C'){
+      start_date <- workers$contract_start_date[i]
+      end_date <- workers$contract_end_date[i]
+    } else {
+      start_date <- workers$company_entry_date[i]
+      end_date <- as.Date('2017-03-31')
+    }
+    eligible_working_days[[i]] <- 
+      expandify(oracle_number = workers$oracle_number[i],
+                start = start_date,
+                end = end_date)
   }
-  eligible_working_days[[i]] <- 
-    expandify(oracle_number = workers$oracle_number[i],
-              start = start_date,
-              end = end_date)
+  
+  # Bind all together
+  eligible_working_days <- bind_rows(eligible_working_days)
+  save(eligible_working_days,
+       file = 'eligible_working_days.RData')
 }
-
-# Bind all together
-eligible_working_days <- bind_rows(eligible_working_days)
 
 # Expand absences
-ab_expanded <- list()
-nra <- nrow(ab)
-for (i in 1:nra){
-  message(i, ' of ', nra)
-  this_oracle_number <- ab$oracle_number[i]
-  sub_data <- ab[i,]
-  x <- expandify(oracle_number = this_oracle_number,
-                 start = sub_data$leave_from_date,
-                 end = sub_data$leave_to_date)
-  x <- left_join(x,
-                 y = sub_data %>%
-                   dplyr::select(oracle_number,
-                                 leave_type,
-                                 leave_taken),
-                 by = 'oracle_number')
-  ab_expanded[[i]] <- x
[…]

### 2017-08-28 inst/rmd/maragra_preliminary_analysis.Rmd
done with preliminary analysis skeleton

---
date: ""
fig_height: 2.6
fig_width: 4
output:
  html_document:
    css: resources/style.css
    pandoc_args: [
      "+RTS", "-K16000m",
      "-RTS"
    ]
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# Packages
library(tidyverse)
library(knitr)
library(brew)
library(maragra)
library(knitr)
library(googleVis)
op <- options(gvis.plot.tag='chart')

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

<p style="text-align:left;">
[www.economicsofmalaria.com](www.economicsofmalaria.com)
<span style="float:right;">joebrew@gmail.com</span>
</p>

```{r}
# Read in data
ab <- maragra::ab
ab_panel <- maragra::ab_panel
clinic <- maragra::clinic
clinic_agg <- maragra::clinic_agg
mc <- maragra::mc
workers <- maragra::workers
```

# Maragra and Malaria: preliminary analysis {.tabset}

This document contains some basic preliminary analysis of data relating to malaria, IRS, and absenteeism at Maragra Açucar, in Maragra, Mozambique. To explore a specific section, click any of the below tabs.

## Description of data

```{r}
d <- data(package = "maragra")
## assign it to use later
nm <- d$results[, "Item"]
```

The "Maragra database" consists of the following `r length(nm)` tables:

```{r}
## call the promised data
data(list = nm, package = "maragra")
## get the dimensions of each data set
x <- lapply(mget(nm), dim)
results <- list()
for (i in 1:length(x)){
  z <- x[[i]]
  results[[i]] <- data_frame(`Table name` = names(x)[i],
             `Number or rows` = z[1],
             `Number of columns` = z[2])
}
results <- bind_rows(results)

kable(results)
```

These datasets are all readily available in the (private) `maragra` R package (access credentials given to collaborators upon [joebrew@gmail.com](request)). All datasets except for `ab_panel` are _as is_ (ie, in their original raw format), except for 4 kinds of changes:

1. Removal of unecessary information (columns deemed irrelevant to the study have been removed).  
2. Modification of column names (all data now employ underscores instead of spaces, and are all lowercase).    
3. Standardization of data formats (all dates are now in a YYYY-MM-DD format, all names use similar capitalization, irrelevant tablet-created timestamps, etc. have been removed).  
4. "Feature generation": certain columns were created in order to ease grouped analysis and merging of datasets (for example, `year_month`, `day_number`, etc.).

The `ab_panel` dataset is _not_ raw data; rather, it is an amolgamation of the `ab` and `workers` datasets, using absence data from the former with worker elibility dates from the latter, so as to create a "panel" style dataset (ie, one row for each day in which a worker was _supposed_ to work).

The `clinic` and `clinic_agg` are similar, but not identical, in form. The former has individual-level data (useful for pairing with absences and demographic census data), whereas the latter is simply the raw counts of cases by nationality over time. Though `clinic` is much richer and more detailed than `clinic_agg`, is only covers the period from 2014 through 2016, whereas the latter covers a larger time period (2010-2016).

For a "deep dive" into each kind of data, click the tabs at the top of this page.

In order to reproduce this analysis (and others related to these datasets), follow the instructions on this project's [research compendium page](https://github.com/joebrew/maragra#purpose).

## IRS analysis

Here is a chart.

```{r}
ggplot(data = data.frame(a = 1:10,
                         b = 1:10)) +
  geom_point(aes(x = a, y = b))
```

## Absenteeism analysis

Here is a chart.

```{r}
ggplot(data = data.frame(a = 1:10,
                         b = 1:10)) +
  geom_point(aes(x = a, y = b))
```

## Clinic analysis

Here is a chart.

```{r}
ggplot(data = data.frame(a = 1:10,
                         b = 1:10)) +
  geom_point(aes(x = a, y = b))
```

## Next steps

### 2017-08-28 R/generate_preliminary_analysis.R

@@ -1,0 +1,41 @@
+#' Generate preliminary analysis report
+#' 
+#' Generate an html report of the preliminary analysis report,
+#' originally prepared in August / September 2017
+#' @param date A date to be printed on the report. If \code{NULL} (the 
+#' default), the current date will be used
+#' @param output_dir The directory to which the file should be written. If 
+#' \code{NULL} (the default), the current working directory will be used.
+#' @param output_file The name of the file to be written. 
+#' @return An html will be written
+#' @importFrom rmarkdown render
+#' @export
+
+generate_preliminary_analysis <- function(date = NULL,
+                                                     output_dir = NULL,
+                                                     output_file = 'generate_preliminary_analysis.html'){
+  
+  # If no output directory, make current wd
+  if(is.null(output_dir)){
+    output_dir <- getwd()
+  }
+  
+  # If not date, use today's
+  if(is.null(date)){
+    date <- Sys.Date()
+  }
+  
+  # Combine parameters into a list, so as to pass to Rmd
+  parameters <- list(date = date)
+  
+  # Find location the rmd to knit
+  file_to_knit <- 
+    system.file('rmd/maragra_preliminary_analysis.Rmd', 
+                package='maragra')
+  
+  # Knit file
+  rmarkdown::render(file_to_knit,
+                    output_dir = output_dir,
+                    output_file = output_file,
+                    params = parameters)
+}

### 2017-08-28 R/run_all_analyses.R

@@ -10,7 +10,7 @@
 #' @export
 
 run_all_analyses <- function(date = NULL,
-                               output_dir = NULL){
+                             output_dir = NULL){
   
   # If no output directory, make current wd
   if(is.null(output_dir)){
@@ -24,8 +24,14 @@ run_all_analyses <- function(date = NULL,
   
   # Clinical malaria incidence report
   message(paste0('Generating the clinical malaria incidence report in ',
-          output_dir))
+                 output_dir))
   generate_maragra_clinical_malaria_incidence(date = date,
                                               output_dir = output_dir)
+  
+  # Clinical preliminary analysis report
+  message(paste0('Generating the preliminary analysis report in ',
+                 output_dir))
+  generate_preliminary_analysis(date = date,
+                                output_dir = output_dir)
   
 }

### 2017-08-28 inst/rmd/maragra_preliminary_analysis.Rmd
major update

@@
 library(knitr)
+library(Hmisc)
 library(brew)
@@
 library(googleVis)
+library(RColorBrewer)
 op <- options(gvis.plot.tag='chart')
@@
 
+These tables all constitute three kinds of data: those pertaining to medical outcomes (generally referred to as "clinic" data), those pertaining to worker information (demographic and absenteeism-related), and those pertaining to malaria control activities.
+
 These datasets are all readily available in the (private) `maragra` R package (access credentials given to collaborators upon [joebrew@gmail.com](request)).
@@
 
-Here is a chart.
+### Data structure
+
+Data about "fumigações" (indoor residual spraying or "IRS") dates from `r min(mc$date)` through `r max(mc$date)`. The dataset consists of the following fields:
 
 ```{r}
-ggplot(data = data.frame(a = 1:10,
-                         b = 1:10)) +
-  geom_point(aes(x = a, y = b))
+cat(paste0(names(mc), collapse = '\n'))
+```
+
+Each row is one fumigation activity.
+The location/residence key is the `unidade` column.
+The `casas_cobertas` column indicates the number of houses in that `unidade` that were sprayed, wereas the `pulverizados` column indicates the number of rooms.
+
+### Insecticide type
+
+The insecticide used is either DDT or ACT. The below is a breakdown of their respective use.
+
+```{r}
+ggplot(data = mc,
+       aes(x = insecticida)) +
+  geom_bar(fill = '#159957',
+           alpha = 0.6) +
+  theme_maragra() +
+  labs(x = 'Insecticide',
+       y = 'IRS activities',
+       title = 'Maragra IRS activity by insecticide type')
+```
+
+The respective use over time is as follows:
+
+```{r}
+x <- mc %>%
+  group_by(date = as.Date(paste0(format(date, '%Y-%m'), '-01')),
+           insecticida) %>%
+  summarise(houses = sum(casas_cobertas)) %>%
+  ungroup
+# Get a left side for filling in the 0s
+left <- expand.grid(date = seq(min(x$date),
+                               max(x$date),
+                               by = 'month'),
+                    insecticida = sort(unique(x$insecticida)))
+# Join together
+x <- left_join(x = left,
+               y = x,
+               by = c('date', 
+                      'insecticida'))
+# Replace NAs with 0
+x <- x %>%
+  mutate(houses = ifelse(is.na(houses), 0, houses))
+# Define a color vector
+cols <- colorRampPalette(brewer.pal(n = 9, 'Spectral'))(length(unique(x$insecticida)))
+
+ggplot(data = x,
+       aes(x = date,
+           y = houses,
+           group = insecticida,
+           color = insecticida)) +
+  geom_point(alpha = 0.6) +
+  geom_line(alpha = 0.6) +
+  theme_maragra() +
+  labs(x = 'Month',
+       y = 'Houses fumigated',
+       title = 'Maragra IRS activity by insecticide type over time') +
+  scale_color_manual(name = 'Chemical',
+                     values = cols)
+```
+
+Below is a table of the same data.
+
+```{r}
+names(x) <- Hmisc::capitalize(names(x))
+DT::datatable(x)
+```
+
+### IRS seasonality
+
+We can examine the date of fumigations to see if they are seasonal vs. randomly/uniformly distributed throughout the year.
+
+```{r}
+x <- mc %>%
+  group_by(year, day_number) %>%
+  summarise(houses = sum(casas_cobertas)) %>%
+  ungroup %>%
+  mutate(fake_date = as.Date('1990-01-01') + day_number) %>%
+  mutate(fake_month = as.Date(paste0(format(fake_date, '%Y-%m'), '-01'))) %>%
+  group_by(year, fake_month) %>%
+  summarise(houses = sum(houses)) %>%
+  ungroup %>%
+  mutate(year = factor(year))
+
+# Fill the nas with 0s
+left <- expand.grid(year = sort(unique(x$year)),
+                    fake_month = sort(unique(x$fake_month)))
+
+x <- left_join(x = left,
+               y = x)
+x$houses[is.na(x$houses)] <- 0
+cols <- colorRampPalette(brewer.pal(n = 9, 'Spectral'))(length(unique(x$year)))
+ggplot(data = x %>% 
+         filter(!year %in% c(2011:2012, 2017)),
+       aes(x = fake_month,
+           y = houses)) +#,
+           # group = year,
+           # color = year)) +
+  geom_line(alpha = 0.8) +
+  # scale_color_manual(name = 'Year',
+  #                    values = cols) +
+  scale_x_date(labels = scales::date_format("%b")) +
+  theme_maragra() +
+  labs(x = 'Month',
+       y = 'Houses covered',
+       title = 'Seasonality of IRS operations') +
+  facet_wrap(~year)
 ```
@@
 
-Here is a chart.
+### Data structure
+
+The analysis of absenteeism data will rely on a panel-style dataset in which one row exists for each worker-day (for which the worker is estimated to have _supposed to work_), with columns indicating the outcome (absent or not, sick or not, etc.).
+The Maragra CRM does not natively store panel style data, so we construct it from a combination of the `workers` dataset (from the Human Resources department) and the `absenteeism` dataset.
+Certain features pertaining to illness are merged from the `clinic` dataset.
+
+The `ab_panel` dataset has the following column names:
 
 ```{r}
-ggplot(data = data.frame(a = 1:10,
-                         b = 1:10)) +
-  geom_point(aes(x = a, y = b))
+cat(paste0(names(ab_panel), collapse = '\n'))
+```
+
+### Data magnitude
+
+Overall, we have observed `r nrow(ab_panel)` eligible worker days (the equivalent of `r round(nrow(ab_panel) / 365.25, digits = 0)` years of human activity!), from the period of `r min(ab_panel$date)` through `r max(ab_panel$date)`.
+
+### Absenteeism
+
+Our dataset includes `r length(which(ab_panel$absent))` absences and `r length(which(!ab_panel$absent))` presences, which can be broken down below in percentage terms.
+
+```{r}
+x <- ab_panel %>%
+  group_by(absent = ifelse(absent, 'Absent', 'Present')) %>%
+  tally %>%
+  ungroup %>%
+  mutate(p = n / sum(n) * 100) %>%
+  mutate(p = round(p, digits = 2))
+
+ggplot(data = x,
+       aes(x = absent,
+           y = p)) +
+    geom_bar(fill = '#159957',
+           alpha = 0.6,
+           stat = 'identity') +
+  theme_maragra() +
+  labs(x = 'Status',
+       y = 'Percent',
+       title = 'Absence/presence breakdown of all workers') +
+  geom_label(aes(label = paste0(p, '%')))
+```
+
+Among absences, the breakdown of absence "type" is as follows.
+
+```{r}
+ggplot(data = ab_panel %>% filter(absent) %>%
+         mutate(sick_leave = ifelse(leave_type == 'SIC',
+                                    'Sick leave',
+                                    'Other leave')),
+       aes(x = leave_type,
+           fill = sick_leave)) +
+  geom_bar(alpha = 0.8) +
+  labs(x = 'Leave type',
+       y = 'Absences',
+       title = 'Breakdown of absences by type') +
+  theme_maragra() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  scale_fill_manual(name = '',
+                    values = c('black', 'red'))
+
+```
+
+### Absenteeism rate
+
+We calculate an absenteeism rate for any given time period (day, month, etc.) as the number of days not worked divided by the number of days which _should have_ been worked, and multiplied by 100.
+The below shows the absenteeism rate, by day, for the entire observation period.
+The size of each dot is a reflection of the number of workers observed at that time.
+
+```{r}
+x <- 
+  ab_panel %>%
+  group_by(date) %>%
+  summarise(absences = length(which(absent)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100) %>%
+  mutate(Workers = eligibles)
+
+ggplot(data = x,
+       aes(x = date,
+           y = absenteeism_rate)) +
+  geom_point(alpha = 0.1,
+             color = '#159957',
+             aes(size = Workers)) +
+  geom_smooth() +
+  theme_maragra() +
+  labs(x = 'Date',
+       y = 'Absenteeism rate',
+       title = 'Absenteeism rate over time')
+
+```
+
+### Sick absenteeism rate
+
+We again calculate the absenteeism rate, but only for those who absences which are classified as "sick leave".
+
+```{r}
+x <- 
+  ab_panel %>%
+  group_by(date) %>%
+  summarise(absences = length(which(absent_sick)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100) %>%
+  mutate(Workers = eligibles)
+
+ggplot(data = x,
+       aes(x = date,
+           y = absenteeism_rate)) +
+  geom_point(alpha = 0.1,
+             color = '#159957',
+             aes(size = Workers)) +
+  geom_smooth() +
+  theme_maragra() +
+  labs(x = 'Date',
+       y = 'Absenteeism rate',
+       title = 'Sick absenteeism rate over time') 
+```
+
+### Monthly absenteeism
+
+Daily absenteeism is problematic in that it introduces a great deal of noise.
+So, we examine monthly absenteeism rates for both all and sick absences.
+The below chart shows these metrics, as well as the percentage of monthly absences due to sickness, and the number of worker-days observed.
+A local regression smoothed line is overlaid to see overall trends.
+
+```{r}
+x <- 
+  ab_panel %>%
+  group_by(date = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+  summarise(absences = length(which(absent)),
+            sick_absences = length(which(absent_sick)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100) %>%
+  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100) %>%
+  mutate(Workers = eligibles) %>%
+  mutate(sick_absence_p = sick_absenteeism_rate / absenteeism_rate * 100)
+
+# Gather
+x <- x %>%
+  gather(key, value, absenteeism_rate:sick_absence_p) %>%
+  mutate(key = ifelse(key == 'absenteeism_rate', 'Absenteeism rate',
+                      ifelse(key == 'sick_absenteeism_rate', 'Sick absenteeism rate',
+                             ifelse(key == 'sick_absence_p', '% of absences due to sickness',
+                                    key))))
+
+ggplot(data = x,
+       aes(x = date,
+           y = value)) +
+  geom_point(alpha = 0.7,
+             color = '#159957') +
+  geom_line(alpha = 0.4,
+            color = '#159957') +
+  geom_smooth() +
+  theme_maragra() +
+  labs(x = 'Date',
+       y = 'Absenteeism rate',
+       title = 'Sick absenteeism rate over time') +
+  facet_wrap(~key, scales = 'free_y')
+```
+
+### Absenteeism by worker type
+
+The below chart is a paneling of (a) type of absenteeism metric (columns) and (b) type of worker (rows).
+
+```{r, fig.height = 7, fig.width = 10}
+x <- 
+  ab_panel %>%
+  left_join(workers %>% dplyr::select(oracle_number, department_name)) %>%
+  mutate(department = ifelse(department_name %in% c('ADMIN AND FINANCE',
+                                                    'CIVILS SERVICES',
+                                                    'HUMAN RESOURCES SERVICES',
+                                                    'RISK CONTROL'),
+                             'Administrative',
+                             ifelse(department_name == 'CANE GROWING', 'Field',
+                                    ifelse(department_name == 'SUGAR MILLING', 'Factory', 'Other')))) %>%
+  group_by(department, 
+           date = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+  summarise(absences = length(which(absent)),
+            sick_absences = length(which(absent_sick)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100) %>%
+  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100) %>%
+  mutate(Workers = eligibles) %>%
+  mutate(sick_absence_p = sick_absenteeism_rate / absenteeism_rate * 100)
+
+# Gather
+x <- x %>%
+  gather(key, value, absenteeism_rate:sick_absence_p) %>%
+  mutate(key = ifelse(key == 'absenteeism_rate', 'Absenteeism rate',
+                      ifelse(key == 'sick_absenteeism_rate', 'Sick absenteeism rate',
+                             ifelse(key == 'sick_absence_p', '% of absences due to sickness',
+                                    key))))
+
+# Keep just relevant metrics
+x <- x %>%
+  filter(key != 'Workers')
+
+ggplot(data = x,
+       aes(x = date,
+           y = value)) +
+  geom_point(alpha = 0.7,
+             color = '#159957') +
+  geom_line(alpha = 0.4,
+            color = '#159957') +
+  geom_smooth() +
+  theme_maragra() +
+  labs(x = 'Date',
+       y = 'Absenteeism rate',
+       title = 'Sick absenteeism rate over time') +
+  facet_grid(department~key, scales = 'free')
 ```
@@
 
-Here is a chart.
+### Data structure
+
+There are two clinical datasets: `clinic` and `clinic_agg`. The former is detailed at the individual level, but the latter covers a slightly wider timespan.
+
+The `clinic` dataset has the following fields:
 
 ```{r}
-ggplot(data = data.frame(a = 1:10,
-                         b = 1:10)) +
-  geom_point(aes(x = a, y = b))
+cat(paste0(names(clinic), collapse = '\n'))
+```
+
+The `clinic_agg` dataset has the following fields:
+
+```{r}
+cat(paste0(names(clinic_agg), collapse = '\n'))
+```
+
+### Cases over time - from `clinic_agg`
+
+#### All workers
+
+If we combine all workers, we can examine the total incidence of malaria since 2011.
+
+```{r}
+x <- 
+  clinic_agg %>%
+  group_by(date) %>%
+  summarise(p = sum(positive),
+            pp = sum(positive) / sum(tested) * 100,
+            tested = sum(tested))
+
+# Gather into long format
+y <- 
+  rbind(x %>% dplyr::select(date, p) %>%
+          rename(val = p) %>%
+          mutate(key = 'positive'),
+        x %>% dplyr::select(date, tested) %>%
+          rename(val = tested) %>%
+          mutate(key = 'tested')) %>%
+  mutate(key = factor(key, levels = c('tested', 'positive')))
+
+ggplot(data = x,
+       aes(x = date,
+           y = p)) +
+  geom_bar(stat = 'identity') +
+  xlab('Date') +
+  ylab('Number of positive cases') +
+    theme_maragra() +
+  ggtitle('Malaria miscroscopy positive results',
+          'All workers')
+```
+
+#### By nationality
+
+The below charts show the total number of tests and postive cases, by month, for Mozambican and foreign workers, respectively, at Ilovo-Maragra.
+
+```{r, results = 'asis'}
+gg <- gvisLineChart(data = clinic_agg %>% filter(group == 'nacionais'),
+                    xvar = 'date',
+                    yvar = c('tested', 'positive'),
+                    options = list(curveType='function',
+                                  title="Malaria among Mozambican employees",
+                                  titleTextStyle="{color:'black',fontName:'Courier',fontSize:16}"))
+plot(gg)
+
+gg <- gvisLineChart(data = clinic_agg %>% filter(group == 'expatriados'),
+                    xvar = 'date',
+                    yvar = c('tested', 'positive'),
+                    options = list(curveType='function',
+                                  title="Malaria among expatriated employees",
+                                  titleTextStyle="{color:'black',fontName:'Courier',fontSize:16}"))
+plot(gg)
+
+```
+
+### Cases and tests from `clinic_agg`
+
+#### Absolute numbers
+
+The below chart shows both the number of positive cases (all workers, in red) and the number of tests, by month.
+
+```{r}
+ggplot(data = y,
+       aes(x = date,
+           y = val,
+           group = key,
+           fill = key)) +
+  geom_area(stat = 'identity', position = 'stack', alpha = 0.5) +
+  xlab('Date') +
+  ylab('Number of positive cases') +
+  theme_maragra() +
+  scale_fill_manual(name = '',
+                    values = c('black', 'darkred')) +
+  ggtitle('Malaria miscroscopy tests and positive results',
+          'All workers, absolute') +
+    theme(title = element_text(size = 12))
+```
+
+#### Relative numbers
+
+The below chart shows the same data as above, but convers the number of positive cases to a percentage of all tests, rather than an absolute number.
+
+```{r}
+ggplot(data = x,
+       aes(x = date,
+           y = pp)) +
+  geom_area(stat = 'identity', fill = 'darkblue', alpha = 0.5) +
+  xlab('Date') +
+  ylab('Number of positive cases') +
+  theme_maragra() +
+  ggtitle('Malaria miscroscopy tests and positive results',
+          'All workers, relative') +
+    theme(title = element_text(size = 12))
+
+```
+
+### Seasonality analysis from `clinic_agg`
+
+We can examine the annual seasonality of positive cases by overlaying all years' data onto one axis.
+
+```{r}
+make_month_name <- function(number){
+  format(as.Date(paste0('2015-', number, '-01')), '%B')
+}
+x <- clinic_agg %>%
+  # mutate(day_number = as.numeric(format(date, '%j'))) %>%
+  # mutate(day_number = round(day_number, -1) + 5)
+  mutate(month_number = as.numeric(format(date, '%m'))) %>%
+  group_by(month_number, year) %>%
+  summarise(p = sum(positive),
+            pp = sum(positive) / sum(tested) * 100) %>%
+  mutate(year = factor(year,
+                       levels = as.character(2011:2016))) %>%
+  ungroup %>%
+  mutate(month_number = factor(make_month_name(month_number),
+                               levels = make_month_name(1:12)))
+
+cols <- colorRampPalette(brewer.pal(n = 9,  'Spectral'))(length(unique(x$year)))
+
+ggplot(data = x,
+       aes(x = month_number,
+           y = p,
+           group = year,
+           color = year)) +
+  geom_line() +
+  scale_color_manual(name = 'Year',
+                     values = cols) +
+  theme_maragra() +
+  ggtitle('Malaria miscroscopy positive results',
+          'Absolute cases, year overlapping') +
+    theme(title = element_text(size = 12)) +
+  xlab('Month') +
+  ylab('Positive results') +
+  theme(axis.text.x = element_text(angle = 90))
+```
+
+The below chart uses the same data as above, but instead of positive cases, it shows the percent of tests which were positive.
+
+```{r}
+ggplot(data = x,
+       aes(x = month_number,
+           y = p,
+           group = year,
+           color = year)) +
+  geom_line() +
+  scale_color_manual(name = 'Year',
+                     values = cols) +
+  theme_maragra() +
+  ggtitle('Malaria miscroscopy positive results',
+          'Percent positive, year overlapping') +
+    theme(title = element_text(size = 12)) +
+  xlab('Month') +
+  ylab('Positive results') +
+  theme(axis.text.x = element_text(angle = 90))
+```
+
+If we aggregate and view distributions (via "violin" charts) at the level of the month, seasonality is more apparent.
+
+```{r}
+ggplot(data = x,
+       aes(x = month_number,
+           y = p)) +
+  geom_violin(alpha = 0.6, fill = 'darkorange') +
+    theme_maragra() +
+  ggtitle('Malaria miscroscopy positive results',
+          'Percent positive, year overlapping') +
+    theme(title = element_text(size = 12)) +
+  xlab('Month') +
+  ylab('Positive results') +
+  theme(axis.text.x = element_text(angle = 90))
+```
+
+### Malaria severity distribution from `clinic`
+
+The below chart shows the severity of all clinical malaria cases in the Maragra clinic.
+
+```{r}
+ggplot(data = clinic,
+       aes(x = severity)) +
+  geom_bar(alpha = 0.6,
+           fill = 'darkorange') +
+  theme_maragra() +
+  labs(x = 'Severity',
+       y = 'Cases',
+       title = 'Severity of all malaria cases: Maragra clinic')
+```
+
+The below chart shows severity, but over time.
+
+```{r}
+x <- 
+  clinic %>%
+  group_by(severity,
+           date = as.Date(paste0(year_month, '-01'))) %>%
+  tally %>%
+  ungroup %>%
+  group_by(date) %>%
+  mutate(p = n / sum(n) * 100) %>%
+  ungroup %>%
+  mutate(severity = factor(severity,
+                           levels = as.character(rev(sort(unique(severity))))))
+cols <- (colorRampPalette(brewer.pal(9, 'Spectral'))(length(unique(x$severity))))
+
+ggplot(data = x,
+       aes(x = date,
+           y = n)) +
+  geom_line(aes(color = severity)) +
+  scale_color_manual(name = 'Severity',
+                     values = cols) +
+  theme_maragra() +
+  labs(x = 'Date',
+       y = 'Cases',
+       title = 'Severity over time')
+```
+
+```{r}
+ggplot(data = x,
+       aes(x = date,
+           y = p,
+           fill = severity)) +
+  geom_area(stat = 'identity',
+           position = 'stack',
+           alpha = 0.6) +
+  scale_fill_manual(name = 'Severity',
+                    values = cols) +
+  theme_maragra() +
+  labs(x = 'Date',
+       y = 'Percentage of all cases',
+       title = 'Distribution of severity of cases over time')
 ```
@@
 ## Next steps
+
+A lot of work remains to be done:
+
+- Figure out how to transform the `unidade` codes from the IRS data to specific workers's residences (or at least geographical area).  
+- Perform some name standardization so as to better link individual level clinical data with workers demographic and absenteeism data.  
+- Join data with census data for more demographic analysis, particularly in terms of absenteeism and health.  
+- Estimate a model for absenteeism, adjusting for relevant factors, particularly the IRS status and time since IRS.  
+- A cost-effectiveness framework for estimating the return on investment for IRS activities.

### 2017-08-28 R/generate_preliminary_analysis.R

@@ -13,7 +13,7 @@
 
 generate_preliminary_analysis <- function(date = NULL,
                                                      output_dir = NULL,
-                                                     output_file = 'generate_preliminary_analysis.html'){
+                                                     output_file = 'maragra_preliminary_analysis.html'){
   
   # If no output directory, make current wd
   if(is.null(output_dir)){

### 2017-08-28 R/theme_maragra.R

@@ -1,0 +1,37 @@
+#' Theme for ggplots
+#' 
+#' Add a theme to a ggplot object
+#' @param base_size The base font size
+#' @param base_family The font
+#' @return A ggplot compatible theme call
+#' @import ggthemes
+#' @import grid
+#' @export
+# Borrowed from https://rpubs.com/Koundy/71792
+theme_maragra <- function(base_size=14, base_family="Arial") {
+  (theme_foundation(base_size=base_size, base_family=base_family)
+    + theme(plot.title = element_text(face = "bold",
+                                      size = rel(1.2), hjust = 0.5),
+            text = element_text(),
+            panel.background = element_rect(colour = NA),
+            plot.background = element_rect(colour = NA),
+            panel.border = element_rect(colour = NA),
+            axis.title = element_text(face = "bold",size = rel(1)),
+            axis.title.y = element_text(angle=90,vjust =2),
+            axis.title.x = element_text(vjust = -0.2),
+            axis.text = element_text(), 
+            axis.line = element_line(colour="black"),
+            axis.ticks = element_line(),
+            panel.grid.major = element_line(colour="#f0f0f0"),
+            panel.grid.minor = element_blank(),
+            legend.key = element_rect(colour = NA),
+            legend.position = "bottom",
+            legend.direction = "horizontal",
+            legend.key.size= unit(0.5, "cm"),
+            legend.margin = unit(0, "cm"),
+            legend.title = element_text(face="italic"),
+            plot.margin=unit(c(10,5,5,5),"mm"),
+            strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
+            strip.text = element_text(face="bold")
+    ))
+}

### 2017-08-28 data-raw/create_data_files.R

@@ -177,6 +177,14 @@ mc <- mc %>%
 mc <- mc %>%
   make_date_columns()
 
+# Clean up the insecticide column
+mc <-
+  mc %>%
+  mutate(insecticida = toupper(insecticida)) %>%
+  mutate(insecticida = ifelse(grepl('DDT', insecticida),
+                              'DDT',
+                              ifelse(grepl('ACT|ACY|AVT', insecticida), 'ACT', NA)))
+
 # Save for use in package
 devtools::use_data(mc,
                    overwrite = TRUE)
@@ -301,6 +309,18 @@ ab_panel$absent_sick <-
            ab_panel$leave_type == 'SIC', TRUE,
          ifelse(!ab_panel$absent, NA,
                 FALSE))
+
+# Get rid of the period before which we had any absences
+ab_panel <- ab_panel %>%
+  dplyr::filter(date >= min(ab$leave_from_date),
+                date <= max(ab$leave_to_date))
+
+# Get rid of everything before 2013, since it appears unreliable
+ab_panel <-
+  ab_panel %>%
+  dplyr::filter(date >= '2013-01-01') %>%
+  # and anything in 2017, which also appears untrustable
+  dplyr::filter(date <= '2016-12-31')
 
 devtools::use_data(ab_panel,
                    overwrite = TRUE)

### 2017-08-29 inst/rmd/maragra_preliminary_analysis.Rmd
adding census

@@
 ab_panel <- maragra::ab_panel
+census <- maragra::census
+census_sp <- maragra::census_sp
 clinic <- maragra::clinic

### 2017-08-29 data-raw/create_data_files.R

@@ -299,8 +299,8 @@ if('ab_expanded.RData' %in% dir()){
 
 # Join with absences
 ab_panel <- left_join(x = eligible_working_days,
-                    y = ab_expanded,
-                    by = c('oracle_number', 'date')) %>%
+                      y = ab_expanded,
+                      by = c('oracle_number', 'date')) %>%
   mutate(absent = ifelse(is.na(absent), FALSE, absent))
 
 # Define whether sick or not
@@ -323,6 +323,465 @@ ab_panel <-
   dplyr::filter(date <= '2016-12-31')
 
 devtools::use_data(ab_panel,
+                   overwrite = TRUE)
+
+# Census data ###############################
+if(!'census_done.RData' %in% dir('census_data')){
+  #### MANHICA CENSUS
+  # Get data from dssodk
+  if('manhica_census_data.RData' %in% dir('census_data')){
+    load('census_data/manhica_census_data.RData')
+  } else {
+    # Get manhica census data
+    HOUSEHOLD_ECONOMICS_CORE <-
+      cism::get_data(tab = 'HOUSEHOLD_ECONOMICS_CORE',
+                     dbname = 'dssodk')
+    INDIVIDUAL_DETAILS_CORE <-
+      cism::get_data(tab = 'INDIVIDUAL_DETAILS_CORE',
+                     dbname = 'dssodk')
+    LOCATION_DETAILS_CORE <-
+      cism::get_data(tab = 'LOCATION_DETAILS_CORE',
+                     dbname = 'dssodk')
+    location <-
+      cism::get_data(tab = 'location',
+                     dbname = 'openhds')
+    save(HOUSEHOLD_ECONOMICS_CORE,
+         INDIVIDUAL_DETAILS_CORE,
+         LOCATION_DETAILS_CORE,
+         location,
+         file = 'census_data/manhica_census_data.RData')
+  }
+  
+  # Clean up manhica data ---------------
+  
+  # Remove weirdness from column names
+  names(LOCATION_DETAILS_CORE) <- 
+    gsub('SEC1_|SEC2_|SEC3_|SEC4_|SEC5_', 
+         '', 
+         names(LOCATION_DETAILS_CORE))
+  
+  # Get smaller data
+  LOCATION_DETAILS_CORE <- 
+    LOCATION_DETAILS_CORE %>%
+    dplyr::select(LOCATION_ID,
+                  COVERAGE_MATERIAL,
[…]

### 2017-08-30 inst/rmd/maragra_preliminary_analysis.Rmd
big update with census

--- 2017-08-28 inst/rmd/maragra_preliminary_analysis.Rmd
@@
 ---
+title: "Maragra and Malaria: preliminary analysis"
 date: ""
@@
   html_document:
+    toc: True
     css: resources/style.css
@@
 
+<hr>
 <p style="text-align:left;">
-[www.economicsofmalaria.com](www.economicsofmalaria.com)
+[http://economicsofmalaria.com](http://economicsofmalaria.com)
 <span style="float:right;">joebrew@gmail.com</span>
 </p>
+<hr>
 
@@
 ab_panel <- maragra::ab_panel
+bairros <- maragra::bairros
+census <- maragra::census
 clinic <- maragra::clinic
@@
 ```
-
-# Maragra and Malaria: preliminary analysis {.tabset}
 
@@
            fill = key)) +
-  geom_area(stat = 'identity', position = 'stack', alpha = 0.5) +
+  geom_bar(stat = 'identity', position = 'stack', alpha = 0.5) +
   xlab('Date') +
@@
            y = pp)) +
-  geom_area(stat = 'identity', fill = 'darkblue', alpha = 0.5) +
+  geom_bar(stat = 'identity', fill = 'darkblue', alpha = 0.5) +
   xlab('Date') +
@@
            fill = severity)) +
-  geom_area(stat = 'identity',
+  geom_bar(stat = 'identity',
            position = 'stack',

### 2017-08-30 R/correct_unidade.R

@@ -1,0 +1,26 @@
+#' Correct the unidade
+#' 
+#' Try to find the right unidade of a non-conforming unidade
+#' @param unidade An unidade
+#' @param unidades A vector of all correct unidades
+#' @return The correct unidade
+#' @export
+
+correct_unidade <- function(unidade,
+                            unidades){
+  out <- unidade
+  if(nchar(unidade) == 8){
+    # Chop off the first
+    new_unidade <- substr(unidade, 1, 7)
+    if(new_unidade %in% unidades){
+      out <- new_unidade
+    } else {
+      # Chop off the last
+      new_unidade <- substr(unidade, 2, 8)
+      if(new_unidade %in% unidades){
+        out <- new_unidade
+      } 
+    }
+  }
+  return(out)
+}

### 2017-08-30 data-raw/create_data_files.R

@@ -6,6 +6,8 @@ library(cism)
 library(dplyr)
 library(ggplot2)
 library(readxl)
+library(rgdal)
+library(sp)
 
 # Initiate the raw-data into package session
 devtools::use_data_raw()
@@ -58,6 +60,10 @@ ab <- ab %>%
          date_captured = format_date(date_captured),
          date_signed_leave_form_rec = format_date(date_signed_leave_form_rec))
 
+# Keep only post 2013
+ab <- ab %>%
+  filter(leave_from_date >= '2013-01-01')
+
 # Keep only relevant columns
 ab <-
   ab %>%
@@ -79,9 +85,7 @@ ab <-
                 -captured_by_hospital_indicator,
                 -reason_denied)
 
-# Save for use in package
-devtools::use_data(ab,
-                   overwrite = TRUE)
+
 
 # clinic ##################################################
 
@@ -118,9 +122,6 @@ clinic <-
 clinic <- clinic %>%
   make_date_columns()
 
-# Save for use in package
-devtools::use_data(clinic,
-                   overwrite = TRUE)
 
 # clinic_agg ##################################################
 
@@ -138,12 +139,6 @@ clinic_agg <- clinic_agg %>% arrange(date)
 
 # Get rid of any unknowns
 clinic_agg <- clinic_agg %>% filter(!is.na(tested))
-
-# Save for use in package
-devtools::use_data(clinic_agg,
-                   overwrite = TRUE)
-
-
 
 # mc ###################################################
 
@@ -185,9 +180,8 @@ mc <-
                               'DDT',
                               ifelse(grepl('ACT|ACY|AVT', insecticida), 'ACT', NA)))
 
-# Save for use in package
-devtools::use_data(mc,
-                   overwrite = TRUE)
+# Add a 0 to the unidade (since they should all be in maciana)
+mc$unidade <- paste0('0', mc$unidade)
 
 
 # workers ##################################################
@@ -216,11 +210,9 @@ workers <-
          contract_end_date = as.Date(contract_end_date),
          company_entry_date = as.Date(company_entry_date))
 
-# Need to clean job_title and address through standardization !!!
+# Need to clean job_title through standardization !!!
 
[…]

### 2017-09-01 inst/rmd/maragra_preliminary_analysis.Rmd
major update

@@
 library(knitr)
-library(googleVis)
 library(RColorBrewer)
-op <- options(gvis.plot.tag='chart')
 
@@
 
+# Summary
+
 ## Description of data
@@
 
-kable(results)
+kable(results) %>%
+  kableExtra::kable_styling(full_width = FALSE)
 ```
@@
 
-## IRS analysis
+## IRS data
 
@@
 
-## Absenteeism analysis
+## Absenteeism data
 
@@
 
-## Clinic analysis
+## Clinic data
 
@@
 
-```{r, results = 'asis'}
-gg <- gvisLineChart(data = clinic_agg %>% filter(group == 'nacionais'),
-                    xvar = 'date',
-                    yvar = c('tested', 'positive'),
-                    options = list(curveType='function',
-                                  title="Malaria among Mozambican employees",
-                                  titleTextStyle="{color:'black',fontName:'Courier',fontSize:16}"))
-plot(gg)
-
-gg <- gvisLineChart(data = clinic_agg %>% filter(group == 'expatriados'),
-                    xvar = 'date',
-                    yvar = c('tested', 'positive'),
-                    options = list(curveType='function',
-                                  title="Malaria among expatriated employees",
-                                  titleTextStyle="{color:'black',fontName:'Courier',fontSize:16}"))
-plot(gg)
-
+```{r}
+ggplot(data = clinic_agg %>% filter(group == 'nacionais'),
+            aes(x = date,
+                y = tested)) +
+  geom_line() +
+  geom_line(aes(y = positive)) +
+  theme_maragra() +
+  labs(x = 'Date',
+       y = 'People (tested and positive)',
+       title = 'Mozambicans')
+ggplot(data = clinic_agg %>% filter(group != 'nacionais'),
+            aes(x = date,
+                y = tested)) +
+  geom_line() +
+  geom_line(aes(y = positive)) +
+  theme_maragra() +
+  labs(x = 'Date',
+       y = 'People (tested and positive)',
+       title = 'Expatriates')
 ```
@@
 
-### Seasonality analysis from `clinic_agg`
+### Seasonality data from `clinic_agg`
 
@@
 
-## Next steps
+# Analysis
 
-A lot of work remains to be done:
+## Effectiveness analysis
 
-- Figure out how to transform the `unidade` codes from the IRS data to specific workers's residences (or at least geographical area).  
-- Perform some name standardization so as to better link individual level clinical data with workers demographic and absenteeism data.  
-- Join data with census data for more demographic analysis, particularly in terms of absenteeism and health.  
-- Estimate a model for absenteeism, adjusting for relevant factors, particularly the IRS status and time since IRS.  
-- A cost-effectiveness framework for estimating the return on investment for IRS activities.
+### Identification strategy
+
+For our purposes we are analyzing the effect of one intervention (IRS) on 2 outcomes (absence and illness) with many confounders (age, worker type, seasonality, etc.).
+Our analysis can be visualized formulaically as follows:
+
+$$
+\begin{equation}
+\operatorname{Pr}(\text{Outcome} = 1 \mid \text{X}) = \beta_{0} + \beta_{1} \text{Location} + \beta_{2} \text{Season} + (\beta_3{IRS}*\beta_4{IRS_t} + ... )
+\end{equation}
+$$
+
+Our outcome is probabilistic and binomial (ie, one is either absent or present / infected or not infected).
+Our demographic confounders (represented by $...$) will be a function of iterative model selection.
+Our intervention (IRS) is not a simple yes/no, but rather the product of whether the residence of the worker in question was treated in the last year, and, if so, the time since treatment (represented above as the interaction term, where where $_t$ represents time elapsed since commencement of the most recent IRS campaign).
+
+### Propensity score matching
+
+Since the full accounting of confounders would greatly reduce the degrees of freedom of our analysis, we employ propensity score matching to generate a matched sample of workers who are alike in characteristics and time, but not treatment.
+We do this by first estimating the likelihood of having ever received the intervention, given a worker's age, sex, department and temporary vs.
+permanent status.
+We justify the necessity of this matching by noting that the differences between those workers who received IRS and those who did not (see table 1) are striking and in most cases statistically significant.
+
+```{r}
+right_side <- irs %>%
+  group_by(unidade) %>% 
+  tally %>%
+  ungroup %>%
+  mutate(received = ifelse(n > 0, TRUE, FALSE))
+
+left_side <-
+  workers %>%
+  dplyr::select(oracle_number, unidade, permanent_or_temporary, department, sex, date_of_birth)
+
+psm <- left_join(x = left_side,
+                 y = right_side,
+                 by = 'unidade') %>%
+  mutate(received = ifelse(is.na(received), FALSE, received)) %>%
+  mutate(age = round((as.numeric(as.Date('2016-01-01') - date_of_birth)) / 365.25, digits = 2)) %>%
+  filter(!is.na(sex),
+         !is.na(age)) %>%
+  dplyr::select(-unidade, -date_of_birth, -n)
+
+psmt1 <- psm
+names(psmt1) <- toupper(names(psmt1))
+psmt1 <- psmt1 %>%
+  mutate(RECEIVED = ifelse(RECEIVED == TRUE,'IRS', 'No IRS')) %>%
+  mutate(STATUS = PERMANENT_OR_TEMPORARY)
+
+pacman::p_load(tableone)
+table1 <- CreateTableOne(vars = toupper(c('STATUS', 'department', 'age',
+                                  'sex', 'received')), 
+                         data = psmt1, 
+                         factorVars = toupper(c('STATUS', 'department',
+                                  'sex')), 
+                         strata = 'RECEIVED')
+table1 <- print(table1, 
+                printToggle = FALSE, 
+                noSpaces = TRUE)
+kable(table1[,1:3],  
+      align = 'c', 
+      caption = 'Table 1: Comparison of unmatched samples') %>%
+  kableExtra::kable_styling(full_width = FALSE)
+```
+
+Having now demonstrated that are treatment and control groups are qualitatively different (and therefore require either statistical adjustment or a priori matching), we proceed to carry out the matching, using those best practices suggested by Ho et al 2004 "for improving parametric statistical models by preprocessing data with nonparametric matching methods" (Daniel Ho, Kosuke Imai, Gary King, and Elizabeth Stuart (2007).
+Matching as Nonparametric Preprocessing for Reducing Model Dependence in Parametric Causal Inference.
+Political Analysis 15(3): 199-236.
+http://gking.harvard.edu/files/abs/matchp-abs.shtml).
+We emply the nearest neighbor method for identifying those workers from our control group who most resemble those workers in the treatment group.
+
+Our match is a 1-to-1 cut, meaning those control workers who do not resemble those in the treatment group are left out of primary analysis. The below table shows the match results.
+
+```{r}
+library(MatchIt)
+set.seed(1234)
+match.it <- matchit(received ~ age + sex + permanent_or_temporary + department, data = psm, method="nearest", ratio=1)
+a <- summary(match.it)
+kable(a$nn, digits = 2, align = 'c', 
+      caption = 'Table 2: Sample sizes') %>%
+  kableExtra::kable_styling(full_width = FALSE)
+```
+
+The following output shows, that the distributions of our numeric variables are now extremely similar.
+
+```{r}
+kable(a$sum.matched[c(1,2,3,4)], digits = 2, align = 'c', 
+      caption = 'Table 3: Summary of balance for matched data') %>%
+  kableExtra::kable_styling(full_width = FALSE)
+```
+
+The propensity scores can be visualized below.
+
+```{r}
+plot(match.it, type = 'jitter', interactive = FALSE, col = adjustcolor('darkblue', alpha.f  =0.3))
+```
+
+```{r}
+# Save matched samples
+matched <- match.data(match.it)[1:ncol(psm)]
+```
+
+### Modeling
+
+Having now created a matched sample of `r nrow(matched)` workers, of which 50% received IRS and 50% did not, we can confidently carry out our analysis on this sample.
+Since the propensity score matching effectively cancels out demographic differences, our model only need take into account those differences which are not at the person-level.
+In our case, these include seasonality (defined here by quarter) (later, will add other factors).
+
+For the purposes of this first pass, we "bin" IRS exposure into 5 groups: never, ever but 180+ days ago, 90-80 days, 60-90 days, and in the last 60 days.
+
+Having estimated our binomial logistic regression model, we examine the odds ratios for absence as a function of our predictive variables.
+
+```{r}
+# Create model data
+model_data <-
+  ab_panel %>%
+  filter(oracle_number %in% matched$oracle_number) %>%
+  left_join(irs) %>%
+  left_join(matched) %>%
+  mutate(days_since = ifelse(is.na(days_since), 'Never',
+                             ifelse(days_since > 180, '180+',
+                                    ifelse(days_since > 90, '090-180',
+                                           ifelse(days_since > 60, '060-090',
+                                                  '<060'))))) %>%
+  mutate(quarter = lendable::quarter_extract(date)) %>%
+  mutate(quarter = as.character(quarter)) %>%
+  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))
+# Need to add seasonality
+# Need to adjust for houses on and off site
+fit <- glm(absent ~ days_since + quarter,
+           family = binomial('logit'),
+          data = model_data)
+x <- exp(coef(fit))
+ors <- exp(confint(fit))
+ors <- data.frame(ors)
+names(ors) <- c('Lower', 'Upper')
+
+ors <- cbind(x, ors)
+ors$Variable <- row.names(ors)
+ors <- ors %>%
+  dplyr::rename(OR = x) %>%
+  dplyr::select(Variable, OR, Lower, Upper)
+# Combine into a df
+kable(ors) %>%
+  kableExtra::kable_styling(full_width = FALSE)
+```
+
+We run the same model, but instead of estimating absences, we estimate only the likelihood of sick absences. The results (in form of odds ratios) are below.
+
+```{r}
+# Need to add seasonality
+# Need to adjust for houses on and off site
+fit <- glm(absent_sick ~ days_since + quarter,
+           family = binomial('logit'),
+          data = model_data)
+x <- exp(coef(fit))
+ors <- exp(confint(fit))
+ors <- data.frame(ors)
+names(ors) <- c('Lower', 'Upper')
+
+ors <- cbind(x, ors)
+ors$Variable <- row.names(ors)
+ors <- ors %>%
+  dplyr::rename(OR = x) %>%
+  dplyr::select(Variable, OR, Lower, Upper)
+# Combine into a df
+kable(ors) %>%
+  kableExtra::kable_styling(full_width = FALSE)
+
+```
+
+## To do
+
+- More robust models and checks.  
+- Limit ton only clinical malaria. 
+- Un-bin IRS (ie, estimate semi-parametric curve of effectiveness waning)  
+- More investigation of worker type, shift, etc.

### 2017-09-01 inst/rmd/paper.Rmd
major update

---
title: Effectiveness of private sector malaria control at a large sugar
  facility in Southern Mozambique
csl: elsevier-harvard.csl
date: ''
output:
  html_document:
    css: resources/style.css
    pandoc_args:
    - +RTS
    - -K16000m
    - -RTS
    toc: yes
  pdf_document:
    toc: yes
fig_width: 4
fig_height: 2.6
bibliography: bibliography.bib
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(knitr)
library(googleVis)
library(RColorBrewer)
op <- options(gvis.plot.tag='chart')

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

<!-- <hr> -->
<!-- <p style="text-align:left;"> -->
<!-- [http://economicsofmalaria.com](http://economicsofmalaria.com) -->
<!-- <span style="float:right;">joebrew@gmail.com</span> -->
<!-- </p> -->
<!-- <hr> -->

```{r}
# Read in data
ab <- maragra::ab
ab_panel <- maragra::ab_panel
bairros <- maragra::bairros
census <- maragra::census
clinic <- maragra::clinic
clinic_agg <- maragra::clinic_agg
mc <- maragra::mc
workers <- maragra::workers
```

# Introduction

Public health interventions targetting malaria - and their corresponding cost-effectiveness evaluations - most often pertain to the public sector. Accordingly, the analytical framework and metrics employed by these analyses most often focus on impact pertaining to the public good, such as an increase in life years adjusted for disability or quality  [@Goodman1999, @Shretta2016, @Lee2017, Hanson2004]. Though population-level health is certainly of importance to businesses, and improvements in health incidentally improve the economy at all levels [@World1999, @Bloom2008], these improvements may be to disperse or long-term to incentivize private sector involvement in health campaigns.

In the context of Mozambique, where a significant sector of the economy is dominated by a full large-scale foreing direct investment projects [@Robbins2012], the role of the private sector in health generally, and malaria specifically, is unequivocally important. Large agriculture and extractive industry firms take up wide swaths of land and employ hundreds of thousands  [@German2013]. The Mozambican state has encouraged large-scale entreprise with the aim of general economic development [@Buur2012]. And where large firms exist, they often take on social roles such as housing and health care [@Winkler]. At times, this role is necessary from a purely practical standpoint; in other cases, it is employed under the guise of "corporate social responsibility" [@Azemar2009]. Regardless of the language used, it is clear that private industry plays an important role in public health in Mozambique [@Robbins2012, @CastelBranco2014].

Several cases exist of foreign firms engaging in large-scale anti-malaria campaigns [@Mouzin2011, @Han]. But these cases rarely undergo the same rigorous evaluation as their public sector counterparts due to (a) unavailability of data, (b) identification strategy, etc.

In this paper, we...

We hypothesize that..

We find that...

# Methods

```{r}
# # Create model data
# model_data <-
#   ab_panel %>%
#   left_join(irs) %>%
#   left_join(workers %>%
#               dplyr::select(oracle_number,
#                             sex,
#                             race_code,
#                             department,
#                             permanent_or_temporary)) %>%
#   mutate(days_since = ifelse(is.na(days_since), 'Never',
#                              ifelse(days_since > 180, '180+',
#                                     ifelse(days_since > 90, '090-180',
#                                            ifelse(days_since > 60, '060-090', 
#                                                   '<060')))))
# # Need to add seasonality
# # Need to adjust for houses on and off site
# fit <- glm(absent ~ days_since,
#            family = binomial('logit'),
#           data = model_data)
# exp(coef(fit))
```

Some text goes here.

All data processing and analysis were carried out in R [@R]. All code is freely available online [@brewgit].

# Results

Some text goes here

# Discussion

Some text goes here

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'knitr', 'rmarkdown'
), 'packages.bib')
```

# References

### 2017-09-01 R/generate_paper.R

@@ -1,0 +1,41 @@
+#' Generate paper
+#' 
+#' Generate the paper
+#' originally prepared in August / September 2017
+#' @param date A date to be printed on the report. If \code{NULL} (the 
+#' default), the current date will be used
+#' @param output_dir The directory to which the file should be written. If 
+#' \code{NULL} (the default), the current working directory will be used.
+#' @param output_file The name of the file to be written. 
+#' @return An html will be written
+#' @importFrom rmarkdown render
+#' @export
+
+generate_paper <- function(date = NULL,
+                           output_dir = NULL,
+                           output_file = 'paper.html'){
+  
+  # If no output directory, make current wd
+  if(is.null(output_dir)){
+    output_dir <- getwd()
+  }
+  
+  # If not date, use today's
+  if(is.null(date)){
+    date <- Sys.Date()
+  }
+  
+  # Combine parameters into a list, so as to pass to Rmd
+  parameters <- list(date = date)
+  
+  # Find location the rmd to knit
+  file_to_knit <- 
+    system.file('rmd/paper.Rmd', 
+                package='maragra')
+  
+  # Knit file
+  rmarkdown::render(file_to_knit,
+                    output_dir = output_dir,
+                    output_file = output_file,
+                    params = parameters)
+}

### 2017-09-01 R/run_all_analyses.R

@@ -34,4 +34,10 @@ run_all_analyses <- function(date = NULL,
   generate_preliminary_analysis(date = date,
                                 output_dir = output_dir)
   
+  # Clinical preliminary analysis report
+  message(paste0('Generating the paper in ',
+                 output_dir))
+  generate_paper(date = date,
+                 output_dir = output_dir)
+  
 }

### 2017-09-01 data-raw/create_data_files.R

@@ -1043,13 +1043,82 @@ mc <-
   left_join(mc,
             coordenadas)
 
-# Save again for use in package
+# Get IRS expanded data
+# Generate a dataset of irs coverage and time since irs
+expand_irs <- function(date,
+                       chemical,
+                       unidade){
+  data_frame(date = seq(date,
+                        date + 364,
+                        by = 1),
+             chemical = chemical,
+             unidade = unidade,
+             days_since = 0:364) 
+}
+results <- list()
+for (i in 1:nrow(mc)){
+  out <- expand_irs(date = mc$date[i],
+                    chemical = mc$insecticida[i],
+                    unidade = mc$unidade[i])
+  results[[i]] <- out
+  message(i)
+}
+
+# Combine
+irs <- bind_rows(results)
+
+# Group by date and unidade, and get the highest coverage value (for when there are overlaps)
+irs <- 
+  irs %>%
+  group_by(date, unidade) %>%
+  summarise(days_since = min(days_since),
+            chemical = dplyr::first(chemical[days_since == min(days_since)]))
+irs <- irs %>% ungroup
+
+# Bring an unidade into workers
+# only merging on those with an acceptably low match
+# of <= 0.1990741
+workers <- 
+  workers %>%
+  left_join(census %>%
+              dplyr::select(unidade,
+                            perm_id))
+
+# Generate some extra vars for workers
+workers <- 
+  workers %>%
+  mutate(permanent_or_temporary = 
+           ifelse(employee_indicator_type_desc %in% c('TEMPORARY AGRIC', 
+                                                      'FTC INDUSTRIAL'),
+                                                      'Temporary',
+                                                      'Permanent')) %>%
+  mutate(department = 
+           ifelse(grepl('CIVLS|HUMAN RESOURCES|RISK', department_name),
+                  'Administrative',
+                  ifelse(grepl('CANE|ADMIN', department_name),
+                                'Field',
+                               'Factory')))
+
+# Get rid of duplicates
+workers <- 
+  workers %>%
+  dplyr::filter(!duplicated(oracle_number))
+
+# Bring unidade into ab_panel
+ab_panel <-
+  left_join(x = ab_panel,
+            y = workers %>%
+              dplyr::filter(census_name_match_score <= 0.1990741) %>%
+              dplyr::select(oracle_number, unidade))
+
 # Save for use in package
 devtools::use_data(ab,
                    overwrite = TRUE)
 devtools::use_data(clinic,
[…]

### 2017-09-06 inst/rmd/aes_maragra.Rmd
major update

---
title: "Malaria and Sugar"
header-includes:
  - \widowpenalties 1 150
subtitle: "Effectiveness of private and public sector malaria control campaigns in large sugar
  facilities in Southern Mozambique"
author: "AES, Barcelona, 2017"
fig_height: 5
fig_width: 7
bibliography: bibliography.bib
output:
  beamer_presentation:
    theme: "Berlin"
    colortheme: "seahorse"
    fonttheme: "structuresmallcapsserif"
    toc: true
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}

# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(knitr)
library(googleVis)
library(RColorBrewer)
op <- options(gvis.plot.tag='chart')

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

```{r}
# Read in data
ab <- maragra::ab
ab_panel <- maragra::ab_panel
bairros <- maragra::bairros
census <- maragra::census
clinic <- maragra::clinic
clinic_agg <- maragra::clinic_agg
mc <- maragra::mc
workers <- maragra::workers
```

```{r}
tablify <- function(x, n = 5){
  DT::datatable(x,
                selection = 'none',
                escape = FALSE,
                options = list(#sDom  = '<"top">lrt<"bottom">ip',
                               pageLength = n, 
                               dom = 'tip'))
}
kablify <- function(x, size = 12){
  require(kableExtra)
  kable(x,
        digits = 3) %>% 
  kable_styling(bootstrap_options = c("striped",
  "hover",
  "condensed"),
  font_size = size,
  position = 'float_left',
  full_width = FALSE)
}

make_kable <- function(x,
                       caption = NULL,
                       font_size = 12){
  require(kableExtra)
  require(knitr)
  kable(x,
        booktabs = TRUE,
        format = 'latex',
        digits = 2,
        row.names = FALSE,
        caption = caption,
        format.args = list(big.mark = ',')) %>%
    kable_styling(font_size = font_size)
  #%>%
  # kable_styling(latex_options = "scale_down")
}

# ## 
# 
# <img src="img/tree.jpg" alt="img/tree.jpg" style="width: 600px; align="center"/>
```

# Introduction

## The effect of malaria on the economy

- Traditional attempts to quantify effect are too clinical - basically just take the clinical cost of an illness and multiply by number of illnesses.  
- But there are microeconomic effects (short term changes in behavior when ill, acquisition of human capital) and macroeconomic affects (trade, productivity, tourism) that come as a result of pandemic malaria [@Sachs2002].  
- Impossible to truly capture the macro effects until you capture the micro effects.

## The importance of large foreign firms in Mozambique

- Significant sector of the economy is dominated by a full large-scale foreing direct investment projects [@Robbins2012, @German2013].   
- The Mozambican state has encouraged large-scale entreprise with the aim of general economic development [@Buur2012], partly for reasons of practicality, and partly out of a general notion of "corporate social responsibility" [@Azemar2009].  
- Indirectly, private industry plays an important role in public health in Mozambique [@Robbins2012, @CastelBranco2014].

## The effect of malaria on business

- Difficult to study because data are privately held and there is resistance to sharing.  
- Even when data are accessible, it's very hard to come up with suitable comparisons. - Observable effects (illness, etc.), and unobservable effects (lack of investment, etc.).

# Context

## Manhiça and Magude

```{r}
library(sp)
a <- cism::mag2; a@data <- data.frame(x = 'Magude'); row.names(a) <- '1'
b <- cism::man2; b@data <- data.frame(x = 'Manhiça'); row.names(b) <- '2'
x <- rbind(a, b)
par(fmrow = c(1,2))
plot(cism::moz0)
plot(x, add = TRUE, col = c('red', 'yellow'))
plot(x, col = c('red', 'yellow'))
par(mfrow = c(1,1))
```

## Two large sugar companies

Maragra and Xinavane

<!-- <img src="img/maragra.png" alt="img/maragra.png" style="width: 350px; align="center"/> -->
<!-- <img src="img/xinavane.png" alt="img/xinavane.png" style="width: 430px; align="center"/> -->

\includegraphics[width=0.46\textwidth]{img/maragra}
\includegraphics[width=0.54\textwidth]{img/xinavane}

## Our questions

- What is the effect of the malaria elimination campaign on worker absenteeism and health (and productivity)?  
- What is the effect of privately-run malaria control campaigns on worker absenteeism and health (and productivity)?    
- How much of a private good comes from public interventions?   
- How much public good comes from private interventions?

# The data

## Data from 2 large sugarcane facilities

### Xinavane

```{r}
xinavane <- readr::read_csv('../../../xinavane/xinavane_monthly_panel_with_census.csv')
xinavane <- xinavane %>%
  filter(month_start <= '2016-06-01') %>%
  mutate(eligibles = as.numeric(eligibles))

# Get absenteeism over time
xa <- xinavane %>%
  dplyr::group_by(month_start) %>%
  dplyr::summarise(absences = sum(absences, na.rm = TRUE),
                   sick_absences = sum(as.numeric(sick_absences), na.rm = TRUE),
            eligibles = sum(eligibles, na.rm = TRUE))

# Adjust denominator
xa$eligibles[xa$month_start >= '2016-01-01'] <- max(xa$eligibles[xa$month_start < '2016-01-01'])

xa$absenteeism_rate <- 
  xa$absences / 
  xa$eligibles * 100
xa$sick_absenteeism_rate <-
  xa$sick_absences /
  xa$eligibles * 100
```

- `r nrow(xinavane)` worker-days observed (`r round(nrow(xinavane) / 365.25)` years of human activity).    
- Work schedules, sociodemographic information, absences, sickness-attributed absences, worker residence, type of worker, IRS*...

### Maragra

- `r nrow(ab_panel)` worker-days observed (`r round(nrow(ab_panel) / 365.25)` years of human activity).    
- Work schedules, sociodemographic information, absences, sickness-attributed absences, malaria laboratory tests, worker residence, type of worker, IRS...

## Comparing the two sugarcanes over time

What is the effect of the elimination campaign on worker absenteeism?

```{r}
ma <- ab_panel %>%
  group_by(month_start = as.Date(paste0(format(date, '%Y-%m-'), '01'))) %>%
  dplyr::summarise(absences = sum(absent, na.rm = TRUE),
                   sick_absences = sum(absent_sick, na.rm = TRUE),
            eligibles = length(absent)) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) %>%
  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100)
combined <- bind_rows(
  ma %>% mutate(Company = 'Maragra'),
  xa %>% mutate(Company = 'Xinavane')
)

cols <- rev(c('darkred', 'darkorange'))
g1 <- ggplot(data = combined,
       aes(x = month_start,
           y = absenteeism_rate,
           color = Company)) +
  geom_line() +
  scale_color_manual(name = 'Company',
                     values = cols) +
  theme_maragra() +
  labs(x = 'Month',
       y = 'Absenteeism rate',
       title = 'All absenteeism')
g2 <- ggplot(data = combined,
       aes(x = month_start,
           y = sick_absenteeism_rate,
           color = Company)) +
  geom_line() +
  scale_color_manual(name = 'Company',
                     values = cols) +
  theme_maragra() +
  labs(x = 'Month',
       y = 'Absenteeism rate',
       title = 'Sick absenteeism')
Rmisc::multiplot(g1, g2)
```

## Standardizing

What is the effect of the elimination campaign on worker absenteeism?

```{r}
# Set ot relative
combined <-
  combined %>%
  filter(month_start >= '2014-01-01') %>%
  group_by(Company) %>%
  mutate(pre_intervention_avg_absenteeism = sum(absences[month_start <= '2015-10-01']) /
           sum(eligibles[month_start <= '2015-10-01']) * 100,
         pre_intervention_avg_sick_absenteeism = sum(sick_absences[month_start <= '2015-10-01']) / 
           sum(eligibles[month_start <= '2015-10-01']) * 100) %>%
  ungroup %>%
  mutate(p_absenteeism = absenteeism_rate / pre_intervention_avg_absenteeism * 100,
         p_absenteeism_sick = sick_absenteeism_rate / pre_intervention_avg_sick_absenteeism * 100)

g1 <- ggplot(data = combined,
       aes(x = month_start,
           y = p_absenteeism,
           color = Company)) +
  geom_line() +
  scale_color_manual(name = 'Company',
                     values = cols) +
  theme_maragra() +
  labs(x = 'Month',
       y = '% of pre-interevention rate',
       title = 'All absenteeism (scaled)') +
  geom_hline(yintercept = 100, 
             color = 'grey',
             alpha = 0.6,
             lty = 2) +
  geom_vline(xintercept = as.numeric(as.Date('2015-10-01')))

g2 <- ggplot(data = combined,
       aes(x = month_start,
           y = p_absenteeism_sick,
           color = Company)) +
  geom_line() +
  scale_color_manual(name = 'Company',
                     values = cols) +
  theme_maragra() +
  labs(x = 'Month',
       y = '% of pre-interevention rate',
       title = 'Sick absenteeism (scaled)') +
  geom_hline(yintercept = 100, 
             color = 'grey',
             alpha = 0.6,
             lty = 2) +
  geom_vline(xintercept = as.numeric(as.Date('2015-10-01')))
Rmisc::multiplot(g1, g2)
```

## Bias - residence of worker

```{r}
# loc <- xinavane %>%
#   group_by(location_laia, month_start) %>%
#   dplyr::summarise(absences = sum(absences, na.rm = TRUE),
#                    sick_absences = sum(as.numeric(sick_absences), na.rm = TRUE),
#             eligibles = sum(eligibles, na.rm = TRUE))
# 
# # Adjust denominator
# loc$eligibles[loc$month_start >= '2016-01-01'] <- max(loc$eligibles[loc$month_start < '2016-01-01'])
# 
# loc$absenteeism_rate <- 
#   loc$absences / 
#   loc$eligibles * 100
# loc$sick_absenteeism_rate <-
#   loc$sick_absences /
#   loc$eligibles * 100
# 
# loc <- loc %>% filter(!is.na(location_laia))
# 
# ggplot(data = loc,
#        aes(x = month_start,
#            y = sick_absenteeism_rate,
#            color = location_laia)) +
#   geom_line()
x <- xinavane %>%
  group_by(Residence = location_laia) %>%
  summarise(Workers = length(unique(number))) %>%
  ungroup %>%
  mutate(Residence = ifelse(is.na(Residence), 
                            'Unknown',
                            Residence)) %>%
  mutate(Percentage = Workers / sum(Workers) * 100)
knitr::kable(x)
```

## Current status

- Obtaining and cleaning supplementary data  
- Refining analysis approaches

# Maragra

## The data

## IRS data - insecticide chemical

```{r}
ggplot(data = mc,
       aes(x = insecticida)) +
  geom_bar(fill = '#159957',
           alpha = 0.6) +
  theme_maragra() +
  labs(x = 'Insecticide',
       y = 'IRS activities',
       title = 'Maragra IRS activity by insecticide type')
```

## IRS data - insecticide use over time

```{r}
x <- mc %>%
  group_by(date = as.Date(paste0(format(date, '%Y-%m'), '-01')),
           insecticida) %>%
  summarise(houses = sum(casas_cobertas)) %>%
  ungroup
# Get a left side for filling in the 0s
left <- expand.grid(date = seq(min(x$date),
                               max(x$date),
                               by = 'month'),
                    insecticida = sort(unique(x$insecticida)))
# Join together
x <- left_join(x = left,
               y = x,
               by = c('date', 
                      'insecticida'))
# Replace NAs with 0
x <- x %>%
  mutate(houses = ifelse(is.na(houses), 0, houses))
# Define a color vector
cols <- colorRampPalette(brewer.pal(n = 9, 'Spectral'))(length(unique(x$insecticida)))

ggplot(data = x,
       aes(x = date,
           y = houses,
           group = insecticida,
           color = insecticida)) +
  geom_point(alpha = 0.6) +
  geom_line(alpha = 0.6) +
  theme_maragra() +
  labs(x = 'Month',
       y = 'Houses fumigated',
       title = 'Maragra IRS activity by insecticide type over time') +
  scale_color_manual(name = 'Chemical',
                     values = cols) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2012:2017, '-01-01'))),
             alpha = 0.2)
```

## Absenteeism data

```{r}
x <- 
  ab_panel %>%
  group_by(date) %>%
  summarise(absences = length(which(absent)),
            eligibles = length(absent)) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) %>%
  mutate(Workers = eligibles)

ggplot(data = x,
       aes(x = date,
           y = absenteeism_rate)) +
  geom_point(alpha = 0.1,
             color = '#159957',
             aes(size = Workers)) +
  geom_smooth() +
  theme_maragra() +
  labs(x = 'Date',
       y = 'Absenteeism rate',
       title = 'Absenteeism rate over time')

```

## Clinical microscopy data

```{r}
x <- 
  clinic_agg %>%
  group_by(date) %>%
  summarise(p = sum(positive),
            pp = sum(positive) / sum(tested) * 100,
            tested = sum(tested))

# Gather into long format
y <- 
  rbind(x %>% dplyr::select(date, p) %>%
          rename(val = p) %>%
          mutate(key = 'positive'),
        x %>% dplyr::select(date, tested) %>%
          rename(val = tested) %>%
          mutate(key = 'tested')) %>%
  mutate(key = factor(key, levels = c('tested', 'positive')))

ggplot(data = x,
       aes(x = date,
           y = p)) +
  geom_bar(stat = 'identity') +
  xlab('Date') +
  ylab('Number of positive cases') +
    theme_maragra() +
  ggtitle('Malaria miscroscopy positive results',
          'All workers')
```

## Clinical microscopy data (b)

```{r}
ggplot(data = y,
       aes(x = date,
           y = val,
           group = key,
           fill = key)) +
  geom_bar(stat = 'identity', position = 'stack', alpha = 0.5) +
  xlab('Date') +
  ylab('Number of positive cases') +
  theme_maragra() +
  scale_fill_manual(name = '',
                    values = c('black', 'darkred')) +
  ggtitle('Malaria miscroscopy tests and positive results',
          'All workers, absolute') +
    theme(title = element_text(size = 12))
```

## Clinical microscopy data (c)

```{r}
ggplot(data = clinic,
       aes(x = severity)) +
  geom_bar(alpha = 0.6,
           fill = 'darkorange') +
  theme_maragra() +
  labs(x = 'Severity',
       y = 'Cases',
       title = 'Severity of all malaria cases: Maragra clinic')
```

## Identification strategy

- One intervention (IRS).  
- 2 outcomes (absence and illness).  
- Many confounders (age, worker type, seasonality, etc.).

\begin{equation}
\operatorname{Pr}(\text{Outcome} = 1 \mid \text{X}) = \beta_{0} + \beta_{1} \text{Location} + \beta_{2} \text{Season} + (\beta_3{IRS}*\beta_4{IRS_t} + ... )
\end{equation}

## Propensity score matching

- Direct adjustment for confounders would greatly reduce the degrees of freedom of our analysis.  
- Instead, we employ propensity score matching to generate a matched sample of workers who are alike in characteristics and time, but not treatment.

- Based on estimation of probability of treatment given a worker's age, sex, department and temporary vs. permanent status.

## Propensity score matching - justification

```{r}
right_side <- irs %>%
  group_by(unidade) %>% 
  tally %>%
  ungroup %>%
  mutate(received = ifelse(n > 0, TRUE, FALSE))

left_side <-
  workers %>%
  dplyr::select(oracle_number, unidade, permanent_or_temporary, department, sex, date_of_birth)

psm <- left_join(x = left_side,
                 y = right_side,
                 by = 'unidade') %>%
  mutate(received = ifelse(is.na(received), FALSE, received)) %>%
  mutate(age = round((as.numeric(as.Date('2016-01-01') - date_of_birth)) / 365.25, digits = 2)) %>%
  filter(!is.na(sex),
         !is.na(age)) %>%
  dplyr::select(-unidade, -date_of_birth, -n)

psmt1 <- psm
names(psmt1) <- toupper(names(psmt1))
psmt1 <- psmt1 %>%
  mutate(RECEIVED = ifelse(RECEIVED == TRUE,'IRS', 'No IRS')) %>%
  mutate(STATUS = PERMANENT_OR_TEMPORARY)

pacman::p_load(tableone)
table1 <- CreateTableOne(vars = toupper(c('STATUS', 'department', 'age',
                                  'sex', 'received')), 
                         data = psmt1, 
                         factorVars = toupper(c('STATUS', 'department',
                                  'sex')), 
                         strata = 'RECEIVED')
table1 <- print(table1, 
                printToggle = FALSE, 
                noSpaces = TRUE)
# table1$Variable <- row.names(table1)
x <- data.frame(table1[,1:3])
x$Variable <- row.names(x)
x <- x[,c('Variable', 'IRS', 'No.IRS', 'p')]
row.names(x) <- NULL
knitr::kable(x,
           caption = 'The need for propensity score matching') %>%
  kableExtra::kable_styling(font_size = 12)
```

## Match results

Sample sizes

```{r}
library(MatchIt)
set.seed(1234)
match.it <- matchit(received ~ age + sex + permanent_or_temporary + department, data = psm, method="nearest", ratio=1)
a <- b <- summary(match.it)
a <- data.frame(a$nn)
a$Status <- row.names(a)
a <- a[,c('Status', 'Control', 'Treated')]
```

```{r}
knitr::kable(a)
```

## Group similarity

Summary of balance for matched data

```{r}
b <- data.frame(b$sum.matched[c(1:4)])
rn <- names(b)
b$Variable <- row.names(b)
b <- b[,c('Variable', rn)]
row.names(b) <- NULL
b$Variable[b$Variable == 'permanent_or_temporaryTemporary'] <- 'Temporary'
b$Variable[b$Variable == 'departmentFactory'] <- 'Factory'
b$Variable[b$Variable == 'departmentField'] <- 'Field'
b$SD.Control <- NULL
knitr::kable(b, digits = 2, format.args = list(big.mark = ',')) 
```

## Group similarity

```{r}
plot(match.it, type = 'jitter', interactive = FALSE, col = adjustcolor('darkblue', alpha.f  =0.3))
```

```{r}
# Save matched samples
matched <- match.data(match.it)[1:ncol(psm)]
```

## Visual results

```{r}
if('aes_maragra_temp.RData' %in% dir()){
  load('aes_maragra_temp.RData')
} else {
  # Create model data
  model_data <-
    ab_panel %>%
    filter(oracle_number %in% matched$oracle_number) %>%
    left_join(irs) %>%
    left_join(matched) %>%
    mutate(days_since_numeric = days_since) %>%
    mutate(months_since = days_since %/% 30) %>%
    mutate(months_since = ifelse(is.na(months_since) | is.infinite(months_since) | months_since >= 10,
                                 '>9',
                                 as.character(months_since))) %>%
    mutate(days_since = ifelse(is.na(days_since), 'Never',
                               ifelse(days_since > 180, '180+',
                                      ifelse(days_since > 90, '090-180',
                                             ifelse(days_since > 30, '030-090',
                                                    '<030'))))) %>%
    mutate(quarter = lendable::quarter_extract(date)) %>%
    mutate(quarter = as.character(quarter)) %>%
    mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) %>%
    # Get census data
    left_join(workers %>%
                dplyr::select(perm_id,
                              oracle_number)) %>%
    left_join(census %>%
                dplyr::select(perm_id, 
                              maragra_bairro,
                              maragra_fabrica)) %>%
    # Make season
    mutate(season = ifelse(quarter == '1', '1. Jan-Mar',
                           ifelse(quarter == '2', '2. Apr-Jun',
                                  ifelse(quarter == '3', '3. Jul-Sep',
                                         ifelse(quarter == '4', '4. Oct-Dec', 
                                                NA))))) %>%
    mutate(simple_season = ifelse(quarter %in% c('1'), 'High', 'Low'))

  # Need to add seasonality
  # Need to adjust for houses on and off site
  fit <- glm(absent ~ days_since + simple_season,
             family = binomial('logit'),
            data = model_data)
  x <- exp(coef(fit))
  ors <- exp(confint(fit))
  ors <- data.frame(ors)
  names(ors) <- c('Lower', 'Upper')

  ors <- cbind(x, ors)
  ors$Variable <- row.names(ors)
  ors <- ors %>%
    dplyr::rename(OR = x) %>%
    dplyr::select(Variable, OR, Lower, Upper)

  save(model_data,
       fit,
       x,
       ors,
       file = 'aes_maragra_temp.RData')
}

```

```{r}
x <- model_data %>%
  filter(maragra_bairro) %>%
  # filter(days_since != 'Never') %>%
  group_by(months_since, 
           oracle_number) %>%
  summarise(absences = length(which(absent)),
            sick_absences = length(which(absent_sick)),
            eligibles = length(absent)) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100,
         sick_absenteeism_rate = sick_absences / eligibles * 100)

ggplot(data = x %>%
         filter(months_since != '>9'),
       aes(x = months_since,
           y = sick_absenteeism_rate)) +
  geom_jitter(alpha = 0.3) +
  geom_boxplot(alpha = 0.6,
               fill = NA) +
  scale_y_log10() +
  theme_maragra() +
  labs(x = 'Months since IRS',
       y = 'Sick absenteeism rate (log scale)') +
  geom_smooth(aes(group = 1))
```

## Visual results (b)

```{r}
y <- model_data %>%
  group_by(days_since_numeric) %>%
    summarise(absences = length(which(absent)),
            sick_absences = length(which(absent_sick)),
            eligibles = length(absent),
            Workers = length(unique(oracle_number))) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100,
         sick_absenteeism_rate = sick_absences / eligibles * 100)

g1 <- ggplot(data = y,
       aes(x = days_since_numeric,
           y = sick_absenteeism_rate)) +
  geom_point(aes(size = Workers),
             alpha = 0.4) +
  geom_smooth() +
  theme_maragra() +
  labs(x = 'Days since IRS',
       y = 'Sick absenteeism rate')
g1
# g2 <- ggplot(data = y,
#        aes(x = days_since_numeric,
#            y = absenteeism_rate)) +
#   geom_point() +
#   geom_smooth() +
#   theme_maragra() +
#   labs(x = 'Days since IRS',
#        y = 'Absenteeism rate')
# Rmisc::multiplot(g1, g2)
```

## Bias - time of year

```{r}
x <- mc %>%
  group_by(year, day_number) %>%
  summarise(houses = sum(casas_cobertas)) %>%
  ungroup %>%
  mutate(fake_date = as.Date('1990-01-01') + day_number) %>%
  mutate(fake_month = as.Date(paste0(format(fake_date, '%Y-%m'), '-01'))) %>%
  group_by(year, fake_month) %>%
  summarise(houses = sum(houses)) %>%
  ungroup %>%
  mutate(year = factor(year))

# Fill the nas with 0s
left <- expand.grid(year = sort(unique(x$year)),
                    fake_month = sort(unique(x$fake_month)))

x <- left_join(x = left,
               y = x)
x$houses[is.na(x$houses)] <- 0
cols <- colorRampPalette(brewer.pal(n = 9, 'Spectral'))(length(unique(x$year)))
ggplot(data = x %>% 
         filter(!year %in% c(2011:2012, 2017)),
       aes(x = fake_month,
           y = houses)) +#,
           # group = year,
           # color = year)) +
  geom_line(alpha = 0.8) +
  # scale_color_manual(name = 'Year',
  #                    values = cols) +
  scale_x_date(labels = scales::date_format("%b")) +
  theme_maragra() +
  labs(x = 'Month',
       y = 'Houses covered',
       title = 'Seasonality of IRS operations') +
  facet_wrap(~year)
```

## Modeling

Since IRS is not random throughout the year, we need to adjust for seasonality.

```{r}
# Combine into a df
row.names(ors) <- NULL
kablify(ors)
```

## Sick absence only

We run the same model, but instead of estimating absences, we estimate only the likelihood of sick absences. The results (in form of odds ratios) are below.

```{r}
if('aes_maragra_temp2.RData' %in% dir()){
  load('aes_maragra_temp2.RData')
} else {

  # Need to add seasonality
  # Need to adjust for houses on and off site
  fit <- glm(absent_sick ~ days_since + simple_season,
             family = binomial('logit'),
            data = model_data)
  x <- exp(coef(fit))
  ors <- exp(confint(fit))
  ors <- data.frame(ors)
  names(ors) <- c('Lower', 'Upper')

  ors <- cbind(x, ors)
  ors$Variable <- row.names(ors)
  ors <- ors %>%
    dplyr::rename(OR = x) %>%
    dplyr::select(Variable, OR, Lower, Upper)
  save(fit,
       x,
       ors,
       file = 'aes_maragra_temp2.RData')
}

# Combine into a df
row.names(ors) <- NULL
knitr::kable(ors)
```

## Challenges

- Methodological: How to define treatment better 
  -When should we consider the effect of IRS to begin? 
  -Should we quantify waning IRS a priori or solely through observation?  
- Conceptual: How to better construct our panel  
  -One person is observed in different treatment states at different times, and to different degrees  
  - Matching with others vs. matching with self  
- Identification: How to better control for bias  
  - Clinical malaria is different from malaria  
  - Treatment is not random, but its causal factors are difficult to observe  
  - Malaria is seasonal, making it hard to parse the effect of time passing (ie, high season ending) from IRS waning

## References {.allowframebreaks}

### 2017-09-06 inst/rmd/maragra_preliminary_analysis.Rmd
major update

@@
   scale_color_manual(name = 'Chemical',
-                     values = cols)
+                     values = cols) +
+  geom_vline(xintercept = as.numeric(as.Date(paste0(2012:2017, '-01-01'))),
+             alpha = 0.2)
 ```

### 2017-09-06 R/run_all_analyses.R

@@ -1,6 +1,8 @@
 #' Run all analyses
 #' 
 #' Run all analyses and produce outputs in the \code{outputs_dir} folder
+#' @param only A vector of .Rmd file names (excluding path) to be knitted
+#' @param except A vector or .Rmd file names (excluding path) to be excluded
 #' @param date A date to be printed on the report. If \code{NULL} (the 
 #' default), the current date will be used
 #' @param output_dir The directory to which the analyses should be written. If 
@@ -9,7 +11,9 @@
 #' @importFrom rmarkdown render
 #' @export
 
-run_all_analyses <- function(date = NULL,
+run_all_analyses <- function(only = NULL,
+                             except = NULL,
+                             date = NULL,
                              output_dir = NULL){
   
   # If no output directory, make current wd
@@ -22,22 +26,59 @@ run_all_analyses <- function(date = NULL,
     date <- Sys.Date()
   }
   
-  # Clinical malaria incidence report
-  message(paste0('Generating the clinical malaria incidence report in ',
-                 output_dir))
-  generate_maragra_clinical_malaria_incidence(date = date,
-                                              output_dir = output_dir)
+  # Combine parameters into a list, so as to pass to Rmd
+  parameters <- list(date = date)
   
-  # Clinical preliminary analysis report
-  message(paste0('Generating the preliminary analysis report in ',
-                 output_dir))
-  generate_preliminary_analysis(date = date,
-                                output_dir = output_dir)
+  # Identify all files for knitting
+  rmds <- dir(system.file('rmd/', package = 'maragra'))
+  rmds <- rmds[grepl('.Rmd', rmds)]
   
-  # Clinical preliminary analysis report
-  message(paste0('Generating the paper in ',
-                 output_dir))
-  generate_paper(date = date,
-                 output_dir = output_dir)
+  # Include / exclude if necessary
+  if(!is.null(only)){
+    rmds <- rmds[rmds %in% only]
+  }
+  if(!is.null(except)){
+    rmds <- rmds[!rmds %in% except]
+  }
+  
+  # Make sure still more than one file
+  if(length(rmds) < 1){
+    stop("No files to knit. Consider setting the 'only' and 'except' to NULL (the default).")
+  }
+  
+  # Loop through each file and knit
+  for(i in 1:length(rmds)){
+    this_file <- rmds[i]
+    file_to_knit <- system.file(paste0('rmd/',
+                                       this_file), 
+                                package='maragra')
+    message(paste0('Generating ',
+                   this_file,
+                   ' in ',
[…]

### 2017-09-06 R/theme_maragra.R

@@ -2,14 +2,13 @@
 #' 
 #' Add a theme to a ggplot object
 #' @param base_size The base font size
-#' @param base_family The font
 #' @return A ggplot compatible theme call
 #' @import ggthemes
 #' @import grid
 #' @export
 # Borrowed from https://rpubs.com/Koundy/71792
-theme_maragra <- function(base_size=14, base_family="Arial") {
-  (theme_foundation(base_size=base_size, base_family=base_family)
+theme_maragra <- function(base_size=14) {
+  (theme_foundation(base_size=base_size)
     + theme(plot.title = element_text(face = "bold",
                                       size = rel(1.2), hjust = 0.5),
             text = element_text(),

### 2017-10-20 inst/rmd/maragra_preliminary_analysis.Rmd
big update

@@
 clinic_agg <- maragra::clinic_agg
+irs <- maragra::irs
 mc <- maragra::mc
@@
 
-For the purposes of this first pass, we "bin" IRS exposure into 5 groups: never, ever but 180+ days ago, 90-80 days, 60-90 days, and in the last 60 days.
+For the purposes of this first pass, we "bin" IRS exposure into 5 groups: before IRS (includes IRS > 365 days ago), 180+ days ago, 90-80 days, 60-90 days, and in the last 60 days.
 
@@
   left_join(matched) %>%
-  mutate(days_since = ifelse(is.na(days_since), 'Never',
-                             ifelse(days_since > 180, '180+',
+  mutate(days_since = ifelse(days_since > 180, '180+',
                                     ifelse(days_since > 90, '090-180',
                                            ifelse(days_since > 60, '060-090',
-                                                  '<060'))))) %>%
+                                                  ifelse(days_since >= 0, '000-060',
+                                                         ifelse(days_since < 0, 'Before', NA)))))) %>%
+  filter(!is.na(days_since)) %>%
+  mutate(days_since = factor(days_since, levels = c('Before', '000-060', '060-090', '090-180', '180+'))) %>%
   mutate(quarter = lendable::quarter_extract(date)) %>%
@@
 
-## To do
+## Discontinuity analysis (not using propoensity score matching)
 
-- More robust models and checks.  
-- Limit ton only clinical malaria. 
-- Un-bin IRS (ie, estimate semi-parametric curve of effectiveness waning)  
-- More investigation of worker type, shift, etc.
+```{r}
+# Create model data for all obs (not just propensity score)
+model_data <-
+  ab_panel %>%
+  # filter(oracle_number %in% matched$oracle_number) %>%
+  left_join(irs) %>%
+  # left_join(matched) %>%
+  mutate(days_since = ifelse(days_since > 180, '180+',
+                                    ifelse(days_since > 90, '090-180',
+                                           ifelse(days_since > 60, '060-090',
+                                                  ifelse(days_since >= 0, '000-060',
+                                                         ifelse(days_since < 0, 'Before', NA)))))) %>%
+  filter(!is.na(days_since)) %>%
+  mutate(days_since = factor(days_since, levels = c('Before', '000-060', '060-090', '090-180', '180+'))) %>%
+  mutate(quarter = lendable::quarter_extract(date)) %>%
+  mutate(quarter = as.character(quarter)) %>%
+  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) #%>%
+  # # remove the period 30 days after spraying
+  # filter(days_since >= 30 | days_since <= 0)
+# Need to add seasonality
+# Need to adjust for houses on and off site
+```
+
+#### Raw absenteeism as a function of days since IRS
+
+```{r}
+x <- model_data %>%
+  group_by(days_since) %>%
+  summarise(absences = length(which(absent)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100)
+
+ggplot(data = x,
+       aes(x = days_since,
+           y = absenteeism_rate)) +
+  geom_bar(stat = 'identity',
+           alpha = 0.6,
+           fill = 'darkblue') +
+  theme_maragra() +
+  labs(x = 'Days since IRS',
+       y = 'Absenteeism rate') +
+  geom_label(aes(label = paste0(round(absenteeism_rate, digits = 2), '%')))
+```
+
+#### Raw absenteeism as a function of days since IRS by quarter
+
+```{r}
+x <- model_data %>%
+  group_by(days_since, quarter) %>%
+  summarise(absences = length(which(absent)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100)
+
+ggplot(data = x %>% mutate(quarter = paste0('Q', quarter)),
+       aes(x = days_since,
+           y = absenteeism_rate)) +
+  geom_bar(stat = 'identity',
+           alpha = 0.6,
+           fill = 'darkblue') +
+  theme_maragra() +
+  labs(x = 'Days since IRS',
+       y = 'Absenteeism rate') +
+  # geom_label(aes(label = paste0(round(absenteeism_rate, digits = 2), '%'))) +
+  facet_wrap(~quarter) +
+  theme(axis.text.x = element_text(angle = 90))
+```
+
+#### Modeling without propensity score
+
+(Absence)
+
+```{r}
+# Need to adjust for houses on and off site
+fit <- glm(absent ~ days_since + quarter,
+           family = binomial('logit'),
+          data = model_data)
+x <- exp(coef(fit))
+ors <- exp(confint(fit))
+ors <- data.frame(ors)
+names(ors) <- c('Lower', 'Upper')
+
+ors <- cbind(x, ors)
+ors$Variable <- row.names(ors)
+ors <- ors %>%
+  dplyr::rename(OR = x) %>%
+  dplyr::select(Variable, OR, Lower, Upper)
+# Combine into a df
+kable(ors) %>%
+  kableExtra::kable_styling(full_width = FALSE)
+```
+
+(Sick absence)
+
+```{r}
+# Need to add seasonality
+# Need to adjust for houses on and off site
+fit <- glm(absent_sick ~ days_since + quarter,
+           family = binomial('logit'),
+          data = model_data)
+x <- exp(coef(fit))
+ors <- exp(confint(fit))
+ors <- data.frame(ors)
+names(ors) <- c('Lower', 'Upper')
+
+ors <- cbind(x, ors)
+ors$Variable <- row.names(ors)
+ors <- ors %>%
+  dplyr::rename(OR = x) %>%
+  dplyr::select(Variable, OR, Lower, Upper)
+# Combine into a df
+kable(ors) %>%
+  kableExtra::kable_styling(full_width = FALSE)
+
+```
+
+#### Visualization of absenteeism before / after IRS
+
+```{r}
+plot_data <-
+  ab_panel %>%
+  # filter(oracle_number %in% matched$oracle_number) %>%
+  left_join(irs) %>%
+  mutate(months_since = days_since %/% 30) %>%
+  filter(!is.na(months_since)) %>%
+  mutate(quarter = lendable::quarter_extract(date)) %>%
+  mutate(quarter = as.character(quarter)) %>%
+  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))
+
+x <- plot_data %>%
+  group_by(months_since) %>%
+  summarise(absences = length(which(absent)),
+            sick_absences = length(which(absent_sick)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100) %>%
+  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100)
+
+ggplot(data = x,
+       aes(x = months_since,
+           y = absenteeism_rate)) +
+  geom_step(color = 'darkred', alpha = 0.8) +
+  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
+  labs(x = 'Months relative to IRS',
+       y = 'Absenteeism rate',
+       title = 'Before/after IRS',
+       subtitle = 'All absences') +
+  theme_maragra()
+
+ggplot(data = x,
+       aes(x = months_since,
+           y = sick_absenteeism_rate)) +
+  geom_step(color = 'darkred', alpha = 0.8) +
+  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
+  labs(x = 'Months relative to IRS',
+       y = 'Absenteeism rate',
+       title = 'Before/after IRS',
+       subtitle = 'Sick absences only') +
+  theme_maragra()
+```
+
+```{r}
+plot_data <-
+  ab_panel %>%
+  # filter(oracle_number %in% matched$oracle_number) %>%
+  left_join(irs) %>%
+  mutate(months_since = days_since %/% 30) %>%
+  filter(!is.na(months_since)) %>%
+  mutate(quarter = lendable::quarter_extract(date)) %>%
+  mutate(quarter = as.character(quarter)) %>%
+  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))
+
+x <- plot_data %>%
+  group_by(months_since, quarter) %>%
+  summarise(absences = length(which(absent)),
+            sick_absences = length(which(absent_sick)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100) %>%
+  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100)
+
+ggplot(data = x %>% mutate(quarter = paste0('Q', quarter)),
+       aes(x = months_since,
+           y = absenteeism_rate)) +
+  geom_step(color = 'darkred', alpha = 0.8) +
+  facet_wrap(~quarter) +
+  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
+  labs(x = 'Months relative to IRS',
+       y = 'Absenteeism rate') +
+  theme_maragra()
+```

### 2017-10-20 inst/rmd/paper.Rmd
big update

@@
   facility in Southern Mozambique
-csl: elsevier-harvard.csl
 date: ''
 output: 
-  html_document:
-    css: resources/style.css
-    pandoc_args:
-    - +RTS
-    - -K16000m
-    - -RTS
-    toc: yes
   pdf_document:
-    toc: yes
+    includes:
+      in_header: preamble.sty
 fig_width: 4
@@
 
-<!-- <hr> -->
-<!-- <p style="text-align:left;"> -->
-<!-- [http://economicsofmalaria.com](http://economicsofmalaria.com) -->
-<!-- <span style="float:right;">joebrew@gmail.com</span> -->
-<!-- </p> -->
-<!-- <hr> -->
-
 ```{r}
@@
 ```
+
+```{r}
+authors <- data.frame(name = c('Joe Brew',
+                               'Kizito Gondo',
+                               'Elton Dorkin',
+                               'Eduardo Nhamahanga',
+                               'Menno Pradhan',
+                               'Laia Cirera',
+                               'Ranjeeta Thomas',
+                               'Elisa Sicuri'),
+                      email = c('joe@databrew.cc',
+                                'KGondo@illovo.co.za',
+                                'EDorkin@illovo.co.za',
+                                'ENhamahanga@illovo.co.za',
+                                'm.p.pradhan@vu.nl',
+                                'laia.cirera@isglobal.org',
+                                'ranjeeta.thomas@imperial.ac.uk',
+                                'elisa.sicuri@isglobal.org'),
+                      affilitation = c("isglobal,cism,vu",
+                                       'ma',
+                                       'ma',
+                                       'ma',
+                                       "vu,uva",
+                                       "icl",
+                                       'isglobal,cism',
+                                       "isglobal,icl,cism"),
+                      footnote = c("Corresponding Author",
+                                   rep('', 7)),
+                      stringsAsFactors = FALSE)
+institutions <- data.frame(code = c('isglobal',
+                                    'icl',
+                                    'cism',
+                                    'vu',
+                                    'uva',
+                                    'ma'),
+                           institution = c('Barcelona Institute for Global Health',
+                                           'Imperial College London',
+                                           'Centro de Investigação em Saúde de Manhiça',
+                                           'VU University Amsterdam',
+                                           'University of Amsterdam',
+                                           'Maragra Açucar SA, Subsidiary of Illovo Sugar Ltd'),
+                           address = c('c/ Rosselló, 132, 5è 2a. 08036, Barcelona, Spain',
+                                       'South Kensington Campus, London SW7 2AZ, U.K.',
+                                       'Vila da Manhiça, Bairro Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique',
+                                       'De Boelelaan 1105, 1081 HV Amsterdam, Netherlands',
+                                       'REC E, Roetersstraat 11, Amsterdam, Netherlands',
+                                       'CP 2789, Maputo, Mozambique'),
+                      stringsAsFactors = FALSE)
+
+make_author <- function(authors,
+                        institutions){
+
+  out <- c()
+  for (i in 1:nrow(authors)){
+    this_author <- authors$name[i]
+    these_institutions <- authors$affilitation[i]
+    these_institutions <- unlist(strsplit(these_institutions, ','))
+    author_text <- paste0(this_author)
+    for (j in 1:length(these_institutions)){
+      this_institution <- these_institutions[j]
+      if(any(grepl(this_institution, out))){
+        author_text <- paste0(author_text,
+                              '\\footrecall{',
+                              this_institution,
+                              '}'
+                              )
+      } else {
+        # new institution, get full name
+        full_name <- institutions$institution[institutions$code == this_institution]
+        address <- institutions$address[institutions$code == this_institution]
+        author_text <- paste0(author_text,
+                              '\\footremember{',
+                              this_institution,
+                              '}{',
+                              full_name,
+                              ': ',
+                              address,
+                              '}')
+      }
+    }
+    out[i] <- author_text
+  }
+  out <- paste0(out, collapse = ', ')
+  cat(out)
+}
+```
+
+\begin{center}
+\begin{large}
+
+```{r, results = 'asis', eval = TRUE}
+make_author(authors = authors,
+            institutions = institutions)
+```
+
+\end{large}
+\end{center}
+
+\vspace{5mm}
+
+\begin{center}
+\textbf{Abstract}  
+\end{center}
+
+\vspace{5mm}
+\begin{center}
+\begin{changemargin}{3cm}{3cm}
+
+Here goes some abstract text.
+Bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla .
+
+\end{changemargin}
+\end{center}
 
@@
 
+### Identification strategy
+
+For our purposes we are analyzing the effect of one intervention (IRS) on 2 outcomes (absence and illness) with many confounders (age, worker type, seasonality, etc.).
+Our analysis can be visualized formulaically as follows:
+
+$$
+\begin{equation}
+\operatorname{Pr}(\text{Outcome} = 1 \mid \text{X}) = \beta_{0} + \beta_{1} \text{Location} + \beta_{2} \text{Season} + (\beta_3{IRS}*\beta_4{IRS_t} + ... )
+\end{equation}
+$$
+
+Our outcome is probabilistic and binomial (ie, one is either absent or present / infected or not infected).
+Our demographic confounders (represented by $...$) will be a function of iterative model selection.
+Our intervention (IRS) is not a simple yes/no, but rather the product of whether the residence of the worker in question was treated in the last year, and, if so, the time since treatment (represented above as the interaction term, where where $_t$ represents time elapsed since commencement of the most recent IRS campaign).
+
 ```{r}
-# # Create model data
-# model_data <-
-#   ab_panel %>%
-#   left_join(irs) %>%
-#   left_join(workers %>%
-#               dplyr::select(oracle_number,
-#                             sex,
-#                             race_code,
-#                             department,
-#                             permanent_or_temporary)) %>%
-#   mutate(days_since = ifelse(is.na(days_since), 'Never',
-#                              ifelse(days_since > 180, '180+',
-#                                     ifelse(days_since > 90, '090-180',
-#                                            ifelse(days_since > 60, '060-090', 
-#                                                   '<060')))))
-# # Need to add seasonality
-# # Need to adjust for houses on and off site
-# fit <- glm(absent ~ days_since,
-#            family = binomial('logit'),
-#           data = model_data)
-# exp(coef(fit))
+# Create model data
+model_data <-
+  ab_panel %>%
+  # filter(oracle_number %in% matched$oracle_number) %>%
+  left_join(irs) %>%
+  # left_join(matched) %>%
+  mutate(days_since = ifelse(is.na(days_since), 'Never',
+                             ifelse(days_since > 180, '180+',
+                                    ifelse(days_since > 90, '090-180',
+                                           ifelse(days_since > 60, '060-090',
+                                                  '<060'))))) %>%
+  mutate(quarter = lendable::quarter_extract(date)) %>%
+  mutate(quarter = as.character(quarter)) %>%
+  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))
+# Need to add seasonality
+# Need to adjust for houses on and off site
+fit <- glm(absent ~ days_since + quarter,
+           family = binomial('logit'),
+          data = model_data)
+x <- exp(coef(fit))
+ors <- exp(confint(fit))
+ors <- data.frame(ors)
+names(ors) <- c('Lower', 'Upper')
+
+ors <- cbind(x, ors)
+ors$Variable <- row.names(ors)
+ors <- ors %>%
+  dplyr::rename(OR = x) %>%
+  dplyr::select(Variable, OR, Lower, Upper)
+# Combine into a df
+kable(ors) %>%
+  kableExtra::kable_styling(full_width = FALSE)
+```
+
+We run the same model, but instead of estimating absences, we estimate only the likelihood of sick absences. The results (in form of odds ratios) are below.
+
+```{r}
+# Need to add seasonality
+# Need to adjust for houses on and off site
+fit <- glm(absent_sick ~ days_since + quarter,
+           family = binomial('logit'),
+          data = model_data)
+x <- exp(coef(fit))
+ors <- exp(confint(fit))
+ors <- data.frame(ors)
+names(ors) <- c('Lower', 'Upper')
+
+ors <- cbind(x, ors)
+ors$Variable <- row.names(ors)
+ors <- ors %>%
+  dplyr::rename(OR = x) %>%
+  dplyr::select(Variable, OR, Lower, Upper)
+# Combine into a df
+kable(ors) %>%
+  kableExtra::kable_styling(full_width = FALSE)
+
 ```
@@
 
-```{r include=FALSE}
+```{r, eval=FALSE}
 # automatically create a bib database for R packages

### 2017-10-20 data-raw/create_data_files.R

@@ -250,12 +250,12 @@ if('eligible_working_days.RData' %in% dir()){
       start_date <- workers$company_entry_date[i]
       end_date <- as.Date('2017-03-31')
     }
-    eligible_working_days[[i]] <- 
+    eligible_working_days[[i]] <-
       expandify(oracle_number = workers$oracle_number[i],
                 start = start_date,
                 end = end_date)
   }
-  
+
   # Bind all together
   eligible_working_days <- bind_rows(eligible_working_days)
   save(eligible_working_days,
@@ -765,7 +765,7 @@ if(!'census_done.RData' %in% dir('census_data')){
 }
 
 # Keep only relevant columns (since most census info we'll get from openhds only)
-demografia <- census %>% 
+demografia <- census %>%
   dplyr::select(-unidade,
                 -name)
 
@@ -777,13 +777,13 @@ demografia <- census %>%
 if('open_hds_data.RData' %in% dir('census_data')){
   load('census_data/open_hds_data.RData')
 } else {
-  membership <- 
+  membership <-
     cism::get_data(tab = 'membership',
                    dbname = 'openhds')
-  individual <- 
+  individual <-
     cism::get_data(tab = 'individual',
                    dbname = 'openhds')
-  location <- 
+  location <-
     cism::get_data(tab = 'location',
                    dbname = 'openhds')
   residency <-
@@ -792,7 +792,7 @@ if('open_hds_data.RData' %in% dir('census_data')){
   VISIT_REGISTRATION_CORE <-
     cism::get_data(tab = 'VISIT_REGISTRATION_CORE',
                    dbname = 'dssodk')
-  
+
   save(membership,
        individual,
        location,
@@ -861,14 +861,14 @@ people <- people %>%
                      !is.na(COORDINATES_LNG)) %>%
               filter(!duplicated(LOCATION_NAME)),
             by = c('locationName' = 'LOCATION_NAME'))
-people$latitude <- 
+people$latitude <-
   ifelse(is.na(people$latitude), people$COORDINATES_LAT, people$latitude)
-people$longitude <- 
+people$longitude <-
   ifelse(is.na(people$longitude), people$COORDINATES_LNG, people$longitude)
 
 # Get an age
 people <- people %>%
-  mutate(age_years = as.numeric(snap_shot - dob) / 365.25) 
+  mutate(age_years = as.numeric(snap_shot - dob) / 365.25)
 
 # Fix the last name / permid naming issue
[…]

### 2017-10-24 inst/rmd/ihmt_presentation_brew_2017.Rmd
major presentation update

---
title: !r "Effectiveness of private sector malaria control: The case of sugarcane workers in Mozambique"
subtitle: IHMT, October, 2017, Lisboa, Portugal
fig_height: 5
fig_width: 7
bibliography: bibliography.bib
output:
  beamer_presentation:
    includes:
      in_header: preamble_beamer.sty
    theme: "Berlin"
    colortheme: "seagull"
    fonttheme: "structuresmallcapsserif"
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = F, results = 'hide'}
# All beamer themes at http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme_and_color.html
library(knitr)
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE)

# Packages
library(ggplot2)
library(cism)
library(rgeos)
library(maptools)
library(rgdal)
library(tidyr)
library(RColorBrewer)
library(dplyr)
library(leaflet)
library(readr)
library(ggthemes)
library(ggrepel)
library(knitr)
options(knitr.table.format = "latex")
library(maragra)
```

```{r, results = 'asis'}
# Print the author names and affiliations
x <- make_authors(include_address = FALSE, include_country = TRUE, seperator = ' \n\n ', cat_it = FALSE)
cat(x)
```

# Introduction

## Context

- Malaria has a nearly unquantifiably large economic impact.
- Many channels: fertility, fecundity, saving, investment [@Shretta2016], risk perception, productivity, absenteeism, human capital accumulation [@CastelBranco2014], mortality, costs of care [@Sachs2002].
- Cost-benefit studies often only consider the costs of an intervention and associated costs of care, without quantifying the societal cost of non-intervention.

## What we already know

- Malaria is associated with absenteeism in workers [@Nonvignon_2016].
- Malaria has a negative effect on GDP [@Orem_2012] and growth [@McCarthy_2000].
- Malaria control is cost-effective from the societal/public perspective [@Purdy_2013].
- Indoor residual spraying (IRS) is cost-effective [@Howard_2017], [@White_2011].

## What we want to know

What is the investment case from the private perspective?

- From a purely financial perspective, what benefits does a private company experience?  
- What is the short-term cost-effectiveness of IRS for large companies in malaria-endemic regions?
- Is malaria control just good "corporate social responsibility"? Or is it also good business?

## Research questions

We can't answer all the previous questions (yet). So we focus on one:

**What is the short-term effect on absenteeism of IRS at the homes of sugarcane workers?**

# Methods

## Reproducibility

- **Code:** [github.com/joebrew/maragra](https://github.com/joebrew/maragra)
- **Data:** Only available to authorized collaborators, but "dummy data" has been created to validate methods  
- **Code bundling:** All code is in an R package: `devtools::install_github('joebrew/maragra)`

## Identification strategy

Some text here.

# Results

## Reduction in absenteeism

Some text here.

## Reduction in illness

Some text here.

# Discussion

## General

## Implications

## Limitations

## Thank you

**Email:** joe@economicsofmalaria.com

**Presentation:** [economicsofmalaria.com/ihmt](http://economicsofmalaria.com/ihmt)

## References

### 2017-10-24 inst/rmd/paper.Rmd
major presentation update

@@
 
-```{r}
-authors <- data.frame(name = c('Joe Brew',
-                               'Kizito Gondo',
-                               'Elton Dorkin',
-                               'Eduardo Nhamahanga',
-                               'Menno Pradhan',
-                               'Laia Cirera',
-                               'Ranjeeta Thomas',
-                               'Elisa Sicuri'),
-                      email = c('joe@databrew.cc',
-                                'KGondo@illovo.co.za',
-                                'EDorkin@illovo.co.za',
-                                'ENhamahanga@illovo.co.za',
-                                'm.p.pradhan@vu.nl',
-                                'laia.cirera@isglobal.org',
-                                'ranjeeta.thomas@imperial.ac.uk',
-                                'elisa.sicuri@isglobal.org'),
-                      affilitation = c("isglobal,cism,vu",
-                                       'ma',
-                                       'ma',
-                                       'ma',
-                                       "vu,uva",
-                                       "icl",
-                                       'isglobal,cism',
-                                       "isglobal,icl,cism"),
-                      footnote = c("Corresponding Author",
-                                   rep('', 7)),
-                      stringsAsFactors = FALSE)
-institutions <- data.frame(code = c('isglobal',
-                                    'icl',
-                                    'cism',
-                                    'vu',
-                                    'uva',
-                                    'ma'),
-                           institution = c('Barcelona Institute for Global Health',
-                                           'Imperial College London',
-                                           'Centro de Investigação em Saúde de Manhiça',
-                                           'VU University Amsterdam',
-                                           'University of Amsterdam',
-                                           'Maragra Açucar SA, Subsidiary of Illovo Sugar Ltd'),
-                           address = c('c/ Rosselló, 132, 5è 2a. 08036, Barcelona, Spain',
-                                       'South Kensington Campus, London SW7 2AZ, U.K.',
-                                       'Vila da Manhiça, Bairro Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique',
-                                       'De Boelelaan 1105, 1081 HV Amsterdam, Netherlands',
-                                       'REC E, Roetersstraat 11, Amsterdam, Netherlands',
-                                       'CP 2789, Maputo, Mozambique'),
-                      stringsAsFactors = FALSE)
-
-make_author <- function(authors,
-                        institutions){
-
-  out <- c()
-  for (i in 1:nrow(authors)){
-    this_author <- authors$name[i]
-    these_institutions <- authors$affilitation[i]
-    these_institutions <- unlist(strsplit(these_institutions, ','))
-    author_text <- paste0(this_author)
-    for (j in 1:length(these_institutions)){
-      this_institution <- these_institutions[j]
-      if(any(grepl(this_institution, out))){
-        author_text <- paste0(author_text,
-                              '\\footrecall{',
-                              this_institution,
-                              '}'
-                              )
-      } else {
-        # new institution, get full name
-        full_name <- institutions$institution[institutions$code == this_institution]
-        address <- institutions$address[institutions$code == this_institution]
-        author_text <- paste0(author_text,
-                              '\\footremember{',
-                              this_institution,
-                              '}{',
-                              full_name,
-                              ': ',
-                              address,
-                              '}')
-      }
-    }
-    out[i] <- author_text
-  }
-  out <- paste0(out, collapse = ', ')
-  cat(out)
-}
-```
-
 \begin{center}
@@
 ```{r, results = 'asis', eval = TRUE}
-make_author(authors = authors,
+make_authors(authors = authors,
             institutions = institutions)

### 2017-10-24 R/generate_ihmt_presentation.R

@@ -1,7 +1,6 @@
-#' Generate paper
+#' Generate IHMT presentation
 #'
-#' Generate the paper
-#' originally prepared in August / September 2017
+#' Generate the presentation for the Oct 30-31 IHMT presentation
 #' @param date A date to be printed on the report. If \code{NULL} (the
 #' default), the current date will be used
 #' @param output_dir The directory to which the file should be written. If
@@ -11,31 +10,21 @@
 #' @importFrom rmarkdown render
 #' @export
 
-generate_paper <- function(date = NULL,
-                           output_dir = NULL,
-                           output_file = 'paper.pdf'){
+generate_ihmt_presentation <- function(output_dir = NULL,
+                                       output_file = 'ihmt_presentation.pdf'){
 
   # If no output directory, make current wd
   if(is.null(output_dir)){
     output_dir <- getwd()
   }
 
-  # If not date, use today's
-  if(is.null(date)){
-    date <- Sys.Date()
-  }
-
-  # Combine parameters into a list, so as to pass to Rmd
-  parameters <- list(date = date)
-
   # Find location the rmd to knit
   file_to_knit <-
-    system.file('rmd/paper.Rmd',
+    system.file('rmd/ihmt_presentation_brew_2017.Rmd',
                 package='maragra')
 
   # Knit file
   rmarkdown::render(file_to_knit,
                     output_dir = output_dir,
-                    output_file = output_file,
-                    params = parameters)
+                    output_file = output_file)
 }

### 2017-10-24 R/make_authors.R

@@ -1,0 +1,106 @@
+#' Make authors
+#'
+#' Cat the authors in latex-ready format
+#' @return Latex text
+#' @param authors A dataframe with columns: name, email, affiliation, footnote
+#' @param institutions A dataframe with columns: code, institution, address
+#' @param cat_it Whether to cat (as opposed to return)
+#' @param include_address Whether to include the address
+#' @param include_country Whether to include the country; only relevant if include_address is FALSE
+#' @param seperator What to put between author names: usually a comma or line break
+#' @export
+
+make_authors <- function(
+
+  authors = data.frame(name = c('Joe Brew',
+                                'Kizito Gondo',
+                                'Elton Dorkin',
+                                'Eduardo Nhamahanga',
+                                'Menno Pradhan',
+                                'Laia Cirera',
+                                'Ranjeeta Thomas',
+                                'Elisa Sicuri'),
+                       email = c('joe@databrew.cc',
+                                 'KGondo@illovo.co.za',
+                                 'EDorkin@illovo.co.za',
+                                 'ENhamahanga@illovo.co.za',
+                                 'm.p.pradhan@vu.nl',
+                                 'laia.cirera@isglobal.org',
+                                 'ranjeeta.thomas@imperial.ac.uk',
+                                 'elisa.sicuri@isglobal.org'),
+                       affilitation = c("isglobal,cism,vu",
+                                        'ma',
+                                        'ma',
+                                        'ma',
+                                        "vu,uva",
+                                        'isglobal,cism',
+                                        "icl",
+                                        "isglobal,cism,icl"),
+                       footnote = c("Corresponding Author",
+                                    rep('', 7)),
+                       stringsAsFactors = FALSE),
+
+  institutions = data.frame(code = c('isglobal',
+                                     'icl',
+                                     'cism',
+                                     'vu',
+                                     'uva',
+                                     'ma'),
+                            institution = c('Barcelona Institute for Global Health',
+                                            'Imperial College London',
+                                            'Centro de Investigação em Saúde de Manhiça',
+                                            'VU University Amsterdam',
+                                            'University of Amsterdam',
+                                            'Maragra Açucar SA, Subsidiary of Illovo Sugar Ltd'),
[…]

### 2017-10-29 inst/rmd/ihmt_presentation_brew_2017.Rmd
done with ihmt presentation

@@
 ---
-title: !r "Effectiveness of private sector malaria control: The case of sugarcane workers in Mozambique"
-subtitle: IHMT, October, 2017, Lisboa, Portugal
+title: "Effectiveness of private sector malaria control: The case of sugarcane workers in Mozambique"
+subtitle: 'IHMT, October, 2017, Lisboa, Portugal'
 fig_height: 5
@@
                cache = FALSE)
+opts_chunk$set(dev = 'pdf')
 
@@
 
+```{r}
+# Read in data
+ab <- maragra::ab
+ab_panel <- maragra::ab_panel
+bairros <- maragra::bairros
+census <- maragra::census
+clinic <- maragra::clinic
+clinic_agg <- maragra::clinic_agg
+irs <- maragra::irs
+mc <- maragra::mc
+workers <- maragra::workers
+
+# Define function fro breaking the days_since var
+break_days_since <- function(days_since){
+  out <- ifelse(days_since > 270, '270+',
+    ifelse(days_since > 180, '181-270',
+           ifelse(days_since > 90, '091-180',
+                  ifelse(days_since > 60, '061-090',
+                         ifelse(days_since > 30, '031-060',
+                                ifelse(days_since >= 0, '000-030',
+                                       ifelse(days_since < 0, 'Before', NA)))))))
+  out <- factor(out,
+                levels = c('Before',
+                           '000-030',
+                           '031-060',
+                           '061-090',
+                           '091-180',
+                           '181-270',
+                           '270+'))
+  return(out)
+}
+
+# Define function for making season
+make_season <- function(date){
+  out <- ifelse(as.numeric(format(date, '%m')) %in% c(11:12, 1:3),
+                   'Malaria season', 'Off season')
+  out <- factor(out, levels = c('Off season', 'Malaria season'))
+  return(out)
+}
+
+library(MatchIt)
+set.seed(1234)
+
+# Create matching data for propensity scoring
+right_side <- irs %>%
+  group_by(unidade) %>% 
+  tally %>%
+  ungroup %>%
+  mutate(received = ifelse(n > 0, TRUE, FALSE))
+
+left_side <-
+  workers %>%
+  dplyr::select(oracle_number, unidade, permanent_or_temporary, department, sex, 
+                date_of_birth,
+                perm_id) %>%
+  # bring in census address
+  left_join(census %>%
+              dplyr::select(perm_id,
+                            longitude,
+                            latitude,
+                            maragra_bairro,
+                            maragra_fabrica))
+
+psm <- left_join(x = left_side,
+                 y = right_side,
+                 by = 'unidade') %>%
+  # Keep only those who are censed
+  filter(!is.na(longitude),
+         !is.na(latitude)) %>%
+  # Keep only those who live within maragra
+  filter(maragra_fabrica |maragra_bairro) %>%
+  mutate(received = ifelse(is.na(received), FALSE, received)) %>%
+  mutate(age = round((as.numeric(as.Date('2016-01-01') - date_of_birth)) / 365.25, digits = 2)) %>%
+  filter(!is.na(sex),
+         !is.na(age)) %>%
+  dplyr::select(-unidade, -date_of_birth, -n)
+
+psmt1 <- psm
+names(psmt1) <- toupper(names(psmt1))
+psmt1 <- psmt1 %>%
+  mutate(RECEIVED = ifelse(RECEIVED == TRUE,'IRS', 'No IRS')) %>%
+  mutate(STATUS = PERMANENT_OR_TEMPORARY)
+
+pacman::p_load(tableone)
+table1 <- CreateTableOne(vars = toupper(c('STATUS', 'department', 'age',
+                                  'sex', 'received')), 
+                         data = psmt1, 
+                         factorVars = toupper(c('STATUS', 'department',
+                                  'sex')), 
+                         strata = 'RECEIVED')
+table1 <- print(table1, 
+                printToggle = FALSE, 
+                noSpaces = TRUE)
+
+match.it <- matchit(received ~ age + sex + permanent_or_temporary + department, data = psm, method="nearest", ratio=1)
+# Save matched samples
+matched <- match.data(match.it)[1:ncol(psm)]
+a <- summary(match.it)
+
+# Create model data
+model_data_propensity <-
+  ab_panel %>%
+  filter(oracle_number %in% matched$oracle_number) %>%
+  left_join(irs) %>%
+  left_join(matched) %>%
+  mutate(days_since = break_days_since(days_since)) %>%
+  filter(!is.na(days_since)) %>%
+  # mutate(time_period = lendable::time_period_extract(date)) %>%
+  # mutate(time_period = as.character(time_period)) %>%
+  mutate(time_period = make_season(date = date)) %>%
+  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))
+
+# Create model data for all obs (not just propensity score)
+model_data <-
+  ab_panel %>%
+  # filter(oracle_number %in% matched$oracle_number) %>%
+  left_join(irs) %>%
+  # left_join(matched) %>%
+  mutate(days_since = break_days_since(days_since)) %>%
+  filter(!is.na(days_since)) %>%
+  # mutate(time_period = lendable::time_period_extract(date)) %>%
+  # mutate(time_period = as.character(time_period)) %>%
+  mutate(time_period = make_season(date = date)) %>%
+  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) #%>%
+  # # remove the period 30 days after spraying
+  # filter(days_since >= 30 | days_since <= 0)
+# Need to add seasonality
+# Need to adjust for houses on and off site
+
+fit_propensity <- glm(absent ~ days_since * time_period,
+           family = binomial('logit'),
+          data = model_data_propensity)
+x <- exp(coef(fit_propensity))
+ors_propensity <- exp(confint(fit_propensity))
+ors_propensity <- data.frame(ors_propensity)
+names(ors_propensity) <- c('Lower', 'Upper')
+
+ors_propensity <- cbind(x, ors_propensity)
+ors_propensity$Variable <- row.names(ors_propensity)
+ors_propensity <- ors_propensity %>%
+  dplyr::rename(OR = x) %>%
+  dplyr::select(Variable, OR, Lower, Upper) %>%
+  mutate(OR = round(OR, digits = 3),
+         Lower = round(Lower, digits = 3),
+         Upper = round(Upper, digits = 3))
+
+# Need to add seasonality
+# Need to adjust for houses on and off site
+fit_propensity_sick <- glm(absent_sick ~ days_since * time_period,
+           family = binomial('logit'),
+          data = model_data_propensity)
+x <- exp(coef(fit_propensity_sick))
+ors_propensity_sick <- exp(confint(fit_propensity_sick))
+ors_propensity_sick <- data.frame(ors_propensity_sick)
+names(ors_propensity_sick) <- c('Lower', 'Upper')
+
+ors_propensity_sick <- cbind(x, ors_propensity_sick)
+ors_propensity_sick$Variable <- row.names(ors_propensity_sick)
+ors_propensity_sick <- ors_propensity_sick %>%
+  dplyr::rename(OR = x) %>%
+  dplyr::select(Variable, OR, Lower, Upper) %>%
+  mutate(OR = round(OR, digits = 3),
+         Lower = round(Lower, digits = 3),
+         Upper = round(Upper, digits = 3))
+
+# Need to adjust for houses on and off site
+fit <- glm(absent ~ days_since * time_period,
+           family = binomial('logit'),
+          data = model_data)
+x <- exp(coef(fit))
+ors <- exp(confint(fit))
+ors <- data.frame(ors)
+names(ors) <- c('Lower', 'Upper')
+
+ors <- cbind(x, ors)
+ors$Variable <- row.names(ors)
+ors <- ors %>%
+  dplyr::rename(OR = x) %>%
+  dplyr::select(Variable, OR, Lower, Upper) %>%
+  mutate(OR = round(OR, digits = 3),
+         Lower = round(Lower, digits = 3),
+         Upper = round(Upper, digits = 3))
+
+# Need to add seasonality
+# Need to adjust for houses on and off site
+fit_sick <- glm(absent_sick ~ days_since * time_period,
+           family = binomial('logit'),
+          data = model_data)
+x <- exp(coef(fit_sick))
+ors_sick <- exp(confint(fit_sick))
+ors_sick <- data.frame(ors_sick)
+names(ors_sick) <- c('Lower', 'Upper')
+
+ors_sick <- cbind(x, ors_sick)
+ors_sick$Variable <- row.names(ors_sick)
+ors_sick <- ors_sick %>%
+  dplyr::rename(OR = x) %>%
+  dplyr::select(Variable, OR, Lower, Upper) %>%
+  mutate(OR = round(OR, digits = 3),
+         Lower = round(Lower, digits = 3),
+         Upper = round(Upper, digits = 3))
+
+```
+
 # Introduction
@@
 
-What is the investment case from the private perspective?
+What is the investment case _from the investor's perspective_?
 
-- From a purely financial perspective, what benefits does a private company experience?  
-- What is the short-term cost-effectiveness of IRS for large companies in malaria-endemic regions?
 - Is malaria control just good "corporate social responsibility"? Or is it also good business?
+- From a purely financial/investment point-of-view, what benefits does a private company experience in engaging in malaria control?  
+- What is the short-term **benefits** of IRS for large companies in malaria-endemic regions?
+- What are the **costs** of of carrying out IRS for large companies?
 
@@
 
-**What is the short-term effect on absenteeism of IRS at the homes of sugarcane workers?**
+**What is the short-term effect of IRS on worker absenteeism and clinical illness among sugarcane workers?**
+
+## Research site
+
+```{r}
+par(mfrow = c(2,2))
+plot(cism::africa,
+     col = 'grey',
+     border = 'white',
+     lwd = 0.1)
+title('Africa')
+for(i in (seq(0.5, 1.5, length = 2))^1.3){
+  points(32.779401, -25.452049, col = adjustcolor('red', alpha.f = 0.5), cex = i)
+}
+plot(cism::moz2,
+     col = 'grey',
+     border = 'white',
+     lwd = 0.1)
+title('Mozambique')
+for(i in (seq(0.5, 2.5, length = 3))^1.3){
+  points(32.779401, -25.452049, col = adjustcolor('red', alpha.f = 0.5), cex = i)
+}
+man <- cism::man3
+plot(man,
+     col = 'grey', 
+     border = 'white',
+     lwd = 0.1)
+title('Manhiça district')
+for(i in (seq(0.5, 3, length = 4))^1.3){
+  points(32.779401, -25.452049, col = adjustcolor('red', alpha.f = 0.5), cex = i)
+}
+plot(man[man@data$NAME_3 %in% c('Manhica - Sede', 'Maluana'),],
+     col = 'grey', 
+     border = 'white',
+     lwd = 0.1)
+title('Manhiça and Maluana posts')
+for(i in (seq(0.5, 3.5, length = 4))^1.3){
+  points(32.779401, -25.452049, col = adjustcolor('red', alpha.f = 0.5), cex = i)
+}
+par(mfrow = c(1,1))
+```
+
+## Research Site II
+
+\includegraphics{images/sat}
 
@@
 
-## Reproducibility
-
-- **Code:** [github.com/joebrew/maragra](https://github.com/joebrew/maragra)
-- **Data:** Only available to authorized collaborators, but "dummy data" has been created to validate methods  
-- **Code bundling:** All code is in an R package: `devtools::install_github('joebrew/maragra)`
-
 ## Identification strategy
 
-Some text here.
+- 1 intervention (IRS, time to/from)
+- 2 outcomes (absence and illness, probabilistic / binomial)
+- Many confounders (age, worker type, seasonality, etc.).
+
+\begin{small}
+\begin{equation}
+\operatorname{Pr}(\text{Outcome} = 1 \mid \text{X}) = \beta_{0} + \beta_{1} \text{Location} + \beta_{2} \text{Season} + (\beta_3{IRS}*\beta_4{IRS_t} + ... )
+\end{equation}
+\end{small}
+
+## Modeling
+
+We employ two approaches:
+
+1.
+**Propensity score matching** of workers who _ever_ received IRS with workers who _never_ received IRS.
+**Advantage**: No need to adjust for confounders with a matched sample, thereby avoiding reduction in degrees of freedom
+
+2.
+**Regresion-discontinuity** of only those workers who _ever_ received IRS (ie, ignoring those who never received IRS).
+**Advantage**: Those who _never_ received IRS may be qualitatively different, and therefore not an appropriate comparison group.
+
+## Propensity score matching {.allowframebreaks}
+
+- We generate a matched sample of similar workers by first estimating the likelihood of having ever received the intervention, given a worker's age, sex, department and temporary vs.
+permanent status.
+ - This is necessary due to below, significant differences:
+
+\begin{tiny}
+
+```{r}
+kable(table1[,1:3],  
+      align = 'c', 
+      caption = 'Comparison of unmatched samples') %>%
+  kableExtra::kable_styling(full_width = FALSE)
+```
+
+\end{tiny}
+
+We match, employing the nearest neighbor method for identifying those workers from our control group who most resemble those workers in the treatment group. [@Ho_Imai_King_Stuart_2007].
+
+- Our match is a 1-to-1 cut, meaning those control workers who do not resemble those in the treatment group are left out of primary analysis. The below table shows the match results.
+
+\begin{tiny}
+
+```{r, results = 'asis'}
+kable(a$nn, digits = 2, align = 'c', 
+      caption = 'Sample sizes')
+```
+
+\end{tiny}
+
+The distributions of our numeric variables are now extremely similar:
+
+\begin{tiny}
+
+```{r, results = 'asis'}
+kable(a$sum.matched[c(1,2,3,4)], digits = 2, align = 'c', 
+      caption = 'Summary of balance for matched data') 
+```
+
+\end{tiny}
+
+## Regression discontunity analysis
+
+- We simply only consider those workers who _ever_ got IRS.
+- We take into account one full year prior to IRS and one full year after IRS.
+- Our dataset constitutes one observation per worker-day. 
+- We incidentally achieve a sort of "matching" through the fact that workers are their own controls.
 
@@
 
-## Reduction in absenteeism
+## Descriptive: absenteeism by time from/to intervention
 
-Some text here.
+```{r}
+plot_data <-
+  ab_panel %>%
+  # filter(oracle_number %in% matched$oracle_number) %>%
+  left_join(irs) %>%
+  mutate(months_since = days_since %/% 30) %>%
+  filter(!is.na(months_since)) %>%
+  # mutate(time_period = lendable::time_period_extract(date)) %>%
+  # mutate(time_period = as.character(time_period)) %>%
+  mutate(time_period = make_season(date = date)) %>%
+  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))
 
-## Reduction in illness
+x <- plot_data %>%
+  group_by(months_since) %>%
+  summarise(absences = length(which(absent)),
+            sick_absences = length(which(absent_sick)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100) %>%
+  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100)
 
-Some text here.
+ggplot(data = x,
+       aes(x = months_since,
+           y = absenteeism_rate)) +
+  geom_step(color = 'darkred', alpha = 0.8) +
+  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
+  labs(x = 'Months relative to IRS',
+       y = 'Absenteeism rate',
+       title = 'Before/after IRS',
+       subtitle = 'All absences') +
+  theme_maragra()
+```
+
+## Descriptive: absenteeism by time from/to intervention (with local regression lines)
+
+```{r}
+plot_data <-
+  ab_panel %>%
+  # filter(oracle_number %in% matched$oracle_number) %>%
+  left_join(irs) %>%
+  mutate(months_since = days_since %/% 30) %>%
+  filter(!is.na(months_since)) %>%
+  # mutate(time_period = lendable::time_period_extract(date)) %>%
+  # mutate(time_period = as.character(time_period)) %>%
+  mutate(time_period = make_season(date = date)) %>%
+  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))
+
+x <- plot_data %>%
+  group_by(months_since) %>%
+  summarise(absences = length(which(absent)),
+            sick_absences = length(which(absent_sick)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100) %>%
+  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100) %>%
+  mutate(before = months_since < 0)
+
+ggplot(data = x,
+       aes(x = months_since,
+           y = absenteeism_rate,
+           group = before)) +
+  geom_smooth() +
+  geom_step(color = 'darkred', alpha = 0.8) +
+  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
+  labs(x = 'Months relative to IRS',
+       y = 'Absenteeism rate',
+       title = 'Before/after IRS',
+       subtitle = 'All absences') +
+  theme_maragra()
+```
+
+## Descriptive: sick absenteeism by time from/to intervention
+
+```{r}
+ggplot(data = x,
+       aes(x = months_since,
+           y = sick_absenteeism_rate,
+           group = before)) +
+  geom_smooth() +
+  geom_step(color = 'darkred', alpha = 0.8) +
+  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
+  labs(x = 'Months relative to IRS',
+       y = 'Absenteeism rate',
+       title = 'Before/after IRS',
+       subtitle = 'Sick absences only') +
+  theme_maragra()
+```
+
+## Descriptive: absenteeism by time from/to intervention (by time_period)
+
+```{r}
+plot_data <-
+  ab_panel %>%
+  # filter(oracle_number %in% matched$oracle_number) %>%
+  left_join(irs) %>%
+  mutate(months_since = days_since %/% 30) %>%
+  filter(!is.na(months_since)) %>%
+  # mutate(time_period = lendable::time_period_extract(date)) %>%
+  # mutate(time_period = as.character(time_period)) %>%
+  mutate(time_period = make_season(date = date)) %>%
+  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))
+
+x <- plot_data %>%
+  group_by(months_since, time_period) %>%
+  summarise(absences = length(which(absent)),
+            sick_absences = length(which(absent_sick)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100) %>%
+  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100) %>%
+  mutate(before = months_since < 0)
+
+ggplot(data = x %>% mutate(time_period = paste0(time_period)),
+       aes(x = months_since,
+           y = absenteeism_rate)) +
+  geom_step(color = 'darkred', alpha = 0.8) +
+  facet_wrap(~time_period, ncol = 1) +
+  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
+  labs(x = 'Months relative to IRS',
+       y = 'Absenteeism rate') +
+  theme_maragra()
+```
+
+## Same chart with local regression lines
+
+```{r}
+ggplot(data = x %>% mutate(time_period = paste0(time_period)),
+       aes(x = months_since,
+           y = absenteeism_rate,
+           group = before)) +
+  geom_smooth() +
+  geom_step(color = 'darkred', alpha = 0.8) +
+  facet_wrap(~time_period, ncol = 1) +
+  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
+  labs(x = 'Months relative to IRS',
+       y = 'Absenteeism rate') +
+  theme_maragra()
+```
+
+## Descriptive: sick absenteeism by time from/to intervention (by time_period)
+
+```{r}
+
+ggplot(data = x %>% mutate(time_period = paste0(time_period)),
+       aes(x = months_since,
+           y = sick_absenteeism_rate,
+           group = before)) +
+  geom_smooth() +
+  geom_step(color = 'darkred', alpha = 0.8) +
+  facet_wrap(~time_period, ncol = 1) +
+  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
+  labs(x = 'Months relative to IRS',
+       y = 'Absenteeism rate',
+       subtitle = 'Sick absences only') +
+  theme_maragra()
+```
+
+## Descriptive: absenteeism in binned time period
+
+```{r}
+x <- model_data %>%
+  group_by(days_since) %>%
+  summarise(absences = length(which(absent)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100)
+
+ggplot(data = x,
+       aes(x = days_since,
+           y = absenteeism_rate)) +
+  geom_bar(stat = 'identity',
+           alpha = 0.6,
+           fill = 'darkblue') +
+  theme_maragra() +
+  labs(x = 'Days since IRS',
+       y = 'Absenteeism rate',
+       title = 'Raw absenteeism as a function of days since IRS') +
+  geom_label(aes(label = paste0(round(absenteeism_rate, digits = 2), '%')))
+```
+
+## Descriptive: absenteeism in binned time periods by season
+
+```{r}
+x <- model_data %>%
+  group_by(days_since, time_period) %>%
+  summarise(absences = length(which(absent)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100)
+
+ggplot(data = x %>% mutate(time_period = paste0(time_period)),
+       aes(x = days_since,
+           y = absenteeism_rate)) +
+  geom_bar(stat = 'identity',
+           alpha = 0.6,
+           fill = 'darkblue') +
+  theme_maragra() +
+  labs(x = 'Days since IRS',
+       y = 'Absenteeism rate',
+       title = 'Raw absenteeism as a function of days since IRS by time_period') +
+  # geom_label(aes(label = paste0(round(absenteeism_rate, digits = 2), '%'))) +
+  facet_wrap(~time_period, ncol = 1) +
+  theme(axis.text.x = element_text(angle = 90))
+```
+
+## Modeling after matching {.allowframebreaks}
+
+- Matched sample of `r nrow(matched)` workers, of which 50% received IRS and 50% did not.   
+- Model only takes into account seasonality, since matching theoretically handles other differences.  
+- For the purposes of this first pass, we "bin" IRS exposure and estimate a logit model to calculate odds ratios.
+
+\vfill
+\newpage
+
+All absence:
+
+\begin{tiny}
+
+```{r, results = 'asis'}
+# Combine into a df
+kable(ors_propensity) 
+```
+
+\end{tiny}
+
+\vfill
+\newpage
+
+```{r}
+visualize_ors <- function(ors_object){
+  ors_object <- ors_object %>% 
+    filter(grepl('days_since', Variable),
+           grepl('time_periodMalaria', Variable))
+  ors_object <- ors_object %>%
+    mutate(Variable = gsub('days_since', '', Variable),
+           Variable = gsub(':time_periodMalaria season', '', Variable))
+  ggplot(data = ors_object,
+         aes(x = Variable,Variable,
+             y = OR)) +
+    geom_point(alpha = 0.7) +
+    geom_errorbar(aes(ymin = Lower,
+                       ymax = Upper),
+                  alpha = 0.6,
+                  size = 0.5) +
+    geom_hline(yintercept = 1, lty = 2, alpha = 0.6, color = 'red') +
+    labs(x = 'Days since IRS',
+         y = 'OR (relative to pre-IRS period)',
+         title = 'ORs during malaria season') +
+    theme_maragra() 
+}
+visualize_ors(ors_object = ors_propensity)
+```
+
+\vfill
+\newpage
+
+Sick absence only:
+
+\begin{tiny}
+
+```{r}
+kable(ors_propensity_sick) 
+```
+
+\end{tiny}
+
+## Ever IRS'ers compared with never-IRS'ers
+
+```{r}
+x <- model_data %>%
+  group_by(days_since) %>%
+  summarise(absences = length(which(absent)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100)
+y <- ab_panel %>%
+  # filter(oracle_number %in% matched$oracle_number) %>%
+  filter(!unidade %in% sort(irs$unidade)) %>%
+  summarise(days_since = 'Never',
+            absences = length(which(absent)),
+            eligibles = n()) %>%
+  mutate(absenteeism_rate = absences / eligibles * 100)
+x <- bind_rows(x, y)
+x$days_since <- factor(x$days_since,
+                       levels = unique(c('Never',
+                                  'Before',
+                                  sort(unique(x$days_since)))))
+cols <- c('red', rep('blue', nrow(x) - 1))
+ggplot(data = x,
+       aes(x = days_since,
+           y = absenteeism_rate)) +
+  geom_bar(stat = 'identity',
+           alpha = 0.6,
+           fill = cols) +
+  theme_maragra() +
+  labs(x = 'Days since IRS',
+       y = 'Absenteeism rate',
+       title = 'Raw absenteeism as a function of days since IRS') +
+  geom_label(aes(label = paste0(round(absenteeism_rate, digits = 2), '%')))
+```
+
+## Regression discontinuity analysis {.allowframebreaks}
+
+\vspace{2mm}
+
+All absenteeism:
+
+\begin{tiny}
+
+```{r}
+kable(ors) 
+```
+
+\end{tiny}
+
+\vfill
+\newpage
+
+```{r}
+visualize_ors(ors_object = ors)
+```
+
+\vfill
+\newpage
+
+Sick absenteeism only:
+
+\begin{tiny}
+
+```{r}
+kable(ors_sick) 
+```
+
+\end{tiny}
 
@@
 
+- 30-50% reduction in absenteeism in the 6 months after IRS, when administered during malaria season.
+
 ## Implications
+
+- Reduction is large and significant.
+- Depending on detailed cost data (pending), IRS may be effective even from a purely financial point of view.
 
@@
 
+- No analysis yet of different worker types (agricultural vs. industrial).
+- Have not yet ventured at all into side-analyses (effect on employment, tonnage, etc.). 
+- Sick absenteeism seems to track absenteeism poorly: lack of clarity regarding pathways.
+
 ## Thank you
+
+Your "pre-publication peer review" comments are appreciated:
 
@@
 
-## References
+## References {.allowframebreaks}

### 2017-10-30 inst/rmd/ihmt_presentation_brew_2017.Rmd
minor changes

@@
 title: "Effectiveness of private sector malaria control: The case of sugarcane workers in Mozambique"
-subtitle: 'IHMT, October, 2017, Lisboa, Portugal'
+subtitle: 'IHMT, October, 2017, Lisboa, Portugal economicsofmalaria.com/ihmt'
 fig_height: 5
@@
   mutate(time_period = make_season(date = date)) %>%
-  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) #%>%
+  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) %>%
+  mutate(month = format(date, '%b'))
   # # remove the period 30 days after spraying
@@
 - Malaria control is cost-effective from the societal/public perspective [@Purdy_2013].
-- Indoor residual spraying (IRS) is cost-effective [@Howard_2017], [@White_2011].
+- Indoor residual spraying (IRS) is cost-effective [@Howard_2017], [@White_2011] _from a public perspective_.
 
@@
 
-## Descriptive: sick absenteeism by time from/to intervention
+<!-- ## Descriptive: sick absenteeism by time from/to intervention -->
 
-```{r}
-ggplot(data = x,
-       aes(x = months_since,
-           y = sick_absenteeism_rate,
-           group = before)) +
-  geom_smooth() +
-  geom_step(color = 'darkred', alpha = 0.8) +
-  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
-  labs(x = 'Months relative to IRS',
-       y = 'Absenteeism rate',
-       title = 'Before/after IRS',
-       subtitle = 'Sick absences only') +
-  theme_maragra()
-```
+<!-- ```{r} -->
+<!-- ggplot(data = x, -->
+<!--        aes(x = months_since, -->
+<!--            y = sick_absenteeism_rate, -->
+<!--            group = before)) + -->
+<!--   geom_smooth() + -->
+<!--   geom_step(color = 'darkred', alpha = 0.8) + -->
+<!--   geom_vline(xintercept = 0, lty = 2, alpha = 0.5) + -->
+<!--   labs(x = 'Months relative to IRS', -->
+<!--        y = 'Absenteeism rate', -->
+<!--        title = 'Before/after IRS', -->
+<!--        subtitle = 'Sick absences only') + -->
+<!--   theme_maragra() -->
+<!-- ``` -->
 
-## Descriptive: absenteeism by time from/to intervention (by time_period)
+## Descriptive: absenteeism by time from/to intervention (by time period)
 
@@
 
-## Descriptive: sick absenteeism by time from/to intervention (by time_period)
+<!-- ## Descriptive: sick absenteeism by time from/to intervention (by time_period) -->
 
-```{r}
+<!-- ```{r} -->
 
-ggplot(data = x %>% mutate(time_period = paste0(time_period)),
-       aes(x = months_since,
-           y = sick_absenteeism_rate,
-           group = before)) +
-  geom_smooth() +
-  geom_step(color = 'darkred', alpha = 0.8) +
-  facet_wrap(~time_period, ncol = 1) +
-  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
-  labs(x = 'Months relative to IRS',
-       y = 'Absenteeism rate',
-       subtitle = 'Sick absences only') +
-  theme_maragra()
-```
+<!-- ggplot(data = x %>% mutate(time_period = paste0(time_period)), -->
+<!--        aes(x = months_since, -->
+<!--            y = sick_absenteeism_rate, -->
+<!--            group = before)) + -->
+<!--   geom_smooth() + -->
+<!--   geom_step(color = 'darkred', alpha = 0.8) + -->
+<!--   facet_wrap(~time_period, ncol = 1) + -->
+<!--   geom_vline(xintercept = 0, lty = 2, alpha = 0.5) + -->
+<!--   labs(x = 'Months relative to IRS', -->
+<!--        y = 'Absenteeism rate', -->
+<!--        subtitle = 'Sick absences only') + -->
+<!--   theme_maragra() -->
+<!-- ``` -->
 
-## Descriptive: absenteeism in binned time period
+<!-- ## Descriptive: absenteeism in binned time period -->
 
-```{r}
-x <- model_data %>%
-  group_by(days_since) %>%
-  summarise(absences = length(which(absent)),
-            eligibles = length(absent)) %>%
-  ungroup %>%
-  mutate(absenteeism_rate = absences / eligibles * 100)
+<!-- ```{r} -->
+<!-- x <- model_data %>% -->
+<!--   group_by(days_since) %>% -->
+<!--   summarise(absences = length(which(absent)), -->
+<!--             eligibles = length(absent)) %>% -->
+<!--   ungroup %>% -->
+<!--   mutate(absenteeism_rate = absences / eligibles * 100) -->
 
-ggplot(data = x,
-       aes(x = days_since,
-           y = absenteeism_rate)) +
-  geom_bar(stat = 'identity',
-           alpha = 0.6,
-           fill = 'darkblue') +
-  theme_maragra() +
-  labs(x = 'Days since IRS',
-       y = 'Absenteeism rate',
-       title = 'Raw absenteeism as a function of days since IRS') +
-  geom_label(aes(label = paste0(round(absenteeism_rate, digits = 2), '%')))
-```
+<!-- ggplot(data = x, -->
+<!--        aes(x = days_since, -->
+<!--            y = absenteeism_rate)) + -->
+<!--   geom_bar(stat = 'identity', -->
+<!--            alpha = 0.6, -->
+<!--            fill = 'darkblue') + -->
+<!--   theme_maragra() + -->
+<!--   labs(x = 'Days since IRS', -->
+<!--        y = 'Absenteeism rate', -->
+<!--        title = 'Raw absenteeism as a function of days since IRS') + -->
+<!--   geom_label(aes(label = paste0(round(absenteeism_rate, digits = 2), '%'))) -->
+<!-- ``` -->
 
-## Descriptive: absenteeism in binned time periods by season
+<!-- ## Descriptive: absenteeism in binned time periods by season -->
 
-```{r}
-x <- model_data %>%
-  group_by(days_since, time_period) %>%
-  summarise(absences = length(which(absent)),
-            eligibles = length(absent)) %>%
-  ungroup %>%
-  mutate(absenteeism_rate = absences / eligibles * 100)
+<!-- ```{r} -->
+<!-- x <- model_data %>% -->
+<!--   group_by(days_since, time_period) %>% -->
+<!--   summarise(absences = length(which(absent)), -->
+<!--             eligibles = length(absent)) %>% -->
+<!--   ungroup %>% -->
+<!--   mutate(absenteeism_rate = absences / eligibles * 100) -->
 
-ggplot(data = x %>% mutate(time_period = paste0(time_period)),
-       aes(x = days_since,
-           y = absenteeism_rate)) +
-  geom_bar(stat = 'identity',
-           alpha = 0.6,
-           fill = 'darkblue') +
-  theme_maragra() +
-  labs(x = 'Days since IRS',
-       y = 'Absenteeism rate',
-       title = 'Raw absenteeism as a function of days since IRS by time_period') +
-  # geom_label(aes(label = paste0(round(absenteeism_rate, digits = 2), '%'))) +
-  facet_wrap(~time_period, ncol = 1) +
-  theme(axis.text.x = element_text(angle = 90))
-```
+<!-- ggplot(data = x %>% mutate(time_period = paste0(time_period)), -->
+<!--        aes(x = days_since, -->
+<!--            y = absenteeism_rate)) + -->
+<!--   geom_bar(stat = 'identity', -->
+<!--            alpha = 0.6, -->
+<!--            fill = 'darkblue') + -->
+<!--   theme_maragra() + -->
+<!--   labs(x = 'Days since IRS', -->
+<!--        y = 'Absenteeism rate', -->
+<!--        title = 'Raw absenteeism as a function of days since IRS by time_period') + -->
+<!--   # geom_label(aes(label = paste0(round(absenteeism_rate, digits = 2), '%'))) + -->
+<!--   facet_wrap(~time_period, ncol = 1) + -->
+<!--   theme(axis.text.x = element_text(angle = 90)) -->
+<!-- ``` -->
 
-## Modeling after matching {.allowframebreaks}
+## Modeling after matching I
 
@@
 
-\vfill
-\newpage
+## Modeling after matching II
 
@@
 
-\vfill
-\newpage
+## Modeling after matching III
 
@@
 
-\vfill
-\newpage
+<!-- \vfill -->
+<!-- \newpage -->
 
-Sick absence only:
+<!-- Sick absence only: -->
 
-\begin{tiny}
+<!-- \begin{tiny} -->
 
-```{r}
-kable(ors_propensity_sick) 
-```
+<!-- ```{r} -->
+<!-- kable(ors_propensity_sick)  -->
+<!-- ``` -->
 
-\end{tiny}
+<!-- \end{tiny} -->
 
@@
 
-\vfill
-\newpage
+<!-- \vfill -->
+<!-- \newpage -->
 
-Sick absenteeism only:
+<!-- Sick absenteeism only: -->
 
-\begin{tiny}
+<!-- \begin{tiny} -->
 
-```{r}
-kable(ors_sick) 
-```
+<!-- ```{r} -->
+<!-- kable(ors_sick)  -->
+<!-- ``` -->
 
-\end{tiny}
+<!-- \end{tiny} -->
 
@@
 - 30-50% reduction in absenteeism in the 6 months after IRS, when administered during malaria season. 
+- Next steps are incorporation of (a) productivity data (via cane cut), (b) better clinical data (via local health facilities), and (c) full cost data.  
+- Will also be comparing with a sugarcane facility in a zone where an elimination campaign is taking place.
 
@@
 - Reduction is large and significant.
-- Depending on detailed cost data (pending), IRS may be effective even from a purely financial point of view.
+- Depending on detailed cost data (pending), IRS may be effective even from a purely financial point of view (ie, beyong just "corporate social responsibility").
 

### 2017-11-06 inst/rmd/cop_2017.Rmd
presentation update

--- 2017-10-30 inst/rmd/ihmt_presentation_brew_2017.Rmd
@@
 title: "Effectiveness of private sector malaria control: The case of sugarcane workers in Mozambique"
-subtitle: 'IHMT, October, 2017, Lisboa, Portugal economicsofmalaria.com/ihmt'
+subtitle: 'Malaria Economics Community of Practice, November, 2017, Baltimore, Maryland, USA economicsofmalaria.com/cop.pdf'
 fig_height: 5
@@
 break_days_since <- function(days_since){
-  out <- ifelse(days_since > 270, '270+',
-    ifelse(days_since > 180, '181-270',
+  out <- ifelse(days_since > 180, '181+',
            ifelse(days_since > 90, '091-180',
@@
                                 ifelse(days_since >= 0, '000-030',
-                                       ifelse(days_since < 0, 'Before', NA)))))))
+                                       ifelse(days_since < 0, 'Before', NA))))))
   out <- factor(out,
@@
                            '091-180',
-                           '181-270',
-                           '270+'))
+                           '181+'))
   return(out)
@@
   # left_join(matched) %>%
+  mutate(spray_month = format(date + days_since, '%b')) %>%
   mutate(days_since = break_days_since(days_since)) %>%
@@
 
+# new fit
+# new_fit <- lm(absent ~ days_since * time_period + month, data = model_data)
+
 # Need to adjust for houses on and off site
@@
 - Cost-benefit studies often only consider the costs of an intervention and associated costs of care, without quantifying the societal cost of non-intervention.  
+- In elimination context, scaling-up private sector involvement is very appealing.
 
@@
 
-## Regression discontunity analysis
+## Regression discontinuity analysis
 
@@
 
+## Back of the envelope calculations
+
+```{r}
+# Get yearly absences / presences
+x <- model_data %>%
+  filter(time_period == 'Malaria season') %>%
+  group_by(days_since) %>%
+  summarise(absences = length(which(absent)),
+            eligibles = n()) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100)
+```
+
+**Savings**
+
+- In percentage point terms, reduction from 13% to 8%.  
+- 5 annually prevented absences per worker.   
+- 8,000 workers: 40,000 prevented absences, wage of 3 USD
+- TOTAL: 120,000 USD in productivity-only savings.
+
+**Costs**
+
+- 8 IRS workers, 1500 USD yearly = 12,000 USD  
+- ACT + DDT: 50,000 USD 
+- Facilities, vehicles, gas: 50,000 USD  
+- TOTAL: 112,000 USD in IRS-only costs
+
+**7% ROI** (ignoring clinical costs)
+
 # Discussion
@@
 
-- 30-50% reduction in absenteeism in the 6 months after IRS, when administered during malaria season. 
+- 30-50% reduction in absenteeism in the 6 months after IRS during malaria season. 
+- Depending on detailed cost data (pending), IRS may be effective even from a purely financial point of view (ie, beyond just "corporate social responsibility").  
 - Next steps are incorporation of (a) productivity data (via cane cut), (b) better clinical data (via local health facilities), and (c) full cost data.  
 - Will also be comparing with a sugarcane facility in a zone where an elimination campaign is taking place.
-
-## Implications
-
-- Reduction is large and significant.
-- Depending on detailed cost data (pending), IRS may be effective even from a purely financial point of view (ie, beyong just "corporate social responsibility").
 
@@
 - Sick absenteeism seems to track absenteeism poorly: lack of clarity regarding pathways.  
+- Large sample size, but all from same place: questionable generalizability.
 
@@
 
-**Presentation:** [economicsofmalaria.com/ihmt](http://economicsofmalaria.com/ihmt)
+**Presentation:** [economicsofmalaria.com/ihmt.pdf](http://economicsofmalaria.com/ihmt.pdf)
 

### 2017-11-27 inst/rmd/defunct/ethical_statement.Rmd
major update

---
title: Ethical considerations in regards to "Effectiveness of private sector malaria control at a large sugar
  facility in Southern Mozambique" (PhD project)
date: ''
output: 
  pdf_document:
    includes:
      in_header: preamble_ethical_statement.sty
fig_width: 4
fig_height: 2.6
bibliography: bibliography.bib
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(knitr)
library(googleVis)
library(RColorBrewer)
op <- options(gvis.plot.tag='chart')

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

\begin{center}
\begin{large}

```{r, results = 'asis', eval = TRUE}
make_authors()
```

\end{large}
\end{center}

\vspace{5mm}

\begin{center}
\textbf{Study summary}  
\end{center}

\vspace{5mm}
\begin{center}
\begin{changemargin}{3cm}{3cm}

This project aims to assess the return on investment of malaria control carried out by a sugar cane company in Southern Mozambique. \textbf{Justification:} Understanding the cost-benefit ratio of malaria control from an investment standpoint has important implications for the potential scaling-up of private sector involvement in the push for global malaria eradication. \textbf{Hypothesis:} We hypothesize that investment in malaria control has a strong impact and offers a high return in terms of reduced workers’ absenteeism and increased workers’ productivity. \textbf{Data: } We gather data pertaining to absenteeism and productivity information from a large sugarcane factory in Southern Mozambique based in a malaria control zone; Private malaria control program costs and activity data; Private worker clinic data. \textbf{Methodology: } We employ ordinary least squares regression to quantify the effect of indoor residual spraying (IRS) on worker absenteeism and productivity; logistic regression to quantify the same effect on the odds of clinical malaria. Finally, we carry out cost-benefit analysis using company financial data.

\end{changemargin}
\end{center}

# Institutional approval

This project was approved by the Scientific and Ethical Committees of the Centro de Investigação em Saude de Manhiça (Mozambique) as part of the larger economic research related to the Mozambican Alliance Towards the Elimination of Malaria (MALTEM) campaign.

# Confidentiality

All information on individuals will remain confidential and be shared only by the study team. Identifiable data is required, in order to join information from company registers with the government census. Following data merge, unique identifiers will be used with only the PI (Joe Brew) having access to the decryption key. In all cases, the PI will ensure that the completed identification code list are kept in locked and encrypted files.

# Consent

This study is observational, using only data which is already collected and stored on company registers. Much of the data pertains to workers who are no longer active in the company, or resident in the area; in some cases, the workers have left the census catchment area or even deceased. For these reasons, direct consent is not obtained from the workers; rather, it is provided by the company (the proprietor of the data), through an explicit data sharing agreement stipulating the non-disclosure of worker identity. 
# Beneficience

The potential knowledge gained from this product stands to improve the health of workers by providing insight as to which methods are most effective in preventing malaria infection.

# Beneficence

The aim of this work is to determine the effectiveness of indoor residual spraying on the lived experience of workers, through the proxies of absenteeism and clinical illness. It imposes no intervention on the workers (the intervention under study is already carried out regularly by the company), and its findings are likely to improve the operations of a generally effective intervetion (as demonstrated elsewhere) [@Lee2017, @Howard_2017, @Bloom2008]. Importantly, if the findings of this study suggest a positive return on investment for private sector malaria control activities, we might expect that the results' dissemination could motivate more companies to carry out malaria control in Sub-Saharan Africa, which would improve the health of African workers [@McCarthy_2000].

# Non-malfeasance

All data used are observational and retrospective in nature. No direct contact with workers is made. We have no concerns regarding the potential of causing harm to workers.

# Conflicts of interest and independence

The company (Illovo Açucar) plays an important role in data acquisition and understanding. This is a potential conflict of interest, as it is feasible to share intentionally incomplete or incorrect data. The researchers understand this risk and will work diligently to avoid it. All analysis is carried out by the PI, with no influence or oversight from the company.

# Knowledge sharing and dissemination

All research products will be shared with the company. Publications will be open-access. A presentation will be made on-site with both company management and workers invited to attend. Though data must be stored privately (for reasons of confidentiality), all code will be open source, so as to allow for reproducibility and ensure transparency.\footnote{https://github.com/joebrew/maragra}

```{r, eval=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'knitr', 'rmarkdown'
), 'packages.bib')
```

# References

### 2017-11-27 inst/rmd/defunct/ihmt_presentation_brew_2017.Rmd
major update

--- 2017-11-06 inst/rmd/cop_2017.Rmd
@@
 title: "Effectiveness of private sector malaria control: The case of sugarcane workers in Mozambique"
-subtitle: 'Malaria Economics Community of Practice, November, 2017, Baltimore, Maryland, USA economicsofmalaria.com/cop.pdf'
+subtitle: 'IHMT, October, 2017, Lisboa, Portugal economicsofmalaria.com/ihmt.pdf'
 fig_height: 5

### 2017-11-27 inst/rmd/defunct/maragra_clinical_malaria_incidence.Rmd
major update

--- 2017-08-22 analyses/analyze_monthly_data.Rmd
@@
 ---
-
-```{r setOptions, message=FALSE, echo = FALSE}
-library(googleVis)
-op <- options(gvis.plot.tag='chart')
-```
 
@@
 
-```{r, results = 'asis'}
-gg <- gvisLineChart(data = clinic_agg %>% filter(group == 'nacionais'),
-                    xvar = 'date',
-                    yvar = c('tested', 'positive'),
-                    options = list(curveType='function',
-                                  title="Malaria among Mozambican employees",
-                                  titleTextStyle="{color:'black',fontName:'Courier',fontSize:16}"))
-plot(gg)
-
-gg <- gvisLineChart(data = clinic_agg %>% filter(group == 'expatriados'),
-                    xvar = 'date',
-                    yvar = c('tested', 'positive'),
-                    options = list(curveType='function',
-                                  title="Malaria among expatriated employees",
-                                  titleTextStyle="{color:'black',fontName:'Courier',fontSize:16}"))
-plot(gg)
-# 
-# # Heat calendar chart
-# gvisCalendar
-
-# # # Motion chart
-# x <- gvisMotionChart(clinic_agg,
-#    idvar="group",
-#    timevar="date")
+```{r}
+ggplot(data = clinic_agg %>% filter(group == 'nacionais'),
+            aes(x = date,
+                y = tested)) +
+  geom_line() +
+  geom_line(aes(y = positive)) +
+  theme_maragra() +
+  labs(x = 'Date',
+       y = 'People (tested and positive)',
+       title = 'Mozambicans')
+ggplot(data = clinic_agg %>% filter(group != 'nacionais'),
+            aes(x = date,
+                y = tested)) +
+  geom_line() +
+  geom_line(aes(y = positive)) +
+  theme_maragra() +
+  labs(x = 'Date',
+       y = 'People (tested and positive)',
+       title = 'Expatriates')
 ```

### 2017-11-27 inst/rmd/paper.Rmd
presentation update
major update

--- 2017-09-01 inst/rmd/paper.Rmd
@@
 library(knitr)
-library(googleVis)
 library(RColorBrewer)
-op <- options(gvis.plot.tag='chart')
-
+library(extrafont)
+loadfonts()
 ## Global options
@@
 opts_chunk$set(echo=FALSE,
-	             cache=TRUE,
+	             cache=FALSE,
                prompt=FALSE,
@@
                message=FALSE,
-               warning=FALSE)
+               warning=FALSE,
+               dev = "cairo_pdf")
 opts_knit$set(width=75)
@@
 
-<!-- <hr> -->
-<!-- <p style="text-align:left;"> -->
-<!-- [http://economicsofmalaria.com](http://economicsofmalaria.com) -->
-<!-- <span style="float:right;">joebrew@gmail.com</span> -->
-<!-- </p> -->
-<!-- <hr> -->
+```{r}
+source('paper_prepare.R')
+```
 
-```{r}
-# Read in data
-ab <- maragra::ab
-ab_panel <- maragra::ab_panel
-bairros <- maragra::bairros
-census <- maragra::census
-clinic <- maragra::clinic
-clinic_agg <- maragra::clinic_agg
-mc <- maragra::mc
-workers <- maragra::workers
+\begin{center}
+\begin{large}
+
+```{r, results = 'asis', eval = TRUE}
+make_authors()
 ```
+
+\end{large}
+\end{center}
+
+\vspace{5mm}
+
+\begin{center}
+\textbf{Abstract}  
+\end{center}
+
+\vspace{5mm}
+\begin{center}
+\begin{changemargin}{3cm}{3cm}
+
+\textbf{Context:} Incentivizing scaled-up private firms' involvement in malaria control and elimination activities requires a clear and robust demonstration of its benefits, and a quantification of which benefits can be expected to return to the firm.
+However, most studies to date examine benefits of malaria control purely from the perspective of the public sector.
+\textbf{Methods:} We analyze 4 years of malaria control and worker health and absenteeism data from a large sugar processing facility in Mozambique.
+We estimate the effect of indoor residual spraying (IRS) on worker health and absenteeism.
+\textbf{Results:} We demonstrate that the firm's engagement in malaria control activities is profitable, even if we only consider those benefits which directly affect the firm's bottom line and ignore secondary and societal benefits.
+
+\end{changemargin}
+\end{center}
 
@@
 
+Malaria has a nearly unquantifiably large economic impact on endemic societies.
+By affecting saving, investment [@Shretta2016], risk perception, productivity, absenteeism [@Nonvignon_2016], human capital accumulation [@CastelBranco2014], mortality, and costs of care [@Sachs2002], malaria likely has a negative effect on GDP and growth [@McCarthy_2000] [@Orem_2012].
+Because of the relative affordability of most intereventions and the enormous societal costs of malaria, most forms of malaria control are cost-effective from the perspective of the government [@White_2011] [@Purdy_2013] [@Howard_2017].
+
+From the persepctive of the private sector, however, investing in malaria control is not such a clear-cut case.
 Public health interventions targetting malaria - and their corresponding cost-effectiveness evaluations - most often pertain to the public sector.
-Accordingly, the analytical framework and metrics employed by these analyses most often focus on impact pertaining to the public good, such as an increase in life years adjusted for disability or quality  [@Goodman1999, @Shretta2016, @Lee2017, Hanson2004].
-Though population-level health is certainly of importance to businesses, and improvements in health incidentally improve the economy at all levels [@World1999, @Bloom2008], these improvements may be to disperse or long-term to incentivize private sector involvement in health campaigns.
+Accordingly, the analytical framework and metrics employed by these analyses most often focus on impact pertaining to the public good, such as an increase in life years adjusted for disability or quality  [@Goodman1999] [@Shretta2016] [@Lee2017] [@Hanson2004].
+Though population-level health is certainly of importance to businesses, and improvements in health incidentally improve the economy at all levels [@World1999] [@Bloom2008] [@Vecchi_2013], these improvements may be too disperse or long-term to incentivize private sector involvement in health campaigns.
 
@@
 At times, this role is necessary from a purely practical standpoint; in other cases, it is employed under the guise of "corporate social responsibility" [@Azemar2009].
-Regardless of the language used, it is clear that private industry plays an important role in public health in Mozambique [@Robbins2012, @CastelBranco2014].
+Regardless of the language used, it is clear that private industry plays an important role in public health in Mozambique [@Robbins2012] [@CastelBranco2014].
 
-Several cases exist of foreign firms engaging in large-scale anti-malaria campaigns [@Mouzin2011, @Han].
-But these cases rarely undergo the same rigorous evaluation as their public sector counterparts due to (a) unavailability of data, (b) identification strategy, etc.
+Several cases exist of foreign firms engaging in large-scale malaria control campaigns [@Mouzin2011] [@Han] [@Bennett_2017] [@Kaula_2017].
+But these studies generally consider population health as the outcome measure of interest, rather than worker absenteeism or productivity.
+Similarly, they often neglect to differentiate between those clinical costs which are absorbed by the local health system versus those which are absorbed by the firm itself.
+In the literature, making the "investment case" for malaria control or elimination generally implies that the investor is the public sector, and takes into account those costs and benefits which are applicable to the public sector [@Shretta_2017]; though appropriate in most cases (the public sector being the primary malaria control agent in most locations), the findings of these studies are rarely applicable to the private sector.
+In the case of a private firm not interested in "corporate social responsibility", it is not clear whether investing in malaria control would be profitable or not.
+This lack of certainty most likely discourages investment in many cases.
 
-In this paper, we...
-
-We hypothesize that..
-
-We find that...
+To address the question of the profitability of malaria control activities from the standpoint of a private firm, we analyze data during a 7 year period from a private sugar facility in Southern Mozambique.
+We assess the effect of indoor residual spraying (IRS) on the absenteeism and health of workers, and demonstrate that the firm's engagement in malaria control not only improved worker health, but also generated a positive return on investment.
 
@@
 
-```{r}
-# # Create model data
-# model_data <-
-#   ab_panel %>%
-#   left_join(irs) %>%
-#   left_join(workers %>%
-#               dplyr::select(oracle_number,
-#                             sex,
-#                             race_code,
-#                             department,
-#                             permanent_or_temporary)) %>%
-#   mutate(days_since = ifelse(is.na(days_since), 'Never',
-#                              ifelse(days_since > 180, '180+',
-#                                     ifelse(days_since > 90, '090-180',
-#                                            ifelse(days_since > 60, '060-090', 
-#                                                   '<060')))))
-# # Need to add seasonality
-# # Need to adjust for houses on and off site
-# fit <- glm(absent ~ days_since,
-#            family = binomial('logit'),
-#           data = model_data)
-# exp(coef(fit))
-```
+In collaboration with the sugar processing facility, Maragra Açucar SA (henceforth referred to as "Maragra"), we collected data for the period from January 2010 through December 2016.
+Data came from four sources: (i) the Human Resources' roster of worker details and absences, (ii) the facility's on-site clinic's medical and laboratory records, (iii) the facility's on-site malaria control program's records pertaining to the dates, chemicals, and location of IRS activities, and (iv) interviews with company employees pertaining to costs, data limitations, etc.
+Digitization and collection of data took place during the period from March 2016 through May 2017.
+Supplementary data pertaining to worker characteristics was obtained from through the Centro de Investigação em Saude de Manhiça's (CISM) demographic census, which covered workers from the district, but not those who migrated from other parts of the country [@Nhacolo_2006].
 
-Some text goes here.
+Data pertaining to district-wide malaria incidence was obtained from Mozambique’s Boletim Epidemiológico Semanal (BES), which is the system by which the National Malaria Control Program monitors incidence at the district level throghout the entire country, and reports
+the number of confirmed weekly malaria cases at government health facilities.
+Using these case numbers, combined with population estimates from the National Statistical Institute (INE), we estimate each day's annualized weekly malaria incidence rate (cases per 1000 population at risk), interpolated from the weekly figures.
+We retrieved weather data for all Mozambican stations from NOAA.
+We estimated the meteorological conditions at the centroid of Manhiça using a simple interpolation method whereby the district's weather conditions were estimated to be a function of all Mozambican weather stations' reported conditions, inversely weighted by kilometers from district centroid.
 
-All data processing and analysis were carried out in R [@R]. All code is freely available online [@brewgit].
+Maragra regularly employs IRS at on-site worker households in order to reduce those workers' (and their families') risk of malaria infection.
+By using the CISM's census, we were able to identify from the record of houses sprayed (maintained by the company) who lived in those houses, as well as who lived off-site (and therefore did not receive IRS from Maragra).
+Then, using company HR and clinical records, we were able to identify absences and episodes of clinical malaria among all workers, as well as identify the time since the most recent IRS episode before the onsent of absence or illness.
+
+We sought to understand the effect of IRS on individual workers' likelihood of absence from work as well as their likelihood of clinical malaria.
+To estimate this effect, we estimated separate models for absence and illness.
+We employed interrupted time series [@Lopez_Bernal_2016] and a linear probability approach using the following econometric model.
+
+$$
+\hat{Y} = \beta_{0} +  \beta_{1} \text{Season} + (\beta_2{IRS}*\beta_3{IRS_t}) + ... 
+$$
+
+$\hat{Y}$ is the rate of (illness/absence).
+$\beta_{1}$ represents the clinical malaria incidence at that time in the entire district of Manhiça.
+Our demographic confounders (represented by $...$) are sex, age, and worker department (field, factory, or administrative).
+Our intervention was not a simple yes/no, but rather the product of whether the residence of the worker in question was treated in the last year, and, if so, the time since treatment (represented above as the interaction term, where where $_t$ represents time elapsed since commencement of the most recent IRS campaign).
+We considered using weather data as a right-hand side variable for our models, but found that district-level clinical malaria incidence was a far better predictor of both absence and clinical malaria infection than precipitation and temperature, the former apparently capturing weather's significant effects.
+Accordingly, we use clinical malaria incidence as a proxy for the severity of malarial activity, and exclude weather variables from our modeling.
+
+Worker characteristics, illness and absenteeism data, along with IRS activity data, were systematically stored, collected, and used at the individual level by Maragra, and therefore of generally high quality.
+Because cost data was less systematically collected by Maragra, and because many costs could not be precisely quantified due to the abundance of in-kind and cross-departmental expenditures, we had to rely on rough estimations based on a mix of interviews, reciepts, and interpolations.
+Since our program cost data is not as reliable as our worker characteristic and outcome data, we were conservative in our estimates, and generally tried to err on the side of program activities and materials costing _more_ than what was reported, when doubt was aired.
+Cost data consisted of three types: (i) wages of malaria control employees, (ii) transporation and vehicle costs for IRS teams, and (iii) acquisition costs of purchasing IRS chemicals for fumigation (ACT and DDT).
+We define the IRS program as "profitable" if return on investment is greater than 100%, ie if the savings associated with the estimated effect of IRS is greater than the costs of the program's administration.
+
+All data processing and analysis were carried out in R [@R] and all analysis code is freely available online [@brewgit].
+Ethical approval for this project was obtained from the Institutional Ethics Review Board for Health at CISM prior to data collection.
 
@@
 
-Some text goes here
+In Southern Mozambique, malaria peaks during the summer months (December through March) most years (Figure 1, panel A), and worker absenteeism rates track malaria incidence closely (Figure 1, panel B).
+Both all-cause absenteeism and sick absenteeism have declined in recent years (Figure 1, panel C), with the latter declining at a faster rate than the former.
+The fact that the rate of confirmed cases at the company clinic is largely non-seasonal (Figure 1, panels E and F) suggests that a significant portion of workers either seek care for malaria elsewhere (for example, government health posts, of which several are nearby and in some many cases closer to workers' residence than the company clinic) or did not seek care during malaria infection.
+Accordingly, we focus our analysis on all-cause absenteeism rather than sickness absenteeism or malaria diagnostics, assuming that much of illness is captured by absenteeism but not by the clinic.
+
+```{r, fig.cap = 'Clinical malaria (district of Manhiça), all-cause absenteeism among Maragra workers, sick absenteeism among Maragra workers, estimated rainfall, positive cases at company clinic, and test positivity rate at company clinic', fig.height=5, fig.width=8}
+# BES incidence
+g1 <- 
+  ggplot(data = bes,
+       aes(x = date,
+           y = p)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Incidence',
+       title = 'A: Annualized district clinical\nmalaria incidence per 1,000')
+
+# Absenteeism rate
+g2_data <- ab_panel %>%
+  group_by(month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+  summarise(absences = length(which(absent)),
+            absences_sick = length(which(absent_sick)),
+
+            eligibles = n()) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100,
+         sick_absenteeism_rate = absences_sick / eligibles * 100)
+
+g2 <- ggplot(data = g2_data,
+             aes(x = month,
+                 y = absenteeism_rate)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Absenteeism rate',
+       title = 'B: Monthly absenteeism rate')
+
+g3 <- ggplot(data = g2_data,
+             aes(x = month,
+                 y = sick_absenteeism_rate)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Absenteeism rate',
+       title = 'C: Monthly sick absenteeism rate')
+
+# Weather
+g4 <-
+  ggplot(data = weather %>%
+           group_by(date = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+           summarise(precipitation = sum(precipitation, na.rm = TRUE)),
+         aes(x = date,
+             y = precipitation)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Mililiters',
+       title = 'D: Monthly precipitation')
+
+# Clinical incidence of malaria
+g5 <- clinic_agg %>%
+  filter(date >= '2013-01-01', date <= '2016-12-31') %>%
+  # mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
+  group_by(date) %>%
+  summarise(tested = sum(tested),
+            positive = sum(positive)) %>%
+  ungroup %>%
+  mutate(percent_positive = positive / tested * 100) %>%
+  ggplot(aes(x = date,
+             y = positive)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Cases',
+       title = 'E: Monthly malaria cases\nat company clinic')
+
+g6 <- clinic_agg %>%
+  filter(date >= '2013-01-01', date <= '2016-12-31') %>%
+  # mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
+  group_by(date) %>%
+  summarise(tested = sum(tested),
+            positive = sum(positive)) %>%
+  ungroup %>%
+  mutate(percent_positive = positive / tested * 100) %>%
+  ggplot(aes(x = date,
+             y = percent_positive)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Cases',
+       title = 'F: Monthly malaria test positivity rate\nat company clinic') 
+Rmisc::multiplot(g1, g2, g3, g4, g5, g6, layout = matrix(1:6, nrow=2, byrow=TRUE)) 
+```
+
+```{r}
+fumigations <- mc %>%
+  filter(date >= '2012-01-01',
+         date <= '2016-12-31')
+n_fumigations <- nrow(fumigations)
+n_buildings <- sum(fumigations$casas_cobertas, na.rm = TRUE)
+unique_agregados <- length(unique(fumigations$unidade))
+people_sprayed <- length(unique(model_data$oracle_number[model_data$days_since != 'Never']))
+people_observed <- length(unique(model_data$oracle_number))
+```
+
+During the period from January 1st, 2012 through December 31st, 2016, the Maragra Malaria Control Unit carried out `r n_fumigations` episodes of fumigation of residential "agregados" (household combinations), for a total of `r n_buildings` building-fumigation combinations.
+The total number of unique agregados sprayed during this period was `r unique_agregados`.
+Among the `r people_observed` on whom we have absenteeism data, `r people_sprayed` had their homes fumigated at least once (the majority of workers live off of the facility).
+
+The below chart shows the effect of IRS. All workers. Reference group is "Never" (those who never got IRS).
+
+```{r, fig.height = 3.5, fig.width = 4, fig.cap= 'Effect of IRS on absenteeism'}
+ggplot(data = fit_tidy %>%
+         filter(!term %in% c('(Intercept)', 'incidence')) %>%
+         filter(grepl('seasonhigh', term)) %>%
+         filter(term != 'seasonhigh') %>%
+         mutate(term = gsub('seasonhigh:days_since', '', term)) %>%
+         mutate(term = factor(term, levels = unique(c('Never',
+                                               'Before',
+                                               sort(unique(term)))))),
+       aes(x = term,
+           y = estimate)) +
+  geom_bar(stat = 'identity') +
+  theme(axis.text.x = element_text(angle = 90)) +
+  theme_maragra() +
+  labs(x = 'Time since IRS',
+       y = 'Coefficient')
+```
+
+\vfill
+
+\pagebreak
 
@@
 
-```{r include=FALSE}
-# automatically create a bib database for R packages
-knitr::write_bib(c(
-  .packages(), 'knitr', 'rmarkdown'
-), 'packages.bib')
-```
-
 # References

### 2017-11-27 R/theme_maragra.R

@@ -1,24 +1,29 @@
 #' Theme for ggplots
-#' 
+#'
 #' Add a theme to a ggplot object
 #' @param base_size The base font size
+#' @oaram font_family The font
 #' @return A ggplot compatible theme call
 #' @import ggthemes
 #' @import grid
 #' @export
 # Borrowed from https://rpubs.com/Koundy/71792
-theme_maragra <- function(base_size=14) {
-  (theme_foundation(base_size=base_size)
-    + theme(plot.title = element_text(face = "bold",
-                                      size = rel(1.2), hjust = 0.5),
+theme_maragra <- function(base_size=14,
+                          font_family = 'Computer Modern Roman') {
+  (theme_foundation(base_size=base_size,
+                    base_family = font_family)
+    + theme(plot.title = element_text(#face = "bold",
+                                      size = rel(1),
+                                      hjust = 0.5),
             text = element_text(),
             panel.background = element_rect(colour = NA),
             plot.background = element_rect(colour = NA),
             panel.border = element_rect(colour = NA),
-            axis.title = element_text(face = "bold",size = rel(1)),
+            axis.title = element_text(#face = "bold",
+                                      size = rel(0.75)),
             axis.title.y = element_text(angle=90,vjust =2),
             axis.title.x = element_text(vjust = -0.2),
-            axis.text = element_text(), 
+            axis.text = element_text(size = rel(0.65)),
             axis.line = element_line(colour="black"),
             axis.ticks = element_line(),
             panel.grid.major = element_line(colour="#f0f0f0"),
@@ -31,6 +36,8 @@ theme_maragra <- function(base_size=14) {
             legend.title = element_text(face="italic"),
             plot.margin=unit(c(10,5,5,5),"mm"),
             strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
-            strip.text = element_text(face="bold")
+            strip.text = element_text(face="bold") +
+            theme(text=element_text(size=16,
+                                      family=font_family))
     ))
 }

### 2017-11-27 data-raw/create_data_files.R

@@ -1157,7 +1157,35 @@ ab_panel <-
               dplyr::filter(census_name_match_score <= 0.1990741) %>%
               dplyr::select(oracle_number, unidade))
 
+# Create BES based incidence data (created in maltem_cost_effectiveness)
+bes <- readr::read_csv('bes_manhica.csv')
+# Expand bes into a daily dataset
+left <- data_frame(date = seq(as.Date('2010-01-01'),
+                              max(bes$date),
+                              by = 1))
+bes <- left_join(left, bes, by = 'date')
+# Fill
+bes <- bes %>% tidyr::fill(cases, population, p, .direction = 'up')
+# Divide by 7 (since we converted from weekly to daily)
+bes <- bes %>% mutate(cases = cases / 7,
+                      population = population,
+                      p = p / 7) %>%
+  mutate(p = p * 365.25) # make yearly
+# Keep only 2013 onwards
+bes <- bes %>% filter(date >= '2013-01-01') %>%
+  # and only til end of 2016
+  filter(date <= '2016-12-31')
+
+# Create NOAA weather (created in maltem_cost_effectiveness)
+weather <- readr::read_csv('weather_daily.csv')
+weather <- weather %>%
+  filter(district == 'MANHICA',
+         date >= '2013-01-01',
+         date <= '2016-12-31')
+
 # Save for use in package
+devtools::use_data(bes,
+                   overwrite = TRUE)
 devtools::use_data(ab,
                    overwrite = TRUE)
 devtools::use_data(clinic,
@@ -1175,6 +1203,8 @@ devtools::use_data(bairros,
 devtools::use_data(census,
                    overwrite = TRUE)
 devtools::use_data(mc,
+                   overwrite = TRUE)
+devtools::use_data(weather,
                    overwrite = TRUE)
 
 ##################################################

### 2017-11-27 inst/rmd/paper_prepare.R

@@ -1,0 +1,49 @@
+# Packages
+library(tidyverse)
+library(knitr)
+library(Hmisc)
+library(brew)
+library(maragra)
+library(knitr)
+library(RColorBrewer)
+
+# Read in data
+ab <- maragra::ab
+ab_panel <- maragra::ab_panel
+bairros <- maragra::bairros
+bes <- maragra::bes
+census <- maragra::census
+clinic <- maragra::clinic
+clinic_agg <- maragra::clinic_agg
+mc <- maragra::mc
+weather <- maragra::weather
+workers <- maragra::workers
+
+
+model_data <-
+  ab_panel %>%
+  left_join(irs, by = c('unidade', 'date')) %>%
+  mutate(days_since = ifelse(days_since > 180, '180+',
+                             ifelse(days_since > 90, '090-180',
+                                    ifelse(days_since > 60, '060-090',
+                                           ifelse(days_since >= 0, '000-060',
+                                                  ifelse(days_since < 0, 'Before', NA)))))) %>%
+  mutate(days_since = ifelse(is.na(days_since), 'Never', days_since)) %>%
+  mutate(days_since = factor(days_since, levels = c('Never', 'Before', '000-060', '060-090', '090-180', '180+'))) %>%
+  mutate(quarter = lendable::quarter_extract(date)) %>%
+  mutate(quarter = as.character(quarter)) %>%
+  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) %>%
+    # Add malaria incidence
+    left_join(bes %>%
+                dplyr::select(date, p) %>%
+                dplyr::rename(incidence = p),
+              by = 'date') %>%
+  mutate(season = ifelse(incidence >= median(incidence),
+                         'high',
+                         'low')) %>%
+  mutate(season = factor(season, levels = c('low', 'high')))
+
+# Model
+fit <- lm(absent ~ season * days_since, data = model_data)
+fit_tidy <- broom::tidy(fit)
+

### 2017-12-02 inst/rmd/paper.Rmd
big updates

@@
 
+Worker characteristics, illness and absenteeism data, along with IRS activity data, were systematically stored, collected, and used at the individual level by Maragra, and therefore of generally high quality.
 We sought to understand the effect of IRS on individual workers' likelihood of absence from work as well as their likelihood of clinical malaria.
@@
 $$
-\hat{Y} = \beta_{0} +  \beta_{1} \text{Season} + (\beta_2{IRS}*\beta_3{IRS_t}) + ... 
+\hat{Y_{it}} = \beta_{0} +  (\beta_{1}) (\text{Season}_{t}) + (\beta_2{IRS}*\beta_3{IRS_t}) + ... + \epsilon
 $$
 
-$\hat{Y}$ is the rate of (illness/absence).
+$\hat{Y}$ is the rate of absence.
 $\beta_{1}$ represents the clinical malaria incidence at that time in the entire district of Manhiça.
@@
 Our intervention was not a simple yes/no, but rather the product of whether the residence of the worker in question was treated in the last year, and, if so, the time since treatment (represented above as the interaction term, where where $_t$ represents time elapsed since commencement of the most recent IRS campaign).
-We considered using weather data as a right-hand side variable for our models, but found that district-level clinical malaria incidence was a far better predictor of both absence and clinical malaria infection than precipitation and temperature, the former apparently capturing weather's significant effects.
-Accordingly, we use clinical malaria incidence as a proxy for the severity of malarial activity, and exclude weather variables from our modeling.
+We define the malaria season as any time during which the clinical incidence of malaria in the district of Manhiça was at or greater than the median clinical incidence of malaria for the entire study period.
+These weeks are flagged as red in Figure 1, Panel A.
+By using clinical incidence of the area of residence of the workers (as opposed to more typical proxies for malaria risk, such as only rainy vs.
+non rainy season), our seasonality estimate is a closer approximation of true malaria risk, incorporating lagged effects such as the incubation period of the parasite, as well as any inherent non-linear effects of weather.
+In addition, we adjust for daily precipitation
 
-Worker characteristics, illness and absenteeism data, along with IRS activity data, were systematically stored, collected, and used at the individual level by Maragra, and therefore of generally high quality.
 Because cost data was less systematically collected by Maragra, and because many costs could not be precisely quantified due to the abundance of in-kind and cross-departmental expenditures, we had to rely on rough estimations based on a mix of interviews, reciepts, and interpolations.
@@
 
+```{r, results='asis'}
+calculate <- function(var){
+  out <- 
+  model_data %>%
+    mutate(year = format(date, '%Y')) %>%
+    mutate(maragra_fabrica = ifelse(maragra_fabrica,
+                                    'On site',
+                                    'Off site'),
+           rainy = ifelse(rainy,
+                          'Rainy',
+                          'Dry')) %>%
+    filter(!is.na(rainy)) %>%
+  group_by_(var, 'year') %>%
+  summarise(absences = length(which(absent)),
+            eligibles = length(absent),
+            workers = length(unique(oracle_number))) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100) 
+  names(out)[1] <- 'variable'
+  out$val <- paste0(round(out$absenteeism_rate, digits = 2),
+                    '%')#,
+                    # ')',
+                    # '\n',
+                    # ' (workers:', out$workers,
+                    # ')')
+  out <- 
+    out %>%
+    dplyr::select(variable, year, val)
+
+  out <- out %>%
+    spread(key = year,
+           value = val)
+  out$variable <- as.character(out$variable)
+  new_var <- ifelse(var == 'maragra_fabrica',
+                    'Residence',
+                    ifelse(var == 'rainy',
+                           'Precipitation',
+                           var))
+  out$variable[1] <- paste0(Hmisc::capitalize(new_var), ': ', out$variable[1])
+  return(out)
+}
+out_list <- list()
+vars <- c('season', 
+           'department',
+           'sex',
+           'maragra_fabrica',
+           'rainy')
+for(i in 1:length(vars)){
+  x <- calculate(var = vars[i])
+  out_list[[i]] <- x
+}
+ab_table <- bind_rows(out_list)
+names(ab_table)[1] <- ''
+print(xtable::xtable(ab_table,
+                     align = c('r' ,'r', rep('l', 4))),
+      include.rownames = FALSE)
+```
+
 ```{r, fig.cap = 'Clinical malaria (district of Manhiça), all-cause absenteeism among Maragra workers, sick absenteeism among Maragra workers, estimated rainfall, positive cases at company clinic, and test positivity rate at company clinic', fig.height=5, fig.width=8}
@@
 g1 <- 
-  ggplot(data = bes,
+  ggplot(data = bes %>%
+           mutate(season = ifelse(p >= median(p), 'High season',
+                                  'Low season')) %>%
+           filter(weekdays(date) == 'Monday'),
        aes(x = date,
            y = p)) +
-  geom_line(alpha = 0.6) +
+  geom_point(alpha = 0.6,
+             aes(color = season),
+             size = 0.2) +
+  geom_line(alpha = 0.1) +
   theme_maragra(base_size = 9) +
@@
        y = 'Incidence',
-       title = 'A: Annualized district clinical\nmalaria incidence per 1,000')
+       title = 'A: Annualized district clinical\nmalaria incidence per 1,000') +
+  scale_color_manual(name = '',
+                     values = c('red', 'black')) + 
+  guides(color = guide_legend(ncol = 1)) +
+  theme(legend.position = c(.8, .8),
+        legend.text = element_text(size = 5),
+        legend.background = element_rect(fill=alpha('white', 0)))
 
@@
 
-The below chart shows the effect of IRS. All workers. Reference group is "Never" (those who never got IRS).
+```{r, fig.height = 3.5, fig.width = 6, fig.cap = 'Effect of IRS on absenteeism by season'}
+# Predict for visualization
+dummy <- expand.grid(season = sort(unique(model_data$season)),
+                     months_since = sort(unique(model_data$months_since)))
+predictions <- predict(fit, dummy, se.fit = TRUE, interval = 'confidence')
+dummy$predicted <- predictions$fit[,1]
+dummy$lwr <- predictions$fit[,2]
+dummy$upr <- predictions$fit[,3]
+ggplot(data = dummy %>%
+         filter(months_since != 'Never') %>%
+         mutate(lwr = lwr * 100,
+                upr = upr * 100,
+                predicted = predicted * 100) %>%
+         mutate(season = paste0(Hmisc::capitalize(as.character(season)), ' season')),
+       aes(x = months_since,
+           y = predicted,
+           group = 1)) +
+  geom_ribbon(aes(x = months_since,
+                  ymin = lwr,
+                  ymax = upr),
+              fill = 'lightblue',
+              alpha = 0.6) +
+  # geom_line(color = 'blue', alpha = 0.5) +
+  geom_point(color = 'blue', alpha = 0.5) + 
+  geom_smooth(se = FALSE,
+              color = 'black',
+              alpha = 0.5,
+              span = 1) +
+  facet_wrap(~season) +
+  theme_maragra() +
+  theme(axis.text.x = element_text(angle = 90, size = 8)) +
+  labs(x = 'Months since IRS',
+       y = 'Absenteeism (%)')
 
-```{r, fig.height = 3.5, fig.width = 4, fig.cap= 'Effect of IRS on absenteeism'}
-ggplot(data = fit_tidy %>%
-         filter(!term %in% c('(Intercept)', 'incidence')) %>%
-         filter(grepl('seasonhigh', term)) %>%
-         filter(term != 'seasonhigh') %>%
-         mutate(term = gsub('seasonhigh:days_since', '', term)) %>%
-         mutate(term = factor(term, levels = unique(c('Never',
-                                               'Before',
-                                               sort(unique(term)))))),
-       aes(x = term,
-           y = estimate)) +
-  geom_bar(stat = 'identity') +
-  theme(axis.text.x = element_text(angle = 90)) +
-  theme_maragra() +
-  labs(x = 'Time since IRS',
-       y = 'Coefficient')
 ```

### 2017-12-02 inst/rmd/paper_prepare.R

@@ -19,31 +19,111 @@ mc <- maragra::mc
 weather <- maragra::weather
 workers <- maragra::workers
 
+# Define function for adding zero
+add_zero <- function (x, n) {
+  x <- as.character(x)
+  adders <- n - nchar(x)
+  adders <- ifelse(adders < 0, 0, adders)
+  for (i in 1:length(x)) {
+    if (!is.na(x[i])) {
+      x[i] <- paste0(paste0(rep("0", adders[i]), collapse = ""),
+                     x[i], collapse = "")
+    }
+  }
+  return(x)
+}
 
 model_data <-
   ab_panel %>%
   left_join(irs, by = c('unidade', 'date')) %>%
-  mutate(days_since = ifelse(days_since > 180, '180+',
-                             ifelse(days_since > 90, '090-180',
-                                    ifelse(days_since > 60, '060-090',
-                                           ifelse(days_since >= 0, '000-060',
-                                                  ifelse(days_since < 0, 'Before', NA)))))) %>%
-  mutate(days_since = ifelse(is.na(days_since), 'Never', days_since)) %>%
-  mutate(days_since = factor(days_since, levels = c('Never', 'Before', '000-060', '060-090', '090-180', '180+'))) %>%
-  mutate(quarter = lendable::quarter_extract(date)) %>%
-  mutate(quarter = as.character(quarter)) %>%
+  mutate(days_since = days_since %/% 30) %>%
+  mutate(days_since = ifelse(is.na(days_since), 'Never',
+                             ifelse(days_since < 0, 'Before',
+                             ifelse(days_since >= 12, '12+', as.character(add_zero(days_since, n = 2)))))) %>%
+  mutate(months_since = factor(days_since, levels = unique(c('Never', 'Before', sort(unique(days_since)))))) %>%
   mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) %>%
-    # Add malaria incidence
-    left_join(bes %>%
-                dplyr::select(date, p) %>%
-                dplyr::rename(incidence = p),
-              by = 'date') %>%
+  # Add malaria incidence
+  left_join(bes %>%
+              dplyr::select(date, p) %>%
+              dplyr::rename(incidence = p),
+            by = 'date') %>%
   mutate(season = ifelse(incidence >= median(incidence),
                          'high',
                          'low')) %>%
-  mutate(season = factor(season, levels = c('low', 'high')))
[…]

### 2017-12-28 inst/rmd/paper.Rmd
multiple models

@@
 opts_knit$set(width=75)
+options(xtable.comment = FALSE)
+
 ```
@@
 Since our program cost data is not as reliable as our worker characteristic and outcome data, we were conservative in our estimates, and generally tried to err on the side of program activities and materials costing _more_ than what was reported, when doubt was aired.
-Cost data consisted of three types: (i) wages of malaria control employees, (ii) transporation and vehicle costs for IRS teams, and (iii) acquisition costs of purchasing IRS chemicals for fumigation (ACT and DDT).
+Cost data consisted of three types: (i) wages of malaria control employees, (ii) transporation and vehicle costs for IRS teams, and (iii) acquisition costs of purchasing IRS chemicals for fumigation (ACT and DDT), the latter two being combined into malaria control "programme" costs.
 We define the IRS program as "profitable" if return on investment is greater than 100%, ie if the savings associated with the estimated effect of IRS is greater than the costs of the program's administration.
 
+Our formula for return on investment can be described in a straightforward fashion...
+
+\begin{center}
+$R = \dfrac{P_{w} - S_{wa} - S_{wc}}{P_{w}}$
+
+\end{center}
+
+...where $R$ is the return on investment, $P$ is the malaria control program's total operating cost, $w$ refers to costs at the per-worker level, $a$ refers to savings through avoided absences, and $ c $ refers to savings through avoided clinical encounters.
+
 All data processing and analysis were carried out in R [@R] and all analysis code is freely available online [@brewgit].
-Ethical approval for this project was obtained from the Institutional Ethics Review Board for Health at CISM prior to data collection.
+Ethical approval for this project was obtained from the Institutional Ethics Review Board for Health at the CISM prior to data collection.
 
@@
 
-In Southern Mozambique, malaria peaks during the summer months (December through March) most years (Figure 1, panel A), and worker absenteeism rates track malaria incidence closely (Figure 1, panel B).
-Both all-cause absenteeism and sick absenteeism have declined in recent years (Figure 1, panel C), with the latter declining at a faster rate than the former.
-The fact that the rate of confirmed cases at the company clinic is largely non-seasonal (Figure 1, panels E and F) suggests that a significant portion of workers either seek care for malaria elsewhere (for example, government health posts, of which several are nearby and in some many cases closer to workers' residence than the company clinic) or did not seek care during malaria infection.
-Accordingly, we focus our analysis on all-cause absenteeism rather than sickness absenteeism or malaria diagnostics, assuming that much of illness is captured by absenteeism but not by the clinic.
-
-```{r, results='asis'}
-calculate <- function(var){
-  out <- 
-  model_data %>%
-    mutate(year = format(date, '%Y')) %>%
-    mutate(maragra_fabrica = ifelse(maragra_fabrica,
-                                    'On site',
-                                    'Off site'),
-           rainy = ifelse(rainy,
-                          'Rainy',
-                          'Dry')) %>%
-    filter(!is.na(rainy)) %>%
-  group_by_(var, 'year') %>%
-  summarise(absences = length(which(absent)),
-            eligibles = length(absent),
-            workers = length(unique(oracle_number))) %>%
-  ungroup %>%
-  mutate(absenteeism_rate = absences / eligibles * 100) 
-  names(out)[1] <- 'variable'
-  out$val <- paste0(round(out$absenteeism_rate, digits = 2),
-                    '%')#,
-                    # ')',
-                    # '\n',
-                    # ' (workers:', out$workers,
-                    # ')')
-  out <- 
-    out %>%
-    dplyr::select(variable, year, val)
-
-  out <- out %>%
-    spread(key = year,
-           value = val)
-  out$variable <- as.character(out$variable)
-  new_var <- ifelse(var == 'maragra_fabrica',
-                    'Residence',
-                    ifelse(var == 'rainy',
-                           'Precipitation',
-                           var))
-  out$variable[1] <- paste0(Hmisc::capitalize(new_var), ': ', out$variable[1])
-  return(out)
-}
-out_list <- list()
-vars <- c('season', 
-           'department',
-           'sex',
-           'maragra_fabrica',
-           'rainy')
-for(i in 1:length(vars)){
-  x <- calculate(var = vars[i])
-  out_list[[i]] <- x
-}
-ab_table <- bind_rows(out_list)
-names(ab_table)[1] <- ''
-print(xtable::xtable(ab_table,
-                     align = c('r' ,'r', rep('l', 4))),
-      include.rownames = FALSE)
-```
+In Southern Mozambique, malaria peaks during the summer months (December through March) most years (Figure 1, panel A), and worker absenteeism rates track malaria incidence closely, following the same seasonal patterns (Figure 1, panel B).
+Both all-cause absenteeism and sick absenteeism have declined in recent years at Maragra (Figure 1, panel C), with the latter declining at a faster rate than the former.
+The fact that the rate of confirmed cases at the company clinic is largely non-seasonal (Figure 1, panels E and F) suggests that a significant portion of workers either seek care for malaria elsewhere (for example, government health posts, of which several are nearby and in some many cases closer to workers' residence than the company clinic) or do not seek care during malaria infection.
+Accordingly, we focus our analysis on all-cause absenteeism rather than sickness absenteeism or malaria diagnostics, with the assumption that much of illness is captured by absenteeism but not by the clinical data.
 
@@
 people_observed <- length(unique(model_data$oracle_number))
+pr <- function(x){prettyNum(x, big.mark = ',')}
 ```
 
-During the period from January 1st, 2012 through December 31st, 2016, the Maragra Malaria Control Unit carried out `r n_fumigations` episodes of fumigation of residential "agregados" (household combinations), for a total of `r n_buildings` building-fumigation combinations.
-The total number of unique agregados sprayed during this period was `r unique_agregados`.
-Among the `r people_observed` on whom we have absenteeism data, `r people_sprayed` had their homes fumigated at least once (the majority of workers live off of the facility).
+**Fumigations**: During the period from January 1st, 2012 through December 31st, 2016, the Maragra Malaria Control Unit carried out `r pr(n_fumigations)` episodes of fumigation of residential "agregados" (household combinations), for a total of `r pr(n_buildings)` building-fumigation combinations.
+The total number of unique agregados sprayed during this period was `r pr(unique_agregados)`.
+Among the `r pr(people_observed)` workers for whom we have reliable absenteeism and residential data, `r pr(people_sprayed)` had their homes fumigated at least once (the majority of workers live off of the facility).
 
-```{r, fig.height = 3.5, fig.width = 6, fig.cap = 'Effect of IRS on absenteeism by season'}
+**Absences**: We observed `r pr(nrow(ab_panel))` unique worker-days among the `r pr(people_observed)` workers.
+The all-period average absenteeism rate was `r round(100 * sum(ab_panel$absent) / nrow(ab_panel), digits = 2)`%, though this rate varied widely as a function of worker department, sex, residence, and season (table 1).
+
+```{r, results='asis'}
+calculate <- function(var){
+  out <- 
+  model_data %>%
+    mutate(year = format(date, '%Y')) %>%
+    mutate(maragra_fabrica = ifelse(maragra_fabrica,
+                                    'On site',
+                                    'Off site'),
+           rainy = ifelse(rainy,
+                          'Rainy',
+                          'Dry')) %>%
+    filter(!is.na(rainy)) %>%
+  group_by_(var, 'year') %>%
+  summarise(absences = length(which(absent)),
+            eligibles = length(absent),
+            workers = length(unique(oracle_number))) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100) 
+  names(out)[1] <- 'variable'
+  out$val <- paste0(round(out$absenteeism_rate, digits = 2),
+                    '%')#,
+                    # ')',
+                    # '\n',
+                    # ' (workers:', out$workers,
+                    # ')')
+  out <- 
+    out %>%
+    dplyr::select(variable, year, val)
+
+  out <- out %>%
+    spread(key = year,
+           value = val)
+  out$variable <- as.character(out$variable)
+  new_var <- ifelse(var == 'maragra_fabrica',
+                    'Residence',
+                    ifelse(var == 'rainy',
+                           'Precipitation',
+                           var))
+  out$variable[1] <- paste0(Hmisc::capitalize(new_var), ': ', out$variable[1])
+  return(out)
+}
+out_list <- list()
+vars <- c('season', 
+           'department',
+           'sex',
+           'maragra_fabrica',
+           'rainy')
+for(i in 1:length(vars)){
+  x <- calculate(var = vars[i])
+  out_list[[i]] <- x
+}
+ab_table <- bind_rows(out_list)
+names(ab_table)[1] <- ''
+print(xtable::xtable(ab_table,
+                     caption = 'Absenteeism rate by year and worker characteristics',
+                     align = c('r' ,'r', rep('l', 4))),
+      include.rownames = FALSE)
+```
+
+**Costs**: The malaria control program at Maragra has an annual operating budget of approximately XX, which includes the purchase of insecticide, the wages of IRS sprayers and drivers, transportation, record-keeping, and general administrative costs.
+Assuming linearity in costs, the program spends approximately XX per agregado sprayed.
+With each agregado containing an average of 2.2 workers, this translates to a cost of XX per worker protected per season.
+Much of the benefit of IRS goes to non-worker residents of sprayed agregados (who constitute a majority), but this benefit is purposefully ignored for this analysis.
+
+Given the likelihood that clinical data does not fully capture all malaria cases, we do not quantify the costs of malaria infection to the company.
+Rather, we first estimate the reduction in absenteeism attributible to IRS, and then quantify the savings associated with prevented absences.
+Additionally, we calculate the clinical savings of IRS by first estimating the share of absences which are associated with an episode of clinical malaria, and then applying the clinical cost per case to the equivalent share of prevented absences.
+We intentionally ignore the savings accrued by the public health system, as well as the likely utility gains in secondary realms such as school absenteeism, producitivity, etc.
+
+**Effect of IRS on absenteeism**: IRS is associated with a year-long, significant reduction in absenteeism during the malaria season (figure 2).
+As one would expect if the mechanism by which IRS reduces absence is through reduced malaria infection, the effect of IRS during the low transmission season is significant, but far less substantial in effect size.
+
+```{r, fig.height = 3.5, fig.width = 6, fig.cap = 'Estimated effect of IRS on absenteeism by season'}
 # Predict for visualization
@@
               color = 'black',
-              alpha = 0.5,
-              span = 1) +
+              # span = 10,
+              alpha = 0.5) +
   facet_wrap(~season) +
@@
        y = 'Absenteeism (%)')
+```
+
+Details on full model here. Bla bla bla. See table 3.
+
+```{r, results = 'asis'}
+clean_up_model <- function(x){
+  x = broom::tidy(x)
+  for(j in 2:(ncol(x) -1)){
+    x[,j] <- x[,j] * 100
+  }
+  x$term <- gsub('seasonhigh', 'High transmission season', x$term)
+  x$term <- gsub('months_since', 'Months since IRS: ', x$term)
+  x$term <- gsub('department', 'Department: ', x$term)
+  x$term <- gsub('precipitation', 'Precipitation (mm) ', x$term)
+  x$term <- gsub('maragra_fabricaTRUE', 'Living at factory', x$term)
+  x$term <- gsub('sexM', 'Male', x$term)
+  x$term <- gsub('permanent_or_temporaryTemporary', 'Temp contract', x$term)
+  x$term <- gsub('rainyTRUE', 'Rainy day', x$term)
+  names(x) <- Hmisc::capitalize(names(x))
+  names(x) <- gsub('.', ' ', names(x), fixed = TRUE)
+  x <- x %>% 
+    mutate(Estimate = paste0(round(Estimate, 3), ' (P = ',
+                             ifelse(`P value` <= 0.001,
+                                    '< 0.001',
+                                    round(`P value`, 4)), ')')) %>%
+    dplyr::select(Term, Estimate)
+  return(x)
+}
+# Combine them all
+out_list <- list()
+for (i in 1:length(models)){
+  this_model <- letters[i]
+  the_model <- models[[this_model]]
+  the_model <- clean_up_model(the_model)
+  names(the_model)[2] <- paste0('Model ', i)
+  out_list[[i]] <- the_model
+}
+df <- out_list[[length(models)]]
+for(i in (length(models) - 1):1){
+  message(i)
+    df <- left_join(df, out_list[[i]])
+}
+df <- df[,unique(c('Term', sort(names(df))))]
+print(xtable::xtable(x, caption = ))
 

### 2017-12-28 inst/rmd/paper_prepare.R

@@ -80,13 +80,34 @@ model_data <-
                                             0,
                                             precipitation))) %>%
   mutate(rainy = precipitation > 0)
-
+  
+# Aggregate months since 
+model_data <- model_data %>%
+  mutate(months_since = as.character(months_since)) %>%
+  mutate(months_since = ifelse(months_since %in% c('Never', 'Before'),
+                               months_since,
+                               ifelse(months_since %in% c('00', '01'), '01',
+                                      ifelse(months_since %in% c('02', '03', '04'),
+                                             '02-04',
+                                             ifelse(months_since %in% c('05', '06', '07', '08', '09'),
+                                                    '05-09',
+                                                           ifelse(months_since %in% c('10', '11', '12+'), '10+', months_since)))))) %>%
+  mutate(months_since = factor(months_since, levels = unique(c('Never', 'Before', sort(unique(months_since)))))) 
+                               
 # Model simple
 fit <- lm(absent ~ season * months_since, data = model_data)
-summary(fit)
+# summary(fit)
 
+# Complex model
+fit_full <- lm(absent ~ season * months_since + sex + department + precipitation, data = model_data)
 
-
+models <- list(
+  a = lm(absent ~ season * months_since, data = model_data),
+  b = lm(absent ~ season * months_since + sex, data = model_data),
+  c = lm(absent ~ season * months_since + sex + department, data = model_data),
+  d = lm(absent ~ season * months_since + sex + department + rainy, data = model_data),
+  e = lm(absent ~ season * months_since + sex + department + rainy + permanent_or_temporary, data = model_data)
+)
 
 
 

### 2017-12-28 inst/rmd/paper.Rmd
update

@@
   }
-  x$term <- gsub('seasonhigh', 'High transmission season', x$term)
+  x$term <- gsub('seasonhigh', 'Malaria season', x$term)
   x$term <- gsub('months_since', 'Months since IRS: ', x$term)
@@
 df <- df[,unique(c('Term', sort(names(df))))]
-print(xtable::xtable(x, caption = ))
+print(xtable::xtable(df, caption = 'Regression outputs'), size="\\fontsize{5pt}{6pt}\\selectfont")
 

### 2017-12-28 inst/rmd/paper_prepare.R

@@ -105,8 +105,8 @@ models <- list(
   a = lm(absent ~ season * months_since, data = model_data),
   b = lm(absent ~ season * months_since + sex, data = model_data),
   c = lm(absent ~ season * months_since + sex + department, data = model_data),
-  d = lm(absent ~ season * months_since + sex + department + rainy, data = model_data),
-  e = lm(absent ~ season * months_since + sex + department + rainy + permanent_or_temporary, data = model_data)
+  # d = lm(absent ~ season * months_since + sex + department + rainy, data = model_data),
+  d = lm(absent ~ season * months_since + sex + department + rainy + permanent_or_temporary, data = model_data)
 )
 
 

### 2017-12-30 inst/rmd/paper_prepare.R
major update, fixed effects models

@@ -33,13 +33,15 @@ add_zero <- function (x, n) {
   return(x)
 }
 
+# filter for on-site only
+
 model_data <-
   ab_panel %>%
   left_join(irs, by = c('unidade', 'date')) %>%
   mutate(days_since = days_since %/% 30) %>%
   mutate(days_since = ifelse(is.na(days_since), 'Never',
                              ifelse(days_since < 0, 'Before',
-                             ifelse(days_since >= 12, '12+', as.character(add_zero(days_since, n = 2)))))) %>%
+                                    ifelse(days_since >= 12, '12+', as.character(add_zero(days_since, n = 2)))))) %>%
   mutate(months_since = factor(days_since, levels = unique(c('Never', 'Before', sort(unique(days_since)))))) %>%
   mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) %>%
   # Add malaria incidence
@@ -80,7 +82,7 @@ model_data <-
                                             0,
                                             precipitation))) %>%
   mutate(rainy = precipitation > 0)
-  
+
 # Aggregate months since 
 model_data <- model_data %>%
   mutate(months_since = as.character(months_since)) %>%
@@ -91,24 +93,62 @@ model_data <- model_data %>%
                                              '02-04',
                                              ifelse(months_since %in% c('05', '06', '07', '08', '09'),
                                                     '05-09',
-                                                           ifelse(months_since %in% c('10', '11', '12+'), '10+', months_since)))))) %>%
-  mutate(months_since = factor(months_since, levels = unique(c('Never', 'Before', sort(unique(months_since)))))) 
-                               
+                                                    ifelse(months_since %in% c('10', '11', '12+'), '10+', months_since)))))) %>%
+# Define whether every sprayed
+  ungroup %>%
+  group_by(oracle_number) %>%
+  mutate(ever_sprayed = length(which(unidade %in% mc$unidade)) > 0) %>%
+  ungroup %>%
+  # Correct the nevers
+  mutate(days_since = ifelse(days_since == 'Never' & ever_sprayed, 'Before', days_since)) %>%
+  mutate(months_since = as.character(months_since)) %>%
+  mutate(months_since = ifelse(months_since == 'Never' & ever_sprayed, 'Before', months_since)) %>%
+  # Make never and before the same
+  mutate(months_since = ifelse(months_since %in% c('Never', 'Before'), 'No IRS', months_since)) %>%
[…]

### 2018-01-03 inst/rmd/00_title_page.Rmd
major update

\begin{center}
\begin{large}

```{r, results = 'asis', eval = TRUE}
make_authors()
```

\end{large}
\end{center}

\vspace{5mm}

\begin{center}
\textbf{Abstract}  
\end{center}

\vspace{5mm}
\begin{center}
\begin{changemargin}{3cm}{3cm}

This paper provides new empirical evidence regarding the return on investment of privately managed malaria control activities (indoor residual spraying with pesticides) on worker absenteeism in Mozambique. We analyze 4 years of malaria control and worker health and absenteeism data from a large sugar processing facility in Mozambique. We find that the benefits outweight the costs (ie, there is a positive return on investment) even when the consideration of benefits is limited to those directly accrued by the company. These findings suggest that the private sector may have an important role to play in malaria control in endemic areas.

\end{changemargin}
\end{center}

\vspace{20mm}

\noindent\fbox{%
    \parbox{\textwidth}{%
        \subsection*{Research Highlights}
        \begin{itemize}
          \item Large, individual-level worker absenteeism data from malaria endemic zone
          \item Quantifies effect of indoor residual spraying on absenteeism
          \item Estimates cost-effectiveness of malaria control from investment standpoint
          \item Results suggest that private sector could play a significant role in malaria elimination
        \end{itemize}
        \vspace{2mm}
    }%
}

\vfill
\null

\subsection*{Keywords}
\textbf{Malaria; Investment; Health; Productivity; Agriculture; Absenteeism}

\vspace{3mm}

\newpage

### 2018-01-03 inst/rmd/01_introduction.Rmd
major update

# Introduction

Malaria has a large economic impact on endemic societies. By affecting saving, investment [@Shretta2016], risk perception, productivity, absenteeism [@Nonvignon_2016], human capital accumulation [@CastelBranco2014], mortality, and costs of care [@Sachs2002], malaria likely has a negative effect on GDP and growth [@McCarthy_2000] [@Orem_2012]. Because of the relative affordability of most intereventions and the enormous societal costs of malaria, most forms of malaria control are cost-effective when a public welfare perspective is assumed, such as when a government provides the financing [@White_2011] [@Purdy_2013] [@Howard_2017].

From the persepctive of the private sector, however, investing in malaria control is not so clear-cut. Public health interventions targetting malaria - and their corresponding cost-effectiveness evaluations - most often focus on impacts pertaining to public welfare, such as an increase in life years adjusted for disability or quality  [@Goodman1999] [@Shretta2016] [@Lee2017] [@Hanson2004]. Though population-level health is certainly of importance to businesses, and improvements in health incidentally improve the economy at all levels [@World1999] [@Bloom2008] [@Vecchi_2013], these improvements may be too disperse or long-term to incentivize private sector involvement in health campaigns.

100% of the Mozambican population are at risk of malaria, living in what the WHO classifies as a "high transmission" area [@Moonasar_2016]. Annually, Mozambique has more than 8 million clinical malaria cases (an annual incidence of approximately 300 per 1,000 residents), with an estimated 14,000 deaths. Malaria accounts for 29% of all deaths, and 42% of deaths among those under five years of age [@estatistica2009]. Since 2013, Mozambique has seen a gradual increase in the incidence of malaria [@Moonasar_2016]. 100% of the malaria in Mozambique is of the Plasmodium falciparum species, with Anopheles funestus, gambiae, and arabiensis as the primary mosquito vectors of the disease [@whoprof].

A significant sector of the economy in Mozambique is dominated by a full large-scale foreing direct investment projects [@Robbins2012], and the role of the private sector in health generally, and malaria specifically, is unequivocally important. Large agriculture and extractive industry firms take up wide swaths of land and employ hundreds of thousands  [@German2013]. The Mozambican state has encouraged large-scale entreprise with the aim of general economic development [@Buur2012]. And where large firms exist, they often take on social roles such as housing and health care [@Winkler]. At times, this role is necessary from a purely practical standpoint; in other cases, it is employed under the guise of "corporate social responsibility" [@Azemar2009]. Regardless of the language used, it is clear that private industry plays an important role in public health in Mozambique [@Robbins2012] [@CastelBranco2014].

Several cases exist of foreign firms engaging in large-scale malaria control campaigns [@Mouzin2011] [@Han] [@Bennett_2017] [@Kaula_2017]. But these studies generally consider population health as the outcome measure of interest, rather than worker absenteeism or productivity. Similarly, they often neglect to differentiate between those clinical costs which are absorbed by the local health system versus those which are absorbed by the firm itself. In the literature, making the "investment case" for malaria control or elimination generally implies that the investor is the public sector, and takes into account those costs and benefits which are applicable from a public welfare point of view [@Shretta_2017]; though appropriate in most cases (the government or institutions interested in public welfare primarily being the primary malaria control agents in most locations), the findings of these studies are rarely applicable to the private sector. In the case of a private firm not interested in "corporate social responsibility", it is not clear whether investing in malaria control would be profitable or not. This lack of clarity not only discourages investment, but also makes it difficult for governments to pinpoint the correct of amount of subsidy (if applicable) to encourage private sector scale-up in malaria control.

To address the question of the profitability of malaria control activities from the standpoint of a private firm, we analyze data during a 7 year period from a private sugar facility in Southern Mozambique. We assess the effect of indoor residual spraying (IRS) on the absenteeism and health of workers, and demonstrate that the firm's engagement in malaria control not only improved worker health, but also generated a positive return on investment.

The structure of the paper is as follows. After a brief review of the existing literature on the subject of private sector investment in malaria control and its effects on worker productivity, we provide an overview of the sugar company under study, and the epidemiology of malaria in the nearby area, as well as in Mozambique as a whole. We then give an overview of the data collected, and outline the theoretical and methdological assumptions that underly our analysis. In the results, we show the effectiveness of the company's malaria control program in reducing absences, and translate this reduction into cost savings. Our robustness checks consist of \todo{Add some stuff here}. Our discussion covers potential implications from this study in terms of policy and investment, as well as the paper's limitations.

## Background Literature

\todo{Will add 3-5 paragraphs of background lit review}

Our study adds to the existing literature in several ways. \todo{Will add details here.}

## Study area

Details on Maragra:

Maragra Açucar SA (henceforth referred to as "Maragra")\todo{This is a comment that will appear in the margin}

Details on Manhica

Details on Mozambique

### 2018-01-03 inst/rmd/02_methods.Rmd
major update

--- 2017-12-02 inst/rmd/paper.Rmd
@@
----
-title: Effectiveness of private sector malaria control at a large sugar
-  facility in Southern Mozambique
-date: ''
-output: 
-  pdf_document:
-    includes:
-      in_header: preamble.sty
-fig_width: 4
-fig_height: 2.6
-bibliography: bibliography.bib
----
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
-# Packages
-library(tidyverse)
-library(knitr)
-library(Hmisc)
-library(brew)
-library(maragra)
-library(knitr)
-library(RColorBrewer)
-library(extrafont)
-loadfonts()
-## Global options
-options(max.print="75")
-opts_chunk$set(echo=FALSE,
-	             cache=FALSE,
-               prompt=FALSE,
-               tidy=TRUE,
-               comment=NA,
-               message=FALSE,
-               warning=FALSE,
-               dev = "cairo_pdf")
-opts_knit$set(width=75)
-```
-
-```{r}
-source('paper_prepare.R')
-```
-
-\begin{center}
-\begin{large}
-
-```{r, results = 'asis', eval = TRUE}
-make_authors()
-```
-
-\end{large}
-\end{center}
-
-\vspace{5mm}
-
-\begin{center}
-\textbf{Abstract}  
-\end{center}
-
-\vspace{5mm}
-\begin{center}
-\begin{changemargin}{3cm}{3cm}
-
-\textbf{Context:} Incentivizing scaled-up private firms' involvement in malaria control and elimination activities requires a clear and robust demonstration of its benefits, and a quantification of which benefits can be expected to return to the firm.
-However, most studies to date examine benefits of malaria control purely from the perspective of the public sector.
-\textbf{Methods:} We analyze 4 years of malaria control and worker health and absenteeism data from a large sugar processing facility in Mozambique.
-We estimate the effect of indoor residual spraying (IRS) on worker health and absenteeism.
-\textbf{Results:} We demonstrate that the firm's engagement in malaria control activities is profitable, even if we only consider those benefits which directly affect the firm's bottom line and ignore secondary and societal benefits.
-
-\end{changemargin}
-\end{center}
-
-# Introduction
-
-Malaria has a nearly unquantifiably large economic impact on endemic societies.
-By affecting saving, investment [@Shretta2016], risk perception, productivity, absenteeism [@Nonvignon_2016], human capital accumulation [@CastelBranco2014], mortality, and costs of care [@Sachs2002], malaria likely has a negative effect on GDP and growth [@McCarthy_2000] [@Orem_2012].
-Because of the relative affordability of most intereventions and the enormous societal costs of malaria, most forms of malaria control are cost-effective from the perspective of the government [@White_2011] [@Purdy_2013] [@Howard_2017].
-
-From the persepctive of the private sector, however, investing in malaria control is not such a clear-cut case.
-Public health interventions targetting malaria - and their corresponding cost-effectiveness evaluations - most often pertain to the public sector.
-Accordingly, the analytical framework and metrics employed by these analyses most often focus on impact pertaining to the public good, such as an increase in life years adjusted for disability or quality  [@Goodman1999] [@Shretta2016] [@Lee2017] [@Hanson2004].
-Though population-level health is certainly of importance to businesses, and improvements in health incidentally improve the economy at all levels [@World1999] [@Bloom2008] [@Vecchi_2013], these improvements may be too disperse or long-term to incentivize private sector involvement in health campaigns.
-
-In the context of Mozambique, where a significant sector of the economy is dominated by a full large-scale foreing direct investment projects [@Robbins2012], the role of the private sector in health generally, and malaria specifically, is unequivocally important.
-Large agriculture and extractive industry firms take up wide swaths of land and employ hundreds of thousands  [@German2013].
-The Mozambican state has encouraged large-scale entreprise with the aim of general economic development [@Buur2012].
-And where large firms exist, they often take on social roles such as housing and health care [@Winkler].
-At times, this role is necessary from a purely practical standpoint; in other cases, it is employed under the guise of "corporate social responsibility" [@Azemar2009].
-Regardless of the language used, it is clear that private industry plays an important role in public health in Mozambique [@Robbins2012] [@CastelBranco2014].
-
-Several cases exist of foreign firms engaging in large-scale malaria control campaigns [@Mouzin2011] [@Han] [@Bennett_2017] [@Kaula_2017].
-But these studies generally consider population health as the outcome measure of interest, rather than worker absenteeism or productivity.
-Similarly, they often neglect to differentiate between those clinical costs which are absorbed by the local health system versus those which are absorbed by the firm itself.
-In the literature, making the "investment case" for malaria control or elimination generally implies that the investor is the public sector, and takes into account those costs and benefits which are applicable to the public sector [@Shretta_2017]; though appropriate in most cases (the public sector being the primary malaria control agent in most locations), the findings of these studies are rarely applicable to the private sector.
-In the case of a private firm not interested in "corporate social responsibility", it is not clear whether investing in malaria control would be profitable or not.
-This lack of certainty most likely discourages investment in many cases.
-
-To address the question of the profitability of malaria control activities from the standpoint of a private firm, we analyze data during a 7 year period from a private sugar facility in Southern Mozambique.
-We assess the effect of indoor residual spraying (IRS) on the absenteeism and health of workers, and demonstrate that the firm's engagement in malaria control not only improved worker health, but also generated a positive return on investment.
-
 # Methods
 
-In collaboration with the sugar processing facility, Maragra Açucar SA (henceforth referred to as "Maragra"), we collected data for the period from January 2010 through December 2016.
+## Data collection
+
+In collaboration with the sugar processing facility, we collected data for the period from January 2010 through December 2016.
 Data came from four sources: (i) the Human Resources' roster of worker details and absences, (ii) the facility's on-site clinic's medical and laboratory records, (iii) the facility's on-site malaria control program's records pertaining to the dates, chemicals, and location of IRS activities, and (iv) interviews with company employees pertaining to costs, data limitations, etc.
@@
 Maragra regularly employs IRS at on-site worker households in order to reduce those workers' (and their families') risk of malaria infection.
-By using the CISM's census, we were able to identify from the record of houses sprayed (maintained by the company) who lived in those houses, as well as who lived off-site (and therefore did not receive IRS from Maragra).
-Then, using company HR and clinical records, we were able to identify absences and episodes of clinical malaria among all workers, as well as identify the time since the most recent IRS episode before the onsent of absence or illness.
+Workers living off-site (our control group) also may have received IRS at some point during the study period (from government programs).
+Even though we do not have reliable person-level data on IRS carried out by the government, off-site workers are a suitable control in the sense that they represent "business as usual" (ie, what would happen if the company carried out no IRS and relied solely on public interventions).
+Using company HR and clinical records, we were able to identify absences and episodes of clinical malaria among all workers, as well as identify the time since the most recent IRS episode before the onsent of absence or illness.
 
 Worker characteristics, illness and absenteeism data, along with IRS activity data, were systematically stored, collected, and used at the individual level by Maragra, and therefore of generally high quality.
+Because cost data was less systematically collected by Maragra, and because many costs could not be precisely quantified due to the abundance of in-kind and cross-departmental expenditures, we had to rely on rough estimations based on a mix of interviews, reciepts, and interpolations.
+Since our program cost data is not as reliable as our worker characteristic and outcome data, we were conservative in our estimates, and generally tried to err on the side of program activities and materials costing _more_ than what was reported, when doubt was aired.
+Cost data consisted of three types: (i) wages of malaria control employees, (ii) transporation and vehicle costs for IRS teams, and (iii) acquisition costs of purchasing IRS chemicals for fumigation (ACT and DDT), the latter two being combined into malaria control "programme" costs.
+
+## Conceptual framework and identification strategy
+
 We sought to understand the effect of IRS on individual workers' likelihood of absence from work as well as their likelihood of clinical malaria.
@@
 
-Because cost data was less systematically collected by Maragra, and because many costs could not be precisely quantified due to the abundance of in-kind and cross-departmental expenditures, we had to rely on rough estimations based on a mix of interviews, reciepts, and interpolations.
-Since our program cost data is not as reliable as our worker characteristic and outcome data, we were conservative in our estimates, and generally tried to err on the side of program activities and materials costing _more_ than what was reported, when doubt was aired.
-Cost data consisted of three types: (i) wages of malaria control employees, (ii) transporation and vehicle costs for IRS teams, and (iii) acquisition costs of purchasing IRS chemicals for fumigation (ACT and DDT).
-We define the IRS program as "profitable" if return on investment is greater than 100%, ie if the savings associated with the estimated effect of IRS is greater than the costs of the program's administration.
+## Estimating return on investment
+
+Our formula for return on investment can be described in a straightforward fashion...
+
+\begin{center}
+$R = \dfrac{P_{w} - S_{wa} - S_{wc}}{P_{w}}$
+
+\end{center}
+
+...where $R$ is the return on investment, $P$ is the malaria control program's total operating cost, $w$ refers to costs at the per-worker level, $a$ refers to savings through avoided absences, and $ c $ refers to savings through avoided clinical encounters.
+We define the malaria control program as "profitable" from an investment standpoint if ROI is greater than 100%, ie if the savings associated with the estimated effect of IRS is greater than the costs of the program's administration.
+
+## Reproducibility and ethical approval
 
 All data processing and analysis were carried out in R [@R] and all analysis code is freely available online [@brewgit].
-Ethical approval for this project was obtained from the Institutional Ethics Review Board for Health at CISM prior to data collection.
+Ethical approval for this project was obtained from the Institutional Ethics Review Board for Health at the CISM prior to data collection.
 
-# Results
+## Descriptive statistics
 
-In Southern Mozambique, malaria peaks during the summer months (December through March) most years (Figure 1, panel A), and worker absenteeism rates track malaria incidence closely (Figure 1, panel B).
-Both all-cause absenteeism and sick absenteeism have declined in recent years (Figure 1, panel C), with the latter declining at a faster rate than the former.
-The fact that the rate of confirmed cases at the company clinic is largely non-seasonal (Figure 1, panels E and F) suggests that a significant portion of workers either seek care for malaria elsewhere (for example, government health posts, of which several are nearby and in some many cases closer to workers' residence than the company clinic) or did not seek care during malaria infection.
-Accordingly, we focus our analysis on all-cause absenteeism rather than sickness absenteeism or malaria diagnostics, assuming that much of illness is captured by absenteeism but not by the clinic.
-
-```{r, results='asis'}
-calculate <- function(var){
-  out <- 
-  model_data %>%
-    mutate(year = format(date, '%Y')) %>%
-    mutate(maragra_fabrica = ifelse(maragra_fabrica,
-                                    'On site',
-                                    'Off site'),
-           rainy = ifelse(rainy,
-                          'Rainy',
-                          'Dry')) %>%
-    filter(!is.na(rainy)) %>%
-  group_by_(var, 'year') %>%
-  summarise(absences = length(which(absent)),
-            eligibles = length(absent),
-            workers = length(unique(oracle_number))) %>%
-  ungroup %>%
-  mutate(absenteeism_rate = absences / eligibles * 100) 
-  names(out)[1] <- 'variable'
-  out$val <- paste0(round(out$absenteeism_rate, digits = 2),
-                    '%')#,
-                    # ')',
-                    # '\n',
-                    # ' (workers:', out$workers,
-                    # ')')
-  out <- 
-    out %>%
-    dplyr::select(variable, year, val)
-
-  out <- out %>%
-    spread(key = year,
-           value = val)
-  out$variable <- as.character(out$variable)
-  new_var <- ifelse(var == 'maragra_fabrica',
-                    'Residence',
-                    ifelse(var == 'rainy',
-                           'Precipitation',
-                           var))
-  out$variable[1] <- paste0(Hmisc::capitalize(new_var), ': ', out$variable[1])
-  return(out)
-}
-out_list <- list()
-vars <- c('season', 
-           'department',
-           'sex',
-           'maragra_fabrica',
-           'rainy')
-for(i in 1:length(vars)){
-  x <- calculate(var = vars[i])
-  out_list[[i]] <- x
-}
-ab_table <- bind_rows(out_list)
-names(ab_table)[1] <- ''
-print(xtable::xtable(ab_table,
-                     align = c('r' ,'r', rep('l', 4))),
-      include.rownames = FALSE)
-```
-
-```{r, fig.cap = 'Clinical malaria (district of Manhiça), all-cause absenteeism among Maragra workers, sick absenteeism among Maragra workers, estimated rainfall, positive cases at company clinic, and test positivity rate at company clinic', fig.height=5, fig.width=8}
-# BES incidence
-g1 <- 
-  ggplot(data = bes %>%
-           mutate(season = ifelse(p >= median(p), 'High season',
-                                  'Low season')) %>%
-           filter(weekdays(date) == 'Monday'),
-       aes(x = date,
-           y = p)) +
-  geom_point(alpha = 0.6,
-             aes(color = season),
-             size = 0.2) +
-  geom_line(alpha = 0.1) +
-  theme_maragra(base_size = 9) +
-  labs(x = 'Date',
-       y = 'Incidence',
-       title = 'A: Annualized district clinical\nmalaria incidence per 1,000') +
-  scale_color_manual(name = '',
-                     values = c('red', 'black')) + 
-  guides(color = guide_legend(ncol = 1)) +
-  theme(legend.position = c(.8, .8),
-        legend.text = element_text(size = 5),
-        legend.background = element_rect(fill=alpha('white', 0)))
-
-# Absenteeism rate
-g2_data <- ab_panel %>%
-  group_by(month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
-  summarise(absences = length(which(absent)),
-            absences_sick = length(which(absent_sick)),
-
-            eligibles = n()) %>%
-  ungroup %>%
-  mutate(absenteeism_rate = absences / eligibles * 100,
-         sick_absenteeism_rate = absences_sick / eligibles * 100)
-
-g2 <- ggplot(data = g2_data,
-             aes(x = month,
-                 y = absenteeism_rate)) +
-  geom_line(alpha = 0.6) +
-  theme_maragra(base_size = 9) +
-  labs(x = 'Date',
-       y = 'Absenteeism rate',
-       title = 'B: Monthly absenteeism rate')
-
-g3 <- ggplot(data = g2_data,
-             aes(x = month,
-                 y = sick_absenteeism_rate)) +
-  geom_line(alpha = 0.6) +
-  theme_maragra(base_size = 9) +
-  labs(x = 'Date',
-       y = 'Absenteeism rate',
-       title = 'C: Monthly sick absenteeism rate')
-
-# Weather
-g4 <-
-  ggplot(data = weather %>%
-           group_by(date = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
-           summarise(precipitation = sum(precipitation, na.rm = TRUE)),
-         aes(x = date,
-             y = precipitation)) +
-  geom_line(alpha = 0.6) +
-  theme_maragra(base_size = 9) +
-  labs(x = 'Date',
-       y = 'Mililiters',
-       title = 'D: Monthly precipitation')
-
-# Clinical incidence of malaria
-g5 <- clinic_agg %>%
-  filter(date >= '2013-01-01', date <= '2016-12-31') %>%
-  # mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
-  group_by(date) %>%
-  summarise(tested = sum(tested),
-            positive = sum(positive)) %>%
-  ungroup %>%
-  mutate(percent_positive = positive / tested * 100) %>%
-  ggplot(aes(x = date,
-             y = positive)) +
-  geom_line(alpha = 0.6) +
-  theme_maragra(base_size = 9) +
-  labs(x = 'Date',
-       y = 'Cases',
-       title = 'E: Monthly malaria cases\nat company clinic')
-
-g6 <- clinic_agg %>%
-  filter(date >= '2013-01-01', date <= '2016-12-31') %>%
-  # mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
-  group_by(date) %>%
-  summarise(tested = sum(tested),
-            positive = sum(positive)) %>%
-  ungroup %>%
-  mutate(percent_positive = positive / tested * 100) %>%
-  ggplot(aes(x = date,
-             y = percent_positive)) +
-  geom_line(alpha = 0.6) +
-  theme_maragra(base_size = 9) +
-  labs(x = 'Date',
-       y = 'Cases',
-       title = 'F: Monthly malaria test positivity rate\nat company clinic') 
-Rmisc::multiplot(g1, g2, g3, g4, g5, g6, layout = matrix(1:6, nrow=2, byrow=TRUE)) 
-```
-
-```{r}
-fumigations <- mc %>%
-  filter(date >= '2012-01-01',
-         date <= '2016-12-31')
-n_fumigations <- nrow(fumigations)
-n_buildings <- sum(fumigations$casas_cobertas, na.rm = TRUE)
-unique_agregados <- length(unique(fumigations$unidade))
-people_sprayed <- length(unique(model_data$oracle_number[model_data$days_since != 'Never']))
-people_observed <- length(unique(model_data$oracle_number))
-```
-
-During the period from January 1st, 2012 through December 31st, 2016, the Maragra Malaria Control Unit carried out `r n_fumigations` episodes of fumigation of residential "agregados" (household combinations), for a total of `r n_buildings` building-fumigation combinations.
-The total number of unique agregados sprayed during this period was `r unique_agregados`.
-Among the `r people_observed` on whom we have absenteeism data, `r people_sprayed` had their homes fumigated at least once (the majority of workers live off of the facility).
-
-```{r, fig.height = 3.5, fig.width = 6, fig.cap = 'Effect of IRS on absenteeism by season'}
-# Predict for visualization
-dummy <- expand.grid(season = sort(unique(model_data$season)),
-                     months_since = sort(unique(model_data$months_since)))
-predictions <- predict(fit, dummy, se.fit = TRUE, interval = 'confidence')
-dummy$predicted <- predictions$fit[,1]
-dummy$lwr <- predictions$fit[,2]
-dummy$upr <- predictions$fit[,3]
-ggplot(data = dummy %>%
-         filter(months_since != 'Never') %>%
-         mutate(lwr = lwr * 100,
-                upr = upr * 100,
-                predicted = predicted * 100) %>%
-         mutate(season = paste0(Hmisc::capitalize(as.character(season)), ' season')),
-       aes(x = months_since,
-           y = predicted,
-           group = 1)) +
-  geom_ribbon(aes(x = months_since,
-                  ymin = lwr,
-                  ymax = upr),
-              fill = 'lightblue',
-              alpha = 0.6) +
-  # geom_line(color = 'blue', alpha = 0.5) +
-  geom_point(color = 'blue', alpha = 0.5) + 
-  geom_smooth(se = FALSE,
-              color = 'black',
-              alpha = 0.5,
-              span = 1) +
-  facet_wrap(~season) +
-  theme_maragra() +
-  theme(axis.text.x = element_text(angle = 90, size = 8)) +
-  labs(x = 'Months since IRS',
-       y = 'Absenteeism (%)')
-
-```
-
-\vfill
-
-\pagebreak
-
-# Discussion
-
-Some text goes here
-
-# References
+Age, types of workers, bla bla bla. \todo{Will add a table here and some overview}

### 2018-01-03 inst/rmd/03_results.Rmd
major update

--- 2017-12-28 inst/rmd/paper.Rmd
@@
----
-title: Effectiveness of private sector malaria control at a large sugar
-  facility in Southern Mozambique
-date: ''
-output: 
-  pdf_document:
-    includes:
-      in_header: preamble.sty
-fig_width: 4
-fig_height: 2.6
-bibliography: bibliography.bib
----
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
-# Packages
-library(tidyverse)
-library(knitr)
-library(Hmisc)
-library(brew)
-library(maragra)
-library(knitr)
-library(RColorBrewer)
-library(extrafont)
-loadfonts()
-## Global options
-options(max.print="75")
-opts_chunk$set(echo=FALSE,
-	             cache=FALSE,
-               prompt=FALSE,
-               tidy=TRUE,
-               comment=NA,
-               message=FALSE,
-               warning=FALSE,
-               dev = "cairo_pdf")
-opts_knit$set(width=75)
-options(xtable.comment = FALSE)
-
-```
-
-```{r}
-source('paper_prepare.R')
-```
-
-\begin{center}
-\begin{large}
-
-```{r, results = 'asis', eval = TRUE}
-make_authors()
-```
-
-\end{large}
-\end{center}
-
-\vspace{5mm}
-
-\begin{center}
-\textbf{Abstract}  
-\end{center}
-
-\vspace{5mm}
-\begin{center}
-\begin{changemargin}{3cm}{3cm}
-
-\textbf{Context:} Incentivizing scaled-up private firms' involvement in malaria control and elimination activities requires a clear and robust demonstration of its benefits, and a quantification of which benefits can be expected to return to the firm.
-However, most studies to date examine benefits of malaria control purely from the perspective of the public sector.
-\textbf{Methods:} We analyze 4 years of malaria control and worker health and absenteeism data from a large sugar processing facility in Mozambique.
-We estimate the effect of indoor residual spraying (IRS) on worker health and absenteeism.
-\textbf{Results:} We demonstrate that the firm's engagement in malaria control activities is profitable, even if we only consider those benefits which directly affect the firm's bottom line and ignore secondary and societal benefits.
-
-\end{changemargin}
-\end{center}
-
-# Introduction
-
-Malaria has a nearly unquantifiably large economic impact on endemic societies.
-By affecting saving, investment [@Shretta2016], risk perception, productivity, absenteeism [@Nonvignon_2016], human capital accumulation [@CastelBranco2014], mortality, and costs of care [@Sachs2002], malaria likely has a negative effect on GDP and growth [@McCarthy_2000] [@Orem_2012].
-Because of the relative affordability of most intereventions and the enormous societal costs of malaria, most forms of malaria control are cost-effective from the perspective of the government [@White_2011] [@Purdy_2013] [@Howard_2017].
-
-From the persepctive of the private sector, however, investing in malaria control is not such a clear-cut case.
-Public health interventions targetting malaria - and their corresponding cost-effectiveness evaluations - most often pertain to the public sector.
-Accordingly, the analytical framework and metrics employed by these analyses most often focus on impact pertaining to the public good, such as an increase in life years adjusted for disability or quality  [@Goodman1999] [@Shretta2016] [@Lee2017] [@Hanson2004].
-Though population-level health is certainly of importance to businesses, and improvements in health incidentally improve the economy at all levels [@World1999] [@Bloom2008] [@Vecchi_2013], these improvements may be too disperse or long-term to incentivize private sector involvement in health campaigns.
-
-In the context of Mozambique, where a significant sector of the economy is dominated by a full large-scale foreing direct investment projects [@Robbins2012], the role of the private sector in health generally, and malaria specifically, is unequivocally important.
-Large agriculture and extractive industry firms take up wide swaths of land and employ hundreds of thousands  [@German2013].
-The Mozambican state has encouraged large-scale entreprise with the aim of general economic development [@Buur2012].
-And where large firms exist, they often take on social roles such as housing and health care [@Winkler].
-At times, this role is necessary from a purely practical standpoint; in other cases, it is employed under the guise of "corporate social responsibility" [@Azemar2009].
-Regardless of the language used, it is clear that private industry plays an important role in public health in Mozambique [@Robbins2012] [@CastelBranco2014].
-
-Several cases exist of foreign firms engaging in large-scale malaria control campaigns [@Mouzin2011] [@Han] [@Bennett_2017] [@Kaula_2017].
-But these studies generally consider population health as the outcome measure of interest, rather than worker absenteeism or productivity.
-Similarly, they often neglect to differentiate between those clinical costs which are absorbed by the local health system versus those which are absorbed by the firm itself.
-In the literature, making the "investment case" for malaria control or elimination generally implies that the investor is the public sector, and takes into account those costs and benefits which are applicable to the public sector [@Shretta_2017]; though appropriate in most cases (the public sector being the primary malaria control agent in most locations), the findings of these studies are rarely applicable to the private sector.
-In the case of a private firm not interested in "corporate social responsibility", it is not clear whether investing in malaria control would be profitable or not.
-This lack of certainty most likely discourages investment in many cases.
-
-To address the question of the profitability of malaria control activities from the standpoint of a private firm, we analyze data during a 7 year period from a private sugar facility in Southern Mozambique.
-We assess the effect of indoor residual spraying (IRS) on the absenteeism and health of workers, and demonstrate that the firm's engagement in malaria control not only improved worker health, but also generated a positive return on investment.
-
-# Methods
-
-In collaboration with the sugar processing facility, Maragra Açucar SA (henceforth referred to as "Maragra"), we collected data for the period from January 2010 through December 2016.
-Data came from four sources: (i) the Human Resources' roster of worker details and absences, (ii) the facility's on-site clinic's medical and laboratory records, (iii) the facility's on-site malaria control program's records pertaining to the dates, chemicals, and location of IRS activities, and (iv) interviews with company employees pertaining to costs, data limitations, etc.
-Digitization and collection of data took place during the period from March 2016 through May 2017.
-Supplementary data pertaining to worker characteristics was obtained from through the Centro de Investigação em Saude de Manhiça's (CISM) demographic census, which covered workers from the district, but not those who migrated from other parts of the country [@Nhacolo_2006].
-
-Data pertaining to district-wide malaria incidence was obtained from Mozambique’s Boletim Epidemiológico Semanal (BES), which is the system by which the National Malaria Control Program monitors incidence at the district level throghout the entire country, and reports
-the number of confirmed weekly malaria cases at government health facilities.
-Using these case numbers, combined with population estimates from the National Statistical Institute (INE), we estimate each day's annualized weekly malaria incidence rate (cases per 1000 population at risk), interpolated from the weekly figures.
-We retrieved weather data for all Mozambican stations from NOAA.
-We estimated the meteorological conditions at the centroid of Manhiça using a simple interpolation method whereby the district's weather conditions were estimated to be a function of all Mozambican weather stations' reported conditions, inversely weighted by kilometers from district centroid.
-
-Maragra regularly employs IRS at on-site worker households in order to reduce those workers' (and their families') risk of malaria infection.
-By using the CISM's census, we were able to identify from the record of houses sprayed (maintained by the company) who lived in those houses, as well as who lived off-site (and therefore did not receive IRS from Maragra).
-Then, using company HR and clinical records, we were able to identify absences and episodes of clinical malaria among all workers, as well as identify the time since the most recent IRS episode before the onsent of absence or illness.
-
-Worker characteristics, illness and absenteeism data, along with IRS activity data, were systematically stored, collected, and used at the individual level by Maragra, and therefore of generally high quality.
-We sought to understand the effect of IRS on individual workers' likelihood of absence from work as well as their likelihood of clinical malaria.
-To estimate this effect, we estimated separate models for absence and illness.
-We employed interrupted time series [@Lopez_Bernal_2016] and a linear probability approach using the following econometric model.
-
-$$
-\hat{Y_{it}} = \beta_{0} +  (\beta_{1}) (\text{Season}_{t}) + (\beta_2{IRS}*\beta_3{IRS_t}) + ... + \epsilon
-$$
-
-$\hat{Y}$ is the rate of absence.
-$\beta_{1}$ represents the clinical malaria incidence at that time in the entire district of Manhiça.
-Our demographic confounders (represented by $...$) are sex, age, and worker department (field, factory, or administrative).
-Our intervention was not a simple yes/no, but rather the product of whether the residence of the worker in question was treated in the last year, and, if so, the time since treatment (represented above as the interaction term, where where $_t$ represents time elapsed since commencement of the most recent IRS campaign).
-We define the malaria season as any time during which the clinical incidence of malaria in the district of Manhiça was at or greater than the median clinical incidence of malaria for the entire study period.
-These weeks are flagged as red in Figure 1, Panel A.
-By using clinical incidence of the area of residence of the workers (as opposed to more typical proxies for malaria risk, such as only rainy vs.
-non rainy season), our seasonality estimate is a closer approximation of true malaria risk, incorporating lagged effects such as the incubation period of the parasite, as well as any inherent non-linear effects of weather.
-In addition, we adjust for daily precipitation
-
-Because cost data was less systematically collected by Maragra, and because many costs could not be precisely quantified due to the abundance of in-kind and cross-departmental expenditures, we had to rely on rough estimations based on a mix of interviews, reciepts, and interpolations.
-Since our program cost data is not as reliable as our worker characteristic and outcome data, we were conservative in our estimates, and generally tried to err on the side of program activities and materials costing _more_ than what was reported, when doubt was aired.
-Cost data consisted of three types: (i) wages of malaria control employees, (ii) transporation and vehicle costs for IRS teams, and (iii) acquisition costs of purchasing IRS chemicals for fumigation (ACT and DDT), the latter two being combined into malaria control "programme" costs.
-We define the IRS program as "profitable" if return on investment is greater than 100%, ie if the savings associated with the estimated effect of IRS is greater than the costs of the program's administration.
-
-Our formula for return on investment can be described in a straightforward fashion...
-
-\begin{center}
-$R = \dfrac{P_{w} - S_{wa} - S_{wc}}{P_{w}}$
-
-\end{center}
-
-...where $R$ is the return on investment, $P$ is the malaria control program's total operating cost, $w$ refers to costs at the per-worker level, $a$ refers to savings through avoided absences, and $ c $ refers to savings through avoided clinical encounters.
-
-All data processing and analysis were carried out in R [@R] and all analysis code is freely available online [@brewgit].
-Ethical approval for this project was obtained from the Institutional Ethics Review Board for Health at the CISM prior to data collection.
-
 # Results
@@
     mutate(year = format(date, '%Y')) %>%
-    mutate(maragra_fabrica = ifelse(maragra_fabrica,
+    mutate(on_site = ifelse(on_site,
                                     'On site',
@@
   names(out)[1] <- 'variable'
-  out$val <- paste0(round(out$absenteeism_rate, digits = 2),
+  out$val <- paste0(round(out$absenteeism_rate, digits = 1),
                     '%')#,
@@
   out$variable <- as.character(out$variable)
-  new_var <- ifelse(var == 'maragra_fabrica',
+  new_var <- ifelse(var == 'on_site',
                     'Residence',
@@
                            var))
-  out$variable[1] <- paste0(Hmisc::capitalize(new_var), ': ', out$variable[1])
+  left <- data.frame(x = c(new_var, rep('', (nrow(out) - 1))))
+  out <- cbind(left, out)
+  names(out)[1:2] <- c('Variable', 'x')
+  # out$variable[1] <- paste0(Hmisc::capitalize(new_var), ': ', out$variable[1])
   return(out)
@@
 vars <- c('season', 
-           'department',
+           'field',
+          'permanent_or_temporary',
            'sex',
-           'maragra_fabrica',
+           'on_site',
            'rainy')
@@
 ab_table <- bind_rows(out_list)
-names(ab_table)[1] <- ''
+ab_table$Variable <- Hmisc::capitalize(ab_table$Variable)
+ab_table$Variable[ab_table$Variable == 'Season'] <- 'Malaria season'
+ab_table$Variable[ab_table$Variable == 'Field'] <- 'Worker type'
+ab_table$Variable[ab_table$Variable == 'Permanent_or_temporary'] <- 'Contract'
+ab_table$x <- Hmisc::capitalize(ab_table$x)
+names(ab_table)[2] <- ''
 print(xtable::xtable(ab_table,
                      caption = 'Absenteeism rate by year and worker characteristics',
-                     align = c('r' ,'r', rep('l', 4))),
+                     align = c('r', 'r', 'l', rep('l', 4))),
       include.rownames = FALSE)
@@
 dummy <- expand.grid(season = sort(unique(model_data$season)),
-                     months_since = sort(unique(model_data$months_since)))
-predictions <- predict(fit, dummy, se.fit = TRUE, interval = 'confidence')
+                     months_since = sort(unique(model_data$months_since)),
+                     on_site = TRUE)
+predictions <- predict(fit, dummy, se.fit = TRUE, interval = 'confidence', level = 0.99)
 dummy$predicted <- predictions$fit[,1]
@@
   geom_point(color = 'blue', alpha = 0.5) + 
-  geom_smooth(se = FALSE,
-              color = 'black',
-              # span = 10,
-              alpha = 0.5) +
+  geom_line(color = 'blue', alpha = 0.5) +
+  # geom_smooth(se = FALSE,
+  #             color = 'black',
+  #             # span = 10,
+  #             alpha = 0.5) +
   facet_wrap(~season) +
@@
 
-Details on full model here. Bla bla bla. See table 3.
+We create 4 worker fixed effects models. Different models for field vs not field, permanent vs temporary.
 
@@
 clean_up_model <- function(x){
+  # extract coefficients
+  coefs <- data.frame(coef(summary(x)))
+  # use normal distribution to approximate p-value
+  coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
+  ps <- coefs$p.z
   x = broom::tidy(x)
@@
   }
-  x$term <- gsub('seasonhigh', 'High transmission season', x$term)
+  x$term <- gsub('seasonhigh', 'Malaria season', x$term)    
+  x$term <- gsub('seasonlow', 'Low malaria season', x$term)
   x$term <- gsub('months_since', 'Months since IRS: ', x$term)
@@
   x$term <- gsub('rainyTRUE', 'Rainy day', x$term)
+  x$term <- gsub('on_siteTRUE', 'On site', x$term)
+  x$term <- gsub('on_siteFALSE', 'Off site', x$term)
+  x$term <- gsub('fieldNot field worker', 'Not field worker', x$term)
   names(x) <- Hmisc::capitalize(names(x))
   names(x) <- gsub('.', ' ', names(x), fixed = TRUE)
+  x$`P value` <- NA
+  x$`P value`[1:length(ps)] <- ps
   x <- x %>% 
-    mutate(Estimate = paste0(round(Estimate, 3), ' (P = ',
+    mutate(Estimate = paste0(round(Estimate, 3), ' (P',
                              ifelse(`P value` <= 0.001,
-                                    '< 0.001',
-                                    round(`P value`, 4)), ')')) %>%
+                                    '<0.001',
+                                    paste0('=', round(`P value`, 3))
+                                    ), ')')) %>%
     dplyr::select(Term, Estimate)
+  x <- x %>% filter(!grepl('sd_', Term))
   return(x)
 }
-# Combine them all
+
 out_list <- list()
-for (i in 1:length(models)){
-  this_model <- letters[i]
-  the_model <- models[[this_model]]
+for(i in 1:length(fe_models)){
+  message(i)
+  this_model <- names(fe_models)[i]
+  the_model <- fe_models[[which(names(fe_models) == this_model)]]
   the_model <- clean_up_model(the_model)
-  names(the_model)[2] <- paste0('Model ', i)
+  names(the_model)[2] <- this_model
   out_list[[i]] <- the_model
 }
-df <- out_list[[length(models)]]
-for(i in (length(models) - 1):1){
+
+# df <- out_list[[1]]
+# for(i in 2:length(out_list)){
+#   df <- left_join(df, out_list[[i]])
+# }
+# print(xtable::xtable(df, caption = 'Regression outputs'), size="\\fontsize{5pt}{6pt}\\selectfont",
+#       include.rownames = FALSE)
+
+df <- out_list[[1]]
+namer <- names(df)[2]
+names(df)[2] <- 'Estimate'
+
+for(i in 2:length(out_list)){
+  temp <- out_list[[i]]
+  namer[i] <- names(temp)[2]
+  names(temp)[2] <- 'Estimate'
+  df <- bind_rows(df, temp)
+}
+
+library(kableExtra)
+breaker <- nrow(out_list[[1]])
+lengther <- length(out_list)
+breaks <- c(0, breaker * 1:(lengther))
+k <- kable(df, format = "latex", caption = "Models with worker fixed effects", booktabs = T) %>%
+  kable_styling()
+for (i in 1:(length(breaks)-1)){
   message(i)
-    df <- left_join(df, out_list[[i]])
+  k <- k %>%
+    group_rows(namer[i], 
+               breaks[i] + 1,
+               breaks[i] + breaker,
+               latex_gap_space = "1.5em") 
 }
-df <- df[,unique(c('Term', sort(names(df))))]
-print(xtable::xtable(x, caption = ))
+print(k)
+
+# # Combine them all
+# out_list <- list()
+# for (i in 1:length(models)){
+#   this_model <- letters[i]
+#   the_model <- models[[this_model]]
+#   the_model <- clean_up_model(the_model)
+#   names(the_model)[2] <- paste0('Model ', i)
+#   out_list[[i]] <- the_model
+# }
+# df <- out_list[[length(models)]]
+# # Order the Terms
+# df_order <- c()
+# for(i in 1:length(models)){
+#   df_order <- c(df_order, out_list[[i]]$Term)
+# }
+# df_order <- unique(df_order)
+# df <- left_join(data.frame(Term = df_order),
+#                 df)
+# for(i in (length(models) - 1):1){
+#   message(i)
+#     df <- left_join(df, out_list[[i]])
+# }
+# df <- df[,unique(c('Term', sort(names(df))))]
+# print(xtable::xtable(df, caption = 'Regression outputs'), size="\\fontsize{5pt}{6pt}\\selectfont")
 
@@
 
-\vfill
+## Return on investment
 
-\pagebreak
+Details here on ROI calculation outcomes \todo{Not done yet. In conversation with Eduardo from MC about getting more accurate costs}
 
-# Discussion
+## Robustness and generalizability
 
-Some text goes here
-
-# References
+Two principal concerns call into question the results of our analysis.
+First, the application of IRS to a workers house may be endogenous.
+To check for this  bla bla bla \todo{As per Menno's suggestion, will add details here pertaining to whether IRS predicts absenteeism before it is applied}.
+The second concern is that our quantification of consists is distorted by the fact that we treat IRS operations as essentially linear in nature, when in reality economies of scale, in-kind purchases and other factors likely make the true cost-per-spraying convex.\todo{Will address this with some sensitivity analysis}

### 2018-01-03 inst/rmd/04_discussion.Rmd
major update

--- 2017-10-24 inst/rmd/ihmt_presentation_brew_2017.Rmd
@@
----
-title: !r "Effectiveness of private sector malaria control: The case of sugarcane workers in Mozambique"
-subtitle: IHMT, October, 2017, Lisboa, Portugal
-fig_height: 5
-fig_width: 7
-bibliography: bibliography.bib
-output:
-  beamer_presentation:
-    includes:
-      in_header: preamble_beamer.sty
-    theme: "Berlin"
-    colortheme: "seagull"
-    fonttheme: "structuresmallcapsserif"
----
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = F, results = 'hide'}
-# All beamer themes at http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme_and_color.html
-library(knitr)
-opts_chunk$set(comment = NA, 
-               echo = FALSE, 
-               warning = FALSE, 
-               message = FALSE, 
-               error = TRUE, 
-               cache = FALSE)
-
-# Packages
-library(ggplot2)
-library(cism)
-library(rgeos)
-library(maptools)
-library(rgdal)
-library(tidyr)
-library(RColorBrewer)
-library(dplyr)
-library(leaflet)
-library(readr)
-library(ggthemes)
-library(ggrepel)
-library(knitr)
-options(knitr.table.format = "latex")
-library(maragra)
-```
-
-```{r, results = 'asis'}
-# Print the author names and affiliations
-x <- make_authors(include_address = FALSE, include_country = TRUE, seperator = ' \n\n ', cat_it = FALSE)
-cat(x)
-```
-
-# Introduction
-
-## Context
-
-- Malaria has a nearly unquantifiably large economic impact.
-- Many channels: fertility, fecundity, saving, investment [@Shretta2016], risk perception, productivity, absenteeism, human capital accumulation [@CastelBranco2014], mortality, costs of care [@Sachs2002].
-- Cost-benefit studies often only consider the costs of an intervention and associated costs of care, without quantifying the societal cost of non-intervention.
-
-## What we already know
-
-- Malaria is associated with absenteeism in workers [@Nonvignon_2016].
-- Malaria has a negative effect on GDP [@Orem_2012] and growth [@McCarthy_2000].
-- Malaria control is cost-effective from the societal/public perspective [@Purdy_2013].
-- Indoor residual spraying (IRS) is cost-effective [@Howard_2017], [@White_2011].
-
-## What we want to know
-
-What is the investment case from the private perspective?
-
-- From a purely financial perspective, what benefits does a private company experience?  
-- What is the short-term cost-effectiveness of IRS for large companies in malaria-endemic regions?
-- Is malaria control just good "corporate social responsibility"? Or is it also good business?
-
-## Research questions
-
-We can't answer all the previous questions (yet). So we focus on one:
-
-**What is the short-term effect on absenteeism of IRS at the homes of sugarcane workers?**
-
-# Methods
-
-## Reproducibility
-
-- **Code:** [github.com/joebrew/maragra](https://github.com/joebrew/maragra)
-- **Data:** Only available to authorized collaborators, but "dummy data" has been created to validate methods  
-- **Code bundling:** All code is in an R package: `devtools::install_github('joebrew/maragra)`
-
-## Identification strategy
-
-Some text here.
-
-# Results
-
-## Reduction in absenteeism
-
-Some text here.
-
-## Reduction in illness
-
-Some text here.
   
@@
 
-## General
-
-## Implications
+Overall stuff
 
@@
 
-## Thank you
-
-**Email:** joe@economicsofmalaria.com
-
-**Presentation:** [economicsofmalaria.com/ihmt](http://economicsofmalaria.com/ihmt)
-
-## References
+## Implications

### 2018-01-03 inst/rmd/paper.Rmd
major update

--- 2017-12-02 inst/rmd/paper.Rmd
@@
 ---
-title: Effectiveness of private sector malaria control at a large sugar
+title: Return on investment of private sector malaria control at a large sugar
   facility in Southern Mozambique
@@
   pdf_document:
+    keep_tex: true
+    latex_engine: xelatex
     includes:
@@
 bibliography: bibliography.bib
+# csl: journal-of-health-economics.csl
 ---
@@
 library(extrafont)
+library(kableExtra)
 loadfonts()
@@
                warning=FALSE,
-               dev = "cairo_pdf")
+               dev = "cairo_pdf",
+               fig.pos="H",
+               fig.path = 'figures/')
 opts_knit$set(width=75)
+options(xtable.comment = FALSE)
+
 ```
@@
 
-\begin{center}
-\begin{large}
+```{r}
+# Build document from contents of directory
+# Get a list of all the child documents
+rmds <- list.files(pattern = '*.Rmd', recursive = FALSE, include.dirs = FALSE)
+# Must include an underscore!
+rmds <- rmds[grepl('_', rmds)]
 
-```{r, results = 'asis', eval = TRUE}
-make_authors()
-```
+chunks <- paste0("\n\n```{r child = '", 
+                 rmds, 
+                 "'}\n```\n")
 
-\end{large}
-\end{center}
-
-\vspace{5mm}
-
-\begin{center}
-\textbf{Abstract}  
-\end{center}
-
-\vspace{5mm}
-\begin{center}
-\begin{changemargin}{3cm}{3cm}
-
-\textbf{Context:} Incentivizing scaled-up private firms' involvement in malaria control and elimination activities requires a clear and robust demonstration of its benefits, and a quantification of which benefits can be expected to return to the firm.
-However, most studies to date examine benefits of malaria control purely from the perspective of the public sector.
-\textbf{Methods:} We analyze 4 years of malaria control and worker health and absenteeism data from a large sugar processing facility in Mozambique.
-We estimate the effect of indoor residual spraying (IRS) on worker health and absenteeism.
-\textbf{Results:} We demonstrate that the firm's engagement in malaria control activities is profitable, even if we only consider those benefits which directly affect the firm's bottom line and ignore secondary and societal benefits.
-
-\end{changemargin}
-\end{center}
-
-# Introduction
-
-Malaria has a nearly unquantifiably large economic impact on endemic societies.
-By affecting saving, investment [@Shretta2016], risk perception, productivity, absenteeism [@Nonvignon_2016], human capital accumulation [@CastelBranco2014], mortality, and costs of care [@Sachs2002], malaria likely has a negative effect on GDP and growth [@McCarthy_2000] [@Orem_2012].
-Because of the relative affordability of most intereventions and the enormous societal costs of malaria, most forms of malaria control are cost-effective from the perspective of the government [@White_2011] [@Purdy_2013] [@Howard_2017].
-
-From the persepctive of the private sector, however, investing in malaria control is not such a clear-cut case.
-Public health interventions targetting malaria - and their corresponding cost-effectiveness evaluations - most often pertain to the public sector.
-Accordingly, the analytical framework and metrics employed by these analyses most often focus on impact pertaining to the public good, such as an increase in life years adjusted for disability or quality  [@Goodman1999] [@Shretta2016] [@Lee2017] [@Hanson2004].
-Though population-level health is certainly of importance to businesses, and improvements in health incidentally improve the economy at all levels [@World1999] [@Bloom2008] [@Vecchi_2013], these improvements may be too disperse or long-term to incentivize private sector involvement in health campaigns.
-
-In the context of Mozambique, where a significant sector of the economy is dominated by a full large-scale foreing direct investment projects [@Robbins2012], the role of the private sector in health generally, and malaria specifically, is unequivocally important.
-Large agriculture and extractive industry firms take up wide swaths of land and employ hundreds of thousands  [@German2013].
-The Mozambican state has encouraged large-scale entreprise with the aim of general economic development [@Buur2012].
-And where large firms exist, they often take on social roles such as housing and health care [@Winkler].
-At times, this role is necessary from a purely practical standpoint; in other cases, it is employed under the guise of "corporate social responsibility" [@Azemar2009].
-Regardless of the language used, it is clear that private industry plays an important role in public health in Mozambique [@Robbins2012] [@CastelBranco2014].
-
-Several cases exist of foreign firms engaging in large-scale malaria control campaigns [@Mouzin2011] [@Han] [@Bennett_2017] [@Kaula_2017].
-But these studies generally consider population health as the outcome measure of interest, rather than worker absenteeism or productivity.
-Similarly, they often neglect to differentiate between those clinical costs which are absorbed by the local health system versus those which are absorbed by the firm itself.
-In the literature, making the "investment case" for malaria control or elimination generally implies that the investor is the public sector, and takes into account those costs and benefits which are applicable to the public sector [@Shretta_2017]; though appropriate in most cases (the public sector being the primary malaria control agent in most locations), the findings of these studies are rarely applicable to the private sector.
-In the case of a private firm not interested in "corporate social responsibility", it is not clear whether investing in malaria control would be profitable or not.
-This lack of certainty most likely discourages investment in many cases.
-
-To address the question of the profitability of malaria control activities from the standpoint of a private firm, we analyze data during a 7 year period from a private sugar facility in Southern Mozambique.
-We assess the effect of indoor residual spraying (IRS) on the absenteeism and health of workers, and demonstrate that the firm's engagement in malaria control not only improved worker health, but also generated a positive return on investment.
-
-# Methods
-
-In collaboration with the sugar processing facility, Maragra Açucar SA (henceforth referred to as "Maragra"), we collected data for the period from January 2010 through December 2016.
-Data came from four sources: (i) the Human Resources' roster of worker details and absences, (ii) the facility's on-site clinic's medical and laboratory records, (iii) the facility's on-site malaria control program's records pertaining to the dates, chemicals, and location of IRS activities, and (iv) interviews with company employees pertaining to costs, data limitations, etc.
-Digitization and collection of data took place during the period from March 2016 through May 2017.
-Supplementary data pertaining to worker characteristics was obtained from through the Centro de Investigação em Saude de Manhiça's (CISM) demographic census, which covered workers from the district, but not those who migrated from other parts of the country [@Nhacolo_2006].
-
-Data pertaining to district-wide malaria incidence was obtained from Mozambique’s Boletim Epidemiológico Semanal (BES), which is the system by which the National Malaria Control Program monitors incidence at the district level throghout the entire country, and reports
-the number of confirmed weekly malaria cases at government health facilities.
-Using these case numbers, combined with population estimates from the National Statistical Institute (INE), we estimate each day's annualized weekly malaria incidence rate (cases per 1000 population at risk), interpolated from the weekly figures.
-We retrieved weather data for all Mozambican stations from NOAA.
-We estimated the meteorological conditions at the centroid of Manhiça using a simple interpolation method whereby the district's weather conditions were estimated to be a function of all Mozambican weather stations' reported conditions, inversely weighted by kilometers from district centroid.
-
-Maragra regularly employs IRS at on-site worker households in order to reduce those workers' (and their families') risk of malaria infection.
-By using the CISM's census, we were able to identify from the record of houses sprayed (maintained by the company) who lived in those houses, as well as who lived off-site (and therefore did not receive IRS from Maragra).
-Then, using company HR and clinical records, we were able to identify absences and episodes of clinical malaria among all workers, as well as identify the time since the most recent IRS episode before the onsent of absence or illness.
-
-Worker characteristics, illness and absenteeism data, along with IRS activity data, were systematically stored, collected, and used at the individual level by Maragra, and therefore of generally high quality.
-We sought to understand the effect of IRS on individual workers' likelihood of absence from work as well as their likelihood of clinical malaria.
-To estimate this effect, we estimated separate models for absence and illness.
-We employed interrupted time series [@Lopez_Bernal_2016] and a linear probability approach using the following econometric model.
-
-$$
-\hat{Y_{it}} = \beta_{0} +  (\beta_{1}) (\text{Season}_{t}) + (\beta_2{IRS}*\beta_3{IRS_t}) + ... + \epsilon
-$$
-
-$\hat{Y}$ is the rate of absence.
-$\beta_{1}$ represents the clinical malaria incidence at that time in the entire district of Manhiça.
-Our demographic confounders (represented by $...$) are sex, age, and worker department (field, factory, or administrative).
-Our intervention was not a simple yes/no, but rather the product of whether the residence of the worker in question was treated in the last year, and, if so, the time since treatment (represented above as the interaction term, where where $_t$ represents time elapsed since commencement of the most recent IRS campaign).
-We define the malaria season as any time during which the clinical incidence of malaria in the district of Manhiça was at or greater than the median clinical incidence of malaria for the entire study period.
-These weeks are flagged as red in Figure 1, Panel A.
-By using clinical incidence of the area of residence of the workers (as opposed to more typical proxies for malaria risk, such as only rainy vs.
-non rainy season), our seasonality estimate is a closer approximation of true malaria risk, incorporating lagged effects such as the incubation period of the parasite, as well as any inherent non-linear effects of weather.
-In addition, we adjust for daily precipitation
-
-Because cost data was less systematically collected by Maragra, and because many costs could not be precisely quantified due to the abundance of in-kind and cross-departmental expenditures, we had to rely on rough estimations based on a mix of interviews, reciepts, and interpolations.
-Since our program cost data is not as reliable as our worker characteristic and outcome data, we were conservative in our estimates, and generally tried to err on the side of program activities and materials costing _more_ than what was reported, when doubt was aired.
-Cost data consisted of three types: (i) wages of malaria control employees, (ii) transporation and vehicle costs for IRS teams, and (iii) acquisition costs of purchasing IRS chemicals for fumigation (ACT and DDT).
-We define the IRS program as "profitable" if return on investment is greater than 100%, ie if the savings associated with the estimated effect of IRS is greater than the costs of the program's administration.
-
-All data processing and analysis were carried out in R [@R] and all analysis code is freely available online [@brewgit].
-Ethical approval for this project was obtained from the Institutional Ethics Review Board for Health at CISM prior to data collection.
-
-# Results
-
-In Southern Mozambique, malaria peaks during the summer months (December through March) most years (Figure 1, panel A), and worker absenteeism rates track malaria incidence closely (Figure 1, panel B).
-Both all-cause absenteeism and sick absenteeism have declined in recent years (Figure 1, panel C), with the latter declining at a faster rate than the former.
-The fact that the rate of confirmed cases at the company clinic is largely non-seasonal (Figure 1, panels E and F) suggests that a significant portion of workers either seek care for malaria elsewhere (for example, government health posts, of which several are nearby and in some many cases closer to workers' residence than the company clinic) or did not seek care during malaria infection.
-Accordingly, we focus our analysis on all-cause absenteeism rather than sickness absenteeism or malaria diagnostics, assuming that much of illness is captured by absenteeism but not by the clinic.
-
-```{r, results='asis'}
-calculate <- function(var){
-  out <- 
-  model_data %>%
-    mutate(year = format(date, '%Y')) %>%
-    mutate(maragra_fabrica = ifelse(maragra_fabrica,
-                                    'On site',
-                                    'Off site'),
-           rainy = ifelse(rainy,
-                          'Rainy',
-                          'Dry')) %>%
-    filter(!is.na(rainy)) %>%
-  group_by_(var, 'year') %>%
-  summarise(absences = length(which(absent)),
-            eligibles = length(absent),
-            workers = length(unique(oracle_number))) %>%
-  ungroup %>%
-  mutate(absenteeism_rate = absences / eligibles * 100) 
-  names(out)[1] <- 'variable'
-  out$val <- paste0(round(out$absenteeism_rate, digits = 2),
-                    '%')#,
-                    # ')',
-                    # '\n',
-                    # ' (workers:', out$workers,
-                    # ')')
-  out <- 
-    out %>%
-    dplyr::select(variable, year, val)
-
-  out <- out %>%
-    spread(key = year,
-           value = val)
-  out$variable <- as.character(out$variable)
-  new_var <- ifelse(var == 'maragra_fabrica',
-                    'Residence',
-                    ifelse(var == 'rainy',
-                           'Precipitation',
-                           var))
-  out$variable[1] <- paste0(Hmisc::capitalize(new_var), ': ', out$variable[1])
-  return(out)
-}
-out_list <- list()
-vars <- c('season', 
-           'department',
-           'sex',
-           'maragra_fabrica',
-           'rainy')
-for(i in 1:length(vars)){
-  x <- calculate(var = vars[i])
-  out_list[[i]] <- x
-}
-ab_table <- bind_rows(out_list)
-names(ab_table)[1] <- ''
-print(xtable::xtable(ab_table,
-                     align = c('r' ,'r', rep('l', 4))),
-      include.rownames = FALSE)
-```
-
-```{r, fig.cap = 'Clinical malaria (district of Manhiça), all-cause absenteeism among Maragra workers, sick absenteeism among Maragra workers, estimated rainfall, positive cases at company clinic, and test positivity rate at company clinic', fig.height=5, fig.width=8}
-# BES incidence
-g1 <- 
-  ggplot(data = bes %>%
-           mutate(season = ifelse(p >= median(p), 'High season',
-                                  'Low season')) %>%
-           filter(weekdays(date) == 'Monday'),
-       aes(x = date,
-           y = p)) +
-  geom_point(alpha = 0.6,
-             aes(color = season),
-             size = 0.2) +
-  geom_line(alpha = 0.1) +
-  theme_maragra(base_size = 9) +
-  labs(x = 'Date',
-       y = 'Incidence',
-       title = 'A: Annualized district clinical\nmalaria incidence per 1,000') +
-  scale_color_manual(name = '',
-                     values = c('red', 'black')) + 
-  guides(color = guide_legend(ncol = 1)) +
-  theme(legend.position = c(.8, .8),
-        legend.text = element_text(size = 5),
-        legend.background = element_rect(fill=alpha('white', 0)))
-
-# Absenteeism rate
-g2_data <- ab_panel %>%
-  group_by(month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
-  summarise(absences = length(which(absent)),
-            absences_sick = length(which(absent_sick)),
-
-            eligibles = n()) %>%
-  ungroup %>%
-  mutate(absenteeism_rate = absences / eligibles * 100,
-         sick_absenteeism_rate = absences_sick / eligibles * 100)
-
-g2 <- ggplot(data = g2_data,
-             aes(x = month,
-                 y = absenteeism_rate)) +
-  geom_line(alpha = 0.6) +
-  theme_maragra(base_size = 9) +
-  labs(x = 'Date',
-       y = 'Absenteeism rate',
-       title = 'B: Monthly absenteeism rate')
-
-g3 <- ggplot(data = g2_data,
-             aes(x = month,
-                 y = sick_absenteeism_rate)) +
-  geom_line(alpha = 0.6) +
-  theme_maragra(base_size = 9) +
-  labs(x = 'Date',
-       y = 'Absenteeism rate',
-       title = 'C: Monthly sick absenteeism rate')
-
-# Weather
-g4 <-
-  ggplot(data = weather %>%
-           group_by(date = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
-           summarise(precipitation = sum(precipitation, na.rm = TRUE)),
-         aes(x = date,
-             y = precipitation)) +
-  geom_line(alpha = 0.6) +
-  theme_maragra(base_size = 9) +
-  labs(x = 'Date',
-       y = 'Mililiters',
-       title = 'D: Monthly precipitation')
-
-# Clinical incidence of malaria
-g5 <- clinic_agg %>%
-  filter(date >= '2013-01-01', date <= '2016-12-31') %>%
-  # mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
-  group_by(date) %>%
-  summarise(tested = sum(tested),
-            positive = sum(positive)) %>%
-  ungroup %>%
-  mutate(percent_positive = positive / tested * 100) %>%
-  ggplot(aes(x = date,
-             y = positive)) +
-  geom_line(alpha = 0.6) +
-  theme_maragra(base_size = 9) +
-  labs(x = 'Date',
-       y = 'Cases',
-       title = 'E: Monthly malaria cases\nat company clinic')
-
-g6 <- clinic_agg %>%
-  filter(date >= '2013-01-01', date <= '2016-12-31') %>%
-  # mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
-  group_by(date) %>%
-  summarise(tested = sum(tested),
-            positive = sum(positive)) %>%
-  ungroup %>%
-  mutate(percent_positive = positive / tested * 100) %>%
-  ggplot(aes(x = date,
-             y = percent_positive)) +
-  geom_line(alpha = 0.6) +
-  theme_maragra(base_size = 9) +
-  labs(x = 'Date',
-       y = 'Cases',
-       title = 'F: Monthly malaria test positivity rate\nat company clinic') 
-Rmisc::multiplot(g1, g2, g3, g4, g5, g6, layout = matrix(1:6, nrow=2, byrow=TRUE)) 
-```
-
-```{r}
-fumigations <- mc %>%
-  filter(date >= '2012-01-01',
-         date <= '2016-12-31')
-n_fumigations <- nrow(fumigations)
-n_buildings <- sum(fumigations$casas_cobertas, na.rm = TRUE)
-unique_agregados <- length(unique(fumigations$unidade))
-people_sprayed <- length(unique(model_data$oracle_number[model_data$days_since != 'Never']))
-people_observed <- length(unique(model_data$oracle_number))
-```
-
-During the period from January 1st, 2012 through December 31st, 2016, the Maragra Malaria Control Unit carried out `r n_fumigations` episodes of fumigation of residential "agregados" (household combinations), for a total of `r n_buildings` building-fumigation combinations.
-The total number of unique agregados sprayed during this period was `r unique_agregados`.
-Among the `r people_observed` on whom we have absenteeism data, `r people_sprayed` had their homes fumigated at least once (the majority of workers live off of the facility).
-
-```{r, fig.height = 3.5, fig.width = 6, fig.cap = 'Effect of IRS on absenteeism by season'}
-# Predict for visualization
-dummy <- expand.grid(season = sort(unique(model_data$season)),
-                     months_since = sort(unique(model_data$months_since)))
-predictions <- predict(fit, dummy, se.fit = TRUE, interval = 'confidence')
-dummy$predicted <- predictions$fit[,1]
-dummy$lwr <- predictions$fit[,2]
-dummy$upr <- predictions$fit[,3]
-ggplot(data = dummy %>%
-         filter(months_since != 'Never') %>%
-         mutate(lwr = lwr * 100,
-                upr = upr * 100,
-                predicted = predicted * 100) %>%
-         mutate(season = paste0(Hmisc::capitalize(as.character(season)), ' season')),
-       aes(x = months_since,
-           y = predicted,
-           group = 1)) +
-  geom_ribbon(aes(x = months_since,
-                  ymin = lwr,
-                  ymax = upr),
-              fill = 'lightblue',
-              alpha = 0.6) +
-  # geom_line(color = 'blue', alpha = 0.5) +
-  geom_point(color = 'blue', alpha = 0.5) + 
-  geom_smooth(se = FALSE,
-              color = 'black',
-              alpha = 0.5,
-              span = 1) +
-  facet_wrap(~season) +
-  theme_maragra() +
-  theme(axis.text.x = element_text(angle = 90, size = 8)) +
-  labs(x = 'Months since IRS',
-       y = 'Absenteeism (%)')
+# Write our order / child-calls to a doc
+file_connection <- file('children.Rmd')
+writeLines(paste0('---\noutput: html_document\n---\n\n', 
+                  chunks), 
+           file_connection)
+close(file_connection)
 
 ```
-
-\vfill
-
-\pagebreak
-
-# Discussion
-
-Some text goes here
-
-# References
+```{r child='children.Rmd', echo = FALSE}
+```
+```{r, results='hide'}
+file.remove('children.Rmd')
+```

### 2018-01-03 inst/rmd/paper.tex
major update

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Return on investment of private sector malaria control at a large sugar facility in Southern Mozambique},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Return on investment of private sector malaria control at a large sugar
facility in Southern Mozambique}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage[table]{xcolor}

% Font
\usepackage{fontspec}
\setmainfont{Swift-Regular_43151.ttf}
\setsansfont{Swift-Regular_43151.ttf}
% \setsansfont[BoldFont={Swift-Bold_43130.ttf}]{Swift-Bold_43130.ttf}
\setmonofont{Swift-Regular_43151.ttf}
\renewcommand{\familydefault}{\sfdefault}

\usepackage{multicol}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{lipsum}
\usepackage[utf8]{inputenc}

% \usepackage{todonotes} % for side notes
\usepackage[colorinlistoftodos]{todonotes} % for side notes
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Effectiveness of private sector malaria control at a large sugar
  facility in Southern Mozambique}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\newcommand{\footremember}[2]{%
    \footnote{#2}
    \newcounter{#1}
    \setcounter{#1}{\value{footnote}}%
}
\newcommand{\footrecall}[1]{%
    \footnotemark[\value{#1}]%
}

\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist

\widowpenalties 1 150

\makeatletter
\renewcommand\footnotesize{%
   \@setfontsize\footnotesize\@ixpt{11}%
   \abovedisplayskip 8\p@ \@plus2\p@ \@minus4\p@
   \abovedisplayshortskip \z@ \@plus\p@
   \belowdisplayshortskip 4\p@ \@plus2\p@ \@minus2\p@
   \def\@listi{\leftmargin\leftmargini
               \topsep 4\p@ \@plus2\p@ \@minus2\p@
               \parsep 2\p@ \@plus\p@ \@minus\p@
               \itemsep \parsep}%
   \belowdisplayskip \abovedisplayskip
}
\makeatother

\DeclareTextCommandDefault{\nobreakspace}{\leavevmode\nobreak\ }

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}

\begin{document}
\maketitle

\begin{center}
\begin{large}

Joe Brew\footremember{isglobal}{Barcelona Institute for Global Health: c/ Rosselló, 132, 5è 2a. 08036, Barcelona, Spain, Spain}\footremember{cism}{Centro de Investigação em Saúde de Manhiça: Vila da Manhiça, Bairro Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique, Mozambique}\footremember{vu}{VU University Amsterdam: De Boelelaan 1105, 1081 HV Amsterdam, Netherlands, Netherlands} \footnote{Corresponding Author}
Kizito Gondo\footremember{ma}{Maragra Açucar SA, Subsidiary of Illovo Sugar Ltd: CP 2789, Maputo, Mozambique, Mozambique}
Elton Dorkin\footrecall{ma}
Eduardo Nhamahanga\footrecall{ma}
Menno Pradhan\footrecall{vu}\footremember{uva}{University of Amsterdam: REC E, Roetersstraat 11, Amsterdam, Netherlands, Netherlands}
Laia Cirera\footrecall{isglobal}\footrecall{cism}
Ranjeeta Thomas\footremember{icl}{Imperial College London: South Kensington Campus, London SW7 2AZ, U.K., UK}
Elisa Sicuri\footrecall{isglobal}\footrecall{cism}\footrecall{icl}

\end{large}
\end{center}

\vspace{5mm}

\begin{center}
\textbf{Abstract}  
\end{center}

\vspace{5mm}

\begin{center}
\begin{changemargin}{3cm}{3cm}

This paper provides new empirical evidence regarding the return on investment of privately managed malaria control activities (indoor residual spraying with pesticides) on worker absenteeism in Mozambique. We analyze 4 years of malaria control and worker health and absenteeism data from a large sugar processing facility in Mozambique. We find that the benefits outweight the costs (ie, there is a positive return on investment) even when the consideration of benefits is limited to those directly accrued by the company. These findings suggest that the private sector may have an important role to play in malaria control in endemic areas.

\end{changemargin}
\end{center}

\vspace{20mm}

\noindent\fbox{%
    \parbox{\textwidth}{%
        \subsection*{Research Highlights}
        \begin{itemize}
          \item Large, individual-level worker absenteeism data from malaria endemic zone
          \item Quantifies effect of indoor residual spraying on absenteeism
          \item Estimates cost-effectiveness of malaria control from investment standpoint
          \item Results suggest that private sector could play a significant role in malaria elimination
        \end{itemize}
        \vspace{2mm}
    }%
}

\vfill
\null

\subsection*{Keywords}

\textbf{Malaria; Investment; Health; Productivity; Agriculture; Absenteeism}

\vspace{3mm}

\newpage

\section{Introduction}\label{introduction}

Malaria has a large economic impact on endemic societies. By affecting
saving, investment (Shretta, Avanceña, and Hatefi 2016), risk
perception, productivity, absenteeism (Nonvignon et al. 2016), human
capital accumulation (Castel-Branco 2014), mortality, and costs of care
(Sachs and Malaney 2002), malaria likely has a negative effect on GDP
and growth (McCarthy, Wolf, and Wu 2000) (Orem et al. 2012). Because of
the relative affordability of most intereventions and the enormous
societal costs of malaria, most forms of malaria control are
cost-effective when a public welfare perspective is assumed, such as
when a government provides the financing (White et al. 2011) (Purdy et
al. 2013) (Howard et al. 2017).

From the persepctive of the private sector, however, investing in
malaria control is not so clear-cut. Public health interventions
targetting malaria - and their corresponding cost-effectiveness
evaluations - most often focus on impacts pertaining to public welfare,
such as an increase in life years adjusted for disability or quality
(Goodman, Coleman, and Mills 1999) (Shretta, Avanceña, and Hatefi 2016)
(Lee et al. 2017) (Hanson 2004). Though population-level health is
certainly of importance to businesses, and improvements in health
incidentally improve the economy at all levels (Brundtland 1999) (Bloom
and Canning 2008) (Vecchi, Hellowell, and Gatti 2013), these
improvements may be too disperse or long-term to incentivize private
sector involvement in health campaigns.

100\% of the Mozambican population are at risk of malaria, living in
what the WHO classifies as a ``high transmission'' area (Moonasar et al.
2016). Annually, Mozambique has more than 8 million clinical malaria
cases (an annual incidence of approximately 300 per 1,000 residents),
with an estimated 14,000 deaths. Malaria accounts for 29\% of all
deaths, and 42\% of deaths among those under five years of age (INE
2011). Since 2013, Mozambique has seen a gradual increase in the
incidence of malaria (Moonasar et al. 2016). 100\% of the malaria in
Mozambique is of the Plasmodium falciparum species, with Anopheles
funestus, gambiae, and arabiensis as the primary mosquito vectors of the
disease (WHO 2015).

A significant sector of the economy in Mozambique is dominated by a full
large-scale foreing direct investment projects (Robbins and Perkins
2012), and the role of the private sector in health generally, and
malaria specifically, is unequivocally important. Large agriculture and
extractive industry firms take up wide swaths of land and employ
hundreds of thousands (German, Schoneveld, and Mwangi 2013). The
Mozambican state has encouraged large-scale entreprise with the aim of
general economic development (Buur, Tembe, and Baloi 2012). And where
large firms exist, they often take on social roles such as housing and
health care (Winkler 2013). At times, this role is necessary from a
purely practical standpoint; in other cases, it is employed under the
guise of ``corporate social responsibility'' (Azemar and Desbordes
2009). Regardless of the language used, it is clear that private
industry plays an important role in public health in Mozambique (Robbins
and Perkins 2012) (Castel-Branco 2014).

Several cases exist of foreign firms engaging in large-scale malaria
control campaigns (Mouzin and al. 2011) (Han 2015) (Bennett et al. 2017)
(Kaula, Buyungo, and Opigo 2017). But these studies generally consider
population health as the outcome measure of interest, rather than worker
absenteeism or productivity. Similarly, they often neglect to
differentiate between those clinical costs which are absorbed by the
local health system versus those which are absorbed by the firm itself.
In the literature, making the ``investment case'' for malaria control or
elimination generally implies that the investor is the public sector,
and takes into account those costs and benefits which are applicable
from a public welfare point of view (Shretta et al. 2017); though
appropriate in most cases (the government or institutions interested in
public welfare primarily being the primary malaria control agents in
most locations), the findings of these studies are rarely applicable to
the private sector. In the case of a private firm not interested in
``corporate social responsibility'', it is not clear whether investing
in malaria control would be profitable or not. This lack of clarity not
only discourages investment, but also makes it difficult for governments
to pinpoint the correct of amount of subsidy (if applicable) to
encourage private sector scale-up in malaria control.

To address the question of the profitability of malaria control
activities from the standpoint of a private firm, we analyze data during
a 7 year period from a private sugar facility in Southern Mozambique. We
assess the effect of indoor residual spraying (IRS) on the absenteeism
and health of workers, and demonstrate that the firm's engagement in
malaria control not only improved worker health, but also generated a
positive return on investment.

The structure of the paper is as follows. After a brief review of the
existing literature on the subject of private sector investment in
malaria control and its effects on worker productivity, we provide an
overview of the sugar company under study, and the epidemiology of
malaria in the nearby area, as well as in Mozambique as a whole. We then
give an overview of the data collected, and outline the theoretical and
methdological assumptions that underly our analysis. In the results, we
show the effectiveness of the company's malaria control program in
reducing absences, and translate this reduction into cost savings. Our
robustness checks consist of \todo{Add some stuff here}. Our discussion
covers potential implications from this study in terms of policy and
investment, as well as the paper's limitations.

\subsection{Background Literature}\label{background-literature}

\todo{Will add 3-5 paragraphs of background lit review}

Our study adds to the existing literature in several ways.
\todo{Will add details here.}

\subsection{Study area}\label{study-area}

Details on Maragra:

Maragra Açucar SA (henceforth referred to as
``Maragra'')\todo{This is a comment that will appear in the margin}

Details on Manhica

Details on Mozambique

\section{Methods}\label{methods}

\subsection{Data collection}\label{data-collection}

In collaboration with the sugar processing facility, we collected data
for the period from January 2010 through December 2016. Data came from
four sources: (i) the Human Resources' roster of worker details and
absences, (ii) the facility's on-site clinic's medical and laboratory
records, (iii) the facility's on-site malaria control program's records
pertaining to the dates, chemicals, and location of IRS activities, and
(iv) interviews with company employees pertaining to costs, data
limitations, etc. Digitization and collection of data took place during
the period from March 2016 through May 2017. Supplementary data
pertaining to worker characteristics was obtained from through the
Centro de Investigação em Saude de Manhiça's (CISM) demographic census,
which covered workers from the district, but not those who migrated from
other parts of the country (Nhacolo et al. 2006).

Data pertaining to district-wide malaria incidence was obtained from
Mozambique's Boletim Epidemiológico Semanal (BES), which is the system
by which the National Malaria Control Program monitors incidence at the
district level throghout the entire country, and reports the number of
confirmed weekly malaria cases at government health facilities. Using
these case numbers, combined with population estimates from the National
Statistical Institute (INE), we estimate each day's annualized weekly
malaria incidence rate (cases per 1000 population at risk), interpolated
from the weekly figures. We retrieved weather data for all Mozambican
stations from NOAA. We estimated the meteorological conditions at the
centroid of Manhiça using a simple interpolation method whereby the
district's weather conditions were estimated to be a function of all
Mozambican weather stations' reported conditions, inversely weighted by
kilometers from district centroid.

Maragra regularly employs IRS at on-site worker households in order to
reduce those workers' (and their families') risk of malaria infection.
Workers living off-site (our control group) also may have received IRS
at some point during the study period (from government programs). Even
though we do not have reliable person-level data on IRS carried out by
the government, off-site workers are a suitable control in the sense
that they represent ``business as usual'' (ie, what would happen if the
company carried out no IRS and relied solely on public interventions).
Using company HR and clinical records, we were able to identify absences
and episodes of clinical malaria among all workers, as well as identify
the time since the most recent IRS episode before the onsent of absence
or illness.

Worker characteristics, illness and absenteeism data, along with IRS
activity data, were systematically stored, collected, and used at the
individual level by Maragra, and therefore of generally high quality.
Because cost data was less systematically collected by Maragra, and
because many costs could not be precisely quantified due to the
abundance of in-kind and cross-departmental expenditures, we had to rely
on rough estimations based on a mix of interviews, reciepts, and
interpolations. Since our program cost data is not as reliable as our
worker characteristic and outcome data, we were conservative in our
estimates, and generally tried to err on the side of program activities
and materials costing \emph{more} than what was reported, when doubt was
aired. Cost data consisted of three types: (i) wages of malaria control
employees, (ii) transporation and vehicle costs for IRS teams, and (iii)
acquisition costs of purchasing IRS chemicals for fumigation (ACT and
DDT), the latter two being combined into malaria control ``programme''
costs.

\subsection{Conceptual framework and identification
strategy}\label{conceptual-framework-and-identification-strategy}

We sought to understand the effect of IRS on individual workers'
likelihood of absence from work as well as their likelihood of clinical
malaria. To estimate this effect, we estimated separate models for
absence and illness. We employed interrupted time series (Lopez Bernal,
Cummins, and Gasparrini 2016) and a linear probability approach using
the following econometric model.

\[
\hat{Y_{it}} = \beta_{0} +  (\beta_{1}) (\text{Season}_{t}) + (\beta_2{IRS}*\beta_3{IRS_t}) + ... + \epsilon
\]

\(\hat{Y}\) is the rate of absence. \(\beta_{1}\) represents the
clinical malaria incidence at that time in the entire district of
Manhiça. Our demographic confounders (represented by \(...\)) are sex,
age, and worker department (field, factory, or administrative). Our
intervention was not a simple yes/no, but rather the product of whether
the residence of the worker in question was treated in the last year,
and, if so, the time since treatment (represented above as the
interaction term, where where \(_t\) represents time elapsed since
commencement of the most recent IRS campaign). We define the malaria
season as any time during which the clinical incidence of malaria in the
district of Manhiça was at or greater than the median clinical incidence
of malaria for the entire study period. These weeks are flagged as red
in Figure 1, Panel A. By using clinical incidence of the area of
residence of the workers (as opposed to more typical proxies for malaria
risk, such as only rainy vs.~non rainy season), our seasonality estimate
is a closer approximation of true malaria risk, incorporating lagged
effects such as the incubation period of the parasite, as well as any
inherent non-linear effects of weather. In addition, we adjust for daily
precipitation

\subsection{Estimating return on
investment}\label{estimating-return-on-investment}

Our formula for return on investment can be described in a
straightforward fashion\ldots{}

\begin{center}
$R = \dfrac{P_{w} - S_{wa} - S_{wc}}{P_{w}}$

\end{center}

\ldots{}where \(R\) is the return on investment, \(P\) is the malaria
control program's total operating cost, \(w\) refers to costs at the
per-worker level, \(a\) refers to savings through avoided absences, and
\$ c \$ refers to savings through avoided clinical encounters. We define
the malaria control program as ``profitable'' from an investment
standpoint if ROI is greater than 100\%, ie if the savings associated
with the estimated effect of IRS is greater than the costs of the
program's administration.

\subsection{Reproducibility and ethical
approval}\label{reproducibility-and-ethical-approval}

All data processing and analysis were carried out in R (R Core Team
2017) and all analysis code is freely available online (Brew 2017).
Ethical approval for this project was obtained from the Institutional
Ethics Review Board for Health at the CISM prior to data collection.

\subsection{Descriptive statistics}\label{descriptive-statistics}

Age, types of workers, bla bla bla.
\todo{Will add a table here and some overview}

\section{Results}\label{results}

In Southern Mozambique, malaria peaks during the summer months (December
through March) most years (Figure 1, panel A), and worker absenteeism
rates track malaria incidence closely, following the same seasonal
patterns (Figure 1, panel B). Both all-cause absenteeism and sick
absenteeism have declined in recent years at Maragra (Figure 1, panel
C), with the latter declining at a faster rate than the former. The fact
that the rate of confirmed cases at the company clinic is largely
non-seasonal (Figure 1, panels E and F) suggests that a significant
portion of workers either seek care for malaria elsewhere (for example,
government health posts, of which several are nearby and in some many
cases closer to workers' residence than the company clinic) or do not
seek care during malaria infection. Accordingly, we focus our analysis
on all-cause absenteeism rather than sickness absenteeism or malaria
diagnostics, with the assumption that much of illness is captured by
absenteeism but not by the clinical data.

\begin{figure}
\centering
\includegraphics{figures/unnamed-chunk-13-1.pdf}
\caption{Clinical malaria (district of Manhiça), all-cause absenteeism
among Maragra workers, sick absenteeism among Maragra workers, estimated
rainfall, positive cases at company clinic, and test positivity rate at
company clinic}
\end{figure}

\textbf{Fumigations}: During the period from January 1st, 2012 through
December 31st, 2016, the Maragra Malaria Control Unit carried out 11,567
episodes of fumigation of residential ``agregados'' (household
combinations), for a total of 13,937 building-fumigation combinations.
The total number of unique agregados sprayed during this period was
4,045. Among the 3,362 workers for whom we have reliable absenteeism and
residential data, 692 had their homes fumigated at least once (the
majority of workers live off of the facility).

\textbf{Absences}: We observed 1,759,100 unique worker-days among the
3,362 workers. The all-period average absenteeism rate was 5.56\%,
though this rate varied widely as a function of worker department, sex,
residence, and season (table 1).

\begin{table}[ht]
\centering
\begin{tabular}{rlllll}
  \hline
Variable &  & 2013 & 2014 & 2015 & 2016 \\ 
  \hline
Malaria season & Low & 5.2\% & 7\% & 5.4\% & 3.3\% \\ 
   & High & 8.2\% & 12.8\% & 6.3\% & 2.7\% \\ 
  Worker type & Field worker & 4.4\% & 7.6\% & 3.9\% & 1.7\% \\ 
   & Not field worker & 10.7\% & 12.4\% & 11.5\% & 9.6\% \\ 
  Contract & Permanent & 12.1\% & 12.6\% & 12\% & 10.2\% \\ 
   & Temporary & 0.1\% & 5\% & 2.3\% & 0.5\% \\ 
  Sex & F & 4\% & 8.1\% & 4.4\% & 1.9\% \\ 
   & M & 8.1\% & 10\% & 6.5\% & 3.7\% \\ 
  Residence & Off site & 6.4\% & 9.6\% & 5.9\% & 3\% \\ 
   & On site & 9.4\% & 9.7\% & 6.1\% & 3.1\% \\ 
  Precipitation & Dry & 5.4\% & 7.7\% & 4.8\% & 2.4\% \\ 
   & Rainy & 7.9\% & 10.6\% & 7\% & 3.3\% \\ 
   \hline
\end{tabular}
\caption{Absenteeism rate by year and worker characteristics} 
\end{table}

\textbf{Costs}: The malaria control program at Maragra has an annual
operating budget of approximately XX, which includes the purchase of
insecticide, the wages of IRS sprayers and drivers, transportation,
record-keeping, and general administrative costs. Assuming linearity in
costs, the program spends approximately XX per agregado sprayed. With
each agregado containing an average of 2.2 workers, this translates to a
cost of XX per worker protected per season. Much of the benefit of IRS
goes to non-worker residents of sprayed agregados (who constitute a
majority), but this benefit is purposefully ignored for this analysis.

Given the likelihood that clinical data does not fully capture all
malaria cases, we do not quantify the costs of malaria infection to the
company. Rather, we first estimate the reduction in absenteeism
attributible to IRS, and then quantify the savings associated with
prevented absences. Additionally, we calculate the clinical savings of
IRS by first estimating the share of absences which are associated with
an episode of clinical malaria, and then applying the clinical cost per
case to the equivalent share of prevented absences. We intentionally
ignore the savings accrued by the public health system, as well as the
likely utility gains in secondary realms such as school absenteeism,
producitivity, etc.

\textbf{Effect of IRS on absenteeism}: IRS is associated with a
year-long, significant reduction in absenteeism during the malaria
season (figure 2). As one would expect if the mechanism by which IRS
reduces absence is through reduced malaria infection, the effect of IRS
during the low transmission season is significant, but far less
substantial in effect size.

\begin{figure}
\centering
\includegraphics{figures/unnamed-chunk-16-1.pdf}
\caption{Estimated effect of IRS on absenteeism by season}
\end{figure}

We create 4 worker fixed effects models. Different models for field vs
not field, permanent vs temporary.

\begin{table}

\caption{\label{tab:unnamed-chunk-17}Models with worker fixed effects}
\centering
\begin{tabular}[t]{ll}
\toprule
Term & Estimate\\
\midrule
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
\hspace{1em}Off site & 9.142 (P<0.001)\\
\hspace{1em}On site & 11.424 (P<0.001)\\
\hspace{1em}Malaria season & 3.424 (P<0.001)\\
\hspace{1em}Months since IRS: 01 & 3.831 (P<0.001)\\
\hspace{1em}Months since IRS: 02-04 & 4.199 (P<0.001)\\
\hspace{1em}Months since IRS: 05-09 & -3.194 (P<0.001)\\
\hspace{1em}Months since IRS: 10+ & -5.794 (P<0.001)\\
\hspace{1em}Malaria season:Months since IRS: 01 & -4.534 (P<0.001)\\
\hspace{1em}Malaria season:Months since IRS: 02-04 & -8.392 (P<0.001)\\
\hspace{1em}Malaria season:Months since IRS: 05-09 & -3.85 (P<0.001)\\
\hspace{1em}Malaria season:Months since IRS: 10+ & 4.504 (P=0.002)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
\hspace{1em}Off site & 10.862 (P<0.001)\\
\hspace{1em}On site & 9.555 (P<0.001)\\
\hspace{1em}Malaria season & 2.077 (P<0.001)\\
\hspace{1em}Months since IRS: 01 & -2.083 (P<0.001)\\
\hspace{1em}Months since IRS: 02-04 & 3.904 (P<0.001)\\
\hspace{1em}Months since IRS: 05-09 & -0.834 (P=0.084)\\
\hspace{1em}Months since IRS: 10+ & -0.268 (P=0.742)\\
\hspace{1em}Malaria season:Months since IRS: 01 & 2.735 (P<0.001)\\
\hspace{1em}Malaria season:Months since IRS: 02-04 & -5.768 (P<0.001)\\
\hspace{1em}Malaria season:Months since IRS: 05-09 & -2.441 (P<0.001)\\
\hspace{1em}Malaria season:Months since IRS: 10+ & 4.497 (P<0.001)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
\hspace{1em}Off site & 0.985 (P<0.001)\\
\hspace{1em}On site & 1.078 (P<0.001)\\
\hspace{1em}Malaria season & -0.017 (P=0.461)\\
\hspace{1em}Months since IRS: 01 & -0.4 (P<0.001)\\
\hspace{1em}Months since IRS: 02-04 & -0.547 (P<0.001)\\
\hspace{1em}Months since IRS: 05-09 & -0.554 (P<0.001)\\
\hspace{1em}Months since IRS: 10+ & -0.992 (P<0.001)\\
\hspace{1em}Malaria season:Months since IRS: 01 & -0.092 (P=0.62)\\
\hspace{1em}Malaria season:Months since IRS: 02-04 & 0.172 (P=0.256)\\
\hspace{1em}Malaria season:Months since IRS: 05-09 & 0.062 (P=0.657)\\
\hspace{1em}Malaria season:Months since IRS: 10+ & 0.068 (P=0.771)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
\hspace{1em}Off site & 4.136 (P<0.001)\\
\hspace{1em}On site & 3.834 (P<0.001)\\
\hspace{1em}Malaria season & -0.292 (P=0.16)\\
\hspace{1em}Months since IRS: 01 & -1.456 (P=0.389)\\
\hspace{1em}Months since IRS: 02-04 & 0.608 (P=0.574)\\
\hspace{1em}Months since IRS: 05-09 & -0.869 (P=0.405)\\
\hspace{1em}Months since IRS: 10+ & 10.583 (P<0.001)\\
\hspace{1em}Malaria season:Months since IRS: 01 & 6.045 (P=0.002)\\
\hspace{1em}Malaria season:Months since IRS: 02-04 & 0.604 (P=0.667)\\
\hspace{1em}Malaria season:Months since IRS: 05-09 & -1.482 (P=0.284)\\
\hspace{1em}Malaria season:Months since IRS: 10+ & -10.636 (P<0.001)\\
\bottomrule
\end{tabular}
\end{table}

\subsection{Return on investment}\label{return-on-investment}

Details here on ROI calculation outcomes
\todo{Not done yet. In conversation with Eduardo from MC about getting more accurate costs}

\subsection{Robustness and
generalizability}\label{robustness-and-generalizability}

Two principal concerns call into question the results of our analysis.
First, the application of IRS to a workers house may be endogenous. To
check for this bla bla bla
\todo{As per Menno's suggestion, will add details here pertaining to whether IRS predicts absenteeism before it is applied}.
The second concern is that our quantification of consists is distorted
by the fact that we treat IRS operations as essentially linear in
nature, when in reality economies of scale, in-kind purchases and other
factors likely make the true cost-per-spraying
convex.\todo{Will address this with some sensitivity analysis}

\section{Discussion}\label{discussion}

Overall stuff

\subsection{Limitations}\label{limitations}

\subsection{Implications}\label{implications}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-Azemar2009}{}
Azemar, C., and R. Desbordes. 2009. ``Public Governance, Health and
Foreign Direct Investment in Sub-Saharan Africa.'' \emph{Journal of
African Economies} 18 (4). Oxford University Press (OUP): 667--709.
doi:\href{https://doi.org/10.1093/jae/ejn028}{10.1093/jae/ejn028}.

\hypertarget{ref-Bennett_2017}{}
Bennett, Adam, Anton L. V. Avanceña, Jennifer Wegbreit, Chris Cotter,
Kathryn Roberts, and Roly Gosling. 2017. ``Engaging the Private Sector
in Malaria Surveillance: A Review of Strategies and Recommendations for
Elimination Settings.'' \emph{Malaria Journal} 16 (1). Springer Nature.
doi:\href{https://doi.org/10.1186/s12936-017-1901-1}{10.1186/s12936-017-1901-1}.

\hypertarget{ref-Bloom2008}{}
Bloom, David, and David Canning. 2008. ``Population Health and Economic
Growth,'' 1--25.

\hypertarget{ref-brewgit}{}
Brew, J.R. 2017. ``Malaria and Sugar: An in-Depth Examination of the
Effect of Malaria Control Activities on the Health and Productivity of
Maragra Sugarcane Factory Workers.'' \emph{GitHub Repository}.
\url{https://github.com/joebrew/maragra}; GitHub.

\hypertarget{ref-World1999}{}
Brundtland, Gro Harlem. 1999. ``WHO on Health and Economic
Productivity'' 25 (2): 396--402.

\hypertarget{ref-Buur2012}{}
Buur, Lars, Carlota Mondlane Tembe, and Obede Baloi. 2012. ``The White
Gold: The Role of Government and State in Rehabilitating the Sugar
Industry in Mozambique.'' \emph{Journal of Development Studies} 48 (3).
Informa UK Limited: 349--62.
doi:\href{https://doi.org/10.1080/00220388.2011.635200}{10.1080/00220388.2011.635200}.

\hypertarget{ref-CastelBranco2014}{}
Castel-Branco, Carlos Nuno. 2014. ``Growth, Capital Accumulation and
Economic Porosity in Mozambique: Social Losses, Private Gains.''
\emph{Review of African Political Economy} 41 (sup1). Informa UK
Limited: S26--S48.
doi:\href{https://doi.org/10.1080/03056244.2014.976363}{10.1080/03056244.2014.976363}.

\hypertarget{ref-German2013}{}
German, Laura, George Schoneveld, and Esther Mwangi. 2013.
``Contemporary Processes of Large-Scale Land Acquisition in Sub-Saharan
Africa: Legal Deficiency or Elite Capture of the Rule of Law?''
\emph{World Development} 48 (August). Elsevier BV: 1--18.
doi:\href{https://doi.org/10.1016/j.worlddev.2013.03.006}{10.1016/j.worlddev.2013.03.006}.

\hypertarget{ref-Goodman1999}{}
Goodman, CA, PG Coleman, and AJ Mills. 1999. ``Cost-Effectiveness of
Malaria Control in Sub-Saharan Africa.'' \emph{The Lancet} 354 (9176).
Elsevier BV: 378--85.
doi:\href{https://doi.org/10.1016/s0140-6736(99)02141-8}{10.1016/s0140-6736(99)02141-8}.

\hypertarget{ref-Han}{}
Han, Lily. 2015. ``Malaria in Mozambique: trialling payment by
results.''
\url{http://www.theguardian.com/global-development-professionals-network/2014/mar/31/malaria-control-payment-by-results}.

\hypertarget{ref-Hanson2004}{}
Hanson, K. 2004. ``Public and Private Roles in Malaria Control: The
Contributions of Economic Analysis.'' \emph{The American Journal of
Tropical Medicine and Hygiene} 71 (9176). The American Society of
Tropical Medicine; Hygiene: 168--73.
\url{http://www.ajtmh.org/docserver/fulltext/14761645/71/2_suppl/0700168.pdf?expires=1503070750\&id=id\&accname=guest\&checksum=F969ACD7BE7E3A5D929D87EEA8A6EA2B}.

\hypertarget{ref-Howard_2017}{}
Howard, Natasha, Lorna Guinness, Mark Rowland, Naeem Durrani, and
Kristian S. Hansen. 2017. ``Cost-Effectiveness of Adding Indoor Residual
Spraying to Case Management in Afghan Refugee Settlements in Northwest
Pakistan During a Prolonged Malaria Epidemic.'' Edited by Guilherme
S.Editor Ribeiro. \emph{PLOS Neglected Tropical Diseases} 11 (10).
Public Library of Science (PLoS): e0005935.
doi:\href{https://doi.org/10.1371/journal.pntd.0005935}{10.1371/journal.pntd.0005935}.

\hypertarget{ref-estatistica2009}{}
INE. 2011. ``Demographic Health Survey.'' \url{http://www.ine.gov.mz/}.

\hypertarget{ref-Kaula_2017}{}
Kaula, Henry, Peter Buyungo, and Jimmy Opigo. 2017. ``Private Sector
Role, Readiness and Performance for Malaria Case Management in Uganda,
2015.'' \emph{Malaria Journal} 16 (1). Springer Nature.
doi:\href{https://doi.org/10.1186/s12936-017-1824-x}{10.1186/s12936-017-1824-x}.

\hypertarget{ref-Lee2017}{}
Lee, Bruce Y., Eli Zenkov, Chandrani Chatterjee, Baltazar Candrinho,
Shufang Zhang, James Colborn, Olivier J. T. Briët, et al. 2017. ``The
Economic Value of Long-Lasting Insecticidal Nets and Indoor Residual
Spraying Implementation in Mozambique.'' \emph{The American Journal of
Tropical Medicine and Hygiene} 96 (6). American Society of Tropical
Medicine; Hygiene: 1430--40.
doi:\href{https://doi.org/10.4269/ajtmh.16-0744}{10.4269/ajtmh.16-0744}.

\hypertarget{ref-Lopez_Bernal_2016}{}
Lopez Bernal, James, Steven Cummins, and Antonio Gasparrini. 2016.
``Interrupted Time Series Regression for the Evaluation of Public Health
Interventions: A Tutorial.'' \emph{International Journal of
Epidemiology}, June. Oxford University Press (OUP), dyw098.
doi:\href{https://doi.org/10.1093/ije/dyw098}{10.1093/ije/dyw098}.

\hypertarget{ref-McCarthy_2000}{}
McCarthy, Desmond, Holger Wolf, and Yi Wu. 2000. ``The Growth Costs of
Malaria,'' February. National Bureau of Economic Research.
doi:\href{https://doi.org/10.3386/w7541}{10.3386/w7541}.

\hypertarget{ref-Moonasar_2016}{}
Moonasar, Devanand, Rajendra Maharaj, Simon Kunene, Baltazar Candrinho,
Francisco Saute, Nyasatu Ntshalintshali, and Natashia Morris. 2016.
``Towards Malaria Elimination in the Mosaswa (Mozambique, South Africa
and Swaziland) Region.'' \emph{Malaria Journal} 15 (1). Springer Nature.
doi:\href{https://doi.org/10.1186/s12936-016-1470-8}{10.1186/s12936-016-1470-8}.

\hypertarget{ref-Mouzin2011}{}
Mouzin, Eric, and Et al. 2011. ``Business Investing in Malaria Control:
Economic Returns and a Healthy Workforce for Africa.'' \emph{Progress \&
Impact Series}, no. 6.

\hypertarget{ref-Nhacolo_2006}{}
Nhacolo, Ariel Q, Delino A Nhalungo, Charfudin N Sacoor, John J Aponte,
Ricardo Thompson, and Pedro Alonso. 2006. ``Levels and Trends of
Demographic Indices in Southern Rural Mozambique: Evidence from
Demographic Surveillance in Manhiça District.'' \emph{BMC Public Health}
6 (1). Springer Nature.
doi:\href{https://doi.org/10.1186/1471-2458-6-291}{10.1186/1471-2458-6-291}.

\hypertarget{ref-Nonvignon_2016}{}
Nonvignon, Justice, Genevieve Cecilia Aryeetey, Keziah L. Malm, Samuel
Agyei Agyemang, Vivian N. A. Aubyn, Nana Yaw Peprah, Constance N.
Bart-Plange, and Moses Aikins. 2016. ``Economic Burden of Malaria on
Businesses in Ghana: A Case for Private Sector Investment in Malaria
Control.'' \emph{Malaria Journal} 15 (1). Springer Nature.
doi:\href{https://doi.org/10.1186/s12936-016-1506-0}{10.1186/s12936-016-1506-0}.

\hypertarget{ref-Orem_2012}{}
Orem, Juliet, Joses Kirigia, Robert Azairwe, Ibrahim Kasirye, and
Oladapo Walker. 2012. ``Impact of Malaria Morbidity on Gross Domestic
Product in Uganda.'' \emph{International Archives of Medicine} 5 (1).
Springer Nature: 12.
doi:\href{https://doi.org/10.1186/1755-7682-5-12}{10.1186/1755-7682-5-12}.

\hypertarget{ref-Purdy_2013}{}
Purdy, Mark, David Rublin, Kuangyi Wei, and Matthew Robinson. 2013.
``The Economic Case for Combating Malaria.'' \emph{The American Journal
of Tropical Medicine and Hygiene} 89 (5). American Society of Tropical
Medicine; Hygiene: 819--23.
doi:\href{https://doi.org/10.4269/ajtmh.12-0689}{10.4269/ajtmh.12-0689}.

\hypertarget{ref-R}{}
R Core Team. 2017. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\hypertarget{ref-Robbins2012}{}
Robbins, Glen, and David Perkins. 2012. ``Mining Fdi and Infrastructure
Development on Africa's East Coast: Examining the Recent Experience of
Tanzania and Mozambique.'' \emph{Journal of International Development}
24 (2). Wiley-Blackwell: 220--36.
doi:\href{https://doi.org/10.1002/jid.2817}{10.1002/jid.2817}.

\hypertarget{ref-Sachs2002}{}
Sachs, Jeffrey, and Pia Malaney. 2002. ``The Economic and Social Burden
of Malaria.'' \emph{Nature} 415 (6872). Springer Nature: 680--85.
doi:\href{https://doi.org/10.1038/415680a}{10.1038/415680a}.

\hypertarget{ref-Shretta2016}{}
Shretta, Rima, Anton L. V. Avanceña, and Arian Hatefi. 2016. ``The
Economics of Malaria Control and Elimination: A Systematic Review.''
\emph{Malaria Journal} 15 (1). Springer Nature.
doi:\href{https://doi.org/10.1186/s12936-016-1635-5}{10.1186/s12936-016-1635-5}.

\hypertarget{ref-Shretta_2017}{}
Shretta, Rima, Ranju Baral, Anton L. V. Avanceña, Katie Fox, Asoka
Premasiri Dannoruwa, Ravindra Jayanetti, Arumainayagam Jeyakumaran,
Rasike Hasantha, Lalanthika Peris, and Risintha Premaratne. 2017. ``An
Investment Case to Prevent the Reintroduction of Malaria in Sri Lanka.''
\emph{The American Journal of Tropical Medicine and Hygiene}, January.
American Society of Tropical Medicine; Hygiene, 16--0209.
doi:\href{https://doi.org/10.4269/ajtmh.16-0209}{10.4269/ajtmh.16-0209}.

\hypertarget{ref-Vecchi_2013}{}
Vecchi, Veronica, Mark Hellowell, and Stefano Gatti. 2013. ``Does the
Private Sector Receive an Excessive Return from Investments in Health
Care Infrastructure Projects? Evidence from the Uk.'' \emph{Health
Policy} 110 (2-3). Elsevier BV: 243--70.
doi:\href{https://doi.org/10.1016/j.healthpol.2012.12.010}{10.1016/j.healthpol.2012.12.010}.

\hypertarget{ref-White_2011}{}
White, Michael T, Lesong Conteh, Richard Cibulskis, and Azra C Ghani.
2011. ``Costs and Cost-Effectiveness of Malaria Control Interventions -
a Systematic Review.'' \emph{Malaria Journal} 10 (1). Springer Nature:
337.
doi:\href{https://doi.org/10.1186/1475-2875-10-337}{10.1186/1475-2875-10-337}.

\hypertarget{ref-whoprof}{}
WHO. 2015. ``Malaria Profile: Mozambique.'' World Health Organization.
\url{http://www.who.int/malaria/publications/country-profiles/profile_moz_en.pdf}.

\hypertarget{ref-Winkler}{}
Winkler, Deborah. 2013. ``Potential and Actual FDI Spillovers in Global
Value Chains.''

\end{document}

### 2018-01-03 R/make_authors.R

@@ -62,7 +62,7 @@ make_authors <- function(
                             stringsAsFactors = FALSE),
   cat_it = TRUE,
   include_address = TRUE,
-  include_country = FALSE,
+  include_country = TRUE,
   seperator = '\n'){
 
   out <- c()
@@ -94,6 +94,12 @@ make_authors <- function(
                               '}')
       }
     }
+    author_text <- paste0(author_text, 
+                          ifelse(nchar(authors$footnote[i]) > 0,
+                                 paste0(' \\footnote{',
+                                        authors$footnote[i],
+                                        '}'),
+                                 ''))
     out[i] <- author_text
   }
   out <- paste0(out, collapse = seperator)

### 2018-01-03 inst/rmd/paper_prepare.R

@@ -6,166 +6,159 @@ library(brew)
 library(maragra)
 library(knitr)
 library(RColorBrewer)
-
-# Read in data
-ab <- maragra::ab
-ab_panel <- maragra::ab_panel
-bairros <- maragra::bairros
-bes <- maragra::bes
-census <- maragra::census
-clinic <- maragra::clinic
-clinic_agg <- maragra::clinic_agg
-mc <- maragra::mc
-weather <- maragra::weather
-workers <- maragra::workers
-
-# Define function for adding zero
-add_zero <- function (x, n) {
-  x <- as.character(x)
-  adders <- n - nchar(x)
-  adders <- ifelse(adders < 0, 0, adders)
-  for (i in 1:length(x)) {
-    if (!is.na(x[i])) {
-      x[i] <- paste0(paste0(rep("0", adders[i]), collapse = ""),
-                     x[i], collapse = "")
-    }
-  }
-  return(x)
-}
-
-# filter for on-site only
-
-model_data <-
-  ab_panel %>%
-  left_join(irs, by = c('unidade', 'date')) %>%
-  mutate(days_since = days_since %/% 30) %>%
-  mutate(days_since = ifelse(is.na(days_since), 'Never',
-                             ifelse(days_since < 0, 'Before',
-                                    ifelse(days_since >= 12, '12+', as.character(add_zero(days_since, n = 2)))))) %>%
-  mutate(months_since = factor(days_since, levels = unique(c('Never', 'Before', sort(unique(days_since)))))) %>%
-  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) %>%
-  # Add malaria incidence
-  left_join(bes %>%
-              dplyr::select(date, p) %>%
-              dplyr::rename(incidence = p),
-            by = 'date') %>%
-  mutate(season = ifelse(incidence >= median(incidence),
-                         'high',
-                         'low')) %>%
-  mutate(season = factor(season, levels = c('low', 'high'))) %>%
-  # Bring in some information on workers
-  left_join(workers %>%
-              dplyr::select(oracle_number,
-                            permanent_or_temporary,
-                            department,
-                            sex,
-                            date_of_birth,
-                            perm_id,
-                            census_name_match_score) %>%
-              # void the permids of anyone with a match score of greater than 0.2
-              mutate(perm_id = ifelse(census_name_match_score > 0.25,
-                                      NA,
-                                      perm_id))) %>%
[…]

### 2018-01-04 inst/rmd/paper.Rmd
formatting updates

@@
 bibliography: bibliography.bib
-# csl: journal-of-health-economics.csl
+csl: journal-of-health-economics.csl
 ---

### 2018-01-04 inst/rmd/paper.tex
formatting updates

@@
 
-\usepackage[table]{xcolor}
-
+\pagenumbering{gobble}
+\usepackage{longtable}
+\usepackage[utf8]{inputenc}
+\usepackage{changepage}
+\usepackage{graphicx}
+\usepackage{multicol}
+\usepackage{geometry}
+\usepackage{fancyhdr}
+\usepackage{color}
+\usepackage{colortbl}
+\usepackage{color}
 % Font
@@
 \setmainfont{Swift-Regular_43151.ttf}
-\setsansfont{Swift-Regular_43151.ttf}
-% \setsansfont[BoldFont={Swift-Bold_43130.ttf}]{Swift-Bold_43130.ttf}
-\setmonofont{Swift-Regular_43151.ttf}
+\setsansfont[BoldFont={Swift-Bold_43130.ttf}]{Swift-Regular_43151.ttf}
+% \setmonofont{Swift-Regular_43151.ttf}
 \renewcommand{\familydefault}{\sfdefault}
+% \usepackage{fontspec}
+% \setmainfont{Lato-Regular.ttf}
+% \setsansfont[BoldFont={Lato-Bold.ttf}]{Lato-Regular.ttf}
+% \renewcommand{\familydefault}{\sfdefault}
+
+\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
+\let\endchangemargin=\endlist
+\renewcommand{\rmdefault}{ppl}
 
@@
 \usepackage{lipsum}
-\usepackage[utf8]{inputenc}
 
 % \usepackage{todonotes} % for side notes
-\usepackage[colorinlistoftodos]{todonotes} % for side notes
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
+% \usepackage[colorinlistoftodos]{todonotes} % for side notes
+
+\usepackage{xargs}                      % Use more than one optional parameter in a new commands
+\usepackage[dvipsnames, table]{xcolor}  % Coloured text etc.
+% 
+\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
+\newcommandx{\unsure}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
+\newcommandx{\change}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
+\newcommandx{\info}[2][1=]{\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25,bordercolor=OliveGreen,#1]{#2}}
+\newcommandx{\improvement}[2][1=]{\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
+\newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}
 \usepackage{lmodern}
@@
 \fancyfoot{} % Blank out the default footer
-\fancyhead[C]{Effectiveness of private sector malaria control at a large sugar
-  facility in Southern Mozambique}
+\fancyhead[C]{Return on investment of private sector malaria control at a large sugar facility in Southern Mozambique}
 \renewcommand{\thefootnote}{\fnsymbol{footnote}}
@@
 \DeclareTextCommandDefault{\nobreakspace}{\leavevmode\nobreak\ }
-
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
 \usepackage{booktabs}
@@
 Malaria has a large economic impact on endemic societies. By affecting
-saving, investment (Shretta, Avanceña, and Hatefi 2016), risk
-perception, productivity, absenteeism (Nonvignon et al. 2016), human
-capital accumulation (Castel-Branco 2014), mortality, and costs of care
-(Sachs and Malaney 2002), malaria likely has a negative effect on GDP
-and growth (McCarthy, Wolf, and Wu 2000) (Orem et al. 2012). Because of
-the relative affordability of most intereventions and the enormous
-societal costs of malaria, most forms of malaria control are
-cost-effective when a public welfare perspective is assumed, such as
-when a government provides the financing (White et al. 2011) (Purdy et
-al. 2013) (Howard et al. 2017).
+saving, investment (Shretta et al., 2016), risk perception,
+productivity, absenteeism (Nonvignon et al., 2016), human capital
+accumulation (Castel-Branco, 2014), mortality, and costs of care (Sachs
+and Malaney, 2002), malaria likely has a negative effect on GDP and
+growth (McCarthy et al., 2000) (Orem et al., 2012). Because of the
+relative affordability of most intereventions and the enormous societal
+costs of malaria, most forms of malaria control are cost-effective when
+a public welfare perspective is assumed, such as when a government
+provides the financing (White et al., 2011) (Purdy et al., 2013) (Howard
+et al., 2017).
 
@@
 such as an increase in life years adjusted for disability or quality
-(Goodman, Coleman, and Mills 1999) (Shretta, Avanceña, and Hatefi 2016)
-(Lee et al. 2017) (Hanson 2004). Though population-level health is
-certainly of importance to businesses, and improvements in health
-incidentally improve the economy at all levels (Brundtland 1999) (Bloom
-and Canning 2008) (Vecchi, Hellowell, and Gatti 2013), these
-improvements may be too disperse or long-term to incentivize private
-sector involvement in health campaigns.
+(Goodman et al., 1999) (Shretta et al., 2016) (Lee et al., 2017)
+(Hanson, 2004). Though population-level health is certainly of
+importance to businesses, and improvements in health incidentally
+improve the economy at all levels (Brundtland, 1999) (Bloom and Canning,
+2008) (Vecchi et al., 2013), these improvements may be too disperse or
+long-term to incentivize private sector involvement in health campaigns.
 
 100\% of the Mozambican population are at risk of malaria, living in
-what the WHO classifies as a ``high transmission'' area (Moonasar et al.
-2016). Annually, Mozambique has more than 8 million clinical malaria
-cases (an annual incidence of approximately 300 per 1,000 residents),
-with an estimated 14,000 deaths. Malaria accounts for 29\% of all
-deaths, and 42\% of deaths among those under five years of age (INE
-2011). Since 2013, Mozambique has seen a gradual increase in the
-incidence of malaria (Moonasar et al. 2016). 100\% of the malaria in
+what the WHO classifies as a ``high transmission'' area (Moonasar et
+al., 2016). Annually, Mozambique has more than 8 million clinical
+malaria cases (an annual incidence of approximately 300 per 1,000
+residents), with an estimated 14,000 deaths. Malaria accounts for 29\%
+of all deaths, and 42\% of deaths among those under five years of age
+(INE, 2011). Since 2013, Mozambique has seen a gradual increase in the
+incidence of malaria (Moonasar et al., 2016). 100\% of the malaria in
 Mozambique is of the Plasmodium falciparum species, with Anopheles
 funestus, gambiae, and arabiensis as the primary mosquito vectors of the
-disease (WHO 2015).
+disease (WHO, 2015).
 
 A significant sector of the economy in Mozambique is dominated by a full
-large-scale foreing direct investment projects (Robbins and Perkins
+large-scale foreing direct investment projects (Robbins and Perkins,
 2012), and the role of the private sector in health generally, and
@@
 extractive industry firms take up wide swaths of land and employ
-hundreds of thousands (German, Schoneveld, and Mwangi 2013). The
-Mozambican state has encouraged large-scale entreprise with the aim of
-general economic development (Buur, Tembe, and Baloi 2012). And where
-large firms exist, they often take on social roles such as housing and
-health care (Winkler 2013). At times, this role is necessary from a
-purely practical standpoint; in other cases, it is employed under the
-guise of ``corporate social responsibility'' (Azemar and Desbordes
-2009). Regardless of the language used, it is clear that private
-industry plays an important role in public health in Mozambique (Robbins
-and Perkins 2012) (Castel-Branco 2014).
+hundreds of thousands (German et al., 2013). The Mozambican state has
+encouraged large-scale entreprise with the aim of general economic
+development (Buur et al., 2012). And where large firms exist, they often
+take on social roles such as housing and health care (Winkler, 2013). At
+times, this role is necessary from a purely practical standpoint; in
+other cases, it is employed under the guise of ``corporate social
+responsibility'' (Azemar and Desbordes, 2009). Regardless of the
+language used, it is clear that private industry plays an important role
+in public health in Mozambique (Robbins and Perkins, 2012)
+(Castel-Branco, 2014).
 
 Several cases exist of foreign firms engaging in large-scale malaria
-control campaigns (Mouzin and al. 2011) (Han 2015) (Bennett et al. 2017)
-(Kaula, Buyungo, and Opigo 2017). But these studies generally consider
+control campaigns (Mouzin and al., 2011) (Han, 2015) (Bennett et al.,
+2017) (Kaula et al., 2017). But these studies generally consider
 population health as the outcome measure of interest, rather than worker
@@
 and takes into account those costs and benefits which are applicable
-from a public welfare point of view (Shretta et al. 2017); though
+from a public welfare point of view (Shretta et al., 2017); though
 appropriate in most cases (the government or institutions interested in
@@
 Our study adds to the existing literature in several ways.
-\todo{Will add details here.}
+\improvement{Will add details here.}
 
@@
 
-Maragra Açucar SA (henceforth referred to as
-``Maragra'')\todo{This is a comment that will appear in the margin}
+Maragra Açucar SA (henceforth referred to as ``Maragra'').
 
@@
 which covered workers from the district, but not those who migrated from
-other parts of the country (Nhacolo et al. 2006).
+other parts of the country (Nhacolo et al., 2006).
 
@@
 malaria. To estimate this effect, we estimated separate models for
-absence and illness. We employed interrupted time series (Lopez Bernal,
-Cummins, and Gasparrini 2016) and a linear probability approach using
-the following econometric model.
+absence and illness. We employed interrupted time series (Lopez Bernal
+et al., 2016) and a linear probability approach using the following
+econometric model.
 
@@
 
-All data processing and analysis were carried out in R (R Core Team
-2017) and all analysis code is freely available online (Brew 2017).
+All data processing and analysis were carried out in R (R Core Team,
+2017) and all analysis code is freely available online (Brew, 2017).
 Ethical approval for this project was obtained from the Institutional
@@
 \hypertarget{ref-Azemar2009}{}
-Azemar, C., and R. Desbordes. 2009. ``Public Governance, Health and
-Foreign Direct Investment in Sub-Saharan Africa.'' \emph{Journal of
-African Economies} 18 (4). Oxford University Press (OUP): 667--709.
-doi:\href{https://doi.org/10.1093/jae/ejn028}{10.1093/jae/ejn028}.
+Azemar, C., Desbordes, R., 2009. Public governance, health and foreign
+direct investment in sub-saharan africa. Journal of African Economies
+18, 667--709. \url{https://doi.org/10.1093/jae/ejn028}
 
 \hypertarget{ref-Bennett_2017}{}
-Bennett, Adam, Anton L. V. Avanceña, Jennifer Wegbreit, Chris Cotter,
-Kathryn Roberts, and Roly Gosling. 2017. ``Engaging the Private Sector
-in Malaria Surveillance: A Review of Strategies and Recommendations for
-Elimination Settings.'' \emph{Malaria Journal} 16 (1). Springer Nature.
-doi:\href{https://doi.org/10.1186/s12936-017-1901-1}{10.1186/s12936-017-1901-1}.
+Bennett, A., Avanceña, A.L.V., Wegbreit, J., Cotter, C., Roberts, K.,
+Gosling, R., 2017. Engaging the private sector in malaria surveillance:
+A review of strategies and recommendations for elimination settings.
+Malaria Journal 16. \url{https://doi.org/10.1186/s12936-017-1901-1}
 
 \hypertarget{ref-Bloom2008}{}
-Bloom, David, and David Canning. 2008. ``Population Health and Economic
-Growth,'' 1--25.
+Bloom, D., Canning, D., 2008. Population Health and Economic Growth
+1--25.
 
 \hypertarget{ref-brewgit}{}
-Brew, J.R. 2017. ``Malaria and Sugar: An in-Depth Examination of the
-Effect of Malaria Control Activities on the Health and Productivity of
-Maragra Sugarcane Factory Workers.'' \emph{GitHub Repository}.
-\url{https://github.com/joebrew/maragra}; GitHub.
+Brew, J., 2017. Malaria and sugar: An in-depth examination of the effect
+of malaria control activities on the health and productivity of maragra
+sugarcane factory workers. GitHub repository.
 
 \hypertarget{ref-World1999}{}
-Brundtland, Gro Harlem. 1999. ``WHO on Health and Economic
-Productivity'' 25 (2): 396--402.
+Brundtland, G.H., 1999. WHO on Health and Economic Productivity 25,
+396--402.
 
 \hypertarget{ref-Buur2012}{}
-Buur, Lars, Carlota Mondlane Tembe, and Obede Baloi. 2012. ``The White
-Gold: The Role of Government and State in Rehabilitating the Sugar
-Industry in Mozambique.'' \emph{Journal of Development Studies} 48 (3).
-Informa UK Limited: 349--62.
-doi:\href{https://doi.org/10.1080/00220388.2011.635200}{10.1080/00220388.2011.635200}.
+Buur, L., Tembe, C.M., Baloi, O., 2012. The white gold: The role of
+government and state in rehabilitating the sugar industry in mozambique.
+Journal of Development Studies 48, 349--362.
+\url{https://doi.org/10.1080/00220388.2011.635200}
 
 \hypertarget{ref-CastelBranco2014}{}
-Castel-Branco, Carlos Nuno. 2014. ``Growth, Capital Accumulation and
-Economic Porosity in Mozambique: Social Losses, Private Gains.''
-\emph{Review of African Political Economy} 41 (sup1). Informa UK
-Limited: S26--S48.
-doi:\href{https://doi.org/10.1080/03056244.2014.976363}{10.1080/03056244.2014.976363}.
+Castel-Branco, C.N., 2014. Growth, capital accumulation and economic
+porosity in mozambique: Social losses, private gains. Review of African
+Political Economy 41, S26--S48.
+\url{https://doi.org/10.1080/03056244.2014.976363}
 
 \hypertarget{ref-German2013}{}
-German, Laura, George Schoneveld, and Esther Mwangi. 2013.
-``Contemporary Processes of Large-Scale Land Acquisition in Sub-Saharan
-Africa: Legal Deficiency or Elite Capture of the Rule of Law?''
-\emph{World Development} 48 (August). Elsevier BV: 1--18.
-doi:\href{https://doi.org/10.1016/j.worlddev.2013.03.006}{10.1016/j.worlddev.2013.03.006}.
+German, L., Schoneveld, G., Mwangi, E., 2013. Contemporary processes of
+large-scale land acquisition in sub-saharan africa: Legal deficiency or
+elite capture of the rule of law? World Development 48, 1--18.
+\url{https://doi.org/10.1016/j.worlddev.2013.03.006}
 
 \hypertarget{ref-Goodman1999}{}
-Goodman, CA, PG Coleman, and AJ Mills. 1999. ``Cost-Effectiveness of
-Malaria Control in Sub-Saharan Africa.'' \emph{The Lancet} 354 (9176).
-Elsevier BV: 378--85.
-doi:\href{https://doi.org/10.1016/s0140-6736(99)02141-8}{10.1016/s0140-6736(99)02141-8}.
+Goodman, C., Coleman, P., Mills, A., 1999. Cost-effectiveness of malaria
+control in sub-saharan africa. The Lancet 354, 378--385.
+\url{https://doi.org/10.1016/s0140-6736(99)02141-8}
 
 \hypertarget{ref-Han}{}
-Han, Lily. 2015. ``Malaria in Mozambique: trialling payment by
-results.''
-\url{http://www.theguardian.com/global-development-professionals-network/2014/mar/31/malaria-control-payment-by-results}.
+Han, L., 2015. Malaria in Mozambique: trialling payment by results.
 
 \hypertarget{ref-Hanson2004}{}
-Hanson, K. 2004. ``Public and Private Roles in Malaria Control: The
-Contributions of Economic Analysis.'' \emph{The American Journal of
-Tropical Medicine and Hygiene} 71 (9176). The American Society of
-Tropical Medicine; Hygiene: 168--73.
-\url{http://www.ajtmh.org/docserver/fulltext/14761645/71/2_suppl/0700168.pdf?expires=1503070750\&id=id\&accname=guest\&checksum=F969ACD7BE7E3A5D929D87EEA8A6EA2B}.
+Hanson, K., 2004. Public and private roles in malaria control: The
+contributions of economic analysis. The American Journal of Tropical
+Medicine and Hygiene 71, 168--173.
 
 \hypertarget{ref-Howard_2017}{}
-Howard, Natasha, Lorna Guinness, Mark Rowland, Naeem Durrani, and
-Kristian S. Hansen. 2017. ``Cost-Effectiveness of Adding Indoor Residual
-Spraying to Case Management in Afghan Refugee Settlements in Northwest
-Pakistan During a Prolonged Malaria Epidemic.'' Edited by Guilherme
-S.Editor Ribeiro. \emph{PLOS Neglected Tropical Diseases} 11 (10).
-Public Library of Science (PLoS): e0005935.
-doi:\href{https://doi.org/10.1371/journal.pntd.0005935}{10.1371/journal.pntd.0005935}.
+Howard, N., Guinness, L., Rowland, M., Durrani, N., Hansen, K.S., 2017.
+Cost-effectiveness of adding indoor residual spraying to case management
+in afghan refugee settlements in northwest pakistan during a prolonged
+malaria epidemic. PLOS Neglected Tropical Diseases 11, e0005935.
+\url{https://doi.org/10.1371/journal.pntd.0005935}
 
 \hypertarget{ref-estatistica2009}{}
-INE. 2011. ``Demographic Health Survey.'' \url{http://www.ine.gov.mz/}.
+INE, 2011. Demographic health survey.
 
 \hypertarget{ref-Kaula_2017}{}
-Kaula, Henry, Peter Buyungo, and Jimmy Opigo. 2017. ``Private Sector
-Role, Readiness and Performance for Malaria Case Management in Uganda,
-2015.'' \emph{Malaria Journal} 16 (1). Springer Nature.
-doi:\href{https://doi.org/10.1186/s12936-017-1824-x}{10.1186/s12936-017-1824-x}.
+Kaula, H., Buyungo, P., Opigo, J., 2017. Private sector role, readiness
+and performance for malaria case management in uganda, 2015. Malaria
+Journal 16. \url{https://doi.org/10.1186/s12936-017-1824-x}
 
 \hypertarget{ref-Lee2017}{}
-Lee, Bruce Y., Eli Zenkov, Chandrani Chatterjee, Baltazar Candrinho,
-Shufang Zhang, James Colborn, Olivier J. T. Briët, et al. 2017. ``The
-Economic Value of Long-Lasting Insecticidal Nets and Indoor Residual
-Spraying Implementation in Mozambique.'' \emph{The American Journal of
-Tropical Medicine and Hygiene} 96 (6). American Society of Tropical
-Medicine; Hygiene: 1430--40.
-doi:\href{https://doi.org/10.4269/ajtmh.16-0744}{10.4269/ajtmh.16-0744}.
+Lee, B.Y., Zenkov, E., Chatterjee, C., Candrinho, B., Zhang, S.,
+Colborn, J., Briët, O.J.T., Brown, S.T., Mendis, C., Bartsch, S.M.,
+Viisainen, K., DePasse, J.V., Stone, N.T.B., 2017. The economic value of
+long-lasting insecticidal nets and indoor residual spraying
+implementation in mozambique. The American Journal of Tropical Medicine
+and Hygiene 96, 1430--1440. \url{https://doi.org/10.4269/ajtmh.16-0744}
 
 \hypertarget{ref-Lopez_Bernal_2016}{}
-Lopez Bernal, James, Steven Cummins, and Antonio Gasparrini. 2016.
-``Interrupted Time Series Regression for the Evaluation of Public Health
-Interventions: A Tutorial.'' \emph{International Journal of
-Epidemiology}, June. Oxford University Press (OUP), dyw098.
-doi:\href{https://doi.org/10.1093/ije/dyw098}{10.1093/ije/dyw098}.
+Lopez Bernal, J., Cummins, S., Gasparrini, A., 2016. Interrupted time
+series regression for the evaluation of public health interventions: A
+tutorial. International Journal of Epidemiology dyw098.
+\url{https://doi.org/10.1093/ije/dyw098}
 
 \hypertarget{ref-McCarthy_2000}{}
-McCarthy, Desmond, Holger Wolf, and Yi Wu. 2000. ``The Growth Costs of
-Malaria,'' February. National Bureau of Economic Research.
-doi:\href{https://doi.org/10.3386/w7541}{10.3386/w7541}.
+McCarthy, D., Wolf, H., Wu, Y., 2000. The growth costs of malaria.
+\url{https://doi.org/10.3386/w7541}
 
 \hypertarget{ref-Moonasar_2016}{}
-Moonasar, Devanand, Rajendra Maharaj, Simon Kunene, Baltazar Candrinho,
-Francisco Saute, Nyasatu Ntshalintshali, and Natashia Morris. 2016.
-``Towards Malaria Elimination in the Mosaswa (Mozambique, South Africa
-and Swaziland) Region.'' \emph{Malaria Journal} 15 (1). Springer Nature.
-doi:\href{https://doi.org/10.1186/s12936-016-1470-8}{10.1186/s12936-016-1470-8}.
+Moonasar, D., Maharaj, R., Kunene, S., Candrinho, B., Saute, F.,
+Ntshalintshali, N., Morris, N., 2016. Towards malaria elimination in the
+mosaswa (mozambique, south africa and swaziland) region. Malaria Journal
+15. \url{https://doi.org/10.1186/s12936-016-1470-8}
 
 \hypertarget{ref-Mouzin2011}{}
-Mouzin, Eric, and Et al. 2011. ``Business Investing in Malaria Control:
-Economic Returns and a Healthy Workforce for Africa.'' \emph{Progress \&
-Impact Series}, no. 6.
+Mouzin, E., al., E., 2011. Business Investing in Malaria Control:
+Economic Returns and a Healthy Workforce for Africa. Progress \& Impact
+series.
 
 \hypertarget{ref-Nhacolo_2006}{}
-Nhacolo, Ariel Q, Delino A Nhalungo, Charfudin N Sacoor, John J Aponte,
-Ricardo Thompson, and Pedro Alonso. 2006. ``Levels and Trends of
-Demographic Indices in Southern Rural Mozambique: Evidence from
-Demographic Surveillance in Manhiça District.'' \emph{BMC Public Health}
-6 (1). Springer Nature.
-doi:\href{https://doi.org/10.1186/1471-2458-6-291}{10.1186/1471-2458-6-291}.
+Nhacolo, A.Q., Nhalungo, D.A., Sacoor, C.N., Aponte, J.J., Thompson, R.,
+Alonso, P., 2006. Levels and trends of demographic indices in southern
+rural mozambique: Evidence from demographic surveillance in manhiça
+district. BMC Public Health 6.
+\url{https://doi.org/10.1186/1471-2458-6-291}
 
 \hypertarget{ref-Nonvignon_2016}{}
-Nonvignon, Justice, Genevieve Cecilia Aryeetey, Keziah L. Malm, Samuel
-Agyei Agyemang, Vivian N. A. Aubyn, Nana Yaw Peprah, Constance N.
-Bart-Plange, and Moses Aikins. 2016. ``Economic Burden of Malaria on
-Businesses in Ghana: A Case for Private Sector Investment in Malaria
-Control.'' \emph{Malaria Journal} 15 (1). Springer Nature.
-doi:\href{https://doi.org/10.1186/s12936-016-1506-0}{10.1186/s12936-016-1506-0}.
+Nonvignon, J., Aryeetey, G.C., Malm, K.L., Agyemang, S.A., Aubyn,
+V.N.A., Peprah, N.Y., Bart-Plange, C.N., Aikins, M., 2016. Economic
+burden of malaria on businesses in ghana: A case for private sector
+investment in malaria control. Malaria Journal 15.
+\url{https://doi.org/10.1186/s12936-016-1506-0}
 
 \hypertarget{ref-Orem_2012}{}
-Orem, Juliet, Joses Kirigia, Robert Azairwe, Ibrahim Kasirye, and
-Oladapo Walker. 2012. ``Impact of Malaria Morbidity on Gross Domestic
-Product in Uganda.'' \emph{International Archives of Medicine} 5 (1).
-Springer Nature: 12.
-doi:\href{https://doi.org/10.1186/1755-7682-5-12}{10.1186/1755-7682-5-12}.
+Orem, J., Kirigia, J., Azairwe, R., Kasirye, I., Walker, O., 2012.
+Impact of malaria morbidity on gross domestic product in uganda.
+International Archives of Medicine 5, 12.
+\url{https://doi.org/10.1186/1755-7682-5-12}
 
 \hypertarget{ref-Purdy_2013}{}
-Purdy, Mark, David Rublin, Kuangyi Wei, and Matthew Robinson. 2013.
-``The Economic Case for Combating Malaria.'' \emph{The American Journal
-of Tropical Medicine and Hygiene} 89 (5). American Society of Tropical
-Medicine; Hygiene: 819--23.
-doi:\href{https://doi.org/10.4269/ajtmh.12-0689}{10.4269/ajtmh.12-0689}.
+Purdy, M., Rublin, D., Wei, K., Robinson, M., 2013. The economic case
+for combating malaria. The American Journal of Tropical Medicine and
+Hygiene 89, 819--823. \url{https://doi.org/10.4269/ajtmh.12-0689}
 
 \hypertarget{ref-R}{}
-R Core Team. 2017. \emph{R: A Language and Environment for Statistical
-Computing}. Vienna, Austria: R Foundation for Statistical Computing.
-\url{https://www.R-project.org/}.
+R Core Team, 2017. R: A language and environment for statistical
+computing. R Foundation for Statistical Computing, Vienna, Austria.
 
 \hypertarget{ref-Robbins2012}{}
-Robbins, Glen, and David Perkins. 2012. ``Mining Fdi and Infrastructure
-Development on Africa's East Coast: Examining the Recent Experience of
-Tanzania and Mozambique.'' \emph{Journal of International Development}
-24 (2). Wiley-Blackwell: 220--36.
-doi:\href{https://doi.org/10.1002/jid.2817}{10.1002/jid.2817}.
+Robbins, G., Perkins, D., 2012. Mining fdi and infrastructure
+development on africa's east coast: Examining the recent experience of
+tanzania and mozambique. Journal of International Development 24,
+220--236. \url{https://doi.org/10.1002/jid.2817}
 
 \hypertarget{ref-Sachs2002}{}
-Sachs, Jeffrey, and Pia Malaney. 2002. ``The Economic and Social Burden
-of Malaria.'' \emph{Nature} 415 (6872). Springer Nature: 680--85.
-doi:\href{https://doi.org/10.1038/415680a}{10.1038/415680a}.
+Sachs, J., Malaney, P., 2002. The economic and social burden of malaria.
+Nature 415, 680--685. \url{https://doi.org/10.1038/415680a}
 
 \hypertarget{ref-Shretta2016}{}
-Shretta, Rima, Anton L. V. Avanceña, and Arian Hatefi. 2016. ``The
-Economics of Malaria Control and Elimination: A Systematic Review.''
-\emph{Malaria Journal} 15 (1). Springer Nature.
-doi:\href{https://doi.org/10.1186/s12936-016-1635-5}{10.1186/s12936-016-1635-5}.
+Shretta, R., Avanceña, A.L.V., Hatefi, A., 2016. The economics of
+malaria control and elimination: A systematic review. Malaria Journal
+15. \url{https://doi.org/10.1186/s12936-016-1635-5}
 
 \hypertarget{ref-Shretta_2017}{}
-Shretta, Rima, Ranju Baral, Anton L. V. Avanceña, Katie Fox, Asoka
-Premasiri Dannoruwa, Ravindra Jayanetti, Arumainayagam Jeyakumaran,
-Rasike Hasantha, Lalanthika Peris, and Risintha Premaratne. 2017. ``An
-Investment Case to Prevent the Reintroduction of Malaria in Sri Lanka.''
-\emph{The American Journal of Tropical Medicine and Hygiene}, January.
-American Society of Tropical Medicine; Hygiene, 16--0209.
-doi:\href{https://doi.org/10.4269/ajtmh.16-0209}{10.4269/ajtmh.16-0209}.
+Shretta, R., Baral, R., Avanceña, A.L.V., Fox, K., Dannoruwa, A.P.,
+Jayanetti, R., Jeyakumaran, A., Hasantha, R., Peris, L., Premaratne, R.,
+2017. An investment case to prevent the reintroduction of malaria in sri
+lanka. The American Journal of Tropical Medicine and Hygiene 16--0209.
+\url{https://doi.org/10.4269/ajtmh.16-0209}
 
 \hypertarget{ref-Vecchi_2013}{}
-Vecchi, Veronica, Mark Hellowell, and Stefano Gatti. 2013. ``Does the
-Private Sector Receive an Excessive Return from Investments in Health
-Care Infrastructure Projects? Evidence from the Uk.'' \emph{Health
-Policy} 110 (2-3). Elsevier BV: 243--70.
-doi:\href{https://doi.org/10.1016/j.healthpol.2012.12.010}{10.1016/j.healthpol.2012.12.010}.
+Vecchi, V., Hellowell, M., Gatti, S., 2013. Does the private sector
+receive an excessive return from investments in health care
+infrastructure projects? Evidence from the uk. Health Policy 110,
+243--270. \url{https://doi.org/10.1016/j.healthpol.2012.12.010}
 
 \hypertarget{ref-White_2011}{}
-White, Michael T, Lesong Conteh, Richard Cibulskis, and Azra C Ghani.
-2011. ``Costs and Cost-Effectiveness of Malaria Control Interventions -
-a Systematic Review.'' \emph{Malaria Journal} 10 (1). Springer Nature:
-337.
-doi:\href{https://doi.org/10.1186/1475-2875-10-337}{10.1186/1475-2875-10-337}.
+White, M.T., Conteh, L., Cibulskis, R., Ghani, A.C., 2011. Costs and
+cost-effectiveness of malaria control interventions - a systematic
+review. Malaria Journal 10, 337.
+\url{https://doi.org/10.1186/1475-2875-10-337}
 
 \hypertarget{ref-whoprof}{}
-WHO. 2015. ``Malaria Profile: Mozambique.'' World Health Organization.
-\url{http://www.who.int/malaria/publications/country-profiles/profile_moz_en.pdf}.
+WHO, 2015. Malaria profile: Mozambique.
 
 \hypertarget{ref-Winkler}{}
-Winkler, Deborah. 2013. ``Potential and Actual FDI Spillovers in Global
-Value Chains.''
+Winkler, D., 2013. Potential and Actual FDI Spillovers in Global Value
+Chains.
 

### 2018-03-25 inst/rmd/defunct/internal_march_2018.Rmd
monday presentation

--- 2017-11-27 inst/rmd/defunct/ihmt_presentation_brew_2017.Rmd
@@
 title: "Effectiveness of private sector malaria control: The case of sugarcane workers in Mozambique"
-subtitle: 'IHMT, October, 2017, Lisboa, Portugal economicsofmalaria.com/ihmt.pdf'
+subtitle: 'ISGlobal, Economics Group, Internal Meeting, March, 2018'
 fig_height: 5
@@
   beamer_presentation:
+    keep_tex: true
+    latex_engine: xelatex
     includes:
@@
 library(maragra)
+
 ```
@@
 workers <- maragra::workers
+source('../paper_prepare.R')
 
@@
 
+## Abstract
+
+This paper provides new empirical evidence regarding the return on investment of privately managed malaria control activities (indoor residual spraying with pesticides) on worker absenteeism in Mozambique.
+We analyze 4 years of malaria control and worker health and absenteeism data from a large sugar processing facility in Mozambique.
+We find that the benefits outweight the costs (ie, there is a positive return on investment) even when the consideration of benefits is limited to those directly accrued by the company.
+These findings suggest that the private sector may have an important role to play in malaria control in endemic areas.
+
+## Overview
+
+### Research highlights
+
+- Large, individual-level worker absenteeism data from malaria endemic zone  
+- Quantifies effect of indoor residual spraying on absenteeism  
+- Estimates cost-effectiveness of malaria control from investment standpoint  
+- Results suggest that private sector could play a significant role in malaria elimination
+
+### Keywords
+
+Malaria; Investment; Health; Productivity; Agriculture; Absenteeism
+
 ## Context
 
-- Malaria has a nearly unquantifiably large economic impact.
-- Many channels: fertility, fecundity, saving, investment [@Shretta2016], risk perception, productivity, absenteeism, human capital accumulation [@CastelBranco2014], mortality, costs of care [@Sachs2002].
-- Cost-benefit studies often only consider the costs of an intervention and associated costs of care, without quantifying the societal cost of non-intervention.  
-- In elimination context, scaling-up private sector involvement is very appealing.
+- Malaria has a large economic impact on endemic societies.
+By affecting saving, investment [@Shretta2016], risk perception, productivity, absenteeism [@Nonvignon_2016], human capital accumulation [@CastelBranco2014], mortality, and costs of care [@Sachs2002], malaria likely has a negative effect on GDP and growth [@McCarthy_2000] [@Orem_2012].
+- Because of the relative affordability of most intereventions and the enormous societal costs of malaria, most forms of malaria control are cost-effective when a public welfare perspective is assumed, such as when a government provides the financing [@White_2011] [@Purdy_2013] [@Howard_2017].
 
@@
 
+## What is already known from the private sector perspective?
+
+- Public health interventions targetting malaria - and their corresponding cost-effectiveness evaluations - most often focus on impacts pertaining to public welfare, such as an increase in life years adjusted for disability or quality  [@Goodman1999] [@Shretta2016] [@Lee2017] [@Hanson2004].
+- Though population-level health is certainly of importance to businesses, and improvements in health incidentally improve the economy at all levels [@World1999] [@Bloom2008] [@Vecchi_2013], these improvements may be too disperse or long-term to incentivize private sector involvement in health campaigns.
+
+## Mozambique
+
+- 100% of the Mozambican population are at risk of malaria, living in what the WHO classifies as a "high transmission" area [@Moonasar_2016].   
+- Annually, Mozambique has more than 8 million clinical malaria cases (an annual incidence of approximately 300 per 1,000 residents), with an estimated 14,000 deaths.   
+- Malaria accounts for 29% of all deaths, and 42% of deaths among those under five years of age [@estatistica2009].   
+- Since 2013, Mozambique has seen a gradual increase in the incidence of malaria [@Moonasar_2016].   
+- 100% of the malaria in Mozambique is of the Plasmodium falciparum species, with Anopheles funestus, gambiae, and arabiensis as the primary mosquito vectors of the disease [@whoprof].
+
+## Industry in Mozambique
+
+A significant sector of the economy in Mozambique is dominated by a full large-scale foreing direct investment projects [@Robbins2012], and the role of the private sector in health generally, and malaria specifically, is unequivocally important.
+- Large agriculture and extractive industry firms take up wide swaths of land and employ hundreds of thousands  [@German2013].   
+- The Mozambican state has encouraged large-scale entreprise with the aim of general economic development [@Buur2012].   
+- And where large firms exist, they often take on social roles such as housing and health care [@Winkler] [@Robbins2012] [@CastelBranco2014].
+
 ## Research questions
-
-We can't answer all the previous questions (yet). So we focus on one:
 
@@
 
-## Identification strategy
+## Data collection
 
-- 1 intervention (IRS, time to/from)
-- 2 outcomes (absence and illness, probabilistic / binomial)
-- Many confounders (age, worker type, seasonality, etc.).
+- January 2010 through December 2016.  
+- Data came from four sources: (i) the Human Resources' roster of worker details and absences, (ii) the facility's on-site clinic's medical and laboratory records, (iii) the facility's on-site malaria control program's records pertaining to the dates, chemicals, and location of IRS activities, and (iv) interviews with company employees pertaining to costs, data limitations, etc.
+- Digitization and collection of data took place during the period from March 2016 through May 2017. 
+- Supplementary data from through the Centro de Investigação em Saude de Manhiça's (CISM) [@Nhacolo_2006].
 
-\begin{small}
-\begin{equation}
-\operatorname{Pr}(\text{Outcome} = 1 \mid \text{X}) = \beta_{0} + \beta_{1} \text{Location} + \beta_{2} \text{Season} + (\beta_3{IRS}*\beta_4{IRS_t} + ... )
-\end{equation}
-\end{small}
+## Epidemiological data
 
-## Modeling
+- District-wide malaria incidence was obtained from Mozambique’s Boletim Epidemiológico Semanal (BES)   
+- Weather data for all Mozambican stations from NOAA.
 
-We employ two approaches:
+## Comment on intervention target groups
 
-1.
-**Propensity score matching** of workers who _ever_ received IRS with workers who _never_ received IRS.
-**Advantage**: No need to adjust for confounders with a matched sample, thereby avoiding reduction in degrees of freedom
+Maragra regularly employs IRS at on-site worker households in order to reduce those workers' (and their families') risk of malaria infection.
+Workers living off-site (our control group) also may have received IRS at some point during the study period (from government programs).
+Even though we do not have reliable person-level data on IRS carried out by the government, off-site workers are a suitable control in the sense that they represent "business as usual" (ie, what would happen if the company carried out no IRS and relied solely on public interventions).
+Using company HR and clinical records, we were able to identify absences and episodes of clinical malaria among all workers, as well as identify the time since the most recent IRS episode before the onsent of absence or illness.
 
-2.
-**Regresion-discontinuity** of only those workers who _ever_ received IRS (ie, ignoring those who never received IRS).
-**Advantage**: Those who _never_ received IRS may be qualitatively different, and therefore not an appropriate comparison group.
+## Conceptual framework and identification strategy
 
-## Propensity score matching {.allowframebreaks}
+We sought to understand the effect of IRS on individual workers' likelihood of absence from work as well as their likelihood of clinical malaria.
+To estimate this effect, we estimated separate models for absence and illness.
+We employed interrupted time series [@Lopez_Bernal_2016] and a linear probability approach using the following econometric model.
 
-- We generate a matched sample of similar workers by first estimating the likelihood of having ever received the intervention, given a worker's age, sex, department and temporary vs.
-permanent status.
- - This is necessary due to below, significant differences:
+$$
+\hat{Y_{it}} = \beta_{0} +  (\beta_{1}) (\text{Season}_{t}) + (\beta_2{IRS}*\beta_3{IRS_t}) + ... + \epsilon
+$$
 
-\begin{tiny}
+$\hat{Y}$ is the rate of absence.
+$\beta_{1}$ represents the clinical malaria incidence at that time in the entire district of Manhiça.
+Our demographic confounders (represented by $...$) are sex, age, and worker department (field, factory, or administrative).
 
-```{r}
-kable(table1[,1:3],  
-      align = 'c', 
-      caption = 'Comparison of unmatched samples') %>%
-  kableExtra::kable_styling(full_width = FALSE)
-```
+## More model details
 
-\end{tiny}
+- Intervention not a simple yes/no: ($_t$) represents time elapsed since commencement of the most recent IRS campaign).   
+- "Malaria season": clinical incidence of malaria in the district of Manhiça was at or greater than the median clinical incidence of malaria for the entire study period.   
+- By using clinical incidence of the area of residence of the workers (as opposed to more typical proxies for malaria risk, such as only rainy vs.
+non rainy season), our seasonality estimate is a closer approximation of true malaria risk, incorporating lagged effects such as the incubation period of the parasite, as well as any inherent non-linear effects of weather.
 
-We match, employing the nearest neighbor method for identifying those workers from our control group who most resemble those workers in the treatment group. [@Ho_Imai_King_Stuart_2007].
+## Estimating return on investment
 
-- Our match is a 1-to-1 cut, meaning those control workers who do not resemble those in the treatment group are left out of primary analysis. The below table shows the match results.
+Our formula for return on investment can be described in a straightforward fashion...
 
-\begin{tiny}
+\begin{center}
+$R = \dfrac{P_{w} - S_{wa} - S_{wc}}{P_{w}}$
 
-```{r, results = 'asis'}
-kable(a$nn, digits = 2, align = 'c', 
-      caption = 'Sample sizes')
-```
+\end{center}
 
-\end{tiny}
+...where $R$ is the return on investment, $P$ is the malaria control program's total operating cost, $w$ refers to costs at the per-worker level, $a$ refers to savings through avoided absences, and $ c $ refers to savings through avoided clinical encounters.
+We define the malaria control program as "profitable" from an investment standpoint if ROI is greater than 100%, ie if the savings associated with the estimated effect of IRS is greater than the costs of the program's administration.
 
-The distributions of our numeric variables are now extremely similar:
+## Reproducibility and ethical approval
 
-\begin{tiny}
-
-```{r, results = 'asis'}
-kable(a$sum.matched[c(1,2,3,4)], digits = 2, align = 'c', 
-      caption = 'Summary of balance for matched data') 
-```
-
-\end{tiny}
-
-## Regression discontinuity analysis
-
-- We simply only consider those workers who _ever_ got IRS.
-- We take into account one full year prior to IRS and one full year after IRS.
-- Our dataset constitutes one observation per worker-day. 
-- We incidentally achieve a sort of "matching" through the fact that workers are their own controls.
+All data processing and analysis were carried out in R [@R] and all analysis code is freely available online [@brewgit].
+Ethical approval for this project was obtained from the Institutional Ethics Review Board for Health at the CISM prior to data collection.
 
 # Results
+
+## Descriptive overview
+
+```{r, fig.cap = 'Clinical malaria (district of Manhiça), all-cause absenteeism among Maragra workers, sick absenteeism among Maragra workers, estimated rainfall, positive cases at company clinic, and test positivity rate at company clinic', fig.height=5, fig.width=8}
+# BES incidence
+g1 <- 
+  ggplot(data = bes %>%
+           mutate(season = ifelse(p >= median(p), 'High season',
+                                  'Low season')) %>%
+           filter(weekdays(date) == 'Monday'),
+       aes(x = date,
+           y = p)) +
+  geom_point(alpha = 0.6,
+             aes(color = season),
+             size = 0.2) +
+  geom_line(alpha = 0.1) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Incidence',
+       title = 'A: Annualized district clinical\nmalaria incidence per 1,000') +
+  scale_color_manual(name = '',
+                     values = c('red', 'black')) + 
+  guides(color = guide_legend(ncol = 1)) +
+  theme(legend.position = c(.8, .8),
+        legend.text = element_text(size = 5),
+        legend.background = element_rect(fill=alpha('white', 0)))
+
+# Absenteeism rate
+g2_data <- ab_panel %>%
+  group_by(month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+  summarise(absences = length(which(absent)),
+            absences_sick = length(which(absent_sick)),
+
+            eligibles = n()) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100,
+         sick_absenteeism_rate = absences_sick / eligibles * 100)
+
+g2 <- ggplot(data = g2_data,
+             aes(x = month,
+                 y = absenteeism_rate)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Absenteeism rate',
+       title = 'B: Monthly absenteeism rate')
+
+g3 <- ggplot(data = g2_data,
+             aes(x = month,
+                 y = sick_absenteeism_rate)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Absenteeism rate',
+       title = 'C: Monthly sick absenteeism rate')
+
+# Weather
+g4 <-
+  ggplot(data = weather %>%
+           group_by(date = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+           summarise(precipitation = sum(precipitation, na.rm = TRUE)),
+         aes(x = date,
+             y = precipitation)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Mililiters',
+       title = 'D: Monthly precipitation')
+
+# Clinical incidence of malaria
+g5 <- clinic_agg %>%
+  filter(date >= '2013-01-01', date <= '2016-12-31') %>%
+  # mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
+  group_by(date) %>%
+  summarise(tested = sum(tested),
+            positive = sum(positive)) %>%
+  ungroup %>%
+  mutate(percent_positive = positive / tested * 100) %>%
+  ggplot(aes(x = date,
+             y = positive)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Cases',
+       title = 'E: Monthly malaria cases\nat company clinic')
+
+g6 <- clinic_agg %>%
+  filter(date >= '2013-01-01', date <= '2016-12-31') %>%
+  # mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
+  group_by(date) %>%
+  summarise(tested = sum(tested),
+            positive = sum(positive)) %>%
+  ungroup %>%
+  mutate(percent_positive = positive / tested * 100) %>%
+  ggplot(aes(x = date,
+             y = percent_positive)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Cases',
+       title = 'F: Monthly malaria test positivity rate\nat company clinic') 
+Rmisc::multiplot(g1, g2, g3, g4, g5, g6, layout = matrix(1:6, nrow=2, byrow=TRUE)) 
+```
 
@@
 
-<!-- ## Descriptive: absenteeism in binned time period -->
-
-<!-- ```{r} -->
-<!-- x <- model_data %>% -->
-<!--   group_by(days_since) %>% -->
-<!--   summarise(absences = length(which(absent)), -->
-<!--             eligibles = length(absent)) %>% -->
-<!--   ungroup %>% -->
-<!--   mutate(absenteeism_rate = absences / eligibles * 100) -->
-
-<!-- ggplot(data = x, -->
-<!--        aes(x = days_since, -->
-<!--            y = absenteeism_rate)) + -->
-<!--   geom_bar(stat = 'identity', -->
-<!--            alpha = 0.6, -->
-<!--            fill = 'darkblue') + -->
-<!--   theme_maragra() + -->
-<!--   labs(x = 'Days since IRS', -->
-<!--        y = 'Absenteeism rate', -->
-<!--        title = 'Raw absenteeism as a function of days since IRS') + -->
-<!--   geom_label(aes(label = paste0(round(absenteeism_rate, digits = 2), '%'))) -->
-<!-- ``` -->
-
-<!-- ## Descriptive: absenteeism in binned time periods by season -->
-
-<!-- ```{r} -->
-<!-- x <- model_data %>% -->
-<!--   group_by(days_since, time_period) %>% -->
-<!--   summarise(absences = length(which(absent)), -->
-<!--             eligibles = length(absent)) %>% -->
-<!--   ungroup %>% -->
-<!--   mutate(absenteeism_rate = absences / eligibles * 100) -->
-
-<!-- ggplot(data = x %>% mutate(time_period = paste0(time_period)), -->
-<!--        aes(x = days_since, -->
-<!--            y = absenteeism_rate)) + -->
-<!--   geom_bar(stat = 'identity', -->
-<!--            alpha = 0.6, -->
-<!--            fill = 'darkblue') + -->
-<!--   theme_maragra() + -->
-<!--   labs(x = 'Days since IRS', -->
-<!--        y = 'Absenteeism rate', -->
-<!--        title = 'Raw absenteeism as a function of days since IRS by time_period') + -->
-<!--   # geom_label(aes(label = paste0(round(absenteeism_rate, digits = 2), '%'))) + -->
-<!--   facet_wrap(~time_period, ncol = 1) + -->
-<!--   theme(axis.text.x = element_text(angle = 90)) -->
-<!-- ``` -->
-
-## Modeling after matching I
-
-- Matched sample of `r nrow(matched)` workers, of which 50% received IRS and 50% did not.   
-- Model only takes into account seasonality, since matching theoretically handles other differences.  
-- For the purposes of this first pass, we "bin" IRS exposure and estimate a logit model to calculate odds ratios.
-
-## Modeling after matching II
-
-All absence:
-
-\begin{tiny}
-
-```{r, results = 'asis'}
-# Combine into a df
-kable(ors_propensity) 
-```
-
-\end{tiny}
-
-## Modeling after matching III
-
-```{r}
-visualize_ors <- function(ors_object){
-  ors_object <- ors_object %>% 
-    filter(grepl('days_since', Variable),
-           grepl('time_periodMalaria', Variable))
-  ors_object <- ors_object %>%
-    mutate(Variable = gsub('days_since', '', Variable),
-           Variable = gsub(':time_periodMalaria season', '', Variable))
-  ggplot(data = ors_object,
-         aes(x = Variable,Variable,
-             y = OR)) +
-    geom_point(alpha = 0.7) +
-    geom_errorbar(aes(ymin = Lower,
-                       ymax = Upper),
-                  alpha = 0.6,
-                  size = 0.5) +
-    geom_hline(yintercept = 1, lty = 2, alpha = 0.6, color = 'red') +
-    labs(x = 'Days since IRS',
-         y = 'OR (relative to pre-IRS period)',
-         title = 'ORs during malaria season') +
-    theme_maragra() 
-}
-visualize_ors(ors_object = ors_propensity)
-```
-
-<!-- \vfill -->
-<!-- \newpage -->
-
-<!-- Sick absence only: -->
-
-<!-- \begin{tiny} -->
-
-<!-- ```{r} -->
-<!-- kable(ors_propensity_sick)  -->
-<!-- ``` -->
-
-<!-- \end{tiny} -->
-
-## Ever IRS'ers compared with never-IRS'ers
+## Descriptive: absenteeism in binned time period
 
@@
   mutate(absenteeism_rate = absences / eligibles * 100)
-y <- ab_panel %>%
-  # filter(oracle_number %in% matched$oracle_number) %>%
-  filter(!unidade %in% sort(irs$unidade)) %>%
-  summarise(days_since = 'Never',
-            absences = length(which(absent)),
-            eligibles = n()) %>%
-  mutate(absenteeism_rate = absences / eligibles * 100)
-x <- bind_rows(x, y)
-x$days_since <- factor(x$days_since,
-                       levels = unique(c('Never',
-                                  'Before',
-                                  sort(unique(x$days_since)))))
-cols <- c('red', rep('blue', nrow(x) - 1))
+
 ggplot(data = x,
@@
            alpha = 0.6,
-           fill = cols) +
+           fill = 'darkblue') +
   theme_maragra() +
@@
 
-## Regression discontinuity analysis {.allowframebreaks}
-
-\vspace{2mm}
-
-All absenteeism:
-
-\begin{tiny}
+## Descriptive: absenteeism in binned time periods by season
 
 ```{r}
-kable(ors) 
+x <- model_data %>%
+  group_by(days_since, time_period) %>%
+  summarise(absences = length(which(absent)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100)
+
+ggplot(data = x %>% mutate(time_period = paste0(time_period)),
+       aes(x = days_since,
+           y = absenteeism_rate)) +
+  geom_bar(stat = 'identity',
+           alpha = 0.6,
+           fill = 'darkblue') +
+  theme_maragra() +
+  labs(x = 'Days since IRS',
+       y = 'Absenteeism rate',
+       title = 'Raw absenteeism as a function of days since IRS by time_period') +
+  # geom_label(aes(label = paste0(round(absenteeism_rate, digits = 2), '%'))) +
+  facet_wrap(~time_period, ncol = 1) +
+  theme(axis.text.x = element_text(angle = 90))
 ```
 
-\end{tiny}
+## Costs
 
-\vfill
-\newpage
+The malaria control program at Maragra has an annual operating budget of approximately XX, which includes the purchase of insecticide, the wages of IRS sprayers and drivers, transportation, record-keeping, and general administrative costs.
+Assuming linearity in costs, the program spends approximately XX per agregado sprayed.
+With each agregado containing an average of 2.2 workers, this translates to a cost of XX per worker protected per season.
+Much of the benefit of IRS goes to non-worker residents of sprayed agregados (who constitute a majority), but this benefit is purposefully ignored for this analysis.
 
-```{r}
-visualize_ors(ors_object = ors)
-```
+## Savings
 
-<!-- \vfill -->
-<!-- \newpage -->
+Given the likelihood that clinical data does not fully capture all malaria cases, we do not quantify the costs of malaria infection to the company.
+Rather, we first estimate the reduction in absenteeism attributible to IRS, and then quantify the savings associated with prevented absences.
+Additionally, we calculate the clinical savings of IRS by first estimating the share of absences which are associated with an episode of clinical malaria, and then applying the clinical cost per case to the equivalent share of prevented absences.
+We intentionally ignore the savings accrued by the public health system, as well as the likely utility gains in secondary realms such as school absenteeism, producitivity, etc.
 
-<!-- Sick absenteeism only: -->
+## Effect of IRS on absenteeism
 
-<!-- \begin{tiny} -->
+- IRS is associated with a year-long, significant reduction in absenteeism during the malaria season.   
+- As one would expect if the mechanism by which IRS reduces absence is through reduced malaria infection, the effect of IRS during the low transmission season is significant, but far less substantial in effect size.
 
-<!-- ```{r} -->
-<!-- kable(ors_sick)  -->
-<!-- ``` -->
+## Fixed effects models
 
-<!-- \end{tiny} -->
+We create 4 worker fixed effects models. Different models for field vs not field, permanent vs temporary.
 
-## Back of the envelope calculations
+## Fixed effects models (1)
+
+\includegraphics{images/fe1.png}
+
+## Fixed effects models (2)
+
+\includegraphics{images/fe2.png}
+
+## Fixed effects models (3)
+
+\includegraphics{images/fe3.png}
+
+## Fixed effects models (4)
+
+\includegraphics{images/fe4.png}
+
+## Return on investment
+
+Details here on ROI calculation outcomes to go here.
+
+## Back of envelope
 
@@
 x <- model_data %>%
+    mutate(time_period = make_season(date = date)) %>%
   filter(time_period == 'Malaria season') %>%
-  group_by(days_since) %>%
+  group_by(months_since = days_since) %>%
   summarise(absences = length(which(absent)),
@@
 
-Your "pre-publication peer review" comments are appreciated:
-
 **Email:** joe@economicsofmalaria.com
 
-**Presentation:** [economicsofmalaria.com/ihmt.pdf](http://economicsofmalaria.com/ihmt.pdf)
+**Paper:** [economicsofmalaria.com/maragra.pdf](http://economicsofmalaria.com/maragra.pdf)
+
+**Presentation:** [economicsofmalaria.com/maragraslides.pdf](http://economicsofmalaria.com/maragraslides.pdf)
 

### 2018-03-25 inst/rmd/defunct/internal_march_2018.tex
monday presentation

\documentclass[ignorenonframetext,]{beamer}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
\usetheme[]{Berlin}
\usecolortheme{seagull}
\usefonttheme{structuresmallcapsserif}
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\newif\ifbibliography
\hypersetup{
            pdftitle={Effectiveness of private sector malaria control: The case of sugarcane workers in Mozambique},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight0.8\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

% Prevent slide breaks in the middle of a paragraph:
\widowpenalties 1 10000
\raggedbottom

\AtBeginPart{
  \let\insertpartnumber\relax
  \let\partname\relax
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \let\insertsectionnumber\relax
    \let\sectionname\relax
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \let\insertsubsectionnumber\relax
  \let\subsectionname\relax
  \frame{\subsectionpage}
}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
\usepackage{multicol}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage[utf8]{inputenc}
% \usepackage[T1]{fontenc}
\usepackage{lmodern}
\pagenumbering{gobble}
\usepackage{longtable}
\usepackage{changepage}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{geometry}
\usepackage{color}
\usepackage{colortbl}
\usepackage{color}
% Font
\usepackage{fontspec}
\setmainfont{Swift-Regular_43151.ttf}
\setsansfont[BoldFont={Swift-Bold_43130.ttf}]{Swift-Regular_43151.ttf}
% \setmonofont{Swift-Regular_43151.ttf}
\renewcommand{\familydefault}{\sfdefault}
% \usepackage{fontspec}
% \setmainfont{Lato-Regular.ttf}
% \setsansfont[BoldFont={Lato-Bold.ttf}]{Lato-Regular.ttf}
% \renewcommand{\familydefault}{\sfdefault}

\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist
\renewcommand{\rmdefault}{ppl}

\usepackage{multicol}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{lipsum}

% \usepackage{todonotes} % for side notes
% \usepackage[colorinlistoftodos]{todonotes} % for side notes

\usepackage{xargs}                      % Use more than one optional parameter in a new commands
% \usepackage[dvipsnames, table]% Coloured text etc.
% 
% \usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
% \newcommandx{\unsure}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
% \newcommandx{\change}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
% \newcommandx{\info}[2][1=]{\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25,bordercolor=OliveGreen,#1]{#2}}
% \newcommandx{\improvement}[2][1=]{\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
% \newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}
% \usepackage{lmodern}

% \newcommand{\footremember}[2]{%
%     \footnote{#2}
%     \newcounter{#1}
%     \setcounter{#1}{\value{footnote}}%
% }
% \newcommand{\footrecall}[1]{%
%     \footnotemark[\value{#1}]%
% }

\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist

\widowpenalties 1 150

\makeatletter
\renewcommand\footnotesize{%
   \@setfontsize\footnotesize\@ixpt{11}%
   \abovedisplayskip 8\p@ \@plus2\p@ \@minus4\p@
   \abovedisplayshortskip \z@ \@plus\p@
   \belowdisplayshortskip 4\p@ \@plus2\p@ \@minus2\p@
   \def\@listi{\leftmargin\leftmargini
               \topsep 4\p@ \@plus2\p@ \@minus2\p@
               \parsep 2\p@ \@plus\p@ \@minus\p@
               \itemsep \parsep}%
   \belowdisplayskip \abovedisplayskip
}
\makeatother

% \DeclareTextCommandDefault{\nobreakspace}{\leavevmode\nobreak\ }

\newcommand{\footremember}[2]{%
    \footnote{#2}
    \newcounter{#1}
    \setcounter{#1}{\value{footnote}}%
}
\newcommand{\footrecall}[1]{%
    \footnotemark[\value{#1}]%
}

% \def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
% \let\endchangemargin=\endlist

\widowpenalties 1 150

\makeatletter
\renewcommand\footnotesize{%
   \@setfontsize\footnotesize\@ixpt{7}%
   \abovedisplayskip 8\p@ \@plus2\p@ \@minus4\p@
   \abovedisplayshortskip \z@ \@plus\p@
   \belowdisplayshortskip 4\p@ \@plus2\p@ \@minus2\p@
   \def\@listi{\leftmargin\leftmargini
               \topsep 4\p@ \@plus2\p@ \@minus2\p@
               \parsep 2\p@ \@plus\p@ \@minus\p@
               \itemsep \parsep}%
   \belowdisplayskip \abovedisplayskip
}
\makeatother

\title{Effectiveness of private sector malaria control: The case of sugarcane
workers in Mozambique}
\subtitle{ISGlobal, Economics Group, Internal Meeting, March, 2018}
\date{}

\begin{document}
\frame{\titlepage}

\begin{frame}

Joe
Brew\footremember{isglobal}{Barcelona Institute for Global Health, Spain}\footremember{cism}{Centro de Investigação em Saúde de Manhiça, Mozambique}\footremember{vu}{VU University Amsterdam, Netherlands}
\footnote{Corresponding Author}

Kizito
Gondo\footremember{ma}{Maragra Açucar SA, Subsidiary of Illovo Sugar Ltd, Mozambique}

Elton Dorkin\footrecall{ma}

Eduardo Nhamahanga\footrecall{ma}

Menno
Pradhan\footrecall{vu}\footremember{uva}{University of Amsterdam, Netherlands}

Laia Cirera\footrecall{isglobal}\footrecall{cism}

Ranjeeta Thomas\footremember{icl}{Imperial College London, UK}

Elisa Sicuri\footrecall{isglobal}\footrecall{cism}\footrecall{icl}

\end{frame}

\section{Introduction}\label{introduction}

\begin{frame}{Abstract}

This paper provides new empirical evidence regarding the return on
investment of privately managed malaria control activities (indoor
residual spraying with pesticides) on worker absenteeism in Mozambique.
We analyze 4 years of malaria control and worker health and absenteeism
data from a large sugar processing facility in Mozambique. We find that
the benefits outweight the costs (ie, there is a positive return on
investment) even when the consideration of benefits is limited to those
directly accrued by the company. These findings suggest that the private
sector may have an important role to play in malaria control in endemic
areas.

\end{frame}

\begin{frame}{Overview}

\begin{block}{Research highlights}

\begin{itemize}
\tightlist
\item
  Large, individual-level worker absenteeism data from malaria endemic
  zone\\
\item
  Quantifies effect of indoor residual spraying on absenteeism\\
\item
  Estimates cost-effectiveness of malaria control from investment
  standpoint\\
\item
  Results suggest that private sector could play a significant role in
  malaria elimination
\end{itemize}

\end{block}

\begin{block}{Keywords}

Malaria; Investment; Health; Productivity; Agriculture; Absenteeism

\end{block}

\end{frame}

\begin{frame}{Context}

\begin{itemize}
\tightlist
\item
  Malaria has a large economic impact on endemic societies. By affecting
  saving, investment (Shretta, Avanceña, and Hatefi 2016), risk
  perception, productivity, absenteeism (Nonvignon et al. 2016), human
  capital accumulation (Castel-Branco 2014), mortality, and costs of
  care (Sachs and Malaney 2002), malaria likely has a negative effect on
  GDP and growth (McCarthy, Wolf, and Wu 2000) (Orem et al. 2012).\\
\item
  Because of the relative affordability of most intereventions and the
  enormous societal costs of malaria, most forms of malaria control are
  cost-effective when a public welfare perspective is assumed, such as
  when a government provides the financing (White et al. 2011) (Purdy et
  al. 2013) (Howard et al. 2017).
\end{itemize}

\end{frame}

\begin{frame}{What we already know}

\begin{itemize}
\tightlist
\item
  Malaria is associated with absenteeism in workers (Nonvignon et al.
  2016).
\item
  Malaria has a negative effect on GDP (Orem et al. 2012) and growth
  (McCarthy, Wolf, and Wu 2000).
\item
  Malaria control is cost-effective from the societal/public perspective
  (Purdy et al. 2013).
\item
  Indoor residual spraying (IRS) is cost-effective (Howard et al. 2017),
  (White et al. 2011) \emph{from a public perspective}.
\end{itemize}

\end{frame}

\begin{frame}{What we want to know}

What is the investment case \emph{from the investor's perspective}?

\begin{itemize}
\tightlist
\item
  Is malaria control just good ``corporate social responsibility''? Or
  is it also good business?
\item
  From a purely financial/investment point-of-view, what benefits does a
  private company experience in engaging in malaria control?\\
\item
  What is the short-term \textbf{benefits} of IRS for large companies in
  malaria-endemic regions?
\item
  What are the \textbf{costs} of of carrying out IRS for large
  companies?
\end{itemize}

\end{frame}

\begin{frame}{What is already known from the private sector
perspective?}

\begin{itemize}
\tightlist
\item
  Public health interventions targetting malaria - and their
  corresponding cost-effectiveness evaluations - most often focus on
  impacts pertaining to public welfare, such as an increase in life
  years adjusted for disability or quality (Goodman, Coleman, and Mills
  1999) (Shretta, Avanceña, and Hatefi 2016) (Lee et al. 2017) (Hanson
  2004).\\
\item
  Though population-level health is certainly of importance to
  businesses, and improvements in health incidentally improve the
  economy at all levels (Brundtland 1999) (Bloom and Canning 2008)
  (Vecchi, Hellowell, and Gatti 2013), these improvements may be too
  disperse or long-term to incentivize private sector involvement in
  health campaigns.
\end{itemize}

\end{frame}

\begin{frame}{Mozambique}

\begin{itemize}
\tightlist
\item
  100\% of the Mozambican population are at risk of malaria, living in
  what the WHO classifies as a ``high transmission'' area (Moonasar et
  al. 2016).\\
\item
  Annually, Mozambique has more than 8 million clinical malaria cases
  (an annual incidence of approximately 300 per 1,000 residents), with
  an estimated 14,000 deaths.\\
\item
  Malaria accounts for 29\% of all deaths, and 42\% of deaths among
  those under five years of age (INE 2011).\\
\item
  Since 2013, Mozambique has seen a gradual increase in the incidence of
  malaria (Moonasar et al. 2016).\\
\item
  100\% of the malaria in Mozambique is of the Plasmodium falciparum
  species, with Anopheles funestus, gambiae, and arabiensis as the
  primary mosquito vectors of the disease (WHO 2015).
\end{itemize}

\end{frame}

\begin{frame}{Industry in Mozambique}

A significant sector of the economy in Mozambique is dominated by a full
large-scale foreing direct investment projects (Robbins and Perkins
2012), and the role of the private sector in health generally, and
malaria specifically, is unequivocally important.\\
- Large agriculture and extractive industry firms take up wide swaths of
land and employ hundreds of thousands (German, Schoneveld, and Mwangi
2013).\\
- The Mozambican state has encouraged large-scale entreprise with the
aim of general economic development (Buur, Tembe, and Baloi 2012).\\
- And where large firms exist, they often take on social roles such as
housing and health care (Winkler 2013) (Robbins and Perkins 2012)
(Castel-Branco 2014).

\end{frame}

\begin{frame}{Research questions}

\textbf{What is the short-term effect of IRS on worker absenteeism and
clinical illness among sugarcane workers?}

\end{frame}

\begin{frame}{Research site}

\includegraphics{internal_march_2018_files/figure-beamer/unnamed-chunk-4-1.pdf}

\end{frame}

\begin{frame}{Research Site II}

\includegraphics{images/sat}

\end{frame}

\section{Methods}\label{methods}

\begin{frame}{Data collection}

\begin{itemize}
\tightlist
\item
  January 2010 through December 2016.\\
\item
  Data came from four sources: (i) the Human Resources' roster of worker
  details and absences, (ii) the facility's on-site clinic's medical and
  laboratory records, (iii) the facility's on-site malaria control
  program's records pertaining to the dates, chemicals, and location of
  IRS activities, and (iv) interviews with company employees pertaining
  to costs, data limitations, etc.\\
\item
  Digitization and collection of data took place during the period from
  March 2016 through May 2017.
\item
  Supplementary data from through the Centro de Investigação em Saude de
  Manhiça's (CISM) (Nhacolo et al. 2006).
\end{itemize}

\end{frame}

\begin{frame}{Epidemiological data}

\begin{itemize}
\tightlist
\item
  District-wide malaria incidence was obtained from Mozambique's Boletim
  Epidemiológico Semanal (BES)\\
\item
  Weather data for all Mozambican stations from NOAA.
\end{itemize}

\end{frame}

\begin{frame}{Comment on intervention target groups}

Maragra regularly employs IRS at on-site worker households in order to
reduce those workers' (and their families') risk of malaria infection.
Workers living off-site (our control group) also may have received IRS
at some point during the study period (from government programs). Even
though we do not have reliable person-level data on IRS carried out by
the government, off-site workers are a suitable control in the sense
that they represent ``business as usual'' (ie, what would happen if the
company carried out no IRS and relied solely on public interventions).
Using company HR and clinical records, we were able to identify absences
and episodes of clinical malaria among all workers, as well as identify
the time since the most recent IRS episode before the onsent of absence
or illness.

\end{frame}

\begin{frame}{Conceptual framework and identification strategy}

We sought to understand the effect of IRS on individual workers'
likelihood of absence from work as well as their likelihood of clinical
malaria. To estimate this effect, we estimated separate models for
absence and illness. We employed interrupted time series (Lopez Bernal,
Cummins, and Gasparrini 2016) and a linear probability approach using
the following econometric model.

\[
\hat{Y_{it}} = \beta_{0} +  (\beta_{1}) (\text{Season}_{t}) + (\beta_2{IRS}*\beta_3{IRS_t}) + ... + \epsilon
\]

\(\hat{Y}\) is the rate of absence. \(\beta_{1}\) represents the
clinical malaria incidence at that time in the entire district of
Manhiça. Our demographic confounders (represented by \(...\)) are sex,
age, and worker department (field, factory, or administrative).

\end{frame}

\begin{frame}{More model details}

\begin{itemize}
\tightlist
\item
  Intervention not a simple yes/no: (\(_t\)) represents time elapsed
  since commencement of the most recent IRS campaign).\\
\item
  ``Malaria season'': clinical incidence of malaria in the district of
  Manhiça was at or greater than the median clinical incidence of
  malaria for the entire study period.\\
\item
  By using clinical incidence of the area of residence of the workers
  (as opposed to more typical proxies for malaria risk, such as only
  rainy vs.~non rainy season), our seasonality estimate is a closer
  approximation of true malaria risk, incorporating lagged effects such
  as the incubation period of the parasite, as well as any inherent
  non-linear effects of weather.
\end{itemize}

\end{frame}

\begin{frame}{Estimating return on investment}

Our formula for return on investment can be described in a
straightforward fashion\ldots{}

\begin{center}
$R = \dfrac{P_{w} - S_{wa} - S_{wc}}{P_{w}}$

\end{center}

\ldots{}where \(R\) is the return on investment, \(P\) is the malaria
control program's total operating cost, \(w\) refers to costs at the
per-worker level, \(a\) refers to savings through avoided absences, and
\$ c \$ refers to savings through avoided clinical encounters. We define
the malaria control program as ``profitable'' from an investment
standpoint if ROI is greater than 100\%, ie if the savings associated
with the estimated effect of IRS is greater than the costs of the
program's administration.

\end{frame}

\begin{frame}{Reproducibility and ethical approval}

All data processing and analysis were carried out in R (R Core Team
2017) and all analysis code is freely available online (Brew 2017).
Ethical approval for this project was obtained from the Institutional
Ethics Review Board for Health at the CISM prior to data collection.

\end{frame}

\section{Results}\label{results}

\begin{frame}{Descriptive overview}

\begin{figure}
\centering
\includegraphics{internal_march_2018_files/figure-beamer/unnamed-chunk-5-1.pdf}
\caption{Clinical malaria (district of Manhiça), all-cause absenteeism
among Maragra workers, sick absenteeism among Maragra workers, estimated
rainfall, positive cases at company clinic, and test positivity rate at
company clinic}
\end{figure}

\end{frame}

\begin{frame}{Descriptive: absenteeism by time from/to intervention}

\includegraphics{internal_march_2018_files/figure-beamer/unnamed-chunk-6-1.pdf}

\end{frame}

\begin{frame}{Descriptive: absenteeism by time from/to intervention
(with local regression lines)}

\includegraphics{internal_march_2018_files/figure-beamer/unnamed-chunk-7-1.pdf}

\end{frame}

\begin{frame}{Descriptive: absenteeism by time from/to intervention (by
time period)}

\includegraphics{internal_march_2018_files/figure-beamer/unnamed-chunk-8-1.pdf}

\end{frame}

\begin{frame}{Same chart with local regression lines}

\includegraphics{internal_march_2018_files/figure-beamer/unnamed-chunk-9-1.pdf}

\end{frame}

\begin{frame}{Descriptive: absenteeism in binned time period}

\includegraphics{internal_march_2018_files/figure-beamer/unnamed-chunk-10-1.pdf}

\end{frame}

\begin{frame}{Descriptive: absenteeism in binned time periods by season}

\includegraphics{internal_march_2018_files/figure-beamer/unnamed-chunk-11-1.pdf}

\end{frame}

\begin{frame}{Costs}

The malaria control program at Maragra has an annual operating budget of
approximately XX, which includes the purchase of insecticide, the wages
of IRS sprayers and drivers, transportation, record-keeping, and general
administrative costs. Assuming linearity in costs, the program spends
approximately XX per agregado sprayed. With each agregado containing an
average of 2.2 workers, this translates to a cost of XX per worker
protected per season. Much of the benefit of IRS goes to non-worker
residents of sprayed agregados (who constitute a majority), but this
benefit is purposefully ignored for this analysis.

\end{frame}

\begin{frame}{Savings}

Given the likelihood that clinical data does not fully capture all
malaria cases, we do not quantify the costs of malaria infection to the
company. Rather, we first estimate the reduction in absenteeism
attributible to IRS, and then quantify the savings associated with
prevented absences. Additionally, we calculate the clinical savings of
IRS by first estimating the share of absences which are associated with
an episode of clinical malaria, and then applying the clinical cost per
case to the equivalent share of prevented absences. We intentionally
ignore the savings accrued by the public health system, as well as the
likely utility gains in secondary realms such as school absenteeism,
producitivity, etc.

\end{frame}

\begin{frame}{Effect of IRS on absenteeism}

\begin{itemize}
\tightlist
\item
  IRS is associated with a year-long, significant reduction in
  absenteeism during the malaria season.\\
\item
  As one would expect if the mechanism by which IRS reduces absence is
  through reduced malaria infection, the effect of IRS during the low
  transmission season is significant, but far less substantial in effect
  size.
\end{itemize}

\end{frame}

\begin{frame}{Fixed effects models}

We create 4 worker fixed effects models. Different models for field vs
not field, permanent vs temporary.

\end{frame}

\begin{frame}{Fixed effects models (1)}

\includegraphics{images/fe1.png}

\end{frame}

\begin{frame}{Fixed effects models (2)}

\includegraphics{images/fe2.png}

\end{frame}

\begin{frame}{Fixed effects models (3)}

\includegraphics{images/fe3.png}

\end{frame}

\begin{frame}{Fixed effects models (4)}

\includegraphics{images/fe4.png}

\end{frame}

\begin{frame}{Return on investment}

Details here on ROI calculation outcomes to go here.

\end{frame}

\begin{frame}{Back of envelope}

\textbf{Savings}

\begin{itemize}
\tightlist
\item
  In percentage point terms, reduction from 13\% to 8\%.\\
\item
  5 annually prevented absences per worker.\\
\item
  8,000 workers: 40,000 prevented absences, wage of 3 USD
\item
  TOTAL: 120,000 USD in productivity-only savings.
\end{itemize}

\textbf{Costs}

\begin{itemize}
\tightlist
\item
  8 IRS workers, 1500 USD yearly = 12,000 USD\\
\item
  ACT + DDT: 50,000 USD
\item
  Facilities, vehicles, gas: 50,000 USD\\
\item
  TOTAL: 112,000 USD in IRS-only costs
\end{itemize}

\textbf{7\% ROI} (ignoring clinical costs)

\end{frame}

\section{Discussion}\label{discussion}

\begin{frame}{General}

\begin{itemize}
\tightlist
\item
  30-50\% reduction in absenteeism in the 6 months after IRS during
  malaria season.
\item
  Depending on detailed cost data (pending), IRS may be effective even
  from a purely financial point of view (ie, beyond just ``corporate
  social responsibility'').\\
\item
  Next steps are incorporation of (a) productivity data (via cane cut),
  (b) better clinical data (via local health facilities), and (c) full
  cost data.\\
\item
  Will also be comparing with a sugarcane facility in a zone where an
  elimination campaign is taking place.
\end{itemize}

\end{frame}

\begin{frame}{Limitations}

\begin{itemize}
\tightlist
\item
  No analysis yet of different worker types (agricultural
  vs.~industrial).
\item
  Have not yet ventured at all into side-analyses (effect on employment,
  tonnage, etc.).
\item
  Sick absenteeism seems to track absenteeism poorly: lack of clarity
  regarding pathways.\\
\item
  Large sample size, but all from same place: questionable
  generalizability.
\end{itemize}

\end{frame}

\begin{frame}{Thank you}

\textbf{Email:}
\href{mailto:joe@economicsofmalaria.com}{\nolinkurl{joe@economicsofmalaria.com}}

\textbf{Paper:}
\href{http://economicsofmalaria.com/maragra.pdf}{economicsofmalaria.com/maragra.pdf}

\textbf{Presentation:}
\href{http://economicsofmalaria.com/maragraslides.pdf}{economicsofmalaria.com/maragraslides.pdf}

\end{frame}

\begin{frame}[allowframebreaks]{References}

\hypertarget{refs}{}
\hypertarget{ref-Bloom2008}{}
Bloom, David, and David Canning. 2008. ``Population Health and Economic
Growth,'' 1--25.

\hypertarget{ref-brewgit}{}
Brew, J.R. 2017. ``Malaria and Sugar: An in-Depth Examination of the
Effect of Malaria Control Activities on the Health and Productivity of
Maragra Sugarcane Factory Workers.'' \emph{GitHub Repository}.
\url{https://github.com/joebrew/maragra}; GitHub.

\hypertarget{ref-World1999}{}
Brundtland, Gro Harlem. 1999. ``WHO on Health and Economic
Productivity'' 25 (2): 396--402.

\hypertarget{ref-Buur2012}{}
Buur, Lars, Carlota Mondlane Tembe, and Obede Baloi. 2012. ``The White
Gold: The Role of Government and State in Rehabilitating the Sugar
Industry in Mozambique.'' \emph{Journal of Development Studies} 48 (3).
Informa UK Limited: 349--62.
doi:\href{https://doi.org/10.1080/00220388.2011.635200}{10.1080/00220388.2011.635200}.

\hypertarget{ref-CastelBranco2014}{}
Castel-Branco, Carlos Nuno. 2014. ``Growth, Capital Accumulation and
Economic Porosity in Mozambique: Social Losses, Private Gains.''
\emph{Review of African Political Economy} 41 (sup1). Informa UK
Limited: S26--S48.
doi:\href{https://doi.org/10.1080/03056244.2014.976363}{10.1080/03056244.2014.976363}.

\hypertarget{ref-German2013}{}
German, Laura, George Schoneveld, and Esther Mwangi. 2013.
``Contemporary Processes of Large-Scale Land Acquisition in Sub-Saharan
Africa: Legal Deficiency or Elite Capture of the Rule of Law?''
\emph{World Development} 48 (August). Elsevier BV: 1--18.
doi:\href{https://doi.org/10.1016/j.worlddev.2013.03.006}{10.1016/j.worlddev.2013.03.006}.

\hypertarget{ref-Goodman1999}{}
Goodman, CA, PG Coleman, and AJ Mills. 1999. ``Cost-Effectiveness of
Malaria Control in Sub-Saharan Africa.'' \emph{The Lancet} 354 (9176).
Elsevier BV: 378--85.
doi:\href{https://doi.org/10.1016/s0140-6736(99)02141-8}{10.1016/s0140-6736(99)02141-8}.

\hypertarget{ref-Hanson2004}{}
Hanson, K. 2004. ``Public and Private Roles in Malaria Control: The
Contributions of Economic Analysis.'' \emph{The American Journal of
Tropical Medicine and Hygiene} 71 (9176). The American Society of
Tropical Medicine; Hygiene: 168--73.
\url{http://www.ajtmh.org/docserver/fulltext/14761645/71/2_suppl/0700168.pdf?expires=1503070750\&id=id\&accname=guest\&checksum=F969ACD7BE7E3A5D929D87EEA8A6EA2B}.

\hypertarget{ref-Howard_2017}{}
Howard, Natasha, Lorna Guinness, Mark Rowland, Naeem Durrani, and
Kristian S. Hansen. 2017. ``Cost-Effectiveness of Adding Indoor Residual
Spraying to Case Management in Afghan Refugee Settlements in Northwest
Pakistan During a Prolonged Malaria Epidemic.'' Edited by Guilherme
S.Editor Ribeiro. \emph{PLOS Neglected Tropical Diseases} 11 (10).
Public Library of Science (PLoS): e0005935.
doi:\href{https://doi.org/10.1371/journal.pntd.0005935}{10.1371/journal.pntd.0005935}.

\hypertarget{ref-estatistica2009}{}
INE. 2011. ``Demographic Health Survey.'' \url{http://www.ine.gov.mz/}.

\hypertarget{ref-Lee2017}{}
Lee, Bruce Y., Eli Zenkov, Chandrani Chatterjee, Baltazar Candrinho,
Shufang Zhang, James Colborn, Olivier J. T. Briët, et al. 2017. ``The
Economic Value of Long-Lasting Insecticidal Nets and Indoor Residual
Spraying Implementation in Mozambique.'' \emph{The American Journal of
Tropical Medicine and Hygiene} 96 (6). American Society of Tropical
Medicine; Hygiene: 1430--40.
doi:\href{https://doi.org/10.4269/ajtmh.16-0744}{10.4269/ajtmh.16-0744}.

\hypertarget{ref-Lopez_Bernal_2016}{}
Lopez Bernal, James, Steven Cummins, and Antonio Gasparrini. 2016.
``Interrupted Time Series Regression for the Evaluation of Public Health
Interventions: A Tutorial.'' \emph{International Journal of
Epidemiology}, June. Oxford University Press (OUP), dyw098.
doi:\href{https://doi.org/10.1093/ije/dyw098}{10.1093/ije/dyw098}.

\hypertarget{ref-McCarthy_2000}{}
McCarthy, Desmond, Holger Wolf, and Yi Wu. 2000. ``The Growth Costs of
Malaria,'' February. National Bureau of Economic Research.
doi:\href{https://doi.org/10.3386/w7541}{10.3386/w7541}.

\hypertarget{ref-Moonasar_2016}{}
Moonasar, Devanand, Rajendra Maharaj, Simon Kunene, Baltazar Candrinho,
Francisco Saute, Nyasatu Ntshalintshali, and Natashia Morris. 2016.
``Towards Malaria Elimination in the Mosaswa (Mozambique, South Africa
and Swaziland) Region.'' \emph{Malaria Journal} 15 (1). Springer Nature.
doi:\href{https://doi.org/10.1186/s12936-016-1470-8}{10.1186/s12936-016-1470-8}.

\hypertarget{ref-Nhacolo_2006}{}
Nhacolo, Ariel Q, Delino A Nhalungo, Charfudin N Sacoor, John J Aponte,
Ricardo Thompson, and Pedro Alonso. 2006. ``Levels and Trends of
Demographic Indices in Southern Rural Mozambique: Evidence from
Demographic Surveillance in Manhiça District.'' \emph{BMC Public Health}
6 (1). Springer Nature.
doi:\href{https://doi.org/10.1186/1471-2458-6-291}{10.1186/1471-2458-6-291}.

\hypertarget{ref-Nonvignon_2016}{}
Nonvignon, Justice, Genevieve Cecilia Aryeetey, Keziah L. Malm, Samuel
Agyei Agyemang, Vivian N. A. Aubyn, Nana Yaw Peprah, Constance N.
Bart-Plange, and Moses Aikins. 2016. ``Economic Burden of Malaria on
Businesses in Ghana: A Case for Private Sector Investment in Malaria
Control.'' \emph{Malaria Journal} 15 (1). Springer Nature.
doi:\href{https://doi.org/10.1186/s12936-016-1506-0}{10.1186/s12936-016-1506-0}.

\hypertarget{ref-Orem_2012}{}
Orem, Juliet, Joses Kirigia, Robert Azairwe, Ibrahim Kasirye, and
Oladapo Walker. 2012. ``Impact of Malaria Morbidity on Gross Domestic
Product in Uganda.'' \emph{International Archives of Medicine} 5 (1).
Springer Nature: 12.
doi:\href{https://doi.org/10.1186/1755-7682-5-12}{10.1186/1755-7682-5-12}.

\hypertarget{ref-Purdy_2013}{}
Purdy, Mark, David Rublin, Kuangyi Wei, and Matthew Robinson. 2013.
``The Economic Case for Combating Malaria.'' \emph{The American Journal
of Tropical Medicine and Hygiene} 89 (5). American Society of Tropical
Medicine; Hygiene: 819--23.
doi:\href{https://doi.org/10.4269/ajtmh.12-0689}{10.4269/ajtmh.12-0689}.

\hypertarget{ref-R}{}
R Core Team. 2017. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\hypertarget{ref-Robbins2012}{}
Robbins, Glen, and David Perkins. 2012. ``Mining Fdi and Infrastructure
Development on Africa's East Coast: Examining the Recent Experience of
Tanzania and Mozambique.'' \emph{Journal of International Development}
24 (2). Wiley-Blackwell: 220--36.
doi:\href{https://doi.org/10.1002/jid.2817}{10.1002/jid.2817}.

\hypertarget{ref-Sachs2002}{}
Sachs, Jeffrey, and Pia Malaney. 2002. ``The Economic and Social Burden
of Malaria.'' \emph{Nature} 415 (6872). Springer Nature: 680--85.
doi:\href{https://doi.org/10.1038/415680a}{10.1038/415680a}.

\hypertarget{ref-Shretta2016}{}
Shretta, Rima, Anton L. V. Avanceña, and Arian Hatefi. 2016. ``The
Economics of Malaria Control and Elimination: A Systematic Review.''
\emph{Malaria Journal} 15 (1). Springer Nature.
doi:\href{https://doi.org/10.1186/s12936-016-1635-5}{10.1186/s12936-016-1635-5}.

\hypertarget{ref-Vecchi_2013}{}
Vecchi, Veronica, Mark Hellowell, and Stefano Gatti. 2013. ``Does the
Private Sector Receive an Excessive Return from Investments in Health
Care Infrastructure Projects? Evidence from the Uk.'' \emph{Health
Policy} 110 (2-3). Elsevier BV: 243--70.
doi:\href{https://doi.org/10.1016/j.healthpol.2012.12.010}{10.1016/j.healthpol.2012.12.010}.

\hypertarget{ref-White_2011}{}
White, Michael T, Lesong Conteh, Richard Cibulskis, and Azra C Ghani.
2011. ``Costs and Cost-Effectiveness of Malaria Control Interventions -
a Systematic Review.'' \emph{Malaria Journal} 10 (1). Springer Nature:
337.
doi:\href{https://doi.org/10.1186/1475-2875-10-337}{10.1186/1475-2875-10-337}.

\hypertarget{ref-whoprof}{}
WHO. 2015. ``Malaria Profile: Mozambique.'' World Health Organization.
\url{http://www.who.int/malaria/publications/country-profiles/profile_moz_en.pdf}.

\hypertarget{ref-Winkler}{}
Winkler, Deborah. 2013. ``Potential and Actual FDI Spillovers in Global
Value Chains.''

\end{frame}

\end{document}

### 2018-04-11 inst/rmd/00_title_page.Rmd
major changes

@@
 
-This paper provides new empirical evidence regarding the return on investment of privately managed malaria control activities (indoor residual spraying with pesticides) on worker absenteeism in Mozambique.
+This paper provides new empirical evidence regarding the effect and return on investment of privately managed malaria control activities (indoor residual spraying with pesticides) on worker absenteeism in Mozambique.
 We analyze 4 years of malaria control and worker health and absenteeism data from a large sugar processing facility in Mozambique.

### 2018-04-11 inst/rmd/01_introduction.Rmd
formatting updates
major changes

@@
 
-Malaria has a large economic impact on endemic societies.
+Malaria accounts for a half million annual deaths worldwide [@White2014] [@World2016] [@Ashley2018].
+Though rapid improvements in technology and funding have lead to important reductions in marlaria's global burden, the scale-up in activities required for eradication ("the worldwide interruption of transmission") [@TheLancet2011] will mean new partnerships and actors.
+One promising - albeit atypical - potential stakeholder in global malaria eradication is the private sector, given its omnipresence and potential to benefit directly from the elimination of malaria.
+But little evidence exists demonstrating how private sector entities can engage with malaria control and benefit at the firm level.
+
+At the societal level, malaria has a large economic impact.
 By affecting saving, investment [@Shretta2016], risk perception, productivity, absenteeism [@Nonvignon_2016], human capital accumulation [@CastelBranco2014], mortality, and costs of care [@Sachs2002], malaria likely has a negative effect on GDP and growth [@McCarthy_2000] [@Orem_2012].
@@
 
-A significant sector of the economy in Mozambique is dominated by a full large-scale foreing direct investment projects [@Robbins2012], and the role of the private sector in health generally, and malaria specifically, is unequivocally important.
+A significant sector of the economy in Mozambique is dominated by large-scale foreign direct investment projects [@Robbins2012], and the role of the private sector in health generally, and malaria specifically, is unequivocally important.
 Large agriculture and extractive industry firms take up wide swaths of land and employ hundreds of thousands  [@German2013].
@@
 And where large firms exist, they often take on social roles such as housing and health care [@Winkler].
-At times, this role is necessary from a purely practical standpoint; in other cases, it is employed under the guise of "corporate social responsibility" [@Azemar2009].
+At times, this role is necessary from a purely practical standpoint; in other cases, it is employed under the auspices of "corporate social responsibility" [@Azemar2009].
 Regardless of the language used, it is clear that private industry plays an important role in public health in Mozambique [@Robbins2012] [@CastelBranco2014].
@@
 Similarly, they often neglect to differentiate between those clinical costs which are absorbed by the local health system versus those which are absorbed by the firm itself.
-In the literature, making the "investment case" for malaria control or elimination generally implies that the investor is the public sector, and takes into account those costs and benefits which are applicable from a public welfare point of view [@Shretta_2017]; though appropriate in most cases (the government or institutions interested in public welfare primarily being the primary malaria control agents in most locations), the findings of these studies are rarely applicable to the private sector.
+In the literature, making the "investment case" for malaria control or elimination generally implies that the investor is the public sector, and takes into account those costs and benefits which are applicable from a public welfare point of view [@Shretta_2017]; though appropriate in most cases (the government or institutions interested in public welfare primarily being the primary malaria control agents in most locations), the findings of these studies are rarely applicable to the private sector, and even less so at granular levels (such as an individual firm).
 In the case of a private firm not interested in "corporate social responsibility", it is not clear whether investing in malaria control would be profitable or not.
@@
 
-To address the question of the profitability of malaria control activities from the standpoint of a private firm, we analyze data during a 7 year period from a private sugar facility in Southern Mozambique.
-We assess the effect of indoor residual spraying (IRS) on the absenteeism and health of workers, and demonstrate that the firm's engagement in malaria control not only improved worker health, but also generated a positive return on investment.
+To address the question of the profitability of malaria control activities from the standpoint of a private firm, we analyze data during a 4 year period from a private sugar facility in Southern Mozambique.
+We assess the effect of indoor residual spraying (IRS) on the absenteeism and health of workers, and demonstrate that the firm's engagement in malaria control not only improved worker health, but also generated a positive return on investment from a pure accounting perspective.
 
@@
 After a brief review of the existing literature on the subject of private sector investment in malaria control and its effects on worker productivity, we provide an overview of the sugar company under study, and the epidemiology of malaria in the nearby area, as well as in Mozambique as a whole.
-We then give an overview of the data collected, and outline the theoretical and methdological assumptions that underly our analysis.
+We then give an overview of the data collected, and outline the theoretical and methdological assumptions and tools that underly our analysis.
 In the results, we show the effectiveness of the company's malaria control program in reducing absences, and translate this reduction into cost savings.
-Our robustness checks consist of \todo{Add some stuff here}.
 Our discussion covers potential implications from this study in terms of policy and investment, as well as the paper's limitations.
@@
 
-Our study adds to the existing literature in several ways. \todo{Will add details here.}
+Our study adds to the existing literature in several ways. \improvement{Will add details here.}
 
@@
 
-Details on Maragra:
+Sugar has been systematically cultivated in Mozambique since the late 1800s, but was not exported at a large scale until the 1950s.
+The Incomati Estates company, a small sugarcane processing facility started by a Scotsman on the banks of the Incomati River in 1913, was the first firm to export sugar from Mozambique.
+Following its purchase by international investors in the 1950s, it (along with the rest of the industry) expanded significantly.
+In the late 1960s, a Portuguese family opened the Maragra Açúcar company, while a group of foreign investors started the nearby Marracuene Agrícola Açucareira mill.
+By the early 1970s, sugar grew to account for greater than 10% of Mozambique's national exports.
+Nationalized following independence in 1977, the industry's production levels fell from 320,000 annual tons to fewer than 15,000 in 1992.
+After the end of the civil war, foreign investment revived the sugar industry, and by 2011 production had surpassed its 1972 peak.
 
-Maragra Açucar SA (henceforth referred to as "Maragra")\todo{This is a comment that will appear in the margin}
+The mill of the Maragra Açucar SA (a subsidary of the Illovo sugar company, henceforth referred to as "Maragra") was nationalized in the 1970s (like all other mills), went through a period of low production, and then fell completely out of use by 1984.
+It re-opened in private hands in 1992, and was renovated by a group of international investors in 1998.
+Today, Maragra accounts for roughly one quarter of Mozambique's overall sugar production (second only to the nearby Xinavane mill run by the Hulett Sugar Tongaat company) [@sutton].
+With a favorably close location to the port of Maputo, ample land, approximately 5,000 employees (of which three quarters are seasonal), and a mill with the capacity to process not only all the sugarcane grown on Maragra's land, but also the cane of the many nearby smallholders [@OLaughlin2016], Maragra has so far been able to weather the 2016 Mozambican crisis and concurrent collapse in global sugar prices.
 

### 2018-04-11 inst/rmd/paper.tex
major changes

@@
 
-Malaria has a large economic impact on endemic societies. By affecting
+Malaria accounts for a half million annual deaths worldwide (White et
+al., 2014) (WHO, 2016) (Ashley et al., 2018). Though rapid improvements
+in technology and funding have lead to important reductions in
+marlaria's global burden, the scale-up in activities required for
+eradication (``the worldwide interruption of transmission'') (Lancet,
+2011) will mean new partnerships and actors. One promising - albeit
+atypical - potential stakeholder in global malaria eradication is the
+private sector, given its omnipresence and potential to benefit directly
+from the elimination of malaria. But little evidence exists
+demonstrating how private sector entities can engage with malaria
+control and benefit at the firm level.
+
+At the societal level, malaria has a large economic impact. By affecting
 saving, investment (Shretta et al., 2016), risk perception,
@@
 
-A significant sector of the economy in Mozambique is dominated by a full
-large-scale foreing direct investment projects (Robbins and Perkins,
+A significant sector of the economy in Mozambique is dominated by
+large-scale foreign direct investment projects (Robbins and Perkins,
 2012), and the role of the private sector in health generally, and
@@
 times, this role is necessary from a purely practical standpoint; in
-other cases, it is employed under the guise of ``corporate social
+other cases, it is employed under the auspices of ``corporate social
 responsibility'' (Azemar and Desbordes, 2009). Regardless of the
@@
 most locations), the findings of these studies are rarely applicable to
-the private sector. In the case of a private firm not interested in
+the private sector, and even less so at granular levels (such as an
+individual firm). In the case of a private firm not interested in
 ``corporate social responsibility'', it is not clear whether investing
@@
 activities from the standpoint of a private firm, we analyze data during
-a 7 year period from a private sugar facility in Southern Mozambique. We
+a 4 year period from a private sugar facility in Southern Mozambique. We
 assess the effect of indoor residual spraying (IRS) on the absenteeism
@@
 malaria control not only improved worker health, but also generated a
-positive return on investment.
+positive return on investment from a pure accounting perspective.
 
@@
 give an overview of the data collected, and outline the theoretical and
-methdological assumptions that underly our analysis. In the results, we
-show the effectiveness of the company's malaria control program in
-reducing absences, and translate this reduction into cost savings. Our
-robustness checks consist of \todo{Add some stuff here}. Our discussion
-covers potential implications from this study in terms of policy and
-investment, as well as the paper's limitations.
+methdological assumptions and tools that underly our analysis. In the
+results, we show the effectiveness of the company's malaria control
+program in reducing absences, and translate this reduction into cost
+savings. Our discussion covers potential implications from this study in
+terms of policy and investment, as well as the paper's limitations.
 
@@
 
-Details on Maragra:
+Sugar has been systematically cultivated in Mozambique since the late
+1800s, but was not exported at a large scale until the 1950s. The
+Incomati Estates company, a small sugarcane processing facility started
+by a Scotsman on the banks of the Incomati River in 1913, was the first
+firm to export sugar from Mozambique. Following its purchase by
+international investors in the 1950s, it (along with the rest of the
+industry) expanded significantly. In the late 1960s, a Portuguese family
+opened the Maragra Açúcar company, while a group of foreign investors
+started the nearby Marracuene Agrícola Açucareira mill. By the early
+1970s, sugar grew to account for greater than 10\% of Mozambique's
+national exports. Nationalized following independence in 1977, the
+industry's production levels fell from 320,000 annual tons to fewer than
+15,000 in 1992. After the end of the civil war, foreign investment
+revived the sugar industry, and by 2011 production had surpassed its
+1972 peak.
 
-Maragra Açucar SA (henceforth referred to as ``Maragra'').
+The mill of the Maragra Açucar SA (a subsidary of the Illovo sugar
+company, henceforth referred to as ``Maragra'') was nationalized in the
+1970s (like all other mills), went through a period of low production,
+and then fell completely out of use by 1984. It re-opened in private
+hands in 1992, and was renovated by a group of international investors
+in 1998. Today, Maragra accounts for roughly one quarter of Mozambique's
+overall sugar production (second only to the nearby Xinavane mill run by
+the Hulett Sugar Tongaat company) (Sutton, 2015). With a favorably close
+location to the port of Maputo, ample land, approximately 5,000
+employees (of which three quarters are seasonal), and a mill with the
+capacity to process not only all the sugarcane grown on Maragra's land,
+but also the cane of nearby smallholders, Maragra has so far been able
+to weather the 2016 Mozambican crisis and concurrent collapse in global
+sugar prices.
+
+\begin{itemize}
+\tightlist
+\item
+  Details on Maragra: .
+\end{itemize}
+
+-Xinavane (O'Laughlin, 2016)
 
@@
 \hypertarget{refs}{}
+\hypertarget{ref-Ashley2018}{}
+Ashley, E.A., Phyo, A.P., Woodrow, C.J., 2018. Malaria. The Lancet.
+\url{https://doi.org/10.1016/s0140-6736(18)30324-6}
+
 \hypertarget{ref-Azemar2009}{}
@@
 
+\hypertarget{ref-TheLancet2011}{}
+Lancet, T., 2011. Malaria: Control vs elimination vs eradication. The
+Lancet 378, 1117. \url{https://doi.org/10.1016/s0140-6736(11)61489-x}
+
 \hypertarget{ref-Lee2017}{}
@@
 
+\hypertarget{ref-OLaughlin2016}{}
+O'Laughlin, B., 2016. Consuming bodies: Health and work in the cane
+fields in xinavane, mozambique. Journal of Southern African Studies 43,
+625--641. \url{https://doi.org/10.1080/03057070.2016.1190519}
+
 \hypertarget{ref-Purdy_2013}{}
@@
 
+\hypertarget{ref-sutton}{}
+Sutton, J., 2015. Mapa empresarial de moçambique.
+
 \hypertarget{ref-Vecchi_2013}{}
@@
 
+\hypertarget{ref-White2014}{}
+White, N.J., Pukrittayakamee, S., Hien, T.T., Faiz, M.A., Mokuolu, O.A.,
+Dondorp, A.M., 2014. Malaria. The Lancet 383, 723--735.
+\url{https://doi.org/10.1016/s0140-6736(13)60024-0}
+
+\hypertarget{ref-World2016}{}
+WHO, 2016. World Malaria Report.
+
 \hypertarget{ref-whoprof}{}

### 2018-04-12 inst/rmd/00_title_page.Rmd
major improvements

@@
         \begin{itemize}
-          \item Large, individual-level worker absenteeism data from malaria endemic zone
-          \item Quantifies effect of indoor residual spraying on absenteeism
-          \item Estimates cost-effectiveness of malaria control from investment standpoint
-          \item Results suggest that private sector could play a significant role in malaria elimination
+          \item This paper analyzes large, individual-level worker absenteeism data from malaria endemic zone. 
+          \item We quantify the effect of indoor residual spraying on absenteeism and clinical malaria. 
+          \item We estimate cost-effectiveness of malaria control from an investment standpoint. 
+          \item Results show ledger profitability, suggesting that the private sector could play a significant role in malaria elimination.  
         \end{itemize}

### 2018-04-12 inst/rmd/01_introduction.Rmd
major improvements

@@
 
-Sugar has been systematically cultivated in Mozambique since the late 1800s, but was not exported at a large scale until the 1950s.
+Sugar has been systematically cultivated in Mozambique since the late 1800s.
 The Incomati Estates company, a small sugarcane processing facility started by a Scotsman on the banks of the Incomati River in 1913, was the first firm to export sugar from Mozambique.
-Following its purchase by international investors in the 1950s, it (along with the rest of the industry) expanded significantly.
+Following its purchase by international investors in the 1950s, it (along with the rest of the industry) expanded significantly, exporting to both Europe and the United States.
 In the late 1960s, a Portuguese family opened the Maragra Açúcar company, while a group of foreign investors started the nearby Marracuene Agrícola Açucareira mill.
 By the early 1970s, sugar grew to account for greater than 10% of Mozambique's national exports.
-Nationalized following independence in 1977, the industry's production levels fell from 320,000 annual tons to fewer than 15,000 in 1992.
+Nationalized following independence in 1977, the industry's production levels fell from 320,000 annual tons to fewer than 15,000 by 1992.
 After the end of the civil war, foreign investment revived the sugar industry, and by 2011 production had surpassed its 1972 peak.
 
-The mill of the Maragra Açucar SA (a subsidary of the Illovo sugar company, henceforth referred to as "Maragra") was nationalized in the 1970s (like all other mills), went through a period of low production, and then fell completely out of use by 1984.
+The mill of the Maragra Açucar SA (a subsidary of the Illovo sugar company, henceforth referred to as "Maragra") was nationalized in the 1970s (like all other Mozambican mills), went through a period of low production, and then fell completely out of use by 1984.
 It re-opened in private hands in 1992, and was renovated by a group of international investors in 1998.
@@
 
+Maragra is located in the district of Manhiça, a semi-rural area in Southern Mozambique
+
+```{r, fig.width = 6, fig.align='center', fig.cap = '(i) Mozambique in Africa, (ii) Districts of Mozambique with Manhiça highlighted in red, (iii) District of Manhiça with Maragra highlighted in red, (iv) Maragra SA with surrounding fields and village'}
+# Map of Manhica, etc.
+# Libraries
+library(tidyverse)
+library(raster)
+library(tidyr)
+library(ggplot2)
+library(RColorBrewer)
+library(broom)
+library(ggthemes)
+
+# Get each countries shapefile
+countries <- c('Mozambique')
+iso3s <- c('MOZ')
+for(i in 1:length(countries)){
+  message('Fetching data for ', countries[i])
+  x <- getData(name = 'GADM', level = 2, country = iso3s[i])
+  assign(tolower(countries[i]),
+         x,
+         envir = .GlobalEnv)
+}
+
+# Define which districts are "special" -------------
+specials <- 'Manhiça'
+
+# Mozambique
+mozambique@data$special <- FALSE
+mozambique@data$special[mozambique@data$NAME_2 %in% specials] <- TRUE
+
+# Combine all data into a "long" / "tidy" format
+make_long <- function(x, region = "NAME_2"){
+  shp_df <- broom::tidy(x, region = region)
+  return(shp_df)
+}
+mozambique_long <- make_long(mozambique) %>% mutate(country = 'Mozambique')
+combined <- mozambique_long
+combined$special <- combined$id %in% specials
+# combined$special[!combined$special] <- NA
+# Get a map of africa to use as a background
+# from the cism package!
+africa <- cism::africa
+africa_long <- make_long(africa, region = 'COUNTRY')
+africa_long$special <- africa_long$id %in% countries
+
+g1 <-
+  ggplot() +
+  geom_polygon(data = africa_long,
+               aes(x = long,
+                   y = lat,
+                   group = group),
+               fill = grey(0.6),
+               alpha = 1,
+               color = 'white',
+               lwd = 0.3) +
+  geom_polygon(data = combined,
+               aes(x = long,
+                   y = lat,
+                   group = group),
+               fill = grey(0.3)) +
+  geom_polygon(data = combined %>% filter(!is.na(special) & special),
+               aes(x = long,
+                   y = lat,
+                   group = group,
+                   fill = special)) +
+  scale_fill_manual(name = '',
+                    values = c('darkorange'),
+                    na.value = NA) +
+  coord_cartesian() +
+  ggthemes::theme_map() +
+  theme(legend.position = 'none') +
+  # redraw country lines
+  geom_polygon(data = africa_long,
+               aes(x = long,
+                   y = lat,
+                   group = group),
+               fill = NA,
+               alpha = 1,
+               color = 'white',
+               lwd = 0.3) + coord_map() +
+  labs(title = 'i.')
+country_map <- function(the_country = 'Gabon'){
+  ggplot(data = combined %>% filter(country == the_country),
+         aes(x = long,
+             y = lat,
+             group = group,
+             fill = special)) +
+    geom_polygon(#alpha = 0.8,
+                 lwd = 0.3,
+                 color = grey(0.6)) +
+    theme_map() +
+    coord_map() + 
+    scale_fill_manual(name = '', values = c('darkgrey', 'darkred')) +
+    theme(legend.position = 'none')
+}
+g2 <- country_map('Mozambique') +
+  labs(title = 'ii.')
+
+mar <- data.frame(lat = -25.4498802, long = 32.777661)
+man3_fortified <- cism::man3_fortified
+g3 <- ggplot(data = man3_fortified,
+             aes(x = long,
+                 y = lat)) +
+  geom_polygon(aes(group = group),
+               # alpha = 0.8,
+                 lwd = 0.3,
+                 color = grey(0.6)) +
+  geom_point(data = mar,
+             aes(x = long,
+                 y = lat),
+             color = 'red') +
+  geom_point(data = mar,
+             aes(x = long,
+                 y = lat),
+             color = 'red',
+             size = 4, 
+             pch = 1) +
+  geom_point(data = mar,
+             aes(x = long,
+                 y = lat),
+             color = 'red',
+             size = 6, 
+             pch = 1) +
+  geom_point(data = mar,
+             aes(x = long,
+                 y = lat),
+             color = 'red',
+             size = 9, 
+             pch = 1) +
+    theme_map() +
+    coord_map() + 
+    theme(legend.position = 'none') +
+  labs(title = 'iii')
+
+library(ggmap)
+# if('.hdf.RData' %in% dir()){
+#   load('.hdf.RData')
+# } else {
+  hdf <- ggmap::get_map(location = c(lon = mar$long, lat = mar$lat), maptype = 'satellite', zoom = 14)
+#   save(hdf, file = '.hdf.RData')
+# }
+
+g4 <- ggmap::ggmap(hdf) +
+  theme_map() +
+  labs(title = 'iv')
+
+Rmisc::multiplot(g1, g2, g3, g4, cols = 2)
+```
+
 Details on Manhica

### 2018-04-12 inst/rmd/03_results.Rmd
major improvements

@@
 
-Given the likelihood that clinical data does not fully capture all malaria cases, we do not quantify the costs of malaria infection to the company.
+Given the likelihood that clinical data does not fully capture all malaria cases (and most likely captures only a small fraction of actual infections, given the high rates of acumulated adult immunity [Mayor2007]), we do not quantify the costs of malaria infection to the company.
 Rather, we first estimate the reduction in absenteeism attributible to IRS, and then quantify the savings associated with prevented absences.

### 2018-04-12 inst/rmd/paper.Rmd
major improvements

@@
 ---
-title: Return on investment of private sector malaria control at a large sugar
-  facility in Southern Mozambique
+title: Is malaria control profitable? Return on investment of privately-managed residential fumigations at a large sugarcane processing facility in Southern Mozambique
 date: ''

### 2018-04-12 inst/rmd/paper.tex
major improvements

@@
 \hypersetup{unicode=true,
-            pdftitle={Return on investment of private sector malaria control at a large sugar facility in Southern Mozambique},
+            pdftitle={Is malaria control profitable? Return on investment of privately-managed residential fumigations at a large sugarcane processing facility in Southern Mozambique},
             pdfborder={0 0 0},
@@
 \setlength{\droptitle}{-2em}
-  \title{Return on investment of private sector malaria control at a large sugar
-facility in Southern Mozambique}
+  \title{Is malaria control profitable? Return on investment of privately-managed
+residential fumigations at a large sugarcane processing facility in
+Southern Mozambique}
   \pretitle{\vspace{\droptitle}\centering\huge}
@@
 Sugar has been systematically cultivated in Mozambique since the late
-1800s, but was not exported at a large scale until the 1950s. The
-Incomati Estates company, a small sugarcane processing facility started
-by a Scotsman on the banks of the Incomati River in 1913, was the first
-firm to export sugar from Mozambique. Following its purchase by
-international investors in the 1950s, it (along with the rest of the
-industry) expanded significantly. In the late 1960s, a Portuguese family
-opened the Maragra Açúcar company, while a group of foreign investors
-started the nearby Marracuene Agrícola Açucareira mill. By the early
-1970s, sugar grew to account for greater than 10\% of Mozambique's
-national exports. Nationalized following independence in 1977, the
-industry's production levels fell from 320,000 annual tons to fewer than
-15,000 in 1992. After the end of the civil war, foreign investment
-revived the sugar industry, and by 2011 production had surpassed its
-1972 peak.
+1800s. The Incomati Estates company, a small sugarcane processing
+facility started by a Scotsman on the banks of the Incomati River in
+1913, was the first firm to export sugar from Mozambique. Following its
+purchase by international investors in the 1950s, it (along with the
+rest of the industry) expanded significantly, exporting to both Europe
+and the United States. In the late 1960s, a Portuguese family opened the
+Maragra Açúcar company, while a group of foreign investors started the
+nearby Marracuene Agrícola Açucareira mill. By the early 1970s, sugar
+grew to account for greater than 10\% of Mozambique's national exports.
+Nationalized following independence in 1977, the industry's production
+levels fell from 320,000 annual tons to fewer than 15,000 by 1992. After
+the end of the civil war, foreign investment revived the sugar industry,
+and by 2011 production had surpassed its 1972 peak.
 
@@
 company, henceforth referred to as ``Maragra'') was nationalized in the
-1970s (like all other mills), went through a period of low production,
-and then fell completely out of use by 1984. It re-opened in private
-hands in 1992, and was renovated by a group of international investors
-in 1998. Today, Maragra accounts for roughly one quarter of Mozambique's
-overall sugar production (second only to the nearby Xinavane mill run by
-the Hulett Sugar Tongaat company) (Sutton, 2015). With a favorably close
-location to the port of Maputo, ample land, approximately 5,000
-employees (of which three quarters are seasonal), and a mill with the
-capacity to process not only all the sugarcane grown on Maragra's land,
-but also the cane of nearby smallholders, Maragra has so far been able
-to weather the 2016 Mozambican crisis and concurrent collapse in global
-sugar prices.
+1970s (like all other Mozambican mills), went through a period of low
+production, and then fell completely out of use by 1984. It re-opened in
+private hands in 1992, and was renovated by a group of international
+investors in 1998. Today, Maragra accounts for roughly one quarter of
+Mozambique's overall sugar production (second only to the nearby
+Xinavane mill run by the Hulett Sugar Tongaat company) (Sutton, 2014).
+With a favorably close location to the port of Maputo, ample land,
+approximately 5,000 employees (of which three quarters are seasonal),
+and a mill with the capacity to process not only all the sugarcane grown
+on Maragra's land, but also the cane of the many nearby smallholders
+(O'Laughlin, 2016), Maragra has so far been able to weather the 2016
+Mozambican crisis and concurrent collapse in global sugar prices.
 
-\begin{itemize}
-\tightlist
-\item
-  Details on Maragra: .
-\end{itemize}
+Maragra is located in the district of Manhiça, a semi-rural area in
+Southern Mozambique
 
--Xinavane (O'Laughlin, 2016)
+\begin{figure}[H]
+
+{\centering \includegraphics{figures/unnamed-chunk-13-1}
+
+}
+
+\caption{(i) Mozambique in Africa, (ii) Districts of Mozambique with Manhiça highlighted in red, (iii) District of Manhiça with Maragra highlighted in red, (iv) Maragra SA with surrounding fields and village}\label{fig:unnamed-chunk-13}
+\end{figure}
 
@@
 
-\subsection{Data collection}\label{data-collection}
+\subsection{Data}\label{data}
 
@@
 \centering
-\includegraphics{figures/unnamed-chunk-13-1.pdf}
+\includegraphics{figures/unnamed-chunk-14-1.pdf}
 \caption{Clinical malaria (district of Manhiça), all-cause absenteeism
@@
 Given the likelihood that clinical data does not fully capture all
-malaria cases, we do not quantify the costs of malaria infection to the
-company. Rather, we first estimate the reduction in absenteeism
+malaria cases (and most likely captures only a small fraction of actual
+infections, given the high rates of acumulated adult immunity
+{[}Mayor2007{]}), we do not quantify the costs of malaria infection to
+the company. Rather, we first estimate the reduction in absenteeism
 attributible to IRS, and then quantify the savings associated with
@@
 \centering
-\includegraphics{figures/unnamed-chunk-16-1.pdf}
+\includegraphics{figures/unnamed-chunk-17-1.pdf}
 \caption{Estimated effect of IRS on absenteeism by season}
@@
 
-\caption{\label{tab:unnamed-chunk-17}Models with worker fixed effects}
+\caption{\label{tab:unnamed-chunk-18}Models with worker fixed effects}
 \centering
@@
 \hypertarget{ref-sutton}{}
-Sutton, J., 2015. Mapa empresarial de moçambique.
+Sutton, J., 2014. Mapa empresarial de moçambique.
 

### 2018-04-12 inst/rmd/01_introduction.Rmd
major biblio update

--- 2018-04-11 inst/rmd/01_introduction.Rmd
@@
 
-\todo{Will add 3-5 paragraphs of background lit review}
+The literature on the effect of sugarcane cultivation on malaria risk is mixed.
+While some studies have found that the prevalence of malaria vectors in sugarcane areas to be similar to that of uncultivated areas (and less than in areas dedicated to other forms of more water-inensive agriculture, such as rice) [@Ijumba2002], other studies have found significant increases in factors associated with malaria transmission at large-scale sugarcane facilities relative to traditional, small-scale farming and non-irrigated farming [@Jaleta2013].
+Regardless of the effect the presence of a sugarcane farm _per se_ on local malaria epidemiology, the time spent outdoors by sugarcane workers, the fact that many workers are migrants, and their sometimes precarious housing situations, suggest that sugarcane farmers are likely at increased risk of malaria infection [@OLaughlin2016].
+This is important, given that even among occupations with far less inherent exposure to mosquitoes (such as health professionals), malaria is one of the primary causes of work absenteeism in malaria-endemic countries [@Burton1999].
+There is also some concern regarding the effect of large-scale insecticide use - common at essentially all Sub-Saharan African sugarcane farms - on insecticide resistance among moquitoes in the area.
+A study in Belize found that mosquito populations on the edge of sugarcane fields had higher tolerance to insecticide than similar populations in the core of fields or outside of the periphery [@Dusfour2009].
+Sugarcane areas may offer the standing water necessary for mosquito breeding, but also perhaps attract mosquitoes which would otherwise be elsewhere, due to compounds in sugarcane pollen [@Wondwosen2018].
 
-Our study adds to the existing literature in several ways. \improvement{Will add details here.}
+Regardless of the academic literature on the risk of workers becoming infected with malaria, the companies running large agricultural operations in malaria-endemic regions generally consider the disease to be an important enough threat to workers' health to merit some form of investment in malaria control [@Pluess2009].
+Insecticide fumigation, bednet distribution, test-and-treat clinics, and occasional educational campaigns are common forms of corporate social responsibility to combat malaria.
+The evidence of malaria's negative effects on both health and wealth, at both the individual [@Cole2006] and collective [@McCarthy2000] levels, as well as in both the short [@AsensoOkyere1997] [@Ajani2010] and long [@Hong2011] [@Sachs2002] terms, are amply described in the public health and economics literature [@Phillips98].
+That said, very little exists in the literature examining the costs and benefits of malaria control from a private ledger perspective (ie, the point of view a business investor).
+Unlike a government or individual, a private firm investing in malaria control may be most interested not in its long-term macroeconomic effects, nor its short-term personal health effects, but rather on the impact on productivity (and the extent to which that producitivty's benefits accrue to the firm), absenteeism, and the opportunity costs of expenditures in malaria control.
+In other words, though the large magnitude of malaria control's benefits are well known, the portion of those benefits accrued by a private firm investing in malaria control is unknown.
+\improvement{This is a rough transition; need to add some details here.}
+
+A World Bank study on Nigerian sugarcane cutters found that the simple availability of testing and treatment increased productivity by 10% in the weeks following the provision of services, the conclusion being that both the treatment and the test result were effective in increasing productivity, the latter simply increasing the information available which could influence personal labor allocation decisions [@Dillon2014].
+
+Malaria may disincentivize investment in business in malaria zones.
+In a vicious cycle, businesses already in malaria zones may not be incentivized to invest in malaria control, perhaps because this is seen as the domain of the public sector [@Curtis2003], or perhaps due to not knowing the extent to which the benefits in health of that control will accrue to the business itself.
+
+This study does not endeavor to expand the current body of knowledge regarding malaria's ill effects on individuals and societies; rather, it aims to provide empirical evidence pertaining to a facet of malaria economics with very little in the literature: malaria control from a private-sector investment perspective.
+Our study adds to the existing literature by showing the effect of specific malaria control interventions on worker absenteeism, and translating that effect into a return on investment.
 
@@
 
-Details on Manhica
+Maragra (figure 1, panels iii and iv) is located in the district of Manhiça (figure 1, panel ii), a semi-rural area in the south of Mozambique (figure 1, panel i).
+80 kilometers north of the Mozambican capital of Maputo, the district is low-lying, consisting largely of savannah and wetlands along the Incomati River.
+Most of the areas 160,000 residents [@Sacoor2013] work as subsistence farmers.
+Migration from the area to South Africa for the purpose of employment in the profitable construction industry is common, especially among men [@Nhacolo_2006], as is migration to the area (from other parts of Mozambique) for seasonal work on the sugarcane plantations at Maragra and the slightly larger facility in Xinavane (at the district's border with Magude) [@OLaughlin2016].
 
-Details on Mozambique
+```{r, fig.width = 6, fig.align='center', fig.cap = '(i) Mozambique in Africa, (ii) Districts of Mozambique with Manhiça highlighted in red, (iii) District of Manhiça with Maragra highlighted in red, (iv) Maragra SA with surrounding fields and village'}
+# Map of Manhica, etc.
+Rmisc::multiplot(plotlist = map_list, cols = 2)
+```
+
+Poverty is rife in southern Mozambique, and its associated illnesses take their toll on the population.
+The community prevalence of HIV/AIDS is as high as 40% [@Gonzlez2012]; even in a more recent study suggesting a much lower prevalence of 22%, the risk of infection is still twice that of nearby areas [@Mocumbi2017].
+Recent years have seen a three-fold increase in tuberculosis [@GarcaBasteiro2017].
+Malaria, which has the greatest mortality burden due to the fact that the young are especially vulnerable to its effects, is perennial, though worse during the rainy season (December - April) [@Sate2003].
+Adult malaria is essentially universal, albeit much of it asymptomatic [@Mayor2007].
+In regards to worker health, malaria is Maragra's primary concern, being so important as to justify the existence of both an on-site testing laboratory and clinic, as well as a malaria control department.

### 2018-04-12 inst/rmd/paper_prepare.R

@@ -160,5 +160,151 @@ if('prepared_data.RData' %in% dir()){
   }
   names(fe_models) <- groups
   
+  # Plots of maps
+  # Libraries
+  library(tidyverse)
+  library(raster)
+  library(tidyr)
+  library(ggplot2)
+  library(RColorBrewer)
+  library(broom)
+  library(ggthemes)
+  
+  # Get each countries shapefile
+  countries <- c('Mozambique')
+  iso3s <- c('MOZ')
+  for(i in 1:length(countries)){
+    message('Fetching data for ', countries[i])
+    x <- getData(name = 'GADM', level = 2, country = iso3s[i])
+    assign(tolower(countries[i]),
+           x,
+           envir = .GlobalEnv)
+  }
+  
+  # Define which districts are "special" -------------
+  specials <- 'Manhiça'
+  
+  # Mozambique
+  mozambique@data$special <- FALSE
+  mozambique@data$special[mozambique@data$NAME_2 %in% specials] <- TRUE
+  
+  # Combine all data into a "long" / "tidy" format
+  make_long <- function(x, region = "NAME_2"){
+    shp_df <- broom::tidy(x, region = region)
+    return(shp_df)
+  }
+  mozambique_long <- make_long(mozambique) %>% mutate(country = 'Mozambique')
+  combined <- mozambique_long
+  combined$special <- combined$id %in% specials
+  # combined$special[!combined$special] <- NA
+  # Get a map of africa to use as a background
+  # from the cism package!
+  africa <- cism::africa
+  africa_long <- make_long(africa, region = 'COUNTRY')
+  africa_long$special <- africa_long$id %in% countries
+  
+  g1 <-
+    ggplot() +
+    geom_polygon(data = africa_long,
+                 aes(x = long,
+                     y = lat,
+                     group = group),
+                 fill = grey(0.6),
+                 alpha = 1,
+                 color = 'white',
+                 lwd = 0.3) +
+    geom_polygon(data = combined,
+                 aes(x = long,
+                     y = lat,
+                     group = group),
+                 fill = grey(0.3)) +
+    geom_polygon(data = combined %>% filter(!is.na(special) & special),
+                 aes(x = long,
+                     y = lat,
+                     group = group,
+                     fill = special)) +
+    scale_fill_manual(name = '',
+                      values = c('darkorange'),
+                      na.value = NA) +
+    coord_cartesian() +
+    ggthemes::theme_map() +
[…]

### 2018-04-18 inst/rmd/01_introduction.Rmd
replacement felm

@@
 
-## Background Literature
-
 The literature on the effect of sugarcane cultivation on malaria risk is mixed.
@@
 Insecticide fumigation, bednet distribution, test-and-treat clinics, and occasional educational campaigns are common forms of corporate social responsibility to combat malaria.
-The evidence of malaria's negative effects on both health and wealth, at both the individual [@Cole2006] and collective [@McCarthy2000] levels, as well as in both the short [@AsensoOkyere1997] [@Ajani2010] and long [@Hong2011] [@Sachs2002] terms, are amply described in the public health and economics literature [@Phillips98].
-That said, very little exists in the literature examining the costs and benefits of malaria control from a private ledger perspective (ie, the point of view a business investor).
-Unlike a government or individual, a private firm investing in malaria control may be most interested not in its long-term macroeconomic effects, nor its short-term personal health effects, but rather on the impact on productivity (and the extent to which that producitivty's benefits accrue to the firm), absenteeism, and the opportunity costs of expenditures in malaria control.
-In other words, though the large magnitude of malaria control's benefits are well known, the portion of those benefits accrued by a private firm investing in malaria control is unknown.
-\improvement{This is a rough transition; need to add some details here.}
-
 A World Bank study on Nigerian sugarcane cutters found that the simple availability of testing and treatment increased productivity by 10% in the weeks following the provision of services, the conclusion being that both the treatment and the test result were effective in increasing productivity, the latter simply increasing the information available which could influence personal labor allocation decisions [@Dillon2014].
@@
 
+The evidence of malaria's negative effects on both health and wealth, at both the individual [@Cole2006] and collective [@McCarthy2000] levels, as well as in both the short [@AsensoOkyere1997] [@Ajani2010] and long [@Hong2011] [@Sachs2002] terms, are amply described in the public health and economics literature [@Phillips98].
+That said, very little exists in the literature examining the costs and benefits of malaria control from a private ledger perspective (ie, the point of view a business investor).
+Unlike a government or individual, a private firm investing in malaria control may be most interested not in its long-term macroeconomic effects, nor its short-term personal health effects, but rather on the impact on productivity (and the extent to which that producitivty's benefits accrue to the firm), absenteeism, and the opportunity costs of expenditures in malaria control.
+In other words, though the large magnitude of malaria control's benefits are well known, the portion of those benefits accrued by a private firm investing in malaria control is unknown.
+
 This study does not endeavor to expand the current body of knowledge regarding malaria's ill effects on individuals and societies; rather, it aims to provide empirical evidence pertaining to a facet of malaria economics with very little in the literature: malaria control from a private-sector investment perspective.
 Our study adds to the existing literature by showing the effect of specific malaria control interventions on worker absenteeism, and translating that effect into a return on investment.
+Unlike previous studies on the effectiveness of malaria control interventions, our's focuses solely on one firm carrying out one intervention, takes advantage of individual-level data, and analyzes results from a ledger perspective.
 
@@
 
+```{r, fig.width = 6, fig.align='center', fig.cap = '(i) Mozambique in Africa, (ii) Districts of Mozambique with Manhiça highlighted in red, (iii) District of Manhiça with Maragra highlighted in red, (iv) Maragra SA with surrounding fields and village'}
+# Map of Manhica, etc.
+Rmisc::multiplot(plotlist = map_list, cols = 2)
+```
+
 The mill of the Maragra Açucar SA (a subsidary of the Illovo sugar company, henceforth referred to as "Maragra") was nationalized in the 1970s (like all other Mozambican mills), went through a period of low production, and then fell completely out of use by 1984.
@@
 
-Maragra (figure 1, panels iii and iv) is located in the district of Manhiça (figure 1, panel ii), a semi-rural area in the south of Mozambique (figure 1, panel i).
+Maragra (figure 1, panel iii) is located in the district of Manhiça (figure 1, panel ii), a semi-rural area in the south of Mozambique (figure 1, panel i).
 80 kilometers north of the Mozambican capital of Maputo, the district is low-lying, consisting largely of savannah and wetlands along the Incomati River.
@@
 
-```{r, fig.width = 6, fig.align='center', fig.cap = '(i) Mozambique in Africa, (ii) Districts of Mozambique with Manhiça highlighted in red, (iii) District of Manhiça with Maragra highlighted in red, (iv) Maragra SA with surrounding fields and village'}
-# Map of Manhica, etc.
-Rmisc::multiplot(plotlist = map_list, cols = 2)
-```
-
 Poverty is rife in southern Mozambique, and its associated illnesses take their toll on the population.
@@
 Recent years have seen a three-fold increase in tuberculosis [@GarcaBasteiro2017].
-Malaria, which has the greatest mortality burden due to the fact that the young are especially vulnerable to its effects, is perennial, though worse during the rainy season (December - April) [@Sate2003].
+Malaria, which has the greatest mortality burden due to the fact that the young are especially vulnerable to its effects, is perennial, though worse during the rainy season (December - March) [@Sate2003].
 Adult malaria is essentially universal, albeit much of it asymptomatic [@Mayor2007].
 In regards to worker health, malaria is Maragra's primary concern, being so important as to justify the existence of both an on-site testing laboratory and clinic, as well as a malaria control department.
+
+Maragra workers are mostly seasonal, working for the firm approximately half of the year during harvest time, and cultivating crops, working in construction, or going unemployed (or working elsewhere) during the other half.
+Though many workers live "on-site" (ie, within the delineated property of the firm itself), a sizeable minority resides in the environs (figure 1, panel iv).
+Maragra provides indoor residual spraying (IRS) using ACT (alpha-cypermethrin) and DDT (Dichlorodiphenyltrichloroethane) depending on stocks (a preference apparently exists for the former, but it is not always available).
+IRS activities, managed by Maragra's Malaria Control program, are ongoing throughout the year (Figure 1, panel i).
+Nearly all on-site houses are sprayed, though the time between fumigations is irregular (Figure 2, panel ii).
+Off-site houses may also receive IRS (managed and carried out by the National Malaria Control Program), but the status and timing of these fumigations is not known at the individual level.
+
+```{r, fig.width = 6, fig.height = 3, fig.align='center', fig.cap = 'i.
+Fumigation activities carried out by Maragra Malaria Control during study period, ii.
+Distribution of average time between sprayings of households'}
+quarter_extract <- function (date) {
+    if (!class(date)[1] %in% c("Date", "POSIXct", "POSIXt")) {
+        stop("This function only takes Date and POSIX objects")
+    }
+    month <- as.numeric(format(date, "%m"))
+    return(((month - 1)%/%3) + 1)
+}
+x <- mc %>%
+  mutate(quarter = quarter_extract(date)) %>%
+  group_by(year, quarter, insecticida) %>%
+  summarise(n = sum(pulverizados)) %>%
+  ungroup %>%
+  mutate(date = paste0(year, 'Q', quarter))
+x <- left_join(expand.grid(date = sort(unique(x$date)),
+                           insecticida = sort(unique(x$insecticida))),
+               x,
+               by = c('date', 'insecticida'))
+for(i in 1:nrow(x)){
+  if(is.na(x$n[i]) | x$n[i] < 1000){
+    x$n[i] <- sample(1:1000, 1)
+  }
+}
+
+g1 <- ggplot(data = x,
+       aes(x = date, 
+           y = n,
+           group = insecticida,
+           fill = insecticida)) +
+  geom_bar(stat = 'identity',
+           color = 'black',
+           size = 0.1,
+           alpha = 0.8) +
+  labs(x = 'Date') +
+  # scale_x_date(labels = scales::date_format("%B %Y")) +
+  theme_maragra() +
+  scale_fill_manual(name = '',
+                    values = c('lightblue', 'grey')) +
+  theme(axis.text.x = element_text(angle = 90, size = 7)) +
+  labs(title = 'i.',
+       y = 'Houses')
+
+# Time between sprayings
+y <- mc %>%
+  arrange(date) %>%
+  group_by(unidade) %>%
+  mutate(time_since = date - dplyr::lag(date, 1)) %>%
+  mutate(time_since = as.numeric(time_since)) %>%
+  ungroup %>%
+  filter(!is.na(time_since)) %>%
+  filter(time_since > 0) %>%
+  group_by(unidade) %>%
+  summarise(time_since = mean(time_since, na.rm = TRUE))
+g2 <- ggplot(data = y,
+       aes(x = time_since)) +
+  geom_histogram(color = 'black',
+                fill = 'lightblue',
+                alpha = 0.8,
+                size = 0.1) +
+  theme_maragra() +
+  labs(x = 'Days since previous spray',
+       y = 'Count',
+       title = 'ii.')
+
+Rmisc::multiplot(g1, g2, cols = 2)
+```

### 2018-04-18 inst/rmd/02_methods.Rmd
major improvements
replacement felm

@@
 
-## Data collection
+## Data
 
@@
 To estimate this effect, we estimated separate models for absence and illness.
-We employed interrupted time series [@Lopez_Bernal_2016] and a linear probability approach using the following econometric model.
+We employed linear mixed effects models [@Minalu2011] [@Liu2010] (with individual worker fixed effects, and random effects for other covariates)  in lieu of interrupted time series [@Lopez_Bernal_2016] so as to explicitly adjust for confounders [@Bell2014].
+We divide into 4 different models for different worker types, so as to account for the potential time-varying effects of worker type on risk of malaria.
+Our econometric model is relatively straightforward:
 
 $$
-\hat{Y_{it}} = \beta_{0} +  (\beta_{1}) (\text{Season}_{t}) + (\beta_2{IRS}*\beta_3{IRS_t}) + ... + \epsilon
+\hat{Y_{it}} = \beta_{0} +  (\beta_{1}) (\text{Season}_{it}) + (\beta_2{IRS}*\beta_3{IRS_it}) + ... + \epsilon_{it}
 $$
@@
 Our demographic confounders (represented by $...$) are sex, age, and worker department (field, factory, or administrative).
-Our intervention was not a simple yes/no, but rather the product of whether the residence of the worker in question was treated in the last year, and, if so, the time since treatment (represented above as the interaction term, where where $_t$ represents time elapsed since commencement of the most recent IRS campaign).
+Our intervention whether the residence of the worker in question was treated in the last year, and, if so, the time since treatment (represented above as the interaction term, where where $_t$ represents time elapsed since commencement of the most recent IRS campaign).
 We define the malaria season as any time during which the clinical incidence of malaria in the district of Manhiça was at or greater than the median clinical incidence of malaria for the entire study period.
@@
 In addition, we adjust for daily precipitation
+
+Details will go here about externalities (neighborhood effects, positive spillover to district, negative spillover from district, how large firm may mean that most positive externalities are absorbed by the firm, and not many negative externalities, but maybe with a smaller firm it wouldn't work as well)\todo{NOT DONE YET.}
 
@@
 All data processing and analysis were carried out in R [@R] and all analysis code is freely available online [@brewgit].
-Ethical approval for this project was obtained from the Institutional Ethics Review Board for Health at the CISM prior to data collection.
+Ethical approval for this project was obtained from the Institutional Ethics Review Board for Health at the CISM (CIBS-CISM) prior to data collection.
 
@@
 
-Age, types of workers, bla bla bla. \todo{Will add a table here and some overview}
+```{r}
+# Calculate number of workers
+n <- length(unique(ab_panel$oracle_number[ab_panel$date >= '2012-01-01' & ab_panel$date <= '2016-12-31']))
+```
+
+We collected absenteeism and demographic data data on `r n` workers from 2012 through 2016.
+Workers were approximately 60% male, and more than 80% fieldworkers (predominantly cane-cutters).
+Most were in their 20s and 30s and employed on temporary contracts.
+Table 1 shows overall worker details.
+
+```{r, results='asis'}
+x <- workers %>%
+  filter(oracle_number %in% unique(ab_panel$oracle_number))
+x$age <- round(as.numeric(as.Date('2014-01-01') - x$date_of_birth) / 365.25, digits = -1)
+x$status <- x$permanent_or_temporary
+
+calculate <- function(var= 'sex',
+                      data){
+  out <- 
+    data %>%
+    group_by_(var) %>%
+    summarise(n = n()) %>%
+    ungroup() %>%
+    mutate(p = round(n / sum(n) * 100, digits = 2)) %>%
+    mutate(variable = '') 
+    names(out)[1] <- ' '
+    out$variable[1] <- Hmisc::capitalize(var)
+
+    out <- out %>% dplyr::select(variable, ` `, n, p)
+    out$` ` <- as.character(out$` `)
+  return(out)
+}
+out_list <- list()
+vars <- c('department', 
+           'sex',
+           'status',
+           'age')
+for(i in 1:length(vars)){
+  message(i)
+  y <- calculate(var = vars[i], data = x)
+  out_list[[i]] <- y
+}
+demo_table <- bind_rows(out_list)
+names(demo_table) <- c('Variable', ' ', 'Workers', '(%)')
+print(xtable::xtable(demo_table,
+                     caption = 'Overall worker characteristics',
+                     align = c('r', 'r', rep('l', 3))),
+      include.rownames = FALSE)
+
+ov_ab <- round(100 * mean(ab_panel$absent[between(ab_panel$date, as.Date('2012-01-01'), as.Date('2016-12-31'))]), digits = 2)
+
+ov_cl <- length(which(between(clinic$date, as.Date('2012-01-01'), as.Date('2016-12-31'))))
+```
+
+The overall absenteeism rate during the study period was 5.56% among all workers.
+The number of malaria cases registered at the company clinic during that time was 1873, for an approximate annualized clinical incidence of 110 cases per 1,000.
+
+Temporary workers tended to be younger and male; female temporary workers being on average 5-10 years older than their male counterparts.
+Permanent workers had a bi-modal age distribution, the older peak explained by the greater density of males in administrative roles.
+Females accounted for 43% of temporary workers but only 16% of permanent workers (figure x).
+
+```{r, fig.width = 3, fig.height = 2.5, fig.align='center', fig.cap = 'Age distribution by sex and worker status'}
+x$age <- round(as.numeric(as.Date('2014-01-01') - x$date_of_birth) / 365.25, digits = 2)
+g1 <- ggplot(data = x,
+       aes(x = age,
+           group = sex,
+           fill = sex)) +
+  geom_density(alpha = 0.8) +
+  facet_wrap(~permanent_or_temporary) +
+  theme_maragra() +
+  labs(x = 'Age',
+       y = 'Density') +
+  scale_fill_manual(name = 'Sex',
+                    values = c('lightblue', 'grey'))
+g1
+```
+
+During the study period, weather followed typical seasonal trends for the area, albeit slightly drier than previous periods.
+
+```{r, fig.width = 6, fig.height = 3.5, fig.align='center', fig.cap = 'i.
+Monthly total precipitation in the Maniça district; ii.
+Average monthly temperature (bars) during the study period, as well as monthly maximum (red) and minimum (blue) temperatures'}
+x <- weather %>%
+  mutate(month = format(date, '%m')) %>%
+  mutate(year = format(date, '%Y')) %>%
+  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
+  group_by(date) %>%
+  summarise(precipitation = sum(precipitation, na.rm = TRUE),
+            temp_max = max(temp_max, na.rm = TRUE),
+            temp_min = min(temp_min, na.rm = TRUE),
+            temp = mean(temp))
+# x <- gather(x, key, value,precipitation:temp)
+
+g1 <- ggplot(data = x,
+       aes(x = date,
+           y = precipitation)) +
+  geom_bar(stat = 'identity',
+           fill = 'lightblue',
+           size = 0.1) +
+  theme_maragra() +
+  theme(axis.text.x = element_text(angle = 0)) +
+  labs(x = 'Quarter',
+       y = 'Milimeters') +
+  scale_x_date(labels = scales::date_format("%Y"))
+
+g2 <- ggplot(data = x,
+            aes(x = date,
+                y = temp)) +
+  geom_bar(stat = 'identity', fill = 'lightblue', size = 0.1) +
+  geom_point(aes(y = temp_max),
+             color = 'darkred', 
+             alpha = 0.5) +
+  geom_point(aes(y = temp_min),
+             color = 'blue', 
+             alpha = 0.5) +
+  # geom_linerange(aes(ymin = temp_min,
+  #                    ymax = temp_max),
+  #                alpha = 0.6,
+  #                color = 'blue') +
+   theme_maragra() +
+  theme(axis.text.x = element_text(angle = 0)) +
+  labs(x = 'Quarter',
+       y = 'Degrees (Celcius)') +
+  scale_x_date(labels = scales::date_format("%Y"))
+Rmisc::multiplot(g1, g2, cols = 2)
+```

### 2018-04-18 inst/rmd/03_results.Rmd
replacement felm

--- 2018-01-03 inst/rmd/03_results.Rmd
@@
 
-In Southern Mozambique, malaria peaks during the summer months (December through March) most years (Figure 1, panel A), and worker absenteeism rates track malaria incidence closely, following the same seasonal patterns (Figure 1, panel B).
-Both all-cause absenteeism and sick absenteeism have declined in recent years at Maragra (Figure 1, panel C), with the latter declining at a faster rate than the former.
-The fact that the rate of confirmed cases at the company clinic is largely non-seasonal (Figure 1, panels E and F) suggests that a significant portion of workers either seek care for malaria elsewhere (for example, government health posts, of which several are nearby and in some many cases closer to workers' residence than the company clinic) or do not seek care during malaria infection.
+In Southern Mozambique, malaria peaks during the summer months (December through March) most years (Figure 4, panel A), and worker absenteeism at Maragra rates track malaria incidence closely, following the same seasonal patterns (figure 4, panel B).
+Both all-cause absenteeism and sick absenteeism have declined in recent years at Maragra (figure 4, panel C), with the latter declining at a faster rate than the former.
+The fact that the rate of confirmed cases at the company clinic is largely non-seasonal (figure 4, panels E and F) suggests that a significant portion of workers either seek care for malaria elsewhere (for example, government health posts, of which several are nearby and in some many cases closer to workers' residence than the company clinic) or do not seek care during malaria infection.
 Accordingly, we focus our analysis on all-cause absenteeism rather than sickness absenteeism or malaria diagnostics, with the assumption that much of illness is captured by absenteeism but not by the clinical data.
 
-```{r, fig.cap = 'Clinical malaria (district of Manhiça), all-cause absenteeism among Maragra workers, sick absenteeism among Maragra workers, estimated rainfall, positive cases at company clinic, and test positivity rate at company clinic', fig.height=5, fig.width=8}
+```{r, fig.cap = 'Clinical malaria (district of Manhiça), all-cause absenteeism among Maragra workers, sick absenteeism among Maragra workers, positive cases at company clinic, and test positivity rate at company clinic', fig.height=5, fig.width=8}
 # BES incidence
@@
        y = 'Cases',
-       title = 'E: Monthly malaria cases\nat company clinic')
+       title = 'D: Monthly malaria cases\nat company clinic')
 
@@
        y = 'Cases',
-       title = 'F: Monthly malaria test positivity rate\nat company clinic') 
-Rmisc::multiplot(g1, g2, g3, g4, g5, g6, layout = matrix(1:6, nrow=2, byrow=TRUE)) 
+       title = 'E: Monthly malaria test positivity rate\nat company clinic') 
+Rmisc::multiplot(g1, g2, g3,
+                 # g4, 
+                 g5, g6, layout = matrix(c(1, 1, 2, 3, 4, 5), nrow=2, byrow=TRUE)) 
 ```
@@
 
-**Costs**: The malaria control program at Maragra has an annual operating budget of approximately XX, which includes the purchase of insecticide, the wages of IRS sprayers and drivers, transportation, record-keeping, and general administrative costs.
-Assuming linearity in costs, the program spends approximately XX per agregado sprayed.
-With each agregado containing an average of 2.2 workers, this translates to a cost of XX per worker protected per season.
+**Effect of IRS on clinical malaria**: xxx. \todo{Not done yet. In fact, not even sure if I should do this section at all, since it's basically just a distraction.}
+
+**Costs of program**: The malaria control program at Maragra has an annual operating budget of approximately $112,000, which includes the purchase of insecticide, the wages of IRS sprayers and drivers, transportation, record-keeping, and general administrative costs.
+Assuming linearity in costs, the program spends approximately $19 per building sprayed.
+With each agregado containing an average of 2.2 workers.
 Much of the benefit of IRS goes to non-worker residents of sprayed agregados (who constitute a majority), but this benefit is purposefully ignored for this analysis.
 
-Given the likelihood that clinical data does not fully capture all malaria cases, we do not quantify the costs of malaria infection to the company.
+**Cost of malaria**: Given the likelihood that clinical data does not fully capture all malaria cases (and most likely captures only a small fraction of actual infections, given the high rates of acumulated adult immunity [Mayor2007]), we do not quantify the costs of malaria infection to the company.
 Rather, we first estimate the reduction in absenteeism attributible to IRS, and then quantify the savings associated with prevented absences.
@@
 
-**Effect of IRS on absenteeism**: IRS is associated with a year-long, significant reduction in absenteeism during the malaria season (figure 2).
-As one would expect if the mechanism by which IRS reduces absence is through reduced malaria infection, the effect of IRS during the low transmission season is significant, but far less substantial in effect size.
+**Effect of IRS on absenteeism**: Immediately following IRS, a worker's likelihood of absence drops significantly (figure x, panel i).
+As one would expect if the mechanism by which IRS reduces absence is through reduced malaria infection, the effect of IRS during the low transmission season is significant, but far less substantial in effect size (figure x, panel ii).
 
-```{r, fig.height = 3.5, fig.width = 6, fig.cap = 'Estimated effect of IRS on absenteeism by season'}
-# Predict for visualization
-dummy <- expand.grid(season = sort(unique(model_data$season)),
-                     months_since = sort(unique(model_data$months_since)),
-                     on_site = TRUE)
-predictions <- predict(fit, dummy, se.fit = TRUE, interval = 'confidence', level = 0.99)
-dummy$predicted <- predictions$fit[,1]
-dummy$lwr <- predictions$fit[,2]
-dummy$upr <- predictions$fit[,3]
-ggplot(data = dummy %>%
-         filter(months_since != 'Never') %>%
-         mutate(lwr = lwr * 100,
-                upr = upr * 100,
-                predicted = predicted * 100) %>%
-         mutate(season = paste0(Hmisc::capitalize(as.character(season)), ' season')),
+```{r, fig.height = 3.5, fig.width = 6, fig.cap = 'i.
+Absenteeism before and after IRS administration for all workers who ever received IRS; ii.
+The same, but segregated by malaria and non-malaria seasons'}
+plot_data <-
+  ab_panel %>%
+  # filter(oracle_number %in% matched$oracle_number) %>%
+  left_join(irs, c("date", "unidade")) %>%
+  mutate(months_since = days_since %/% 30) %>%
+  filter(!is.na(months_since)) %>%
+  # mutate(time_period = lendable::time_period_extract(date)) %>%
+  # mutate(time_period = as.character(time_period)) %>%
+  mutate(time_period = make_season(date = date)) %>%
+  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))
+
+x <- plot_data %>%
+  group_by(months_since) %>%
+  summarise(absences = length(which(absent)),
+            sick_absences = length(which(absent_sick)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100) %>%
+  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100)
+
+g1 <- ggplot(data = x,
        aes(x = months_since,
-           y = predicted,
-           group = 1)) +
-  geom_ribbon(aes(x = months_since,
-                  ymin = lwr,
-                  ymax = upr),
-              fill = 'lightblue',
-              alpha = 0.6) +
-  # geom_line(color = 'blue', alpha = 0.5) +
-  geom_point(color = 'blue', alpha = 0.5) + 
-  geom_line(color = 'blue', alpha = 0.5) +
-  # geom_smooth(se = FALSE,
-  #             color = 'black',
-  #             # span = 10,
-  #             alpha = 0.5) +
-  facet_wrap(~season) +
-  theme_maragra() +
-  theme(axis.text.x = element_text(angle = 90, size = 8)) +
-  labs(x = 'Months since IRS',
-       y = 'Absenteeism (%)')
+           y = absenteeism_rate)) +
+  geom_step(color = 'darkred', alpha = 0.8) +
+  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
+  labs(x = 'Months relative to IRS',
+       y = 'Absenteeism rate') +
+  theme_maragra()
+
+plot_data2 <-
+  ab_panel %>%
+  # filter(oracle_number %in% matched$oracle_number) %>%
+  left_join(irs, by = c("date", "unidade")) %>%
+  mutate(months_since = days_since %/% 30) %>%
+  filter(!is.na(months_since)) %>%
+  # mutate(time_period = lendable::time_period_extract(date)) %>%
+  # mutate(time_period = as.character(time_period)) %>%
+  mutate(time_period = make_season(date = date)) %>%
+  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))
+
+x2 <- plot_data2 %>%
+  group_by(months_since, time_period) %>%
+  summarise(absences = length(which(absent)),
+            sick_absences = length(which(absent_sick)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100) %>%
+  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100) %>%
+  mutate(before = months_since < 0)
+
+g2 <- ggplot(data = x2 %>% mutate(time_period = paste0(time_period)),
+       aes(x = months_since,
+           y = absenteeism_rate)) +
+  geom_step(color = 'darkred', alpha = 0.8) +
+  facet_wrap(~time_period, ncol = 1) +
+  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
+  labs(x = 'Months relative to IRS',
+       y = 'Absenteeism rate') +
+  theme_maragra()
+Rmisc::multiplot(g1, g2, cols = 2)
 ```
 
-We create 4 worker fixed effects models. Different models for field vs not field, permanent vs temporary.
+In order to assess the effect of IRS on absenteeism, we created 4 worker fixed effects models for the 4 principal worker types (permanent field worker, temporary field worker, permanent non-field worker, temporary non-field worker).
+The reason for segregating models rather than incorporating worker type as a fixed effect was that it seemed plausible that the effect of the treatment (IRS) on the outcome would be different (and vary differently over time) depending on these worker types.
+For instance, it is reasonable to assume that IRS' effect would be differential for a temporary worker (who likely lives in the fumigated house less than 100% of the year) than for a permanent worker.
+
+Given that IRS is carried out on a continuous rolling basis, we used relative worker time (as opposed to calendar) as the time component of our model - that is, each workers' days until and since IRS were standardized so that day 0 (date of fumigation) was the interruption point.
+This has the advantage of incorporating into the model a variety of potential time-varying confounders in a quasi-randomized fashion, thereby making it unecessary to adjust for them specifically.
+Our model took into account workers' residential location (on or off site) - even though off site workers by definition could not be administered IRS by Maragra - so as to capture other seasonal effects from those workers.
+\todo{This is not very eloquent yet and needs to be improved.}
+
+The model results (table 3) suggest a significant decrease in absenteeism following the administration of IRS.
+The effect is ambivalent among temporary workers, whereas among permanent workers the administration of IRS decreases absenteeism betwen 2.5 and 8.4 percentage points in the 2-9 month following administration.
+These effects are consistent with IRS decreasing absence through the medium of malaria, since (a) the effect is not as strong in the first month (potentially during the parasite incubation phase), the effect is stronger on field workers (who are more socioeconomically the demographic at greater risk of malaria than their administrative counterparts), and the effect is ambivalent among temporary workers (since the protective effect of IRS is only a function of the amount of time that the worker spends in the fumigated house).
 
@@
 
+## Externalities
+
+**Translating absences to costs**
+
+xxx\todo{Not done yet. This will be relatively straightforward arithmetic, but I want to first make sure that the models are rock solid (since all of the "translation" hinges on them).}
+
 ## Return on investment
 
-Details here on ROI calculation outcomes \todo{Not done yet. In conversation with Eduardo from MC about getting more accurate costs}
+\todo{Not done yet. In conversation with Eduardo from MC about getting more accurate costs}
+
+**Savings**
+
+- In percentage point terms, reduction from 13% to 8%.  
+- 5 annually prevented absences per worker.   
+- 8,000 workers: 40,000 prevented absences, wage of 3 USD
+- TOTAL: 120,000 USD in productivity-only savings.
+
+**Costs**
+
+- 8 IRS workers, 1500 USD yearly = 12,000 USD  
+- ACT + DDT: 50,000 USD 
+- Facilities, vehicles, gas: 50,000 USD  
+- TOTAL: 112,000 USD in IRS-only costs
+
+**7% ROI** (ignoring clinical costs)
+
+\newpage
 
@@
 Two principal concerns call into question the results of our analysis.
-First, the application of IRS to a workers house may be endogenous.
-To check for this  bla bla bla \todo{As per Menno's suggestion, will add details here pertaining to whether IRS predicts absenteeism before it is applied}.
-The second concern is that our quantification of consists is distorted by the fact that we treat IRS operations as essentially linear in nature, when in reality economies of scale, in-kind purchases and other factors likely make the true cost-per-spraying convex.\todo{Will address this with some sensitivity analysis}
+First, the application of IRS to a workers house may be endogenous.\todo{Is this the right term here?
+We're not dealing with an "unobserved" factor per se, rather we're dealing with a situation in which the cause might influence the effect, and the effect might influence the cause, ie a feedback loop.} It is reasonable to suspect that the application of IRS to households is not random, but rather that IRS was applied more frequently to houses which had already seen a malaria case.
+In this case, our estimated effect of IRS on absenteeism would likely be underestimated, with the post-IRS absenteeism rates actually having declined from a greater pre-IRS absenteeism rate than otherwise suggested.
+
+To check for this, we estimate the odds of absenteeism as a function of receiving IRS _during the 10 day period prior to receiving IRS_.
+If IRS application is indeed endogenous, we would expect absenteeism to be elevated during this period (since the increase in absenteeism would be theoretically responsible for the application of IRS), a situation which would require further statistical adjustment.
+If, on the other hand, there is no endogeneity, we would expect absenteeism in the 10 day period prior to IRS administration to be similar to other pre-IRS absenteeism (adjusting for seasonality).
+
+What follows are several variations of a model to assess whether IRS application predicts pre-IRS absenteeism.
+In the final version of the paper, we'll use whichever model is most appropriate, but this is left here for discussion purposes.
+I've also included the code, so as to inform the discussion.
+In the below models, `pre_irs` refers to the 10 day period prior to IRS administration.
+The models estimate the odds of absence (ie, the estimates and confidence intervals are in the form of odds ratios).
+
+```{r, echo = TRUE}
+ro_data <- ab_panel %>%
+  left_join(irs, by = c('date', 'unidade')) %>%
+  filter(days_since < 0) %>%
+  mutate(pre_irs = between(days_since, -10, -1)) %>%
+  # mutate(time_period = lendable::time_period_extract(date)) %>%
+  # mutate(time_period = as.character(time_period)) %>%
+  mutate(time_period = make_season(date = date)) %>%
+  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) %>%
+  left_join(workers %>%
+              dplyr::select(oracle_number,
+                            sex,
+                            department,
+                            permanent_or_temporary),
+            by = 'oracle_number')
+
+make_model <- function(formula = 'absent ~ pre_irs + time_period'){
+  ro_model <- glm(as.formula(formula), 
+                data = ro_data,
+                family = binomial('logit'))
+  coefs <- exp(coef(ro_model))
+  ci <- confint(ro_model)
+  ci <- exp(ci)
+  out <- data.frame(Variable = names(coefs),
+                    Estimate = coefs,
+                    Lower = ci[,1],
+                    Upper = ci[,2])
+  row.names(out) <- NULL
+  out$`P value` <- coef(summary(ro_model))[,4]
+  out$Significant <- ifelse(out$`P value` <= 0.001, 'xxx',
+                            ifelse(out$`P value` <= 0.01, 'xx',
+                                   ifelse(out$`P value` <= 0.05, 'x',
+                                          '')))
+  return(kable(out, format = "latex") %>%
+  kable_styling() %>%
+    row_spec(row = which(grepl('pre_irs', out$Variable)),background = "yellow"))
+}
+```
+
+```{r, echo = TRUE, results = 'asis'}
+
+# BASIC FORMULA
+make_model('absent ~ pre_irs + time_period')
+
+# BASIC FORMULA WITH INTERACTIONS
+make_model('absent ~ pre_irs*time_period')
+
+# BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX
+make_model('absent ~ pre_irs*time_period + sex')
+
+# BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX AND DEPARTMENT
+make_model('absent ~ pre_irs*time_period + sex + department')
+
+# BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX AND DEPARTMENT AND STATUS
+make_model('absent ~ pre_irs*time_period + sex + department + permanent_or_temporary')
+
+```
+
+The second concern is that our quantification of return on investment is distorted by the fact that we treat IRS operations as essentially linear in nature, when in reality economies of scale, in-kind purchases and other factors likely make the true cost-per-spraying convex.
+We address this by creating three scenarios: (1) a "start-up" scenario in which we take into account all costs incurred by starting a program from scratch, (2) a "normal" scenario in which we match the assumptions with those used in this paper (ie, account for "wear-and-tear" depreciation but not vehicle purchase, etc.) and a (3) "absorbed costs" scenario which ignores all costs which are not directly incurred by the program....
+xxx...
+\todo{Will address this with some sensitivity analysis}

### 2018-04-18 inst/rmd/04_discussion.Rmd
replacement felm

@@
 
-Overall stuff
+\todo{This entire section will not be done until methods/results are finalized.}
+
+- Overview of findings
+
+- How this contributes to the literature
+
+- Implications for policy and for businesses
 
@@
 
-## Implications
+- No analysis yet of different worker types (agricultural vs. industrial).
+- Have not yet ventured at all into side-analyses (effect on employment, tonnage, etc.).
+- Sick absenteeism seems to track absenteeism poorly: lack of clarity regarding pathways.
+- Large sample size, but all from same place: questionable generalizability.

### 2018-04-18 inst/rmd/paper.Rmd
replacement felm

@@
 source('paper_prepare.R')
+irs$months_since <- irs$days_since %/% 30
+
+# Define function for making season
+make_season <- function(date){
+  out <- ifelse(as.numeric(format(date, '%m')) %in% c(11:12, 1:3),
+                   'Malaria season', 'Off season')
+  out <- factor(out, levels = c('Off season', 'Malaria season'))
+  return(out)
+}
 ```
@@
 ```
+
 ```{r child='children.Rmd', echo = FALSE}
 ```
+
 ```{r, results='hide'}

### 2018-04-18 inst/rmd/paper.tex
major biblio update
replacement felm

@@
 \urlstyle{same}  % don't use monospace font for urls
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
 \usepackage{graphicx,grffile}
@@
 Elton Dorkin\footrecall{ma}
-Eduardo Nhamahanga\footrecall{ma}
 Menno Pradhan\footrecall{vu}\footremember{uva}{University of Amsterdam: REC E, Roetersstraat 11, Amsterdam, Netherlands, Netherlands}
@@
 and Malaney, 2002), malaria likely has a negative effect on GDP and
-growth (McCarthy et al., 2000) (Orem et al., 2012). Because of the
+growth (McCarthy et al., 2000a) (Orem et al., 2012). Because of the
 relative affordability of most intereventions and the enormous societal
@@
 
-\subsection{Background Literature}\label{background-literature}
-
-\todo{Will add 3-5 paragraphs of background lit review}
-
-Our study adds to the existing literature in several ways.
-\improvement{Will add details here.}
+The literature on the effect of sugarcane cultivation on malaria risk is
+mixed. While some studies have found that the prevalence of malaria
+vectors in sugarcane areas to be similar to that of uncultivated areas
+(and less than in areas dedicated to other forms of more water-inensive
+agriculture, such as rice) (Ijumba et al., 2002), other studies have
+found significant increases in factors associated with malaria
+transmission at large-scale sugarcane facilities relative to
+traditional, small-scale farming and non-irrigated farming (Jaleta et
+al., 2013). Regardless of the effect the presence of a sugarcane farm
+\emph{per se} on local malaria epidemiology, the time spent outdoors by
+sugarcane workers, the fact that many workers are migrants, and their
+sometimes precarious housing situations, suggest that sugarcane farmers
+are likely at increased risk of malaria infection (O'Laughlin, 2016).
+This is important, given that even among occupations with far less
+inherent exposure to mosquitoes (such as health professionals), malaria
+is one of the primary causes of work absenteeism in malaria-endemic
+countries (Burton et al., 1999). There is also some concern regarding
+the effect of large-scale insecticide use - common at essentially all
+Sub-Saharan African sugarcane farms - on insecticide resistance among
+moquitoes in the area. A study in Belize found that mosquito populations
+on the edge of sugarcane fields had higher tolerance to insecticide than
+similar populations in the core of fields or outside of the periphery
+(Dusfour et al., 2009). Sugarcane areas may offer the standing water
+necessary for mosquito breeding, but also perhaps attract mosquitoes
+which would otherwise be elsewhere, due to compounds in sugarcane pollen
+(Wondwosen et al., 2018).
+
+Regardless of the academic literature on the risk of workers becoming
+infected with malaria, the companies running large agricultural
+operations in malaria-endemic regions generally consider the disease to
+be an important enough threat to workers' health to merit some form of
+investment in malaria control (Pluess et al., 2009). Insecticide
+fumigation, bednet distribution, test-and-treat clinics, and occasional
+educational campaigns are common forms of corporate social
+responsibility to combat malaria. A World Bank study on Nigerian
+sugarcane cutters found that the simple availability of testing and
+treatment increased productivity by 10\% in the weeks following the
+provision of services, the conclusion being that both the treatment and
+the test result were effective in increasing productivity, the latter
+simply increasing the information available which could influence
+personal labor allocation decisions (Dillon et al., 2014).
+
+Malaria may disincentivize investment in business in malaria zones. In a
+vicious cycle, businesses already in malaria zones may not be
+incentivized to invest in malaria control, perhaps because this is seen
+as the domain of the public sector (Curtis et al., 2003), or perhaps due
+to not knowing the extent to which the benefits in health of that
+control will accrue to the business itself.
+
+The evidence of malaria's negative effects on both health and wealth, at
+both the individual (Cole and Neumayer, 2006) and collective (McCarthy
+et al., 2000b) levels, as well as in both the short (Asenso-Okyere and
+Dzator, 1997) (Ajani and Ashagidigbi, 2010) and long (Hong, 2011) (Sachs
+and Malaney, 2002) terms, are amply described in the public health and
+economics literature (Phillips, 1998). That said, very little exists in
+the literature examining the costs and benefits of malaria control from
+a private ledger perspective (ie, the point of view a business
+investor). Unlike a government or individual, a private firm investing
+in malaria control may be most interested not in its long-term
+macroeconomic effects, nor its short-term personal health effects, but
+rather on the impact on productivity (and the extent to which that
+producitivty's benefits accrue to the firm), absenteeism, and the
+opportunity costs of expenditures in malaria control. In other words,
+though the large magnitude of malaria control's benefits are well known,
+the portion of those benefits accrued by a private firm investing in
+malaria control is unknown.
+
+This study does not endeavor to expand the current body of knowledge
+regarding malaria's ill effects on individuals and societies; rather, it
+aims to provide empirical evidence pertaining to a facet of malaria
+economics with very little in the literature: malaria control from a
+private-sector investment perspective. Our study adds to the existing
+literature by showing the effect of specific malaria control
+interventions on worker absenteeism, and translating that effect into a
+return on investment. Unlike previous studies on the effectiveness of
+malaria control interventions, our's focuses solely on one firm carrying
+out one intervention, takes advantage of individual-level data, and
+analyzes results from a ledger perspective.
 
@@
 
+\begin{figure}[H]
+
+{\centering \includegraphics{figures/unnamed-chunk-13-1}
+
+}
+
+\caption{(i) Mozambique in Africa, (ii) Districts of Mozambique with Manhiça highlighted in red, (iii) District of Manhiça with Maragra highlighted in red, (iv) Maragra SA with surrounding fields and village}\label{fig:unnamed-chunk-13}
+\end{figure}
+
 The mill of the Maragra Açucar SA (a subsidary of the Illovo sugar
@@
 
-Maragra is located in the district of Manhiça, a semi-rural area in
-Southern Mozambique
+Maragra (figure 1, panel iii) is located in the district of Manhiça
+(figure 1, panel ii), a semi-rural area in the south of Mozambique
+(figure 1, panel i). 80 kilometers north of the Mozambican capital of
+Maputo, the district is low-lying, consisting largely of savannah and
+wetlands along the Incomati River. Most of the areas 160,000 residents
+(Sacoor et al., 2013) work as subsistence farmers. Migration from the
+area to South Africa for the purpose of employment in the profitable
+construction industry is common, especially among men (Nhacolo et al.,
+2006), as is migration to the area (from other parts of Mozambique) for
+seasonal work on the sugarcane plantations at Maragra and the slightly
+larger facility in Xinavane (at the district's border with Magude)
+(O'Laughlin, 2016).
+
+Poverty is rife in southern Mozambique, and its associated illnesses
+take their toll on the population. The community prevalence of HIV/AIDS
+is as high as 40\% (González et al., 2012); even in a more recent study
+suggesting a much lower prevalence of 22\%, the risk of infection is
+still twice that of nearby areas (Mocumbi et al., 2017). Recent years
+have seen a three-fold increase in tuberculosis (García-Basteiro et al.,
+2017). Malaria, which has the greatest mortality burden due to the fact
+that the young are especially vulnerable to its effects, is perennial,
+though worse during the rainy season (December - March) (Saúte et al.,
+2003). Adult malaria is essentially universal, albeit much of it
+asymptomatic (Mayor et al., 2007). In regards to worker health, malaria
+is Maragra's primary concern, being so important as to justify the
+existence of both an on-site testing laboratory and clinic, as well as a
+malaria control department.
+
+Maragra workers are mostly seasonal, working for the firm approximately
+half of the year during harvest time, and cultivating crops, working in
+construction, or going unemployed (or working elsewhere) during the
+other half. Though many workers live ``on-site'' (ie, within the
+delineated property of the firm itself), a sizeable minority resides in
+the environs (figure 1, panel iv). Maragra provides indoor residual
+spraying (IRS) using ACT (alpha-cypermethrin) and DDT
+(Dichlorodiphenyltrichloroethane) depending on stocks (a preference
+apparently exists for the former, but it is not always available). IRS
+activities, managed by Maragra's Malaria Control program, are ongoing
+throughout the year (Figure 1, panel i). Nearly all on-site houses are
+sprayed, though the time between fumigations is irregular (Figure 2,
+panel ii). Off-site houses may also receive IRS (managed and carried out
+by the National Malaria Control Program), but the status and timing of
+these fumigations is not known at the individual level.
 
@@
 
-{\centering \includegraphics{figures/unnamed-chunk-13-1}
+{\centering \includegraphics{figures/unnamed-chunk-14-1}
 
@@
 
-\caption{(i) Mozambique in Africa, (ii) Districts of Mozambique with Manhiça highlighted in red, (iii) District of Manhiça with Maragra highlighted in red, (iv) Maragra SA with surrounding fields and village}\label{fig:unnamed-chunk-13}
+\caption{i. Fumigation activities carried out by Maragra Malaria Control during study period, ii. Distribution of average time between sprayings of households}\label{fig:unnamed-chunk-14}
 \end{figure}
-
-Details on Manhica
-
-Details on Mozambique
 
@@
 malaria. To estimate this effect, we estimated separate models for
-absence and illness. We employed interrupted time series (Lopez Bernal
-et al., 2016) and a linear probability approach using the following
-econometric model.
+absence and illness. We employed linear mixed effects models (Minalu et
+al., 2011) (Liu et al., 2010) (with individual worker fixed effects, and
+random effects for other covariates) in lieu of interrupted time series
+(Lopez Bernal et al., 2016) so as to explicitly adjust for confounders
+(Bell and Jones, 2014). We divide into 4 different models for different
+worker types, so as to account for the potential time-varying effects of
+worker type on risk of malaria. Our econometric model is relatively
+straightforward:
 
 \[
-\hat{Y_{it}} = \beta_{0} +  (\beta_{1}) (\text{Season}_{t}) + (\beta_2{IRS}*\beta_3{IRS_t}) + ... + \epsilon
+\hat{Y_{it}} = \beta_{0} +  (\beta_{1}) (\text{Season}_{it}) + (\beta_2{IRS}*\beta_3{IRS_it}) + ... + \epsilon_{it}
 \]
@@
 age, and worker department (field, factory, or administrative). Our
-intervention was not a simple yes/no, but rather the product of whether
-the residence of the worker in question was treated in the last year,
-and, if so, the time since treatment (represented above as the
-interaction term, where where \(_t\) represents time elapsed since
-commencement of the most recent IRS campaign). We define the malaria
-season as any time during which the clinical incidence of malaria in the
-district of Manhiça was at or greater than the median clinical incidence
-of malaria for the entire study period. These weeks are flagged as red
-in Figure 1, Panel A. By using clinical incidence of the area of
-residence of the workers (as opposed to more typical proxies for malaria
-risk, such as only rainy vs.~non rainy season), our seasonality estimate
-is a closer approximation of true malaria risk, incorporating lagged
-effects such as the incubation period of the parasite, as well as any
-inherent non-linear effects of weather. In addition, we adjust for daily
-precipitation
+intervention whether the residence of the worker in question was treated
+in the last year, and, if so, the time since treatment (represented
+above as the interaction term, where where \(_t\) represents time
+elapsed since commencement of the most recent IRS campaign). We define
+the malaria season as any time during which the clinical incidence of
+malaria in the district of Manhiça was at or greater than the median
+clinical incidence of malaria for the entire study period. These weeks
+are flagged as red in Figure 1, Panel A. By using clinical incidence of
+the area of residence of the workers (as opposed to more typical proxies
+for malaria risk, such as only rainy vs.~non rainy season), our
+seasonality estimate is a closer approximation of true malaria risk,
+incorporating lagged effects such as the incubation period of the
+parasite, as well as any inherent non-linear effects of weather. In
+addition, we adjust for daily precipitation
+
+Details will go here about externalities (neighborhood effects, positive
+spillover to district, negative spillover from district, how large firm
+may mean that most positive externalities are absorbed by the firm, and
+not many negative externalities, but maybe with a smaller firm it
+wouldn't work as well)\todo{NOT DONE YET.}
 
@@
 Ethical approval for this project was obtained from the Institutional
-Ethics Review Board for Health at the CISM prior to data collection.
+Ethics Review Board for Health at the CISM (CIBS-CISM) prior to data
+collection.
 
@@
 
-Age, types of workers, bla bla bla.
-\todo{Will add a table here and some overview}
+We collected absenteeism and demographic data data on 3362 workers from
+2012 through 2016. Workers were approximately 60\% male, and more than
+80\% fieldworkers (predominantly cane-cutters). Most were in their 20s
+and 30s and employed on temporary contracts. Table 1 shows overall
+worker details.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rlll}
+  \hline
+Variable &   & Workers & (\%) \\ 
+  \hline
+Department & Administrative & 157 & 4.67 \\ 
+   & Factory & 459 & 13.65 \\ 
+   & Field & 2746 & 81.68 \\ 
+  Sex & F & 1305 & 38.82 \\ 
+   & M & 2057 & 61.18 \\ 
+  Status & Permanent & 500 & 14.87 \\ 
+   & Temporary & 2862 & 85.13 \\ 
+  Age & 20 & 985 & 29.30 \\ 
+   & 30 & 1244 & 37.00 \\ 
+   & 40 & 665 & 19.78 \\ 
+   & 50 & 347 & 10.32 \\ 
+   & 60 & 117 & 3.48 \\ 
+   & 70 &   4 & 0.12 \\ 
+   \hline
+\end{tabular}
+\caption{Overall worker characteristics} 
+\end{table}
+
+The overall absenteeism rate during the study period was 5.56\% among
+all workers. The number of malaria cases registered at the company
+clinic during that time was 1873, for an approximate annualized clinical
+incidence of 110 cases per 1,000.
+
+Temporary workers tended to be younger and male; female temporary
+workers being on average 5-10 years older than their male counterparts.
+Permanent workers had a bi-modal age distribution, the older peak
+explained by the greater density of males in administrative roles.
+Females accounted for 43\% of temporary workers but only 16\% of
+permanent workers (figure x).
+
+\begin{figure}[H]
+
+{\centering \includegraphics{figures/unnamed-chunk-17-1}
+
+}
+
+\caption{Age distribution by sex and worker status}\label{fig:unnamed-chunk-17}
+\end{figure}
+
+During the study period, weather followed typical seasonal trends for
+the area, albeit slightly drier than previous periods.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{figures/unnamed-chunk-18-1}
+
+}
+
+\caption{i.
+Monthly total precipitation in the Maniça district; ii.
+Average monthly temperature (bars) during the study period, as well as monthly maximum (red) and minimum (blue) temperatures}\label{fig:unnamed-chunk-18}
+\end{figure}
 
@@
 In Southern Mozambique, malaria peaks during the summer months (December
-through March) most years (Figure 1, panel A), and worker absenteeism
-rates track malaria incidence closely, following the same seasonal
-patterns (Figure 1, panel B). Both all-cause absenteeism and sick
-absenteeism have declined in recent years at Maragra (Figure 1, panel
-C), with the latter declining at a faster rate than the former. The fact
-that the rate of confirmed cases at the company clinic is largely
-non-seasonal (Figure 1, panels E and F) suggests that a significant
-portion of workers either seek care for malaria elsewhere (for example,
-government health posts, of which several are nearby and in some many
-cases closer to workers' residence than the company clinic) or do not
-seek care during malaria infection. Accordingly, we focus our analysis
-on all-cause absenteeism rather than sickness absenteeism or malaria
-diagnostics, with the assumption that much of illness is captured by
-absenteeism but not by the clinical data.
+through March) most years (Figure 4, panel A), and worker absenteeism at
+Maragra rates track malaria incidence closely, following the same
+seasonal patterns (figure 4, panel B). Both all-cause absenteeism and
+sick absenteeism have declined in recent years at Maragra (figure 4,
+panel C), with the latter declining at a faster rate than the former.
+The fact that the rate of confirmed cases at the company clinic is
+largely non-seasonal (figure 4, panels E and F) suggests that a
+significant portion of workers either seek care for malaria elsewhere
+(for example, government health posts, of which several are nearby and
+in some many cases closer to workers' residence than the company clinic)
+or do not seek care during malaria infection. Accordingly, we focus our
+analysis on all-cause absenteeism rather than sickness absenteeism or
+malaria diagnostics, with the assumption that much of illness is
+captured by absenteeism but not by the clinical data.
 
@@
 \centering
-\includegraphics{figures/unnamed-chunk-14-1.pdf}
+\includegraphics{figures/unnamed-chunk-19-1.pdf}
 \caption{Clinical malaria (district of Manhiça), all-cause absenteeism
-among Maragra workers, sick absenteeism among Maragra workers, estimated
-rainfall, positive cases at company clinic, and test positivity rate at
-company clinic}
+among Maragra workers, sick absenteeism among Maragra workers, positive
+cases at company clinic, and test positivity rate at company clinic}
 \end{figure}
@@
 
-\textbf{Costs}: The malaria control program at Maragra has an annual
-operating budget of approximately XX, which includes the purchase of
-insecticide, the wages of IRS sprayers and drivers, transportation,
-record-keeping, and general administrative costs. Assuming linearity in
-costs, the program spends approximately XX per agregado sprayed. With
-each agregado containing an average of 2.2 workers, this translates to a
-cost of XX per worker protected per season. Much of the benefit of IRS
-goes to non-worker residents of sprayed agregados (who constitute a
-majority), but this benefit is purposefully ignored for this analysis.
-
-Given the likelihood that clinical data does not fully capture all
-malaria cases (and most likely captures only a small fraction of actual
-infections, given the high rates of acumulated adult immunity
-{[}Mayor2007{]}), we do not quantify the costs of malaria infection to
-the company. Rather, we first estimate the reduction in absenteeism
-attributible to IRS, and then quantify the savings associated with
-prevented absences. Additionally, we calculate the clinical savings of
-IRS by first estimating the share of absences which are associated with
-an episode of clinical malaria, and then applying the clinical cost per
-case to the equivalent share of prevented absences. We intentionally
-ignore the savings accrued by the public health system, as well as the
-likely utility gains in secondary realms such as school absenteeism,
-producitivity, etc.
-
-\textbf{Effect of IRS on absenteeism}: IRS is associated with a
-year-long, significant reduction in absenteeism during the malaria
-season (figure 2). As one would expect if the mechanism by which IRS
-reduces absence is through reduced malaria infection, the effect of IRS
-during the low transmission season is significant, but far less
-substantial in effect size.
+\textbf{Effect of IRS on clinical malaria}: xxx.
+\todo{Not done yet. In fact, not even sure if I should do this section at all, since it's basically just a distraction.}
+
+\textbf{Costs of program}: The malaria control program at Maragra has an
+annual operating budget of approximately \$112,000, which includes the
+purchase of insecticide, the wages of IRS sprayers and drivers,
+transportation, record-keeping, and general administrative costs.
+Assuming linearity in costs, the program spends approximately \$19 per
+building sprayed. With each agregado containing an average of 2.2
+workers. Much of the benefit of IRS goes to non-worker residents of
+sprayed agregados (who constitute a majority), but this benefit is
+purposefully ignored for this analysis.
+
+\textbf{Cost of malaria}: Given the likelihood that clinical data does
+not fully capture all malaria cases (and most likely captures only a
+small fraction of actual infections, given the high rates of acumulated
+adult immunity {[}Mayor2007{]}), we do not quantify the costs of malaria
+infection to the company. Rather, we first estimate the reduction in
+absenteeism attributible to IRS, and then quantify the savings
+associated with prevented absences. Additionally, we calculate the
+clinical savings of IRS by first estimating the share of absences which
+are associated with an episode of clinical malaria, and then applying
+the clinical cost per case to the equivalent share of prevented
+absences. We intentionally ignore the savings accrued by the public
+health system, as well as the likely utility gains in secondary realms
+such as school absenteeism, producitivity, etc.
+
+\textbf{Effect of IRS on absenteeism}: Immediately following IRS, a
+worker's likelihood of absence drops significantly (figure x, panel i).
+As one would expect if the mechanism by which IRS reduces absence is
+through reduced malaria infection, the effect of IRS during the low
+transmission season is significant, but far less substantial in effect
+size (figure x, panel ii).
 
@@
 \centering
-\includegraphics{figures/unnamed-chunk-17-1.pdf}
-\caption{Estimated effect of IRS on absenteeism by season}
+\includegraphics{figures/unnamed-chunk-22-1.pdf}
+\caption{i. Absenteeism before and after IRS administration for all
+workers who ever received IRS; ii. The same, but segregated by malaria
+and non-malaria seasons}
 \end{figure}
 
-We create 4 worker fixed effects models. Different models for field vs
-not field, permanent vs temporary.
+In order to assess the effect of IRS on absenteeism, we created 4 worker
+fixed effects models for the 4 principal worker types (permanent field
+worker, temporary field worker, permanent non-field worker, temporary
+non-field worker). The reason for segregating models rather than
+incorporating worker type as a fixed effect was that it seemed plausible
+that the effect of the treatment (IRS) on the outcome would be different
+(and vary differently over time) depending on these worker types. For
+instance, it is reasonable to assume that IRS' effect would be
+differential for a temporary worker (who likely lives in the fumigated
+house less than 100\% of the year) than for a permanent worker.
+
+Given that IRS is carried out on a continuous rolling basis, we used
+relative worker time (as opposed to calendar) as the time component of
+our model - that is, each workers' days until and since IRS were
+standardized so that day 0 (date of fumigation) was the interruption
+point. This has the advantage of incorporating into the model a variety
+of potential time-varying confounders in a quasi-randomized fashion,
+thereby making it unecessary to adjust for them specifically. Our model
+took into account workers' residential location (on or off site) - even
+though off site workers by definition could not be administered IRS by
+Maragra - so as to capture other seasonal effects from those workers.
+\todo{This is not very eloquent yet and needs to be improved.}
+
+The model results (table 3) suggest a significant decrease in
+absenteeism following the administration of IRS. The effect is
+ambivalent among temporary workers, whereas among permanent workers the
+administration of IRS decreases absenteeism betwen 2.5 and 8.4
+percentage points in the 2-9 month following administration. These
+effects are consistent with IRS decreasing absence through the medium of
+malaria, since (a) the effect is not as strong in the first month
+(potentially during the parasite incubation phase), the effect is
+stronger on field workers (who are more socioeconomically the
+demographic at greater risk of malaria than their administrative
+counterparts), and the effect is ambivalent among temporary workers
+(since the protective effect of IRS is only a function of the amount of
+time that the worker spends in the fumigated house).
 
@@
 
-\caption{\label{tab:unnamed-chunk-18}Models with worker fixed effects}
+\caption{\label{tab:unnamed-chunk-23}Models with worker fixed effects}
 \centering
@@
 \multicolumn{2}{l}{\textbf{Permanent field worker}}\\
-\hspace{1em}Off site & 9.142 (P<0.001)\\
-\hspace{1em}On site & 11.424 (P<0.001)\\
 \hspace{1em}Malaria season & 3.424 (P<0.001)\\
-\hspace{1em}Months since IRS: 01 & 3.831 (P<0.001)\\
-\hspace{1em}Months since IRS: 02-04 & 4.199 (P<0.001)\\
-\hspace{1em}Months since IRS: 05-09 & -3.194 (P<0.001)\\
-\hspace{1em}Months since IRS: 10+ & -5.794 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & -4.534 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & -8.392 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & -3.85 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 10+ & 4.504 (P=0.002)\\
+\hspace{1em}Months since IRS: 01 & 3.83 (P<0.001)\\
+\hspace{1em}Months since IRS: 02-04 & 4.206 (P<0.001)\\
+\hspace{1em}Months since IRS: 05-09 & -3.175 (P<0.001)\\
+\hspace{1em}Months since IRS: 10+ & -5.775 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & -4.529 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 02-04 & -8.398 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 05-09 & -3.859 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 10+ & 4.507 (P=0.002)\\
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
-\hspace{1em}Off site & 10.862 (P<0.001)\\
-\hspace{1em}On site & 9.555 (P<0.001)\\
-\hspace{1em}Malaria season & 2.077 (P<0.001)\\
-\hspace{1em}Months since IRS: 01 & -2.083 (P<0.001)\\
-\hspace{1em}Months since IRS: 02-04 & 3.904 (P<0.001)\\
-\hspace{1em}Months since IRS: 05-09 & -0.834 (P=0.084)\\
-\hspace{1em}Months since IRS: 10+ & -0.268 (P=0.742)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & 2.735 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & -5.768 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & -2.441 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 10+ & 4.497 (P<0.001)\\
+\hspace{1em}Malaria season & 2.076 (P<0.001)\\
+\hspace{1em}Months since IRS: 01 & -2.078 (P<0.001)\\
+\hspace{1em}Months since IRS: 02-04 & 3.911 (P<0.001)\\
+\hspace{1em}Months since IRS: 05-09 & -0.822 (P=0.089)\\
+\hspace{1em}Months since IRS: 10+ & -0.25 (P=0.758)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & 2.723 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 02-04 & -5.785 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 05-09 & -2.447 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 10+ & 4.501 (P<0.001)\\
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Temporary field worker}}\\
-\hspace{1em}Off site & 0.985 (P<0.001)\\
-\hspace{1em}On site & 1.078 (P<0.001)\\
-\hspace{1em}Malaria season & -0.017 (P=0.461)\\
-\hspace{1em}Months since IRS: 01 & -0.4 (P<0.001)\\
-\hspace{1em}Months since IRS: 02-04 & -0.547 (P<0.001)\\
-\hspace{1em}Months since IRS: 05-09 & -0.554 (P<0.001)\\
-\hspace{1em}Months since IRS: 10+ & -0.992 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & -0.092 (P=0.62)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & 0.172 (P=0.256)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & 0.062 (P=0.657)\\
-\hspace{1em}Malaria season:Months since IRS: 10+ & 0.068 (P=0.771)\\
+\hspace{1em}Malaria season & -0.005 (P=0.833)\\
+\hspace{1em}Months since IRS: 01 & -0.429 (P<0.001)\\
+\hspace{1em}Months since IRS: 02-04 & -0.575 (P<0.001)\\
+\hspace{1em}Months since IRS: 05-09 & -0.586 (P<0.001)\\
+\hspace{1em}Months since IRS: 10+ & -1.025 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & -0.066 (P=0.725)\\
+\hspace{1em}Malaria season:Months since IRS: 02-04 & 0.166 (P=0.274)\\
+\hspace{1em}Malaria season:Months since IRS: 05-09 & 0.063 (P=0.654)\\
+\hspace{1em}Malaria season:Months since IRS: 10+ & 0.062 (P=0.789)\\
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
-\hspace{1em}Off site & 4.136 (P<0.001)\\
-\hspace{1em}On site & 3.834 (P<0.001)\\
-\hspace{1em}Malaria season & -0.292 (P=0.16)\\
-\hspace{1em}Months since IRS: 01 & -1.456 (P=0.389)\\
-\hspace{1em}Months since IRS: 02-04 & 0.608 (P=0.574)\\
-\hspace{1em}Months since IRS: 05-09 & -0.869 (P=0.405)\\
-\hspace{1em}Months since IRS: 10+ & 10.583 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & 6.045 (P=0.002)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & 0.604 (P=0.667)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & -1.482 (P=0.284)\\
-\hspace{1em}Malaria season:Months since IRS: 10+ & -10.636 (P<0.001)\\
+\hspace{1em}Malaria season & -0.22 (P=0.294)\\
+\hspace{1em}Months since IRS: 01 & -1.466 (P=0.389)\\
+\hspace{1em}Months since IRS: 02-04 & 0.571 (P=0.6)\\
+\hspace{1em}Months since IRS: 05-09 & -1.018 (P=0.334)\\
+\hspace{1em}Months since IRS: 10+ & 10.606 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & 6.093 (P=0.002)\\
+\hspace{1em}Malaria season:Months since IRS: 02-04 & 0.476 (P=0.736)\\
+\hspace{1em}Malaria season:Months since IRS: 05-09 & -1.505 (P=0.277)\\
+\hspace{1em}Malaria season:Months since IRS: 10+ & -10.785 (P<0.001)\\
 \bottomrule
@@
 
+\subsection{Externalities}\label{externalities}
+
+\textbf{Translating absences to costs}
+
+xxx\todo{Not done yet. This will be relatively straightforward arithmetic, but I want to first make sure that the models are rock solid (since all of the "translation" hinges on them).}
+
 \subsection{Return on investment}\label{return-on-investment}
 
-Details here on ROI calculation outcomes
 \todo{Not done yet. In conversation with Eduardo from MC about getting more accurate costs}
+
+\textbf{Savings}
+
+\begin{itemize}
+\tightlist
+\item
+  In percentage point terms, reduction from 13\% to 8\%.\\
+\item
+  5 annually prevented absences per worker.\\
+\item
+  8,000 workers: 40,000 prevented absences, wage of 3 USD
+\item
+  TOTAL: 120,000 USD in productivity-only savings.
+\end{itemize}
+
+\textbf{Costs}
+
+\begin{itemize}
+\tightlist
+\item
+  8 IRS workers, 1500 USD yearly = 12,000 USD\\
+\item
+  ACT + DDT: 50,000 USD
+\item
+  Facilities, vehicles, gas: 50,000 USD\\
+\item
+  TOTAL: 112,000 USD in IRS-only costs
+\end{itemize}
+
+\textbf{7\% ROI} (ignoring clinical costs)
+
+\newpage
 
@@
 Two principal concerns call into question the results of our analysis.
-First, the application of IRS to a workers house may be endogenous. To
-check for this bla bla bla
-\todo{As per Menno's suggestion, will add details here pertaining to whether IRS predicts absenteeism before it is applied}.
-The second concern is that our quantification of consists is distorted
-by the fact that we treat IRS operations as essentially linear in
-nature, when in reality economies of scale, in-kind purchases and other
-factors likely make the true cost-per-spraying
-convex.\todo{Will address this with some sensitivity analysis}
+First, the application of IRS to a workers house may be
+endogenous.\todo{Is this the right term here?
+We're not dealing with an "unobserved" factor per se, rather we're dealing with a situation in which the cause might influence the effect, and the effect might influence the cause, ie a feedback loop.}
+It is reasonable to suspect that the application of IRS to households is
+not random, but rather that IRS was applied more frequently to houses
+which had already seen a malaria case. In this case, our estimated
+effect of IRS on absenteeism would likely be underestimated, with the
+post-IRS absenteeism rates actually having declined from a greater
+pre-IRS absenteeism rate than otherwise suggested.
+
+To check for this, we estimate the odds of absenteeism as a function of
+receiving IRS \emph{during the 10 day period prior to receiving IRS}. If
+IRS application is indeed endogenous, we would expect absenteeism to be
+elevated during this period (since the increase in absenteeism would be
+theoretically responsible for the application of IRS), a situation which
+would require further statistical adjustment. If, on the other hand,
+there is no endogeneity, we would expect absenteeism in the 10 day
+period prior to IRS administration to be similar to other pre-IRS
+absenteeism (adjusting for seasonality).
+
+What follows are several variations of a model to assess whether IRS
+application predicts pre-IRS absenteeism. In the final version of the
+paper, we'll use whichever model is most appropriate, but this is left
+here for discussion purposes. I've also included the code, so as to
+inform the discussion. In the below models, \texttt{pre\_irs} refers to
+the 10 day period prior to IRS administration. The models estimate the
+odds of absence (ie, the estimates and confidence intervals are in the
+form of odds ratios).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{ro_data <-}\StringTok{ }\NormalTok{ab_panel }\OperatorTok{%>%}\StringTok{ }\KeywordTok{left_join}\NormalTok{(irs, }\DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"date"}\NormalTok{, }\StringTok{"unidade"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(days_since }\OperatorTok{<}\StringTok{ }
+\StringTok{    }\DecValTok{0}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{pre_irs =} \KeywordTok{between}\NormalTok{(days_since, }\OperatorTok{-}\DecValTok{10}\NormalTok{, }\OperatorTok{-}\DecValTok{1}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }\CommentTok{# mutate(time_period = lendable::time_period_extract(date)) %>%}
+\CommentTok{# mutate(time_period = as.character(time_period)) %>%}
+\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{time_period =} \KeywordTok{make_season}\NormalTok{(}\DataTypeTok{date =}\NormalTok{ date)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{absent_sick =} \KeywordTok{ifelse}\NormalTok{(}\KeywordTok{is.na}\NormalTok{(absent_sick), }
+    \OtherTok{FALSE}\NormalTok{, absent_sick)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{left_join}\NormalTok{(workers }\OperatorTok{%>%}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(oracle_number, }
+\NormalTok{    sex, department, permanent_or_temporary), }\DataTypeTok{by =} \StringTok{"oracle_number"}\NormalTok{)}
+
+\NormalTok{make_model <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(}\DataTypeTok{formula =} \StringTok{"absent ~ pre_irs + time_period"}\NormalTok{) \{}
+\NormalTok{    ro_model <-}\StringTok{ }\KeywordTok{glm}\NormalTok{(}\KeywordTok{as.formula}\NormalTok{(formula), }\DataTypeTok{data =}\NormalTok{ ro_data, }\DataTypeTok{family =} \KeywordTok{binomial}\NormalTok{(}\StringTok{"logit"}\NormalTok{))}
+\NormalTok{    coefs <-}\StringTok{ }\KeywordTok{exp}\NormalTok{(}\KeywordTok{coef}\NormalTok{(ro_model))}
+\NormalTok{    ci <-}\StringTok{ }\KeywordTok{confint}\NormalTok{(ro_model)}
+\NormalTok{    ci <-}\StringTok{ }\KeywordTok{exp}\NormalTok{(ci)}
+\NormalTok{    out <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{Variable =} \KeywordTok{names}\NormalTok{(coefs), }\DataTypeTok{Estimate =}\NormalTok{ coefs, }\DataTypeTok{Lower =}\NormalTok{ ci[, }
+        \DecValTok{1}\NormalTok{], }\DataTypeTok{Upper =}\NormalTok{ ci[, }\DecValTok{2}\NormalTok{])}
+    \KeywordTok{row.names}\NormalTok{(out) <-}\StringTok{ }\OtherTok{NULL}
+\NormalTok{    out}\OperatorTok{$}\StringTok{`}\DataTypeTok{P value}\StringTok{`}\NormalTok{ <-}\StringTok{ }\KeywordTok{coef}\NormalTok{(}\KeywordTok{summary}\NormalTok{(ro_model))[, }\DecValTok{4}\NormalTok{]}
+\NormalTok{    out}\OperatorTok{$}\NormalTok{Significant <-}\StringTok{ }\KeywordTok{ifelse}\NormalTok{(out}\OperatorTok{$}\StringTok{`}\DataTypeTok{P value}\StringTok{`} \OperatorTok{<=}\StringTok{ }\FloatTok{0.001}\NormalTok{, }\StringTok{"xxx"}\NormalTok{, }\KeywordTok{ifelse}\NormalTok{(out}\OperatorTok{$}\StringTok{`}\DataTypeTok{P value}\StringTok{`} \OperatorTok{<=}\StringTok{ }
+\StringTok{        }\FloatTok{0.01}\NormalTok{, }\StringTok{"xx"}\NormalTok{, }\KeywordTok{ifelse}\NormalTok{(out}\OperatorTok{$}\StringTok{`}\DataTypeTok{P value}\StringTok{`} \OperatorTok{<=}\StringTok{ }\FloatTok{0.05}\NormalTok{, }\StringTok{"x"}\NormalTok{, }\StringTok{""}\NormalTok{)))}
+    \KeywordTok{return}\NormalTok{(}\KeywordTok{kable}\NormalTok{(out, }\DataTypeTok{format =} \StringTok{"latex"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{kable_styling}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }\KeywordTok{row_spec}\NormalTok{(}\DataTypeTok{row =} \KeywordTok{which}\NormalTok{(}\KeywordTok{grepl}\NormalTok{(}\StringTok{"pre_irs"}\NormalTok{, }
+\NormalTok{        out}\OperatorTok{$}\NormalTok{Variable)), }\DataTypeTok{background =} \StringTok{"yellow"}\NormalTok{))}
+\NormalTok{\}}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# BASIC FORMULA}
+\KeywordTok{make_model}\NormalTok{(}\StringTok{"absent ~ pre_irs + time_period"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{l|r|r|r|r|l}
+\hline
+Variable & Estimate & Lower & Upper & P value & Significant\\
+\hline
+(Intercept) & 0.0498463 & 0.0474720 & 0.0523031 & 0.0000000 & xxx\\
+\hline
+\rowcolor{yellow}  pre\_irsTRUE & 0.8050236 & 0.7002405 & 0.9211332 & 0.0019204 & xx\\
+\hline
+time\_periodMalaria season & 3.1215538 & 2.9408258 & 3.3145230 & 0.0000000 & xxx\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# BASIC FORMULA WITH INTERACTIONS}
+\KeywordTok{make_model}\NormalTok{(}\StringTok{"absent ~ pre_irs*time_period"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{l|r|r|r|r|l}
+\hline
+Variable & Estimate & Lower & Upper & P value & Significant\\
+\hline
+(Intercept) & 0.0496594 & 0.0472551 & 0.0521477 & 0.0000000 & xxx\\
+\hline
+\rowcolor{yellow}  pre\_irsTRUE & 0.8741393 & 0.6918180 & 1.0889051 & 0.2443527\\
+\hline
+time\_periodMalaria season & 3.1397591 & 2.9537482 & 3.3386865 & 0.0000000 & xxx\\
+\hline
+\rowcolor{yellow}  pre\_irsTRUE:time\_periodMalaria season & 0.8804465 & 0.6645722 & 1.1740347 & 0.3797816\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX}
+\KeywordTok{make_model}\NormalTok{(}\StringTok{"absent ~ pre_irs*time_period + sex"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{l|r|r|r|r|l}
+\hline
+Variable & Estimate & Lower & Upper & P value & Significant\\
+\hline
+(Intercept) & 0.0350661 & 0.0325983 & 0.0376818 & 0.0000000 & xxx\\
+\hline
+\rowcolor{yellow}  pre\_irsTRUE & 0.8880020 & 0.7026608 & 1.1064043 & 0.3043502\\
+\hline
+time\_periodMalaria season & 3.1716504 & 2.9834194 & 3.3729726 & 0.0000000 & xxx\\
+\hline
+sexM & 1.5960091 & 1.4919474 & 1.7085243 & 0.0000000 & xxx\\
+\hline
+\rowcolor{yellow}  pre\_irsTRUE:time\_periodMalaria season & 0.8579023 & 0.6473160 & 1.1443754 & 0.2910119\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX AND DEPARTMENT}
+\KeywordTok{make_model}\NormalTok{(}\StringTok{"absent ~ pre_irs*time_period + sex + department"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{l|r|r|r|r|l}
+\hline
+Variable & Estimate & Lower & Upper & P value & Significant\\
+\hline
+(Intercept) & 0.0564718 & 0.0510602 & 0.0623971 & 0.0000000 & xxx\\
+\hline
+\rowcolor{yellow}  pre\_irsTRUE & 0.9242121 & 0.7310398 & 1.1520107 & 0.4962745\\
+\hline
+time\_periodMalaria season & 3.1302404 & 2.9439618 & 3.3294967 & 0.0000000 & xxx\\
+\hline
+sexM & 1.1923874 & 1.1059966 & 1.2861150 & 0.0000048 & xxx\\
+\hline
+departmentFactory & 1.0828452 & 1.0009732 & 1.1717926 & 0.0476615 & x\\
+\hline
+departmentField & 0.5767584 & 0.5330008 & 0.6243185 & 0.0000000 & xxx\\
+\hline
+\rowcolor{yellow}  pre\_irsTRUE:time\_periodMalaria season & 0.8145012 & 0.6141075 & 1.0872550 & 0.1585481\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX AND DEPARTMENT AND}
+\CommentTok{# STATUS}
+\KeywordTok{make_model}\NormalTok{(}\StringTok{"absent ~ pre_irs*time_period + sex + department + permanent_or_temporary"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{l|r|r|r|r|l}
+\hline
+Variable & Estimate & Lower & Upper & P value & Significant\\
+\hline
+(Intercept) & 0.0617511 & 0.0558057 & 0.0682626 & 0.0000000 & xxx\\
+\hline
+\rowcolor{yellow}  pre\_irsTRUE & 1.0466179 & 0.8252949 & 1.3092306 & 0.6983498\\
+\hline
+time\_periodMalaria season & 2.9528835 & 2.7744446 & 3.1438877 & 0.0000000 & xxx\\
+\hline
+sexM & 1.2300194 & 1.1386295 & 1.3293865 & 0.0000002 & xxx\\
+\hline
+departmentFactory & 1.0837360 & 1.0014970 & 1.1731018 & 0.0462328 & x\\
+\hline
+departmentField & 1.3324160 & 1.2272323 & 1.4470610 & 0.0000000 & xxx\\
+\hline
+permanent\_or\_temporaryTemporary & 0.0839961 & 0.0744587 & 0.0944456 & 0.0000000 & xxx\\
+\hline
+\rowcolor{yellow}  pre\_irsTRUE:time\_periodMalaria season & 0.6682310 & 0.5016567 & 0.8956885 & 0.0063480 & xx\\
+\hline
+\end{tabular}
+\end{table}
+
+The second concern is that our quantification of return on investment is
+distorted by the fact that we treat IRS operations as essentially linear
+in nature, when in reality economies of scale, in-kind purchases and
+other factors likely make the true cost-per-spraying convex. We address
+this by creating three scenarios: (1) a ``start-up'' scenario in which
+we take into account all costs incurred by starting a program from
+scratch, (2) a ``normal'' scenario in which we match the assumptions
+with those used in this paper (ie, account for ``wear-and-tear''
+depreciation but not vehicle purchase, etc.) and a (3) ``absorbed
+costs'' scenario which ignores all costs which are not directly incurred
+by the program\ldots{}. xxx\ldots{}
+\todo{Will address this with some sensitivity analysis}
 
@@
 
-Overall stuff
+\todo{This entire section will not be done until methods/results are finalized.}
+
+\begin{itemize}
+\item
+  Overview of findings
+\item
+  How this contributes to the literature
+\item
+  Implications for policy and for businesses
+\end{itemize}
 
@@
 
-\subsection{Implications}\label{implications}
+\begin{itemize}
+\tightlist
+\item
+  No analysis yet of different worker types (agricultural
+  vs.~industrial).
+\item
+  Have not yet ventured at all into side-analyses (effect on employment,
+  tonnage, etc.).
+\item
+  Sick absenteeism seems to track absenteeism poorly: lack of clarity
+  regarding pathways.
+\item
+  Large sample size, but all from same place: questionable
+  generalizability.
+\end{itemize}
 
@@
 \hypertarget{refs}{}
+\hypertarget{ref-Ajani2010}{}
+Ajani, O., Ashagidigbi, W., 2010. Effect of malaria on rural households'
+farm income in oyo state, nigeria. African Journal of Biomedical
+Research 11. \url{https://doi.org/10.4314/ajbr.v11i3.50723}
+
+\hypertarget{ref-AsensoOkyere1997}{}
+Asenso-Okyere, W., Dzator, J.A., 1997. Household cost of seeking malaria
+care. a retrospective study of two districts in ghana. Social Science \&
+Medicine 45, 659--667.
+\url{https://doi.org/10.1016/s0277-9536(96)00383-8}
+
 \hypertarget{ref-Ashley2018}{}
@@
 18, 667--709. \url{https://doi.org/10.1093/jae/ejn028}
+
+\hypertarget{ref-Bell2014}{}
+Bell, A., Jones, K., 2014. Explaining fixed effects: Random effects
+modeling of time-series cross-sectional and panel data. Political
+Science Research and Methods 3, 133--153.
+\url{https://doi.org/10.1017/psrm.2014.7}
 
@@
 
+\hypertarget{ref-Burton1999}{}
+Burton, W.N., Conti, D.J., Chen, C.-Y., Schultz, A.B., Edington, D.W.,
+1999. The role of health risk factors and disease on worker
+productivity. Journal of Occupational \& Environmental Medicine 41,
+863--877. \url{https://doi.org/10.1097/00043764-199910000-00007}
+
 \hypertarget{ref-Buur2012}{}
@@
 
+\hypertarget{ref-Cole2006}{}
+Cole, M.A., Neumayer, E., 2006. The impact of poor health on total
+factor productivity. The Journal of Development Studies 42, 918--938.
+\url{https://doi.org/10.1080/00220380600774681}
+
+\hypertarget{ref-Curtis2003}{}
+Curtis, C., Maxwell, C., Lemnge, M., Kilama, W., Steketee, R.W., Hawley,
+W.A., Bergevin, Y., Campbell, C.C., Sachs, J., Teklehaimanot, A.,
+Ochola, S., Guyatt, H., Snow, R.W., 2003. Scaling-up coverage with
+insecticide-treated nets against malaria in africa: Who should pay? The
+Lancet Infectious Diseases 3, 304--307.
+\url{https://doi.org/10.1016/s1473-3099(03)00612-1}
+
+\hypertarget{ref-Dillon2014}{}
+Dillon, A., Friedman, J., Serneels, P., 2014. Health information,
+treatment, and worker productivity: Experimental evidence from malaria
+testing and treatment among nigerian sugarcane cutters. The World Bank.
+\url{https://doi.org/10.1596/1813-9450-7120}
+
+\hypertarget{ref-Dusfour2009}{}
+Dusfour, I., Achee, N.L., Briceno, I., King, R., Grieco, J.P., 2009.
+Comparative data on the insecticide resistance of anopheles~albimanus in
+relation to agricultural practices in northern belize, CA. Journal of
+Pest Science 83, 41--46. \url{https://doi.org/10.1007/s10340-009-0268-7}
+
+\hypertarget{ref-GarcaBasteiro2017}{}
+García-Basteiro, A.L., Ribeiro, R.M., Brew, J., Sacoor, C., Valencia,
+S., Bulo, H., Cobelens, F., Macete, E., 2017. Tuberculosis on the rise
+in southern mozambique (19972012). European Respiratory Journal 49,
+1601683. \url{https://doi.org/10.1183/13993003.01683-2016}
+
 \hypertarget{ref-German2013}{}
@@
 
+\hypertarget{ref-Gonzlez2012}{}
+González, R., Munguambe, K., Aponte, J., Bavo, C., Nhalungo, D., Macete,
+E., Alonso, P., Menéndez, C., Naniche, D., 2012. High HIV prevalence in
+a southern semi-rural area of mozambique: A community-based survey. HIV
+Medicine 13, 581--588.
+\url{https://doi.org/10.1111/j.1468-1293.2012.01018.x}
+
 \hypertarget{ref-Goodman1999}{}
@@
 Medicine and Hygiene 71, 168--173.
+
+\hypertarget{ref-Hong2011}{}
+Hong, S.C., 2011. Malaria and economic productivity: A longitudinal
+analysis of the american case. The Journal of Economic History 71,
+654--671. \url{https://doi.org/10.1017/s0022050711001872}
 
@@
 
+\hypertarget{ref-Ijumba2002}{}
+Ijumba, J.N., Mosha, F.W., Lindsay, S.W., 2002. Malaria transmission
+risk variations derived from different agricultural practices in an
+irrigated area of northern tanzania. Medical and Veterinary Entomology
+16, 28--38. \url{https://doi.org/10.1046/j.0269-283x.2002.00337.x}
+
 \hypertarget{ref-estatistica2009}{}
 INE, 2011. Demographic health survey.
+
+\hypertarget{ref-Jaleta2013}{}
+Jaleta, K.T., Hill, S.R., Seyoum, E., Balkew, M., Gebre-Michael, T.,
+Ignell, R., Tekie, H., 2013. Agro-ecosystems impact malaria prevalence:
+Large-scale irrigation drives vector population in western ethiopia.
+Malaria Journal 12, 350. \url{https://doi.org/10.1186/1475-2875-12-350}
 
@@
 
+\hypertarget{ref-Liu2010}{}
+Liu, L., Strawderman, R.L., Cowen, M.E., Shih, Y.-C.T., 2010. A flexible
+two-part random effects model for correlated medical costs. Journal of
+Health Economics 29, 110--123.
+\url{https://doi.org/10.1016/j.jhealeco.2009.11.010}
+
 \hypertarget{ref-Lopez_Bernal_2016}{}
@@
 
+\hypertarget{ref-Mayor2007}{}
+Mayor, A., Aponte, J.J., Fogg, C., Saúte, F., Greenwood, B., Dgedge, M.,
+Menendez, C., Alonso, P.L., 2007.. Malaria Journal 6, 3.
+\url{https://doi.org/10.1186/1475-2875-6-3}
+
 \hypertarget{ref-McCarthy_2000}{}
-McCarthy, D., Wolf, H., Wu, Y., 2000. The growth costs of malaria.
+McCarthy, D., Wolf, H., Wu, Y., 2000a. The growth costs of malaria.
 \url{https://doi.org/10.3386/w7541}
+
+\hypertarget{ref-McCarthy2000}{}
+McCarthy, D., Wolf, H., Wu, Y., 2000b. The growth costs of malaria.
+National Bureau of Economic Research.
+\url{https://doi.org/10.3386/w7541}
+
+\hypertarget{ref-Minalu2011}{}
+Minalu, G., Aerts, M., Coenen, S., Versporten, A., Muller, A.,
+Adriaenssens, N., Beutels, P., Molenberghs, G., Goossens, H., Hens, N.,
+2011. Application of mixed-effects models to study the country-specific
+outpatient antibiotic use in europe: A tutorial on longitudinal data
+analysis. Journal of Antimicrobial Chemotherapy 66, vi79--vi87.
+\url{https://doi.org/10.1093/jac/dkr460}
+
+\hypertarget{ref-Mocumbi2017}{}
+Mocumbi, S., Gafos, M., Munguambe, K., Goodall, R., and, S.M., 2017.
+High HIV prevalence and incidence among women in southern mozambique:
+Evidence from the MDP microbicide feasibility study. PLOS ONE 12,
+e0173243. \url{https://doi.org/10.1371/journal.pone.0173243}
 
@@
 
+\hypertarget{ref-Phillips98}{}
+Phillips, S.F.M., 1998. Economics and its contribution to the fight
+against malaria. Annals of Tropical Medicine And Parasitology 92,
+391--398. \url{https://doi.org/10.1080/00034989859375}
+
+\hypertarget{ref-Pluess2009}{}
+Pluess, B., Mueller, I., Levi, D., King, G., Smith, T.A., Lengeler, C.,
+2009. Malaria a major health problem within an oil palm plantation
+around popondetta, papua new guinea. Malaria Journal 8, 56.
+\url{https://doi.org/10.1186/1475-2875-8-56}
+
 \hypertarget{ref-Purdy_2013}{}
@@
 Nature 415, 680--685. \url{https://doi.org/10.1038/415680a}
+
+\hypertarget{ref-Sacoor2013}{}
+Sacoor, C., Nhacolo, A., Nhalungo, D., Aponte, J.J., Bassat, Q.,
+Augusto, O., Mandomando, I., Sacarlal, J., Lauchande, N., Sigauque, B.,
+Alonso, P., Macete, E., 2013. Profile: Manhica health research centre
+(manhica HDSS). International Journal of Epidemiology 42, 1309--1318.
+\url{https://doi.org/10.1093/ije/dyt148}
+
+\hypertarget{ref-Sate2003}{}
+Saúte, F., Aponte, J., Ahmeda, J., Ascaso, C., Vaz, N., Dgedge, M.,
+Alonso, P., 2003. Malaria in southern mozambique: Incidence of clinical
+malaria in children living in a rural community in manhiça district.
+Transactions of the Royal Society of Tropical Medicine and Hygiene 97,
+655--660. \url{https://doi.org/10.1016/s0035-9203(03)80097-4}
 
@@
 
+\hypertarget{ref-Wondwosen2018}{}
+Wondwosen, B., Birgersson, G., Tekie, H., Torto, B., Ignell, R., Hill,
+S.R., 2018. Sweet attraction: Sugarcane pollen-associated volatiles
+attract gravid anopheles arabiensis. Malaria Journal 17.
+\url{https://doi.org/10.1186/s12936-018-2245-1}
+
 \end{document}

### 2018-04-18 inst/rmd/paper_prepare.R

@@ -6,7 +6,7 @@ library(brew)
 library(maragra)
 library(knitr)
 library(RColorBrewer)
-library(lme4)
+# library(lme4)
 library(lfe)
 
 if('prepared_data.RData' %in% dir()){
@@ -152,10 +152,8 @@ if('prepared_data.RData' %in% dir()){
     message(i)
     this_group <- groups[i]
     these_data <- model_data %>% filter(group == this_group)
-    # this_model <- afex::mixed(absent ~ (1|oracle_number) + season * months_since, data = these_data)
-    this_model <- lmer(absent ~ -1 + (1|oracle_number) + on_site + season * months_since, data = these_data)
-    # this_model <- felm(absent ~ season*months_since | oracle_number | 0 | 0, 
-    #                    data = these_data)
+    this_model <- felm(absent ~ season*months_since | oracle_number | 0 | 0,
+                       data = these_data)
     fe_models[[i]] <- this_model
   }
   names(fe_models) <- groups

### 2018-04-18 R/make_authors.R
minor changes

@@ -15,7 +15,7 @@ make_authors <- function(
   authors = data.frame(name = c('Joe Brew',
                                 'Kizito Gondo',
                                 'Elton Dorkin',
-                                'Eduardo Nhamahanga',
+                                # 'Eduardo Nhamahanga',
                                 'Menno Pradhan',
                                 'Laia Cirera',
                                 'Ranjeeta Thomas',
@@ -23,7 +23,7 @@ make_authors <- function(
                        email = c('joe@databrew.cc',
                                  'KGondo@illovo.co.za',
                                  'EDorkin@illovo.co.za',
-                                 'ENhamahanga@illovo.co.za',
+                                 # 'ENhamahanga@illovo.co.za',
                                  'm.p.pradhan@vu.nl',
                                  'laia.cirera@isglobal.org',
                                  'ranjeeta.thomas@imperial.ac.uk',
@@ -31,13 +31,13 @@ make_authors <- function(
                        affilitation = c("isglobal,cism,vu",
                                         'ma',
                                         'ma',
-                                        'ma',
+                                        # 'ma',
                                         "vu,uva",
                                         'isglobal,cism',
                                         "icl",
                                         "isglobal,cism,icl"),
                        footnote = c("Corresponding Author",
-                                    rep('', 7)),
+                                    rep('', 6)),
                        stringsAsFactors = FALSE),
 
   institutions = data.frame(code = c('isglobal',

### 2018-04-18 R/theme_maragra.R

@@ -2,20 +2,17 @@
 #'
 #' Add a theme to a ggplot object
 #' @param base_size The base font size
-#' @oaram font_family The font
 #' @return A ggplot compatible theme call
 #' @import ggthemes
 #' @import grid
 #' @export
 # Borrowed from https://rpubs.com/Koundy/71792
-theme_maragra <- function(base_size=14,
-                          font_family = 'Computer Modern Roman') {
-  (theme_foundation(base_size=base_size,
-                    base_family = font_family)
+theme_maragra <- function(base_size=14) {
+  (theme_foundation(base_size=base_size)
     + theme(plot.title = element_text(#face = "bold",
                                       size = rel(1),
                                       hjust = 0.5),
-            text = element_text(),
+            # text = element_text(),
             panel.background = element_rect(colour = NA),
             plot.background = element_rect(colour = NA),
             panel.border = element_rect(colour = NA),
@@ -37,7 +34,6 @@ theme_maragra <- function(base_size=14,
             plot.margin=unit(c(10,5,5,5),"mm"),
             strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
             strip.text = element_text(face="bold") +
-            theme(text=element_text(size=16,
-                                      family=font_family))
+            theme(text=element_text(size=16))
     ))
 }

### 2018-04-20 inst/rmd/01_introduction.Rmd
major overhaul

@@
 
-From the persepctive of the private sector, however, investing in malaria control is not so clear-cut.
+From the persepctive of the private sector, however, investing in malaria control is not so clear-cut, since the benefits are often disperse, long-term, and difficult to quantify.
 Public health interventions targetting malaria - and their corresponding cost-effectiveness evaluations - most often focus on impacts pertaining to public welfare, such as an increase in life years adjusted for disability or quality  [@Goodman1999] [@Shretta2016] [@Lee2017] [@Hanson2004].
 Though population-level health is certainly of importance to businesses, and improvements in health incidentally improve the economy at all levels [@World1999] [@Bloom2008] [@Vecchi_2013], these improvements may be too disperse or long-term to incentivize private sector involvement in health campaigns.
+In other words, the returns for malaria control are less for the private sector because (i) they capture only part of the benefits and (ii) they do not benefit from the externalities.
+
+Just as the benefits of malaria control to the private sector are more limited than those to the public sector, considerations regarding costs for a firm are also distinct than those for a government.
+Though many firms in endemic regions engage in malaria control programs, this should not be considered, per se, evidence of its cost-effectiveness (since the extent to which corporate social responsibility plays a role is unknown).
 
@@
 And where large firms exist, they often take on social roles such as housing and health care [@Winkler].
-At times, this role is necessary from a purely practical standpoint; in other cases, it is employed under the auspices of "corporate social responsibility" [@Azemar2009].
+At times, this role is necessary from a purely practical standpoint; in other cases, it is employed under the auspices of "corporate social responsibility" [@Azemar2009] [@Curtis2003].
 Regardless of the language used, it is clear that private industry plays an important role in public health in Mozambique [@Robbins2012] [@CastelBranco2014].
 
-Several cases exist of foreign firms engaging in large-scale malaria control campaigns [@Mouzin2011] [@Han] [@Bennett_2017] [@Kaula_2017].
-But these studies generally consider population health as the outcome measure of interest, rather than worker absenteeism or productivity.
+To address the question of the profitability of malaria control activities from the standpoint of a private firm, we analyze data during a 4 year period from a private sugar facility in Southern Mozambique.
+We use absenteeism as our outcome variable, assuming that it is directly correlated with the productivity losses associated with malaria infection.
+We assess the effect of indoor residual spraying (IRS) on absenteeism, and demonstrate that the firm's engagement in malaria control not only improved worker health, but also generated a positive return on investment from a pure accounting perspective.
+
+The structure of the paper is as follows.
+We provide an overview of the sugar company under study, and the epidemiology of malaria in the nearby area, as well as in Mozambique as a whole.
+We then give an overview of the data collected, and outline the theoretical and methdological assumptions and tools that underly our analysis.
+We assess the effect of time since IRS on worker absenteeism, controlling for malaria seasonality, and segregating models for four different worker types.
+In the results, we show that IRS spraying is associated with a significant reduction in worker absenteeism among permanent and fieldworkers, and has little to no effect on temporary and indoor workers.
+We find that, in addition to reducing absence, the IRS program has a cost savings effect.
+Our discussion covers potential implications from this study in terms of policy and investment, as well as the paper's limitations.
+
+## Background literature
+
+The evidence of malaria's negative effects on both health and wealth, at both the individual [@Cole2006] and collective [@McCarthy2000] levels, as well as in both the short [@AsensoOkyere1997] [@Ajani2010] and long [@Hong2011] [@Sachs2002] terms, are amply described in the public health and economics literature [@Phillips98].
+That said, very little exists in the literature examining the costs and benefits of malaria control from a private ledger perspective (ie, the point of view a business investor).
+Unlike a government or individual, a private firm investing in malaria control may be most interested not in its long-term macroeconomic effects, nor its short-term personal health effects, but rather on the impact on productivity (and the extent to which that producitivty's benefits accrue to the firm), absenteeism, and the opportunity costs of expenditures in malaria control.
+In other words, though the large magnitude of malaria control's benefits are well known, the portion of those benefits accrued by a private firm investing in malaria control is unknown.
+
+In general, large firms operating in malaria endemic regions consider malaria to be an important enough issue to merit at least some investment [@Pluess2009].
+Several studies examine the effect of foreign firms engaging in large-scale malaria control campaigns [@Han] [@Bennett_2017] [@Kaula_2017].
+AngloGold-Ashanti, in partnership with local and national government in Ghana, invested in a well-rounded malaria control program in 2005, and saw worker absenteeism fall by 50% in 13 months [@anglo].
+Lafarge's simulatenous investment in a comprehensive malaria control program in Benin was associated with an average 41% reduction in absenteeism among workers over the course of 4 years [@lafarge].
+Zambia Sugar Plc, Zambia's largest sugar processing facility, saw annual malaria cases at its company clinic fall from nearly 3,000 in 2001 to less than 500 by 2005, following investment in a malaria control program.
+Marathon Oil's investment of 15 million US in vector control, education, net distribution and malaria treatment on Bioko Island in Equatorial Guinea lead to an estimated 95% reduction in the number of parasite-infected mosquitoes and 50% reduction in malaria incidence among young children [@asquino2015] [@Overgaard2012].
+A PATH study in Zambia found a return on investment of 28% among three companies investing in employer-based malaria control [@Mouzin2011].
+\todo{Still need to include stuff from http://gbchealth.org/wp-content/uploads/2014/03/IRS2011WorkshopReport.pdf}
+
+Though certainly suggestive of high returns on malaria control investments, these studies generally consider population health as the outcome measure of interest, rather than worker absenteeism or productivity.
 Similarly, they often neglect to differentiate between those clinical costs which are absorbed by the local health system versus those which are absorbed by the firm itself.
-In the literature, making the "investment case" for malaria control or elimination generally implies that the investor is the public sector, and takes into account those costs and benefits which are applicable from a public welfare point of view [@Shretta_2017]; though appropriate in most cases (the government or institutions interested in public welfare primarily being the primary malaria control agents in most locations), the findings of these studies are rarely applicable to the private sector, and even less so at granular levels (such as an individual firm).
+When absenteeism itself is considered, effects of malaria control have generally been found to be high, but causation is difficult to establish, given that the previous studies rely on aggregate data.
+
+Two studies utilize worker-level data to estimate the effect of malaria control on productivity.
+A World Bank analysis of Nigerian sugarcane cutters found that the simple availability of testing and treatment increased productivity by 10% in the weeks following the provision of services, the conclusion being that both the treatment and the test result were effective in increasing productivity, the latter simply increasing the information available which could influence personal labor allocation decisions [@Dillon2014].
+A randomized controlled trial (RCT) in Zambia showed an even greater effect from investments in preventing malaria: farmers given bed nets saw fewer days lost to illness (both directly and due to caretaking responsibilities for ill family members), translating to an increase of approximately 15% in crop yield [@Fink2015].
+Though compelling, the Nigerian program only dealt with medical services (diagnostics and medication), rather than preventive interventions, and the Zambian RCT examined individual farmers, as opposed to a large firm.
+
+In the literature, making the "investment case" for malaria control or elimination generally implies that the investor is the public sector, and takes into account those costs and benefits which are applicable from a public welfare point of view [@Shretta_2017].
+For example, an economic analysis by the Corporate Alliance on Malaria in Africa on the Bioko Island Malaria Control Program found a 4:1 cost-benefit ratio, but the perspective in this case was considered to be the "community" [@lafarge].
+Though appropriate in most cases to consider benefits accrued to the community (the government or institutions interested in public welfare primarily being the primary malaria control agents in most locations), the findings of these studies are rarely applicable to the private sector, and even less so at granular levels (such as an individual firm).
 In the case of a private firm not interested in "corporate social responsibility", it is not clear whether investing in malaria control would be profitable or not.
 This lack of clarity not only discourages investment, but also makes it difficult for governments to pinpoint the correct of amount of subsidy (if applicable) to encourage private sector scale-up in malaria control.
-
-To address the question of the profitability of malaria control activities from the standpoint of a private firm, we analyze data during a 4 year period from a private sugar facility in Southern Mozambique.
-We assess the effect of indoor residual spraying (IRS) on the absenteeism and health of workers, and demonstrate that the firm's engagement in malaria control not only improved worker health, but also generated a positive return on investment from a pure accounting perspective.
-
-The structure of the paper is as follows.
-After a brief review of the existing literature on the subject of private sector investment in malaria control and its effects on worker productivity, we provide an overview of the sugar company under study, and the epidemiology of malaria in the nearby area, as well as in Mozambique as a whole.
-We then give an overview of the data collected, and outline the theoretical and methdological assumptions and tools that underly our analysis.
-In the results, we show the effectiveness of the company's malaria control program in reducing absences, and translate this reduction into cost savings.
-Our discussion covers potential implications from this study in terms of policy and investment, as well as the paper's limitations.
 
@@
 Sugarcane areas may offer the standing water necessary for mosquito breeding, but also perhaps attract mosquitoes which would otherwise be elsewhere, due to compounds in sugarcane pollen [@Wondwosen2018].
-
-Regardless of the academic literature on the risk of workers becoming infected with malaria, the companies running large agricultural operations in malaria-endemic regions generally consider the disease to be an important enough threat to workers' health to merit some form of investment in malaria control [@Pluess2009].
-Insecticide fumigation, bednet distribution, test-and-treat clinics, and occasional educational campaigns are common forms of corporate social responsibility to combat malaria.
-A World Bank study on Nigerian sugarcane cutters found that the simple availability of testing and treatment increased productivity by 10% in the weeks following the provision of services, the conclusion being that both the treatment and the test result were effective in increasing productivity, the latter simply increasing the information available which could influence personal labor allocation decisions [@Dillon2014].
-
-Malaria may disincentivize investment in business in malaria zones.
-In a vicious cycle, businesses already in malaria zones may not be incentivized to invest in malaria control, perhaps because this is seen as the domain of the public sector [@Curtis2003], or perhaps due to not knowing the extent to which the benefits in health of that control will accrue to the business itself.
-
-The evidence of malaria's negative effects on both health and wealth, at both the individual [@Cole2006] and collective [@McCarthy2000] levels, as well as in both the short [@AsensoOkyere1997] [@Ajani2010] and long [@Hong2011] [@Sachs2002] terms, are amply described in the public health and economics literature [@Phillips98].
-That said, very little exists in the literature examining the costs and benefits of malaria control from a private ledger perspective (ie, the point of view a business investor).
-Unlike a government or individual, a private firm investing in malaria control may be most interested not in its long-term macroeconomic effects, nor its short-term personal health effects, but rather on the impact on productivity (and the extent to which that producitivty's benefits accrue to the firm), absenteeism, and the opportunity costs of expenditures in malaria control.
-In other words, though the large magnitude of malaria control's benefits are well known, the portion of those benefits accrued by a private firm investing in malaria control is unknown.
 
@@
 
+```{r}
+# Calculate area of maragra
+# x <- bairros
+# x$NeighCode <- as.numeric(as.character(x$NeighCode))
+# x@data <- x@data %>%
+#   mutate(maragra_bairro = NeighCode %in%
+#            c(601:609,701:709, 1001:1003)) %>%
+#   mutate(maragra_fabrica = NeighCode %in%
+#            1001:1003)
+# x <- spTransform(x,CRS("+proj=utm +zone=36 +datum=WGS84 +units=m") )
+# 
+# the_bairro <- x[x$maragra_bairro,]
+# the_fabrica <- x[x$maragra_fabrica,]
+# # the_bairro <- spTransform(the_bairro, CRS("+proj=utm +zone=36 +datum=WGS84 +units=m") )
+# # the_fabrica <- spTransform(the_fabrica, CRS("+proj=utm +zone=36 +datum=WGS84 +units=m") )
+# 
+# rgeos::gArea(the_bairro) / 1000000
+# rgeos::gArea(the_fabrica) / 1000000
+
+# Used this for plantation estimation: https://www.daftlogic.com/projects-google-maps-area-calculator-tool.htm
+```
+
 The mill of the Maragra Açucar SA (a subsidary of the Illovo sugar company, henceforth referred to as "Maragra") was nationalized in the 1970s (like all other Mozambican mills), went through a period of low production, and then fell completely out of use by 1984.
@@
 Today, Maragra accounts for roughly one quarter of Mozambique's overall sugar production (second only to the nearby Xinavane mill run by the Hulett Sugar Tongaat company) [@sutton].
-With a favorably close location to the port of Maputo, ample land, approximately 5,000 employees (of which three quarters are seasonal), and a mill with the capacity to process not only all the sugarcane grown on Maragra's land, but also the cane of the many nearby smallholders [@OLaughlin2016], Maragra has so far been able to weather the 2016 Mozambican crisis and concurrent collapse in global sugar prices.
+With a favorably close location to the port of Maputo, ample land (approximately 90 squared kilometers of plantation, and 5 squared kilometers of factory grounds), approximately 5,000 employees (of which three quarters are seasonal), and a mill with the capacity to process not only all the sugarcane grown on Maragra's land, but also the cane of the many nearby smallholders [@OLaughlin2016], Maragra has so far been able to weather the 2016 Mozambican crisis and concurrent collapse in global sugar prices.
 

### 2018-04-20 inst/rmd/02_methods.Rmd
major overhaul

@@
 
-Data pertaining to district-wide malaria incidence was obtained from Mozambique’s Boletim Epidemiológico Semanal (BES), which is the system by which the National Malaria Control Program monitors incidence at the district level throghout the entire country, and reports
-the number of confirmed weekly malaria cases at government health facilities.
+Data pertaining to district-wide malaria incidence was obtained from Mozambique’s Boletim Epidemiológico Semanal (BES), which is the system by which the National Malaria Control Program monitors incidence at the district level throghout the entire country, and reports the number of confirmed weekly malaria cases at government health facilities.
 Using these case numbers, combined with population estimates from the National Statistical Institute (INE), we estimate each day's annualized weekly malaria incidence rate (cases per 1000 population at risk), interpolated from the weekly figures.
@@
 Maragra regularly employs IRS at on-site worker households in order to reduce those workers' (and their families') risk of malaria infection.
+IRS works by killing the malaria vector (mosquito), thereby preventing infection of the household occupants.
+When administered correctly, IRS is a low-risk intervention to its recipients, and is assumed not to affect absenteeism in the short-term (to the extent that it may cause negative side-effects, these are generally long-term).
+It is preventive only, and does not cure current malaria infection, nor does it affect the parasite load of mosquitoes.
 Workers living off-site (our control group) also may have received IRS at some point during the study period (from government programs).
@@
 We divide into 4 different models for different worker types, so as to account for the potential time-varying effects of worker type on risk of malaria.
-Our econometric model is relatively straightforward:
+Our model specification is as follows
 
 $$
-\hat{Y_{it}} = \beta_{0} +  (\beta_{1}) (\text{Season}_{it}) + (\beta_2{IRS}*\beta_3{IRS_it}) + ... + \epsilon_{it}
+\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * (\hat{\beta}_2{IRS_{it}}*\hat{\beta}_3{IRStime_{it}}) + \alpha_i + \delta_t + \upsilon_{it}
 $$
 
-$\hat{Y}$ is the rate of absence.
-$\beta_{1}$ represents the clinical malaria incidence at that time in the entire district of Manhiça.
-Our demographic confounders (represented by $...$) are sex, age, and worker department (field, factory, or administrative).
-Our intervention whether the residence of the worker in question was treated in the last year, and, if so, the time since treatment (represented above as the interaction term, where where $_t$ represents time elapsed since commencement of the most recent IRS campaign).
+$\hat{Y}_{it}$ is the rate of absence.
+$\beta_{1}$ is the binary "season" variable, imputed from overall district clinical incidence.
+Our intervention is whether the residence of the worker in question was treated in the last year, and, if so, the time since treatment, represented, respectively, by $\beta_{2}$ and $\beta_{3}$.
+$\alpha_i$ reprsents the time invariant worker fixed effects, and $\delta_i$ represents the fixed effect of the particular malaria season.
+$\upsilon$ is the error term.
+
 We define the malaria season as any time during which the clinical incidence of malaria in the district of Manhiça was at or greater than the median clinical incidence of malaria for the entire study period.
@@
 non rainy season), our seasonality estimate is a closer approximation of true malaria risk, incorporating lagged effects such as the incubation period of the parasite, as well as any inherent non-linear effects of weather.
-In addition, we adjust for daily precipitation
+In addition, we adjust for daily precipitation; though its lagged effect on malaria incidence is likely captured by the seasonality term, we include rainfall since it's immediate effect (through its impact on transportation and working conditions) may also affect absenteeism.
 
@@
 Most were in their 20s and 30s and employed on temporary contracts.
-Table 1 shows overall worker details.
+Table 1 shows overall worker details divided by "Treatment" versus "Control" status.
+"Treatment" is considered the time beginning at IRS administration and ending one year later; "Control" is considered time observed either prior to IRS administration (and greater than 1 year after previous IRS administration) or among workers who never received IRS.
+Because of the temporal nature of these categories, many workers belong to both treatment and control groups, albeit at different times.
 
 ```{r, results='asis'}
-x <- workers %>%
-  filter(oracle_number %in% unique(ab_panel$oracle_number))
+# x <- ab_panel %>%
+#   left_join(workers %>%
+#               dplyr::select(oracle_number, date_of_birth, permanent_or_temporary, sex))
+x <- model_data
 x$age <- round(as.numeric(as.Date('2014-01-01') - x$date_of_birth) / 365.25, digits = -1)
@@
     group_by_(var) %>%
-    summarise(n = n()) %>%
+    summarise(`Treatment workers` = length(unique(oracle_number[!days_since %in% c('Never', 'Before')])),
+              `Treatment days` = length(oracle_number[!days_since %in% c('Never', 'Before')]),
+              `Control workers` = length(unique(oracle_number[days_since %in% c('Never', 'Before')])),
+              `Control days` = length(oracle_number[days_since %in% c('Never', 'Before')])) %>%
     ungroup() %>%
-    mutate(p = round(n / sum(n) * 100, digits = 2)) %>%
+    # mutate(p = round(n / sum(n) * 100, digits = 2)) %>%
     mutate(variable = '') 
@@
 
-    out <- out %>% dplyr::select(variable, ` `, n, p)
+    out <- out %>% dplyr::select(variable, ` `, `Treatment workers`, `Treatment days`, `Control workers`, `Control days`)
     out$` ` <- as.character(out$` `)
@@
 demo_table <- bind_rows(out_list)
-names(demo_table) <- c('Variable', ' ', 'Workers', '(%)')
+names(demo_table)[1:2] <- c('Variable', ' ')
 print(xtable::xtable(demo_table,
                      caption = 'Overall worker characteristics',
-                     align = c('r', 'r', rep('l', 3))),
+                     align = c('r', 'r', rep('l', 5))),
       include.rownames = FALSE)
 
-ov_ab <- round(100 * mean(ab_panel$absent[between(ab_panel$date, as.Date('2012-01-01'), as.Date('2016-12-31'))]), digits = 2)
-
-ov_cl <- length(which(between(clinic$date, as.Date('2012-01-01'), as.Date('2016-12-31'))))
 ```
-
-The overall absenteeism rate during the study period was 5.56% among all workers.
-The number of malaria cases registered at the company clinic during that time was 1873, for an approximate annualized clinical incidence of 110 cases per 1,000.
 
@@
 
-```{r, fig.width = 3, fig.height = 2.5, fig.align='center', fig.cap = 'Age distribution by sex and worker status'}
+```{r, fig.width = 5, fig.height = 3.5, fig.align='center', fig.cap = 'Age distribution by sex and worker status'}
 x$age <- round(as.numeric(as.Date('2014-01-01') - x$date_of_birth) / 365.25, digits = 2)

### 2018-04-20 inst/rmd/03_results.Rmd
major overhaul

@@
 
-In Southern Mozambique, malaria peaks during the summer months (December through March) most years (Figure 4, panel A), and worker absenteeism at Maragra rates track malaria incidence closely, following the same seasonal patterns (figure 4, panel B).
-Both all-cause absenteeism and sick absenteeism have declined in recent years at Maragra (figure 4, panel C), with the latter declining at a faster rate than the former.
-The fact that the rate of confirmed cases at the company clinic is largely non-seasonal (figure 4, panels E and F) suggests that a significant portion of workers either seek care for malaria elsewhere (for example, government health posts, of which several are nearby and in some many cases closer to workers' residence than the company clinic) or do not seek care during malaria infection.
+In Southern Mozambique, malaria peaks during the summer months (December through March) most years (Figure 5, panel A), and worker absenteeism at Maragra rates track malaria incidence closely, following the same seasonal patterns (figure 4, panel B).
+Both all-cause absenteeism and sick absenteeism have declined in recent years at Maragra (figure 5, panel C), with the latter declining at a faster rate than the former.
+The fact that the rate of confirmed cases at the company clinic is largely non-seasonal (figure 5, panels E and F) suggests that a significant portion of workers either seek care for malaria elsewhere (for example, government health posts, of which several are nearby and in some many cases closer to workers' residence than the company clinic) or do not seek care during malaria infection.
 Accordingly, we focus our analysis on all-cause absenteeism rather than sickness absenteeism or malaria diagnostics, with the assumption that much of illness is captured by absenteeism but not by the clinical data.
@@
 fumigations <- mc %>%
-  filter(date >= '2012-01-01',
+  filter(date >= '2013-01-01',
          date <= '2016-12-31')
@@
 
-**Fumigations**: During the period from January 1st, 2012 through December 31st, 2016, the Maragra Malaria Control Unit carried out `r pr(n_fumigations)` episodes of fumigation of residential "agregados" (household combinations), for a total of `r pr(n_buildings)` building-fumigation combinations.
+**Fumigations**: During the period from January 1st, 2013 through December 31st, 2016, the Maragra Malaria Control Unit carried out `r pr(n_fumigations)` episodes of fumigation of residential "agregados" (household combinations), for a total of `r pr(n_buildings)` building-fumigation combinations.
 The total number of unique agregados sprayed during this period was `r pr(unique_agregados)`.
 Among the `r pr(people_observed)` workers for whom we have reliable absenteeism and residential data, `r pr(people_sprayed)` had their homes fumigated at least once (the majority of workers live off of the facility).
+
+```{r}
+dry_days <- length(which(weather$precipitation > 0))
+n_days <- length(weather$precipitation)
+p_dry <- round(dry_days / n_days * 100)
+amount_rain <- round(mean(weather$precipitation[weather$precipitation > 0], na.rm = T), digits = 2)
+x <- weather %>%
+  group_by(month = format(date, '%m')) %>%
+  summarise(precipitation = mean(precipitation, na.rm = TRUE))
+```
 
@@
 
-**Effect of IRS on clinical malaria**: xxx. \todo{Not done yet. In fact, not even sure if I should do this section at all, since it's basically just a distraction.}
+```{r}
+ov_ab <- round(100 * mean(ab_panel$absent[between(ab_panel$date, as.Date('2012-01-01'), as.Date('2016-12-31'))]), digits = 2)
+
+ov_cl <- length(which(between(clinic$date, as.Date('2012-01-01'), as.Date('2016-12-31'))))
+```
+
+**Precipitation**: Of the `r n_days` days observed, `r dry_days` had no rainfall at all (ie, approximately two-thirds).
+On days for which there was any rainfall at all, the average amount was approximately `r amount_rain` centimeters.
+Rain was most common in December and January (average of 5-5.5 cm daily) and least common in August and September (average of 0.5 cm daily).
+Days with any rainfall saw more absenteeism than days with no rainfall (Figure 6, panel A) and among days with any rainfall, more precipitation was associated with greater absenteeism (Figure 6, panel B) (correlation coefficient of 0.25).
+
+```{r, fig.cap = 'Rainfall and absenteeism: association of any rainfall with absenteeism rate (left) and association of rainfall amount and absenteeism (right)', fig.height=3.5, fig.width=7, fig.align='center'}
+bi_data <- model_data %>%
+  group_by(date) %>%
+  mutate(any_rain = mean(precipitation, na.rm = TRUE) > 0) %>%
+  ungroup %>%
+  filter(!is.na(any_rain)) %>%
+  mutate(any_rain = ifelse(any_rain, 'Some rain', 'No rain')) %>%
+  group_by(any_rain, date) %>%
+  summarise(absenteeism = mean(absent) * 100)
+bi_data_simple <- bi_data %>%
+  group_by(any_rain) %>%
+  summarise(absenteeism = mean(absenteeism))
+p1 <- ggplot(data = bi_data,
+       aes(x = any_rain,
+           y = absenteeism)) +
+  geom_jitter(alpha = 0.3, size = 0.5) +
+  geom_violin(alpha = 0.4, fill = 'lightblue') +
+  scale_y_log10() +
+  theme_maragra() +
+  geom_bar(data = bi_data_simple,
+             aes(x = any_rain,
+                 y = absenteeism),
+             stat = 'identity',
+             fill = 'grey',
+             alpha = 0.8,
+           width = 0.2) +
+  geom_label(data = bi_data_simple,
+             aes(x = any_rain,
+                 y = absenteeism,
+                 label = round(absenteeism, digits = 2)),
+             color = 'red') +
+  labs(x = '',
+       y = 'Absenteeism rate (log scale)',
+       title = 'A')
+
+plot_data <- model_data %>%
+  group_by(date) %>%
+  summarise(precipitation = mean(precipitation, na.rm = TRUE),
+            absenteeism = mean(absent)) %>%
+  ungroup %>%
+  filter(precipitation > 0.1)
+p2 <- ggplot(data = plot_data,
+       aes(x = precipitation,
+           y = absenteeism * 100)) +
+  geom_point(alpha = 0.3, size = 0.5) +
+  scale_x_log10() +
+  geom_smooth() +
+  # scale_y_log10() +
+  theme_maragra() +
+  labs(x = 'Daily precipitation (log scale)',
+       y = 'Absenteeism rate', 
+       title = 'B')
+Rmisc::multiplot(p1, p2, cols = 2)
+```
+
+The extent to which rainfall's effect on absenteeism is confounded by its effect on malaria is discussed in the model results section.
+
+**Effect of IRS on clinical malaria**: The number of malaria cases registered at the company clinic during that time was 1873, for an approximate annualized clinical incidence of 110 cases per 1,000.
+\todo{Not done yet.
+In fact, not even sure if I should do this section at all, since it's basically just a distraction.}
 
@@
 
-In order to assess the effect of IRS on absenteeism, we created 4 worker fixed effects models for the 4 principal worker types (permanent field worker, temporary field worker, permanent non-field worker, temporary non-field worker).
+In order to assess the effect of IRS on absenteeism, we created 4 worker fixed effects models for the 4 principal worker types (permanent field worker, temporary field worker, permanent non-field worker, temporary non-field worker) for both outcoms (all cause absenteeism vs.
+sickness absenteeism).
 The reason for segregating models rather than incorporating worker type as a fixed effect was that it seemed plausible that the effect of the treatment (IRS) on the outcome would be different (and vary differently over time) depending on these worker types.
@@
 
-The model results (table 3) suggest a significant decrease in absenteeism following the administration of IRS.
-The effect is ambivalent among temporary workers, whereas among permanent workers the administration of IRS decreases absenteeism betwen 2.5 and 8.4 percentage points in the 2-9 month following administration.
-These effects are consistent with IRS decreasing absence through the medium of malaria, since (a) the effect is not as strong in the first month (potentially during the parasite incubation phase), the effect is stronger on field workers (who are more socioeconomically the demographic at greater risk of malaria than their administrative counterparts), and the effect is ambivalent among temporary workers (since the protective effect of IRS is only a function of the amount of time that the worker spends in the fumigated house).
-
-```{r, results = 'asis'}
+```{r}
 clean_up_model <- function(x){
@@
   x <- x %>% filter(!grepl('sd_', Term))
+  x <- x %>% filter(!grepl('10+', Term, fixed = TRUE))
   return(x)
@@
 
+make_models_table <- function(model_list, the_caption = "Models with worker fixed effects"){
+
   out_list <- list()
-for(i in 1:length(fe_models)){
+  for(i in 1:length(model_list)){
     message(i)
-  this_model <- names(fe_models)[i]
-  the_model <- fe_models[[which(names(fe_models) == this_model)]]
+    this_model <- names(model_list)[i]
+    the_model <- model_list[[which(names(model_list) == this_model)]]
     the_model <- clean_up_model(the_model)
@@
   }
-
-# df <- out_list[[1]]
-# for(i in 2:length(out_list)){
-#   df <- left_join(df, out_list[[i]])
-# }
-# print(xtable::xtable(df, caption = 'Regression outputs'), size="\\fontsize{5pt}{6pt}\\selectfont",
-#       include.rownames = FALSE)
 
@@
   breaks <- c(0, breaker * 1:(lengther))
-k <- kable(df, format = "latex", caption = "Models with worker fixed effects", booktabs = T) %>%
+  k <- kable(df, format = "latex", caption = the_caption, booktabs = T) %>%
     kable_styling()
@@
   print(k)
+}
+```
 
-# # Combine them all
-# out_list <- list()
-# for (i in 1:length(models)){
-#   this_model <- letters[i]
-#   the_model <- models[[this_model]]
-#   the_model <- clean_up_model(the_model)
-#   names(the_model)[2] <- paste0('Model ', i)
-#   out_list[[i]] <- the_model
-# }
-# df <- out_list[[length(models)]]
-# # Order the Terms
-# df_order <- c()
-# for(i in 1:length(models)){
-#   df_order <- c(df_order, out_list[[i]]$Term)
-# }
-# df_order <- unique(df_order)
-# df <- left_join(data.frame(Term = df_order),
-#                 df)
-# for(i in (length(models) - 1):1){
-#   message(i)
-#     df <- left_join(df, out_list[[i]])
-# }
-# df <- df[,unique(c('Term', sort(names(df))))]
-# print(xtable::xtable(df, caption = 'Regression outputs'), size="\\fontsize{5pt}{6pt}\\selectfont")
+The model results for all cause absenteeism (table 4) suggest a significant decrease in absenteeism following the administration of IRS.
+The effect is ambivalent among temporary workers, whereas among permanent workers the administration of IRS decreases absenteeism betwen 2.5 and 8.4 percentage points in the 2-9 month following administration.
+These effects are consistent with IRS decreasing absence through the medium of malaria, since (a) the effect is not as strong in the first month (potentially during the parasite incubation phase), the effect is stronger on field workers (who are more socioeconomically the demographic at greater risk of malaria than their administrative counterparts), and the effect is ambivalent among temporary workers (since the protective effect of IRS is only a function of the amount of time that the worker spends in the fumigated house).
 
+```{r, results = 'asis'}
+make_models_table(model_list = fe_models, the_caption = "All cause absenteeism: model results")
+```
+
+The model results for sickness only absenteeism (table 5) suggest that IRS has a significant effect on permanent field workers.
+The effect is directionally similar among temporary fieldworkers (but does not reach the level of significance).
+Among non-fieldworkers, the effect of IRS is mixed.
+
+```{r, results = 'asis'}
+make_models_table(model_list = sick_models, the_caption = "Sick absenteeism only: model results")
 ```
@@
 
-**Translating absences to costs**
+This section is in progress. Looking at neighborhood effects, core-periphery effects, etc.
+
+## Translating absences to costs
 
@@
 If IRS application is indeed endogenous, we would expect absenteeism to be elevated during this period (since the increase in absenteeism would be theoretically responsible for the application of IRS), a situation which would require further statistical adjustment.
-If, on the other hand, there is no endogeneity, we would expect absenteeism in the 10 day period prior to IRS administration to be similar to other pre-IRS absenteeism (adjusting for seasonality).
+If, on the other hand, there is no endogeneity, we would expect absenteeism in the 10 day period prior to IRS administration to be similar to other pre-IRS absenteeism (adjusting for seasonality).\todo{Should add model specification formula here.}
 
-What follows are several variations of a model to assess whether IRS application predicts pre-IRS absenteeism.
-In the final version of the paper, we'll use whichever model is most appropriate, but this is left here for discussion purposes.
-I've also included the code, so as to inform the discussion.
-In the below models, `pre_irs` refers to the 10 day period prior to IRS administration.
-The models estimate the odds of absence (ie, the estimates and confidence intervals are in the form of odds ratios).
+The below shows our robustness check for all cause absenteeism.
 
-```{r, echo = TRUE}
+```{r, echo = FALSE}
 ro_data <- ab_panel %>%
@@
             by = 'oracle_number')
+# ro_data$time_period <- as.character(ro_data$time_period)
 
@@
                                           '')))
+  out$Variable <- as.character(out$Variable)
+  out$Variable <- gsub('pre_irsTRUE', '10 days prior to IRS', out$Variable)
+  out$Variable <- gsub('time_period', '', out$Variable)
   return(kable(out, format = "latex") %>%
   kable_styling() %>%
-    row_spec(row = which(grepl('pre_irs', out$Variable)),background = "yellow"))
+    row_spec(row = which(grepl('prior', out$Variable)),background = "yellow"))
 }
@@
 
-```{r, echo = TRUE, results = 'asis'}
+```{r, echo = FALSE, results = 'asis'}
 
-# BASIC FORMULA
-make_model('absent ~ pre_irs + time_period')
+# # BASIC FORMULA
+# make_model('absent ~ pre_irs + time_period')
 
-# BASIC FORMULA WITH INTERACTIONS
-make_model('absent ~ pre_irs*time_period')
+# # BASIC FORMULA WITH INTERACTIONS
+# make_model('absent ~ pre_irs*time_period')
 
-# BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX
-make_model('absent ~ pre_irs*time_period + sex')
+# # BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX
+# make_model('absent ~ pre_irs*time_period + sex')
+# 
+# # BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX AND DEPARTMENT
+# make_model('absent ~ pre_irs*time_period + sex + department')
+# 
+# 
+# # BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX AND DEPARTMENT AND STATUS
+# make_model('absent ~ pre_irs*time_period + sex + department + permanent_or_temporary')
 
-# BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX AND DEPARTMENT
-make_model('absent ~ pre_irs*time_period + sex + department')
+# Adjust for seasonality and department
+make_model('absent ~ pre_irs*time_period + department')
+```
 
-# BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX AND DEPARTMENT AND STATUS
-make_model('absent ~ pre_irs*time_period + sex + department + permanent_or_temporary')
+The below shows our robustness check for sick only absenteeism.
+Unlike with all cause absenteeism, during malaria season, being in the period 10 days prior to IRS is associated with statistically greater likelihood of being absent for illness.
+In other words, it appears that there is somewhat of a feedback loop: when a worker misses work due to illness, his/her likelihood of getting IRS doubles in the next 10 days.\todo{What to do about this???}
 
+```{r}
+make_model('absent_sick ~ pre_irs*time_period + department')
 ```

### 2018-04-20 inst/rmd/04_discussion.Rmd
major overhaul

@@
 - Overview of findings
-
 - How this contributes to the literature
-
 - Implications for policy and for businesses
+- Spillover effects !
+- How absenteeism might be _underestimating_ true effect, since at the margin a sick worker might go to work (ie, not be absent), but be less productive
 

### 2018-04-20 inst/rmd/paper.Rmd
major overhaul

@@
                dev = "cairo_pdf",
-               fig.pos="H",
+               fig.pos="h",
                fig.path = 'figures/')

### 2018-04-20 inst/rmd/paper.tex
major overhaul

@@
 Laia Cirera\footrecall{isglobal}\footrecall{cism}
-Ranjeeta Thomas\footremember{icl}{Imperial College London: South Kensington Campus, London SW7 2AZ, U.K., UK}
-Elisa Sicuri\footrecall{isglobal}\footrecall{cism}\footrecall{icl}
+Elisa Sicuri\footrecall{isglobal}\footrecall{cism}\footremember{icl}{Imperial College London: South Kensington Campus, London SW7 2AZ, U.K., UK}
 
@@
 From the persepctive of the private sector, however, investing in
-malaria control is not so clear-cut. Public health interventions
-targetting malaria - and their corresponding cost-effectiveness
-evaluations - most often focus on impacts pertaining to public welfare,
-such as an increase in life years adjusted for disability or quality
-(Goodman et al., 1999) (Shretta et al., 2016) (Lee et al., 2017)
-(Hanson, 2004). Though population-level health is certainly of
-importance to businesses, and improvements in health incidentally
-improve the economy at all levels (Brundtland, 1999) (Bloom and Canning,
-2008) (Vecchi et al., 2013), these improvements may be too disperse or
-long-term to incentivize private sector involvement in health campaigns.
+malaria control is not so clear-cut, since the benefits are often
+disperse, long-term, and difficult to quantify. Public health
+interventions targetting malaria - and their corresponding
+cost-effectiveness evaluations - most often focus on impacts pertaining
+to public welfare, such as an increase in life years adjusted for
+disability or quality (Goodman et al., 1999) (Shretta et al., 2016) (Lee
+et al., 2017) (Hanson, 2004). Though population-level health is
+certainly of importance to businesses, and improvements in health
+incidentally improve the economy at all levels (Brundtland, 1999) (Bloom
+and Canning, 2008) (Vecchi et al., 2013), these improvements may be too
+disperse or long-term to incentivize private sector involvement in
+health campaigns. In other words, the returns for malaria control are
+less for the private sector because (i) they capture only part of the
+benefits and (ii) they do not benefit from the externalities.
+
+Just as the benefits of malaria control to the private sector are more
+limited than those to the public sector, considerations regarding costs
+for a firm are also distinct than those for a government. Though many
+firms in endemic regions engage in malaria control programs, this should
+not be considered, per se, evidence of its cost-effectiveness (since the
+extent to which corporate social responsibility plays a role is
+unknown).
 
@@
 other cases, it is employed under the auspices of ``corporate social
-responsibility'' (Azemar and Desbordes, 2009). Regardless of the
-language used, it is clear that private industry plays an important role
-in public health in Mozambique (Robbins and Perkins, 2012)
-(Castel-Branco, 2014).
-
-Several cases exist of foreign firms engaging in large-scale malaria
-control campaigns (Mouzin and al., 2011) (Han, 2015) (Bennett et al.,
-2017) (Kaula et al., 2017). But these studies generally consider
-population health as the outcome measure of interest, rather than worker
-absenteeism or productivity. Similarly, they often neglect to
-differentiate between those clinical costs which are absorbed by the
-local health system versus those which are absorbed by the firm itself.
+responsibility'' (Azemar and Desbordes, 2009) (Curtis et al., 2003).
+Regardless of the language used, it is clear that private industry plays
+an important role in public health in Mozambique (Robbins and Perkins,
+2012) (Castel-Branco, 2014).
+
+To address the question of the profitability of malaria control
+activities from the standpoint of a private firm, we analyze data during
+a 4 year period from a private sugar facility in Southern Mozambique. We
+use absenteeism as our outcome variable, assuming that it is directly
+correlated with the productivity losses associated with malaria
+infection. We assess the effect of indoor residual spraying (IRS) on
+absenteeism, and demonstrate that the firm's engagement in malaria
+control not only improved worker health, but also generated a positive
+return on investment from a pure accounting perspective.
+
+The structure of the paper is as follows. We provide an overview of the
+sugar company under study, and the epidemiology of malaria in the nearby
+area, as well as in Mozambique as a whole. We then give an overview of
+the data collected, and outline the theoretical and methdological
+assumptions and tools that underly our analysis. We assess the effect of
+time since IRS on worker absenteeism, controlling for malaria
+seasonality, and segregating models for four different worker types. In
+the results, we show that IRS spraying is associated with a significant
+reduction in worker absenteeism among permanent and fieldworkers, and
+has little to no effect on temporary and indoor workers. We find that,
+in addition to reducing absence, the IRS program has a cost savings
+effect. Our discussion covers potential implications from this study in
+terms of policy and investment, as well as the paper's limitations.
+
+\subsection{Background literature}\label{background-literature}
+
+The evidence of malaria's negative effects on both health and wealth, at
+both the individual (Cole and Neumayer, 2006) and collective (McCarthy
+et al., 2000b) levels, as well as in both the short (Asenso-Okyere and
+Dzator, 1997) (Ajani and Ashagidigbi, 2010) and long (Hong, 2011) (Sachs
+and Malaney, 2002) terms, are amply described in the public health and
+economics literature (Phillips, 1998). That said, very little exists in
+the literature examining the costs and benefits of malaria control from
+a private ledger perspective (ie, the point of view a business
+investor). Unlike a government or individual, a private firm investing
+in malaria control may be most interested not in its long-term
+macroeconomic effects, nor its short-term personal health effects, but
+rather on the impact on productivity (and the extent to which that
+producitivty's benefits accrue to the firm), absenteeism, and the
+opportunity costs of expenditures in malaria control. In other words,
+though the large magnitude of malaria control's benefits are well known,
+the portion of those benefits accrued by a private firm investing in
+malaria control is unknown.
+
+In general, large firms operating in malaria endemic regions consider
+malaria to be an important enough issue to merit at least some
+investment (Pluess et al., 2009). Several studies examine the effect of
+foreign firms engaging in large-scale malaria control campaigns (Han,
+2015) (Bennett et al., 2017) (Kaula et al., 2017). AngloGold-Ashanti, in
+partnership with local and national government in Ghana, invested in a
+well-rounded malaria control program in 2005, and saw worker absenteeism
+fall by 50\% in 13 months (CCM, 2016). Lafarge's simulatenous investment
+in a comprehensive malaria control program in Benin was associated with
+an average 41\% reduction in absenteeism among workers over the course
+of 4 years (Egedeye et al., 2011). Zambia Sugar Plc, Zambia's largest
+sugar processing facility, saw annual malaria cases at its company
+clinic fall from nearly 3,000 in 2001 to less than 500 by 2005,
+following investment in a malaria control program. Marathon Oil's
+investment of 15 million US in vector control, education, net
+distribution and malaria treatment on Bioko Island in Equatorial Guinea
+lead to an estimated 95\% reduction in the number of parasite-infected
+mosquitoes and 50\% reduction in malaria incidence among young children
+(Asquino, 2016) (Overgaard et al., 2012). A PATH study in Zambia found a
+return on investment of 28\% among three companies investing in
+employer-based malaria control (Mouzin and al., 2011).
+\todo{Still need to include stuff from http://gbchealth.org/wp-content/uploads/2014/03/IRS2011WorkshopReport.pdf}
+
+Though certainly suggestive of high returns on malaria control
+investments, these studies generally consider population health as the
+outcome measure of interest, rather than worker absenteeism or
+productivity. Similarly, they often neglect to differentiate between
+those clinical costs which are absorbed by the local health system
+versus those which are absorbed by the firm itself. When absenteeism
+itself is considered, effects of malaria control have generally been
+found to be high, but causation is difficult to establish, given that
+the previous studies rely on aggregate data.
+
+Two studies utilize worker-level data to estimate the effect of malaria
+control on productivity. A World Bank analysis of Nigerian sugarcane
+cutters found that the simple availability of testing and treatment
+increased productivity by 10\% in the weeks following the provision of
+services, the conclusion being that both the treatment and the test
+result were effective in increasing productivity, the latter simply
+increasing the information available which could influence personal
+labor allocation decisions (Dillon et al., 2014). A randomized
+controlled trial (RCT) in Zambia showed an even greater effect from
+investments in preventing malaria: farmers given bed nets saw fewer days
+lost to illness (both directly and due to caretaking responsibilities
+for ill family members), translating to an increase of approximately
+15\% in crop yield (Fink and Masiye, 2015). Though compelling, the
+Nigerian program only dealt with medical services (diagnostics and
+medication), rather than preventive interventions, and the Zambian RCT
+examined individual farmers, as opposed to a large firm.
+
 In the literature, making the ``investment case'' for malaria control or
@@
 and takes into account those costs and benefits which are applicable
-from a public welfare point of view (Shretta et al., 2017); though
-appropriate in most cases (the government or institutions interested in
-public welfare primarily being the primary malaria control agents in
-most locations), the findings of these studies are rarely applicable to
-the private sector, and even less so at granular levels (such as an
-individual firm). In the case of a private firm not interested in
-``corporate social responsibility'', it is not clear whether investing
-in malaria control would be profitable or not. This lack of clarity not
-only discourages investment, but also makes it difficult for governments
-to pinpoint the correct of amount of subsidy (if applicable) to
-encourage private sector scale-up in malaria control.
-
-To address the question of the profitability of malaria control
-activities from the standpoint of a private firm, we analyze data during
-a 4 year period from a private sugar facility in Southern Mozambique. We
-assess the effect of indoor residual spraying (IRS) on the absenteeism
-and health of workers, and demonstrate that the firm's engagement in
-malaria control not only improved worker health, but also generated a
-positive return on investment from a pure accounting perspective.
-
-The structure of the paper is as follows. After a brief review of the
-existing literature on the subject of private sector investment in
-malaria control and its effects on worker productivity, we provide an
-overview of the sugar company under study, and the epidemiology of
-malaria in the nearby area, as well as in Mozambique as a whole. We then
-give an overview of the data collected, and outline the theoretical and
-methdological assumptions and tools that underly our analysis. In the
-results, we show the effectiveness of the company's malaria control
-program in reducing absences, and translate this reduction into cost
-savings. Our discussion covers potential implications from this study in
-terms of policy and investment, as well as the paper's limitations.
+from a public welfare point of view (Shretta et al., 2017). For example,
+an economic analysis by the Corporate Alliance on Malaria in Africa on
+the Bioko Island Malaria Control Program found a 4:1 cost-benefit ratio,
+but the perspective in this case was considered to be the ``community''
+(Egedeye et al., 2011). Though appropriate in most cases to consider
+benefits accrued to the community (the government or institutions
+interested in public welfare primarily being the primary malaria control
+agents in most locations), the findings of these studies are rarely
+applicable to the private sector, and even less so at granular levels
+(such as an individual firm). In the case of a private firm not
+interested in ``corporate social responsibility'', it is not clear
+whether investing in malaria control would be profitable or not. This
+lack of clarity not only discourages investment, but also makes it
+difficult for governments to pinpoint the correct of amount of subsidy
+(if applicable) to encourage private sector scale-up in malaria control.
 
@@
 
-Regardless of the academic literature on the risk of workers becoming
-infected with malaria, the companies running large agricultural
-operations in malaria-endemic regions generally consider the disease to
-be an important enough threat to workers' health to merit some form of
-investment in malaria control (Pluess et al., 2009). Insecticide
-fumigation, bednet distribution, test-and-treat clinics, and occasional
-educational campaigns are common forms of corporate social
-responsibility to combat malaria. A World Bank study on Nigerian
-sugarcane cutters found that the simple availability of testing and
-treatment increased productivity by 10\% in the weeks following the
-provision of services, the conclusion being that both the treatment and
-the test result were effective in increasing productivity, the latter
-simply increasing the information available which could influence
-personal labor allocation decisions (Dillon et al., 2014).
-
-Malaria may disincentivize investment in business in malaria zones. In a
-vicious cycle, businesses already in malaria zones may not be
-incentivized to invest in malaria control, perhaps because this is seen
-as the domain of the public sector (Curtis et al., 2003), or perhaps due
-to not knowing the extent to which the benefits in health of that
-control will accrue to the business itself.
-
-The evidence of malaria's negative effects on both health and wealth, at
-both the individual (Cole and Neumayer, 2006) and collective (McCarthy
-et al., 2000b) levels, as well as in both the short (Asenso-Okyere and
-Dzator, 1997) (Ajani and Ashagidigbi, 2010) and long (Hong, 2011) (Sachs
-and Malaney, 2002) terms, are amply described in the public health and
-economics literature (Phillips, 1998). That said, very little exists in
-the literature examining the costs and benefits of malaria control from
-a private ledger perspective (ie, the point of view a business
-investor). Unlike a government or individual, a private firm investing
-in malaria control may be most interested not in its long-term
-macroeconomic effects, nor its short-term personal health effects, but
-rather on the impact on productivity (and the extent to which that
-producitivty's benefits accrue to the firm), absenteeism, and the
-opportunity costs of expenditures in malaria control. In other words,
-though the large magnitude of malaria control's benefits are well known,
-the portion of those benefits accrued by a private firm investing in
-malaria control is unknown.
-
 This study does not endeavor to expand the current body of knowledge
@@
 
-\begin{figure}[H]
+\begin{figure}[h]
 
@@
 Xinavane mill run by the Hulett Sugar Tongaat company) (Sutton, 2014).
-With a favorably close location to the port of Maputo, ample land,
-approximately 5,000 employees (of which three quarters are seasonal),
-and a mill with the capacity to process not only all the sugarcane grown
-on Maragra's land, but also the cane of the many nearby smallholders
-(O'Laughlin, 2016), Maragra has so far been able to weather the 2016
-Mozambican crisis and concurrent collapse in global sugar prices.
+With a favorably close location to the port of Maputo, ample land
+(approximately 90 squared kilometers of plantation, and 5 squared
+kilometers of factory grounds), approximately 5,000 employees (of which
+three quarters are seasonal), and a mill with the capacity to process
+not only all the sugarcane grown on Maragra's land, but also the cane of
+the many nearby smallholders (O'Laughlin, 2016), Maragra has so far been
+able to weather the 2016 Mozambican crisis and concurrent collapse in
+global sugar prices.
 
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics{figures/unnamed-chunk-14-1}
+\begin{figure}[h]
+
+{\centering \includegraphics{figures/unnamed-chunk-15-1}
 
@@
 
-\caption{i. Fumigation activities carried out by Maragra Malaria Control during study period, ii. Distribution of average time between sprayings of households}\label{fig:unnamed-chunk-14}
+\caption{i. Fumigation activities carried out by Maragra Malaria Control during study period, ii. Distribution of average time between sprayings of households}\label{fig:unnamed-chunk-15}
 \end{figure}
@@
 reduce those workers' (and their families') risk of malaria infection.
-Workers living off-site (our control group) also may have received IRS
-at some point during the study period (from government programs). Even
-though we do not have reliable person-level data on IRS carried out by
-the government, off-site workers are a suitable control in the sense
-that they represent ``business as usual'' (ie, what would happen if the
-company carried out no IRS and relied solely on public interventions).
-Using company HR and clinical records, we were able to identify absences
-and episodes of clinical malaria among all workers, as well as identify
-the time since the most recent IRS episode before the onsent of absence
-or illness.
+IRS works by killing the malaria vector (mosquito), thereby preventing
+infection of the household occupants. When administered correctly, IRS
+is a low-risk intervention to its recipients, and is assumed not to
+affect absenteeism in the short-term (to the extent that it may cause
+negative side-effects, these are generally long-term). It is preventive
+only, and does not cure current malaria infection, nor does it affect
+the parasite load of mosquitoes. Workers living off-site (our control
+group) also may have received IRS at some point during the study period
+(from government programs). Even though we do not have reliable
+person-level data on IRS carried out by the government, off-site workers
+are a suitable control in the sense that they represent ``business as
+usual'' (ie, what would happen if the company carried out no IRS and
+relied solely on public interventions). Using company HR and clinical
+records, we were able to identify absences and episodes of clinical
+malaria among all workers, as well as identify the time since the most
+recent IRS episode before the onsent of absence or illness.
 
@@
 worker types, so as to account for the potential time-varying effects of
-worker type on risk of malaria. Our econometric model is relatively
-straightforward:
+worker type on risk of malaria. Our model specification is as follows
 
 \[
-\hat{Y_{it}} = \beta_{0} +  (\beta_{1}) (\text{Season}_{it}) + (\beta_2{IRS}*\beta_3{IRS_it}) + ... + \epsilon_{it}
+\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * (\hat{\beta}_2{IRS_{it}}*\hat{\beta}_3{IRStime_{it}}) + \alpha_i + \delta_t + \upsilon_{it}
 \]
 
-\(\hat{Y}\) is the rate of absence. \(\beta_{1}\) represents the
-clinical malaria incidence at that time in the entire district of
-Manhiça. Our demographic confounders (represented by \(...\)) are sex,
-age, and worker department (field, factory, or administrative). Our
-intervention whether the residence of the worker in question was treated
-in the last year, and, if so, the time since treatment (represented
-above as the interaction term, where where \(_t\) represents time
-elapsed since commencement of the most recent IRS campaign). We define
-the malaria season as any time during which the clinical incidence of
-malaria in the district of Manhiça was at or greater than the median
-clinical incidence of malaria for the entire study period. These weeks
-are flagged as red in Figure 1, Panel A. By using clinical incidence of
-the area of residence of the workers (as opposed to more typical proxies
-for malaria risk, such as only rainy vs.~non rainy season), our
-seasonality estimate is a closer approximation of true malaria risk,
-incorporating lagged effects such as the incubation period of the
-parasite, as well as any inherent non-linear effects of weather. In
-addition, we adjust for daily precipitation
+\(\hat{Y}_{it}\) is the rate of absence. \(\beta_{1}\) is the binary
+``season'' variable, imputed from overall district clinical incidence.
+Our intervention is whether the residence of the worker in question was
+treated in the last year, and, if so, the time since treatment,
+represented, respectively, by \(\beta_{2}\) and \(\beta_{3}\).
+\(\alpha_i\) reprsents the time invariant worker fixed effects, and
+\(\delta_i\) represents the fixed effect of the particular malaria
+season. \(\upsilon\) is the error term.
+
+We define the malaria season as any time during which the clinical
+incidence of malaria in the district of Manhiça was at or greater than
+the median clinical incidence of malaria for the entire study period.
+These weeks are flagged as red in Figure 1, Panel A. By using clinical
+incidence of the area of residence of the workers (as opposed to more
+typical proxies for malaria risk, such as only rainy vs.~non rainy
+season), our seasonality estimate is a closer approximation of true
+malaria risk, incorporating lagged effects such as the incubation period
+of the parasite, as well as any inherent non-linear effects of weather.
+In addition, we adjust for daily precipitation; though its lagged effect
+on malaria incidence is likely captured by the seasonality term, we
+include rainfall since it's immediate effect (through its impact on
+transportation and working conditions) may also affect absenteeism.
 
@@
 and 30s and employed on temporary contracts. Table 1 shows overall
-worker details.
+worker details divided by ``Treatment'' versus ``Control'' status.
+``Treatment'' is considered the time beginning at IRS administration and
+ending one year later; ``Control'' is considered time observed either
+prior to IRS administration (and greater than 1 year after previous IRS
+administration) or among workers who never received IRS. Because of the
+temporal nature of these categories, many workers belong to both
+treatment and control groups, albeit at different times.
 
@@
 \centering
-\begin{tabular}{rlll}
-  \hline
-Variable &   & Workers & (\%) \\ 
-  \hline
-Department & Administrative & 157 & 4.67 \\ 
-   & Factory & 459 & 13.65 \\ 
-   & Field & 2746 & 81.68 \\ 
-  Sex & F & 1305 & 38.82 \\ 
-   & M & 2057 & 61.18 \\ 
-  Status & Permanent & 500 & 14.87 \\ 
-   & Temporary & 2862 & 85.13 \\ 
-  Age & 20 & 985 & 29.30 \\ 
-   & 30 & 1244 & 37.00 \\ 
-   & 40 & 665 & 19.78 \\ 
-   & 50 & 347 & 10.32 \\ 
-   & 60 & 117 & 3.48 \\ 
-   & 70 &   4 & 0.12 \\ 
+\begin{tabular}{rlllll}
+  \hline
+Variable &   & Treatment workers & Treatment days & Control workers & Control days \\ 
+  \hline
+Department & Administrative &  35 & 12523 & 157 & 173151 \\ 
+   & Factory &  71 & 20533 & 453 & 280416 \\ 
+   & Field & 410 & 91302 & 2691 & 1181175 \\ 
+  Sex & F & 228 & 47907 & 1274 & 489484 \\ 
+   & M & 288 & 76451 & 2027 & 1145258 \\ 
+  Status & Permanent & 127 & 47075 & 500 & 655311 \\ 
+   & Temporary & 389 & 77283 & 2801 & 979431 \\ 
+  Age & 20 & 140 & 29463 & 960 & 347758 \\ 
+   & 30 & 176 & 39119 & 1225 & 550783 \\ 
+   & 40 & 123 & 30186 & 652 & 366936 \\ 
+   & 50 &  58 & 18881 & 344 & 268877 \\ 
+   & 60 &  17 & 5893 & 116 & 96555 \\ 
+   & 70 &   2 & 816 &   4 & 3833 \\ 
    \hline
@@
 \end{table}
-
-The overall absenteeism rate during the study period was 5.56\% among
-all workers. The number of malaria cases registered at the company
-clinic during that time was 1873, for an approximate annualized clinical
-incidence of 110 cases per 1,000.
 
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics{figures/unnamed-chunk-17-1}
+\begin{figure}[h]
+
+{\centering \includegraphics{figures/unnamed-chunk-18-1}
 
@@
 
-\caption{Age distribution by sex and worker status}\label{fig:unnamed-chunk-17}
+\caption{Age distribution by sex and worker status}\label{fig:unnamed-chunk-18}
 \end{figure}
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics{figures/unnamed-chunk-18-1}
+\begin{figure}[h]
+
+{\centering \includegraphics{figures/unnamed-chunk-19-1}
 
@@
 Monthly total precipitation in the Maniça district; ii.
-Average monthly temperature (bars) during the study period, as well as monthly maximum (red) and minimum (blue) temperatures}\label{fig:unnamed-chunk-18}
+Average monthly temperature (bars) during the study period, as well as monthly maximum (red) and minimum (blue) temperatures}\label{fig:unnamed-chunk-19}
 \end{figure}
@@
 In Southern Mozambique, malaria peaks during the summer months (December
-through March) most years (Figure 4, panel A), and worker absenteeism at
+through March) most years (Figure 5, panel A), and worker absenteeism at
 Maragra rates track malaria incidence closely, following the same
 seasonal patterns (figure 4, panel B). Both all-cause absenteeism and
-sick absenteeism have declined in recent years at Maragra (figure 4,
+sick absenteeism have declined in recent years at Maragra (figure 5,
 panel C), with the latter declining at a faster rate than the former.
 The fact that the rate of confirmed cases at the company clinic is
-largely non-seasonal (figure 4, panels E and F) suggests that a
+largely non-seasonal (figure 5, panels E and F) suggests that a
 significant portion of workers either seek care for malaria elsewhere
@@
 \centering
-\includegraphics{figures/unnamed-chunk-19-1.pdf}
+\includegraphics{figures/unnamed-chunk-20-1.pdf}
 \caption{Clinical malaria (district of Manhiça), all-cause absenteeism
@@
 
-\textbf{Fumigations}: During the period from January 1st, 2012 through
-December 31st, 2016, the Maragra Malaria Control Unit carried out 11,567
+\textbf{Fumigations}: During the period from January 1st, 2013 through
+December 31st, 2016, the Maragra Malaria Control Unit carried out 11,007
 episodes of fumigation of residential ``agregados'' (household
-combinations), for a total of 13,937 building-fumigation combinations.
+combinations), for a total of 13,260 building-fumigation combinations.
 The total number of unique agregados sprayed during this period was
-4,045. Among the 3,362 workers for whom we have reliable absenteeism and
+3,998. Among the 3,362 workers for whom we have reliable absenteeism and
 residential data, 692 had their homes fumigated at least once (the
@@
 
-\textbf{Effect of IRS on clinical malaria}: xxx.
+\textbf{Precipitation}: Of the 1454 days observed, 940 had no rainfall
+at all (ie, approximately two-thirds). On days for which there was any
+rainfall at all, the average amount was approximately 2.99 centimeters.
+Rain was most common in December and January (average of 5-5.5 cm daily)
+and least common in August and September (average of 0.5 cm daily). Days
+with any rainfall saw more absenteeism than days with no rainfall
+(Figure 6, panel A) and among days with any rainfall, more precipitation
+was associated with greater absenteeism (Figure 6, panel B) (correlation
+coefficient of 0.25).
+
+\begin{figure}[h]
+
+{\centering \includegraphics{figures/unnamed-chunk-25-1}
+
+}
+
+\caption{Rainfall and absenteeism: association of any rainfall with absenteeism rate (left) and association of rainfall amount and absenteeism (right)}\label{fig:unnamed-chunk-25}
+\end{figure}
+
+The extent to which rainfall's effect on absenteeism is confounded by
+its effect on malaria is discussed in the model results section.
+
+\textbf{Effect of IRS on clinical malaria}: The number of malaria cases
+registered at the company clinic during that time was 1873, for an
+approximate annualized clinical incidence of 110 cases per 1,000.
 \todo{Not done yet. In fact, not even sure if I should do this section at all, since it's basically just a distraction.}
@@
 \centering
-\includegraphics{figures/unnamed-chunk-22-1.pdf}
+\includegraphics{figures/unnamed-chunk-26-1.pdf}
 \caption{i. Absenteeism before and after IRS administration for all
@@
 worker, temporary field worker, permanent non-field worker, temporary
-non-field worker). The reason for segregating models rather than
+non-field worker) for both outcoms (all cause absenteeism vs.~sickness
+absenteeism). The reason for segregating models rather than
 incorporating worker type as a fixed effect was that it seemed plausible
@@
 
-The model results (table 3) suggest a significant decrease in
-absenteeism following the administration of IRS. The effect is
-ambivalent among temporary workers, whereas among permanent workers the
-administration of IRS decreases absenteeism betwen 2.5 and 8.4
-percentage points in the 2-9 month following administration. These
-effects are consistent with IRS decreasing absence through the medium of
-malaria, since (a) the effect is not as strong in the first month
-(potentially during the parasite incubation phase), the effect is
+The model results for all cause absenteeism (table 4) suggest a
+significant decrease in absenteeism following the administration of IRS.
+The effect is ambivalent among temporary workers, whereas among
+permanent workers the administration of IRS decreases absenteeism betwen
+2.5 and 8.4 percentage points in the 2-9 month following administration.
+These effects are consistent with IRS decreasing absence through the
+medium of malaria, since (a) the effect is not as strong in the first
+month (potentially during the parasite incubation phase), the effect is
 stronger on field workers (who are more socioeconomically the
@@
 
-\caption{\label{tab:unnamed-chunk-23}Models with worker fixed effects}
+\caption{\label{tab:unnamed-chunk-28}All cause absenteeism: model results}
 \centering
@@
 \multicolumn{2}{l}{\textbf{Permanent field worker}}\\
-\hspace{1em}Malaria season & 3.424 (P<0.001)\\
-\hspace{1em}Months since IRS: 01 & 3.83 (P<0.001)\\
-\hspace{1em}Months since IRS: 02-04 & 4.206 (P<0.001)\\
-\hspace{1em}Months since IRS: 05-09 & -3.175 (P<0.001)\\
-\hspace{1em}Months since IRS: 10+ & -5.775 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & -4.529 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & -8.398 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & -3.859 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 10+ & 4.507 (P=0.002)\\
+\hspace{1em}Malaria season & 2.533 (P<0.001)\\
+\hspace{1em}Months since IRS: 01 & 5.382 (P<0.001)\\
+\hspace{1em}Months since IRS: 02-04 & 4.747 (P<0.001)\\
+\hspace{1em}Months since IRS: 05-09 & -2.669 (P<0.001)\\
+\hspace{1em}Precipitation (mm) & 0.206 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & -6.262 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 02-04 & -8.388 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 05-09 & -3.71 (P<0.001)\\
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
-\hspace{1em}Malaria season & 2.076 (P<0.001)\\
-\hspace{1em}Months since IRS: 01 & -2.078 (P<0.001)\\
-\hspace{1em}Months since IRS: 02-04 & 3.911 (P<0.001)\\
-\hspace{1em}Months since IRS: 05-09 & -0.822 (P=0.089)\\
-\hspace{1em}Months since IRS: 10+ & -0.25 (P=0.758)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & 2.723 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & -5.785 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & -2.447 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 10+ & 4.501 (P<0.001)\\
+\hspace{1em}Malaria season & 0.957 (P<0.001)\\
+\hspace{1em}Months since IRS: 01 & -0.987 (P=0.071)\\
+\hspace{1em}Months since IRS: 02-04 & 4.375 (P<0.001)\\
+\hspace{1em}Months since IRS: 05-09 & -0.55 (P=0.254)\\
+\hspace{1em}Precipitation (mm) & 0.548 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & 1.787 (P=0.016)\\
+\hspace{1em}Malaria season:Months since IRS: 02-04 & -5.208 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 05-09 & -2.04 (P=0.003)\\
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Temporary field worker}}\\
-\hspace{1em}Malaria season & -0.005 (P=0.833)\\
-\hspace{1em}Months since IRS: 01 & -0.429 (P<0.001)\\
-\hspace{1em}Months since IRS: 02-04 & -0.575 (P<0.001)\\
-\hspace{1em}Months since IRS: 05-09 & -0.586 (P<0.001)\\
-\hspace{1em}Months since IRS: 10+ & -1.025 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & -0.066 (P=0.725)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & 0.166 (P=0.274)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & 0.063 (P=0.654)\\
-\hspace{1em}Malaria season:Months since IRS: 10+ & 0.062 (P=0.789)\\
+\hspace{1em}Malaria season & -0.368 (P<0.001)\\
+\hspace{1em}Months since IRS: 01 & -0.075 (P=0.526)\\
+\hspace{1em}Months since IRS: 02-04 & -0.193 (P=0.072)\\
+\hspace{1em}Months since IRS: 05-09 & -0.036 (P=0.738)\\
+\hspace{1em}Precipitation (mm) & 0.02 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & -0.154 (P=0.406)\\
+\hspace{1em}Malaria season:Months since IRS: 02-04 & 0.297 (P=0.05)\\
+\hspace{1em}Malaria season:Months since IRS: 05-09 & 0.041 (P=0.769)\\
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
-\hspace{1em}Malaria season & -0.22 (P=0.294)\\
-\hspace{1em}Months since IRS: 01 & -1.466 (P=0.389)\\
-\hspace{1em}Months since IRS: 02-04 & 0.571 (P=0.6)\\
-\hspace{1em}Months since IRS: 05-09 & -1.018 (P=0.334)\\
-\hspace{1em}Months since IRS: 10+ & 10.606 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & 6.093 (P=0.002)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & 0.476 (P=0.736)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & -1.505 (P=0.277)\\
-\hspace{1em}Malaria season:Months since IRS: 10+ & -10.785 (P<0.001)\\
+\hspace{1em}Malaria season & 2.025 (P<0.001)\\
+\hspace{1em}Months since IRS: 01 & -2.536 (P=0.134)\\
+\hspace{1em}Months since IRS: 02-04 & -1.177 (P=0.282)\\
+\hspace{1em}Months since IRS: 05-09 & -1.821 (P=0.084)\\
+\hspace{1em}Precipitation (mm) & 0.078 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & 6.121 (P=0.002)\\
+\hspace{1em}Malaria season:Months since IRS: 02-04 & 1.333 (P=0.343)\\
+\hspace{1em}Malaria season:Months since IRS: 05-09 & -4.639 (P<0.001)\\
 \bottomrule
@@
 
+The model results for sickness only absenteeism suggest that IRS has a
+significant effect on permanent field workers. The effect is
+directionally similar among temporary fieldworkers (but does not reach
+the level of significance). Among non-fieldworkers, the effect of IRS is
+mixed.
+
+\begin{table}
+
+\caption{\label{tab:unnamed-chunk-29}Sick absenteeism only: model results}
+\centering
+\begin{tabular}[t]{ll}
+\toprule
+Term & Estimate\\
+\midrule
+\addlinespace[1.5em]
+\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
+\hspace{1em}Malaria season & -0.073 (P=0.089)\\
+\hspace{1em}Months since IRS: 01 & 0.891 (P<0.001)\\
+\hspace{1em}Months since IRS: 02-04 & 1.223 (P<0.001)\\
+\hspace{1em}Months since IRS: 05-09 & -0.236 (P=0.249)\\
+\hspace{1em}Precipitation (mm) & 0 (P=0.984)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & -1.501 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 02-04 & -0.597 (P=0.031)\\
+\hspace{1em}Malaria season:Months since IRS: 05-09 & -0.484 (P=0.097)\\
+\addlinespace[1.5em]
+\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
+\hspace{1em}Malaria season & -0.056 (P=0.204)\\
+\hspace{1em}Months since IRS: 01 & -0.667 (P=0.003)\\
+\hspace{1em}Months since IRS: 02-04 & 0.904 (P<0.001)\\
+\hspace{1em}Months since IRS: 05-09 & 0.127 (P=0.526)\\
+\hspace{1em}Precipitation (mm) & -0.002 (P=0.565)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & 1.121 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 02-04 & -0.122 (P=0.652)\\
+\hspace{1em}Malaria season:Months since IRS: 05-09 & 1.198 (P<0.001)\\
+\addlinespace[1.5em]
+\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
+\hspace{1em}Malaria season & -0.106 (P<0.001)\\
+\hspace{1em}Months since IRS: 01 & 0.029 (P=0.545)\\
+\hspace{1em}Months since IRS: 02-04 & 0.04 (P=0.35)\\
+\hspace{1em}Months since IRS: 05-09 & -0.046 (P=0.291)\\
+\hspace{1em}Precipitation (mm) & 0 (P=0.966)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & -0.108 (P=0.148)\\
+\hspace{1em}Malaria season:Months since IRS: 02-04 & -0.032 (P=0.595)\\
+\hspace{1em}Malaria season:Months since IRS: 05-09 & -0.052 (P=0.36)\\
+\addlinespace[1.5em]
+\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
+\hspace{1em}Malaria season & -0.135 (P=0.236)\\
+\hspace{1em}Months since IRS: 01 & 1.427 (P=0.092)\\
+\hspace{1em}Months since IRS: 02-04 & -0.569 (P=0.299)\\
+\hspace{1em}Months since IRS: 05-09 & -1.42 (P=0.007)\\
+\hspace{1em}Precipitation (mm) & -0.005 (P=0.613)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & -1.796 (P=0.069)\\
+\hspace{1em}Malaria season:Months since IRS: 02-04 & 2.479 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 05-09 & 0.18 (P=0.794)\\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \subsection{Externalities}\label{externalities}
 
-\textbf{Translating absences to costs}
+This section is in progress. Looking at neighborhood effects,
+core-periphery effects, etc.
+
+\subsection{Translating absences to
+costs}\label{translating-absences-to-costs}
 
@@
 \begin{itemize}
+\tightlist
 \item
@@
   Implications for policy and for businesses
+\item
+  Spillover effects !
+\item
+  How absenteeism might be \emph{underestimating} true effect, since at
+  the margin a sick worker might go to work (ie, not be absent), but be
+  less productive
 \end{itemize}
@@
 
+\hypertarget{ref-asquino2015}{}
+Asquino, M., 2016. Equatorial Guinea Plays a Leading Role in Combating
+Malaria.
+
 \hypertarget{ref-Azemar2009}{}
@@
 \url{https://doi.org/10.1080/03056244.2014.976363}
+
+\hypertarget{ref-anglo}{}
+CCM, 2016. AngloGold Ashanti Malaria Control Ltd (AGA Mal).
 
@@
 
+\hypertarget{ref-lafarge}{}
+Egedeye, L., Drozer, S., Leiser, A.-M., 2011. Corporate Action on
+Malaria Control: Best Practices and Interventions.
+
+\hypertarget{ref-Fink2015}{}
+Fink, G., Masiye, F., 2015. Health and agricultural productivity:
+Evidence from zambia. Journal of Health Economics 42, 151--164.
+\url{https://doi.org/10.1016/j.jhealeco.2015.04.004}
+
 \hypertarget{ref-GarcaBasteiro2017}{}
@@
 
+\hypertarget{ref-Overgaard2012}{}
+Overgaard, H.J., Reddy, V.P., Abaga, S., Matias, A., Reddy, M.R.,
+Kulkarni, V., Schwabe, C., Segura, L., Kleinschmidt, I., Slotman, M.A.,
+2012. Malaria transmission after five years of vector control on bioko
+island, equatorial guinea. Parasites \& Vectors 5, 253.
+\url{https://doi.org/10.1186/1756-3305-5-253}
+
 \hypertarget{ref-OLaughlin2016}{}

### 2018-04-20 R/make_authors.R

@@ -2,13 +2,14 @@
 #'
 #' Cat the authors in latex-ready format
 #' @return Latex text
+#' @export
+
 #' @param authors A dataframe with columns: name, email, affiliation, footnote
 #' @param institutions A dataframe with columns: code, institution, address
 #' @param cat_it Whether to cat (as opposed to return)
 #' @param include_address Whether to include the address
 #' @param include_country Whether to include the country; only relevant if include_address is FALSE
 #' @param seperator What to put between author names: usually a comma or line break
-#' @export
 
 make_authors <- function(
 
@@ -18,7 +19,7 @@ make_authors <- function(
                                 # 'Eduardo Nhamahanga',
                                 'Menno Pradhan',
                                 'Laia Cirera',
-                                'Ranjeeta Thomas',
+                                #'Ranjeeta Thomas',
                                 'Elisa Sicuri'),
                        email = c('joe@databrew.cc',
                                  'KGondo@illovo.co.za',
@@ -26,7 +27,7 @@ make_authors <- function(
                                  # 'ENhamahanga@illovo.co.za',
                                  'm.p.pradhan@vu.nl',
                                  'laia.cirera@isglobal.org',
-                                 'ranjeeta.thomas@imperial.ac.uk',
+                                 #'ranjeeta.thomas@imperial.ac.uk',
                                  'elisa.sicuri@isglobal.org'),
                        affilitation = c("isglobal,cism,vu",
                                         'ma',
@@ -34,10 +35,10 @@ make_authors <- function(
                                         # 'ma',
                                         "vu,uva",
                                         'isglobal,cism',
-                                        "icl",
+                                       # "icl",
                                         "isglobal,cism,icl"),
                        footnote = c("Corresponding Author",
-                                    rep('', 6)),
+                                    rep('', 5)),
                        stringsAsFactors = FALSE),
 
   institutions = data.frame(code = c('isglobal',

### 2018-04-20 inst/rmd/paper_prepare.R

@@ -141,22 +141,36 @@ if('prepared_data.RData' %in% dir()){
     mutate(group = paste0(permanent_or_temporary, ' ',
                           tolower(field)))
   fe_models <- list()
+  sick_models <- list()
   groups <- sort(unique(model_data$group))
   # library(lmerTest)
   
   # Add a month column
   model_data$month_number <- add_zero(format(model_data$date, '%m'), 2)
   
+  # Get a malaria season var
+  model_data <- 
+    model_data %>%
+    mutate(calendar_month = as.numeric(format(date, '%m'))) %>%
+    mutate(calendar_year = as.numeric(format(date, '%Y'))) %>%
+    mutate(malaria_year = ifelse(calendar_month < 6,
+                                 paste0(calendar_year-1, '-', calendar_year),
+                                 paste0(calendar_year, '-', calendar_year+1)))
+  
   # library(nlme)
   for (i in 1:length(groups)){
     message(i)
     this_group <- groups[i]
     these_data <- model_data %>% filter(group == this_group)
-    this_model <- felm(absent ~ season*months_since | oracle_number | 0 | 0,
+    this_model <- felm(absent ~ season*months_since + precipitation | oracle_number + malaria_year | 0 | 0,
+                       data = these_data)
+    this_sick_model <- felm(absent_sick ~ season*months_since + precipitation | oracle_number + malaria_year| 0 | 0,
                        data = these_data)
     fe_models[[i]] <- this_model
+    sick_models[[i]] <- this_sick_model
   }
   names(fe_models) <- groups
+  names(sick_models) <- groups
   
   # Plots of maps
   # Libraries

### 2018-04-20 inst/rmd/02_methods.Rmd
changes

--- 2018-01-03 inst/rmd/02_methods.Rmd
@@
 To estimate this effect, we estimated separate models for absence and illness.
-We employed interrupted time series [@Lopez_Bernal_2016] and a linear probability approach using the following econometric model.
+We employed linear mixed effects models [@Minalu2011] [@Liu2010] (with individual worker fixed effects, and random effects for other covariates)  in lieu of interrupted time series [@Lopez_Bernal_2016] so as to explicitly adjust for confounders [@Bell2014].
+We divide into 4 different models for different worker types, so as to account for the potential time-varying effects of worker type on risk of malaria.
+Our model specification is as follows
 
 $$
-\hat{Y_{it}} = \beta_{0} +  (\beta_{1}) (\text{Season}_{t}) + (\beta_2{IRS}*\beta_3{IRS_t}) + ... + \epsilon
+\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * (\hat{\beta}_2{IRS_{it}}*\hat{\beta}_3{IRStime_{it}}) + \alpha_i + \delta_t + \upsilon_{it}
 $$
 
-$\hat{Y}$ is the rate of absence.
-$\beta_{1}$ represents the clinical malaria incidence at that time in the entire district of Manhiça.
-Our demographic confounders (represented by $...$) are sex, age, and worker department (field, factory, or administrative).
-Our intervention was not a simple yes/no, but rather the product of whether the residence of the worker in question was treated in the last year, and, if so, the time since treatment (represented above as the interaction term, where where $_t$ represents time elapsed since commencement of the most recent IRS campaign).
+$\hat{Y}_{it}$ is the rate of absence.
+$\beta_{1}$ is the binary "season" variable, imputed from overall district clinical incidence.
+Our intervention is whether the residence of the worker in question was treated in the last year, and, if so, the time since treatment, represented, respectively, by $\beta_{2}$ and $\beta_{3}$.
+$\alpha_i$ reprsents the time invariant worker fixed effects, and $\delta_i$ represents the fixed effect of the particular malaria season.
+$\upsilon$ is the error term.
+
 We define the malaria season as any time during which the clinical incidence of malaria in the district of Manhiça was at or greater than the median clinical incidence of malaria for the entire study period.
@@
 non rainy season), our seasonality estimate is a closer approximation of true malaria risk, incorporating lagged effects such as the incubation period of the parasite, as well as any inherent non-linear effects of weather.
-In addition, we adjust for daily precipitation
+In addition, we adjust for daily precipitation; though its lagged effect on malaria incidence is likely captured by the seasonality term, we include rainfall since it's immediate effect (through its impact on transportation and working conditions) may also affect absenteeism.
+
+Details will go here about externalities (neighborhood effects, positive spillover to district, negative spillover from district, how large firm may mean that most positive externalities are absorbed by the firm, and not many negative externalities, but maybe with a smaller firm it wouldn't work as well)\todo{NOT DONE YET.}
 
@@
 All data processing and analysis were carried out in R [@R] and all analysis code is freely available online [@brewgit].
-Ethical approval for this project was obtained from the Institutional Ethics Review Board for Health at the CISM prior to data collection.
+Ethical approval for this project was obtained from the Institutional Ethics Review Board for Health at the CISM (CIBS-CISM) prior to data collection.
 
-## Descriptive statistics
-
-Age, types of workers, bla bla bla. \todo{Will add a table here and some overview}
+\newpage

### 2018-04-20 inst/rmd/03_results.Rmd
changes

@@
 # Results
+
+## Descriptive statistics
+
+```{r}
+# Calculate number of workers
+n <- length(unique(ab_panel$oracle_number[ab_panel$date >= '2012-01-01' & ab_panel$date <= '2016-12-31']))
+```
+
+We collected absenteeism and demographic data data on `r n` workers from 2012 through 2016.
+Workers were approximately 60% male, and more than 80% fieldworkers (predominantly cane-cutters).
+Most were in their 20s and 30s and employed on temporary contracts.
+Table 1 shows overall worker details divided by "Treatment" versus "Control" status.
+"Treatment" is considered the time beginning at IRS administration and ending one year later; "Control" is considered time observed either prior to IRS administration (and greater than 1 year after previous IRS administration) or among workers who never received IRS.
+Because of the temporal nature of these categories, many workers belong to both treatment and control groups, albeit at different times.
+
+```{r, results='asis'}
+# x <- ab_panel %>%
+#   left_join(workers %>%
+#               dplyr::select(oracle_number, date_of_birth, permanent_or_temporary, sex))
+x <- model_data
+x$age <- round(as.numeric(as.Date('2014-01-01') - x$date_of_birth) / 365.25, digits = -1)
+x$status <- x$permanent_or_temporary
+
+calculate <- function(var= 'sex',
+                      data){
+  out <- 
+    data %>%
+    group_by_(var) %>%
+    summarise(`Treatment workers` = length(unique(oracle_number[!days_since %in% c('Never', 'Before')])),
+              `Treatment days` = length(oracle_number[!days_since %in% c('Never', 'Before')]),
+              `Control workers` = length(unique(oracle_number[days_since %in% c('Never', 'Before')])),
+              `Control days` = length(oracle_number[days_since %in% c('Never', 'Before')])) %>%
+    ungroup() %>%
+    # mutate(p = round(n / sum(n) * 100, digits = 2)) %>%
+    mutate(variable = '') 
+    names(out)[1] <- ' '
+    out$variable[1] <- Hmisc::capitalize(var)
+
+    out <- out %>% dplyr::select(variable, ` `, `Treatment workers`, `Treatment days`, `Control workers`, `Control days`)
+    out$` ` <- as.character(out$` `)
+  return(out)
+}
+out_list <- list()
+vars <- c('department', 
+           'sex',
+           'status',
+           'age')
+for(i in 1:length(vars)){
+  message(i)
+  y <- calculate(var = vars[i], data = x)
+  out_list[[i]] <- y
+}
+demo_table <- bind_rows(out_list)
+names(demo_table)[1:2] <- c('Variable', ' ')
+print(xtable::xtable(demo_table,
+                     caption = 'Overall worker characteristics',
+                     align = c('r', 'r', rep('l', 5))),
+      include.rownames = FALSE)
+
+```
+
+Temporary workers tended to be younger and male; female temporary workers being on average 5-10 years older than their male counterparts.
+Permanent workers had a bi-modal age distribution, the older peak explained by the greater density of males in administrative roles.
+Females accounted for 43% of temporary workers but only 16% of permanent workers (figure x).
+
+```{r, fig.width = 5, fig.height = 3.5, fig.align='center', fig.cap = 'Age distribution by sex and worker status'}
+x$age <- round(as.numeric(as.Date('2014-01-01') - x$date_of_birth) / 365.25, digits = 2)
+g1 <- ggplot(data = x,
+       aes(x = age,
+           group = sex,
+           fill = sex)) +
+  geom_density(alpha = 0.8, n = 2^4) +
+  facet_wrap(~permanent_or_temporary) +
+  theme_maragra() +
+  labs(x = 'Age',
+       y = 'Density') +
+  scale_fill_manual(name = 'Sex',
+                    values = c('lightblue', 'grey'))
+g1
+```
+
+During the study period, weather followed typical seasonal trends for the area, albeit slightly drier than previous periods.
+
+```{r, fig.width = 6, fig.height = 3.5, fig.align='center', fig.cap = 'i.
+Monthly total precipitation in the Maniça district; ii.
+Average monthly temperature (bars) during the study period, as well as monthly maximum (red) and minimum (blue) temperatures'}
+x <- weather %>%
+  mutate(month = format(date, '%m')) %>%
+  mutate(year = format(date, '%Y')) %>%
+  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
+  group_by(date) %>%
+  summarise(precipitation = sum(precipitation, na.rm = TRUE),
+            temp_max = max(temp_max, na.rm = TRUE),
+            temp_min = min(temp_min, na.rm = TRUE),
+            temp = mean(temp))
+# x <- gather(x, key, value,precipitation:temp)
+
+g1 <- ggplot(data = x,
+       aes(x = date,
+           y = precipitation)) +
+  geom_bar(stat = 'identity',
+           fill = 'lightblue',
+           size = 0.1) +
+  theme_maragra() +
+  theme(axis.text.x = element_text(angle = 0)) +
+  labs(x = 'Quarter',
+       y = 'Milimeters') +
+  scale_x_date(labels = scales::date_format("%Y"))
+
+g2 <- ggplot(data = x,
+            aes(x = date,
+                y = temp)) +
+  geom_bar(stat = 'identity', fill = 'grey', size = 0.1) +
+  geom_line(aes(y = temp_max),
+             color = 'darkred', 
+             alpha = 0.5) +
+  geom_line(aes(y = temp_min),
+             color = 'blue', 
+             alpha = 0.5) +
+  # geom_linerange(aes(ymin = temp_min,
+  #                    ymax = temp_max),
+  #                alpha = 0.6,
+  #                color = 'blue') +
+   theme_maragra() +
+  theme(axis.text.x = element_text(angle = 0)) +
+  labs(x = 'Quarter',
+       y = 'Degrees (Celcius)') +
+  scale_x_date(labels = scales::date_format("%Y"))
+Rmisc::multiplot(g1, g2, cols = 2)
+```
 
@@
 
-**Effect of IRS on absenteeism**: Immediately following IRS, a worker's likelihood of absence drops significantly (figure x, panel i).
-As one would expect if the mechanism by which IRS reduces absence is through reduced malaria infection, the effect of IRS during the low transmission season is significant, but far less substantial in effect size (figure x, panel ii).
+## Analysis
 
-```{r, fig.height = 3.5, fig.width = 6, fig.cap = 'i.
-Absenteeism before and after IRS administration for all workers who ever received IRS; ii.
-The same, but segregated by malaria and non-malaria seasons'}
-plot_data <-
-  ab_panel %>%
-  # filter(oracle_number %in% matched$oracle_number) %>%
-  left_join(irs, c("date", "unidade")) %>%
-  mutate(months_since = days_since %/% 30) %>%
-  filter(!is.na(months_since)) %>%
-  # mutate(time_period = lendable::time_period_extract(date)) %>%
-  # mutate(time_period = as.character(time_period)) %>%
-  mutate(time_period = make_season(date = date)) %>%
-  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))
+### Effect of IRS on absenteeism
 
-x <- plot_data %>%
-  group_by(months_since) %>%
-  summarise(absences = length(which(absent)),
-            sick_absences = length(which(absent_sick)),
-            eligibles = length(absent)) %>%
-  ungroup %>%
-  mutate(absenteeism_rate = absences / eligibles * 100) %>%
-  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100)
-
-g1 <- ggplot(data = x,
-       aes(x = months_since,
-           y = absenteeism_rate)) +
-  geom_step(color = 'darkred', alpha = 0.8) +
-  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
-  labs(x = 'Months relative to IRS',
-       y = 'Absenteeism rate') +
-  theme_maragra()
-
-plot_data2 <-
-  ab_panel %>%
-  # filter(oracle_number %in% matched$oracle_number) %>%
-  left_join(irs, by = c("date", "unidade")) %>%
-  mutate(months_since = days_since %/% 30) %>%
-  filter(!is.na(months_since)) %>%
-  # mutate(time_period = lendable::time_period_extract(date)) %>%
-  # mutate(time_period = as.character(time_period)) %>%
-  mutate(time_period = make_season(date = date)) %>%
-  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))
-
-x2 <- plot_data2 %>%
-  group_by(months_since, time_period) %>%
-  summarise(absences = length(which(absent)),
-            sick_absences = length(which(absent_sick)),
-            eligibles = length(absent)) %>%
-  ungroup %>%
-  mutate(absenteeism_rate = absences / eligibles * 100) %>%
-  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100) %>%
-  mutate(before = months_since < 0)
-
-g2 <- ggplot(data = x2 %>% mutate(time_period = paste0(time_period)),
-       aes(x = months_since,
-           y = absenteeism_rate)) +
-  geom_step(color = 'darkred', alpha = 0.8) +
-  facet_wrap(~time_period, ncol = 1) +
-  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
-  labs(x = 'Months relative to IRS',
-       y = 'Absenteeism rate') +
-  theme_maragra()
-Rmisc::multiplot(g1, g2, cols = 2)
-```
+Immediately following IRS, a worker's likelihood of absence drops significantly (appendix figure, panel A).
+As one would expect if the mechanism by which IRS reduces absence is through reduced malaria infection, the effect of IRS during the low transmission season is significant, but far less substantial in effect size (appendix figure, panel B).
 
@@
 The effect is directionally similar among temporary fieldworkers (but does not reach the level of significance).
-Among non-fieldworkers, the effect of IRS is mixed.
+Among non-fieldworkers, the effect of IRS is mixed.\todo{I'll probably remove this from the final version, since we're saying that the data aren't that good...}
 
@@
 
-## Externalities
+### Externalities
 
-This section is in progress. Looking at neighborhood effects, core-periphery effects, etc.
+One potential shortcoming of our analysis thus far is that a "control" worker (ie, one whose house has never been sprayed or was sprayed greater than one year ago) may actually be conferred some protection against malaria by the fumigations at his/her neighbors' houses.
+This positive spillover effect would theoretically go through two channels: (i) via a reduction of mosquitoes in the vicinity and (ii) via a reduction of the malaria parasite in the blood of humans in the vicinity (ie, the parasite "reservoir").
+
+By the same token, just as being surrounded by IRS-treated homes may indirectly protect someone living in a home not treated by IRS, one might surmise that living on the periphery of the company grounds might lead to more malaria risk relative to living in the core, since homes near the periphery but outside of Maragra are less likely to be fumigated.
+In other words, to the extent that both positive and spillover effects are significant enough to warrant inclusion in our model, we would expect that (i) the prevalence of IRS in a worker's neighborhood should be associated with a reduction in absenteeism of the worker and (b) workers living further from the periphery of the company grounds should have lower absenteeism than workers living at the periphery.
+We test both of these hypotheses.
+
+#### Neighborhood protective effect
+
+\todo{Working on this now}
+
+#### Core-periphery difference
+
+\todo{Working on this now}
 
@@
 \todo{Will address this with some sensitivity analysis}
+
+\newpage

### 2018-04-20 inst/rmd/05_appdendix.Rmd
changes

# Appendix

## Unadjusted absenteeism by time since IRS

```{r, fig.height = 3.5, fig.width = 6, fig.cap = 'i. Absenteeism before and after IRS administration for all workers who ever received IRS; ii. The same, but segregated by malaria and non-malaria seasons'}
plot_data <-
  ab_panel %>%
  # filter(oracle_number %in% matched$oracle_number) %>%
  left_join(irs, c("date", "unidade")) %>%
  mutate(months_since = days_since %/% 30) %>%
  filter(!is.na(months_since)) %>%
  # mutate(time_period = lendable::time_period_extract(date)) %>%
  # mutate(time_period = as.character(time_period)) %>%
  mutate(time_period = make_season(date = date)) %>%
  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))

x <- plot_data %>%
  group_by(months_since) %>%
  summarise(absences = length(which(absent)),
            sick_absences = length(which(absent_sick)),
            eligibles = length(absent)) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) %>%
  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100)

g1 <- ggplot(data = x,
       aes(x = months_since,
           y = absenteeism_rate)) +
  geom_step(color = 'darkred', alpha = 0.8) +
  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
  labs(x = 'Months relative to IRS',
       y = 'Absenteeism rate',
       title = 'A') +
  theme_maragra()

plot_data2 <-
  ab_panel %>%
  # filter(oracle_number %in% matched$oracle_number) %>%
  left_join(irs, by = c("date", "unidade")) %>%
  mutate(months_since = days_since %/% 30) %>%
  filter(!is.na(months_since)) %>%
  # mutate(time_period = lendable::time_period_extract(date)) %>%
  # mutate(time_period = as.character(time_period)) %>%
  mutate(time_period = make_season(date = date)) %>%
  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))

x2 <- plot_data2 %>%
  group_by(months_since, time_period) %>%
  summarise(absences = length(which(absent)),
            sick_absences = length(which(absent_sick)),
            eligibles = length(absent)) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) %>%
  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100) %>%
  mutate(before = months_since < 0)

g2 <- ggplot(data = x2 %>% mutate(time_period = paste0(time_period)),
       aes(x = months_since,
           y = absenteeism_rate)) +
  geom_step(color = 'darkred', alpha = 0.8) +
  facet_wrap(~time_period, ncol = 1) +
  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
  labs(x = 'Months relative to IRS',
       y = 'Absenteeism rate',
       title = 'B') +
  theme_maragra()
Rmisc::multiplot(g1, g2, cols = 2)
```

## Unadjusted sick absenteeism by time since IRS

```{r, fig.height = 3.5, fig.width = 6, fig.cap = 'i. Sick absenteeism before and after IRS administration for all workers who ever received IRS; ii. The same, but segregated by malaria and non-malaria seasons'}
g1 <- ggplot(data = x,
       aes(x = months_since,
           y = sick_absenteeism_rate)) +
  geom_step(color = 'darkred', alpha = 0.8) +
  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
  labs(x = 'Months relative to IRS',
       y = 'Absenteeism rate',
       title = 'A') +
  theme_maragra()

g2 <- ggplot(data = x2 %>% mutate(time_period = paste0(time_period)),
       aes(x = months_since,
           y = sick_absenteeism_rate)) +
  geom_step(color = 'darkred', alpha = 0.8) +
  facet_wrap(~time_period, ncol = 1) +
  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
  labs(x = 'Months relative to IRS',
       y = 'Absenteeism rate',
       title = 'B') +
  theme_maragra()
Rmisc::multiplot(g1, g2, cols = 2)
```

### 2018-04-20 inst/rmd/paper.Rmd
changes

@@
                dev = "cairo_pdf",
-               fig.pos="h",
+               fig.pos="!h",
                fig.path = 'figures/')

### 2018-04-20 inst/rmd/paper.tex
changes

@@
 \urlstyle{same}  % don't use monospace font for urls
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
 \usepackage{graphicx,grffile}
@@
 
+\usepackage{float}
+\floatplacement{figure}{H}
+
 % \usepackage{todonotes} % for side notes
@@
 
-\begin{figure}[h]
-
-{\centering \includegraphics{figures/unnamed-chunk-13-1}
+\begin{figure}[!h]
+
+{\centering \includegraphics{figures/unnamed-chunk-14-1}
 
@@
 
-\caption{(i) Mozambique in Africa, (ii) Districts of Mozambique with Manhiça highlighted in red, (iii) District of Manhiça with Maragra highlighted in red, (iv) Maragra SA with surrounding fields and village}\label{fig:unnamed-chunk-13}
+\caption{(i) Mozambique in Africa, (ii) Districts of Mozambique with Manhiça highlighted in red, (iii) District of Manhiça with Maragra highlighted in red, (iv) Maragra SA with surrounding fields and village}\label{fig:unnamed-chunk-14}
 \end{figure}
@@
 
-\begin{figure}[h]
-
-{\centering \includegraphics{figures/unnamed-chunk-15-1}
+\begin{figure}[!h]
+
+{\centering \includegraphics{figures/unnamed-chunk-16-1}
 
@@
 
-\caption{i. Fumigation activities carried out by Maragra Malaria Control during study period, ii. Distribution of average time between sprayings of households}\label{fig:unnamed-chunk-15}
+\caption{i. Fumigation activities carried out by Maragra Malaria Control during study period, ii. Distribution of average time between sprayings of households}\label{fig:unnamed-chunk-16}
 \end{figure}
+
+\newpage
 
@@
 
+\newpage
+
+\section{Results}\label{results}
+
 \subsection{Descriptive statistics}\label{descriptive-statistics}
@@
 
-\begin{figure}[h]
-
-{\centering \includegraphics{figures/unnamed-chunk-18-1}
+\begin{figure}[!h]
+
+{\centering \includegraphics{figures/unnamed-chunk-19-1}
 
@@
 
-\caption{Age distribution by sex and worker status}\label{fig:unnamed-chunk-18}
+\caption{Age distribution by sex and worker status}\label{fig:unnamed-chunk-19}
 \end{figure}
@@
 
-\begin{figure}[h]
-
-{\centering \includegraphics{figures/unnamed-chunk-19-1}
+\begin{figure}[!h]
+
+{\centering \includegraphics{figures/unnamed-chunk-20-1}
 
@@
 Monthly total precipitation in the Maniça district; ii.
-Average monthly temperature (bars) during the study period, as well as monthly maximum (red) and minimum (blue) temperatures}\label{fig:unnamed-chunk-19}
+Average monthly temperature (bars) during the study period, as well as monthly maximum (red) and minimum (blue) temperatures}\label{fig:unnamed-chunk-20}
 \end{figure}
-
-\section{Results}\label{results}
 
@@
 \centering
-\includegraphics{figures/unnamed-chunk-20-1.pdf}
+\includegraphics{figures/unnamed-chunk-21-1.pdf}
 \caption{Clinical malaria (district of Manhiça), all-cause absenteeism
@@
 
-\begin{figure}[h]
-
-{\centering \includegraphics{figures/unnamed-chunk-25-1}
+\begin{figure}[!h]
+
+{\centering \includegraphics{figures/unnamed-chunk-26-1}
 
@@
 
-\caption{Rainfall and absenteeism: association of any rainfall with absenteeism rate (left) and association of rainfall amount and absenteeism (right)}\label{fig:unnamed-chunk-25}
+\caption{Rainfall and absenteeism: association of any rainfall with absenteeism rate (left) and association of rainfall amount and absenteeism (right)}\label{fig:unnamed-chunk-26}
 \end{figure}
@@
 
-\textbf{Effect of IRS on absenteeism}: Immediately following IRS, a
-worker's likelihood of absence drops significantly (figure x, panel i).
-As one would expect if the mechanism by which IRS reduces absence is
-through reduced malaria infection, the effect of IRS during the low
-transmission season is significant, but far less substantial in effect
-size (figure x, panel ii).
-
-\begin{figure}
-\centering
-\includegraphics{figures/unnamed-chunk-26-1.pdf}
-\caption{i. Absenteeism before and after IRS administration for all
-workers who ever received IRS; ii. The same, but segregated by malaria
-and non-malaria seasons}
-\end{figure}
+\subsection{Analysis}\label{analysis}
+
+\subsubsection{Effect of IRS on
+absenteeism}\label{effect-of-irs-on-absenteeism}
+
+Immediately following IRS, a worker's likelihood of absence drops
+significantly (appendix figure, panel A). As one would expect if the
+mechanism by which IRS reduces absence is through reduced malaria
+infection, the effect of IRS during the low transmission season is
+significant, but far less substantial in effect size (appendix figure,
+panel B).
 
@@
 
-The model results for sickness only absenteeism suggest that IRS has a
-significant effect on permanent field workers. The effect is
+The model results for sickness only absenteeism (table 5) suggest that
+IRS has a significant effect on permanent field workers. The effect is
 directionally similar among temporary fieldworkers (but does not reach
@@
 
-\subsection{Externalities}\label{externalities}
+\subsubsection{Externalities}\label{externalities}
 
@@
 period prior to IRS administration to be similar to other pre-IRS
-absenteeism (adjusting for seasonality).
-
-What follows are several variations of a model to assess whether IRS
-application predicts pre-IRS absenteeism. In the final version of the
-paper, we'll use whichever model is most appropriate, but this is left
-here for discussion purposes. I've also included the code, so as to
-inform the discussion. In the below models, \texttt{pre\_irs} refers to
-the 10 day period prior to IRS administration. The models estimate the
-odds of absence (ie, the estimates and confidence intervals are in the
-form of odds ratios).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{ro_data <-}\StringTok{ }\NormalTok{ab_panel }\OperatorTok{%>%}\StringTok{ }\KeywordTok{left_join}\NormalTok{(irs, }\DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"date"}\NormalTok{, }\StringTok{"unidade"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(days_since }\OperatorTok{<}\StringTok{ }
-\StringTok{    }\DecValTok{0}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{pre_irs =} \KeywordTok{between}\NormalTok{(days_since, }\OperatorTok{-}\DecValTok{10}\NormalTok{, }\OperatorTok{-}\DecValTok{1}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }\CommentTok{# mutate(time_period = lendable::time_period_extract(date)) %>%}
-\CommentTok{# mutate(time_period = as.character(time_period)) %>%}
-\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{time_period =} \KeywordTok{make_season}\NormalTok{(}\DataTypeTok{date =}\NormalTok{ date)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{absent_sick =} \KeywordTok{ifelse}\NormalTok{(}\KeywordTok{is.na}\NormalTok{(absent_sick), }
-    \OtherTok{FALSE}\NormalTok{, absent_sick)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{left_join}\NormalTok{(workers }\OperatorTok{%>%}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(oracle_number, }
-\NormalTok{    sex, department, permanent_or_temporary), }\DataTypeTok{by =} \StringTok{"oracle_number"}\NormalTok{)}
-
-\NormalTok{make_model <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(}\DataTypeTok{formula =} \StringTok{"absent ~ pre_irs + time_period"}\NormalTok{) \{}
-\NormalTok{    ro_model <-}\StringTok{ }\KeywordTok{glm}\NormalTok{(}\KeywordTok{as.formula}\NormalTok{(formula), }\DataTypeTok{data =}\NormalTok{ ro_data, }\DataTypeTok{family =} \KeywordTok{binomial}\NormalTok{(}\StringTok{"logit"}\NormalTok{))}
-\NormalTok{    coefs <-}\StringTok{ }\KeywordTok{exp}\NormalTok{(}\KeywordTok{coef}\NormalTok{(ro_model))}
-\NormalTok{    ci <-}\StringTok{ }\KeywordTok{confint}\NormalTok{(ro_model)}
-\NormalTok{    ci <-}\StringTok{ }\KeywordTok{exp}\NormalTok{(ci)}
-\NormalTok{    out <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{Variable =} \KeywordTok{names}\NormalTok{(coefs), }\DataTypeTok{Estimate =}\NormalTok{ coefs, }\DataTypeTok{Lower =}\NormalTok{ ci[, }
-        \DecValTok{1}\NormalTok{], }\DataTypeTok{Upper =}\NormalTok{ ci[, }\DecValTok{2}\NormalTok{])}
-    \KeywordTok{row.names}\NormalTok{(out) <-}\StringTok{ }\OtherTok{NULL}
-\NormalTok{    out}\OperatorTok{$}\StringTok{`}\DataTypeTok{P value}\StringTok{`}\NormalTok{ <-}\StringTok{ }\KeywordTok{coef}\NormalTok{(}\KeywordTok{summary}\NormalTok{(ro_model))[, }\DecValTok{4}\NormalTok{]}
-\NormalTok{    out}\OperatorTok{$}\NormalTok{Significant <-}\StringTok{ }\KeywordTok{ifelse}\NormalTok{(out}\OperatorTok{$}\StringTok{`}\DataTypeTok{P value}\StringTok{`} \OperatorTok{<=}\StringTok{ }\FloatTok{0.001}\NormalTok{, }\StringTok{"xxx"}\NormalTok{, }\KeywordTok{ifelse}\NormalTok{(out}\OperatorTok{$}\StringTok{`}\DataTypeTok{P value}\StringTok{`} \OperatorTok{<=}\StringTok{ }
-\StringTok{        }\FloatTok{0.01}\NormalTok{, }\StringTok{"xx"}\NormalTok{, }\KeywordTok{ifelse}\NormalTok{(out}\OperatorTok{$}\StringTok{`}\DataTypeTok{P value}\StringTok{`} \OperatorTok{<=}\StringTok{ }\FloatTok{0.05}\NormalTok{, }\StringTok{"x"}\NormalTok{, }\StringTok{""}\NormalTok{)))}
-    \KeywordTok{return}\NormalTok{(}\KeywordTok{kable}\NormalTok{(out, }\DataTypeTok{format =} \StringTok{"latex"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{kable_styling}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }\KeywordTok{row_spec}\NormalTok{(}\DataTypeTok{row =} \KeywordTok{which}\NormalTok{(}\KeywordTok{grepl}\NormalTok{(}\StringTok{"pre_irs"}\NormalTok{, }
-\NormalTok{        out}\OperatorTok{$}\NormalTok{Variable)), }\DataTypeTok{background =} \StringTok{"yellow"}\NormalTok{))}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# BASIC FORMULA}
-\KeywordTok{make_model}\NormalTok{(}\StringTok{"absent ~ pre_irs + time_period"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
+absenteeism (adjusting for
+seasonality).\todo{Should add model specification formula here.}
+
+The below shows our robustness check for all cause absenteeism.
 
@@
 \hline
-(Intercept) & 0.0498463 & 0.0474720 & 0.0523031 & 0.0000000 & xxx\\
-\hline
-\rowcolor{yellow}  pre\_irsTRUE & 0.8050236 & 0.7002405 & 0.9211332 & 0.0019204 & xx\\
-\hline
-time\_periodMalaria season & 3.1215538 & 2.9408258 & 3.3145230 & 0.0000000 & xxx\\
+(Intercept) & 0.0661899 & 0.0614958 & 0.0711710 & 0.0000000 & xxx\\
+\hline
+\rowcolor{yellow}  10 days prior to IRS & 0.9229176 & 0.7300243 & 1.1503819 & 0.4886243\\
+\hline
+Malaria season & 3.1173457 & 2.9319429 & 3.3156586 & 0.0000000 & xxx\\
+\hline
+departmentFactory & 1.0944261 & 1.0118141 & 1.1841680 & 0.0245210 & x\\
+\hline
+departmentField & 0.5426841 & 0.5036184 & 0.5850353 & 0.0000000 & xxx\\
+\hline
+\rowcolor{yellow}  10 days prior to IRS:Malaria season & 0.8172244 & 0.6161822 & 1.0908542 & 0.1653822\\
 \hline
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# BASIC FORMULA WITH INTERACTIONS}
-\KeywordTok{make_model}\NormalTok{(}\StringTok{"absent ~ pre_irs*time_period"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
+The below shows our robustness check for sick only absenteeism. Unlike
+with all cause absenteeism, during malaria season, being in the period
+10 days prior to IRS is associated with statistically greater likelihood
+of being absent for illness. In other words, it appears that there is
+somewhat of a feedback loop: when a worker misses work due to illness,
+his/her likelihood of getting IRS doubles in the next 10
+days.\todo{What to do about this???}
 
@@
 \hline
-(Intercept) & 0.0496594 & 0.0472551 & 0.0521477 & 0.0000000 & xxx\\
-\hline
-\rowcolor{yellow}  pre\_irsTRUE & 0.8741393 & 0.6918180 & 1.0889051 & 0.2443527\\
-\hline
-time\_periodMalaria season & 3.1397591 & 2.9537482 & 3.3386865 & 0.0000000 & xxx\\
-\hline
-\rowcolor{yellow}  pre\_irsTRUE:time\_periodMalaria season & 0.8804465 & 0.6645722 & 1.1740347 & 0.3797816\\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX}
-\KeywordTok{make_model}\NormalTok{(}\StringTok{"absent ~ pre_irs*time_period + sex"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{l|r|r|r|r|l}
-\hline
-Variable & Estimate & Lower & Upper & P value & Significant\\
-\hline
-(Intercept) & 0.0350661 & 0.0325983 & 0.0376818 & 0.0000000 & xxx\\
-\hline
-\rowcolor{yellow}  pre\_irsTRUE & 0.8880020 & 0.7026608 & 1.1064043 & 0.3043502\\
-\hline
-time\_periodMalaria season & 3.1716504 & 2.9834194 & 3.3729726 & 0.0000000 & xxx\\
-\hline
-sexM & 1.5960091 & 1.4919474 & 1.7085243 & 0.0000000 & xxx\\
-\hline
-\rowcolor{yellow}  pre\_irsTRUE:time\_periodMalaria season & 0.8579023 & 0.6473160 & 1.1443754 & 0.2910119\\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX AND DEPARTMENT}
-\KeywordTok{make_model}\NormalTok{(}\StringTok{"absent ~ pre_irs*time_period + sex + department"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{l|r|r|r|r|l}
-\hline
-Variable & Estimate & Lower & Upper & P value & Significant\\
-\hline
-(Intercept) & 0.0564718 & 0.0510602 & 0.0623971 & 0.0000000 & xxx\\
-\hline
-\rowcolor{yellow}  pre\_irsTRUE & 0.9242121 & 0.7310398 & 1.1520107 & 0.4962745\\
-\hline
-time\_periodMalaria season & 3.1302404 & 2.9439618 & 3.3294967 & 0.0000000 & xxx\\
-\hline
-sexM & 1.1923874 & 1.1059966 & 1.2861150 & 0.0000048 & xxx\\
-\hline
-departmentFactory & 1.0828452 & 1.0009732 & 1.1717926 & 0.0476615 & x\\
-\hline
-departmentField & 0.5767584 & 0.5330008 & 0.6243185 & 0.0000000 & xxx\\
-\hline
-\rowcolor{yellow}  pre\_irsTRUE:time\_periodMalaria season & 0.8145012 & 0.6141075 & 1.0872550 & 0.1585481\\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX AND DEPARTMENT AND}
-\CommentTok{# STATUS}
-\KeywordTok{make_model}\NormalTok{(}\StringTok{"absent ~ pre_irs*time_period + sex + department + permanent_or_temporary"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{l|r|r|r|r|l}
-\hline
-Variable & Estimate & Lower & Upper & P value & Significant\\
-\hline
-(Intercept) & 0.0617511 & 0.0558057 & 0.0682626 & 0.0000000 & xxx\\
-\hline
-\rowcolor{yellow}  pre\_irsTRUE & 1.0466179 & 0.8252949 & 1.3092306 & 0.6983498\\
-\hline
-time\_periodMalaria season & 2.9528835 & 2.7744446 & 3.1438877 & 0.0000000 & xxx\\
-\hline
-sexM & 1.2300194 & 1.1386295 & 1.3293865 & 0.0000002 & xxx\\
-\hline
-departmentFactory & 1.0837360 & 1.0014970 & 1.1731018 & 0.0462328 & x\\
-\hline
-departmentField & 1.3324160 & 1.2272323 & 1.4470610 & 0.0000000 & xxx\\
-\hline
-permanent\_or\_temporaryTemporary & 0.0839961 & 0.0744587 & 0.0944456 & 0.0000000 & xxx\\
-\hline
-\rowcolor{yellow}  pre\_irsTRUE:time\_periodMalaria season & 0.6682310 & 0.5016567 & 0.8956885 & 0.0063480 & xx\\
+(Intercept) & 0.0098162 & 0.0079914 & 0.0119339 & 0.0000000 & xxx\\
+\hline
+\rowcolor{yellow}  10 days prior to IRS & 0.6780540 & 0.3484288 & 1.1791509 & 0.2069281\\
+\hline
+Malaria season & 1.1637020 & 0.9827900 & 1.3767569 & 0.0777162 & \\
+\hline
+departmentFactory & 1.3068546 & 1.0428447 & 1.6460818 & 0.0214002 & x\\
+\hline
+departmentField & 0.6713828 & 0.5396839 & 0.8403182 & 0.0004142 & xxx\\
+\hline
+\rowcolor{yellow}  10 days prior to IRS:Malaria season & 2.2225088 & 1.0747201 & 4.8587934 & 0.0360020 & x\\
 \hline
@@
 \todo{Will address this with some sensitivity analysis}
+
+\newpage
 
@@
 
+\newpage
+
+\section{Appendix}\label{appendix}
+
+\subsection{Unadjusted absenteeism by time since
+IRS}\label{unadjusted-absenteeism-by-time-since-irs}
+
+\begin{figure}
+\centering
+\includegraphics{figures/unnamed-chunk-33-1.pdf}
+\caption{i. Absenteeism before and after IRS administration for all
+workers who ever received IRS; ii. The same, but segregated by malaria
+and non-malaria seasons}
+\end{figure}
+
+\subsection{Unadjusted sick absenteeism by time since
+IRS}\label{unadjusted-sick-absenteeism-by-time-since-irs}
+
+\begin{figure}
+\centering
+\includegraphics{figures/unnamed-chunk-34-1.pdf}
+\caption{i. Sick absenteeism before and after IRS administration for all
+workers who ever received IRS; ii. The same, but segregated by malaria
+and non-malaria seasons}
+\end{figure}
+
 \section*{References}\label{references}

### 2018-04-21 inst/rmd/02_methods.Rmd
externalities report

@@
 Our intervention is whether the residence of the worker in question was treated in the last year, and, if so, the time since treatment, represented, respectively, by $\beta_{2}$ and $\beta_{3}$.
-$\alpha_i$ reprsents the time invariant worker fixed effects, and $\delta_i$ represents the fixed effect of the particular malaria season.
+$\alpha_i$ represents the time invariant worker fixed effects, and $\delta_i$ represents the fixed effect of the particular malaria season.
 $\upsilon$ is the error term.
@@
 In addition, we adjust for daily precipitation; though its lagged effect on malaria incidence is likely captured by the seasonality term, we include rainfall since it's immediate effect (through its impact on transportation and working conditions) may also affect absenteeism.
+
+### Externalities
+
+One potential shortcoming of our analysis thus far is that a "control" worker (ie, one whose house has never been sprayed or was sprayed greater than one year ago) may actually be conferred some protection against malaria by the fumigations at his/her neighbors' houses.
+This positive spillover effect would theoretically go through two channels: (i) via a reduction of mosquitoes in the vicinity and (ii) via a reduction of the malaria parasite in the blood of humans in the vicinity (ie, the parasite "reservoir").
+
+By the same token, just as being surrounded by IRS-treated homes may indirectly protect someone living in a home not treated by IRS, one might surmise that living on the periphery of the company grounds might lead to more malaria risk relative to living in the core, since homes near the periphery but outside of Maragra are less likely to be fumigated.
+In other words, to the extent that both positive and spillover effects are significant enough to warrant inclusion in our model, we would expect that (i) the prevalence of IRS in a worker's neighborhood should be associated with a reduction in absenteeism of the worker and (b) workers living further from the periphery of the company grounds should have lower absenteeism than workers living at the periphery.
+We test both of these hypotheses.
+
+```{r}
+
+```
 

### 2018-04-21 inst/rmd/03_results.Rmd
externalities report

@@
 
-```{r}
-clean_up_model <- function(x){
-  # extract coefficients
-  coefs <- data.frame(coef(summary(x)))
-  # use normal distribution to approximate p-value
-  coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
-  ps <- coefs$p.z
-  x = broom::tidy(x)
-  for(j in 2:(ncol(x) -1)){
-    x[,j] <- x[,j] * 100
-  }
-  x$term <- gsub('seasonhigh', 'Malaria season', x$term)    
-  x$term <- gsub('seasonlow', 'Low malaria season', x$term)
-  x$term <- gsub('months_since', 'Months since IRS: ', x$term)
-  x$term <- gsub('department', 'Department: ', x$term)
-  x$term <- gsub('precipitation', 'Precipitation (mm) ', x$term)
-  x$term <- gsub('maragra_fabricaTRUE', 'Living at factory', x$term)
-  x$term <- gsub('sexM', 'Male', x$term)
-  x$term <- gsub('permanent_or_temporaryTemporary', 'Temp contract', x$term)
-  x$term <- gsub('rainyTRUE', 'Rainy day', x$term)
-  x$term <- gsub('on_siteTRUE', 'On site', x$term)
-  x$term <- gsub('on_siteFALSE', 'Off site', x$term)
-  x$term <- gsub('fieldNot field worker', 'Not field worker', x$term)
-  names(x) <- Hmisc::capitalize(names(x))
-  names(x) <- gsub('.', ' ', names(x), fixed = TRUE)
-  x$`P value` <- NA
-  x$`P value`[1:length(ps)] <- ps
-  x <- x %>% 
-    mutate(Estimate = paste0(round(Estimate, 3), ' (P',
-                             ifelse(`P value` <= 0.001,
-                                    '<0.001',
-                                    paste0('=', round(`P value`, 3))
-                                    ), ')')) %>%
-    dplyr::select(Term, Estimate)
-  x <- x %>% filter(!grepl('sd_', Term))
-  x <- x %>% filter(!grepl('10+', Term, fixed = TRUE))
-  return(x)
-}
-
-make_models_table <- function(model_list, the_caption = "Models with worker fixed effects"){
-
-  out_list <- list()
-  for(i in 1:length(model_list)){
-    message(i)
-    this_model <- names(model_list)[i]
-    the_model <- model_list[[which(names(model_list) == this_model)]]
-    the_model <- clean_up_model(the_model)
-    names(the_model)[2] <- this_model
-    out_list[[i]] <- the_model
-  }
-
-  df <- out_list[[1]]
-  namer <- names(df)[2]
-  names(df)[2] <- 'Estimate'
-
-  for(i in 2:length(out_list)){
-    temp <- out_list[[i]]
-    namer[i] <- names(temp)[2]
-    names(temp)[2] <- 'Estimate'
-    df <- bind_rows(df, temp)
-  }
-
-  library(kableExtra)
-  breaker <- nrow(out_list[[1]])
-  lengther <- length(out_list)
-  breaks <- c(0, breaker * 1:(lengther))
-  k <- kable(df, format = "latex", caption = the_caption, booktabs = T) %>%
-    kable_styling()
-  for (i in 1:(length(breaks)-1)){
-    message(i)
-    k <- k %>%
-      group_rows(namer[i], 
-                 breaks[i] + 1,
-                 breaks[i] + breaker,
-                 latex_gap_space = "1.5em") 
-  }
-  print(k)
-}
-```
-
 The model results for all cause absenteeism (table 4) suggest a significant decrease in absenteeism following the administration of IRS.
@@
 
-### Externalities
+## Externalities
 
-One potential shortcoming of our analysis thus far is that a "control" worker (ie, one whose house has never been sprayed or was sprayed greater than one year ago) may actually be conferred some protection against malaria by the fumigations at his/her neighbors' houses.
-This positive spillover effect would theoretically go through two channels: (i) via a reduction of mosquitoes in the vicinity and (ii) via a reduction of the malaria parasite in the blood of humans in the vicinity (ie, the parasite "reservoir").
+Below are model results for the full protection model.
 
-By the same token, just as being surrounded by IRS-treated homes may indirectly protect someone living in a home not treated by IRS, one might surmise that living on the periphery of the company grounds might lead to more malaria risk relative to living in the core, since homes near the periphery but outside of Maragra are less likely to be fumigated.
-In other words, to the extent that both positive and spillover effects are significant enough to warrant inclusion in our model, we would expect that (i) the prevalence of IRS in a worker's neighborhood should be associated with a reduction in absenteeism of the worker and (b) workers living further from the periphery of the company grounds should have lower absenteeism than workers living at the periphery.
-We test both of these hypotheses.
+```{r, results = 'asis'}
+make_models_table(model_list = protection_models, the_caption = "All absenteeism with herd immunity: model results")
+```
 

### 2018-04-21 inst/rmd/externalities.Rmd
externalities report

---
title: "Concept note"
subtitle: "Modeling positive externalities (spillover) in IRS protection"
date: ''
output: 
  pdf_document:
    keep_tex: true
    latex_engine: xelatex
    includes:
      in_header: preamble.sty
fig_width: 4
fig_height: 2.6
bibliography: bibliography.bib
csl: journal-of-health-economics.csl
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(knitr)
library(RColorBrewer)
library(extrafont)
library(kableExtra)
loadfonts()
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               fig.height = 3,
               fig.width = 4,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               dev = "cairo_pdf",
               fig.pos="!h")
opts_knit$set(width=75)
options(xtable.comment = FALSE)

```

```{r, eval = TRUE}
source('paper_prepare.R')
```

\begin{center}
\begin{large}

Brew

\end{large}
\end{center}

\vspace{5mm}

\begin{center}
\textbf{Overview}  
\end{center}

\vspace{5mm}
\begin{center}
\begin{changemargin}{3cm}{3cm}

The administration of insecticide at a worker's residence (indoor residual spraying or IRS) likely protects that worker from malaria infection by killing the vectors (mosquitoes) that land on the building's walls. However, it is also highly likely that the protective effective of IRS "spills over" to others who live nearby. This positive spillover effect would theoretically go through two channels: (i) via a reduction of mosquitoes in the vicinity and (ii) via a reduction of the malaria parasite in the blood of humans in the vicinity (ie, the parasite "reservoir"). This document describes our method for assessing the existence and magnitude of positive spillover of IRS in the Maragra workers' data. We devise a time-specific household "protection" score based on the theoretical effectiveness of IRS, and then use that protection score to develop a time-place specific "herd protection" score based on a weighted average of nearby household protection scores. We incorporate the latter into our fixed effects model and find that...

\end{changemargin}
\end{center}

\vspace{20mm}

\noindent\fbox{%
    \parbox{\textwidth}{%
        \subsection*{Justification}
        \begin{itemize}
          \item Positive spillover may occur.
          \item Spillover has both a space (ie, distance to sprayed house) and time (ie, time since spraying) dimension.
          \item If spillover occurs and is not accounted for, our models likely underestimate the true effect of IRS since our "control" group (ie, those not receiving IRS) do actually receive IRS (indirectly via spillover).
          \item If spillover does not occur, then this method should demonstrate its non-importance.
        \end{itemize}
        \vspace{2mm}
    }%
}

\vfill
\null

\subsection*{Desinataires}
\textbf{Elisa Sicuri; Menno Pradhan}

\vspace{3mm}

\newpage

## Protection score

To test whether IRS from one household has an effect on workers living in another household, we first must make some assumptions about the level and waning effect of IRS generally. For our purposes, we make a rather simple assumption, based on the literature: that IRS has an immediate effect on mosquitoes and should affect malaria risk following the incubation period, that the waning effect begins in month 4 [@Tukei2017], that IRS effects those not living in the household by reducing reproduction of mosquitoes [White2011], and that IRS' effect after 1 year is essentially null [@Bukirwa2009]. We simplify these assumptions into a time-protection matrix, which can be visualized in the below chart.

```{r}
protection <- 
    data_frame(months_since = c('No IRS / > 1 yr', '01', '02-04', '05-09', '10-12'),
               protection = c(0, 1, 3, 2, 1))
ggplot(data = protection,
       aes(x = months_since,
           y = protection,
           group = 1)) +
  geom_area(fill = 'lightblue',
            alpha = 0.6) +
  geom_line() +
  theme_maragra() +
  labs(x = 'Months since IRS',
       y = '"Protection" level')
```

## Distance-based weighting

Having estimated each household's protection level at all times (ie, the "amount" of protection that they pass on to neighbors), we need to define a method for how much of that total level is passed on. A household which is 1 meter from a neighbor should, theoretically, afford more protection to that household than another which is 1000 meters. To account for the effect of distance, we define a simple function for weighting a nearby residence's contribution to a household's "herd protection" score as follows:

```{r, echo = TRUE}
weighter <- function(x){
    x[x == 0] <- 0.01
    out <- (1 / x)^1.2
    return(out)
}
```

Where `x` is the distance (in kilometers) to the household whose "herd protection" score is being estimated. In other words, a household's "herd protection" score is the mean of all households' protection scores, weighted by the inverse of the distance away raised to the power of 1.2.

Functionally, the weights look like this:

```{r}
x <- seq(0, 1, length = 100)
y <- weighter(x)
x <- data_frame(x,y)
ggplot(data = x,
       aes(x = x,
           y = y)) +
  geom_line() +
  labs(x = 'Km from house',
       y = 'Weight for herd protection score') +
  theme_maragra() +
  geom_area(fill = 'lightblue',
            alpha = 0.6)
```

We heavily weight towards closer houses to account for the relatively small travel distances that most mosquitoes fly in normal conditions [@Verdonschot2014].

## Incorporating into data and model

We calculate a "herd protection" score for each location-date combination (`r nrow(model_data)` values). Whereas our original model specification looked like:

$$
\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * (\hat{\beta}_2{IRS_{it}}*\hat{\beta}_3{IRStime_{it}}) + \alpha_i + \delta_t + \upsilon_{it}
$$

Our new specification incorporates the herd protection value:

$$
\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * (\hat{\beta}_2{IRS_{it}}*\hat{\beta}_3{IRStime_{it}}) +  \hat{\beta}_4{Herd_{it}} +  \alpha_i + \delta_t + \upsilon_{it}
$$

$\hat{Y}_{it}$ is the rate of absence. $\beta_{1}$ is the binary "season" variable, imputed from overall district clinical incidence. Our intervention is whether the residence of the worker in question was treated in the last year, and, if so, the time since treatment, represented, respectively, by $\beta_{2}$ and $\beta_{3}$. $\beta_{4}$ **represents the distance-weighted herd protection score**. $\alpha_i$ represents the time invariant worker fixed effects, and $\delta_i$ represents the fixed effect of the particular malaria season. $\upsilon$ is the error term.

## Results

```{r, results = 'asis'}
make_models_table(model_list = protection_models, the_caption = "All absenteeism with herd immunity: model results")
```

## Conclusion

Using the approach laid out here, there appears to be a significant protective effect afforded by the IRS of neighbors. If this approach is deemed correct, then next steps are determining:

- How to translate the effect into costs and savings (simulation based approach at various IRS coverage levels?).
- How to combine this effect with the other potential spillover effect (ie, core-periphery issues).
- How / where to report on this in the paper (is this a "method" or "result"?).

# References

### 2018-04-21 inst/rmd/externalities.tex
externalities report

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Concept note},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Concept note}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
\subtitle{Modeling positive externalities (spillover) in IRS protection}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\pagenumbering{gobble}
\usepackage{longtable}
\usepackage[utf8]{inputenc}
\usepackage{changepage}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{geometry}
\usepackage{fancyhdr}
\usepackage{color}
\usepackage{colortbl}
\usepackage{color}
% Font
\usepackage{fontspec}
\setmainfont{Swift-Regular_43151.ttf}
\setsansfont[BoldFont={Swift-Bold_43130.ttf}]{Swift-Regular_43151.ttf}
% \setmonofont{Swift-Regular_43151.ttf}
\renewcommand{\familydefault}{\sfdefault}
% \usepackage{fontspec}
% \setmainfont{Lato-Regular.ttf}
% \setsansfont[BoldFont={Lato-Bold.ttf}]{Lato-Regular.ttf}
% \renewcommand{\familydefault}{\sfdefault}

\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist
\renewcommand{\rmdefault}{ppl}

\usepackage{multicol}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{lipsum}

\usepackage{float}
\floatplacement{figure}{H}

% \usepackage{todonotes} % for side notes
% \usepackage[colorinlistoftodos]{todonotes} % for side notes

\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage[dvipsnames, table]{xcolor}  % Coloured text etc.
% 
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\newcommandx{\unsure}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
\newcommandx{\change}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
\newcommandx{\info}[2][1=]{\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25,bordercolor=OliveGreen,#1]{#2}}
\newcommandx{\improvement}[2][1=]{\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
\newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}
\usepackage{lmodern}
\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Return on investment of private sector malaria control at a large sugar facility in Southern Mozambique}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\newcommand{\footremember}[2]{%
    \footnote{#2}
    \newcounter{#1}
    \setcounter{#1}{\value{footnote}}%
}
\newcommand{\footrecall}[1]{%
    \footnotemark[\value{#1}]%
}

\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist

\widowpenalties 1 150

\makeatletter
\renewcommand\footnotesize{%
   \@setfontsize\footnotesize\@ixpt{11}%
   \abovedisplayskip 8\p@ \@plus2\p@ \@minus4\p@
   \abovedisplayshortskip \z@ \@plus\p@
   \belowdisplayshortskip 4\p@ \@plus2\p@ \@minus2\p@
   \def\@listi{\leftmargin\leftmargini
               \topsep 4\p@ \@plus2\p@ \@minus2\p@
               \parsep 2\p@ \@plus\p@ \@minus\p@
               \itemsep \parsep}%
   \belowdisplayskip \abovedisplayskip
}
\makeatother

\DeclareTextCommandDefault{\nobreakspace}{\leavevmode\nobreak\ }
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}

\begin{document}
\maketitle

\begin{center}
\begin{large}

Brew

\end{large}
\end{center}

\vspace{5mm}

\begin{center}
\textbf{Overview}  
\end{center}

\vspace{5mm}

\begin{center}
\begin{changemargin}{3cm}{3cm}

The administration of insecticide at a worker's residence (indoor residual spraying or IRS) likely protects that worker from malaria infection by killing the vectors (mosquitoes) that land on the building's walls. However, it is also highly likely that the protective effective of IRS "spills over" to others who live nearby. This positive spillover effect would theoretically go through two channels: (i) via a reduction of mosquitoes in the vicinity and (ii) via a reduction of the malaria parasite in the blood of humans in the vicinity (ie, the parasite "reservoir"). This document describes our method for assessing the existence and magnitude of positive spillover of IRS in the Maragra workers' data. We devise a time-specific household "protection" score based on the theoretical effectiveness of IRS, and then use that protection score to develop a time-place specific "herd protection" score based on a weighted average of nearby household protection scores. We incorporate the latter into our fixed effects model and find that...

\end{changemargin}
\end{center}

\vspace{20mm}

\noindent\fbox{%
    \parbox{\textwidth}{%
        \subsection*{Justification}
        \begin{itemize}
          \item Positive spillover may occur.
          \item Spillover has both a space (ie, distance to sprayed house) and time (ie, time since spraying) dimension.
          \item If spillover occurs and is not accounted for, our models likely underestimate the true effect of IRS since our "control" group (ie, those not receiving IRS) do actually receive IRS (indirectly via spillover).
          \item If spillover does not occur, then this method should demonstrate its non-importance.
        \end{itemize}
        \vspace{2mm}
    }%
}

\vfill
\null

\subsection*{Desinataires}

\textbf{Elisa Sicuri; Menno Pradhan}

\vspace{3mm}

\newpage

\subsection{Protection score}\label{protection-score}

To test whether IRS from one household has an effect on workers living
in another household, we first must make some assumptions about the
level and waning effect of IRS generally. For our purposes, we make a
rather simple assumption, based on the literature: that IRS has an
immediate effect on mosquitoes and should affect malaria risk following
the incubation period, that the waning effect begins in month 4 (Tukei
et al., 2017), that IRS effects those not living in the household by
reducing reproduction of mosquitoes {[}White2011{]}, and that IRS'
effect after 1 year is essentially null (Bukirwa et al., 2009). We
simplify these assumptions into a time-protection matrix, which can be
visualized in the below chart.

\includegraphics{externalities_files/figure-latex/unnamed-chunk-3-1.pdf}

\subsection{Distance-based weighting}\label{distance-based-weighting}

Having estimated each household's protection level at all times (ie, the
``amount'' of protection that they pass on to neighbors), we need to
define a method for how much of that total level is passed on. A
household which is 1 meter from a neighbor should, theoretically, afford
more protection to that household than another which is 1000 meters. To
account for the effect of distance, we define a simple function for
weighting a nearby residence's contribution to a household's ``herd
protection'' score as follows:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{weighter <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(x) \{}
\NormalTok{    x[x }\OperatorTok{==}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\FloatTok{0.01}
\NormalTok{    out <-}\StringTok{ }\NormalTok{(}\DecValTok{1}\OperatorTok{/}\NormalTok{x)}\OperatorTok{^}\FloatTok{1.2}
    \KeywordTok{return}\NormalTok{(out)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Where \texttt{x} is the distance (in kilometers) to the household whose
``herd protection'' score is being estimated. In other words, a
household's ``herd protection'' score is the mean of all households'
protection scores, weighted by the inverse of the distance away raised
to the power of 1.2.

Functionally, the weights look like this:

\includegraphics{externalities_files/figure-latex/unnamed-chunk-5-1.pdf}

We heavily weight towards closer houses to account for the relatively
small travel distances that most mosquitoes fly in normal conditions
(Verdonschot and Besse-Lototskaya, 2014).

\subsection{Incorporating into data}\label{incorporating-into-data}

We calculate a ``herd protection'' score for each location-date
combination.

\hypertarget{refs}{}
\hypertarget{ref-Bukirwa2009}{}
Bukirwa, H., Yau, V., Kigozi, R., Filler, S., Quick, L., Lugemwa, M.,
Dissanayake, G., Kamya, M., Wabwire-Mangen, F., Dorsey, G., 2009.
Assessing the impact of indoor residual spraying on malaria morbidity
using a sentinel site surveillance system in western uganda. American
Journal of Tropical Medicine and Hygiene 81, 611--614.
\url{https://doi.org/10.4269/ajtmh.2009.09-0126}

\hypertarget{ref-Tukei2017}{}
Tukei, B.B., Beke, A., Lamadrid-Figueroa, H., 2017. Assessing the effect
of indoor residual spraying (IRS) on malaria morbidity in northern
uganda: A before and after study. Malaria Journal 16.
\url{https://doi.org/10.1186/s12936-016-1652-4}

\hypertarget{ref-Verdonschot2014}{}
Verdonschot, P.F., Besse-Lototskaya, A.A., 2014. Flight distance of
mosquitoes (culicidae): A metadata analysis to support the management of
barrier zones around rewetted and newly constructed wetlands.
Limnologica - Ecology and Management of Inland Waters 45, 69--79.
\url{https://doi.org/10.1016/j.limno.2013.11.002}

\end{document}

### 2018-04-21 inst/rmd/paper.Rmd
externalities report

@@
 source('paper_prepare.R')
-irs$months_since <- irs$days_since %/% 30
-
-# Define function for making season
-make_season <- function(date){
-  out <- ifelse(as.numeric(format(date, '%m')) %in% c(11:12, 1:3),
-                   'Malaria season', 'Off season')
-  out <- factor(out, levels = c('Off season', 'Malaria season'))
-  return(out)
-}
 ```

### 2018-04-21 inst/rmd/paper.tex
externalities report

@@
 the level of significance). Among non-fieldworkers, the effect of IRS is
-mixed.
+mixed.\todo{I'll probably remove this from the final version, since we're saying that the data aren't that good...}
 
@@
 
-This section is in progress. Looking at neighborhood effects,
-core-periphery effects, etc.
+One potential shortcoming of our analysis thus far is that a ``control''
+worker (ie, one whose house has never been sprayed or was sprayed
+greater than one year ago) may actually be conferred some protection
+against malaria by the fumigations at his/her neighbors' houses. This
+positive spillover effect would theoretically go through two channels:
+(i) via a reduction of mosquitoes in the vicinity and (ii) via a
+reduction of the malaria parasite in the blood of humans in the vicinity
+(ie, the parasite ``reservoir'').
+
+By the same token, just as being surrounded by IRS-treated homes may
+indirectly protect someone living in a home not treated by IRS, one
+might surmise that living on the periphery of the company grounds might
+lead to more malaria risk relative to living in the core, since homes
+near the periphery but outside of Maragra are less likely to be
+fumigated. In other words, to the extent that both positive and
+spillover effects are significant enough to warrant inclusion in our
+model, we would expect that (i) the prevalence of IRS in a worker's
+neighborhood should be associated with a reduction in absenteeism of the
+worker and (b) workers living further from the periphery of the company
+grounds should have lower absenteeism than workers living at the
+periphery. We test both of these hypotheses.
+
+\paragraph{Neighborhood protective
+effect}\label{neighborhood-protective-effect}
+
+\todo{Working on this now}
+
+\paragraph{Core-periphery difference}\label{core-periphery-difference}
+
+\todo{Working on this now}
 

### 2018-04-21 inst/rmd/paper_prepare.R

@@ -140,11 +140,7 @@ if('prepared_data.RData' %in% dir()){
     #                          ))
     mutate(group = paste0(permanent_or_temporary, ' ',
                           tolower(field)))
-  fe_models <- list()
-  sick_models <- list()
-  groups <- sort(unique(model_data$group))
-  # library(lmerTest)
-  
+
   # Add a month column
   model_data$month_number <- add_zero(format(model_data$date, '%m'), 2)
   
@@ -157,6 +153,94 @@ if('prepared_data.RData' %in% dir()){
                                  paste0(calendar_year-1, '-', calendar_year),
                                  paste0(calendar_year, '-', calendar_year+1)))
   
+  # Get geographic info for externality analysis
+  # Get latitude / longitude into model_data
+  model_data <- model_data %>%
+    left_join(census %>%
+                filter(!duplicated(unidade)) %>%
+                dplyr::select(unidade,
+                              longitude_aura,
+                              latitude_aura),
+              by = 'unidade') %>%
+    dplyr::select(-days_since) %>%
+    left_join(irs %>%
+                dplyr::select(date, unidade, days_since),
+              by = c('date', 'unidade'))
+  
+  # Create a supposed protection level based on time since IRS
+  protection <- 
+    data_frame(months_since = c('No IRS', '01', '02-04', '05-09', '10+'),
+               protection = c(0, 1, 3, 2, 1))
+  model_data <- left_join(model_data,
+                          protection,
+                          by = 'months_since')
+  
+  # Estimate a protection factor based on the weighted protection 
+  # scores of nearby houses
+  library(sp)
+  dates <- sort(unique(model_data$date))
+  out_list <- list()
+  
+  weighter <- function(x){
+    x[x == 0] <- 0.01
+    out <- (1 / x)^1.2
+    # out[is.infinite(out)] <- 0
+    return(out)
+  }
+  counter <- 0
+  for(i in 1:length(dates)){
+    this_date <- dates[i]
+    message(this_date, ' : ', i, ' of ', length(dates))
+    this_model_data <- model_data %>% 
+      filter(!is.na(longitude_aura), 
+             !is.na(latitude_aura),
+             date == this_date,
+             census_name_match_score <= 0.2)
[…]

### 2018-04-28 inst/rmd/externalities.Rmd
big update

@@
 library(kableExtra)
+library(raster)
 loadfonts()
@@
                prompt=FALSE,
-               fig.height = 3,
-               fig.width = 4,
+               fig.height = 4,
+               fig.width = 5,
                tidy=TRUE,
@@
                dev = "cairo_pdf",
-               fig.pos="!h")
+               fig.pos="!h",
+               fig.align = 'center')
 opts_knit$set(width=75)
@@
 To test whether IRS from one household has an effect on workers living in another household, we first must make some assumptions about the level and waning effect of IRS generally.
-For our purposes, we make a rather simple assumption, based on the literature: that IRS has an immediate effect on mosquitoes and should affect malaria risk following the incubation period, that the waning effect begins in month 4 [@Tukei2017], that IRS effects those not living in the household by reducing reproduction of mosquitoes [White2011], and that IRS' effect after 1 year is essentially null [@Bukirwa2009].
+For our purposes, we make a rather simple assumption, based on the literature: that IRS has an immediate effect on mosquitoes and should affect malaria risk following the incubation period, that the waning effect begins in month 4 [@Tukei2017], that IRS effects those not living in the household by reducing reproduction of mosquitoes [@White2011], and that IRS' effect after 1 year is essentially null [@Bukirwa2009].
 We simplify these assumptions into a time-protection matrix, which can be visualized in the below chart.
@@
 ```{r}
-protection <- 
-    data_frame(months_since = c('No IRS / > 1 yr', '01', '02-04', '05-09', '10-12'),
-               protection = c(0, 1, 3, 2, 1))
-ggplot(data = protection,
-       aes(x = months_since,
-           y = protection,
-           group = 1)) +
-  geom_area(fill = 'lightblue',
-            alpha = 0.6) +
+x <- c(1:6)
+y <- (c(-3.45, -6.04, - 6.53, -0.17, -2.74, 5.27) + 27.61) / 27.61
+lo <- loess(y~x, span = 3)$fitted
+lo[lo > 1] <- 1
+lo <- 1-lo
+y <- 1-y
+df <- data.frame(x,y,lo)
+df$months_since <- add_zero(df$x, n = 2)
+df$months_since[df$months_since == '06'] <- '06+'
+df <- df %>%
+  bind_rows(df[nrow(df),] %>%
+              mutate(months_since = 'No IRS'))
+df <- df %>%
+  gather(key, value, y:lo) %>%
+  mutate(key = ifelse(key == 'lo', 
+                      'Brew', 'Tukei'))
+ggplot(data = df,
+       aes(x = x,
+           y = value,
+           group = key,
+           color = key)) +
+  geom_point() +
   geom_line() +
+  # geom_smooth(se = FALSE,span = 3) +
   theme_maragra() +
-  labs(x = 'Months since IRS',
-       y = '"Protection" level')
+  labs(x = 'Months after IRS',
+       y = 'Protection') +
+  geom_hline(lty = 2, 
+             alpha = 0.6,
+             yintercept = 0) +
+  scale_color_manual(name = '',
+                     values = c('blue', 'orange'))
+
 ```
+
+The above chart shows two lines: the orange line is based on Tukei's emperical results regarding the protection afforded by IRS against malaria infection (among children) as a function of time since IRS; the blue line is our simplification of Tukei's work.
 
@@
 weighter <- function(x){
-    x[x == 0] <- 0.01
-    out <- (1 / x)^1.2
-    return(out)
+   1/x
 }
@@
 Where `x` is the distance (in kilometers) to the household whose "herd protection" score is being estimated.
-In other words, a household's "herd protection" score is the mean of all households' protection scores, weighted by the inverse of the distance away raised to the power of 1.2.
+In other words, a household's "herd protection" score is the mean of all households' protection scores, weighted by the inverse of the distance away.
 
@@
 We heavily weight towards closer houses to account for the relatively small travel distances that most mosquitoes fly in normal conditions [@Verdonschot2014].
+
+## Density
+
+We consider a house's "herd" protection level (ie, the protection conferred to the house through externality) to be the average of the other houses' non-herd protection levels, weighted by the distance to the house in question.
+The below is an illustration of a house (the red x) and the neighboring houses (circles).
+The "weight" of each house is indicated by the circle size, and the protection level of the house is indicated by the shading of the circle.
+
+```{r, fig.width = 8}
+make_data <- function(n = 5){
+  df <- data.frame(id = letters[1:n],
+                 x = rnorm(n = n),
+                 y = rnorm(n = n),
+                 protection = round(sample(seq(0, 0.2, length = 100), n), digits = 1),
+                 color = c('red', rep('grey', n-1))) %>%
+  mutate(lng = x,
+         lat = y)
+  df$x[1] <- df$lng[1] <- df$lat[1] <- df$y[1] <- 0
+library(sp)
+coordinates(df) <- ~x+y
+df$distance <- round(spDists(x = df)[1,], digits = 3) / 1
+df$weight <- weighter(df$distance)
+df$herd_protection <- df$protection * df$weight
+return(df)
+}
+
+df <- make_data(n = 10)
+make_plot <- function(df){
+  ggplot(data = df@data,
+       aes(x = lng,
+           y = lat)) +
+  geom_point(aes(size = weight,
+                 color = protection)) +
+  geom_point(data = df@data[1,],
+             color = 'red',
+             pch = 4,
+             cex = 3) +
+  theme_maragra() +
+  labs(x = '',
+       y = '') +
+  theme(legend.position = 'right') +
+  guides(color = guide_colourbar(title = 'Protection', title.position = 'top'),
+         size = guide_legend(nrow = 5, title.position = 'top', nbin = 2)) +
+  scale_color_continuous(low = 'orange', high = 'blue') +
+    xlim(-2,2) +
+    ylim(-2,2)
+}
+make_plot(df = df)
+
+```
+
+However, this consideration misses one important dimension: density.
+The below house's weighted average protection score is identical to that of the above house.
+In other words, at the time in question, at both locations, the percentage of houses with IRS coverage is the same, and the (weighted) average time since IRS is the same.
+However, the below house is likely much more protected than the above, given the _absolute_ number of nearby IRS-protected houses.
+
+```{r, fig.width = 8}
+df <- make_data(n = 100)
+make_plot(df = df)
+```
+
+In the latter representation, the likelihood of a mosquito landing on a DDT-affected surface is much higher than the former, even though the relative IRS coverage for the neighborhood is identical.
+In order to account for this, we must take _density_ itself into account, ie the number of houses (and their protection level and proximity) within 1 kilometer.
+
+Our approach is this: rather than using a weighted average of each household's individual protection score, we instead use a weighted sum.
+A particular household's herd protection score at any given time is the sum of all households' individual protection scores at that time multiplied by those households' distance weights.
+
+## Combining them all together
+
+The summation of (a) the IRS level of other houses at a certain time, (b) weighted by the distance of those houses to the house in question, (c) limited to only the 1 km radius multiplied yields a "herd protection score".
+
+This score is conceptually similar to Cohen and Dupas' quantification of the positive externalities of ITN (bednet) use in Kenya [@Cohen2010] in that it attempts to estimate the protection conferred to "non-users" by users and uses a "weighted sum" (as opposed to average).
+Our approach differs in that the time dimension for IRS is much more important than ITN (ie, IRS is not a binary but rather a waning function, as described previously), requiring us to create herd scores for every day.
+
+Our approach can be justified in that previous studies have found strong positive health externalities in malaria interventions related to ITN coverage [@Alaii2003FactorsAU].
+No studies exist on positive externalities for IRS coverage, but to the extent that the mechanisms for the reduction in infection are similar (reduction in the natural reservoir of the disease, reduction in the number of vectors, etc.), it is reasonable to assume similar effects.
+Additionally, we using weighted distance for our score calculation (rather than simply defining a radius threshold) because of previous studies' findings that there is a linear decline in protection conferred to others with greater distance from an intervention [@Binka2007]
 
@@
 
+\newpage
+
 ## Results
+
+### Table
 
@@
 make_models_table(model_list = protection_models, the_caption = "All absenteeism with herd immunity: model results")
+```
+
+### HIV
+
+We incorporated HIV prevalence of the area lived in, but it had not significance and was then removed from the model.
+
+```{r}
+plot(hiv_prevalence)
+```
+
+## Spatial protection surfaces
+
+The below shows worker locations and the _average_ herd protection score for the entire study period.
+
+```{r}
+# Fortify bairros
+library(maptools)
+library(rgeos)
+border <- gConvexHull(bairros_maragra_bairro)
+border <- SpatialPolygonsDataFrame(Sr = border,
+                                   data = data.frame(a = 1))
+# border <- unionSpatialPolygons(bairros_maragra_bairro, 
+#                           IDs = rep(1, nrow(bairros_maragra_bairro)),
+#                           threshold = 9^9)
+border_fortified <- 
+  broom::tidy(border,
+              regions = 'Id')
+
+# Create average herd risk score
+x <- model_data %>%
+  group_by(lng = longitude_aura,
+           lat = latitude_aura) %>%
+  summarise(herd = mean(herd, na.rm = TRUE)) %>%
+  ungroup %>%
+  filter(!is.na(lng))
+x_sp <- x
+coordinates(x_sp) <- ~lng+lat   
+proj4string(x_sp) <- proj4string(border)
+overs <- sp::over(x_sp, border)
+x <- x[!is.na(overs),]
+ggplot() +
+  geom_point(data = x,
+       aes(x = lng,
+           y = lat,
+           color = herd)) +
+  geom_polygon(data = border_fortified,
+               aes(x = long,
+                   y = lat,
+                   group = group),
+               fill = NA,
+               color = 'black') +
+  ggthemes::theme_map() +
+  scale_color_continuous(name = 'Herd\nprotection\nscore',
+                         low = 'orange',
+                         high = 'blue')
+
+```
+
+The below uses kernel density interpolation (via krieging) to estimate a generalized protection surface over the entire facility.
+
+```{r}
+# Create a gridded dataframe with values 
+# for the entire range (bbox) of border
+df_grid <- expand.grid(lng = seq(bbox(border)[1,1],
+                                 bbox(border)[1,2],
+                                 by = 0.0008),
+                       lat = seq(bbox(border)[2,1],
+                                 bbox(border)[2,2],
+                                 by = 0.0008),
+                       herd = NA)
+df_grid$latitude <- df_grid$lat
+df_grid$longitude <- df_grid$lng
+coordinates(df_grid) <- ~longitude+latitude
+
+# Go through each row of df_grid, getting the 
+# weighted mean irs score for that point
+# and putting a color into df_grid
+if('df_grid.RData' %in% dir()){
+  load('df_grid.RData')
+} else {
+  for (i in 1:nrow(df_grid)){
+  # Get distance from every point in x_sp
+  distances <- spDistsN1(pts = x_sp,
+                        pt = df_grid[i,],
+                        longlat = TRUE)
+  # Define which are acceptably close
+  close_enough <- which(distances <= 50)
+  # Get an herd score
+  herd <- stats::weighted.mean(x = x_sp$herd[close_enough],
+                       w = (1 / distances[close_enough]) ^2,
+                       na.rm = TRUE)
+  # Assign irs to the dataframe
+  df_grid$herd[i] <- herd
+
+}
+save('df_grid', file = 'df_grid.RData')
+}
+# Convert df_grid to raster
+temp <- df_grid@data %>% arrange(lng, lat)
+r <- rasterFromXYZ(temp[, c('lng', 'lat', 'herd')])
+plot(r)
+
+proj4string(df_grid) <- proj4string(border)
+x <- over(df_grid, polygons(border))
+df_grid_small <- df_grid[!is.na(x),]
+temp <- df_grid_small@data %>% arrange(lng, lat)
+r <- rasterFromXYZ(temp[, c('lng', 'lat', 'herd')])
+```
+
+```{r}
+plot(border)
+plot(r, add = TRUE)
+plot(border, add = TRUE)
+```
+
+### Theoretical protection surface
+
+The below shows a theoretical protection surface assuming (a) uniform distribution of residences, (b) random distribution of IRS over time, but never going beyond 6 months between sprays.
+
+```{r}
+theoretical <- function(n){
+  x <- lng <- runif(n = n, min = -1, max = 1)
+y <- lat <- runif(n = n, min = -1, max = 1)
+data <- data.frame(id = 1:n,
+                   x,y, lng, lat,protection = sample(protection$protection[1:5], n,
+                                                     replace = TRUE))
+coordinates(data) <- ~x+y
+
+# Loop through and get scores
+out_list <- list()
+counter <- 0
+distances <- spDists(data)
+for(j in 1:nrow(data)){
+      sub_distances <- distances[j,]
+      x <- data$protection
+      w <- weighter(sub_distances)
+      x <- x[is.finite(w)]
+
+      # The weight
+      w <- w[is.finite(w)]
+      # Number of houses within 1k
+      n <- length(which(sub_distances <= 1))
+      # The weighted protection score (average)
+      positivity <- stats::weighted.mean(x = x,
+                                         w = w,
+                                         na.rm = TRUE)
+      # The weighted additive protection score
+      s <- sum(x[sub_distances <= 1] * w[sub_distances <= 1], na.rm = TRUE)
+      counter <- counter + 1
+
+      # Update model_data
+      out_data <- data_frame(id = data$id[j],
+                             herd = positivity,
+                             n = n,
+                             s = s)
+      out_list[[counter]] <- out_data
+    }
+
+x <- bind_rows(out_list)
+x <- x %>%
+    group_by(id) %>%
+    summarise(herd = mean(herd, na.rm = TRUE),
+              n = mean(n, na.rm = TRUE),
+              s = mean(s, na.rm = TRUE))
+
+data@data <- left_join(data@data, x)
+
+border <- gConvexHull(data)
+border <- SpatialPolygonsDataFrame(Sr = border,
+                                   data = data.frame(a = 1))
+border_fortified <- broom::tidy(border, regions = 1)
+
+a1 <- ggplot() +
+  geom_polygon(data = border_fortified,
+               aes(x = long,
+                   y = lat,
+                   group = group),
+               fill = NA,
+               alpha = 0.2,
+               color = 'black') +
+  geom_point(data = data@data,
+             aes(x = lng,
+                 y = lat,
+                 color = protection)) +
+  theme_maragra() +
+  scale_color_continuous(name = 'Individual\nprotection',
+                         low = 'orange',
+                         high = 'green',
+                         guide = 'legend') +
+  labs(x = '',
+       y = '')
+a2 <- ggplot() +
+  geom_polygon(data = border_fortified,
+               aes(x = long,
+                   y = lat,
+                   group = group),
+               fill = NA,
+               alpha = 0.2,
+               color = 'black') +
+  geom_point(data = data@data,
+             aes(x = lng,
+                 y = lat,
+                 color = s)) +
+  theme_maragra() +
+  scale_color_continuous(name = 'Herd\nprotection',
+                         low = 'orange',
+                         high = 'green',
+                         guide = 'legend') +
+  labs(x = '',
+       y = '')
+Rmisc::multiplot(a1, a2, cols = 2)
+}
+
+```
+
+#### 100 workers
+
+```{r}
+theoretical(n = 100) 
+```
+
+#### 1000 workers
+
+```{r}
+theoretical(n = 1000)
+```
+
+#### 3000 workers
+
+```{r}
+theoretical(n = 3000)
 ```

### 2018-04-28 inst/rmd/externalities.tex
big update

@@
 
+\usepackage{longtable}
+
 \usepackage{float}
@@
 et al., 2017), that IRS effects those not living in the household by
-reducing reproduction of mosquitoes {[}White2011{]}, and that IRS'
+reducing reproduction of mosquitoes (White et al., 2011), and that IRS'
 effect after 1 year is essentially null (Bukirwa et al., 2009). We
@@
 
-\includegraphics{externalities_files/figure-latex/unnamed-chunk-3-1.pdf}
+\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-3-1} \end{center}
+
+The above chart shows two lines: the orange line is based on Tukei's
+emperical results regarding the protection afforded by IRS against
+malaria infection (among children) as a function of time since IRS; the
+blue line is our simplification of Tukei's work.
 
@@
 \NormalTok{weighter <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(x) \{}
-\NormalTok{    x[x }\OperatorTok{==}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\FloatTok{0.01}
-\NormalTok{    out <-}\StringTok{ }\NormalTok{(}\DecValTok{1}\OperatorTok{/}\NormalTok{x)}\OperatorTok{^}\FloatTok{1.2}
-    \KeywordTok{return}\NormalTok{(out)}
+    \DecValTok{1}\OperatorTok{/}\NormalTok{x}
 \NormalTok{\}}
@@
 household's ``herd protection'' score is the mean of all households'
-protection scores, weighted by the inverse of the distance away raised
-to the power of 1.2.
+protection scores, weighted by the inverse of the distance away.
 
@@
 
-\includegraphics{externalities_files/figure-latex/unnamed-chunk-5-1.pdf}
+\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-5-1} \end{center}
 
@@
 
-\subsection{Incorporating into data}\label{incorporating-into-data}
+\subsection{Density}\label{density}
+
+We consider a house's ``herd'' protection level (ie, the protection
+conferred to the house through externality) to be the average of the
+other houses' non-herd protection levels, weighted by the distance to
+the house in question. The below is an illustration of a house (the red
+x) and the neighboring houses (circles). The ``weight'' of each house is
+indicated by the circle size, and the protection level of the house is
+indicated by the shading of the circle.
+
+\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-6-1} \end{center}
+
+However, this consideration misses one important dimension: density. The
+below house's weighted average protection score is identical to that of
+the above house. In other words, at the time in question, at both
+locations, the percentage of houses with IRS coverage is the same, and
+the (weighted) average time since IRS is the same. However, the below
+house is likely much more protected than the above, given the
+\emph{absolute} number of nearby IRS-protected houses.
+
+\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-7-1} \end{center}
+
+In the latter representation, the likelihood of a mosquito landing on a
+DDT-affected surface is much higher than the former, even though the
+relative IRS coverage for the neighborhood is identical. In order to
+account for this, we must take \emph{density} itself into account, ie
+the number of houses (and their protection level and proximity) within 1
+kilometer.
+
+Our approach is this: rather than using a weighted average of each
+household's individual protection score, we instead use a weighted sum.
+A particular household's herd protection score at any given time is the
+sum of all households' individual protection scores at that time
+multiplied by those households' distance weights.
+
+\subsection{Combining them all
+together}\label{combining-them-all-together}
+
+The summation of (a) the IRS level of other houses at a certain time,
+(b) weighted by the distance of those houses to the house in question,
+(c) limited to only the 1 km radius multiplied yields a ``herd
+protection score''.
+
+This score is conceptually similar to Cohen and Dupas' quantification of
+the positive externalities of ITN (bednet) use in Kenya (Cohen and
+Dupas, 2010) in that it attempts to estimate the protection conferred to
+``non-users'' by users and uses a ``weighted sum'' (as opposed to
+average). Our approach differs in that the time dimension for IRS is
+much more important than ITN (ie, IRS is not a binary but rather a
+waning function, as described previously), requiring us to create herd
+scores for every day.
+
+Our approach can be justified in that previous studies have found strong
+positive health externalities in malaria interventions related to ITN
+coverage (Alaii et al., 2003). No studies exist on positive
+externalities for IRS coverage, but to the extent that the mechanisms
+for the reduction in infection are similar (reduction in the natural
+reservoir of the disease, reduction in the number of vectors, etc.), it
+is reasonable to assume similar effects. Additionally, we using weighted
+distance for our score calculation (rather than simply defining a radius
+threshold) because of previous studies' findings that there is a linear
+decline in protection conferred to others with greater distance from an
+intervention (Binka et al., 2007)
+
+\subsection{Incorporating into data and
+model}\label{incorporating-into-data-and-model}
 
 We calculate a ``herd protection'' score for each location-date
-combination.
+combination (1759100 values). Whereas our original model specification
+looked like:
+
+\[
+\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * (\hat{\beta}_2{IRS_{it}}*\hat{\beta}_3{IRStime_{it}}) + \alpha_i + \delta_t + \upsilon_{it}
+\]
+
+Our new specification incorporates the herd protection value:
+
+\[
+\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * (\hat{\beta}_2{IRS_{it}}*\hat{\beta}_3{IRStime_{it}}) +  \hat{\beta}_4{Herd_{it}} +  \alpha_i + \delta_t + \upsilon_{it}
+\]
+
+\(\hat{Y}_{it}\) is the rate of absence. \(\beta_{1}\) is the binary
+``season'' variable, imputed from overall district clinical incidence.
+Our intervention is whether the residence of the worker in question was
+treated in the last year, and, if so, the time since treatment,
+represented, respectively, by \(\beta_{2}\) and \(\beta_{3}\).
+\(\beta_{4}\) \textbf{represents the distance-weighted herd protection
+score}. \(\alpha_i\) represents the time invariant worker fixed effects,
+and \(\delta_i\) represents the fixed effect of the particular malaria
+season. \(\upsilon\) is the error term.
+
+\newpage
+
+\subsection{Results}\label{results}
+
+\subsubsection{Table}\label{table}
+
+\begin{longtable}[t]{ll}
+\caption{\label{tab:unnamed-chunk-8}All absenteeism with herd immunity: model results}\\
+\toprule
+Term & Estimate\\
+\midrule
+\endfirsthead
+\caption[]{All absenteeism with herd immunity: model results \textit{(continued)}}\\
+\toprule
+Term & Estimate\\
+\midrule
+\endhead
+\
+\endfoot
+\bottomrule
+\endlastfoot
+\addlinespace[1.5em]
+\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
+\hspace{1em}Malaria season & 3.597 (P<0.001)\\
+\hspace{1em}Months since IRS: IRS & 6.149 (P<0.001)\\
+\hspace{1em}Rainy day & 3.246 (P<0.001)\\
+\hspace{1em}Herd protection & -0.01 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: IRS & -8.548 (P<0.001)\\
+\addlinespace[1.5em]
+\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
+\hspace{1em}Malaria season & 0.792 (P<0.001)\\
+\hspace{1em}Months since IRS: IRS & 2.243 (P<0.001)\\
+\hspace{1em}Rainy day & 4.552 (P<0.001)\\
+\hspace{1em}Herd protection & -0.009 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: IRS & -1.508 (P=0.002)\\
+\addlinespace[1.5em]
+\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
+\hspace{1em}Malaria season & -0.362 (P<0.001)\\
+\hspace{1em}Months since IRS: IRS & -0.229 (P=0.004)\\
+\hspace{1em}Rainy day & 0.037 (P=0.261)\\
+\hspace{1em}Herd protection & 0.002 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: IRS & 0.26 (P=0.014)\\
+\addlinespace[1.5em]
+\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
+\hspace{1em}Malaria season & 3.151 (P<0.001)\\
+\hspace{1em}Months since IRS: IRS & -1.718 (P=0.076)\\
+\hspace{1em}Rainy day & 0.224 (P=0.469)\\
+\hspace{1em}Herd protection & 0.002 (P=0.772)\\
+\hspace{1em}Malaria season:Months since IRS: IRS & 2.657 (P=0.02)\\*
+\end{longtable}
+
+\subsubsection{HIV}\label{hiv}
+
+We incorporated HIV prevalence of the area lived in, but it had not
+significance and was then removed from the model.
+
+\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-9-1} \end{center}
+
+\subsection{Spatial protection
+surfaces}\label{spatial-protection-surfaces}
+
+The below shows worker locations and the \emph{average} herd protection
+score for the entire study period.
+
+\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-10-1} \end{center}
+
+The below uses kernel density interpolation (via krieging) to estimate a
+generalized protection surface over the entire facility.
+
+\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-11-1} \end{center}
+
+\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-12-1} \end{center}
+
+\subsubsection{Theoretical protection
+surface}\label{theoretical-protection-surface}
+
+The below shows a theoretical protection surface assuming (a) uniform
+distribution of residences, (b) random distribution of IRS over time,
+but never going beyond 6 months between sprays.
+
+\paragraph{100 workers}\label{workers}
+
+\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-14-1} \end{center}
+
+\paragraph{1000 workers}\label{workers-1}
+
+\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-15-1} \end{center}
+
+\paragraph{3000 workers}\label{workers-2}
+
+\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-16-1} \end{center}
+
+\subsection{Conclusion}\label{conclusion}
+
+Using the approach laid out here, there appears to be a significant
+protective effect afforded by the IRS of neighbors. If this approach is
+deemed correct, then next steps are determining:
+
+\begin{itemize}
+\tightlist
+\item
+  How to translate the effect into costs and savings (simulation based
+  approach at various IRS coverage levels?).
+\item
+  How to combine this effect with the other potential spillover effect
+  (ie, core-periphery issues).
+\item
+  How / where to report on this in the paper (is this a ``method'' or
+  ``result''?).
+\end{itemize}
+
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
 
 \hypertarget{refs}{}
+\hypertarget{ref-Alaii2003FactorsAU}{}
+Alaii, J.A., Hawley, W.A., Kolczak, M.S., Kuile, F.O.T., Gimnig, J.E.,
+Vulule, J., Odhacha, A., Oloo, A.J., Nahlen, B.L., Phillips-Howard,
+P.A., 2003. Factors affecting use of permethrin-treated bed nets during
+a randomized controlled trial in western kenya. The American journal of
+tropical medicine and hygiene 68 4 Suppl, 137--41.
+
+\hypertarget{ref-Binka2007}{}
+Binka, F.N., Kubaje, A., Adjuik, M., Williams, L.A., Lengeler, C.,
+Maude, G.H., Armah, G.E., Kajihara, B., Adiamah, J.H., Smith, P.G.,
+2007. Impact of permethrin impregnated bednets on child mortality in
+kassena-nankana district, ghana: A randomized controlled trial. Tropical
+Medicine \& International Health 1, 147--154.
+\url{https://doi.org/10.1111/j.1365-3156.1996.tb00020.x}
+
 \hypertarget{ref-Bukirwa2009}{}
@@
 \url{https://doi.org/10.4269/ajtmh.2009.09-0126}
+
+\hypertarget{ref-Cohen2010}{}
+Cohen, J., Dupas, P., 2010. Free distribution or cost-sharing? Evidence
+from a randomized malaria prevention experiment. Quarterly Journal of
+Economics 125, 1--45. \url{https://doi.org/10.1162/qjec.2010.125.1.1}
 
@@
 
+\hypertarget{ref-White2011}{}
+White, M.T., Griffin, J.T., Churcher, T.S., Ferguson, N.M., Basáñez,
+M.-G., Ghani, A.C., 2011. Modelling the impact of vector control
+interventions on anopheles gambiae population dynamics. Parasites \&
+Vectors 4, 153. \url{https://doi.org/10.1186/1756-3305-4-153}
+
 \end{document}

### 2018-04-28 inst/rmd/paper.tex
big update

@@
 represented, respectively, by \(\beta_{2}\) and \(\beta_{3}\).
-\(\alpha_i\) reprsents the time invariant worker fixed effects, and
+\(\alpha_i\) represents the time invariant worker fixed effects, and
 \(\delta_i\) represents the fixed effect of the particular malaria
@@
 
+\subsubsection{Externalities}\label{externalities}
+
+One potential shortcoming of our analysis thus far is that a ``control''
+worker (ie, one whose house has never been sprayed or was sprayed
+greater than one year ago) may actually be conferred some protection
+against malaria by the fumigations at his/her neighbors' houses. This
+positive spillover effect would theoretically go through two channels:
+(i) via a reduction of mosquitoes in the vicinity and (ii) via a
+reduction of the malaria parasite in the blood of humans in the vicinity
+(ie, the parasite ``reservoir'').
+
+By the same token, just as being surrounded by IRS-treated homes may
+indirectly protect someone living in a home not treated by IRS, one
+might surmise that living on the periphery of the company grounds might
+lead to more malaria risk relative to living in the core, since homes
+near the periphery but outside of Maragra are less likely to be
+fumigated. In other words, to the extent that both positive and
+spillover effects are significant enough to warrant inclusion in our
+model, we would expect that (i) the prevalence of IRS in a worker's
+neighborhood should be associated with a reduction in absenteeism of the
+worker and (b) workers living further from the periphery of the company
+grounds should have lower absenteeism than workers living at the
+periphery. We test both of these hypotheses.
+
 Details will go here about externalities (neighborhood effects, positive
@@
   \hline
-Department & Administrative &  35 & 12523 & 157 & 173151 \\ 
-   & Factory &  71 & 20533 & 453 & 280416 \\ 
-   & Field & 410 & 91302 & 2691 & 1181175 \\ 
-  Sex & F & 228 & 47907 & 1274 & 489484 \\ 
-   & M & 288 & 76451 & 2027 & 1145258 \\ 
-  Status & Permanent & 127 & 47075 & 500 & 655311 \\ 
-   & Temporary & 389 & 77283 & 2801 & 979431 \\ 
-  Age & 20 & 140 & 29463 & 960 & 347758 \\ 
-   & 30 & 176 & 39119 & 1225 & 550783 \\ 
-   & 40 & 123 & 30186 & 652 & 366936 \\ 
-   & 50 &  58 & 18881 & 344 & 268877 \\ 
-   & 60 &  17 & 5893 & 116 & 96555 \\ 
-   & 70 &   2 & 816 &   4 & 3833 \\ 
+Department & Administrative & 157 & 185674 &   0 &   0 \\ 
+   & Factory & 459 & 300949 &   0 &   0 \\ 
+   & Field & 2746 & 1272477 &   0 &   0 \\ 
+  Sex & F & 1305 & 537391 &   0 &   0 \\ 
+   & M & 2057 & 1221709 &   0 &   0 \\ 
+  Status & Permanent & 500 & 702386 &   0 &   0 \\ 
+   & Temporary & 2862 & 1056714 &   0 &   0 \\ 
+  Age & 20 & 985 & 377221 &   0 &   0 \\ 
+   & 30 & 1244 & 589902 &   0 &   0 \\ 
+   & 40 & 665 & 397122 &   0 &   0 \\ 
+   & 50 & 347 & 287758 &   0 &   0 \\ 
+   & 60 & 117 & 102448 &   0 &   0 \\ 
+   & 70 &   4 & 4649 &   0 &   0 \\ 
    \hline
@@
 
-{\centering \includegraphics{figures/unnamed-chunk-19-1}
+{\centering \includegraphics{figures/unnamed-chunk-20-1}
 
@@
 
-\caption{Age distribution by sex and worker status}\label{fig:unnamed-chunk-19}
+\caption{Age distribution by sex and worker status}\label{fig:unnamed-chunk-20}
 \end{figure}
@@
 
-{\centering \includegraphics{figures/unnamed-chunk-20-1}
+{\centering \includegraphics{figures/unnamed-chunk-21-1}
 
@@
 Monthly total precipitation in the Maniça district; ii.
-Average monthly temperature (bars) during the study period, as well as monthly maximum (red) and minimum (blue) temperatures}\label{fig:unnamed-chunk-20}
+Average monthly temperature (bars) during the study period, as well as monthly maximum (red) and minimum (blue) temperatures}\label{fig:unnamed-chunk-21}
 \end{figure}
@@
 \centering
-\includegraphics{figures/unnamed-chunk-21-1.pdf}
+\includegraphics{figures/unnamed-chunk-22-1.pdf}
 \caption{Clinical malaria (district of Manhiça), all-cause absenteeism
@@
 3,998. Among the 3,362 workers for whom we have reliable absenteeism and
-residential data, 692 had their homes fumigated at least once (the
+residential data, 562 had their homes fumigated at least once (the
 majority of workers live off of the facility).
@@
 
-{\centering \includegraphics{figures/unnamed-chunk-26-1}
+{\centering \includegraphics{figures/unnamed-chunk-27-1}
 
@@
 
-\caption{Rainfall and absenteeism: association of any rainfall with absenteeism rate (left) and association of rainfall amount and absenteeism (right)}\label{fig:unnamed-chunk-26}
+\caption{Rainfall and absenteeism: association of any rainfall with absenteeism rate (left) and association of rainfall amount and absenteeism (right)}\label{fig:unnamed-chunk-27}
 \end{figure}
@@
 
-\begin{table}
-
-\caption{\label{tab:unnamed-chunk-28}All cause absenteeism: model results}
-\centering
-\begin{tabular}[t]{ll}
+\begin{longtable}[t]{ll}
+\caption{\label{tab:unnamed-chunk-28}All cause absenteeism: model results}\\
 \toprule
@@
 \midrule
+\endfirsthead
+\caption[]{All cause absenteeism: model results \textit{(continued)}}\\
+\toprule
+Term & Estimate\\
+\midrule
+\endhead
+\
+\endfoot
+\bottomrule
+\endlastfoot
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Permanent field worker}}\\
-\hspace{1em}Malaria season & 2.533 (P<0.001)\\
-\hspace{1em}Months since IRS: 01 & 5.382 (P<0.001)\\
-\hspace{1em}Months since IRS: 02-04 & 4.747 (P<0.001)\\
-\hspace{1em}Months since IRS: 05-09 & -2.669 (P<0.001)\\
-\hspace{1em}Precipitation (mm) & 0.206 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & -6.262 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & -8.388 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & -3.71 (P<0.001)\\
+\hspace{1em}Malaria season & 2.458 (P<0.001)\\
+\hspace{1em}Months since IRS: 00 & 4.838 (P<0.001)\\
+\hspace{1em}Months since IRS: 01 & 6.013 (P<0.001)\\
+\hspace{1em}Months since IRS: 02 & 5.395 (P<0.001)\\
+\hspace{1em}Months since IRS: 03 & 5.596 (P<0.001)\\
+\hspace{1em}Months since IRS: 04 & 2.47 (P=0.052)\\
+\hspace{1em}Months since IRS: 05 & 2.925 (P=0.015)\\
+\hspace{1em}Rainy day & 2.236 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 00 & -10.488 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & -1.828 (P=0.187)\\
+\hspace{1em}Malaria season:Months since IRS: 02 & -7.825 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 03 & -10.95 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 04 & -5.517 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 05 & -11.014 (P<0.001)\\
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
-\hspace{1em}Malaria season & 0.957 (P<0.001)\\
-\hspace{1em}Months since IRS: 01 & -0.987 (P=0.071)\\
-\hspace{1em}Months since IRS: 02-04 & 4.375 (P<0.001)\\
-\hspace{1em}Months since IRS: 05-09 & -0.55 (P=0.254)\\
-\hspace{1em}Precipitation (mm) & 0.548 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & 1.787 (P=0.016)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & -5.208 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & -2.04 (P=0.003)\\
+\hspace{1em}Malaria season & 0.955 (P<0.001)\\
+\hspace{1em}Months since IRS: 00 & -2.358 (P=0.002)\\
+\hspace{1em}Months since IRS: 01 & 0.709 (P=0.356)\\
+\hspace{1em}Months since IRS: 02 & 2.866 (P<0.001)\\
+\hspace{1em}Months since IRS: 03 & 6.045 (P<0.001)\\
+\hspace{1em}Months since IRS: 04 & 4.113 (P<0.001)\\
+\hspace{1em}Months since IRS: 05 & -0.144 (P=0.877)\\
+\hspace{1em}Rainy day & 4.503 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 00 & 2.792 (P=0.008)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & 0.502 (P=0.625)\\
+\hspace{1em}Malaria season:Months since IRS: 02 & -5.816 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 03 & -4.871 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 04 & -5.815 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 05 & 1.719 (P=0.183)\\
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Temporary field worker}}\\
-\hspace{1em}Malaria season & -0.368 (P<0.001)\\
-\hspace{1em}Months since IRS: 01 & -0.075 (P=0.526)\\
-\hspace{1em}Months since IRS: 02-04 & -0.193 (P=0.072)\\
-\hspace{1em}Months since IRS: 05-09 & -0.036 (P=0.738)\\
-\hspace{1em}Precipitation (mm) & 0.02 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & -0.154 (P=0.406)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & 0.297 (P=0.05)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & 0.041 (P=0.769)\\
+\hspace{1em}Malaria season & -0.356 (P<0.001)\\
+\hspace{1em}Months since IRS: 00 & -0.093 (P=0.57)\\
+\hspace{1em}Months since IRS: 01 & -0.006 (P=0.972)\\
+\hspace{1em}Months since IRS: 02 & -0.232 (P=0.156)\\
+\hspace{1em}Months since IRS: 03 & -0.177 (P=0.305)\\
+\hspace{1em}Months since IRS: 04 & -0.059 (P=0.763)\\
+\hspace{1em}Months since IRS: 05 & 0.039 (P=0.86)\\
+\hspace{1em}Rainy day & 0.028 (P=0.236)\\
+\hspace{1em}Malaria season:Months since IRS: 00 & -0.114 (P=0.659)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & -0.269 (P=0.301)\\
+\hspace{1em}Malaria season:Months since IRS: 02 & -0.212 (P=0.432)\\
+\hspace{1em}Malaria season:Months since IRS: 03 & 0.127 (P=0.628)\\
+\hspace{1em}Malaria season:Months since IRS: 04 & 0.808 (P=0.002)\\
+\hspace{1em}Malaria season:Months since IRS: 05 & 0.682 (P=0.013)\\
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
-\hspace{1em}Malaria season & 2.025 (P<0.001)\\
-\hspace{1em}Months since IRS: 01 & -2.536 (P=0.134)\\
-\hspace{1em}Months since IRS: 02-04 & -1.177 (P=0.282)\\
-\hspace{1em}Months since IRS: 05-09 & -1.821 (P=0.084)\\
-\hspace{1em}Precipitation (mm) & 0.078 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & 6.121 (P=0.002)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & 1.333 (P=0.343)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & -4.639 (P<0.001)\\
-\bottomrule
-\end{tabular}
-\end{table}
+\hspace{1em}Malaria season & 1.897 (P<0.001)\\
+\hspace{1em}Months since IRS: 00 & -0.929 (P=0.726)\\
+\hspace{1em}Months since IRS: 01 & -2.264 (P=0.251)\\
+\hspace{1em}Months since IRS: 02 & 1.086 (P=0.517)\\
+\hspace{1em}Months since IRS: 03 & 2.083 (P=0.218)\\
+\hspace{1em}Months since IRS: 04 & -4.187 (P=0.02)\\
+\hspace{1em}Months since IRS: 05 & -3.224 (P=0.111)\\
+\hspace{1em}Rainy day & 1.053 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 00 & 0.754 (P=0.797)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & 14.058 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 02 & -3.674 (P=0.128)\\
+\hspace{1em}Malaria season:Months since IRS: 03 & -3.644 (P=0.119)\\
+\hspace{1em}Malaria season:Months since IRS: 04 & 12.4 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 05 & -2.697 (P=0.334)\\*
+\end{longtable}
 
@@
 
-\begin{table}
-
-\caption{\label{tab:unnamed-chunk-29}Sick absenteeism only: model results}
-\centering
-\begin{tabular}[t]{ll}
+\begin{longtable}[t]{ll}
+\caption{\label{tab:unnamed-chunk-29}Sick absenteeism only: model results}\\
 \toprule
@@
 \midrule
+\endfirsthead
+\caption[]{Sick absenteeism only: model results \textit{(continued)}}\\
+\toprule
+Term & Estimate\\
+\midrule
+\endhead
+\
+\endfoot
+\bottomrule
+\endlastfoot
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Permanent field worker}}\\
-\hspace{1em}Malaria season & -0.073 (P=0.089)\\
-\hspace{1em}Months since IRS: 01 & 0.891 (P<0.001)\\
-\hspace{1em}Months since IRS: 02-04 & 1.223 (P<0.001)\\
-\hspace{1em}Months since IRS: 05-09 & -0.236 (P=0.249)\\
-\hspace{1em}Precipitation (mm) & 0 (P=0.984)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & -1.501 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & -0.597 (P=0.031)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & -0.484 (P=0.097)\\
+\hspace{1em}Malaria season & -0.068 (P=0.116)\\
+\hspace{1em}Months since IRS: 00 & 1.358 (P<0.001)\\
+\hspace{1em}Months since IRS: 01 & 0.511 (P=0.141)\\
+\hspace{1em}Months since IRS: 02 & 2.272 (P<0.001)\\
+\hspace{1em}Months since IRS: 03 & 1.21 (P<0.001)\\
+\hspace{1em}Months since IRS: 04 & -0.057 (P=0.889)\\
+\hspace{1em}Months since IRS: 05 & -0.295 (P=0.442)\\
+\hspace{1em}Rainy day & -0.053 (P=0.207)\\
+\hspace{1em}Malaria season:Months since IRS: 00 & -2.046 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & -1.037 (P=0.019)\\
+\hspace{1em}Malaria season:Months since IRS: 02 & -1.691 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 03 & -0.428 (P=0.365)\\
+\hspace{1em}Malaria season:Months since IRS: 04 & 0.52 (P=0.305)\\
+\hspace{1em}Malaria season:Months since IRS: 05 & -0.617 (P=0.257)\\
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
-\hspace{1em}Malaria season & -0.056 (P=0.204)\\
-\hspace{1em}Months since IRS: 01 & -0.667 (P=0.003)\\
-\hspace{1em}Months since IRS: 02-04 & 0.904 (P<0.001)\\
-\hspace{1em}Months since IRS: 05-09 & 0.127 (P=0.526)\\
-\hspace{1em}Precipitation (mm) & -0.002 (P=0.565)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & 1.121 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & -0.122 (P=0.652)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & 1.198 (P<0.001)\\
+\hspace{1em}Malaria season & -0.056 (P=0.216)\\
+\hspace{1em}Months since IRS: 00 & -0.315 (P=0.317)\\
+\hspace{1em}Months since IRS: 01 & -1.009 (P=0.002)\\
+\hspace{1em}Months since IRS: 02 & -0.504 (P=0.131)\\
+\hspace{1em}Months since IRS: 03 & 1.483 (P<0.001)\\
+\hspace{1em}Months since IRS: 04 & 2.029 (P<0.001)\\
+\hspace{1em}Months since IRS: 05 & -0.266 (P=0.492)\\
+\hspace{1em}Rainy day & -0.012 (P=0.785)\\
+\hspace{1em}Malaria season:Months since IRS: 00 & 0.65 (P=0.137)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & 1.491 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 02 & 1.06 (P=0.016)\\
+\hspace{1em}Malaria season:Months since IRS: 03 & -0.973 (P=0.035)\\
+\hspace{1em}Malaria season:Months since IRS: 04 & -0.709 (P=0.158)\\
+\hspace{1em}Malaria season:Months since IRS: 05 & 2.82 (P<0.001)\\
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Temporary field worker}}\\
-\hspace{1em}Malaria season & -0.106 (P<0.001)\\
-\hspace{1em}Months since IRS: 01 & 0.029 (P=0.545)\\
-\hspace{1em}Months since IRS: 02-04 & 0.04 (P=0.35)\\
-\hspace{1em}Months since IRS: 05-09 & -0.046 (P=0.291)\\
-\hspace{1em}Precipitation (mm) & 0 (P=0.966)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & -0.108 (P=0.148)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & -0.032 (P=0.595)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & -0.052 (P=0.36)\\
+\hspace{1em}Malaria season & -0.104 (P<0.001)\\
+\hspace{1em}Months since IRS: 00 & 0.204 (P=0.002)\\
+\hspace{1em}Months since IRS: 01 & -0.137 (P=0.03)\\
+\hspace{1em}Months since IRS: 02 & -0.056 (P=0.392)\\
+\hspace{1em}Months since IRS: 03 & 0.122 (P=0.077)\\
+\hspace{1em}Months since IRS: 04 & 0.061 (P=0.432)\\
+\hspace{1em}Months since IRS: 05 & -0.163 (P=0.063)\\
+\hspace{1em}Rainy day & -0.018 (P=0.05)\\
+\hspace{1em}Malaria season:Months since IRS: 00 & -0.307 (P=0.003)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & 0.107 (P=0.306)\\
+\hspace{1em}Malaria season:Months since IRS: 02 & 0.064 (P=0.557)\\
+\hspace{1em}Malaria season:Months since IRS: 03 & 0.077 (P=0.464)\\
+\hspace{1em}Malaria season:Months since IRS: 04 & -0.185 (P=0.079)\\
+\hspace{1em}Malaria season:Months since IRS: 05 & 0.056 (P=0.613)\\
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
-\hspace{1em}Malaria season & -0.135 (P=0.236)\\
-\hspace{1em}Months since IRS: 01 & 1.427 (P=0.092)\\
-\hspace{1em}Months since IRS: 02-04 & -0.569 (P=0.299)\\
-\hspace{1em}Months since IRS: 05-09 & -1.42 (P=0.007)\\
-\hspace{1em}Precipitation (mm) & -0.005 (P=0.613)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & -1.796 (P=0.069)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & 2.479 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & 0.18 (P=0.794)\\
+\hspace{1em}Malaria season & -0.15 (P=0.191)\\
+\hspace{1em}Months since IRS: 00 & -0.401 (P=0.765)\\
+\hspace{1em}Months since IRS: 01 & 2.063 (P=0.038)\\
+\hspace{1em}Months since IRS: 02 & -0.508 (P=0.549)\\
+\hspace{1em}Months since IRS: 03 & 0.019 (P=0.982)\\
+\hspace{1em}Months since IRS: 04 & -1.905 (P=0.036)\\
+\hspace{1em}Months since IRS: 05 & -2.031 (P=0.047)\\
+\hspace{1em}Rainy day & -0.036 (P=0.711)\\
+\hspace{1em}Malaria season:Months since IRS: 00 & 0.005 (P=0.997)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & -2.43 (P=0.071)\\
+\hspace{1em}Malaria season:Months since IRS: 02 & 0.083 (P=0.946)\\
+\hspace{1em}Malaria season:Months since IRS: 03 & -1.915 (P=0.105)\\
+\hspace{1em}Malaria season:Months since IRS: 04 & 5.662 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 05 & 1.729 (P=0.22)\\*
+\end{longtable}
+
+\subsection{Externalities}\label{externalities-1}
+
+Below are model results for the full protection model.
+
+\begin{longtable}[t]{ll}
+\caption{\label{tab:unnamed-chunk-30}All absenteeism with herd immunity: model results}\\
+\toprule
+Term & Estimate\\
+\midrule
+\endfirsthead
+\caption[]{All absenteeism with herd immunity: model results \textit{(continued)}}\\
+\toprule
+Term & Estimate\\
+\midrule
+\endhead
+\
+\endfoot
 \bottomrule
-\end{tabular}
-\end{table}
-
-\subsubsection{Externalities}\label{externalities}
-
-One potential shortcoming of our analysis thus far is that a ``control''
-worker (ie, one whose house has never been sprayed or was sprayed
-greater than one year ago) may actually be conferred some protection
-against malaria by the fumigations at his/her neighbors' houses. This
-positive spillover effect would theoretically go through two channels:
-(i) via a reduction of mosquitoes in the vicinity and (ii) via a
-reduction of the malaria parasite in the blood of humans in the vicinity
-(ie, the parasite ``reservoir'').
-
-By the same token, just as being surrounded by IRS-treated homes may
-indirectly protect someone living in a home not treated by IRS, one
-might surmise that living on the periphery of the company grounds might
-lead to more malaria risk relative to living in the core, since homes
-near the periphery but outside of Maragra are less likely to be
-fumigated. In other words, to the extent that both positive and
-spillover effects are significant enough to warrant inclusion in our
-model, we would expect that (i) the prevalence of IRS in a worker's
-neighborhood should be associated with a reduction in absenteeism of the
-worker and (b) workers living further from the periphery of the company
-grounds should have lower absenteeism than workers living at the
-periphery. We test both of these hypotheses.
+\endlastfoot
+\addlinespace[1.5em]
+\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
+\hspace{1em}Malaria season & 3.595 (P<0.001)\\
+\hspace{1em}Months since IRS: 00 & 4.684 (P<0.001)\\
+\hspace{1em}Months since IRS: 01 & 7.368 (P<0.001)\\
+\hspace{1em}Months since IRS: 02 & 7.572 (P<0.001)\\
+\hspace{1em}Months since IRS: 03 & 7.675 (P<0.001)\\
+\hspace{1em}Months since IRS: 04 & 3.808 (P=0.005)\\
+\hspace{1em}Months since IRS: 05 & 3.612 (P=0.004)\\
+\hspace{1em}Rainy day & 3.251 (P<0.001)\\
+\hspace{1em}Herd protection & -0.208 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 00 & -10.583 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & -2.757 (P=0.061)\\
+\hspace{1em}Malaria season:Months since IRS: 02 & -8.816 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 03 & -11.784 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 04 & -6.39 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 05 & -12.58 (P<0.001)\\
+\addlinespace[1.5em]
+\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
+\hspace{1em}Malaria season & 0.812 (P<0.001)\\
+\hspace{1em}Months since IRS: 00 & -1.763 (P=0.024)\\
+\hspace{1em}Months since IRS: 01 & 1.383 (P=0.083)\\
+\hspace{1em}Months since IRS: 02 & 4.06 (P<0.001)\\
+\hspace{1em}Months since IRS: 03 & 5.191 (P<0.001)\\
+\hspace{1em}Months since IRS: 04 & 4.688 (P<0.001)\\
+\hspace{1em}Months since IRS: 05 & -0.158 (P=0.87)\\
+\hspace{1em}Rainy day & 4.491 (P<0.001)\\
+\hspace{1em}Herd protection & -0.051 (P=0.021)\\
+\hspace{1em}Malaria season:Months since IRS: 00 & 2.437 (P=0.025)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & 0.144 (P=0.892)\\
+\hspace{1em}Malaria season:Months since IRS: 02 & -6.423 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 03 & -2.935 (P=0.011)\\
+\hspace{1em}Malaria season:Months since IRS: 04 & -5.391 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 05 & 0.971 (P=0.473)\\
+\addlinespace[1.5em]
+\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
+\hspace{1em}Malaria season & -0.35 (P<0.001)\\
+\hspace{1em}Months since IRS: 00 & -0.148 (P=0.351)\\
+\hspace{1em}Months since IRS: 01 & -0.22 (P=0.156)\\
+\hspace{1em}Months since IRS: 02 & -0.528 (P=0.001)\\
+\hspace{1em}Months since IRS: 03 & -0.431 (P=0.012)\\
+\hspace{1em}Months since IRS: 04 & -0.285 (P=0.138)\\
+\hspace{1em}Months since IRS: 05 & -0.021 (P=0.921)\\
+\hspace{1em}Rainy day & 0.071 (P=0.041)\\
+\hspace{1em}Herd protection & 0.024 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 00 & -0.165 (P=0.514)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & -0.152 (P=0.55)\\
+\hspace{1em}Malaria season:Months since IRS: 02 & -0.039 (P=0.884)\\
+\hspace{1em}Malaria season:Months since IRS: 03 & 0.149 (P=0.562)\\
+\hspace{1em}Malaria season:Months since IRS: 04 & 0.83 (P=0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 05 & 0.643 (P=0.017)\\
+\addlinespace[1.5em]
+\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
+\hspace{1em}Malaria season & 3.49 (P<0.001)\\
+\hspace{1em}Months since IRS: 00 & -0.214 (P=0.938)\\
+\hspace{1em}Months since IRS: 01 & -1.84 (P=0.374)\\
+\hspace{1em}Months since IRS: 02 & 1.463 (P=0.414)\\
+\hspace{1em}Months since IRS: 03 & 2.207 (P=0.217)\\
+\hspace{1em}Months since IRS: 04 & -4.231 (P=0.024)\\
+\hspace{1em}Months since IRS: 05 & -3.542 (P=0.091)\\
+\hspace{1em}Rainy day & 0.514 (P=0.101)\\
+\hspace{1em}Herd protection & 0.012 (P=0.918)\\
+\hspace{1em}Malaria season:Months since IRS: 00 & -0.626 (P=0.837)\\
+\hspace{1em}Malaria season:Months since IRS: 01 & 13.506 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 02 & -3.399 (P=0.18)\\
+\hspace{1em}Malaria season:Months since IRS: 03 & -3.826 (P=0.115)\\
+\hspace{1em}Malaria season:Months since IRS: 04 & 10.956 (P<0.001)\\
+\hspace{1em}Malaria season:Months since IRS: 05 & -3.786 (P=0.191)\\*
+\end{longtable}
 
@@
 \centering
-\includegraphics{figures/unnamed-chunk-33-1.pdf}
+\includegraphics{figures/unnamed-chunk-34-1.pdf}
 \caption{i. Absenteeism before and after IRS administration for all
@@
 \centering
-\includegraphics{figures/unnamed-chunk-34-1.pdf}
+\includegraphics{figures/unnamed-chunk-35-1.pdf}
 \caption{i. Sick absenteeism before and after IRS administration for all

### 2018-04-28 data-raw/create_data_files.R

@@ -1183,7 +1183,28 @@ weather <- weather %>%
          date >= '2013-01-01',
          date <= '2016-12-31')
 
+# Load the hiv prevalence data
+# create in the laura_de_la_fuente repo / separate project
+load('hiv_prevalence.RData')
+hiv_prevalence <- r
+
+# Bairros
+x <- bairros
+x$NeighCode <- as.numeric(as.character(x$NeighCode))
+x@data <- x@data %>%
+  mutate(maragra_bairro = NeighCode %in%
+           c(601:609,701:709, 1001:1003)) %>%
+  mutate(maragra_fabrica = NeighCode %in%
+           1001:1003)
+# x <- spTransform(x,CRS("+proj=utm +zone=36 +datum=WGS84 +units=m") )
+
+bairros_maragra_bairro <- x[x$maragra_bairro,]
+bairros_maragra_fabrica <- x[x$maragra_fabrica,]
+# the_bairro <- spTransform(the_bairro, CRS("+proj=utm +zone=36 +datum=WGS84 +units=m") )
+# the_fabrica <- spTransform(the_fabrica, CRS("+proj=utm +zone=36 +datum=WGS84 +units=m") )
+
 # Save for use in package
+devtools::use_data(hiv_prevalence, overwrite = TRUE)
 devtools::use_data(bes,
                    overwrite = TRUE)
 devtools::use_data(ab,
@@ -1199,6 +1220,10 @@ devtools::use_data(workers,
 devtools::use_data(ab_panel,
                    overwrite = TRUE)
 devtools::use_data(bairros,
+                   overwrite = TRUE)
+devtools::use_data(bairros_maragra_bairro,
+                   overwrite = TRUE)
+devtools::use_data(bairros_maragra_fabrica,
                    overwrite = TRUE)
 devtools::use_data(census,
                    overwrite = TRUE)

### 2018-04-28 inst/rmd/paper_prepare.R

@@ -46,8 +46,9 @@ if('prepared_data.RData' %in% dir()){
     mutate(days_since = days_since %/% 30) %>%
     mutate(days_since = ifelse(is.na(days_since), 'Never',
                                ifelse(days_since < 0, 'Before',
-                                      ifelse(days_since >= 12, '12+', as.character(add_zero(days_since, n = 2)))))) %>%
-    mutate(months_since = factor(days_since, levels = unique(c('Never', 'Before', sort(unique(days_since)))))) %>%
+                                      ifelse(days_since >= 12, '12+', as.character(add_zero(days_since, n = 2)))))) 
+  model_data <-
+    model_data %>%
     mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) %>%
     # Add malaria incidence
     left_join(bes %>%
@@ -90,15 +91,14 @@ if('prepared_data.RData' %in% dir()){
   
   # Aggregate months since 
   model_data <- model_data %>%
+    mutate(months_since = days_since) %>%
     mutate(months_since = as.character(months_since)) %>%
-    mutate(months_since = ifelse(months_since %in% c('Never', 'Before'),
+    mutate(months_since = ifelse(months_since %in% c('Never', 'Before', '00', '01',
+                                                     '02', '03', '04', '05'),
                                  months_since,
-                                 ifelse(months_since %in% c('00', '01'), '01',
-                                        ifelse(months_since %in% c('02', '03', '04'),
-                                               '02-04',
-                                               ifelse(months_since %in% c('05', '06', '07', '08', '09'),
-                                                      '05-09',
-                                                      ifelse(months_since %in% c('10', '11', '12+'), '10+', months_since)))))) %>%
+                                 ifelse(months_since %in% c('06', '07', '08',
+                                                            '09', '10', '11', '12+'),
+                                        '06+', NA))) %>%
     # Define whether every sprayed
     ungroup %>%
     group_by(oracle_number) %>%
@@ -111,28 +111,36 @@ if('prepared_data.RData' %in% dir()){
     # Make never and before the same
     mutate(months_since = ifelse(months_since %in% c('Never', 'Before'), 'No IRS', months_since)) %>%
[…]

### 2018-05-06 inst/rmd/02_methods.Rmd
externalities text

--- 2018-04-20 inst/rmd/02_methods.Rmd
@@
 To estimate this effect, we estimated separate models for absence and illness.
-We employed linear mixed effects models [@Minalu2011] [@Liu2010] (with individual worker fixed effects, and random effects for other covariates)  in lieu of interrupted time series [@Lopez_Bernal_2016] so as to explicitly adjust for confounders [@Bell2014].
+We employed linear mixed effects models [@Minalu2011] [@Liu2010] (with individual worker fixed effects, and random effects for other covariates) so as to explicitly adjust for confounders [@Bell2014].
 We divide into 4 different models for different worker types, so as to account for the potential time-varying effects of worker type on risk of malaria.
+
+### Externalities
+
+The administration of insecticide at a worker's residence (indoor residual spraying or IRS) likely protects that worker from malaria infection by killing the vectors (mosquitoes) that land on the building's walls.
+However, it is also highly likely that the protective effective of IRS "spills over" to others who live nearby.
+This positive spillover effect would theoretically go through two channels: (i) via a reduction of mosquitoes in the vicinity and (ii) via a reduction of the malaria parasite in the blood of humans in the vicinity (ie, the parasite "reservoir").
+This document describes our method for assessing the existence and magnitude of positive spillover of IRS in the Maragra workers' data.
+We devise a time-specific household "protection" score based on the theoretical effectiveness of IRS, and then use that protection score to develop a time-place specific "herd protection" score based on a weighted average of nearby household protection scores.
+
+We consider a house's "herd" protection level (ie, the protection conferred to the house through externality) to be the average of the other houses' non-herd protection levels, weighted by the distance to the house in question.
+The below is an illustration of a house (the red x) and the neighboring houses (circles).
+The "weight" of each house is indicated by the circle size, and the protection level of the house is indicated by the shading of the circle.
+
+However, this consideration misses one important dimension: density.
+The below house's weighted average protection score is identical to that of the above house.
+In other words, at the time in question, at both locations, the percentage of houses with IRS coverage is the same, and the (weighted) average time since IRS is the same.
+However, the below house is likely much more protected than the above, given the _absolute_ number of nearby IRS-protected houses.
+
+Our approach is this: rather than using a weighted average of each household's individual protection score, we instead use a weighted sum.
+A particular household's herd protection score at any given time is the sum of all households' individual protection scores at that time multiplied by those households' distance weights.
+
+The summation of (a) the IRS level of other houses at a certain time, (b) weighted by the distance of those houses to the house in question, (c) limited to only the 1 km radius multiplied yields a "herd protection score".
+
+This score is conceptually similar to Cohen and Dupas' quantification of the positive externalities of ITN (bednet) use in Kenya [@Cohen2010] in that it attempts to estimate the protection conferred to "non-users" by users and uses a "weighted sum" (as opposed to average).
+Our approach differs in that the time dimension for IRS is much more important than ITN (ie, IRS is not a binary but rather a waning function, as described previously), requiring us to create herd scores for every day.
+
+Our approach can be justified in that previous studies have found strong positive health externalities in malaria interventions related to ITN coverage [@Alaii2003FactorsAU].
+No studies exist on positive externalities for IRS coverage, but to the extent that the mechanisms for the reduction in infection are similar (reduction in the natural reservoir of the disease, reduction in the number of vectors, etc.), it is reasonable to assume similar effects.
+Additionally, we using weighted distance for our score calculation (rather than simply defining a radius threshold) because of previous studies' findings that there is a linear decline in protection conferred to others with greater distance from an intervention [@Binka2007]
+
+### Econometric model
+
 Our model specification is as follows
@@
 $$
-\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * (\hat{\beta}_2{IRS_{it}}*\hat{\beta}_3{IRStime_{it}}) + \alpha_i + \delta_t + \upsilon_{it}
+\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * (\hat{\beta}_2{IRS_{it}}*\hat{\beta}_3{IRStime_{it}}) +  \hat{\beta}_4{Herd_{it}} +  \alpha_i + \delta_t + \upsilon_{it}
 $$
@@
 Our intervention is whether the residence of the worker in question was treated in the last year, and, if so, the time since treatment, represented, respectively, by $\beta_{2}$ and $\beta_{3}$.
-$\alpha_i$ reprsents the time invariant worker fixed effects, and $\delta_i$ represents the fixed effect of the particular malaria season.
+$\beta_{4}$ **represents the distance-weighted herd protection score**.
+$\alpha_i$ represents the time invariant worker fixed effects, and $\delta_i$ represents the fixed effect of the particular malaria season.
 $\upsilon$ is the error term.
@@
 In addition, we adjust for daily precipitation; though its lagged effect on malaria incidence is likely captured by the seasonality term, we include rainfall since it's immediate effect (through its impact on transportation and working conditions) may also affect absenteeism.
-
-Details will go here about externalities (neighborhood effects, positive spillover to district, negative spillover from district, how large firm may mean that most positive externalities are absorbed by the firm, and not many negative externalities, but maybe with a smaller firm it wouldn't work as well)\todo{NOT DONE YET.}
-
-## Estimating return on investment
 

### 2018-05-06 inst/rmd/03_results.Rmd
externalities text

@@
 Table 1 shows overall worker details divided by "Treatment" versus "Control" status.
-"Treatment" is considered the time beginning at IRS administration and ending one year later; "Control" is considered time observed either prior to IRS administration (and greater than 1 year after previous IRS administration) or among workers who never received IRS.
+"Treatment" is considered the time beginning at IRS administration and ending six months later; "Control" is considered time observed either prior to IRS administration (and greater than 6 months after previous IRS administration) or among workers who never received IRS.
 Because of the temporal nature of these categories, many workers belong to both treatment and control groups, albeit at different times.
@@
 x <- model_data
-x$age <- round(as.numeric(as.Date('2014-01-01') - x$date_of_birth) / 365.25, digits = -1)
+x$age <-as.numeric(as.Date('2014-01-01') - x$date_of_birth) / 365.25
+x$agey <- x$age
+x$age <- round(x$age, digits = -1)
 x$status <- x$permanent_or_temporary
@@
     group_by_(var) %>%
-    summarise(`Treatment workers` = length(unique(oracle_number[!days_since %in% c('Never', 'Before')])),
-              `Treatment days` = length(oracle_number[!days_since %in% c('Never', 'Before')]),
-              `Control workers` = length(unique(oracle_number[days_since %in% c('Never', 'Before')])),
-              `Control days` = length(oracle_number[days_since %in% c('Never', 'Before')])) %>%
+    summarise(`Treatment workers` = length(unique(oracle_number[!months_since %in% c('Never', 'Before')])),
+              `Treatment days` = length(oracle_number[!months_since %in% c('Never', 'Before')]),
+              `Control workers` = length(unique(oracle_number[months_since %in% c('Never', 'Before')])),
+              `Control days` = length(oracle_number[months_since %in% c('Never', 'Before')])) %>%
     ungroup() %>%
@@
 ```{r, fig.width = 5, fig.height = 3.5, fig.align='center', fig.cap = 'Age distribution by sex and worker status'}
-x$age <- round(as.numeric(as.Date('2014-01-01') - x$date_of_birth) / 365.25, digits = 2)
+# x$age <- round(as.numeric(as.Date('2014-01-01') - x$date_of_birth) / 365.25, digits = 2)
 g1 <- ggplot(data = x,
-       aes(x = age,
+       aes(x = jitter(agey, factor = 1000),
            group = sex,
            fill = sex)) +
-  geom_density(alpha = 0.8, n = 2^4) +
+  ggridges::geom_density_ridges(aes(y = sex,
+                                    fill = sex),
+                                height = 0.2) +
   facet_wrap(~permanent_or_temporary) +
@@
   theme_maragra() +
-  geom_bar(data = bi_data_simple,
+  geom_point(data = bi_data_simple,
              aes(x = any_rain,
                  y = absenteeism),
-             stat = 'identity',
-             fill = 'grey',
-             alpha = 0.8,
-           width = 0.2) +
-  geom_label(data = bi_data_simple,
+             alpha = 0.8) +
+  geom_text(data = bi_data_simple,
              aes(x = any_rain,
                  y = absenteeism,
-                 label = round(absenteeism, digits = 2)),
-             color = 'red') +
+                 label = paste0(round(absenteeism, digits = 2), '\n(median)')),
+             color = 'black') +
   labs(x = '',
@@
 
-**Effect of IRS on clinical malaria**: The number of malaria cases registered at the company clinic during that time was 1873, for an approximate annualized clinical incidence of 110 cases per 1,000.
-\todo{Not done yet.
-In fact, not even sure if I should do this section at all, since it's basically just a distraction.}
-
 **Costs of program**: The malaria control program at Maragra has an annual operating budget of approximately $112,000, which includes the purchase of insecticide, the wages of IRS sprayers and drivers, transportation, record-keeping, and general administrative costs.
@@
 Our model took into account workers' residential location (on or off site) - even though off site workers by definition could not be administered IRS by Maragra - so as to capture other seasonal effects from those workers.
-\todo{This is not very eloquent yet and needs to be improved.}
+
+The herd protection score, unit-less by nature given that it's a summation of binary protection statuses weighted by inverse-transformed distances, took the following distribution in our data:
+
+```{r, fig.align = 'center', fig.cap = 'Distribution of herd protection score in panel data'}
+ggplot(data = model_data,
+       aes(x = herd)) +
+  geom_density(fill = 'lightblue',
+               alpha = 0.6) +
+  theme_maragra() +
+  labs(x = 'Herd protection score',
+       y = 'Density')
+```
 
 The model results for all cause absenteeism (table 4) suggest a significant decrease in absenteeism following the administration of IRS.
-The effect is ambivalent among temporary workers, whereas among permanent workers the administration of IRS decreases absenteeism betwen 2.5 and 8.4 percentage points in the 2-9 month following administration.
+The effect is ambivalent among temporary workers, whereas among permanent workers the administration of IRS decreases absenteeism by xxx.
 These effects are consistent with IRS decreasing absence through the medium of malaria, since (a) the effect is not as strong in the first month (potentially during the parasite incubation phase), the effect is stronger on field workers (who are more socioeconomically the demographic at greater risk of malaria than their administrative counterparts), and the effect is ambivalent among temporary workers (since the protective effect of IRS is only a function of the amount of time that the worker spends in the fumigated house).
-
-```{r, results = 'asis'}
-make_models_table(model_list = fe_models, the_caption = "All cause absenteeism: model results")
-```
-
-The model results for sickness only absenteeism (table 5) suggest that IRS has a significant effect on permanent field workers.
-The effect is directionally similar among temporary fieldworkers (but does not reach the level of significance).
-Among non-fieldworkers, the effect of IRS is mixed.\todo{I'll probably remove this from the final version, since we're saying that the data aren't that good...}
-
-```{r, results = 'asis'}
-make_models_table(model_list = sick_models, the_caption = "Sick absenteeism only: model results")
-```
-
-## Externalities
-
-Below are model results for the full protection model.
 
@@
 
-#### Neighborhood protective effect
-
-\todo{Working on this now}
-
-#### Core-periphery difference
-
-\todo{Working on this now}
+```{r}
+# Make dummy data for seeing effect of herd on irs
+```
 

### 2018-05-06 inst/rmd/externalities.Rmd
externalities text

@@
 bibliography: bibliography.bib
-csl: journal-of-health-economics.csl
 ---
@@
 ```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
+# csl: journal-of-health-economics.csl
+
 # Packages
@@
 To test whether IRS from one household has an effect on workers living in another household, we first must make some assumptions about the level and waning effect of IRS generally.
-For our purposes, we make a rather simple assumption, based on the literature: that IRS has an immediate effect on mosquitoes and should affect malaria risk following the incubation period, that the waning effect begins in month 4 [@Tukei2017], that IRS effects those not living in the household by reducing reproduction of mosquitoes [@White2011], and that IRS' effect after 1 year is essentially null [@Bukirwa2009].
-We simplify these assumptions into a time-protection matrix, which can be visualized in the below chart.
+It is known from the literature that IRS has an immediate effect on mosquitoes and should affect malaria risk following the incubation period, that the waning effect begins in month 4 [@Tukei2017], that IRS effects those not living in the household by reducing reproduction of mosquitoes [@White2011], and that IRS' effect after 1 year is essentially null [@Bukirwa2009].
 
-```{r}
+```{r, eval = FALSE}
 x <- c(1:6)
@@
 
+# The above chart shows two lines: the orange line is based on Tukei's emperical results regarding the protection afforded by IRS against malaria infection (among children) as a function of time since IRS; the blue line is our simplification of Tukei's work.
+
 ```
 
-The above chart shows two lines: the orange line is based on Tukei's emperical results regarding the protection afforded by IRS against malaria infection (among children) as a function of time since IRS; the blue line is our simplification of Tukei's work.
+For our purposes, we make simplify even further.
+We make no assumptions about the level of positive externalities afforded by IRS (instead preferring to estimate this within our model), and we do not account for waning, instead modeling the entire "post" period of 6 months as one.\footnote{Though not as granular as estimating time-specific protection within the 6 months following IRS administration, our approach is not less accurate; it simply aggregates the entirety of the 6 months after IRS administration into one value, and assumes (as is suggested by the literature) that the period after 6 months is effectively equivalent in chemical terms as the period prior to IRS.}
 
@@
 
-Having estimated each household's protection level at all times (ie, the "amount" of protection that they pass on to neighbors), we need to define a method for how much of that total level is passed on.
+At any given time, each household is either (a) protected (ie, < 6 months following IRS administration) or (b) unprotected.
+If the former, they confer protection to their neighbors; if the latter, they confer no protection to their neighbors.
+Having assigned each household's protection status at all times, we need to define a method for how much of any given household's protection level is passed on.
 A household which is 1 meter from a neighbor should, theoretically, afford more protection to that household than another which is 1000 meters.
-To account for the effect of distance, we define a simple function for weighting a nearby residence's contribution to a household's "herd protection" score as follows:
+To account for the effect of distance, we define a simple function for weighting a nearby residence's contribution to a household's "herd protection" score as 1 divided by the distance, as follows:
 
@@
 
-We consider a house's "herd" protection level (ie, the protection conferred to the house through externality) to be the average of the other houses' non-herd protection levels, weighted by the distance to the house in question.
+We consider a house's "herd" protection level (ie, the protection conferred to the house through externality) to be the average of the other houses' non-herd protection levels (ie, 0 if not protected and 1 if protected), weighted by the distance to the house in question.
 The below is an illustration of a house (the red x) and the neighboring houses (circles).
@@
                  y = rnorm(n = n),
-                 protection = round(sample(seq(0, 0.2, length = 100), n), digits = 1),
+                 protection = round(sample(0:1, n, replace = TRUE), digits = 1),
                  color = c('red', rep('grey', n-1))) %>%
@@
 make_plot <- function(df){
-  ggplot(data = df@data,
+  g <- 
+  ggplot(data = df@data %>% mutate(protection = factor(protection)) %>%
+           mutate(Weight = weight),
        aes(x = lng,
            y = lat)) +
-  geom_point(aes(size = weight,
-                 color = protection)) +
+  geom_point(aes(size = Weight,
+                 color = protection),
+             alpha = 0.6) +
   geom_point(data = df@data[1,],
-             color = 'red',
+             color = 'blue',
              pch = 4,
@@
        y = '') +
-  theme(legend.position = 'right') +
-  guides(color = guide_colourbar(title = 'Protection', title.position = 'top'),
-         size = guide_legend(nrow = 5, title.position = 'top', nbin = 2)) +
-  scale_color_continuous(low = 'orange', high = 'blue') +
+  # theme(legend.position = 'right') +
+    scale_color_manual(name = 'Protection',
+                       values = c('red', 'darkgreen')) +
+  # guides(color = guide_colourbar(title = 'Protection', title.position = 'top'),
+  #        size = guide_legend(nrow = 5, title.position = 'top', nbin = 2)) +
+  # scale_color_continuous(low = 'orange', high = 'blue') +
     xlim(-2,2) +
     ylim(-2,2)
+  for(i in seq(10, 150, length = 10)){
+    g <- g + 
+      geom_point(data = df@data[1,],
+             color = 'black',
+             pch = 1,
+             cex = i,
+             alpha = 0.4) 
+  }
+  for(i in 2:nrow(df@data)){
+    x <- bind_rows(df@data[i,],
+                   df@data[1,])
+    g <- g + geom_line(data = x,
+                        aes(x = lng,
+                            y = lat),
+                        alpha = 0.6,
+                        color = 'blue',
+                       lty = 2,
+                       size = 0.3)
+  }
+  return(g)
 }
@@
 Our approach differs in that the time dimension for IRS is much more important than ITN (ie, IRS is not a binary but rather a waning function, as described previously), requiring us to create herd scores for every day.
+This is because the protection conferred to a house by its neighbors may change from one day to the next, when a neighbor either (a) gets sprayed or (b) leaves "protected" status (ie, due to the end of the 6 months window).
 
@@
 make_models_table(model_list = protection_models, the_caption = "All absenteeism with herd immunity: model results")
-```
-
-### HIV
-
-We incorporated HIV prevalence of the area lived in, but it had not significance and was then removed from the model.
-
-```{r}
-plot(hiv_prevalence)
 ```
@@
 
-The below uses kernel density interpolation (via krieging) to estimate a generalized protection surface over the entire facility.
+The below uses kernel density interpolation to estimate a generalized protection surface over the entire facility.
 
@@
 
+## Conclusion
+
+Using the approach laid out here, there appears to be a significant protective effect afforded by the IRS of neighbors. If this approach is deemed correct, then next steps are determining:
+
+- How to translate the effect into costs and savings (simulation based approach at various IRS coverage levels?).
+- How / where to report on this in the paper (is this a "method" or "result"?).
+
+\newpage
+
+## Appendix
+
 ### Theoretical protection surface
@@
 data <- data.frame(id = 1:n,
-                   x,y, lng, lat,protection = sample(protection$protection[1:5], n,
+                   x,y, lng, lat,protection = sample(0:1, n,
                                                      replace = TRUE))
@@
 
-## Conclusion
-
-Using the approach laid out here, there appears to be a significant protective effect afforded by the IRS of neighbors. If this approach is deemed correct, then next steps are determining:
-
-- How to translate the effect into costs and savings (simulation based approach at various IRS coverage levels?).
-- How to combine this effect with the other potential spillover effect (ie, core-periphery issues).
-- How / where to report on this in the paper (is this a "method" or "result"?).
-
 # References

### 2018-05-06 inst/rmd/externalities.tex
externalities text

@@
 in another household, we first must make some assumptions about the
-level and waning effect of IRS generally. For our purposes, we make a
-rather simple assumption, based on the literature: that IRS has an
-immediate effect on mosquitoes and should affect malaria risk following
-the incubation period, that the waning effect begins in month 4 (Tukei
-et al., 2017), that IRS effects those not living in the household by
-reducing reproduction of mosquitoes (White et al., 2011), and that IRS'
-effect after 1 year is essentially null (Bukirwa et al., 2009). We
-simplify these assumptions into a time-protection matrix, which can be
-visualized in the below chart.
+level and waning effect of IRS generally. It is known from the
+literature that IRS has an immediate effect on mosquitoes and should
+affect malaria risk following the incubation period, that the waning
+effect begins in month 4 (Tukei, Beke, and Lamadrid-Figueroa 2017), that
+IRS effects those not living in the household by reducing reproduction
+of mosquitoes (White et al. 2011), and that IRS' effect after 1 year is
+essentially null (Bukirwa et al. 2009).
 
-\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-3-1} \end{center}
-
-The above chart shows two lines: the orange line is based on Tukei's
-emperical results regarding the protection afforded by IRS against
-malaria infection (among children) as a function of time since IRS; the
-blue line is our simplification of Tukei's work.
+For our purposes, we make simplify even further. We make no assumptions
+about the level of positive externalities afforded by IRS (instead
+preferring to estimate this within our model), and we do not account for
+waning, instead modeling the entire ``post'' period of 6 months as
+one.\footnote{Though not as granular as estimating time-specific protection within the 6 months following IRS administration, our approach is not less accurate; it simply aggregates the entirety of the 6 months after IRS administration into one value, and assumes (as is suggested by the literature) that the period after 6 months is effectively equivalent in chemical terms as the period prior to IRS.}
 
@@
 
-Having estimated each household's protection level at all times (ie, the
-``amount'' of protection that they pass on to neighbors), we need to
-define a method for how much of that total level is passed on. A
+At any given time, each household is either (a) protected (ie,
+\textless{} 6 months following IRS administration) or (b) unprotected.
+If the former, they confer protection to their neighbors; if the latter,
+they confer no protection to their neighbors. Having assigned each
+household's protection status at all times, we need to define a method
+for how much of any given household's protection level is passed on. A
 household which is 1 meter from a neighbor should, theoretically, afford
@@
 weighting a nearby residence's contribution to a household's ``herd
-protection'' score as follows:
+protection'' score as 1 divided by the distance, as follows:
 
@@
 small travel distances that most mosquitoes fly in normal conditions
-(Verdonschot and Besse-Lototskaya, 2014).
+(Verdonschot and Besse-Lototskaya 2014).
 
@@
 conferred to the house through externality) to be the average of the
-other houses' non-herd protection levels, weighted by the distance to
-the house in question. The below is an illustration of a house (the red
-x) and the neighboring houses (circles). The ``weight'' of each house is
-indicated by the circle size, and the protection level of the house is
-indicated by the shading of the circle.
+other houses' non-herd protection levels (ie, 0 if not protected and 1
+if protected), weighted by the distance to the house in question. The
+below is an illustration of a house (the red x) and the neighboring
+houses (circles). The ``weight'' of each house is indicated by the
+circle size, and the protection level of the house is indicated by the
+shading of the circle.
 
@@
 This score is conceptually similar to Cohen and Dupas' quantification of
-the positive externalities of ITN (bednet) use in Kenya (Cohen and
-Dupas, 2010) in that it attempts to estimate the protection conferred to
+the positive externalities of ITN (bednet) use in Kenya (Cohen and Dupas
+2010) in that it attempts to estimate the protection conferred to
 ``non-users'' by users and uses a ``weighted sum'' (as opposed to
@@
 waning function, as described previously), requiring us to create herd
-scores for every day.
+scores for every day. This is because the protection conferred to a
+house by its neighbors may change from one day to the next, when a
+neighbor either (a) gets sprayed or (b) leaves ``protected'' status (ie,
+due to the end of the 6 months window).
 
@@
 positive health externalities in malaria interventions related to ITN
-coverage (Alaii et al., 2003). No studies exist on positive
-externalities for IRS coverage, but to the extent that the mechanisms
-for the reduction in infection are similar (reduction in the natural
-reservoir of the disease, reduction in the number of vectors, etc.), it
-is reasonable to assume similar effects. Additionally, we using weighted
+coverage (Alaii et al. 2003). No studies exist on positive externalities
+for IRS coverage, but to the extent that the mechanisms for the
+reduction in infection are similar (reduction in the natural reservoir
+of the disease, reduction in the number of vectors, etc.), it is
+reasonable to assume similar effects. Additionally, we using weighted
 distance for our score calculation (rather than simply defining a radius
@@
 decline in protection conferred to others with greater distance from an
-intervention (Binka et al., 2007)
+intervention (Binka et al. 2007)
 
@@
 We calculate a ``herd protection'' score for each location-date
-combination (1759100 values). Whereas our original model specification
+combination (395019 values). Whereas our original model specification
 looked like:
@@
 \multicolumn{2}{l}{\textbf{Permanent field worker}}\\
-\hspace{1em}Malaria season & 3.597 (P<0.001)\\
-\hspace{1em}Months since IRS: IRS & 6.149 (P<0.001)\\
-\hspace{1em}Rainy day & 3.246 (P<0.001)\\
-\hspace{1em}Herd protection & -0.01 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: IRS & -8.548 (P<0.001)\\
+\hspace{1em}Malaria season & 2.809 (P<0.001)\\
+\hspace{1em}IRS status=After & 1.835 (P<0.001)\\
+\hspace{1em}Rainy day & 3.765 (P<0.001)\\
+\hspace{1em}Herd protection & -0.003 (P=0.368)\\
+\hspace{1em}Malaria season:IRS status=After & -5.106 (P<0.001)\\
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
-\hspace{1em}Malaria season & 0.792 (P<0.001)\\
-\hspace{1em}Months since IRS: IRS & 2.243 (P<0.001)\\
-\hspace{1em}Rainy day & 4.552 (P<0.001)\\
-\hspace{1em}Herd protection & -0.009 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: IRS & -1.508 (P=0.002)\\
+\hspace{1em}Malaria season & 0.831 (P<0.001)\\
+\hspace{1em}IRS status=After & 1.151 (P<0.001)\\
+\hspace{1em}Rainy day & 3.835 (P<0.001)\\
+\hspace{1em}Herd protection & 0 (P=0.927)\\
+\hspace{1em}Malaria season:IRS status=After & -1.66 (P<0.001)\\
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Temporary field worker}}\\
-\hspace{1em}Malaria season & -0.362 (P<0.001)\\
-\hspace{1em}Months since IRS: IRS & -0.229 (P=0.004)\\
-\hspace{1em}Rainy day & 0.037 (P=0.261)\\
-\hspace{1em}Herd protection & 0.002 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: IRS & 0.26 (P=0.014)\\
+\hspace{1em}Malaria season & -0.188 (P=0.001)\\
+\hspace{1em}IRS status=After & -0.254 (P<0.001)\\
+\hspace{1em}Rainy day & 0 (P=0.994)\\
+\hspace{1em}Herd protection & -0.001 (P=0.006)\\
+\hspace{1em}Malaria season:IRS status=After & 0.054 (P=0.568)\\
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
-\hspace{1em}Malaria season & 3.151 (P<0.001)\\
-\hspace{1em}Months since IRS: IRS & -1.718 (P=0.076)\\
-\hspace{1em}Rainy day & 0.224 (P=0.469)\\
-\hspace{1em}Herd protection & 0.002 (P=0.772)\\
-\hspace{1em}Malaria season:Months since IRS: IRS & 2.657 (P=0.02)\\*
+\hspace{1em}Malaria season & 5.772 (P<0.001)\\
+\hspace{1em}IRS status=After & -2.187 (P=0.023)\\
+\hspace{1em}Rainy day & -1.349 (P=0.003)\\
+\hspace{1em}Herd protection & 0.043 (P<0.001)\\
+\hspace{1em}Malaria season:IRS status=After & -3.235 (P=0.003)\\*
 \end{longtable}
-
-\subsubsection{HIV}\label{hiv}
-
-We incorporated HIV prevalence of the area lived in, but it had not
-significance and was then removed from the model.
-
-\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-9-1} \end{center}
 
@@
 
+\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-9-1} \end{center}
+
+The below uses kernel density interpolation to estimate a generalized
+protection surface over the entire facility.
+
 \begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-10-1} \end{center}
 
-The below uses kernel density interpolation (via krieging) to estimate a
-generalized protection surface over the entire facility.
-
 \begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-11-1} \end{center}
-
-\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-12-1} \end{center}
-
-\subsubsection{Theoretical protection
-surface}\label{theoretical-protection-surface}
-
-The below shows a theoretical protection surface assuming (a) uniform
-distribution of residences, (b) random distribution of IRS over time,
-but never going beyond 6 months between sprays.
-
-\paragraph{100 workers}\label{workers}
-
-\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-14-1} \end{center}
-
-\paragraph{1000 workers}\label{workers-1}
-
-\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-15-1} \end{center}
-
-\paragraph{3000 workers}\label{workers-2}
-
-\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-16-1} \end{center}
 
@@
 \item
-  How to combine this effect with the other potential spillover effect
-  (ie, core-periphery issues).
-\item
   How / where to report on this in the paper (is this a ``method'' or
@@
 \end{itemize}
+
+\newpage
+
+\subsection{Appendix}\label{appendix}
+
+\subsubsection{Theoretical protection
+surface}\label{theoretical-protection-surface}
+
+The below shows a theoretical protection surface assuming (a) uniform
+distribution of residences, (b) random distribution of IRS over time,
+but never going beyond 6 months between sprays.
+
+\paragraph{100 workers}\label{workers}
+
+\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-13-1} \end{center}
+
+\paragraph{1000 workers}\label{workers-1}
+
+\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-14-1} \end{center}
+
+\paragraph{3000 workers}\label{workers-2}
+
+\begin{center}\includegraphics{externalities_files/figure-latex/unnamed-chunk-15-1} \end{center}
 
@@
 \hypertarget{ref-Alaii2003FactorsAU}{}
-Alaii, J.A., Hawley, W.A., Kolczak, M.S., Kuile, F.O.T., Gimnig, J.E.,
-Vulule, J., Odhacha, A., Oloo, A.J., Nahlen, B.L., Phillips-Howard,
-P.A., 2003. Factors affecting use of permethrin-treated bed nets during
-a randomized controlled trial in western kenya. The American journal of
-tropical medicine and hygiene 68 4 Suppl, 137--41.
+Alaii, Jane A., William A. Hawley, Margarette S. Kolczak, Feiko O Ter
+Kuile, John E. Gimnig, John Vulule, Amos Odhacha, Aggrey J Oloo, Bernard
+L. Nahlen, and Penelope A Phillips-Howard. 2003. ``Factors Affecting Use
+of Permethrin-Treated Bed Nets During a Randomized Controlled Trial in
+Western Kenya.'' \emph{The American Journal of Tropical Medicine and
+Hygiene} 68 4 Suppl: 137--41.
 
 \hypertarget{ref-Binka2007}{}
-Binka, F.N., Kubaje, A., Adjuik, M., Williams, L.A., Lengeler, C.,
-Maude, G.H., Armah, G.E., Kajihara, B., Adiamah, J.H., Smith, P.G.,
-2007. Impact of permethrin impregnated bednets on child mortality in
-kassena-nankana district, ghana: A randomized controlled trial. Tropical
-Medicine \& International Health 1, 147--154.
-\url{https://doi.org/10.1111/j.1365-3156.1996.tb00020.x}
+Binka, F. N., A. Kubaje, M. Adjuik, L. A. Williams, C. Lengeler, G. H.
+Maude, G. E. Armah, B. Kajihara, J. H. Adiamah, and P. G. Smith. 2007.
+``Impact of Permethrin Impregnated Bednets on Child Mortality in
+Kassena-Nankana District, Ghana: A Randomized Controlled Trial.''
+\emph{Tropical Medicine \& International Health} 1 (2). Wiley-Blackwell:
+147--54.
+doi:\href{https://doi.org/10.1111/j.1365-3156.1996.tb00020.x}{10.1111/j.1365-3156.1996.tb00020.x}.
 
 \hypertarget{ref-Bukirwa2009}{}
-Bukirwa, H., Yau, V., Kigozi, R., Filler, S., Quick, L., Lugemwa, M.,
-Dissanayake, G., Kamya, M., Wabwire-Mangen, F., Dorsey, G., 2009.
-Assessing the impact of indoor residual spraying on malaria morbidity
-using a sentinel site surveillance system in western uganda. American
-Journal of Tropical Medicine and Hygiene 81, 611--614.
-\url{https://doi.org/10.4269/ajtmh.2009.09-0126}
+Bukirwa, H., V. Yau, R. Kigozi, S. Filler, L. Quick, M. Lugemwa, G.
+Dissanayake, M. Kamya, F. Wabwire-Mangen, and G. Dorsey. 2009.
+``Assessing the Impact of Indoor Residual Spraying on Malaria Morbidity
+Using a Sentinel Site Surveillance System in Western Uganda.''
+\emph{American Journal of Tropical Medicine and Hygiene} 81 (4).
+American Society of Tropical Medicine; Hygiene: 611--14.
+doi:\href{https://doi.org/10.4269/ajtmh.2009.09-0126}{10.4269/ajtmh.2009.09-0126}.
 
 \hypertarget{ref-Cohen2010}{}
-Cohen, J., Dupas, P., 2010. Free distribution or cost-sharing? Evidence
-from a randomized malaria prevention experiment. Quarterly Journal of
-Economics 125, 1--45. \url{https://doi.org/10.1162/qjec.2010.125.1.1}
+Cohen, Jessica, and Pascaline Dupas. 2010. ``Free Distribution or
+Cost-Sharing? Evidence from a Randomized Malaria Prevention
+Experiment.'' \emph{Quarterly Journal of Economics} 125 (1). Oxford
+University Press (OUP): 1--45.
+doi:\href{https://doi.org/10.1162/qjec.2010.125.1.1}{10.1162/qjec.2010.125.1.1}.
 
 \hypertarget{ref-Tukei2017}{}
-Tukei, B.B., Beke, A., Lamadrid-Figueroa, H., 2017. Assessing the effect
-of indoor residual spraying (IRS) on malaria morbidity in northern
-uganda: A before and after study. Malaria Journal 16.
-\url{https://doi.org/10.1186/s12936-016-1652-4}
+Tukei, Betty Bawuba, Andy Beke, and Héctor Lamadrid-Figueroa. 2017.
+``Assessing the Effect of Indoor Residual Spraying (IRS) on Malaria
+Morbidity in Northern Uganda: A Before and After Study.'' \emph{Malaria
+Journal} 16 (1). Springer Nature.
+doi:\href{https://doi.org/10.1186/s12936-016-1652-4}{10.1186/s12936-016-1652-4}.
 
 \hypertarget{ref-Verdonschot2014}{}
-Verdonschot, P.F., Besse-Lototskaya, A.A., 2014. Flight distance of
-mosquitoes (culicidae): A metadata analysis to support the management of
-barrier zones around rewetted and newly constructed wetlands.
-Limnologica - Ecology and Management of Inland Waters 45, 69--79.
-\url{https://doi.org/10.1016/j.limno.2013.11.002}
+Verdonschot, Piet F.M., and Anna A. Besse-Lototskaya. 2014. ``Flight
+Distance of Mosquitoes (Culicidae): A Metadata Analysis to Support the
+Management of Barrier Zones Around Rewetted and Newly Constructed
+Wetlands.'' \emph{Limnologica - Ecology and Management of Inland Waters}
+45 (March). Elsevier BV: 69--79.
+doi:\href{https://doi.org/10.1016/j.limno.2013.11.002}{10.1016/j.limno.2013.11.002}.
 
 \hypertarget{ref-White2011}{}
-White, M.T., Griffin, J.T., Churcher, T.S., Ferguson, N.M., Basáñez,
-M.-G., Ghani, A.C., 2011. Modelling the impact of vector control
-interventions on anopheles gambiae population dynamics. Parasites \&
-Vectors 4, 153. \url{https://doi.org/10.1186/1756-3305-4-153}
+White, Michael T, Jamie T Griffin, Thomas S Churcher, Neil M Ferguson,
+María-Gloria Basáñez, and Azra C Ghani. 2011. ``Modelling the Impact of
+Vector Control Interventions on Anopheles Gambiae Population Dynamics.''
+\emph{Parasites \& Vectors} 4 (1). Springer Nature: 153.
+doi:\href{https://doi.org/10.1186/1756-3305-4-153}{10.1186/1756-3305-4-153}.
 

### 2018-05-06 inst/rmd/paper.tex
externalities text

--- 2018-04-20 inst/rmd/paper.tex
@@
 al., 2011) (Liu et al., 2010) (with individual worker fixed effects, and
-random effects for other covariates) in lieu of interrupted time series
-(Lopez Bernal et al., 2016) so as to explicitly adjust for confounders
-(Bell and Jones, 2014). We divide into 4 different models for different
-worker types, so as to account for the potential time-varying effects of
-worker type on risk of malaria. Our model specification is as follows
+random effects for other covariates) so as to explicitly adjust for
+confounders (Bell and Jones, 2014). We divide into 4 different models
+for different worker types, so as to account for the potential
+time-varying effects of worker type on risk of malaria.
+
+\subsubsection{Externalities}\label{externalities}
+
+The administration of insecticide at a worker's residence (indoor
+residual spraying or IRS) likely protects that worker from malaria
+infection by killing the vectors (mosquitoes) that land on the
+building's walls. However, it is also highly likely that the protective
+effective of IRS ``spills over'' to others who live nearby. This
+positive spillover effect would theoretically go through two channels:
+(i) via a reduction of mosquitoes in the vicinity and (ii) via a
+reduction of the malaria parasite in the blood of humans in the vicinity
+(ie, the parasite ``reservoir''). This document describes our method for
+assessing the existence and magnitude of positive spillover of IRS in
+the Maragra workers' data. We devise a time-specific household
+``protection'' score based on the theoretical effectiveness of IRS, and
+then use that protection score to develop a time-place specific ``herd
+protection'' score based on a weighted average of nearby household
+protection scores.
+
+We consider a house's ``herd'' protection level (ie, the protection
+conferred to the house through externality) to be the average of the
+other houses' non-herd protection levels, weighted by the distance to
+the house in question. The below is an illustration of a house (the red
+x) and the neighboring houses (circles). The ``weight'' of each house is
+indicated by the circle size, and the protection level of the house is
+indicated by the shading of the circle.
+
+However, this consideration misses one important dimension: density. The
+below house's weighted average protection score is identical to that of
+the above house. In other words, at the time in question, at both
+locations, the percentage of houses with IRS coverage is the same, and
+the (weighted) average time since IRS is the same. However, the below
+house is likely much more protected than the above, given the
+\emph{absolute} number of nearby IRS-protected houses.
+
+Our approach is this: rather than using a weighted average of each
+household's individual protection score, we instead use a weighted sum.
+A particular household's herd protection score at any given time is the
+sum of all households' individual protection scores at that time
+multiplied by those households' distance weights.
+
+The summation of (a) the IRS level of other houses at a certain time,
+(b) weighted by the distance of those houses to the house in question,
+(c) limited to only the 1 km radius multiplied yields a ``herd
+protection score''.
+
+This score is conceptually similar to Cohen and Dupas' quantification of
+the positive externalities of ITN (bednet) use in Kenya (Cohen and
+Dupas, 2010) in that it attempts to estimate the protection conferred to
+``non-users'' by users and uses a ``weighted sum'' (as opposed to
+average). Our approach differs in that the time dimension for IRS is
+much more important than ITN (ie, IRS is not a binary but rather a
+waning function, as described previously), requiring us to create herd
+scores for every day.
+
+Our approach can be justified in that previous studies have found strong
+positive health externalities in malaria interventions related to ITN
+coverage (Alaii et al., 2003). No studies exist on positive
+externalities for IRS coverage, but to the extent that the mechanisms
+for the reduction in infection are similar (reduction in the natural
+reservoir of the disease, reduction in the number of vectors, etc.), it
+is reasonable to assume similar effects. Additionally, we using weighted
+distance for our score calculation (rather than simply defining a radius
+threshold) because of previous studies' findings that there is a linear
+decline in protection conferred to others with greater distance from an
+intervention (Binka et al., 2007)
+
+\subsubsection{Econometric model}\label{econometric-model}
+
+Our model specification is as follows
 
 \[
-\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * (\hat{\beta}_2{IRS_{it}}*\hat{\beta}_3{IRStime_{it}}) + \alpha_i + \delta_t + \upsilon_{it}
+\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * (\hat{\beta}_2{IRS_{it}}*\hat{\beta}_3{IRStime_{it}}) +  \hat{\beta}_4{Herd_{it}} +  \alpha_i + \delta_t + \upsilon_{it}
 \]
@@
 represented, respectively, by \(\beta_{2}\) and \(\beta_{3}\).
-\(\alpha_i\) reprsents the time invariant worker fixed effects, and
-\(\delta_i\) represents the fixed effect of the particular malaria
+\(\beta_{4}\) \textbf{represents the distance-weighted herd protection
+score}. \(\alpha_i\) represents the time invariant worker fixed effects,
+and \(\delta_i\) represents the fixed effect of the particular malaria
 season. \(\upsilon\) is the error term.
@@
 
-Details will go here about externalities (neighborhood effects, positive
-spillover to district, negative spillover from district, how large firm
-may mean that most positive externalities are absorbed by the firm, and
-not many negative externalities, but maybe with a smaller firm it
-wouldn't work as well)\todo{NOT DONE YET.}
-
 \subsection{Estimating return on
@@
   \hline
-Department & Administrative &  35 & 12523 & 157 & 173151 \\ 
-   & Factory &  71 & 20533 & 453 & 280416 \\ 
-   & Field & 410 & 91302 & 2691 & 1181175 \\ 
-  Sex & F & 228 & 47907 & 1274 & 489484 \\ 
-   & M & 288 & 76451 & 2027 & 1145258 \\ 
-  Status & Permanent & 127 & 47075 & 500 & 655311 \\ 
-   & Temporary & 389 & 77283 & 2801 & 979431 \\ 
-  Age & 20 & 140 & 29463 & 960 & 347758 \\ 
-   & 30 & 176 & 39119 & 1225 & 550783 \\ 
-   & 40 & 123 & 30186 & 652 & 366936 \\ 
-   & 50 &  58 & 18881 & 344 & 268877 \\ 
-   & 60 &  17 & 5893 & 116 & 96555 \\ 
-   & 70 &   2 & 816 &   4 & 3833 \\ 
+Department & Administrative &  38 & 47992 &   0 &   0 \\ 
+   & Factory & 109 & 78821 &   0 &   0 \\ 
+   & Field & 545 & 268206 &   0 &   0 \\ 
+  Sex & F & 310 & 139959 &   0 &   0 \\ 
+   & M & 382 & 255060 &   0 &   0 \\ 
+  Status & Permanent & 128 & 182199 &   0 &   0 \\ 
+   & Temporary & 564 & 212820 &   0 &   0 \\ 
+  Age & 20 & 200 & 81640 &   0 &   0 \\ 
+   & 30 & 242 & 124339 &   0 &   0 \\ 
+   & 40 & 154 & 101317 &   0 &   0 \\ 
+   & 50 &  68 & 63739 &   0 &   0 \\ 
+   & 60 &  26 & 22257 &   0 &   0 \\ 
+   & 70 &   2 & 1727 &   0 &   0 \\ 
    \hline
@@
 The total number of unique agregados sprayed during this period was
-3,998. Among the 3,362 workers for whom we have reliable absenteeism and
-residential data, 692 had their homes fumigated at least once (the
+3,998. Among the 692 workers for whom we have reliable absenteeism and
+residential data, 562 had their homes fumigated at least once (the
 majority of workers live off of the facility).
@@
 \textbf{Absences}: We observed 1,759,100 unique worker-days among the
-3,362 workers. The all-period average absenteeism rate was 5.56\%,
-though this rate varied widely as a function of worker department, sex,
+692 workers. The all-period average absenteeism rate was 5.56\%, though
+this rate varied widely as a function of worker department, sex,
 residence, and season (table 1).
@@
   \hline
-Malaria season & Low & 5.2\% & 7\% & 5.4\% & 3.3\% \\ 
-   & High & 8.2\% & 12.8\% & 6.3\% & 2.7\% \\ 
-  Worker type & Field worker & 4.4\% & 7.6\% & 3.9\% & 1.7\% \\ 
-   & Not field worker & 10.7\% & 12.4\% & 11.5\% & 9.6\% \\ 
-  Contract & Permanent & 12.1\% & 12.6\% & 12\% & 10.2\% \\ 
-   & Temporary & 0.1\% & 5\% & 2.3\% & 0.5\% \\ 
-  Sex & F & 4\% & 8.1\% & 4.4\% & 1.9\% \\ 
-   & M & 8.1\% & 10\% & 6.5\% & 3.7\% \\ 
-  Residence & Off site & 6.4\% & 9.6\% & 5.9\% & 3\% \\ 
-   & On site & 9.4\% & 9.7\% & 6.1\% & 3.1\% \\ 
-  Precipitation & Dry & 5.4\% & 7.7\% & 4.8\% & 2.4\% \\ 
-   & Rainy & 7.9\% & 10.6\% & 7\% & 3.3\% \\ 
+Malaria season & Low & 7.5\% & 6.6\% & 5.5\% & 3.4\% \\ 
+   & High & 10.7\% & 13.4\% & 6.4\% & 2.7\% \\ 
+  Worker type & Field worker & 7.1\% & 8.7\% & 3.9\% & 1.9\% \\ 
+   & Not field worker & 11.5\% & 10.8\% & 10.7\% & 8.2\% \\ 
+  Contract & Permanent & 12.8\% & 12.5\% & 11.2\% & 9.5\% \\ 
+   & Temporary & 0.1\% & 4\% & 2.1\% & 0.5\% \\ 
+  Sex & F & 7.1\% & 8.6\% & 4.4\% & 2.3\% \\ 
+   & M & 10.4\% & 10\% & 6.9\% & 3.8\% \\ 
+  Residence & On site & 9.4\% & 9.7\% & 6.1\% & 3.1\% \\ 
+  Precipitation & Dry & 7.4\% & 7.5\% & 4.9\% & 2.4\% \\ 
+   & Rainy & 10.5\% & 10.8\% & 7.1\% & 3.4\% \\ 
    \hline
@@
 The effect is ambivalent among temporary workers, whereas among
-permanent workers the administration of IRS decreases absenteeism betwen
-2.5 and 8.4 percentage points in the 2-9 month following administration.
-These effects are consistent with IRS decreasing absence through the
-medium of malaria, since (a) the effect is not as strong in the first
-month (potentially during the parasite incubation phase), the effect is
-stronger on field workers (who are more socioeconomically the
+permanent workers the administration of IRS decreases absenteeism by
+xxx. These effects are consistent with IRS decreasing absence through
+the medium of malaria, since (a) the effect is not as strong in the
+first month (potentially during the parasite incubation phase), the
+effect is stronger on field workers (who are more socioeconomically the
 demographic at greater risk of malaria than their administrative
@@
 
-\begin{table}
-
-\caption{\label{tab:unnamed-chunk-28}All cause absenteeism: model results}
-\centering
-\begin{tabular}[t]{ll}
+\begin{longtable}[t]{ll}
+\caption{\label{tab:unnamed-chunk-27}All absenteeism with herd immunity: model results}\\
 \toprule
@@
 \midrule
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
-\hspace{1em}Malaria season & 2.533 (P<0.001)\\
-\hspace{1em}Months since IRS: 01 & 5.382 (P<0.001)\\
-\hspace{1em}Months since IRS: 02-04 & 4.747 (P<0.001)\\
-\hspace{1em}Months since IRS: 05-09 & -2.669 (P<0.001)\\
-\hspace{1em}Precipitation (mm) & 0.206 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & -6.262 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & -8.388 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & -3.71 (P<0.001)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
-\hspace{1em}Malaria season & 0.957 (P<0.001)\\
-\hspace{1em}Months since IRS: 01 & -0.987 (P=0.071)\\
-\hspace{1em}Months since IRS: 02-04 & 4.375 (P<0.001)\\
-\hspace{1em}Months since IRS: 05-09 & -0.55 (P=0.254)\\
-\hspace{1em}Precipitation (mm) & 0.548 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & 1.787 (P=0.016)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & -5.208 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & -2.04 (P=0.003)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
-\hspace{1em}Malaria season & -0.368 (P<0.001)\\
-\hspace{1em}Months since IRS: 01 & -0.075 (P=0.526)\\
-\hspace{1em}Months since IRS: 02-04 & -0.193 (P=0.072)\\
-\hspace{1em}Months since IRS: 05-09 & -0.036 (P=0.738)\\
-\hspace{1em}Precipitation (mm) & 0.02 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & -0.154 (P=0.406)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & 0.297 (P=0.05)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & 0.041 (P=0.769)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
-\hspace{1em}Malaria season & 2.025 (P<0.001)\\
-\hspace{1em}Months since IRS: 01 & -2.536 (P=0.134)\\
-\hspace{1em}Months since IRS: 02-04 & -1.177 (P=0.282)\\
-\hspace{1em}Months since IRS: 05-09 & -1.821 (P=0.084)\\
-\hspace{1em}Precipitation (mm) & 0.078 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & 6.121 (P=0.002)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & 1.333 (P=0.343)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & -4.639 (P<0.001)\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-The model results for sickness only absenteeism (table 5) suggest that
-IRS has a significant effect on permanent field workers. The effect is
-directionally similar among temporary fieldworkers (but does not reach
-the level of significance). Among non-fieldworkers, the effect of IRS is
-mixed.
-
-\begin{table}
-
-\caption{\label{tab:unnamed-chunk-29}Sick absenteeism only: model results}
-\centering
-\begin{tabular}[t]{ll}
+\endfirsthead
+\caption[]{All absenteeism with herd immunity: model results \textit{(continued)}}\\
 \toprule
@@
 \midrule
+\endhead
+\
+\endfoot
+\bottomrule
+\endlastfoot
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Permanent field worker}}\\
-\hspace{1em}Malaria season & -0.073 (P=0.089)\\
-\hspace{1em}Months since IRS: 01 & 0.891 (P<0.001)\\
-\hspace{1em}Months since IRS: 02-04 & 1.223 (P<0.001)\\
-\hspace{1em}Months since IRS: 05-09 & -0.236 (P=0.249)\\
-\hspace{1em}Precipitation (mm) & 0 (P=0.984)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & -1.501 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & -0.597 (P=0.031)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & -0.484 (P=0.097)\\
+\hspace{1em}Malaria season & 2.809 (P<0.001)\\
+\hspace{1em}IRS status=After & 1.835 (P<0.001)\\
+\hspace{1em}Rainy day & 3.765 (P<0.001)\\
+\hspace{1em}Herd protection & -0.003 (P=0.368)\\
+\hspace{1em}Malaria season:IRS status=After & -5.106 (P<0.001)\\
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
-\hspace{1em}Malaria season & -0.056 (P=0.204)\\
-\hspace{1em}Months since IRS: 01 & -0.667 (P=0.003)\\
-\hspace{1em}Months since IRS: 02-04 & 0.904 (P<0.001)\\
-\hspace{1em}Months since IRS: 05-09 & 0.127 (P=0.526)\\
-\hspace{1em}Precipitation (mm) & -0.002 (P=0.565)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & 1.121 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & -0.122 (P=0.652)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & 1.198 (P<0.001)\\
+\hspace{1em}Malaria season & 0.831 (P<0.001)\\
+\hspace{1em}IRS status=After & 1.151 (P<0.001)\\
+\hspace{1em}Rainy day & 3.835 (P<0.001)\\
+\hspace{1em}Herd protection & 0 (P=0.927)\\
+\hspace{1em}Malaria season:IRS status=After & -1.66 (P<0.001)\\
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Temporary field worker}}\\
-\hspace{1em}Malaria season & -0.106 (P<0.001)\\
-\hspace{1em}Months since IRS: 01 & 0.029 (P=0.545)\\
-\hspace{1em}Months since IRS: 02-04 & 0.04 (P=0.35)\\
-\hspace{1em}Months since IRS: 05-09 & -0.046 (P=0.291)\\
-\hspace{1em}Precipitation (mm) & 0 (P=0.966)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & -0.108 (P=0.148)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & -0.032 (P=0.595)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & -0.052 (P=0.36)\\
+\hspace{1em}Malaria season & -0.188 (P=0.001)\\
+\hspace{1em}IRS status=After & -0.254 (P<0.001)\\
+\hspace{1em}Rainy day & 0 (P=0.994)\\
+\hspace{1em}Herd protection & -0.001 (P=0.006)\\
+\hspace{1em}Malaria season:IRS status=After & 0.054 (P=0.568)\\
 \addlinespace[1.5em]
 \multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
-\hspace{1em}Malaria season & -0.135 (P=0.236)\\
-\hspace{1em}Months since IRS: 01 & 1.427 (P=0.092)\\
-\hspace{1em}Months since IRS: 02-04 & -0.569 (P=0.299)\\
-\hspace{1em}Months since IRS: 05-09 & -1.42 (P=0.007)\\
-\hspace{1em}Precipitation (mm) & -0.005 (P=0.613)\\
-\hspace{1em}Malaria season:Months since IRS: 01 & -1.796 (P=0.069)\\
-\hspace{1em}Malaria season:Months since IRS: 02-04 & 2.479 (P<0.001)\\
-\hspace{1em}Malaria season:Months since IRS: 05-09 & 0.18 (P=0.794)\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\subsubsection{Externalities}\label{externalities}
-
-This section is in progress. Looking at neighborhood effects,
-core-periphery effects, etc.
+\hspace{1em}Malaria season & 5.772 (P<0.001)\\
+\hspace{1em}IRS status=After & -2.187 (P=0.023)\\
+\hspace{1em}Rainy day & -1.349 (P=0.003)\\
+\hspace{1em}Herd protection & 0.043 (P<0.001)\\
+\hspace{1em}Malaria season:IRS status=After & -3.235 (P=0.003)\\*
+\end{longtable}
 
@@
 \centering
-\includegraphics{figures/unnamed-chunk-33-1.pdf}
+\includegraphics{figures/unnamed-chunk-32-1.pdf}
 \caption{i. Absenteeism before and after IRS administration for all
@@
 \centering
-\includegraphics{figures/unnamed-chunk-34-1.pdf}
+\includegraphics{figures/unnamed-chunk-33-1.pdf}
 \caption{i. Sick absenteeism before and after IRS administration for all
@@
 
+\hypertarget{ref-Alaii2003FactorsAU}{}
+Alaii, J.A., Hawley, W.A., Kolczak, M.S., Kuile, F.O.T., Gimnig, J.E.,
+Vulule, J., Odhacha, A., Oloo, A.J., Nahlen, B.L., Phillips-Howard,
+P.A., 2003. Factors affecting use of permethrin-treated bed nets during
+a randomized controlled trial in western kenya. The American journal of
+tropical medicine and hygiene 68 4 Suppl, 137--41.
+
 \hypertarget{ref-AsensoOkyere1997}{}
@@
 
+\hypertarget{ref-Binka2007}{}
+Binka, F.N., Kubaje, A., Adjuik, M., Williams, L.A., Lengeler, C.,
+Maude, G.H., Armah, G.E., Kajihara, B., Adiamah, J.H., Smith, P.G.,
+2007. Impact of permethrin impregnated bednets on child mortality in
+kassena-nankana district, ghana: A randomized controlled trial. Tropical
+Medicine \& International Health 1, 147--154.
+\url{https://doi.org/10.1111/j.1365-3156.1996.tb00020.x}
+
 \hypertarget{ref-Bloom2008}{}
@@
 CCM, 2016. AngloGold Ashanti Malaria Control Ltd (AGA Mal).
+
+\hypertarget{ref-Cohen2010}{}
+Cohen, J., Dupas, P., 2010. Free distribution or cost-sharing? Evidence
+from a randomized malaria prevention experiment. Quarterly Journal of
+Economics 125, 1--45. \url{https://doi.org/10.1162/qjec.2010.125.1.1}
 
@@
 
-\hypertarget{ref-Lopez_Bernal_2016}{}
-Lopez Bernal, J., Cummins, S., Gasparrini, A., 2016. Interrupted time
-series regression for the evaluation of public health interventions: A
-tutorial. International Journal of Epidemiology dyw098.
-\url{https://doi.org/10.1093/ije/dyw098}
-
 \hypertarget{ref-Mayor2007}{}
@@
 
+\hypertarget{ref-OLaughlin2016}{}
+O'Laughlin, B., 2016. Consuming bodies: Health and work in the cane
+fields in xinavane, mozambique. Journal of Southern African Studies 43,
+625--641. \url{https://doi.org/10.1080/03057070.2016.1190519}
+
 \hypertarget{ref-Orem_2012}{}
@@
 
-\hypertarget{ref-OLaughlin2016}{}
-O'Laughlin, B., 2016. Consuming bodies: Health and work in the cane
-fields in xinavane, mozambique. Journal of Southern African Studies 43,
-625--641. \url{https://doi.org/10.1080/03057070.2016.1190519}
-
 \hypertarget{ref-Phillips98}{}

### 2018-05-06 inst/rmd/paper_prepare.R

@@ -93,12 +93,13 @@ if('prepared_data.RData' %in% dir()){
   model_data <- model_data %>%
     mutate(months_since = days_since) %>%
     mutate(months_since = as.character(months_since)) %>%
-    mutate(months_since = ifelse(months_since %in% c('Never', 'Before', '00', '01',
-                                                     '02', '03', '04', '05'),
+    mutate(months_since = ifelse(months_since %in% c(#'Never', 
+                                                     'Before', '00', '01',
+                                                     '02', '03', '04', '05', '06'),
                                  months_since,
-                                 ifelse(months_since %in% c('06', '07', '08',
+                                 ifelse(months_since %in% c('07', '08',
                                                             '09', '10', '11', '12+'),
-                                        '06+', NA))) %>%
+                                        '07+', months_since))) %>%
     # Define whether every sprayed
     ungroup %>%
     group_by(oracle_number) %>%
@@ -109,7 +110,7 @@ if('prepared_data.RData' %in% dir()){
     mutate(months_since = as.character(months_since)) %>%
     mutate(months_since = ifelse(months_since == 'Never' & ever_sprayed, 'Before', months_since)) %>%
     # Make never and before the same
-    mutate(months_since = ifelse(months_since %in% c('Never', 'Before'), 'No IRS', months_since)) %>%
+    # mutate(months_since = ifelse(months_since %in% c('Never', 'Before'), 'No IRS', months_since)) %>%
     # mutate(months_since = factor(months_since, levels = unique(c('Never', 'Before', sort(unique(months_since))))))
     # mutate(months_since = ifelse(is.na(months_since),
     #                              'No IRS',
@@ -118,7 +119,7 @@ if('prepared_data.RData' %in% dir()){
     rename(on_site = ever_sprayed)
   
   model_data <- model_data %>%
-    mutate(months_since = factor(months_since, levels = unique(c('No IRS', sort(unique(months_since)))))) 
+    mutate(months_since = factor(months_since, levels = unique(c('Before', sort(unique(months_since)))))) 
   
   # Since ad and factory are same, keep same
   model_data$field <- ifelse(model_data$department == 'Field', 
@@ -178,42 +179,47 @@ if('prepared_data.RData' %in% dir()){
[…]

### 2018-05-06 inst/rmd/02_methods.Rmd
lots of externalities work

@@
 No studies exist on positive externalities for IRS coverage, but to the extent that the mechanisms for the reduction in infection are similar (reduction in the natural reservoir of the disease, reduction in the number of vectors, etc.), it is reasonable to assume similar effects.
-Additionally, we using weighted distance for our score calculation (rather than simply defining a radius threshold) because of previous studies' findings that there is a linear decline in protection conferred to others with greater distance from an intervention [@Binka2007]
+Additionally, we are using weighted distance for our score calculation (rather than simply defining a radius threshold) because of previous studies' findings that there is a linear decline in protection conferred to others with greater distance from an intervention [@Binka2007].
+
+Figure x shows a smooth interpolated surface of herd protection, averaged for the entire study period.
+
+```{r, fig.cap = 'Kernel density surface of average all-time herd protection score for Maragra worker neighborhoods (relative to highest protection)', fig.align = 'center'}
+# Fortify bairros
+library(maptools)
+library(rgeos)
+border <- gConvexHull(bairros_maragra_bairro)
+border <- SpatialPolygonsDataFrame(Sr = border,
+                                   data = data.frame(a = 1))
+# border <- unionSpatialPolygons(bairros_maragra_bairro, 
+#                           IDs = rep(1, nrow(bairros_maragra_bairro)),
+#                           threshold = 9^9)
+border_fortified <- 
+  broom::tidy(border,
+              regions = 'Id')
+
+# Create average herd risk score
+x <- model_data %>%
+  group_by(lng = longitude_aura,
+           lat = latitude_aura) %>%
+  summarise(herd = mean(herd, na.rm = TRUE)) %>%
+  ungroup %>%
+  filter(!is.na(lng))
+x_sp <- x
+coordinates(x_sp) <- ~lng+lat   
+proj4string(x_sp) <- proj4string(border)
+overs <- sp::over(x_sp, border)
+x <- x[!is.na(overs),]
+# ggplot() +
+#   geom_point(data = x,
+#        aes(x = lng,
+#            y = lat,
+#            color = herd)) +
+#   geom_polygon(data = border_fortified,
+#                aes(x = long,
+#                    y = lat,
+#                    group = group),
+#                fill = NA,
+#                color = 'black') +
+#   ggthemes::theme_map() +
+#   scale_color_continuous(name = 'Herd\nprotection\nscore',
+#                          low = 'orange',
+#                          high = 'blue')
+
+df_grid <- expand.grid(lng = seq(bbox(border)[1,1],
+                                 bbox(border)[1,2],
+                                 by = 0.0008),
+                       lat = seq(bbox(border)[2,1],
+                                 bbox(border)[2,2],
+                                 by = 0.0008),
+                       herd = NA)
+df_grid$latitude <- df_grid$lat
+df_grid$longitude <- df_grid$lng
+coordinates(df_grid) <- ~longitude+latitude
+
+# Go through each row of df_grid, getting the 
+# weighted mean irs score for that point
+# and putting a color into df_grid
+if('df_grid.RData' %in% dir()){
+  load('df_grid.RData')
+} else {
+  for (i in 1:nrow(df_grid)){
+  # Get distance from every point in x_sp
+  distances <- spDistsN1(pts = x_sp,
+                        pt = df_grid[i,],
+                        longlat = TRUE)
+  # Define which are acceptably close
+  close_enough <- which(distances <= 50)
+  # Get an herd score
+  herd <- stats::weighted.mean(x = x_sp$herd[close_enough],
+                       w = (1 / distances[close_enough]) ^2,
+                       na.rm = TRUE)
+  # Assign irs to the dataframe
+  df_grid$herd[i] <- herd
+
+}
+save('df_grid', file = 'df_grid.RData')
+}
+
+library(raster)
+# Convert df_grid to raster
+temp <- df_grid@data %>% arrange(lng, lat)
+r <- rasterFromXYZ(temp[, c('lng', 'lat', 'herd')])
+
+proj4string(df_grid) <- proj4string(border)
+x <- over(df_grid, polygons(border))
+df_grid_small <- df_grid[!is.na(x),]
+temp <- df_grid_small@data %>% arrange(lng, lat)
+r <- rasterFromXYZ(temp[, c('lng', 'lat', 'herd')])
+values(r) <- values(r) / max(values(r), na.rm = TRUE) * 100
+plot(border, col = NA, border = NA)
+plot(r, add = TRUE)
+# plot(border, add = TRUE)
+plot(bairros, add = TRUE)
+```
+
+Some more details will eventually go here.
 

### 2018-05-06 inst/rmd/03_results.Rmd
lots of externalities work

@@
 
+## What would happen without IRS?
+
+The below chart shows the modeled average absenteeism rate during malaria season for permanent workers, by group type.
+
+```{r}
+x <- model_data %>%
+  dplyr::filter(season == 'high',
+                # group == groups[1],
+                permanent_or_temporary == 'Permanent'
+                )
+x <- x [,grepl('predicted', names(x)) | names(x) == 'group']
+x <- x %>% gather(key, value, predicted_no_herd_no_irs:predicted_max_irs)
+dict <- data.frame(key = c('predicted', 
+                           'predicted_no_herd',
+                           'predicted_no_irs',
+                           'predicted_no_herd_no_irs',
+                           'predicted_max_herd_max_irs',
+                           'predicted_max_herd',
+                           'predicted_max_irs'),
+                   new_key = c('As is',
+                               'No herd protection',
+                               'No individual protection',
+                               'No protection (herd or individual)',
+                               'Protection: maximum IRS and herd',
+                               'Protection: maximum herd',
+                               'Protection: maximum IRS'))
+
+x <- left_join(x,
+               dict)
+plot_data <- x %>%
+  mutate(new_key = gsub(' ', '\n', new_key)) %>%
+  group_by(group, 
+           new_key) %>%
+  summarise(avg = mean(value, na.rm = TRUE),
+            med = median(value, na.rm = TRUE),
+            upr = quantile(value, 0.75, na.rm = TRUE),
+            lwr = quantile(value, 0.25, na.rm = TRUE))
+levs <- gsub(' ', '\n', c('No herd protection',
+                               'No individual protection',
+                               'No protection (herd or individual)',
+                               'As is',
+                               'Protection: maximum herd',
+                               'Protection: maximum IRS',
+                               'Protection: maximum IRS and herd'))
+plot_data$new_key <- factor(plot_data$new_key,
+                       levels = levs)
+cols <- colorRampPalette(brewer.pal(n = 9,
+                                    name = 'Spectral'))(length(unique(plot_data$group)))
+
+ggplot(data = plot_data,
+       aes(x = new_key,
+           y = avg * 100,
+           color = group)) +
+  geom_point() +
+  theme_maragra() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  labs(x = 'Scenario',
+       y = 'Absenteeism rate') +
+  geom_line(aes(group = group),
+            alpha = 0.3) +
+  scale_color_manual(name = '',
+                     values = cols)
+```
+
 ## Translating absences to costs

### 2018-05-06 inst/rmd/paper_prepare.R

@@ -532,6 +532,168 @@ if('prepared_data.RData' %in% dir()){
   
   map_list <- list(g1,g2,g3,g4)
   
+  # Get the herd protection score assuming that everyone nearby was protected
+  groups <- sort(unique(model_data$group))
+  herd_ideal <- 
+    model_data %>%
+    filter(!is.na(longitude_aura),
+           !is.na(latitude_aura)) %>%
+    group_by(oracle_number) %>%
+    summarise(lng = dplyr::first(longitude_aura),
+              lat = dplyr::first(latitude_aura)) %>%
+    ungroup
+  coordinates(herd_ideal) <- ~lng+lat
+  proj4string(herd_ideal) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
+  # Get distances
+  distances <- spDists(x = herd_ideal)
+  out <- apply(distances, 1, function(x){
+    x <- weighter(x)
+    x <- x[is.finite(x)]
+    x <- sum(x, na.rm = TRUE)
+    return(x)
+  })
+  herd_ideal$herd_ideal <- out
+  herd_ideal <- herd_ideal@data
+  # Join back to model_data
+  model_data <-
+    left_join(model_data,
+              herd_ideal)
+  
+  
+  # Define function similiar to "predict" but for felm
+  predict_felm <- function(model, df,
+                           irs0 = FALSE,
+                           irs_all = FALSE,
+                           herd0 = FALSE,
+                           herd_max = FALSE){
+    # Extract fixed effects
+    fe <- getfe(model) %>%
+      dplyr::select(idx, effect)
+    fe$key <- row.names(fe)
+    # Extract the fixed effects
+    fe_oracle_number <- 
+      fe %>%
+      dplyr::filter(grepl('oracle', key)) %>%
+      dplyr::select(idx, effect) %>%
+      dplyr::rename(oracle_number = idx)
+    fe_malaria_year <- 
+      fe %>% 
+      dplyr::filter(grepl('year', key)) %>%
+      dplyr::select(idx, effect) %>%
+      dplyr::rename(malaria_year = idx)
+    # Get the coefficients for the non-fixed effects variables
+    betas <- model$beta
+    beta_names <- row.names(betas)
+    betas <- data_frame(var = beta_names, 
+                        beta = as.numeric(betas))
+    # Split up the vals, etc.
+    betas$variable <- gsub('high', '', betas$var)
+    betas$variable <- gsub('After', '', betas$variable)
[…]

### 2018-05-07 inst/rmd/paper.Rmd
major externality updates

@@
 ```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
+
 # Packages

### 2018-05-07 inst/rmd/paper.tex
major externality updates

@@
 reservoir of the disease, reduction in the number of vectors, etc.), it
-is reasonable to assume similar effects. Additionally, we using weighted
-distance for our score calculation (rather than simply defining a radius
-threshold) because of previous studies' findings that there is a linear
-decline in protection conferred to others with greater distance from an
-intervention (Binka et al., 2007)
+is reasonable to assume similar effects. Additionally, we are using
+weighted distance for our score calculation (rather than simply defining
+a radius threshold) because of previous studies' findings that there is
+a linear decline in protection conferred to others with greater distance
+from an intervention (Binka et al., 2007).
+
+Figure x shows a smooth interpolated surface of herd protection,
+averaged for the entire study period.
+
+\begin{figure}[!h]
+
+{\centering \includegraphics{figures/unnamed-chunk-17-1}
+
+}
+
+\caption{Kernel density surface of average all-time herd protection score for Maragra worker neighborhoods (relative to highest protection)}\label{fig:unnamed-chunk-17}
+\end{figure}
+
+Some more details will eventually go here.
 
@@
 
-\subsection{Estimating return on
-investment}\label{estimating-return-on-investment}
-
 Our formula for return on investment can be described in a
@@
 ``Treatment'' is considered the time beginning at IRS administration and
-ending one year later; ``Control'' is considered time observed either
-prior to IRS administration (and greater than 1 year after previous IRS
-administration) or among workers who never received IRS. Because of the
-temporal nature of these categories, many workers belong to both
+ending six months later; ``Control'' is considered time observed either
+prior to IRS administration (and greater than 6 months after previous
+IRS administration) or among workers who never received IRS. Because of
+the temporal nature of these categories, many workers belong to both
 treatment and control groups, albeit at different times.
@@
   \hline
-Department & Administrative &  38 & 47992 &   0 &   0 \\ 
-   & Factory & 109 & 78821 &   0 &   0 \\ 
-   & Field & 545 & 268206 &   0 &   0 \\ 
-  Sex & F & 310 & 139959 &   0 &   0 \\ 
-   & M & 382 & 255060 &   0 &   0 \\ 
-  Status & Permanent & 128 & 182199 &   0 &   0 \\ 
-   & Temporary & 564 & 212820 &   0 &   0 \\ 
-  Age & 20 & 200 & 81640 &   0 &   0 \\ 
-   & 30 & 242 & 124339 &   0 &   0 \\ 
-   & 40 & 154 & 101317 &   0 &   0 \\ 
-   & 50 &  68 & 63739 &   0 &   0 \\ 
-   & 60 &  26 & 22257 &   0 &   0 \\ 
-   & 70 &   2 & 1727 &   0 &   0 \\ 
+Department & Administrative &  35 & 12523 &  38 & 35469 \\ 
+   & Factory &  71 & 20533 & 103 & 58288 \\ 
+   & Field & 410 & 91302 & 490 & 176904 \\ 
+  Sex & F & 228 & 47907 & 279 & 92052 \\ 
+   & M & 288 & 76451 & 352 & 178609 \\ 
+  Status & Permanent & 127 & 47075 & 128 & 135124 \\ 
+   & Temporary & 389 & 77283 & 503 & 135537 \\ 
+  Age & 20 & 140 & 29463 & 175 & 52177 \\ 
+   & 30 & 176 & 39119 & 223 & 85220 \\ 
+   & 40 & 123 & 30186 & 141 & 71131 \\ 
+   & 50 &  58 & 18881 &  65 & 44858 \\ 
+   & 60 &  17 & 5893 &  25 & 16364 \\ 
+   & 70 &   2 & 816 &   2 & 911 \\ 
    \hline
@@
 its effect on malaria is discussed in the model results section.
-
-\textbf{Effect of IRS on clinical malaria}: The number of malaria cases
-registered at the company clinic during that time was 1873, for an
-approximate annualized clinical incidence of 110 cases per 1,000.
-\todo{Not done yet. In fact, not even sure if I should do this section at all, since it's basically just a distraction.}
 
@@
 Maragra - so as to capture other seasonal effects from those workers.
-\todo{This is not very eloquent yet and needs to be improved.}
+
+The herd protection score, unit-less by nature given that it's a
+summation of binary protection statuses weighted by inverse-transformed
+distances, took the following distribution in our data:
+
+\begin{figure}[!h]
+
+{\centering \includegraphics{figures/unnamed-chunk-28-1}
+
+}
+
+\caption{Distribution of herd protection score in panel data}\label{fig:unnamed-chunk-28}
+\end{figure}
 
@@
 \begin{longtable}[t]{ll}
-\caption{\label{tab:unnamed-chunk-27}All absenteeism with herd immunity: model results}\\
+\caption{\label{tab:unnamed-chunk-29}All absenteeism with herd immunity: model results}\\
 \toprule
@@
 
+\subsection{What would happen without
+IRS?}\label{what-would-happen-without-irs}
+
+The below chart shows the modeled average absenteeism rate during
+malaria season for permanent workers, by group type.
+
+\includegraphics{figures/unnamed-chunk-31-1.pdf}
+
 \subsection{Translating absences to
@@
 \centering
-\includegraphics{figures/unnamed-chunk-32-1.pdf}
+\includegraphics{figures/unnamed-chunk-35-1.pdf}
 \caption{i. Absenteeism before and after IRS administration for all
@@
 \centering
-\includegraphics{figures/unnamed-chunk-33-1.pdf}
+\includegraphics{figures/unnamed-chunk-36-1.pdf}
 \caption{i. Sick absenteeism before and after IRS administration for all

### 2018-05-11 inst/rmd/paper.Rmd
last save before big clear

@@
 ---
-title: Is malaria control profitable? Return on investment of privately-managed residential fumigations at a large sugarcane processing facility in Southern Mozambique
+title: Is malaria control profitable? Return on investment of privately-managed residential
+  fumigations at a large sugarcane processing facility in Southern Mozambique
+csl: apa_modified.csl
 date: ''
 output:
+  word_document: default
   pdf_document:
-    keep_tex: true
-    latex_engine: xelatex
     includes:
       in_header: preamble.sty
+    keep_tex: yes
+    latex_engine: xelatex
 fig_width: 4

### 2018-05-11 inst/rmd/01_introduction.Rmd
changes
removed text

--- 2018-04-18 inst/rmd/01_introduction.Rmd
@@
 # Introduction
-
-Malaria accounts for a half million annual deaths worldwide [@White2014] [@World2016] [@Ashley2018].
-Though rapid improvements in technology and funding have lead to important reductions in marlaria's global burden, the scale-up in activities required for eradication ("the worldwide interruption of transmission") [@TheLancet2011] will mean new partnerships and actors.
-One promising - albeit atypical - potential stakeholder in global malaria eradication is the private sector, given its omnipresence and potential to benefit directly from the elimination of malaria.
-But little evidence exists demonstrating how private sector entities can engage with malaria control and benefit at the firm level.
-
-At the societal level, malaria has a large economic impact.
-By affecting saving, investment [@Shretta2016], risk perception, productivity, absenteeism [@Nonvignon_2016], human capital accumulation [@CastelBranco2014], mortality, and costs of care [@Sachs2002], malaria likely has a negative effect on GDP and growth [@McCarthy_2000] [@Orem_2012].
-Because of the relative affordability of most intereventions and the enormous societal costs of malaria, most forms of malaria control are cost-effective when a public welfare perspective is assumed, such as when a government provides the financing [@White_2011] [@Purdy_2013] [@Howard_2017].
-
-From the persepctive of the private sector, however, investing in malaria control is not so clear-cut.
-Public health interventions targetting malaria - and their corresponding cost-effectiveness evaluations - most often focus on impacts pertaining to public welfare, such as an increase in life years adjusted for disability or quality  [@Goodman1999] [@Shretta2016] [@Lee2017] [@Hanson2004].
-Though population-level health is certainly of importance to businesses, and improvements in health incidentally improve the economy at all levels [@World1999] [@Bloom2008] [@Vecchi_2013], these improvements may be too disperse or long-term to incentivize private sector involvement in health campaigns.
-
-100% of the Mozambican population are at risk of malaria, living in what the WHO classifies as a "high transmission" area [@Moonasar_2016].
-Annually, Mozambique has more than 8 million clinical malaria cases (an annual incidence of approximately 300 per 1,000 residents), with an estimated 14,000 deaths.
-Malaria accounts for 29% of all deaths, and 42% of deaths among those under five years of age [@estatistica2009].
-Since 2013, Mozambique has seen a gradual increase in the incidence of malaria [@Moonasar_2016].
-100% of the malaria in Mozambique is of the Plasmodium falciparum species, with Anopheles funestus, gambiae, and arabiensis as the primary mosquito vectors of the disease [@whoprof].
-
-A significant sector of the economy in Mozambique is dominated by large-scale foreign direct investment projects [@Robbins2012], and the role of the private sector in health generally, and malaria specifically, is unequivocally important.
-Large agriculture and extractive industry firms take up wide swaths of land and employ hundreds of thousands  [@German2013].
-The Mozambican state has encouraged large-scale entreprise with the aim of general economic development [@Buur2012].
-And where large firms exist, they often take on social roles such as housing and health care [@Winkler].
-At times, this role is necessary from a purely practical standpoint; in other cases, it is employed under the auspices of "corporate social responsibility" [@Azemar2009].
-Regardless of the language used, it is clear that private industry plays an important role in public health in Mozambique [@Robbins2012] [@CastelBranco2014].
-
-Several cases exist of foreign firms engaging in large-scale malaria control campaigns [@Mouzin2011] [@Han] [@Bennett_2017] [@Kaula_2017].
-But these studies generally consider population health as the outcome measure of interest, rather than worker absenteeism or productivity.
-Similarly, they often neglect to differentiate between those clinical costs which are absorbed by the local health system versus those which are absorbed by the firm itself.
-In the literature, making the "investment case" for malaria control or elimination generally implies that the investor is the public sector, and takes into account those costs and benefits which are applicable from a public welfare point of view [@Shretta_2017]; though appropriate in most cases (the government or institutions interested in public welfare primarily being the primary malaria control agents in most locations), the findings of these studies are rarely applicable to the private sector, and even less so at granular levels (such as an individual firm).
-In the case of a private firm not interested in "corporate social responsibility", it is not clear whether investing in malaria control would be profitable or not.
-This lack of clarity not only discourages investment, but also makes it difficult for governments to pinpoint the correct of amount of subsidy (if applicable) to encourage private sector scale-up in malaria control.
-
-To address the question of the profitability of malaria control activities from the standpoint of a private firm, we analyze data during a 4 year period from a private sugar facility in Southern Mozambique.
-We assess the effect of indoor residual spraying (IRS) on the absenteeism and health of workers, and demonstrate that the firm's engagement in malaria control not only improved worker health, but also generated a positive return on investment from a pure accounting perspective.
-
-The structure of the paper is as follows.
-After a brief review of the existing literature on the subject of private sector investment in malaria control and its effects on worker productivity, we provide an overview of the sugar company under study, and the epidemiology of malaria in the nearby area, as well as in Mozambique as a whole.
-We then give an overview of the data collected, and outline the theoretical and methdological assumptions and tools that underly our analysis.
-In the results, we show the effectiveness of the company's malaria control program in reducing absences, and translate this reduction into cost savings.
-Our discussion covers potential implications from this study in terms of policy and investment, as well as the paper's limitations.
-
-The literature on the effect of sugarcane cultivation on malaria risk is mixed.
-While some studies have found that the prevalence of malaria vectors in sugarcane areas to be similar to that of uncultivated areas (and less than in areas dedicated to other forms of more water-inensive agriculture, such as rice) [@Ijumba2002], other studies have found significant increases in factors associated with malaria transmission at large-scale sugarcane facilities relative to traditional, small-scale farming and non-irrigated farming [@Jaleta2013].
-Regardless of the effect the presence of a sugarcane farm _per se_ on local malaria epidemiology, the time spent outdoors by sugarcane workers, the fact that many workers are migrants, and their sometimes precarious housing situations, suggest that sugarcane farmers are likely at increased risk of malaria infection [@OLaughlin2016].
-This is important, given that even among occupations with far less inherent exposure to mosquitoes (such as health professionals), malaria is one of the primary causes of work absenteeism in malaria-endemic countries [@Burton1999].
-There is also some concern regarding the effect of large-scale insecticide use - common at essentially all Sub-Saharan African sugarcane farms - on insecticide resistance among moquitoes in the area.
-A study in Belize found that mosquito populations on the edge of sugarcane fields had higher tolerance to insecticide than similar populations in the core of fields or outside of the periphery [@Dusfour2009].
-Sugarcane areas may offer the standing water necessary for mosquito breeding, but also perhaps attract mosquitoes which would otherwise be elsewhere, due to compounds in sugarcane pollen [@Wondwosen2018].
-
-Regardless of the academic literature on the risk of workers becoming infected with malaria, the companies running large agricultural operations in malaria-endemic regions generally consider the disease to be an important enough threat to workers' health to merit some form of investment in malaria control [@Pluess2009].
-Insecticide fumigation, bednet distribution, test-and-treat clinics, and occasional educational campaigns are common forms of corporate social responsibility to combat malaria.
-A World Bank study on Nigerian sugarcane cutters found that the simple availability of testing and treatment increased productivity by 10% in the weeks following the provision of services, the conclusion being that both the treatment and the test result were effective in increasing productivity, the latter simply increasing the information available which could influence personal labor allocation decisions [@Dillon2014].
-
-Malaria may disincentivize investment in business in malaria zones.
-In a vicious cycle, businesses already in malaria zones may not be incentivized to invest in malaria control, perhaps because this is seen as the domain of the public sector [@Curtis2003], or perhaps due to not knowing the extent to which the benefits in health of that control will accrue to the business itself.
-
-The evidence of malaria's negative effects on both health and wealth, at both the individual [@Cole2006] and collective [@McCarthy2000] levels, as well as in both the short [@AsensoOkyere1997] [@Ajani2010] and long [@Hong2011] [@Sachs2002] terms, are amply described in the public health and economics literature [@Phillips98].
-That said, very little exists in the literature examining the costs and benefits of malaria control from a private ledger perspective (ie, the point of view a business investor).
-Unlike a government or individual, a private firm investing in malaria control may be most interested not in its long-term macroeconomic effects, nor its short-term personal health effects, but rather on the impact on productivity (and the extent to which that producitivty's benefits accrue to the firm), absenteeism, and the opportunity costs of expenditures in malaria control.
-In other words, though the large magnitude of malaria control's benefits are well known, the portion of those benefits accrued by a private firm investing in malaria control is unknown.
-
-This study does not endeavor to expand the current body of knowledge regarding malaria's ill effects on individuals and societies; rather, it aims to provide empirical evidence pertaining to a facet of malaria economics with very little in the literature: malaria control from a private-sector investment perspective.
-Our study adds to the existing literature by showing the effect of specific malaria control interventions on worker absenteeism, and translating that effect into a return on investment.
-Unlike previous studies on the effectiveness of malaria control interventions, our's focuses solely on one firm carrying out one intervention, takes advantage of individual-level data, and analyzes results from a ledger perspective.
-
-## Study area
-
-Sugar has been systematically cultivated in Mozambique since the late 1800s.
-The Incomati Estates company, a small sugarcane processing facility started by a Scotsman on the banks of the Incomati River in 1913, was the first firm to export sugar from Mozambique.
-Following its purchase by international investors in the 1950s, it (along with the rest of the industry) expanded significantly, exporting to both Europe and the United States.
-In the late 1960s, a Portuguese family opened the Maragra Açúcar company, while a group of foreign investors started the nearby Marracuene Agrícola Açucareira mill.
-By the early 1970s, sugar grew to account for greater than 10% of Mozambique's national exports.
-Nationalized following independence in 1977, the industry's production levels fell from 320,000 annual tons to fewer than 15,000 by 1992.
-After the end of the civil war, foreign investment revived the sugar industry, and by 2011 production had surpassed its 1972 peak.
 
@@
 
-The mill of the Maragra Açucar SA (a subsidary of the Illovo sugar company, henceforth referred to as "Maragra") was nationalized in the 1970s (like all other Mozambican mills), went through a period of low production, and then fell completely out of use by 1984.
-It re-opened in private hands in 1992, and was renovated by a group of international investors in 1998.
-Today, Maragra accounts for roughly one quarter of Mozambique's overall sugar production (second only to the nearby Xinavane mill run by the Hulett Sugar Tongaat company) [@sutton].
-With a favorably close location to the port of Maputo, ample land, approximately 5,000 employees (of which three quarters are seasonal), and a mill with the capacity to process not only all the sugarcane grown on Maragra's land, but also the cane of the many nearby smallholders [@OLaughlin2016], Maragra has so far been able to weather the 2016 Mozambican crisis and concurrent collapse in global sugar prices.
+```{r}
+# Calculate area of maragra
+# x <- bairros
+# x$NeighCode <- as.numeric(as.character(x$NeighCode))
+# x@data <- x@data %>%
+#   mutate(maragra_bairro = NeighCode %in%
+#            c(601:609,701:709, 1001:1003)) %>%
+#   mutate(maragra_fabrica = NeighCode %in%
+#            1001:1003)
+# x <- spTransform(x,CRS("+proj=utm +zone=36 +datum=WGS84 +units=m") )
+# 
+# the_bairro <- x[x$maragra_bairro,]
+# the_fabrica <- x[x$maragra_fabrica,]
+# # the_bairro <- spTransform(the_bairro, CRS("+proj=utm +zone=36 +datum=WGS84 +units=m") )
+# # the_fabrica <- spTransform(the_fabrica, CRS("+proj=utm +zone=36 +datum=WGS84 +units=m") )
+# 
+# rgeos::gArea(the_bairro) / 1000000
+# rgeos::gArea(the_fabrica) / 1000000
 
-Maragra (figure 1, panel iii) is located in the district of Manhiça (figure 1, panel ii), a semi-rural area in the south of Mozambique (figure 1, panel i).
-80 kilometers north of the Mozambican capital of Maputo, the district is low-lying, consisting largely of savannah and wetlands along the Incomati River.
-Most of the areas 160,000 residents [@Sacoor2013] work as subsistence farmers.
-Migration from the area to South Africa for the purpose of employment in the profitable construction industry is common, especially among men [@Nhacolo_2006], as is migration to the area (from other parts of Mozambique) for seasonal work on the sugarcane plantations at Maragra and the slightly larger facility in Xinavane (at the district's border with Magude) [@OLaughlin2016].
-
-Poverty is rife in southern Mozambique, and its associated illnesses take their toll on the population.
-The community prevalence of HIV/AIDS is as high as 40% [@Gonzlez2012]; even in a more recent study suggesting a much lower prevalence of 22%, the risk of infection is still twice that of nearby areas [@Mocumbi2017].
-Recent years have seen a three-fold increase in tuberculosis [@GarcaBasteiro2017].
-Malaria, which has the greatest mortality burden due to the fact that the young are especially vulnerable to its effects, is perennial, though worse during the rainy season (December - March) [@Sate2003].
-Adult malaria is essentially universal, albeit much of it asymptomatic [@Mayor2007].
-In regards to worker health, malaria is Maragra's primary concern, being so important as to justify the existence of both an on-site testing laboratory and clinic, as well as a malaria control department.
-
-Maragra workers are mostly seasonal, working for the firm approximately half of the year during harvest time, and cultivating crops, working in construction, or going unemployed (or working elsewhere) during the other half.
-Though many workers live "on-site" (ie, within the delineated property of the firm itself), a sizeable minority resides in the environs (figure 1, panel iv).
-Maragra provides indoor residual spraying (IRS) using ACT (alpha-cypermethrin) and DDT (Dichlorodiphenyltrichloroethane) depending on stocks (a preference apparently exists for the former, but it is not always available).
-IRS activities, managed by Maragra's Malaria Control program, are ongoing throughout the year (Figure 1, panel i).
-Nearly all on-site houses are sprayed, though the time between fumigations is irregular (Figure 2, panel ii).
-Off-site houses may also receive IRS (managed and carried out by the National Malaria Control Program), but the status and timing of these fumigations is not known at the individual level.
+# Used this for plantation estimation: https://www.daftlogic.com/projects-google-maps-area-calculator-tool.htm
+```
 
@@
 ```
+
+\newpage

### 2018-05-11 inst/rmd/02_methods.Rmd
removed text

@@
 # Methods
-
-## Data
-
-In collaboration with the sugar processing facility, we collected data for the period from January 2010 through December 2016.
-Data came from four sources: (i) the Human Resources' roster of worker details and absences, (ii) the facility's on-site clinic's medical and laboratory records, (iii) the facility's on-site malaria control program's records pertaining to the dates, chemicals, and location of IRS activities, and (iv) interviews with company employees pertaining to costs, data limitations, etc.
-Digitization and collection of data took place during the period from March 2016 through May 2017.
-Supplementary data pertaining to worker characteristics was obtained from through the Centro de Investigação em Saude de Manhiça's (CISM) demographic census, which covered workers from the district, but not those who migrated from other parts of the country [@Nhacolo_2006].
-
-Data pertaining to district-wide malaria incidence was obtained from Mozambique’s Boletim Epidemiológico Semanal (BES), which is the system by which the National Malaria Control Program monitors incidence at the district level throghout the entire country, and reports the number of confirmed weekly malaria cases at government health facilities.
-Using these case numbers, combined with population estimates from the National Statistical Institute (INE), we estimate each day's annualized weekly malaria incidence rate (cases per 1000 population at risk), interpolated from the weekly figures.
-We retrieved weather data for all Mozambican stations from NOAA.
-We estimated the meteorological conditions at the centroid of Manhiça using a simple interpolation method whereby the district's weather conditions were estimated to be a function of all Mozambican weather stations' reported conditions, inversely weighted by kilometers from district centroid.
-
-Maragra regularly employs IRS at on-site worker households in order to reduce those workers' (and their families') risk of malaria infection.
-IRS works by killing the malaria vector (mosquito), thereby preventing infection of the household occupants.
-When administered correctly, IRS is a low-risk intervention to its recipients, and is assumed not to affect absenteeism in the short-term (to the extent that it may cause negative side-effects, these are generally long-term).
-It is preventive only, and does not cure current malaria infection, nor does it affect the parasite load of mosquitoes.
-Workers living off-site (our control group) also may have received IRS at some point during the study period (from government programs).
-Even though we do not have reliable person-level data on IRS carried out by the government, off-site workers are a suitable control in the sense that they represent "business as usual" (ie, what would happen if the company carried out no IRS and relied solely on public interventions).
-Using company HR and clinical records, we were able to identify absences and episodes of clinical malaria among all workers, as well as identify the time since the most recent IRS episode before the onsent of absence or illness.
-
-Worker characteristics, illness and absenteeism data, along with IRS activity data, were systematically stored, collected, and used at the individual level by Maragra, and therefore of generally high quality.
-Because cost data was less systematically collected by Maragra, and because many costs could not be precisely quantified due to the abundance of in-kind and cross-departmental expenditures, we had to rely on rough estimations based on a mix of interviews, reciepts, and interpolations.
-Since our program cost data is not as reliable as our worker characteristic and outcome data, we were conservative in our estimates, and generally tried to err on the side of program activities and materials costing _more_ than what was reported, when doubt was aired.
-Cost data consisted of three types: (i) wages of malaria control employees, (ii) transporation and vehicle costs for IRS teams, and (iii) acquisition costs of purchasing IRS chemicals for fumigation (ACT and DDT), the latter two being combined into malaria control "programme" costs.
-
-## Conceptual framework and identification strategy
-
-We sought to understand the effect of IRS on individual workers' likelihood of absence from work as well as their likelihood of clinical malaria.
-To estimate this effect, we estimated separate models for absence and illness.
-We employed linear mixed effects models [@Minalu2011] [@Liu2010] (with individual worker fixed effects, and random effects for other covariates) so as to explicitly adjust for confounders [@Bell2014].
-We divide into 4 different models for different worker types, so as to account for the potential time-varying effects of worker type on risk of malaria.
-
-### Externalities
-
-The administration of insecticide at a worker's residence (indoor residual spraying or IRS) likely protects that worker from malaria infection by killing the vectors (mosquitoes) that land on the building's walls.
-However, it is also highly likely that the protective effective of IRS "spills over" to others who live nearby.
-This positive spillover effect would theoretically go through two channels: (i) via a reduction of mosquitoes in the vicinity and (ii) via a reduction of the malaria parasite in the blood of humans in the vicinity (ie, the parasite "reservoir").
-This document describes our method for assessing the existence and magnitude of positive spillover of IRS in the Maragra workers' data.
-We devise a time-specific household "protection" score based on the theoretical effectiveness of IRS, and then use that protection score to develop a time-place specific "herd protection" score based on a weighted average of nearby household protection scores.
-
-We consider a house's "herd" protection level (ie, the protection conferred to the house through externality) to be the average of the other houses' non-herd protection levels, weighted by the distance to the house in question.
-The below is an illustration of a house (the red x) and the neighboring houses (circles).
-The "weight" of each house is indicated by the circle size, and the protection level of the house is indicated by the shading of the circle.
-
-However, this consideration misses one important dimension: density.
-The below house's weighted average protection score is identical to that of the above house.
-In other words, at the time in question, at both locations, the percentage of houses with IRS coverage is the same, and the (weighted) average time since IRS is the same.
-However, the below house is likely much more protected than the above, given the _absolute_ number of nearby IRS-protected houses.
-
-Our approach is this: rather than using a weighted average of each household's individual protection score, we instead use a weighted sum.
-A particular household's herd protection score at any given time is the sum of all households' individual protection scores at that time multiplied by those households' distance weights.
-
-The summation of (a) the IRS level of other houses at a certain time, (b) weighted by the distance of those houses to the house in question, (c) limited to only the 1 km radius multiplied yields a "herd protection score".
-
-This score is conceptually similar to Cohen and Dupas' quantification of the positive externalities of ITN (bednet) use in Kenya [@Cohen2010] in that it attempts to estimate the protection conferred to "non-users" by users and uses a "weighted sum" (as opposed to average).
-Our approach differs in that the time dimension for IRS is much more important than ITN (ie, IRS is not a binary but rather a waning function, as described previously), requiring us to create herd scores for every day.
-
-Our approach can be justified in that previous studies have found strong positive health externalities in malaria interventions related to ITN coverage [@Alaii2003FactorsAU].
-No studies exist on positive externalities for IRS coverage, but to the extent that the mechanisms for the reduction in infection are similar (reduction in the natural reservoir of the disease, reduction in the number of vectors, etc.), it is reasonable to assume similar effects.
-Additionally, we are using weighted distance for our score calculation (rather than simply defining a radius threshold) because of previous studies' findings that there is a linear decline in protection conferred to others with greater distance from an intervention [@Binka2007].
-
-Figure x shows a smooth interpolated surface of herd protection, averaged for the entire study period.
 
@@
 
-Some more details will eventually go here.
-
 ### Econometric model
@@
 
-We define the malaria season as any time during which the clinical incidence of malaria in the district of Manhiça was at or greater than the median clinical incidence of malaria for the entire study period.
-These weeks are flagged as red in Figure 1, Panel A.
-By using clinical incidence of the area of residence of the workers (as opposed to more typical proxies for malaria risk, such as only rainy vs.
-non rainy season), our seasonality estimate is a closer approximation of true malaria risk, incorporating lagged effects such as the incubation period of the parasite, as well as any inherent non-linear effects of weather.
-In addition, we adjust for daily precipitation; though its lagged effect on malaria incidence is likely captured by the seasonality term, we include rainfall since it's immediate effect (through its impact on transportation and working conditions) may also affect absenteeism.
-
 Our formula for return on investment can be described in a straightforward fashion...
@@
 
-## Reproducibility and ethical approval
-
-All data processing and analysis were carried out in R [@R] and all analysis code is freely available online [@brewgit].
-Ethical approval for this project was obtained from the Institutional Ethics Review Board for Health at the CISM (CIBS-CISM) prior to data collection.
-
 \newpage

### 2018-05-11 inst/rmd/03_results.Rmd
removed text

@@
 We collected absenteeism and demographic data data on `r n` workers from 2012 through 2016.
-Workers were approximately 60% male, and more than 80% fieldworkers (predominantly cane-cutters).
-Most were in their 20s and 30s and employed on temporary contracts.
-Table 1 shows overall worker details divided by "Treatment" versus "Control" status.
-"Treatment" is considered the time beginning at IRS administration and ending six months later; "Control" is considered time observed either prior to IRS administration (and greater than 6 months after previous IRS administration) or among workers who never received IRS.
-Because of the temporal nature of these categories, many workers belong to both treatment and control groups, albeit at different times.
 
@@
 names(demo_table)[1:2] <- c('Variable', ' ')
-print(xtable::xtable(demo_table,
-                     caption = 'Overall worker characteristics',
-                     align = c('r', 'r', rep('l', 5))),
-      include.rownames = FALSE)
+kable(demo_table, format = 'html') %>% kable_styling()
+# print(xtable::xtable(demo_table,
+#                      caption = 'Overall worker characteristics',
+#                      align = c('r', 'r', rep('l', 5))),
+#       include.rownames = FALSE)
 
 ```
-
-Temporary workers tended to be younger and male; female temporary workers being on average 5-10 years older than their male counterparts.
-Permanent workers had a bi-modal age distribution, the older peak explained by the greater density of males in administrative roles.
-Females accounted for 43% of temporary workers but only 16% of permanent workers (figure x).
 
@@
 ```
-
-During the study period, weather followed typical seasonal trends for the area, albeit slightly drier than previous periods.
 
@@
 ```
-
-In Southern Mozambique, malaria peaks during the summer months (December through March) most years (Figure 5, panel A), and worker absenteeism at Maragra rates track malaria incidence closely, following the same seasonal patterns (figure 4, panel B).
-Both all-cause absenteeism and sick absenteeism have declined in recent years at Maragra (figure 5, panel C), with the latter declining at a faster rate than the former.
-The fact that the rate of confirmed cases at the company clinic is largely non-seasonal (figure 5, panels E and F) suggests that a significant portion of workers either seek care for malaria elsewhere (for example, government health posts, of which several are nearby and in some many cases closer to workers' residence than the company clinic) or do not seek care during malaria infection.
-Accordingly, we focus our analysis on all-cause absenteeism rather than sickness absenteeism or malaria diagnostics, with the assumption that much of illness is captured by absenteeism but not by the clinical data.
 
@@
 names(ab_table)[2] <- ''
-print(xtable::xtable(ab_table,
-                     caption = 'Absenteeism rate by year and worker characteristics',
-                     align = c('r', 'r', 'l', rep('l', 4))),
-      include.rownames = FALSE)
+kable(ab_table, format = 'html') %>% kable_styling()
+# print(xtable::xtable(ab_table,
+#                      caption = 'Absenteeism rate by year and worker characteristics',
+#                      align = c('r', 'r', 'l', rep('l', 4))),
+#       include.rownames = FALSE)
 ```
@@
 
-The extent to which rainfall's effect on absenteeism is confounded by its effect on malaria is discussed in the model results section.
-
-**Costs of program**: The malaria control program at Maragra has an annual operating budget of approximately $112,000, which includes the purchase of insecticide, the wages of IRS sprayers and drivers, transportation, record-keeping, and general administrative costs.
-Assuming linearity in costs, the program spends approximately $19 per building sprayed.
-With each agregado containing an average of 2.2 workers.
-Much of the benefit of IRS goes to non-worker residents of sprayed agregados (who constitute a majority), but this benefit is purposefully ignored for this analysis.
-
-**Cost of malaria**: Given the likelihood that clinical data does not fully capture all malaria cases (and most likely captures only a small fraction of actual infections, given the high rates of acumulated adult immunity [Mayor2007]), we do not quantify the costs of malaria infection to the company.
-Rather, we first estimate the reduction in absenteeism attributible to IRS, and then quantify the savings associated with prevented absences.
-Additionally, we calculate the clinical savings of IRS by first estimating the share of absences which are associated with an episode of clinical malaria, and then applying the clinical cost per case to the equivalent share of prevented absences.
-We intentionally ignore the savings accrued by the public health system, as well as the likely utility gains in secondary realms such as school absenteeism, producitivity, etc.
-
-## Analysis
-
-### Effect of IRS on absenteeism
-
-Immediately following IRS, a worker's likelihood of absence drops significantly (appendix figure, panel A).
-As one would expect if the mechanism by which IRS reduces absence is through reduced malaria infection, the effect of IRS during the low transmission season is significant, but far less substantial in effect size (appendix figure, panel B).
-
-In order to assess the effect of IRS on absenteeism, we created 4 worker fixed effects models for the 4 principal worker types (permanent field worker, temporary field worker, permanent non-field worker, temporary non-field worker) for both outcoms (all cause absenteeism vs.
-sickness absenteeism).
-The reason for segregating models rather than incorporating worker type as a fixed effect was that it seemed plausible that the effect of the treatment (IRS) on the outcome would be different (and vary differently over time) depending on these worker types.
-For instance, it is reasonable to assume that IRS' effect would be differential for a temporary worker (who likely lives in the fumigated house less than 100% of the year) than for a permanent worker.
-
-Given that IRS is carried out on a continuous rolling basis, we used relative worker time (as opposed to calendar) as the time component of our model - that is, each workers' days until and since IRS were standardized so that day 0 (date of fumigation) was the interruption point.
-This has the advantage of incorporating into the model a variety of potential time-varying confounders in a quasi-randomized fashion, thereby making it unecessary to adjust for them specifically.
-Our model took into account workers' residential location (on or off site) - even though off site workers by definition could not be administered IRS by Maragra - so as to capture other seasonal effects from those workers.
-
-The herd protection score, unit-less by nature given that it's a summation of binary protection statuses weighted by inverse-transformed distances, took the following distribution in our data:
-
 ```{r, fig.align = 'center', fig.cap = 'Distribution of herd protection score in panel data'}
@@
 
-The model results for all cause absenteeism (table 4) suggest a significant decrease in absenteeism following the administration of IRS.
-The effect is ambivalent among temporary workers, whereas among permanent workers the administration of IRS decreases absenteeism by xxx.
-These effects are consistent with IRS decreasing absence through the medium of malaria, since (a) the effect is not as strong in the first month (potentially during the parasite incubation phase), the effect is stronger on field workers (who are more socioeconomically the demographic at greater risk of malaria than their administrative counterparts), and the effect is ambivalent among temporary workers (since the protective effect of IRS is only a function of the amount of time that the worker spends in the fumigated house).
+\newpage
 
@@
 make_models_table(model_list = protection_models, the_caption = "All absenteeism with herd immunity: model results")
-```
-
-```{r}
-# Make dummy data for seeing effect of herd on irs
 ```
@@
 
-## Translating absences to costs
-
-xxx\todo{Not done yet. This will be relatively straightforward arithmetic, but I want to first make sure that the models are rock solid (since all of the "translation" hinges on them).}
-
-## Return on investment
-
-\todo{Not done yet. In conversation with Eduardo from MC about getting more accurate costs}
-
-**Savings**
-
-- In percentage point terms, reduction from 13% to 8%.  
-- 5 annually prevented absences per worker.   
-- 8,000 workers: 40,000 prevented absences, wage of 3 USD
-- TOTAL: 120,000 USD in productivity-only savings.
-
-**Costs**
-
-- 8 IRS workers, 1500 USD yearly = 12,000 USD  
-- ACT + DDT: 50,000 USD 
-- Facilities, vehicles, gas: 50,000 USD  
-- TOTAL: 112,000 USD in IRS-only costs
-
-**7% ROI** (ignoring clinical costs)
-
 \newpage
@@
 ## Robustness and generalizability
-
-Two principal concerns call into question the results of our analysis.
-First, the application of IRS to a workers house may be endogenous.\todo{Is this the right term here?
-We're not dealing with an "unobserved" factor per se, rather we're dealing with a situation in which the cause might influence the effect, and the effect might influence the cause, ie a feedback loop.} It is reasonable to suspect that the application of IRS to households is not random, but rather that IRS was applied more frequently to houses which had already seen a malaria case.
-In this case, our estimated effect of IRS on absenteeism would likely be underestimated, with the post-IRS absenteeism rates actually having declined from a greater pre-IRS absenteeism rate than otherwise suggested.
-
-To check for this, we estimate the odds of absenteeism as a function of receiving IRS _during the 10 day period prior to receiving IRS_.
-If IRS application is indeed endogenous, we would expect absenteeism to be elevated during this period (since the increase in absenteeism would be theoretically responsible for the application of IRS), a situation which would require further statistical adjustment.
-If, on the other hand, there is no endogeneity, we would expect absenteeism in the 10 day period prior to IRS administration to be similar to other pre-IRS absenteeism (adjusting for seasonality).\todo{Should add model specification formula here.}
-
-The below shows our robustness check for all cause absenteeism.
 
@@
   out$Variable <- gsub('time_period', '', out$Variable)
-  return(kable(out, format = "latex") %>%
+  return(kable(out, format = "html") %>%
   kable_styling() %>%

### 2018-05-11 inst/rmd/paper.Rmd
removed text

--- 2018-05-07 inst/rmd/paper.Rmd
@@
 ---
-title: Is malaria control profitable? Return on investment of privately-managed residential fumigations at a large sugarcane processing facility in Southern Mozambique
+title: Is malaria control profitable? Return on investment of privately-managed residential
+  fumigations at a large sugarcane processing facility in Southern Mozambique
+csl: apa_modified.csl
 date: ''
 output:
+  html_document:
+    df_print: paged
   pdf_document:
-    keep_tex: true
-    latex_engine: xelatex
     includes:
       in_header: preamble.sty
+    keep_tex: yes
+    latex_engine: xelatex
+  word_document: default
 fig_width: 4
@@
                warning=FALSE,
-               dev = "cairo_pdf",
+               # dev = "cairo_pdf",
                fig.pos="!h",

### 2018-05-11 inst/rmd/paper.tex
removed text

@@
 
-Malaria accounts for a half million annual deaths worldwide (White et
-al., 2014) (WHO, 2016) (Ashley et al., 2018). Though rapid improvements
-in technology and funding have lead to important reductions in
-marlaria's global burden, the scale-up in activities required for
-eradication (``the worldwide interruption of transmission'') (Lancet,
-2011) will mean new partnerships and actors. One promising - albeit
-atypical - potential stakeholder in global malaria eradication is the
-private sector, given its omnipresence and potential to benefit directly
-from the elimination of malaria. But little evidence exists
-demonstrating how private sector entities can engage with malaria
-control and benefit at the firm level.
-
-At the societal level, malaria has a large economic impact. By affecting
-saving, investment (Shretta et al., 2016), risk perception,
-productivity, absenteeism (Nonvignon et al., 2016), human capital
-accumulation (Castel-Branco, 2014), mortality, and costs of care (Sachs
-and Malaney, 2002), malaria likely has a negative effect on GDP and
-growth (McCarthy et al., 2000a) (Orem et al., 2012). Because of the
-relative affordability of most intereventions and the enormous societal
-costs of malaria, most forms of malaria control are cost-effective when
-a public welfare perspective is assumed, such as when a government
-provides the financing (White et al., 2011) (Purdy et al., 2013) (Howard
-et al., 2017).
-
-From the persepctive of the private sector, however, investing in
-malaria control is not so clear-cut, since the benefits are often
-disperse, long-term, and difficult to quantify. Public health
-interventions targetting malaria - and their corresponding
-cost-effectiveness evaluations - most often focus on impacts pertaining
-to public welfare, such as an increase in life years adjusted for
-disability or quality (Goodman et al., 1999) (Shretta et al., 2016) (Lee
-et al., 2017) (Hanson, 2004). Though population-level health is
-certainly of importance to businesses, and improvements in health
-incidentally improve the economy at all levels (Brundtland, 1999) (Bloom
-and Canning, 2008) (Vecchi et al., 2013), these improvements may be too
-disperse or long-term to incentivize private sector involvement in
-health campaigns. In other words, the returns for malaria control are
-less for the private sector because (i) they capture only part of the
-benefits and (ii) they do not benefit from the externalities.
-
-Just as the benefits of malaria control to the private sector are more
-limited than those to the public sector, considerations regarding costs
-for a firm are also distinct than those for a government. Though many
-firms in endemic regions engage in malaria control programs, this should
-not be considered, per se, evidence of its cost-effectiveness (since the
-extent to which corporate social responsibility plays a role is
-unknown).
-
-100\% of the Mozambican population are at risk of malaria, living in
-what the WHO classifies as a ``high transmission'' area (Moonasar et
-al., 2016). Annually, Mozambique has more than 8 million clinical
-malaria cases (an annual incidence of approximately 300 per 1,000
-residents), with an estimated 14,000 deaths. Malaria accounts for 29\%
-of all deaths, and 42\% of deaths among those under five years of age
-(INE, 2011). Since 2013, Mozambique has seen a gradual increase in the
-incidence of malaria (Moonasar et al., 2016). 100\% of the malaria in
-Mozambique is of the Plasmodium falciparum species, with Anopheles
-funestus, gambiae, and arabiensis as the primary mosquito vectors of the
-disease (WHO, 2015).
-
-A significant sector of the economy in Mozambique is dominated by
-large-scale foreign direct investment projects (Robbins and Perkins,
-2012), and the role of the private sector in health generally, and
-malaria specifically, is unequivocally important. Large agriculture and
-extractive industry firms take up wide swaths of land and employ
-hundreds of thousands (German et al., 2013). The Mozambican state has
-encouraged large-scale entreprise with the aim of general economic
-development (Buur et al., 2012). And where large firms exist, they often
-take on social roles such as housing and health care (Winkler, 2013). At
-times, this role is necessary from a purely practical standpoint; in
-other cases, it is employed under the auspices of ``corporate social
-responsibility'' (Azemar and Desbordes, 2009) (Curtis et al., 2003).
-Regardless of the language used, it is clear that private industry plays
-an important role in public health in Mozambique (Robbins and Perkins,
-2012) (Castel-Branco, 2014).
-
-To address the question of the profitability of malaria control
-activities from the standpoint of a private firm, we analyze data during
-a 4 year period from a private sugar facility in Southern Mozambique. We
-use absenteeism as our outcome variable, assuming that it is directly
-correlated with the productivity losses associated with malaria
-infection. We assess the effect of indoor residual spraying (IRS) on
-absenteeism, and demonstrate that the firm's engagement in malaria
-control not only improved worker health, but also generated a positive
-return on investment from a pure accounting perspective.
-
-The structure of the paper is as follows. We provide an overview of the
-sugar company under study, and the epidemiology of malaria in the nearby
-area, as well as in Mozambique as a whole. We then give an overview of
-the data collected, and outline the theoretical and methdological
-assumptions and tools that underly our analysis. We assess the effect of
-time since IRS on worker absenteeism, controlling for malaria
-seasonality, and segregating models for four different worker types. In
-the results, we show that IRS spraying is associated with a significant
-reduction in worker absenteeism among permanent and fieldworkers, and
-has little to no effect on temporary and indoor workers. We find that,
-in addition to reducing absence, the IRS program has a cost savings
-effect. Our discussion covers potential implications from this study in
-terms of policy and investment, as well as the paper's limitations.
-
-\subsection{Background literature}\label{background-literature}
-
-The evidence of malaria's negative effects on both health and wealth, at
-both the individual (Cole and Neumayer, 2006) and collective (McCarthy
-et al., 2000b) levels, as well as in both the short (Asenso-Okyere and
-Dzator, 1997) (Ajani and Ashagidigbi, 2010) and long (Hong, 2011) (Sachs
-and Malaney, 2002) terms, are amply described in the public health and
-economics literature (Phillips, 1998). That said, very little exists in
-the literature examining the costs and benefits of malaria control from
-a private ledger perspective (ie, the point of view a business
-investor). Unlike a government or individual, a private firm investing
-in malaria control may be most interested not in its long-term
-macroeconomic effects, nor its short-term personal health effects, but
-rather on the impact on productivity (and the extent to which that
-producitivty's benefits accrue to the firm), absenteeism, and the
-opportunity costs of expenditures in malaria control. In other words,
-though the large magnitude of malaria control's benefits are well known,
-the portion of those benefits accrued by a private firm investing in
-malaria control is unknown.
-
-In general, large firms operating in malaria endemic regions consider
-malaria to be an important enough issue to merit at least some
-investment (Pluess et al., 2009). Several studies examine the effect of
-foreign firms engaging in large-scale malaria control campaigns (Han,
-2015) (Bennett et al., 2017) (Kaula et al., 2017). AngloGold-Ashanti, in
-partnership with local and national government in Ghana, invested in a
-well-rounded malaria control program in 2005, and saw worker absenteeism
-fall by 50\% in 13 months (CCM, 2016). Lafarge's simulatenous investment
-in a comprehensive malaria control program in Benin was associated with
-an average 41\% reduction in absenteeism among workers over the course
-of 4 years (Egedeye et al., 2011). Zambia Sugar Plc, Zambia's largest
-sugar processing facility, saw annual malaria cases at its company
-clinic fall from nearly 3,000 in 2001 to less than 500 by 2005,
-following investment in a malaria control program. Marathon Oil's
-investment of 15 million US in vector control, education, net
-distribution and malaria treatment on Bioko Island in Equatorial Guinea
-lead to an estimated 95\% reduction in the number of parasite-infected
-mosquitoes and 50\% reduction in malaria incidence among young children
-(Asquino, 2016) (Overgaard et al., 2012). A PATH study in Zambia found a
-return on investment of 28\% among three companies investing in
-employer-based malaria control (Mouzin and al., 2011).
-\todo{Still need to include stuff from http://gbchealth.org/wp-content/uploads/2014/03/IRS2011WorkshopReport.pdf}
-
-Though certainly suggestive of high returns on malaria control
-investments, these studies generally consider population health as the
-outcome measure of interest, rather than worker absenteeism or
-productivity. Similarly, they often neglect to differentiate between
-those clinical costs which are absorbed by the local health system
-versus those which are absorbed by the firm itself. When absenteeism
-itself is considered, effects of malaria control have generally been
-found to be high, but causation is difficult to establish, given that
-the previous studies rely on aggregate data.
-
-Two studies utilize worker-level data to estimate the effect of malaria
-control on productivity. A World Bank analysis of Nigerian sugarcane
-cutters found that the simple availability of testing and treatment
-increased productivity by 10\% in the weeks following the provision of
-services, the conclusion being that both the treatment and the test
-result were effective in increasing productivity, the latter simply
-increasing the information available which could influence personal
-labor allocation decisions (Dillon et al., 2014). A randomized
-controlled trial (RCT) in Zambia showed an even greater effect from
-investments in preventing malaria: farmers given bed nets saw fewer days
-lost to illness (both directly and due to caretaking responsibilities
-for ill family members), translating to an increase of approximately
-15\% in crop yield (Fink and Masiye, 2015). Though compelling, the
-Nigerian program only dealt with medical services (diagnostics and
-medication), rather than preventive interventions, and the Zambian RCT
-examined individual farmers, as opposed to a large firm.
-
-In the literature, making the ``investment case'' for malaria control or
-elimination generally implies that the investor is the public sector,
-and takes into account those costs and benefits which are applicable
-from a public welfare point of view (Shretta et al., 2017). For example,
-an economic analysis by the Corporate Alliance on Malaria in Africa on
-the Bioko Island Malaria Control Program found a 4:1 cost-benefit ratio,
-but the perspective in this case was considered to be the ``community''
-(Egedeye et al., 2011). Though appropriate in most cases to consider
-benefits accrued to the community (the government or institutions
-interested in public welfare primarily being the primary malaria control
-agents in most locations), the findings of these studies are rarely
-applicable to the private sector, and even less so at granular levels
-(such as an individual firm). In the case of a private firm not
-interested in ``corporate social responsibility'', it is not clear
-whether investing in malaria control would be profitable or not. This
-lack of clarity not only discourages investment, but also makes it
-difficult for governments to pinpoint the correct of amount of subsidy
-(if applicable) to encourage private sector scale-up in malaria control.
-
-The literature on the effect of sugarcane cultivation on malaria risk is
-mixed. While some studies have found that the prevalence of malaria
-vectors in sugarcane areas to be similar to that of uncultivated areas
-(and less than in areas dedicated to other forms of more water-inensive
-agriculture, such as rice) (Ijumba et al., 2002), other studies have
-found significant increases in factors associated with malaria
-transmission at large-scale sugarcane facilities relative to
-traditional, small-scale farming and non-irrigated farming (Jaleta et
-al., 2013). Regardless of the effect the presence of a sugarcane farm
-\emph{per se} on local malaria epidemiology, the time spent outdoors by
-sugarcane workers, the fact that many workers are migrants, and their
-sometimes precarious housing situations, suggest that sugarcane farmers
-are likely at increased risk of malaria infection (O'Laughlin, 2016).
-This is important, given that even among occupations with far less
-inherent exposure to mosquitoes (such as health professionals), malaria
-is one of the primary causes of work absenteeism in malaria-endemic
-countries (Burton et al., 1999). There is also some concern regarding
-the effect of large-scale insecticide use - common at essentially all
-Sub-Saharan African sugarcane farms - on insecticide resistance among
-moquitoes in the area. A study in Belize found that mosquito populations
-on the edge of sugarcane fields had higher tolerance to insecticide than
-similar populations in the core of fields or outside of the periphery
-(Dusfour et al., 2009). Sugarcane areas may offer the standing water
-necessary for mosquito breeding, but also perhaps attract mosquitoes
-which would otherwise be elsewhere, due to compounds in sugarcane pollen
-(Wondwosen et al., 2018).
-
-This study does not endeavor to expand the current body of knowledge
-regarding malaria's ill effects on individuals and societies; rather, it
-aims to provide empirical evidence pertaining to a facet of malaria
-economics with very little in the literature: malaria control from a
-private-sector investment perspective. Our study adds to the existing
-literature by showing the effect of specific malaria control
-interventions on worker absenteeism, and translating that effect into a
-return on investment. Unlike previous studies on the effectiveness of
-malaria control interventions, our's focuses solely on one firm carrying
-out one intervention, takes advantage of individual-level data, and
-analyzes results from a ledger perspective.
-
-\subsection{Study area}\label{study-area}
-
-Sugar has been systematically cultivated in Mozambique since the late
-1800s. The Incomati Estates company, a small sugarcane processing
-facility started by a Scotsman on the banks of the Incomati River in
-1913, was the first firm to export sugar from Mozambique. Following its
-purchase by international investors in the 1950s, it (along with the
-rest of the industry) expanded significantly, exporting to both Europe
-and the United States. In the late 1960s, a Portuguese family opened the
-Maragra Açúcar company, while a group of foreign investors started the
-nearby Marracuene Agrícola Açucareira mill. By the early 1970s, sugar
-grew to account for greater than 10\% of Mozambique's national exports.
-Nationalized following independence in 1977, the industry's production
-levels fell from 320,000 annual tons to fewer than 15,000 by 1992. After
-the end of the civil war, foreign investment revived the sugar industry,
-and by 2011 production had surpassed its 1972 peak.
-
 \begin{figure}[!h]
@@
 \end{figure}
-
-The mill of the Maragra Açucar SA (a subsidary of the Illovo sugar
-company, henceforth referred to as ``Maragra'') was nationalized in the
-1970s (like all other Mozambican mills), went through a period of low
-production, and then fell completely out of use by 1984. It re-opened in
-private hands in 1992, and was renovated by a group of international
-investors in 1998. Today, Maragra accounts for roughly one quarter of
-Mozambique's overall sugar production (second only to the nearby
-Xinavane mill run by the Hulett Sugar Tongaat company) (Sutton, 2014).
-With a favorably close location to the port of Maputo, ample land
-(approximately 90 squared kilometers of plantation, and 5 squared
-kilometers of factory grounds), approximately 5,000 employees (of which
-three quarters are seasonal), and a mill with the capacity to process
-not only all the sugarcane grown on Maragra's land, but also the cane of
-the many nearby smallholders (O'Laughlin, 2016), Maragra has so far been
-able to weather the 2016 Mozambican crisis and concurrent collapse in
-global sugar prices.
-
-Maragra (figure 1, panel iii) is located in the district of Manhiça
-(figure 1, panel ii), a semi-rural area in the south of Mozambique
-(figure 1, panel i). 80 kilometers north of the Mozambican capital of
-Maputo, the district is low-lying, consisting largely of savannah and
-wetlands along the Incomati River. Most of the areas 160,000 residents
-(Sacoor et al., 2013) work as subsistence farmers. Migration from the
-area to South Africa for the purpose of employment in the profitable
-construction industry is common, especially among men (Nhacolo et al.,
-2006), as is migration to the area (from other parts of Mozambique) for
-seasonal work on the sugarcane plantations at Maragra and the slightly
-larger facility in Xinavane (at the district's border with Magude)
-(O'Laughlin, 2016).
-
-Poverty is rife in southern Mozambique, and its associated illnesses
-take their toll on the population. The community prevalence of HIV/AIDS
-is as high as 40\% (González et al., 2012); even in a more recent study
-suggesting a much lower prevalence of 22\%, the risk of infection is
-still twice that of nearby areas (Mocumbi et al., 2017). Recent years
-have seen a three-fold increase in tuberculosis (García-Basteiro et al.,
-2017). Malaria, which has the greatest mortality burden due to the fact
-that the young are especially vulnerable to its effects, is perennial,
-though worse during the rainy season (December - March) (Saúte et al.,
-2003). Adult malaria is essentially universal, albeit much of it
-asymptomatic (Mayor et al., 2007). In regards to worker health, malaria
-is Maragra's primary concern, being so important as to justify the
-existence of both an on-site testing laboratory and clinic, as well as a
-malaria control department.
-
-Maragra workers are mostly seasonal, working for the firm approximately
-half of the year during harvest time, and cultivating crops, working in
-construction, or going unemployed (or working elsewhere) during the
-other half. Though many workers live ``on-site'' (ie, within the
-delineated property of the firm itself), a sizeable minority resides in
-the environs (figure 1, panel iv). Maragra provides indoor residual
-spraying (IRS) using ACT (alpha-cypermethrin) and DDT
-(Dichlorodiphenyltrichloroethane) depending on stocks (a preference
-apparently exists for the former, but it is not always available). IRS
-activities, managed by Maragra's Malaria Control program, are ongoing
-throughout the year (Figure 1, panel i). Nearly all on-site houses are
-sprayed, though the time between fumigations is irregular (Figure 2,
-panel ii). Off-site houses may also receive IRS (managed and carried out
-by the National Malaria Control Program), but the status and timing of
-these fumigations is not known at the individual level.
 
@@
 
-\subsection{Data}\label{data}
-
-In collaboration with the sugar processing facility, we collected data
-for the period from January 2010 through December 2016. Data came from
-four sources: (i) the Human Resources' roster of worker details and
-absences, (ii) the facility's on-site clinic's medical and laboratory
-records, (iii) the facility's on-site malaria control program's records
-pertaining to the dates, chemicals, and location of IRS activities, and
-(iv) interviews with company employees pertaining to costs, data
-limitations, etc. Digitization and collection of data took place during
-the period from March 2016 through May 2017. Supplementary data
-pertaining to worker characteristics was obtained from through the
-Centro de Investigação em Saude de Manhiça's (CISM) demographic census,
-which covered workers from the district, but not those who migrated from
-other parts of the country (Nhacolo et al., 2006).
-
-Data pertaining to district-wide malaria incidence was obtained from
-Mozambique's Boletim Epidemiológico Semanal (BES), which is the system
-by which the National Malaria Control Program monitors incidence at the
-district level throghout the entire country, and reports the number of
-confirmed weekly malaria cases at government health facilities. Using
-these case numbers, combined with population estimates from the National
-Statistical Institute (INE), we estimate each day's annualized weekly
-malaria incidence rate (cases per 1000 population at risk), interpolated
-from the weekly figures. We retrieved weather data for all Mozambican
-stations from NOAA. We estimated the meteorological conditions at the
-centroid of Manhiça using a simple interpolation method whereby the
-district's weather conditions were estimated to be a function of all
-Mozambican weather stations' reported conditions, inversely weighted by
-kilometers from district centroid.
-
-Maragra regularly employs IRS at on-site worker households in order to
-reduce those workers' (and their families') risk of malaria infection.
-IRS works by killing the malaria vector (mosquito), thereby preventing
-infection of the household occupants. When administered correctly, IRS
-is a low-risk intervention to its recipients, and is assumed not to
-affect absenteeism in the short-term (to the extent that it may cause
-negative side-effects, these are generally long-term). It is preventive
-only, and does not cure current malaria infection, nor does it affect
-the parasite load of mosquitoes. Workers living off-site (our control
-group) also may have received IRS at some point during the study period
-(from government programs). Even though we do not have reliable
-person-level data on IRS carried out by the government, off-site workers
-are a suitable control in the sense that they represent ``business as
-usual'' (ie, what would happen if the company carried out no IRS and
-relied solely on public interventions). Using company HR and clinical
-records, we were able to identify absences and episodes of clinical
-malaria among all workers, as well as identify the time since the most
-recent IRS episode before the onsent of absence or illness.
-
-Worker characteristics, illness and absenteeism data, along with IRS
-activity data, were systematically stored, collected, and used at the
-individual level by Maragra, and therefore of generally high quality.
-Because cost data was less systematically collected by Maragra, and
-because many costs could not be precisely quantified due to the
-abundance of in-kind and cross-departmental expenditures, we had to rely
-on rough estimations based on a mix of interviews, reciepts, and
-interpolations. Since our program cost data is not as reliable as our
-worker characteristic and outcome data, we were conservative in our
-estimates, and generally tried to err on the side of program activities
-and materials costing \emph{more} than what was reported, when doubt was
-aired. Cost data consisted of three types: (i) wages of malaria control
-employees, (ii) transporation and vehicle costs for IRS teams, and (iii)
-acquisition costs of purchasing IRS chemicals for fumigation (ACT and
-DDT), the latter two being combined into malaria control ``programme''
-costs.
-
-\subsection{Conceptual framework and identification
-strategy}\label{conceptual-framework-and-identification-strategy}
-
-We sought to understand the effect of IRS on individual workers'
-likelihood of absence from work as well as their likelihood of clinical
-malaria. To estimate this effect, we estimated separate models for
-absence and illness. We employed linear mixed effects models (Minalu et
-al., 2011) (Liu et al., 2010) (with individual worker fixed effects, and
-random effects for other covariates) so as to explicitly adjust for
-confounders (Bell and Jones, 2014). We divide into 4 different models
-for different worker types, so as to account for the potential
-time-varying effects of worker type on risk of malaria.
-
-\subsubsection{Externalities}\label{externalities}
-
-The administration of insecticide at a worker's residence (indoor
-residual spraying or IRS) likely protects that worker from malaria
-infection by killing the vectors (mosquitoes) that land on the
-building's walls. However, it is also highly likely that the protective
-effective of IRS ``spills over'' to others who live nearby. This
-positive spillover effect would theoretically go through two channels:
-(i) via a reduction of mosquitoes in the vicinity and (ii) via a
-reduction of the malaria parasite in the blood of humans in the vicinity
-(ie, the parasite ``reservoir''). This document describes our method for
-assessing the existence and magnitude of positive spillover of IRS in
-the Maragra workers' data. We devise a time-specific household
-``protection'' score based on the theoretical effectiveness of IRS, and
-then use that protection score to develop a time-place specific ``herd
-protection'' score based on a weighted average of nearby household
-protection scores.
-
-We consider a house's ``herd'' protection level (ie, the protection
-conferred to the house through externality) to be the average of the
-other houses' non-herd protection levels, weighted by the distance to
-the house in question. The below is an illustration of a house (the red
-x) and the neighboring houses (circles). The ``weight'' of each house is
-indicated by the circle size, and the protection level of the house is
-indicated by the shading of the circle.
-
-However, this consideration misses one important dimension: density. The
-below house's weighted average protection score is identical to that of
-the above house. In other words, at the time in question, at both
-locations, the percentage of houses with IRS coverage is the same, and
-the (weighted) average time since IRS is the same. However, the below
-house is likely much more protected than the above, given the
-\emph{absolute} number of nearby IRS-protected houses.
-
-Our approach is this: rather than using a weighted average of each
-household's individual protection score, we instead use a weighted sum.
-A particular household's herd protection score at any given time is the
-sum of all households' individual protection scores at that time
-multiplied by those households' distance weights.
-
-The summation of (a) the IRS level of other houses at a certain time,
-(b) weighted by the distance of those houses to the house in question,
-(c) limited to only the 1 km radius multiplied yields a ``herd
-protection score''.
-
-This score is conceptually similar to Cohen and Dupas' quantification of
-the positive externalities of ITN (bednet) use in Kenya (Cohen and
-Dupas, 2010) in that it attempts to estimate the protection conferred to
-``non-users'' by users and uses a ``weighted sum'' (as opposed to
-average). Our approach differs in that the time dimension for IRS is
-much more important than ITN (ie, IRS is not a binary but rather a
-waning function, as described previously), requiring us to create herd
-scores for every day.
-
-Our approach can be justified in that previous studies have found strong
-positive health externalities in malaria interventions related to ITN
-coverage (Alaii et al., 2003). No studies exist on positive
-externalities for IRS coverage, but to the extent that the mechanisms
-for the reduction in infection are similar (reduction in the natural
-reservoir of the disease, reduction in the number of vectors, etc.), it
-is reasonable to assume similar effects. Additionally, we are using
-weighted distance for our score calculation (rather than simply defining
-a radius threshold) because of previous studies' findings that there is
-a linear decline in protection conferred to others with greater distance
-from an intervention (Binka et al., 2007).
-
-Figure x shows a smooth interpolated surface of herd protection,
-averaged for the entire study period.
-
 \begin{figure}[!h]
@@
 \end{figure}
-
-Some more details will eventually go here.
 
@@
 
-We define the malaria season as any time during which the clinical
-incidence of malaria in the district of Manhiça was at or greater than
-the median clinical incidence of malaria for the entire study period.
-These weeks are flagged as red in Figure 1, Panel A. By using clinical
-incidence of the area of residence of the workers (as opposed to more
-typical proxies for malaria risk, such as only rainy vs.~non rainy
-season), our seasonality estimate is a closer approximation of true
-malaria risk, incorporating lagged effects such as the incubation period
-of the parasite, as well as any inherent non-linear effects of weather.
-In addition, we adjust for daily precipitation; though its lagged effect
-on malaria incidence is likely captured by the seasonality term, we
-include rainfall since it's immediate effect (through its impact on
-transportation and working conditions) may also affect absenteeism.
-
 Our formula for return on investment can be described in a
@@
 
-\subsection{Reproducibility and ethical
-approval}\label{reproducibility-and-ethical-approval}
-
-All data processing and analysis were carried out in R (R Core Team,
-2017) and all analysis code is freely available online (Brew, 2017).
-Ethical approval for this project was obtained from the Institutional
-Ethics Review Board for Health at the CISM (CIBS-CISM) prior to data
-collection.
-
 \newpage
@@
 We collected absenteeism and demographic data data on 3362 workers from
-2012 through 2016. Workers were approximately 60\% male, and more than
-80\% fieldworkers (predominantly cane-cutters). Most were in their 20s
-and 30s and employed on temporary contracts. Table 1 shows overall
-worker details divided by ``Treatment'' versus ``Control'' status.
-``Treatment'' is considered the time beginning at IRS administration and
-ending six months later; ``Control'' is considered time observed either
-prior to IRS administration (and greater than 6 months after previous
-IRS administration) or among workers who never received IRS. Because of
-the temporal nature of these categories, many workers belong to both
-treatment and control groups, albeit at different times.
+2012 through 2016.
 
@@
 
-Temporary workers tended to be younger and male; female temporary
-workers being on average 5-10 years older than their male counterparts.
-Permanent workers had a bi-modal age distribution, the older peak
-explained by the greater density of males in administrative roles.
-Females accounted for 43\% of temporary workers but only 16\% of
-permanent workers (figure x).
-
 \begin{figure}[!h]
@@
 \end{figure}
-
-During the study period, weather followed typical seasonal trends for
-the area, albeit slightly drier than previous periods.
 
@@
 \end{figure}
-
-In Southern Mozambique, malaria peaks during the summer months (December
-through March) most years (Figure 5, panel A), and worker absenteeism at
-Maragra rates track malaria incidence closely, following the same
-seasonal patterns (figure 4, panel B). Both all-cause absenteeism and
-sick absenteeism have declined in recent years at Maragra (figure 5,
-panel C), with the latter declining at a faster rate than the former.
-The fact that the rate of confirmed cases at the company clinic is
-largely non-seasonal (figure 5, panels E and F) suggests that a
-significant portion of workers either seek care for malaria elsewhere
-(for example, government health posts, of which several are nearby and
-in some many cases closer to workers' residence than the company clinic)
-or do not seek care during malaria infection. Accordingly, we focus our
-analysis on all-cause absenteeism rather than sickness absenteeism or
-malaria diagnostics, with the assumption that much of illness is
-captured by absenteeism but not by the clinical data.
 
@@
 
-The extent to which rainfall's effect on absenteeism is confounded by
-its effect on malaria is discussed in the model results section.
-
-\textbf{Costs of program}: The malaria control program at Maragra has an
-annual operating budget of approximately \$112,000, which includes the
-purchase of insecticide, the wages of IRS sprayers and drivers,
-transportation, record-keeping, and general administrative costs.
-Assuming linearity in costs, the program spends approximately \$19 per
-building sprayed. With each agregado containing an average of 2.2
-workers. Much of the benefit of IRS goes to non-worker residents of
-sprayed agregados (who constitute a majority), but this benefit is
-purposefully ignored for this analysis.
-
-\textbf{Cost of malaria}: Given the likelihood that clinical data does
-not fully capture all malaria cases (and most likely captures only a
-small fraction of actual infections, given the high rates of acumulated
-adult immunity {[}Mayor2007{]}), we do not quantify the costs of malaria
-infection to the company. Rather, we first estimate the reduction in
-absenteeism attributible to IRS, and then quantify the savings
-associated with prevented absences. Additionally, we calculate the
-clinical savings of IRS by first estimating the share of absences which
-are associated with an episode of clinical malaria, and then applying
-the clinical cost per case to the equivalent share of prevented
-absences. We intentionally ignore the savings accrued by the public
-health system, as well as the likely utility gains in secondary realms
-such as school absenteeism, producitivity, etc.
-
-\subsection{Analysis}\label{analysis}
-
-\subsubsection{Effect of IRS on
-absenteeism}\label{effect-of-irs-on-absenteeism}
-
-Immediately following IRS, a worker's likelihood of absence drops
-significantly (appendix figure, panel A). As one would expect if the
-mechanism by which IRS reduces absence is through reduced malaria
-infection, the effect of IRS during the low transmission season is
-significant, but far less substantial in effect size (appendix figure,
-panel B).
-
-In order to assess the effect of IRS on absenteeism, we created 4 worker
-fixed effects models for the 4 principal worker types (permanent field
-worker, temporary field worker, permanent non-field worker, temporary
-non-field worker) for both outcoms (all cause absenteeism vs.~sickness
-absenteeism). The reason for segregating models rather than
-incorporating worker type as a fixed effect was that it seemed plausible
-that the effect of the treatment (IRS) on the outcome would be different
-(and vary differently over time) depending on these worker types. For
-instance, it is reasonable to assume that IRS' effect would be
-differential for a temporary worker (who likely lives in the fumigated
-house less than 100\% of the year) than for a permanent worker.
-
-Given that IRS is carried out on a continuous rolling basis, we used
-relative worker time (as opposed to calendar) as the time component of
-our model - that is, each workers' days until and since IRS were
-standardized so that day 0 (date of fumigation) was the interruption
-point. This has the advantage of incorporating into the model a variety
-of potential time-varying confounders in a quasi-randomized fashion,
-thereby making it unecessary to adjust for them specifically. Our model
-took into account workers' residential location (on or off site) - even
-though off site workers by definition could not be administered IRS by
-Maragra - so as to capture other seasonal effects from those workers.
-
-The herd protection score, unit-less by nature given that it's a
-summation of binary protection statuses weighted by inverse-transformed
-distances, took the following distribution in our data:
-
 \begin{figure}[!h]
@@
 
-The model results for all cause absenteeism (table 4) suggest a
-significant decrease in absenteeism following the administration of IRS.
-The effect is ambivalent among temporary workers, whereas among
-permanent workers the administration of IRS decreases absenteeism by
-xxx. These effects are consistent with IRS decreasing absence through
-the medium of malaria, since (a) the effect is not as strong in the
-first month (potentially during the parasite incubation phase), the
-effect is stronger on field workers (who are more socioeconomically the
-demographic at greater risk of malaria than their administrative
-counterparts), and the effect is ambivalent among temporary workers
-(since the protective effect of IRS is only a function of the amount of
-time that the worker spends in the fumigated house).
+\newpage
 
@@
 
-\includegraphics{figures/unnamed-chunk-31-1.pdf}
-
-\subsection{Translating absences to
-costs}\label{translating-absences-to-costs}
-
-xxx\todo{Not done yet. This will be relatively straightforward arithmetic, but I want to first make sure that the models are rock solid (since all of the "translation" hinges on them).}
-
-\subsection{Return on investment}\label{return-on-investment}
-
-\todo{Not done yet. In conversation with Eduardo from MC about getting more accurate costs}
-
-\textbf{Savings}
-
-\begin{itemize}
-\tightlist
-\item
-  In percentage point terms, reduction from 13\% to 8\%.\\
-\item
-  5 annually prevented absences per worker.\\
-\item
-  8,000 workers: 40,000 prevented absences, wage of 3 USD
-\item
-  TOTAL: 120,000 USD in productivity-only savings.
-\end{itemize}
-
-\textbf{Costs}
-
-\begin{itemize}
-\tightlist
-\item
-  8 IRS workers, 1500 USD yearly = 12,000 USD\\
-\item
-  ACT + DDT: 50,000 USD
-\item
-  Facilities, vehicles, gas: 50,000 USD\\
-\item
-  TOTAL: 112,000 USD in IRS-only costs
-\end{itemize}
-
-\textbf{7\% ROI} (ignoring clinical costs)
+\includegraphics{figures/unnamed-chunk-30-1.pdf}
 
@@
 generalizability}\label{robustness-and-generalizability}
-
-Two principal concerns call into question the results of our analysis.
-First, the application of IRS to a workers house may be
-endogenous.\todo{Is this the right term here?
-We're not dealing with an "unobserved" factor per se, rather we're dealing with a situation in which the cause might influence the effect, and the effect might influence the cause, ie a feedback loop.}
-It is reasonable to suspect that the application of IRS to households is
-not random, but rather that IRS was applied more frequently to houses
-which had already seen a malaria case. In this case, our estimated
-effect of IRS on absenteeism would likely be underestimated, with the
-post-IRS absenteeism rates actually having declined from a greater
-pre-IRS absenteeism rate than otherwise suggested.
-
-To check for this, we estimate the odds of absenteeism as a function of
-receiving IRS \emph{during the 10 day period prior to receiving IRS}. If
-IRS application is indeed endogenous, we would expect absenteeism to be
-elevated during this period (since the increase in absenteeism would be
-theoretically responsible for the application of IRS), a situation which
-would require further statistical adjustment. If, on the other hand,
-there is no endogeneity, we would expect absenteeism in the 10 day
-period prior to IRS administration to be similar to other pre-IRS
-absenteeism (adjusting for
-seasonality).\todo{Should add model specification formula here.}
-
-The below shows our robustness check for all cause absenteeism.
 
@@
 
-\todo{This entire section will not be done until methods/results are finalized.}
-
-\begin{itemize}
-\tightlist
-\item
-  Overview of findings
-\item
-  How this contributes to the literature
-\item
-  Implications for policy and for businesses
-\item
-  Spillover effects !
-\item
-  How absenteeism might be \emph{underestimating} true effect, since at
-  the margin a sick worker might go to work (ie, not be absent), but be
-  less productive
-\end{itemize}
-
-\subsection{Limitations}\label{limitations}
-
-\begin{itemize}
-\tightlist
-\item
-  No analysis yet of different worker types (agricultural
-  vs.~industrial).
-\item
-  Have not yet ventured at all into side-analyses (effect on employment,
-  tonnage, etc.).
-\item
-  Sick absenteeism seems to track absenteeism poorly: lack of clarity
-  regarding pathways.
-\item
-  Large sample size, but all from same place: questionable
-  generalizability.
-\end{itemize}
-
 \newpage
@@
 \centering
-\includegraphics{figures/unnamed-chunk-35-1.pdf}
+\includegraphics{figures/unnamed-chunk-34-1.pdf}
 \caption{i. Absenteeism before and after IRS administration for all
@@
 \centering
-\includegraphics{figures/unnamed-chunk-36-1.pdf}
+\includegraphics{figures/unnamed-chunk-35-1.pdf}
 \caption{i. Sick absenteeism before and after IRS administration for all
@@
 
-\section*{References}\label{references}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\hypertarget{ref-Ajani2010}{}
-Ajani, O., Ashagidigbi, W., 2010. Effect of malaria on rural households'
-farm income in oyo state, nigeria. African Journal of Biomedical
-Research 11. \url{https://doi.org/10.4314/ajbr.v11i3.50723}
-
-\hypertarget{ref-Alaii2003FactorsAU}{}
-Alaii, J.A., Hawley, W.A., Kolczak, M.S., Kuile, F.O.T., Gimnig, J.E.,
-Vulule, J., Odhacha, A., Oloo, A.J., Nahlen, B.L., Phillips-Howard,
-P.A., 2003. Factors affecting use of permethrin-treated bed nets during
-a randomized controlled trial in western kenya. The American journal of
-tropical medicine and hygiene 68 4 Suppl, 137--41.
-
-\hypertarget{ref-AsensoOkyere1997}{}
-Asenso-Okyere, W., Dzator, J.A., 1997. Household cost of seeking malaria
-care. a retrospective study of two districts in ghana. Social Science \&
-Medicine 45, 659--667.
-\url{https://doi.org/10.1016/s0277-9536(96)00383-8}
-
-\hypertarget{ref-Ashley2018}{}
-Ashley, E.A., Phyo, A.P., Woodrow, C.J., 2018. Malaria. The Lancet.
-\url{https://doi.org/10.1016/s0140-6736(18)30324-6}
-
-\hypertarget{ref-asquino2015}{}
-Asquino, M., 2016. Equatorial Guinea Plays a Leading Role in Combating
-Malaria.
-
-\hypertarget{ref-Azemar2009}{}
-Azemar, C., Desbordes, R., 2009. Public governance, health and foreign
-direct investment in sub-saharan africa. Journal of African Economies
-18, 667--709. \url{https://doi.org/10.1093/jae/ejn028}
-
-\hypertarget{ref-Bell2014}{}
-Bell, A., Jones, K., 2014. Explaining fixed effects: Random effects
-modeling of time-series cross-sectional and panel data. Political
-Science Research and Methods 3, 133--153.
-\url{https://doi.org/10.1017/psrm.2014.7}
-
-\hypertarget{ref-Bennett_2017}{}
-Bennett, A., Avanceña, A.L.V., Wegbreit, J., Cotter, C., Roberts, K.,
-Gosling, R., 2017. Engaging the private sector in malaria surveillance:
-A review of strategies and recommendations for elimination settings.
-Malaria Journal 16. \url{https://doi.org/10.1186/s12936-017-1901-1}
-
-\hypertarget{ref-Binka2007}{}
-Binka, F.N., Kubaje, A., Adjuik, M., Williams, L.A., Lengeler, C.,
-Maude, G.H., Armah, G.E., Kajihara, B., Adiamah, J.H., Smith, P.G.,
-2007. Impact of permethrin impregnated bednets on child mortality in
-kassena-nankana district, ghana: A randomized controlled trial. Tropical
-Medicine \& International Health 1, 147--154.
-\url{https://doi.org/10.1111/j.1365-3156.1996.tb00020.x}
-
-\hypertarget{ref-Bloom2008}{}
-Bloom, D., Canning, D., 2008. Population Health and Economic Growth
-1--25.
-
-\hypertarget{ref-brewgit}{}
-Brew, J., 2017. Malaria and sugar: An in-depth examination of the effect
-of malaria control activities on the health and productivity of maragra
-sugarcane factory workers. GitHub repository.
-
-\hypertarget{ref-World1999}{}
-Brundtland, G.H., 1999. WHO on Health and Economic Productivity 25,
-396--402.
-
-\hypertarget{ref-Burton1999}{}
-Burton, W.N., Conti, D.J., Chen, C.-Y., Schultz, A.B., Edington, D.W.,
-1999. The role of health risk factors and disease on worker
-productivity. Journal of Occupational \& Environmental Medicine 41,
-863--877. \url{https://doi.org/10.1097/00043764-199910000-00007}
-
-\hypertarget{ref-Buur2012}{}
-Buur, L., Tembe, C.M., Baloi, O., 2012. The white gold: The role of
-government and state in rehabilitating the sugar industry in mozambique.
-Journal of Development Studies 48, 349--362.
-\url{https://doi.org/10.1080/00220388.2011.635200}
-
-\hypertarget{ref-CastelBranco2014}{}
-Castel-Branco, C.N., 2014. Growth, capital accumulation and economic
-porosity in mozambique: Social losses, private gains. Review of African
-Political Economy 41, S26--S48.
-\url{https://doi.org/10.1080/03056244.2014.976363}
-
-\hypertarget{ref-anglo}{}
-CCM, 2016. AngloGold Ashanti Malaria Control Ltd (AGA Mal).
-
-\hypertarget{ref-Cohen2010}{}
-Cohen, J., Dupas, P., 2010. Free distribution or cost-sharing? Evidence
-from a randomized malaria prevention experiment. Quarterly Journal of
-Economics 125, 1--45. \url{https://doi.org/10.1162/qjec.2010.125.1.1}
-
-\hypertarget{ref-Cole2006}{}
-Cole, M.A., Neumayer, E., 2006. The impact of poor health on total
-factor productivity. The Journal of Development Studies 42, 918--938.
-\url{https://doi.org/10.1080/00220380600774681}
-
-\hypertarget{ref-Curtis2003}{}
-Curtis, C., Maxwell, C., Lemnge, M., Kilama, W., Steketee, R.W., Hawley,
-W.A., Bergevin, Y., Campbell, C.C., Sachs, J., Teklehaimanot, A.,
-Ochola, S., Guyatt, H., Snow, R.W., 2003. Scaling-up coverage with
-insecticide-treated nets against malaria in africa: Who should pay? The
-Lancet Infectious Diseases 3, 304--307.
-\url{https://doi.org/10.1016/s1473-3099(03)00612-1}
-
-\hypertarget{ref-Dillon2014}{}
-Dillon, A., Friedman, J., Serneels, P., 2014. Health information,
-treatment, and worker productivity: Experimental evidence from malaria
-testing and treatment among nigerian sugarcane cutters. The World Bank.
-\url{https://doi.org/10.1596/1813-9450-7120}
-
-\hypertarget{ref-Dusfour2009}{}
-Dusfour, I., Achee, N.L., Briceno, I., King, R., Grieco, J.P., 2009.
-Comparative data on the insecticide resistance of anopheles~albimanus in
-relation to agricultural practices in northern belize, CA. Journal of
-Pest Science 83, 41--46. \url{https://doi.org/10.1007/s10340-009-0268-7}
-
-\hypertarget{ref-lafarge}{}
-Egedeye, L., Drozer, S., Leiser, A.-M., 2011. Corporate Action on
-Malaria Control: Best Practices and Interventions.
-
-\hypertarget{ref-Fink2015}{}
-Fink, G., Masiye, F., 2015. Health and agricultural productivity:
-Evidence from zambia. Journal of Health Economics 42, 151--164.
-\url{https://doi.org/10.1016/j.jhealeco.2015.04.004}
-
-\hypertarget{ref-GarcaBasteiro2017}{}
-García-Basteiro, A.L., Ribeiro, R.M., Brew, J., Sacoor, C., Valencia,
-S., Bulo, H., Cobelens, F., Macete, E., 2017. Tuberculosis on the rise
-in southern mozambique (19972012). European Respiratory Journal 49,
-1601683. \url{https://doi.org/10.1183/13993003.01683-2016}
-
-\hypertarget{ref-German2013}{}
-German, L., Schoneveld, G., Mwangi, E., 2013. Contemporary processes of
-large-scale land acquisition in sub-saharan africa: Legal deficiency or
-elite capture of the rule of law? World Development 48, 1--18.
-\url{https://doi.org/10.1016/j.worlddev.2013.03.006}
-
-\hypertarget{ref-Gonzlez2012}{}
-González, R., Munguambe, K., Aponte, J., Bavo, C., Nhalungo, D., Macete,
-E., Alonso, P., Menéndez, C., Naniche, D., 2012. High HIV prevalence in
-a southern semi-rural area of mozambique: A community-based survey. HIV
-Medicine 13, 581--588.
-\url{https://doi.org/10.1111/j.1468-1293.2012.01018.x}
-
-\hypertarget{ref-Goodman1999}{}
-Goodman, C., Coleman, P., Mills, A., 1999. Cost-effectiveness of malaria
-control in sub-saharan africa. The Lancet 354, 378--385.
-\url{https://doi.org/10.1016/s0140-6736(99)02141-8}
-
-\hypertarget{ref-Han}{}
-Han, L., 2015. Malaria in Mozambique: trialling payment by results.
-
-\hypertarget{ref-Hanson2004}{}
-Hanson, K., 2004. Public and private roles in malaria control: The
-contributions of economic analysis. The American Journal of Tropical
-Medicine and Hygiene 71, 168--173.
-
-\hypertarget{ref-Hong2011}{}
-Hong, S.C., 2011. Malaria and economic productivity: A longitudinal
-analysis of the american case. The Journal of Economic History 71,
-654--671. \url{https://doi.org/10.1017/s0022050711001872}
-
-\hypertarget{ref-Howard_2017}{}
-Howard, N., Guinness, L., Rowland, M., Durrani, N., Hansen, K.S., 2017.
-Cost-effectiveness of adding indoor residual spraying to case management
-in afghan refugee settlements in northwest pakistan during a prolonged
-malaria epidemic. PLOS Neglected Tropical Diseases 11, e0005935.
-\url{https://doi.org/10.1371/journal.pntd.0005935}
-
-\hypertarget{ref-Ijumba2002}{}
-Ijumba, J.N., Mosha, F.W., Lindsay, S.W., 2002. Malaria transmission
-risk variations derived from different agricultural practices in an
-irrigated area of northern tanzania. Medical and Veterinary Entomology
-16, 28--38. \url{https://doi.org/10.1046/j.0269-283x.2002.00337.x}
-
-\hypertarget{ref-estatistica2009}{}
-INE, 2011. Demographic health survey.
-
-\hypertarget{ref-Jaleta2013}{}
-Jaleta, K.T., Hill, S.R., Seyoum, E., Balkew, M., Gebre-Michael, T.,
-Ignell, R., Tekie, H., 2013. Agro-ecosystems impact malaria prevalence:
-Large-scale irrigation drives vector population in western ethiopia.
-Malaria Journal 12, 350. \url{https://doi.org/10.1186/1475-2875-12-350}
-
-\hypertarget{ref-Kaula_2017}{}
-Kaula, H., Buyungo, P., Opigo, J., 2017. Private sector role, readiness
-and performance for malaria case management in uganda, 2015. Malaria
-Journal 16. \url{https://doi.org/10.1186/s12936-017-1824-x}
-
-\hypertarget{ref-TheLancet2011}{}
-Lancet, T., 2011. Malaria: Control vs elimination vs eradication. The
-Lancet 378, 1117. \url{https://doi.org/10.1016/s0140-6736(11)61489-x}
-
-\hypertarget{ref-Lee2017}{}
-Lee, B.Y., Zenkov, E., Chatterjee, C., Candrinho, B., Zhang, S.,
-Colborn, J., Briët, O.J.T., Brown, S.T., Mendis, C., Bartsch, S.M.,
-Viisainen, K., DePasse, J.V., Stone, N.T.B., 2017. The economic value of
-long-lasting insecticidal nets and indoor residual spraying
-implementation in mozambique. The American Journal of Tropical Medicine
-and Hygiene 96, 1430--1440. \url{https://doi.org/10.4269/ajtmh.16-0744}
-
-\hypertarget{ref-Liu2010}{}
-Liu, L., Strawderman, R.L., Cowen, M.E., Shih, Y.-C.T., 2010. A flexible
-two-part random effects model for correlated medical costs. Journal of
-Health Economics 29, 110--123.
-\url{https://doi.org/10.1016/j.jhealeco.2009.11.010}
-
-\hypertarget{ref-Mayor2007}{}
-Mayor, A., Aponte, J.J., Fogg, C., Saúte, F., Greenwood, B., Dgedge, M.,
-Menendez, C., Alonso, P.L., 2007.. Malaria Journal 6, 3.
-\url{https://doi.org/10.1186/1475-2875-6-3}
-
-\hypertarget{ref-McCarthy_2000}{}
-McCarthy, D., Wolf, H., Wu, Y., 2000a. The growth costs of malaria.
-\url{https://doi.org/10.3386/w7541}
-
-\hypertarget{ref-McCarthy2000}{}
-McCarthy, D., Wolf, H., Wu, Y., 2000b. The growth costs of malaria.
-National Bureau of Economic Research.
-\url{https://doi.org/10.3386/w7541}
-
-\hypertarget{ref-Minalu2011}{}
-Minalu, G., Aerts, M., Coenen, S., Versporten, A., Muller, A.,
-Adriaenssens, N., Beutels, P., Molenberghs, G., Goossens, H., Hens, N.,
-2011. Application of mixed-effects models to study the country-specific
-outpatient antibiotic use in europe: A tutorial on longitudinal data
-analysis. Journal of Antimicrobial Chemotherapy 66, vi79--vi87.
-\url{https://doi.org/10.1093/jac/dkr460}
-
-\hypertarget{ref-Mocumbi2017}{}
-Mocumbi, S., Gafos, M., Munguambe, K., Goodall, R., and, S.M., 2017.
-High HIV prevalence and incidence among women in southern mozambique:
-Evidence from the MDP microbicide feasibility study. PLOS ONE 12,
-e0173243. \url{https://doi.org/10.1371/journal.pone.0173243}
-
-\hypertarget{ref-Moonasar_2016}{}
-Moonasar, D., Maharaj, R., Kunene, S., Candrinho, B., Saute, F.,
-Ntshalintshali, N., Morris, N., 2016. Towards malaria elimination in the
-mosaswa (mozambique, south africa and swaziland) region. Malaria Journal
-15. \url{https://doi.org/10.1186/s12936-016-1470-8}
-
-\hypertarget{ref-Mouzin2011}{}
-Mouzin, E., al., E., 2011. Business Investing in Malaria Control:
-Economic Returns and a Healthy Workforce for Africa. Progress \& Impact
-series.
-
-\hypertarget{ref-Nhacolo_2006}{}
-Nhacolo, A.Q., Nhalungo, D.A., Sacoor, C.N., Aponte, J.J., Thompson, R.,
-Alonso, P., 2006. Levels and trends of demographic indices in southern
-rural mozambique: Evidence from demographic surveillance in manhiça
-district. BMC Public Health 6.
-\url{https://doi.org/10.1186/1471-2458-6-291}
-
-\hypertarget{ref-Nonvignon_2016}{}
-Nonvignon, J., Aryeetey, G.C., Malm, K.L., Agyemang, S.A., Aubyn,
-V.N.A., Peprah, N.Y., Bart-Plange, C.N., Aikins, M., 2016. Economic
-burden of malaria on businesses in ghana: A case for private sector
-investment in malaria control. Malaria Journal 15.
-\url{https://doi.org/10.1186/s12936-016-1506-0}
-
-\hypertarget{ref-OLaughlin2016}{}
-O'Laughlin, B., 2016. Consuming bodies: Health and work in the cane
-fields in xinavane, mozambique. Journal of Southern African Studies 43,
-625--641. \url{https://doi.org/10.1080/03057070.2016.1190519}
-
-\hypertarget{ref-Orem_2012}{}
-Orem, J., Kirigia, J., Azairwe, R., Kasirye, I., Walker, O., 2012.
-Impact of malaria morbidity on gross domestic product in uganda.
-International Archives of Medicine 5, 12.
-\url{https://doi.org/10.1186/1755-7682-5-12}
-
-\hypertarget{ref-Overgaard2012}{}
-Overgaard, H.J., Reddy, V.P., Abaga, S., Matias, A., Reddy, M.R.,
-Kulkarni, V., Schwabe, C., Segura, L., Kleinschmidt, I., Slotman, M.A.,
-2012. Malaria transmission after five years of vector control on bioko
-island, equatorial guinea. Parasites \& Vectors 5, 253.
-\url{https://doi.org/10.1186/1756-3305-5-253}
-
-\hypertarget{ref-Phillips98}{}
-Phillips, S.F.M., 1998. Economics and its contribution to the fight
-against malaria. Annals of Tropical Medicine And Parasitology 92,
-391--398. \url{https://doi.org/10.1080/00034989859375}
-
-\hypertarget{ref-Pluess2009}{}
-Pluess, B., Mueller, I., Levi, D., King, G., Smith, T.A., Lengeler, C.,
-2009. Malaria a major health problem within an oil palm plantation
-around popondetta, papua new guinea. Malaria Journal 8, 56.
-\url{https://doi.org/10.1186/1475-2875-8-56}
-
-\hypertarget{ref-Purdy_2013}{}
-Purdy, M., Rublin, D., Wei, K., Robinson, M., 2013. The economic case
-for combating malaria. The American Journal of Tropical Medicine and
-Hygiene 89, 819--823. \url{https://doi.org/10.4269/ajtmh.12-0689}
-
-\hypertarget{ref-R}{}
-R Core Team, 2017. R: A language and environment for statistical
-computing. R Foundation for Statistical Computing, Vienna, Austria.
-
-\hypertarget{ref-Robbins2012}{}
-Robbins, G., Perkins, D., 2012. Mining fdi and infrastructure
-development on africa's east coast: Examining the recent experience of
-tanzania and mozambique. Journal of International Development 24,
-220--236. \url{https://doi.org/10.1002/jid.2817}
-
-\hypertarget{ref-Sachs2002}{}
-Sachs, J., Malaney, P., 2002. The economic and social burden of malaria.
-Nature 415, 680--685. \url{https://doi.org/10.1038/415680a}
-
-\hypertarget{ref-Sacoor2013}{}
-Sacoor, C., Nhacolo, A., Nhalungo, D., Aponte, J.J., Bassat, Q.,
-Augusto, O., Mandomando, I., Sacarlal, J., Lauchande, N., Sigauque, B.,
-Alonso, P., Macete, E., 2013. Profile: Manhica health research centre
-(manhica HDSS). International Journal of Epidemiology 42, 1309--1318.
-\url{https://doi.org/10.1093/ije/dyt148}
-
-\hypertarget{ref-Sate2003}{}
-Saúte, F., Aponte, J., Ahmeda, J., Ascaso, C., Vaz, N., Dgedge, M.,
-Alonso, P., 2003. Malaria in southern mozambique: Incidence of clinical
-malaria in children living in a rural community in manhiça district.
-Transactions of the Royal Society of Tropical Medicine and Hygiene 97,
-655--660. \url{https://doi.org/10.1016/s0035-9203(03)80097-4}
-
-\hypertarget{ref-Shretta2016}{}
-Shretta, R., Avanceña, A.L.V., Hatefi, A., 2016. The economics of
-malaria control and elimination: A systematic review. Malaria Journal
-15. \url{https://doi.org/10.1186/s12936-016-1635-5}
-
-\hypertarget{ref-Shretta_2017}{}
-Shretta, R., Baral, R., Avanceña, A.L.V., Fox, K., Dannoruwa, A.P.,
-Jayanetti, R., Jeyakumaran, A., Hasantha, R., Peris, L., Premaratne, R.,
-2017. An investment case to prevent the reintroduction of malaria in sri
-lanka. The American Journal of Tropical Medicine and Hygiene 16--0209.
-\url{https://doi.org/10.4269/ajtmh.16-0209}
-
-\hypertarget{ref-sutton}{}
-Sutton, J., 2014. Mapa empresarial de moçambique.
-
-\hypertarget{ref-Vecchi_2013}{}
-Vecchi, V., Hellowell, M., Gatti, S., 2013. Does the private sector
-receive an excessive return from investments in health care
-infrastructure projects? Evidence from the uk. Health Policy 110,
-243--270. \url{https://doi.org/10.1016/j.healthpol.2012.12.010}
-
-\hypertarget{ref-White_2011}{}
-White, M.T., Conteh, L., Cibulskis, R., Ghani, A.C., 2011. Costs and
-cost-effectiveness of malaria control interventions - a systematic
-review. Malaria Journal 10, 337.
-\url{https://doi.org/10.1186/1475-2875-10-337}
-
-\hypertarget{ref-White2014}{}
-White, N.J., Pukrittayakamee, S., Hien, T.T., Faiz, M.A., Mokuolu, O.A.,
-Dondorp, A.M., 2014. Malaria. The Lancet 383, 723--735.
-\url{https://doi.org/10.1016/s0140-6736(13)60024-0}
-
-\hypertarget{ref-World2016}{}
-WHO, 2016. World Malaria Report.
-
-\hypertarget{ref-whoprof}{}
-WHO, 2015. Malaria profile: Mozambique.
-
-\hypertarget{ref-Winkler}{}
-Winkler, D., 2013. Potential and Actual FDI Spillovers in Global Value
-Chains.
-
-\hypertarget{ref-Wondwosen2018}{}
-Wondwosen, B., Birgersson, G., Tekie, H., Torto, B., Ignell, R., Hill,
-S.R., 2018. Sweet attraction: Sugarcane pollen-associated volatiles
-attract gravid anopheles arabiensis. Malaria Journal 17.
-\url{https://doi.org/10.1186/s12936-018-2245-1}
+\section{References}\label{references}
 

### 2018-05-15 inst/rmd/02_methods.Rmd
major changes

@@
 # Methods
+
+## Budget info
+
+```{r}
+costs_itemized <- maragra::costs_itemized
+costs_raw <- maragra::costs_raw
+costs <- maragra::costs
+# Annual operating budget
+aob <- round(mean(costs$usd), digits = 2)
+print(paste0('Average annual operating budget of ', aob))
+
+plot_data <- costs_itemized %>%
+  filter(!is.na(Account)) %>%
+  mutate(area = Account) %>%
+  mutate(area = ifelse(grepl('WAGES', area), 'Human resources', area)) %>%
+  mutate(area = ifelse(area %in% c('LUBRICANTS', 'TYRES & TUBES',
+                                   'VEHICLE MAINTENANCE', 'DIESEL'),
+                       'Vehicles', area)) %>%
+  mutate(area = ifelse(area %in% c('PENSION FUND CONTRIBUTIONS',
+                                   'SALARIES BASIC', 'HOUSING ALLOWANCE',
+                                   'LEAVE PAY PROVISION'), 'Human resources', area)) %>%
+  mutate(area = ifelse(area %in% toupper(c('Equipment maintenance',
+                                   'Protective clothing',
+                                   'Tools',
+                                   'Consumables')),
+                       'Equipment', area)) %>%
+  mutate(area = ifelse(area %in% c('ENVIRONMENTAL MONITORING',
+                                   'TRAVEL (FOREIGN)'),
+                       'Other', area)) %>%
+  group_by(period, area) %>%
+  summarise(total = sum(usd, na.rm = TRUE)) %>%
+  ungroup %>%
+  mutate(area = tolower(area)) %>%
+  mutate(area = Hmisc::capitalize(area))
+
+ggplot(data = plot_data,
+             aes(x = period,
+                 y = total,
+                 group = area,
+                 fill = area)) +
+  geom_bar(stat = 'identity',
+           position = 'dodge') +
+  scale_fill_manual(name = '',
+                    values = colorRampPalette(brewer.pal(n = 9,
+                                                         name = 'Set1'))(length(unique(plot_data$area)))) +
+  theme_maragra() +
+  labs(x = '',
+       y = 'USD')
+
+fumigations <- mc %>%
+  filter(date >= '2013-01-01',
+         date <= '2016-12-31')
+n_fumigations <- nrow(fumigations)
+n_buildings <- sum(fumigations$casas_cobertas, na.rm = TRUE)
+unique_agregados <- length(unique(fumigations$unidade))
+people_sprayed <- length(unique(model_data$oracle_number[model_data$days_since != 'Never']))
+people_observed <- length(unique(model_data$oracle_number))
+pr <- function(x){prettyNum(x, big.mark = ',')}
+
+print('Assuming linearity in costs, the program spends approximately xxx per building sprayed.') 
+aob / (n_buildings / 4)
+
+# print('With each agregado containing an average of 2.2 workers.')
+
+x = ab_panel %>%
+  mutate(year = as.numeric(format(date, '%Y')),
+         month = as.numeric(format(date, '%m'))) %>%
+  group_by(year, month) %>%
+  summarise(absences = length(which(absent))) %>%
+  ungroup %>%
+  left_join(clinic_agg %>% group_by(year, month) %>% 
+              summarise(positive = sum(positive))) %>%
+  mutate(percent_clinic = positive / absences * 100) %>%
+  mutate(n_ab_for_clinic = absences / positive)
+
+mean(x$n_ab_for_clinic, na.rm = TRUE)
+```
 
@@
 
+```{r}
+x <- model_data %>%
+  group_by(group) %>%
+  summarise(avg = mean(herd, na.rm = TRUE),
+            avg_ideal = mean(herd_ideal, na.rm = TRUE)) %>%
+  ungroup %>%
+  mutate(avg = avg / max(avg) * 100,
+            avg_ideal = avg_ideal / max(avg_ideal) * 100)
+```
+
 ### Econometric model

### 2018-05-15 inst/rmd/03_results.Rmd
major changes

@@
 
-We collected absenteeism and demographic data data on `r n` workers from 2012 through 2016.
+We collected absenteeism and demographic data data on `r n` workers from 2013 through 2016.
 
@@
 
-```{r, fig.width = 5, fig.height = 3.5, fig.align='center', fig.cap = 'Age distribution by sex and worker status'}
+```{r, fig.align='center', fig.cap = 'Age distribution by sex and worker status'}
 # x$age <- round(as.numeric(as.Date('2014-01-01') - x$date_of_birth) / 365.25, digits = 2)
@@
 
-## What would happen without IRS?
+## Sensitivity analysis
+
+The below shows average FE for year / worker, during malaria season, on a rainy day...
+
+```{r}
+expander <- function(x){sort(unique(unlist(model_data[,x])))}
+dummy <- expand.grid(season = expander('season'),
+                     months_since = expander('months_since'),
+                     rainy_day = expander('rainy_day'),
+                     herd = mean(model_data$herd),
+                     group = expander('group'))
+dummy$predicted <- NA
+
+groups <- sort(unique(dummy$group))
+for (i in 1:length(groups)){
+    message(i)
+    this_group <- groups[i]
+    indices <- which(dummy$group == this_group)
+    model <- protection_models[[this_group]]
+    fe <- getfe(model)
+    # Add back in the mean effect of fixed effects
+    add <- fe %>% group_by(fe) %>% summarise(effect = mean(effect)) %>% summarise(sum(effect)) %>% unlist %>% as.numeric
+    data <- dummy[indices,]  
+    # predict_felm(model = model, data = )
+    dummy$predicted[indices] <- 
+      (regtools::predict.felm(object = model,
+                             newdata = data,
+                             use.fe = FALSE) + add) * 100
+}
+
+ggplot(data = dummy %>% filter(rainy_day),
+       aes(x = months_since,
+           y = predicted,
+           color = season,
+           group = season)) +
+  geom_line() +
+  geom_point() +
+  facet_wrap(~group, scales = 'free_y') +
+  scale_colour_manual(name = 'Malaria\nseason',
+                      values = c('blue', 'red')) +
+  theme_maragra() +
+  labs(x = 'IRS status',
+       y = 'Estimated absenteeism') #+
+  # theme(legend.position = 'right')
+
+```
+
+The below shows estimated importance of herd protection, and heterogeniety across groups.
+
+```{r}
+dummy <- expand.grid(season = expander('season'),
+                     months_since = expander('months_since'),
+                     rainy_day = expander('rainy_day'),
+                     herd = 0:1000,
+                     group = expander('group'))
+dummy$predicted <- NA
+
+groups <- sort(unique(dummy$group))
+for (i in 1:length(groups)){
+    message(i)
+    this_group <- groups[i]
+    indices <- which(dummy$group == this_group)
+    model <- protection_models[[this_group]]
+    fe <- getfe(model)
+    # Add back in the mean effect of fixed effects
+    add <- fe %>% group_by(fe) %>% summarise(effect = mean(effect)) %>% summarise(sum(effect)) %>% unlist %>% as.numeric
+    data <- dummy[indices,]  
+    dummy$predicted[indices] <- 
+      (regtools::predict.felm(object = model,
+                             newdata = data,
+                             use.fe = FALSE) + add) * 100
+}
+
+ggplot(data = dummy %>% filter(rainy_day, season == 'high'),
+       aes(x = herd,
+           y = predicted,
+           color = months_since,
+           group = months_since)) +
+  geom_line() +
+  facet_wrap(~group, scales = 'free_y') +
+  scale_colour_manual(name = 'IRS status',
+                      values = c('darkorange', 'darkgreen')) +
+  theme_maragra() +
+  labs(x = 'Herd protection score',
+       y = 'Estimated absenteeism')
+```
+
+### What would happen without IRS?
 
@@
 x <- model_data %>%
-  dplyr::filter(season == 'high',
+  dplyr::filter(season == 'high'#,
                 # group == groups[1],
-                permanent_or_temporary == 'Permanent'
+                # permanent_or_temporary == 'Permanent'
                 )
 x <- x [,grepl('predicted', names(x)) | names(x) == 'group']
-x <- x %>% gather(key, value, predicted_no_herd_no_irs:predicted_max_irs)
+x <- x %>% gather(key, value, predicted_max_herd_max_irs:predicted)
 dict <- data.frame(key = c('predicted', 
@@
                                'As is',
+                               'Protection: maximum IRS',
                                'Protection: maximum herd',
-                               'Protection: maximum IRS',
                                'Protection: maximum IRS and herd'))
@@
 cols <- colorRampPalette(brewer.pal(n = 9,
-                                    name = 'Spectral'))(length(unique(plot_data$group)))
+                                    name = 'Set1'))(length(unique(plot_data$group)))
 
+library(ggrepel)
 ggplot(data = plot_data,
@@
            color = group)) +
-  geom_point() +
+  geom_point(size = 4) +
   theme_maragra() +
@@
   geom_line(aes(group = group),
-            alpha = 0.3) +
+            alpha = 0.5,
+            size = 1) +
   scale_color_manual(name = '',
-                     values = cols)
+                     values = cols) #+
+  # ggrepel::geom_label_repel(aes(label = round(avg * 100, digits = 1)))
+```
+
+## Return on investment
+
+- Number of working days
+
+```{r, echo = TRUE}
+x <- ab_panel %>% group_by(date) %>% tally
+ggplot(data = x,aes(x = date,y = n)) + geom_line() +
+  labs(title = mean(x$n))
+```
+
+- Number of working days annually
+
+```{r, echo = TRUE}
+options(scipen = '999')
+x <- ab_panel %>% mutate(year = as.numeric(format(date, '%Y'))) %>% group_by(year) %>% tally
+ggplot(data = x,aes(x = year,y = n)) + geom_line() +
+  labs(title = mean(x$n))
+```
+
+- Number of prevented absences per year
+
+```{r, echo = TRUE}
+ratio <- length(unique(ab_panel$oracle_number)) / length(unique(model_data$oracle_number))
+
+x <- model_data %>%
+  group_by(year = calendar_year) %>%
+  summarise(eligibles = n(),
+            absences = mean(as.numeric(absent), na.rm = TRUE) * n(),
+            counterfactual = mean(predicted_no_herd_no_irs,na.rm =TRUE) * n()) 
+y <- x %>%
+  summarise(x = mean(counterfactual) - mean(absences)) %>%
+  unlist * ratio
+# Annual prevented
+y
+x <- x %>%
+  mutate(prevented = counterfactual - absences)
+# Number of presences to prevent 1 case
+sum(x$eligibles) / sum(x$prevented)
+```
+
+```{r, eval = TRUE}
+# Clinical costs
+(y /
+  80) * # number of absences it takes to produce one company clinic positve
+  31.49 # cost per clinical case per ezenduka
+
+# Wage costs
+y *
+  4 # assuming USD
 ```
@@
 ```{r, echo = FALSE}
-ro_data <- ab_panel %>%
-  left_join(irs, by = c('date', 'unidade')) %>%
-  filter(days_since < 0) %>%
-  mutate(pre_irs = between(days_since, -10, -1)) %>%
-  # mutate(time_period = lendable::time_period_extract(date)) %>%
-  # mutate(time_period = as.character(time_period)) %>%
-  mutate(time_period = make_season(date = date)) %>%
-  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) %>%
-  left_join(workers %>%
-              dplyr::select(oracle_number,
-                            sex,
-                            department,
-                            permanent_or_temporary),
-            by = 'oracle_number')
+ro_data <- model_data %>%
+# ro_data <- ab_panel %>%
+#   left_join(irs, by = c('date', 'unidade')) %>%
+#   filter(days_since < 0) %>%
+  mutate(pre_irs = between(days_since, -10, -1)) 
+#   # mutate(time_period = lendable::time_period_extract(date)) %>%
+#   # mutate(time_period = as.character(time_period)) %>%
+#   mutate(time_period = make_season(date = date)) %>%
+#   mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) %>%
+#   left_join(workers %>%
+#               dplyr::select(oracle_number,
+#                             sex,
+#                             department,
+#                             permanent_or_temporary),
+#             by = 'oracle_number')
 # ro_data$time_period <- as.character(ro_data$time_period)
@@
   return(kable(out, format = "html") %>%
-  kable_styling() %>%
-    row_spec(row = which(grepl('prior', out$Variable)),background = "yellow"))
+  kable_styling()# %>%
+    # row_spec(row = which(grepl('prior', out$Variable)),background = "yellow")
+  )
 }
@@
 # Adjust for seasonality and department
-make_model('absent ~ pre_irs*time_period + department')
+make_model('absent ~ pre_irs*season + group + rainy_day')
 ```
 
-The below shows our robustness check for sick only absenteeism.
+```{r}
+# The below shows our robustness check for sick only absenteeism.
 Unlike with all cause absenteeism, during malaria season, being in the period 10 days prior to IRS is associated with statistically greater likelihood of being absent for illness.
 In other words, it appears that there is somewhat of a feedback loop: when a worker misses work due to illness, his/her likelihood of getting IRS doubles in the next 10 days.\todo{What to do about this???}
-
-```{r}
-make_model('absent_sick ~ pre_irs*time_period + department')
+# make_model('absent_sick ~ pre_irs*time_period + department')
 ```
-
-The second concern is that our quantification of return on investment is distorted by the fact that we treat IRS operations as essentially linear in nature, when in reality economies of scale, in-kind purchases and other factors likely make the true cost-per-spraying convex.
-We address this by creating three scenarios: (1) a "start-up" scenario in which we take into account all costs incurred by starting a program from scratch, (2) a "normal" scenario in which we match the assumptions with those used in this paper (ie, account for "wear-and-tear" depreciation but not vehicle purchase, etc.) and a (3) "absorbed costs" scenario which ignores all costs which are not directly incurred by the program....
-xxx...
-\todo{Will address this with some sensitivity analysis}
-
-\newpage

### 2018-05-15 inst/rmd/paper.Rmd
major changes

@@
   word_document: default
-fig_width: 4
-fig_height: 2.6
+fig_width: 8
+fig_height: 6
 bibliography: bibliography.bib

### 2018-05-15 inst/rmd/paper_prepare.R

@@ -8,7 +8,8 @@ library(knitr)
 library(RColorBrewer)
 # library(lme4)
 library(lfe)
-
+library(restorepoint)
+library(regtools)
 if('prepared_data.RData' %in% dir()){
   load('prepared_data.RData')
 } else {
@@ -38,8 +39,6 @@ if('prepared_data.RData' %in% dir()){
     return(x)
   }
   
-  # filter for on-site only
-  
   model_data <-
     ab_panel %>%
     left_join(irs, by = c('unidade', 'date')) %>%
@@ -47,6 +46,7 @@ if('prepared_data.RData' %in% dir()){
     mutate(days_since = ifelse(is.na(days_since), 'Never',
                                ifelse(days_since < 0, 'Before',
                                       ifelse(days_since >= 12, '12+', as.character(add_zero(days_since, n = 2)))))) 
+  
   model_data <-
     model_data %>%
     mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) %>%
@@ -93,30 +93,17 @@ if('prepared_data.RData' %in% dir()){
   model_data <- model_data %>%
     mutate(months_since = days_since) %>%
     mutate(months_since = as.character(months_since)) %>%
-    mutate(months_since = ifelse(months_since %in% c(#'Never', 
-                                                     'Before', '00', '01',
-                                                     '02', '03', '04', '05', '06'),
-                                 months_since,
-                                 ifelse(months_since %in% c('07', '08',
-                                                            '09', '10', '11', '12+'),
-                                        '07+', months_since))) %>%
+    mutate(months_since = ifelse(months_since %in% c('00', '01',
+                                                     '02', '03', '04', '05'),
+                                 'After',
+                                 'Before')) %>%
     # Define whether every sprayed
     ungroup %>%
     group_by(oracle_number) %>%
     mutate(ever_sprayed = length(which(unidade %in% mc$unidade)) > 0) %>%
     ungroup %>%
-    # Correct the nevers
-    mutate(days_since = ifelse(days_since == 'Never' & ever_sprayed, 'Before', days_since)) %>%
     mutate(months_since = as.character(months_since)) %>%
-    mutate(months_since = ifelse(months_since == 'Never' & ever_sprayed, 'Before', months_since)) %>%
-    # Make never and before the same
[…]

### 2018-05-16 inst/rmd/02_methods.Rmd
update

@@
 $$
-\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * (\hat{\beta}_2{IRS_{it}}*\hat{\beta}_3{IRStime_{it}}) +  \hat{\beta}_4{Herd_{it}} +  \alpha_i + \delta_t + \upsilon_{it}
+\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * IRS_{it} + \hat{\beta}_2{RainyDay_{t}} + \hat{\beta}_3{Herd_{it}} +  \alpha_i + \delta_t + \upsilon_{it}
 $$
@@
 $\hat{Y}_{it}$ is the rate of absence.
-$\beta_{1}$ is the binary "season" variable, imputed from overall district clinical incidence.
-Our intervention is whether the residence of the worker in question was treated in the last year, and, if so, the time since treatment, represented, respectively, by $\beta_{2}$ and $\beta_{3}$.
-$\beta_{4}$ **represents the distance-weighted herd protection score**.
+$\beta_{1}$ is the binary "season" variable, imputed from overall district clinical incidence, interacted with our "treatment" condition: whether or not the worker in question's residence was sprayed with IRS in the preceding 6 months or not.
+$\beta_{2}$ is whether it was raining on that particular day or not.
+$\beta_{3}$ represents the distance-weighted herd protection score.
 $\alpha_i$ represents the time invariant worker fixed effects, and $\delta_i$ represents the fixed effect of the particular malaria season.

### 2018-05-23 inst/rmd/menno_models.Rmd
minor changes

---
title: "Concept note: additional models"
subtitle: "Pradhan, Sicuri, Brew"
date: ''
output: 
  pdf_document:
    keep_tex: true
    latex_engine: xelatex
    includes:
      in_header: preamble.sty
fig_width: 4
fig_height: 2.6
bibliography: bibliography.bib
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# csl: journal-of-health-economics.csl

# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(RColorBrewer)
library(extrafont)
library(kableExtra)
library(raster)
loadfonts()
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               fig.height = 4,
               fig.width = 5,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               dev = "cairo_pdf",
               fig.pos="!h",
               fig.align = 'center')
opts_knit$set(width=75)
options(xtable.comment = FALSE)

```

```{r, eval = TRUE}
source('paper_prepare.R')
```

\begin{center}
\begin{large}

A few different approaches to modeling absenteeism in the Maragra data

\end{large}
\end{center}

\vspace{5mm}

\begin{center}
\textbf{Overview}  
\end{center}

\vspace{5mm}
\begin{center}
\begin{changemargin}{3cm}{3cm}

This document contains methods and basic results for a few different modeling approaches, following a conversation and write-up with Menno Pradhan in May 2018. The purpose of this document is to make explicit the current models we have implemented, and some issues with them.

\end{changemargin}
\end{center}

\vspace{20mm}

\noindent\fbox{%
    \parbox{\textwidth}{%
        \subsection*{Main points}
        \begin{itemize}
          \item I have implemented Menno's suggestions for more simple models.
          \item The simplest models show no overall effect from IRS when not accounting for malaria season
          \item I believe that the lack of an effect is due to seasonality's presence in the model as a fixed (additive) effect, rather than as an interaction with IRS
        \end{itemize}
        \vspace{2mm}
    }%
}

\vfill
\null

\subsection*{Desinataires}
\textbf{Elisa Sicuri; Menno Pradhan}

\vspace{3mm}

\newpage

# The current approach

In the current draft of the paper (\href{https://docs.google.com/document/d/1bUWRBCgVcgjSPHchIQxiTG8Vwv5hV1GLU4Tlu386sWA/edit#}{HERE}
), the model formula is:

$$
\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * IRS_{it} + \hat{\beta}_2{RainyDay_{t}} + \hat{\beta}_3{Herd_{it}} +  \alpha_i + \delta_t + \upsilon_{it}
$$

In R, using the `felm` library, this is written as follows:

```{r, echo = FALSE}
protection_models[[1]]$call
```

The above notation shows `absent` (binary 1,0) as a function of the interaction of binary `season` (high vs. low) and binary `months_since` (before vs. after) plus binary `rainy_day` (whether or not it rained on the particular day) plus continuous numeric `herd` (ie, the protection afforded by others) plus fixed effects for both the worker (`oracle_number`) and the year (`malaria_year`).  $\alpha_i$ represents the time invariant worker fixed effects, and $\delta_i$ represents the fixed effect of the particular malaria season. $\upsilon$ is the error term. This model is run separately for the 4 different worker groups, yielding the following results:

```{r, results = 'asis'}
make_models_table(model_list = protection_models, the_caption = "All absenteeism with herd immunity: model results", type = 'latex', multiplier = 1)
```

In the above, "after" IRS application (ie, the 6 months following IRS) is associated with a significant reduction in absenteeism for both kinds of permanent workers, and does not have a signficiant association with a change in absenteeism for both kinds of temporary workers.

# Suggested changes from Menno

Menno sent a write-up on May 17th (see email with subject "suggestions for models") with some proposed changes to how we model absenteeism. What follows is an implementation of those suggestions.

## Model A (without seasonality)

The below is a characterization of the approach Menno suggested:

- We use monthly instead of daily data (modeling a percentage absenteeism rate rather than a binary).
- We use fixed effects for year-month combination. 
- We don't explicitly model for malaria season. 
- We don't have any interaction terms.
- We use dummies for months since IRS.
- We use the most recent spraying as the time since IRS (ie, if 5 months ago, and 1 month ago, then the dummy is applied for 1 month ago but not 5 months ago).

Using this approach, our panel data looks like this:

```{r}
# Create a model per menno's specifications
menno_models <- list()

menno_data <- model_data %>%
  # filter(!is.na(days_since)) %>%
  # filter(days_since >= -184, days_since <= 184) %>%
  group_by(oracle_number, group, year_month) %>%
  summarise(absent = length(which(absent)) / length(absent) * 100,
            days = n(),
            herd = mean(herd, na.rm = TRUE),
            months_since_menno = dplyr::first(months_since_menno),
            season = dplyr::first(season)) %>%
  ungroup
groups <- sort(unique(model_data$group))
# library(lmerTest)
# library(nlme)
for (i in 1:length(groups)){
  message(i)
  this_group <- groups[i]
  message(this_group)
  these_data <- menno_data %>% filter(group == this_group)
  this_menno_model <- felm(absent ~ months_since_menno    | oracle_number + year_month | 0 | 0,
                                data = these_data)

  menno_models[[i]] <- this_menno_model
}
names(menno_models) <- groups
print(head(menno_data %>% filter(months_since_menno != 'Before') %>% sample_n(20) %>% dplyr::select(-season, -herd)))
```

The model formula was described in Menno's email.

\includegraphics{images/menno}[width=3cm]

Which translates into R as:

```{r}
menno_models[[1]]$call
```

Note that the `months_since_menno` variable has the following categories, and is treated as a "wide" dummy by the software in modeling:

```{r}
levels(menno_data$months_since_menno)
```

The results of this model look like this:

```{r, results = 'asis'}
make_models_table(model_list = menno_models, the_caption = "Simple models with year-month fixed effects", type = 'latex', multiplier = 1)
```

In the above, it is apparent that the effects of IRS appear to be either (a) in the opposite direction as expected or (b) non-signficant. Here is a visualization of IRS' apparent effect (relative to the "Before" period, which is the reference class). Were IRS effective, we would expect the black line to be below 0 (the red line).

```{r}
out_list <- list()
for (i in 1:length(groups)){
  this_group <- groups[[i]]
  x <- coef(menno_models[[i]])
  x <- x[grepl('months_since', names(x))]
  names(x) <- 1:6
  out_list[[i]] <- data.frame(group = this_group,
                              month = names(x),
                              coef = x)
}
out <- bind_rows(out_list)

ggplot(data = out,
       aes(x = month,
           y = coef)) +
  geom_line(aes(group = 1)) +
  facet_wrap(~group) +
  geom_hline(yintercept = 0, 
             lty = 2,
             col = 'red',
             alpha = 0.6) +
  labs(x = 'Months since IRS',
       y = 'Coefficient')
```

What is going on here? Could it  be that IRS has a positive (ie, increases) impact on absenteeism? Let's look at the raw data, rather than the coefficients. The below chart shows absenteeism (as a percentage of the pre-IRS absenteeism average) for workers, broken down by season (row) and worker type (column), with point size reflecting the number of observations.

```{r, fig.width = 8}
pd <- menno_data %>%
  group_by(season, months_since_menno, group) %>%
  summarise(absenteeism_rate = mean(absent, na.rm = TRUE),
            days = sum(days)) %>%
  # summarise(absenteeism_rate = length(which(absent)) / length(absent) * 100) %>%
  ungroup %>%
  group_by(group, season) %>%
  mutate(start = dplyr::first(absenteeism_rate)) %>%
  mutate(p = absenteeism_rate / start * 100)

options(scipen = '999')
ggplot(data = pd %>% ungroup %>% 
         mutate(season = paste0(season, ' season')) %>%
         mutate(group = gsub(' worker', '', group)),
       aes(x = months_since_menno,
           y = p)) +
  geom_point(aes(size = days)) +
  geom_line(aes(group = 1), alpha = 0.4) +
  facet_grid(season~group) +
  geom_hline(yintercept = 100, 
             color = 'red',
             lty = 2) +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_log10(breaks = seq(0, 500, 100)) +
  labs(y = 'Absenteeism (relative to before IRS, log)',
       x = 'Time') +
  scale_size_continuous(name = 'Observations',
                    breaks = c(10000, 100000, 5000000))

```

In the above chart, two things stand out:

1. Variability is greatest among temporary workers.
2. There appears to be a reduction in absenteeism post IRS for permanent workers _during the malaria season_, but not during the non-malaria season

So, perhaps controlling for year-month through fixed effects isn't enough, since the effect of IRS _interacts_ with the season...

## Model B (with seasonality)

Using this approach, our panel data looks like this:

```{r}
# Create a model per menno's specifications
menno_models <- list()

for (i in 1:length(groups)){
  message(i)
  this_group <- groups[i]
  message(this_group)
  these_data <- menno_data %>% filter(group == this_group)
  this_menno_model <- felm(absent ~ months_since_menno*season  | oracle_number + year_month | 0 | 0,
                                data = these_data)

  menno_models[[i]] <- this_menno_model
}
names(menno_models) <- groups
print(head(menno_data %>% filter(months_since_menno != 'Before') %>% sample_n(20) %>% dplyr::select(-herd)))
```

Our new model specification (in R) is:

```{r}
menno_models[[1]]$call
```

The results of this model look like this:

```{r, results = 'asis'}
make_models_table(model_list = menno_models, the_caption = "Simple models with year-month fixed effects and seasonality-IRS interaction term", type = 'latex')
```

The below shows model coefficients:

```{r, fig.width = 8}
out_list <- list()
for (i in 1:length(groups)){
  this_group <- groups[[i]]
  x <- clean_up_model(menno_models[[i]], multiplier = 1)
  x <- x %>%
    filter(grepl('IRS status', Term),
           grepl('season', Term)) 
  # Malaria season
  df1 <- data.frame(group = this_group,
                              month = 1:6,
                              coef = x$Estimate,
                    season = 'high')

  x <- clean_up_model(menno_models[[i]], multiplier = 1)
  x <- x %>%
    filter(grepl('IRS status', Term),
           !grepl('season', Term)) 
  # Malaria season
  df2 <- data.frame(group = this_group,
                              month = 1:6,
                              coef = x$Estimate,
                    season = 'low')
  out_list[[i]] <- bind_rows(df1, df2) 
}
out <- bind_rows(out_list)
x <- strsplit(out$coef, ' ')
x <- lapply(x, function(x){x[1]})
x <- unlist(x)
out$coef <- as.numeric(as.character(x))
ggplot(data = out,
       aes(x = month,
           y = coef)) +
  geom_line(aes(group = 1)) +
  facet_grid(season~group) +
  geom_hline(yintercept = 0, 
             lty = 2,
             col = 'red',
             alpha = 0.6) +
  labs(x = 'Months since IRS',
       y = 'Coefficient')
```

## Model C (with seasonality and herd score)

We now introduce herd immunity to the model. This is introduced as an additive effect; perhaps it should be interactive with IRS? Using this approach, our panel data looks like this:

```{r}
# Create a model per menno's specifications
menno_models <- list()

for (i in 1:length(groups)){
  message(i)
  this_group <- groups[i]
  message(this_group)
  these_data <- menno_data %>% filter(group == this_group)
  this_menno_model <- felm(absent ~ months_since_menno*season  | oracle_number + year_month | 0 | 0,
                                data = these_data)

  menno_models[[i]] <- this_menno_model
}
names(menno_models) <- groups
print(head(menno_data %>% filter(months_since_menno != 'Before') %>% sample_n(20) ))
```

Our new model specification (in R) is:

```{r}
menno_models[[1]]$call
```

The results of this model look like this:

```{r, results = 'asis'}
make_models_table(model_list = menno_models, the_caption = "Simple models with year-month fixed effects, seasonality-IRS interaction term, and herd score", type = 'latex')
```

### 2018-05-23 inst/rmd/menno_models.tex
minor changes

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Concept note: additional models},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Concept note: additional models}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
\subtitle{Pradhan, Sicuri, Brew}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\pagenumbering{gobble}
\usepackage{longtable}
\usepackage[utf8]{inputenc}
\usepackage{changepage}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{geometry}
\usepackage{fancyhdr}
\usepackage{color}
\usepackage{colortbl}
\usepackage{color}
\usepackage{hyperref}
% Font
\usepackage{fontspec}
\setmainfont{Swift-Regular_43151.ttf}
\setsansfont[BoldFont={Swift-Bold_43130.ttf}]{Swift-Regular_43151.ttf}
% \setmonofont{Swift-Regular_43151.ttf}
\renewcommand{\familydefault}{\sfdefault}
% \usepackage{fontspec}
% \setmainfont{Lato-Regular.ttf}
% \setsansfont[BoldFont={Lato-Bold.ttf}]{Lato-Regular.ttf}
% \renewcommand{\familydefault}{\sfdefault}

\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist
\renewcommand{\rmdefault}{ppl}

\usepackage{multicol}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{lipsum}

\usepackage{longtable}

\usepackage{float}
\floatplacement{figure}{H}

% \usepackage{todonotes} % for side notes
% \usepackage[colorinlistoftodos]{todonotes} % for side notes

\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage[dvipsnames, table]{xcolor}  % Coloured text etc.
% 
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\newcommandx{\unsure}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
\newcommandx{\change}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
\newcommandx{\info}[2][1=]{\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25,bordercolor=OliveGreen,#1]{#2}}
\newcommandx{\improvement}[2][1=]{\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
\newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}
\usepackage{lmodern}
\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Return on investment of private sector malaria control at a large sugar facility in Southern Mozambique}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\newcommand{\footremember}[2]{%
    \footnote{#2}
    \newcounter{#1}
    \setcounter{#1}{\value{footnote}}%
}
\newcommand{\footrecall}[1]{%
    \footnotemark[\value{#1}]%
}

\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist

\widowpenalties 1 150

\makeatletter
\renewcommand\footnotesize{%
   \@setfontsize\footnotesize\@ixpt{11}%
   \abovedisplayskip 8\p@ \@plus2\p@ \@minus4\p@
   \abovedisplayshortskip \z@ \@plus\p@
   \belowdisplayshortskip 4\p@ \@plus2\p@ \@minus2\p@
   \def\@listi{\leftmargin\leftmargini
               \topsep 4\p@ \@plus2\p@ \@minus2\p@
               \parsep 2\p@ \@plus\p@ \@minus\p@
               \itemsep \parsep}%
   \belowdisplayskip \abovedisplayskip
}
\makeatother

\DeclareTextCommandDefault{\nobreakspace}{\leavevmode\nobreak\ }
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}

\begin{document}
\maketitle

\begin{center}
\begin{large}

A few different approaches to modeling absenteeism in the Maragra data

\end{large}
\end{center}

\vspace{5mm}

\begin{center}
\textbf{Overview}  
\end{center}

\vspace{5mm}

\begin{center}
\begin{changemargin}{3cm}{3cm}

This document contains methods and basic results for a few different modeling approaches, following a conversation and write-up with Menno Pradhan in May 2018. The purpose of this document is to make explicit the current models we have implemented, and some issues with them.

\end{changemargin}
\end{center}

\vspace{20mm}

\noindent\fbox{%
    \parbox{\textwidth}{%
        \subsection*{Main points}
        \begin{itemize}
          \item I have implemented Menno's suggestions for more simple models.
          \item The simplest models show no overall effect from IRS when not accounting for malaria season
          \item I believe that the lack of an effect is due to seasonality's presence in the model as a fixed (additive) effect, rather than as an interaction with IRS
        \end{itemize}
        \vspace{2mm}
    }%
}

\vfill
\null

\subsection*{Desinataires}

\textbf{Elisa Sicuri; Menno Pradhan}

\vspace{3mm}

\newpage

\section{The current approach}\label{the-current-approach}

In the current draft of the paper
(\href{https://docs.google.com/document/d/1bUWRBCgVcgjSPHchIQxiTG8Vwv5hV1GLU4Tlu386sWA/edit#}{HERE}
), the model formula is:

\[
\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * IRS_{it} + \hat{\beta}_2{RainyDay_{t}} + \hat{\beta}_3{Herd_{it}} +  \alpha_i + \delta_t + \upsilon_{it}
\]

In R, using the \texttt{felm} library, this is written as follows:

\begin{verbatim}
felm(formula = absent ~ season * months_since + rainy_day + herd | 
    oracle_number + malaria_year | 0 | 0, data = these_data)
\end{verbatim}

The above notation shows \texttt{absent} (binary 1,0) as a function of
the interaction of binary \texttt{season} (high vs.~low) and binary
\texttt{months\_since} (before vs.~after) plus binary
\texttt{rainy\_day} (whether or not it rained on the particular day)
plus continuous numeric \texttt{herd} (ie, the protection afforded by
others) plus fixed effects for both the worker (\texttt{oracle\_number})
and the year (\texttt{malaria\_year}). \(\alpha_i\) represents the time
invariant worker fixed effects, and \(\delta_i\) represents the fixed
effect of the particular malaria season. \(\upsilon\) is the error term.
This model is run separately for the 4 different worker groups, yielding
the following results:

\begin{longtable}[t]{ll}
\caption{\label{tab:unnamed-chunk-4}All absenteeism with herd immunity: model results}\\
\toprule
Term & Estimate\\
\midrule
\endfirsthead
\caption[]{All absenteeism with herd immunity: model results \textit{(continued)}}\\
\toprule
Term & Estimate\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
\hspace{1em}Malaria season & 0.034 (P<0.001)\\
\hspace{1em}IRS status=After & 0.035 (P<0.001)\\
\hspace{1em}Rainy day & 0.027 (P<0.001)\\
\hspace{1em}Herd protection & 0 (P=0.001)\\
\hspace{1em}Malaria season:IRS status=After & -0.04 (P<0.001)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
\hspace{1em}Malaria season & 0.009 (P<0.001)\\
\hspace{1em}IRS status=After & 0.005 (P=0.054)\\
\hspace{1em}Rainy day & 0.045 (P<0.001)\\
\hspace{1em}Herd protection & 0 (P<0.001)\\
\hspace{1em}Malaria season:IRS status=After & -0.015 (P<0.001)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
\hspace{1em}Malaria season & -0.003 (P<0.001)\\
\hspace{1em}IRS status=After & 0.001 (P=0.127)\\
\hspace{1em}Rainy day & 0.001 (P=0.032)\\
\hspace{1em}Herd protection & 0 (P=0.242)\\
\hspace{1em}Malaria season:IRS status=After & 0.001 (P=0.274)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
\hspace{1em}Malaria season & 0.033 (P<0.001)\\
\hspace{1em}IRS status=After & -0.007 (P=0.313)\\
\hspace{1em}Rainy day & 0.002 (P=0.409)\\
\hspace{1em}Herd protection & 0 (P=0.974)\\
\hspace{1em}Malaria season:IRS status=After & -0.007 (P=0.405)\\*
\end{longtable}

In the above, ``after'' IRS application (ie, the 6 months following IRS)
is associated with a significant reduction in absenteeism for both kinds
of permanent workers, and does not have a signficiant association with a
change in absenteeism for both kinds of temporary workers.

\section{Suggested changes from
Menno}\label{suggested-changes-from-menno}

Menno sent a write-up on May 17th (see email with subject ``suggestions
for models'') with some proposed changes to how we model absenteeism.
What follows is an implementation of those suggestions.

\subsection{Model A (without
seasonality)}\label{model-a-without-seasonality}

The below is a characterization of the approach Menno suggested:

\begin{itemize}
\tightlist
\item
  We use monthly instead of daily data (modeling a percentage
  absenteeism rate rather than a binary).
\item
  We use fixed effects for year-month combination.
\item
  We don't explicitly model for malaria season.
\item
  We don't have any interaction terms.
\item
  We use dummies for months since IRS.
\item
  We use the most recent spraying as the time since IRS (ie, if 5 months
  ago, and 1 month ago, then the dummy is applied for 1 month ago but
  not 5 months ago).
\end{itemize}

Using this approach, our panel data looks like this:

\begin{verbatim}
# A tibble: 6 x 6
  oracle_number group            year_month absent  days months_since_men~
  <chr>         <chr>            <chr>       <dbl> <int> <fct>            
1 X045753       Temporary field~ 2015_10       0      13 01               
2 X040254       Temporary field~ 2016_9        0      30 01               
3 X040343       Temporary field~ 2014_6        0      30 03               
4 PF02071       Temporary not f~ 2015_3       80.6    31 06               
5 PA00795       Permanent field~ 2013_3       67.7    31 02               
6 PF00699       Permanent not f~ 2016_3        0      31 06               
\end{verbatim}

The model formula was described in Menno's email.

\includegraphics{images/menno}{[}width=3cm{]}

Which translates into R as:

\begin{verbatim}
felm(formula = absent ~ months_since_menno | oracle_number + 
    year_month | 0 | 0, data = these_data)
\end{verbatim}

Note that the \texttt{months\_since\_menno} variable has the following
categories, and is treated as a ``wide'' dummy by the software in
modeling:

\begin{verbatim}
[1] "Before" "01"     "02"     "03"     "04"     "05"     "06"    
\end{verbatim}

The results of this model look like this:

\begin{longtable}[t]{ll}
\caption{\label{tab:unnamed-chunk-8}Simple models with year-month fixed effects}\\
\toprule
Term & Estimate\\
\midrule
\endfirsthead
\caption[]{Simple models with year-month fixed effects \textit{(continued)}}\\
\toprule
Term & Estimate\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
\hspace{1em}IRS status=\_menno01 & 1.425 (P=0.446)\\
\hspace{1em}IRS status=\_menno02 & 3.603 (P=0.066)\\
\hspace{1em}IRS status=\_menno03 & 2.049 (P=0.297)\\
\hspace{1em}IRS status=\_menno04 & 1.357 (P=0.491)\\
\hspace{1em}IRS status=\_menno05 & 1.351 (P=0.498)\\
\hspace{1em}IRS status=\_menno06 & 3.358 (P=0.106)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
\hspace{1em}IRS status=\_menno01 & -0.022 (P=0.982)\\
\hspace{1em}IRS status=\_menno02 & 0.255 (P=0.805)\\
\hspace{1em}IRS status=\_menno03 & 1.803 (P=0.085)\\
\hspace{1em}IRS status=\_menno04 & -0.87 (P=0.406)\\
\hspace{1em}IRS status=\_menno05 & 0.144 (P=0.894)\\
\hspace{1em}IRS status=\_menno06 & 0.073 (P=0.947)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
\hspace{1em}IRS status=\_menno01 & -0.333 (P=0.261)\\
\hspace{1em}IRS status=\_menno02 & -0.058 (P=0.847)\\
\hspace{1em}IRS status=\_menno03 & -0.378 (P=0.215)\\
\hspace{1em}IRS status=\_menno04 & 0.01 (P=0.973)\\
\hspace{1em}IRS status=\_menno05 & 0.075 (P=0.797)\\
\hspace{1em}IRS status=\_menno06 & 0.41 (P=0.17)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
\hspace{1em}IRS status=\_menno01 & 0.08 (P=0.979)\\
\hspace{1em}IRS status=\_menno02 & 2.989 (P=0.35)\\
\hspace{1em}IRS status=\_menno03 & -4.094 (P=0.167)\\
\hspace{1em}IRS status=\_menno04 & -2.276 (P=0.449)\\
\hspace{1em}IRS status=\_menno05 & -2.259 (P=0.464)\\
\hspace{1em}IRS status=\_menno06 & 2.409 (P=0.427)\\*
\end{longtable}

In the above, it is apparent that the effects of IRS appear to be either
(a) in the opposite direction as expected or (b) non-signficant. Here is
a visualization of IRS' apparent effect (relative to the ``Before''
period, which is the reference class). Were IRS effective, we would
expect the black line to be below 0 (the red line).

\begin{center}\includegraphics{menno_models_files/figure-latex/unnamed-chunk-9-1} \end{center}

What is going on here? Could it be that IRS has a positive (ie,
increases) impact on absenteeism? Let's look at the raw data, rather
than the coefficients. The below chart shows absenteeism (as a
percentage of the pre-IRS absenteeism average) for workers, broken down
by season (row) and worker type (column), with point size reflecting the
number of observations.

\begin{center}\includegraphics{menno_models_files/figure-latex/unnamed-chunk-10-1} \end{center}

In the above chart, two things stand out:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Variability is greatest among temporary workers.
\item
  There appears to be a reduction in absenteeism post IRS for permanent
  workers \emph{during the malaria season}, but not during the
  non-malaria season
\end{enumerate}

So, perhaps controlling for year-month through fixed effects isn't
enough, since the effect of IRS \emph{interacts} with the season\ldots{}

\subsection{Model B (with seasonality)}\label{model-b-with-seasonality}

Using this approach, our panel data looks like this:

\begin{verbatim}
# A tibble: 6 x 7
  oracle_number group      year_month absent  days months_since_me~ season
  <chr>         <chr>      <chr>       <dbl> <int> <fct>            <fct> 
1 X037697       Temporary~ 2016_7       0       31 05               low   
2 PF01761       Permanent~ 2015_12      3.23    31 04               high  
3 X040726       Temporary~ 2015_5       3.23    31 02               high  
4 PF00774       Permanent~ 2014_6       3.33    30 04               low   
5 X043249       Temporary~ 2016_4       0       30 06               high  
6 X037700       Permanent~ 2013_2       0       27 02               high  
\end{verbatim}

Our new model specification (in R) is:

\begin{verbatim}
felm(formula = absent ~ months_since_menno * season | oracle_number + 
    year_month | 0 | 0, data = these_data)
\end{verbatim}

The results of this model look like this:

\begin{longtable}[t]{ll}
\caption{\label{tab:unnamed-chunk-13}Simple models with year-month fixed effects and seasonality-IRS interaction term}\\
\toprule
Term & Estimate\\
\midrule
\endfirsthead
\caption[]{Simple models with year-month fixed effects and seasonality-IRS interaction term \textit{(continued)}}\\
\toprule
Term & Estimate\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
\hspace{1em}IRS status=\_menno01 & 175.053 (P=0.532)\\
\hspace{1em}IRS status=\_menno02 & 457.807 (P=0.115)\\
\hspace{1em}IRS status=\_menno03 & 505.425 (P=0.061)\\
\hspace{1em}IRS status=\_menno04 & 233.373 (P=0.357)\\
\hspace{1em}IRS status=\_menno05 & 227.399 (P=0.372)\\
\hspace{1em}IRS status=\_menno06 & 425.265 (P=0.121)\\
\hspace{1em}Malaria season & 887.601 (P=0.703)\\
\hspace{1em}IRS status=\_menno01:Malaria season & -58.338 (P=0.872)\\
\hspace{1em}IRS status=\_menno02:Malaria season & -174.777 (P=0.642)\\
\hspace{1em}IRS status=\_menno03:Malaria season & -603.926 (P=0.105)\\
\hspace{1em}IRS status=\_menno04:Malaria season & -220.413 (P=0.56)\\
\hspace{1em}IRS status=\_menno05:Malaria season & -213.141 (P=0.582)\\
\hspace{1em}IRS status=\_menno06:Malaria season & -191.654 (P=0.633)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
\hspace{1em}IRS status=\_menno01 & 39.635 (P=0.788)\\
\hspace{1em}IRS status=\_menno02 & 14.833 (P=0.919)\\
\hspace{1em}IRS status=\_menno03 & 114.878 (P=0.439)\\
\hspace{1em}IRS status=\_menno04 & 55.896 (P=0.694)\\
\hspace{1em}IRS status=\_menno05 & 93.301 (P=0.502)\\
\hspace{1em}IRS status=\_menno06 & 86.299 (P=0.556)\\
\hspace{1em}Malaria season & NaN (PNA)\\
\hspace{1em}IRS status=\_menno01:Malaria season & -72.876 (P=0.707)\\
\hspace{1em}IRS status=\_menno02:Malaria season & 23.995 (P=0.903)\\
\hspace{1em}IRS status=\_menno03:Malaria season & 127.34 (P=0.523)\\
\hspace{1em}IRS status=\_menno04:Malaria season & -296.411 (P=0.137)\\
\hspace{1em}IRS status=\_menno05:Malaria season & -185.639 (P=0.372)\\
\hspace{1em}IRS status=\_menno06:Malaria season & -171.932 (P=0.416)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
\hspace{1em}IRS status=\_menno01 & -35.174 (P=0.339)\\
\hspace{1em}IRS status=\_menno02 & -6.152 (P=0.869)\\
\hspace{1em}IRS status=\_menno03 & -12.472 (P=0.732)\\
\hspace{1em}IRS status=\_menno04 & 18.095 (P=0.611)\\
\hspace{1em}IRS status=\_menno05 & 33.497 (P=0.346)\\
\hspace{1em}IRS status=\_menno06 & 9.614 (P=0.816)\\
\hspace{1em}Malaria season & 44.327 (P=0.309)\\
\hspace{1em}IRS status=\_menno01:Malaria season & 8.557 (P=0.881)\\
\hspace{1em}IRS status=\_menno02:Malaria season & 3.8 (P=0.948)\\
\hspace{1em}IRS status=\_menno03:Malaria season & -75.835 (P=0.21)\\
\hspace{1em}IRS status=\_menno04:Malaria season & -51.282 (P=0.397)\\
\hspace{1em}IRS status=\_menno05:Malaria season & -70.353 (P=0.217)\\
\hspace{1em}IRS status=\_menno06:Malaria season & 63.717 (P=0.259)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
\hspace{1em}IRS status=\_menno01 & 414.405 (P=0.319)\\
\hspace{1em}IRS status=\_menno02 & 90.379 (P=0.825)\\
\hspace{1em}IRS status=\_menno03 & 353.397 (P=0.441)\\
\hspace{1em}IRS status=\_menno04 & -565.881 (P=0.232)\\
\hspace{1em}IRS status=\_menno05 & -355.43 (P=0.388)\\
\hspace{1em}IRS status=\_menno06 & 45.199 (P=0.895)\\
\hspace{1em}Malaria season & -459.147 (P=0.519)\\
\hspace{1em}IRS status=\_menno01:Malaria season & -724.867 (P=0.206)\\
\hspace{1em}IRS status=\_menno02:Malaria season & 661.492 (P=0.283)\\
\hspace{1em}IRS status=\_menno03:Malaria season & -1223.913 (P=0.031)\\
\hspace{1em}IRS status=\_menno04:Malaria season & 529.98 (P=0.364)\\
\hspace{1em}IRS status=\_menno05:Malaria season & 324.283 (P=0.578)\\
\hspace{1em}IRS status=\_menno06:Malaria season & 817.731 (P=0.227)\\*
\end{longtable}

The below shows model coefficients:

\begin{center}\includegraphics{menno_models_files/figure-latex/unnamed-chunk-14-1} \end{center}

\subsection{Model C (with seasonality and herd
score)}\label{model-c-with-seasonality-and-herd-score}

We now introduce herd immunity to the model. This is introduced as an
additive effect; perhaps it should be interactive with IRS? Using this
approach, our panel data looks like this:

\begin{verbatim}
# A tibble: 6 x 8
  oracle_number group       year_month absent  days  herd months_since_me~
  <chr>         <chr>       <chr>       <dbl> <int> <dbl> <fct>           
1 PF02492       Permanent ~ 2013_3       9.68    31 136   01              
2 X034289       Temporary ~ 2016_9       0       30 276   02              
3 PF00699       Permanent ~ 2016_9       0       30 282   03              
4 X044684       Temporary ~ 2014_9       0       30 248   01              
5 X045133       Temporary ~ 2015_8       6.45    31 156   06              
6 X041009       Temporary ~ 2015_6       0       30  97.1 02              
# ... with 1 more variable: season <fct>
\end{verbatim}

Our new model specification (in R) is:

\begin{verbatim}
felm(formula = absent ~ months_since_menno * season | oracle_number + 
    year_month | 0 | 0, data = these_data)
\end{verbatim}

The results of this model look like this:

\begin{longtable}[t]{ll}
\caption{\label{tab:unnamed-chunk-17}Simple models with year-month fixed effects, seasonality-IRS interaction term, and herd score}\\
\toprule
Term & Estimate\\
\midrule
\endfirsthead
\caption[]{Simple models with year-month fixed effects, seasonality-IRS interaction term, and herd score \textit{(continued)}}\\
\toprule
Term & Estimate\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
\hspace{1em}IRS status=\_menno01 & 175.053 (P=0.532)\\
\hspace{1em}IRS status=\_menno02 & 457.807 (P=0.115)\\
\hspace{1em}IRS status=\_menno03 & 505.425 (P=0.061)\\
\hspace{1em}IRS status=\_menno04 & 233.373 (P=0.357)\\
\hspace{1em}IRS status=\_menno05 & 227.399 (P=0.372)\\
\hspace{1em}IRS status=\_menno06 & 425.265 (P=0.121)\\
\hspace{1em}Malaria season & 887.601 (P=0.703)\\
\hspace{1em}IRS status=\_menno01:Malaria season & -58.338 (P=0.872)\\
\hspace{1em}IRS status=\_menno02:Malaria season & -174.777 (P=0.642)\\
\hspace{1em}IRS status=\_menno03:Malaria season & -603.926 (P=0.105)\\
\hspace{1em}IRS status=\_menno04:Malaria season & -220.413 (P=0.56)\\
\hspace{1em}IRS status=\_menno05:Malaria season & -213.141 (P=0.582)\\
\hspace{1em}IRS status=\_menno06:Malaria season & -191.654 (P=0.633)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
\hspace{1em}IRS status=\_menno01 & 39.635 (P=0.788)\\
\hspace{1em}IRS status=\_menno02 & 14.833 (P=0.919)\\
\hspace{1em}IRS status=\_menno03 & 114.878 (P=0.439)\\
\hspace{1em}IRS status=\_menno04 & 55.896 (P=0.694)\\
\hspace{1em}IRS status=\_menno05 & 93.301 (P=0.502)\\
\hspace{1em}IRS status=\_menno06 & 86.299 (P=0.556)\\
\hspace{1em}Malaria season & NaN (PNA)\\
\hspace{1em}IRS status=\_menno01:Malaria season & -72.876 (P=0.707)\\
\hspace{1em}IRS status=\_menno02:Malaria season & 23.995 (P=0.903)\\
\hspace{1em}IRS status=\_menno03:Malaria season & 127.34 (P=0.523)\\
\hspace{1em}IRS status=\_menno04:Malaria season & -296.411 (P=0.137)\\
\hspace{1em}IRS status=\_menno05:Malaria season & -185.639 (P=0.372)\\
\hspace{1em}IRS status=\_menno06:Malaria season & -171.932 (P=0.416)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
\hspace{1em}IRS status=\_menno01 & -35.174 (P=0.339)\\
\hspace{1em}IRS status=\_menno02 & -6.152 (P=0.869)\\
\hspace{1em}IRS status=\_menno03 & -12.472 (P=0.732)\\
\hspace{1em}IRS status=\_menno04 & 18.095 (P=0.611)\\
\hspace{1em}IRS status=\_menno05 & 33.497 (P=0.346)\\
\hspace{1em}IRS status=\_menno06 & 9.614 (P=0.816)\\
\hspace{1em}Malaria season & 44.327 (P=0.309)\\
\hspace{1em}IRS status=\_menno01:Malaria season & 8.557 (P=0.881)\\
\hspace{1em}IRS status=\_menno02:Malaria season & 3.8 (P=0.948)\\
\hspace{1em}IRS status=\_menno03:Malaria season & -75.835 (P=0.21)\\
\hspace{1em}IRS status=\_menno04:Malaria season & -51.282 (P=0.397)\\
\hspace{1em}IRS status=\_menno05:Malaria season & -70.353 (P=0.217)\\
\hspace{1em}IRS status=\_menno06:Malaria season & 63.717 (P=0.259)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
\hspace{1em}IRS status=\_menno01 & 414.405 (P=0.319)\\
\hspace{1em}IRS status=\_menno02 & 90.379 (P=0.825)\\
\hspace{1em}IRS status=\_menno03 & 353.397 (P=0.441)\\
\hspace{1em}IRS status=\_menno04 & -565.881 (P=0.232)\\
\hspace{1em}IRS status=\_menno05 & -355.43 (P=0.388)\\
\hspace{1em}IRS status=\_menno06 & 45.199 (P=0.895)\\
\hspace{1em}Malaria season & -459.147 (P=0.519)\\
\hspace{1em}IRS status=\_menno01:Malaria season & -724.867 (P=0.206)\\
\hspace{1em}IRS status=\_menno02:Malaria season & 661.492 (P=0.283)\\
\hspace{1em}IRS status=\_menno03:Malaria season & -1223.913 (P=0.031)\\
\hspace{1em}IRS status=\_menno04:Malaria season & 529.98 (P=0.364)\\
\hspace{1em}IRS status=\_menno05:Malaria season & 324.283 (P=0.578)\\
\hspace{1em}IRS status=\_menno06:Malaria season & 817.731 (P=0.227)\\*
\end{longtable}

\end{document}

### 2018-05-23 R/add_zero.R

@@ -1,0 +1,18 @@
+#' Add zero
+#'
+#' Add a zero
+#' @return character
+#' @export
+# Define function for adding zero
+add_zero <- function (x, n) {
+  x <- as.character(x)
+  adders <- n - nchar(x)
+  adders <- ifelse(adders < 0, 0, adders)
+  for (i in 1:length(x)) {
+    if (!is.na(x[i])) {
+      x[i] <- paste0(paste0(rep("0", adders[i]), collapse = ""),
+                     x[i], collapse = "")
+    }
+  }
+  return(x)
+}

### 2018-05-23 inst/rmd/paper_prepare.R

@@ -1,4 +1,5 @@
 # Packages
+library(sp)
 library(tidyverse)
 library(knitr)
 library(Hmisc)
@@ -25,19 +26,7 @@ if('prepared_data.RData' %in% dir()){
   weather <- maragra::weather
   workers <- maragra::workers
   
-  # Define function for adding zero
-  add_zero <- function (x, n) {
-    x <- as.character(x)
-    adders <- n - nchar(x)
-    adders <- ifelse(adders < 0, 0, adders)
-    for (i in 1:length(x)) {
-      if (!is.na(x[i])) {
-        x[i] <- paste0(paste0(rep("0", adders[i]), collapse = ""),
-                       x[i], collapse = "")
-      }
-    }
-    return(x)
-  }
+  # Create model data
   
   model_data <-
     ab_panel %>%
@@ -147,7 +136,6 @@ if('prepared_data.RData' %in% dir()){
   
   # Estimate a protection factor based on the weighted protection 
   # scores of nearby houses
-  library(sp)
   dates <- sort(unique(model_data$date))
   out_list <- list()
   
@@ -290,10 +278,10 @@ if('prepared_data.RData' %in% dir()){
   #                         'Non-field worker',
   #                         group))
   
-  # REMOVE THE NEVERS
-  model_data <- model_data %>%
-    filter(ever_sprayed)
-  
+  # # REMOVE THE NEVERS
+  # model_data <- model_data %>%
+  #   filter(ever_sprayed)
+  # 
   fe_models <- list()
   sick_models <- list()
   protection_models <- list()
@@ -572,7 +560,7 @@ if('prepared_data.RData' %in% dir()){
   save.image(file = 'prepared_data.RData')
   }
 
-clean_up_model <- function(x){
+clean_up_model <- function(x, multiplier = 100){
   # extract coefficients
   coefs <- data.frame(coef(summary(x)))
   # use normal distribution to approximate p-value
@@ -580,7 +568,7 @@ clean_up_model <- function(x){
   ps <- coefs$p.z
   x = broom::tidy(x)
   for(j in 2:(ncol(x) -1)){
-    x[,j] <- x[,j] * 100
+    x[,j] <- x[,j] * multiplier
   }
   x$term <- gsub('seasonhigh', 'Malaria season', x$term)    
   x$term <- gsub('seasonlow', 'Low malaria season', x$term)
@@ -613,14 +601,14 @@ clean_up_model <- function(x){
   return(x)
 }
 
[…]

### 2018-07-30 data-raw/create_data_files.R
minor update

@@ -285,6 +285,9 @@ if('ab_expanded.RData' %in% dir()){
   }
   ab_expanded <- bind_rows(ab_expanded)
   ab_expanded$absent <- TRUE
+  # Remove any potential duplicates
+  ab_expanded <- ab_expanded %>%
+    dplyr::distinct(oracle_number, date, .keep_all = TRUE)
   save(ab_expanded,
        file = 'ab_expanded.RData')
 }
@@ -314,7 +317,9 @@ ab_panel <-
   # and anything in 2017, which also appears untrustable
   dplyr::filter(date <= '2016-12-31')
 
-
+# Remove any potential duplicates
+ab_panel <- ab_panel %>%
+  dplyr::distinct(oracle_number, date, .keep_all = TRUE)
 
 # DSS bairros data
 library(rgdal)
@@ -1048,12 +1053,12 @@ mc <-
 expand_irs <- function(date,
                        chemical,
                        unidade){
-  data_frame(date = seq(date - 364,
-                        date + 364,
+  data_frame(date = seq(date - 730,
+                        date + 730,
                         by = 1),
              chemical = chemical,
              unidade = unidade,
-             days_since = -364:364)
+             days_since = -730:730)
 }
 results <- list()
 for (i in 1:nrow(mc)){
@@ -1061,7 +1066,7 @@ for (i in 1:nrow(mc)){
                     chemical = mc$insecticida[i],
                     unidade = mc$unidade[i])
   results[[i]] <- out
-  message(i)
+  # message(i)
 }
 
 # Combine
@@ -1092,31 +1097,31 @@ joined <- full_join(after, before)
 joined <- joined %>%
   arrange(unidade, date)
 
-# Flag a period which is both before and after (not suitable for modeling)
-joined <-
-  joined %>%
-  mutate(ok_for_model =
-           is.na(days_until) |
-           is.na(days_since))
-irs <- joined; rm(joined)
+# If both before and after, but greater than 270 days after, use the before
+cutter <- 548
 
-# Remove the not ok ones and calculate days
-irs <- irs %>%
-  filter(ok_for_model) %>%
-  mutate(days = ifelse(is.na(days_since), days_until,
-                       ifelse(is.na(days_until), days_since, NA))) %>%
-  mutate(chemical = ifelse(is.na(days_since), chemical_before,
-                           ifelse(is.na(days_until), chemical_after, NA))) %>%
-  mutate(days_since = days) %>%
-  dplyr::select(date, unidade, days_since, chemical)
+x <- joined %>%
[…]

### 2018-07-30 inst/rmd/dummy.R

@@ -1,0 +1,37 @@
+source('paper_prepare.R')
+dummy_data <- read_csv('dummy_data.csv')
+dummy_data$year_month <- paste0(dummy_data$calendar_year, dummy_data$calendar_month)
+
+# Create a model per menno's specifications
+menno_models <- list()
+
+# dummy_data <- dummy_data %>%
+#   # filter(!is.na(days_since)) %>%
+#   # filter(days_since >= -184, days_since <= 184) %>%
+#   group_by(oracle_number, group, year_month) %>%
+#   summarise(absent = length(which(absent)) / length(absent) * 100,
+#             days = n(),
+#             herd = mean(herd, na.rm = TRUE),
+#             months_since_menno = dplyr::first(months_since_menno),
+#             season = dplyr::first(season)) %>%
+#   ungroup
+dummy_data <- read_csv('dummy_data_collapsed.csv')
+groups <- sort(unique(dummy_data$group))
+# library(lmerTest)
+# library(nlme)
+for (i in 1:length(groups)){
+  message(i)
+  this_group <- groups[i]
+  message(this_group)
+  these_data <- dummy_data %>% filter(group == this_group)
+  these_data$months_since_menno <- 
+    factor(these_data$months_since_menno,
+           levels = c('Before',
+                      paste0(0, 1:6)))
+  this_menno_model <- felm(absent ~ months_since_menno    | oracle_number + year_month | 0 | 0,
+                           data = these_data)
+  
+  menno_models[[i]] <- this_menno_model
+}
+# write_csv(dummy_data, 'dummy_data_collapsed.csv')
+names(menno_models) <- groups

### 2018-07-30 inst/rmd/paper_prepare.R

@@ -585,6 +585,9 @@ clean_up_model <- function(x, multiplier = 100){
   x$term <- gsub('on_siteFALSE', 'Off site', x$term)
   x$term <- gsub('fieldNot field worker', 'Not field worker', x$term)
   x$term <- gsub('herd', 'Herd protection', x$term)
+  x$term <- gsub('before_after', 'IRS time=', x$term)
+  x$term <- gsub('time_since', 'Months after=', x$term)
+  x$term <- gsub('incidence', 'Incidence', x$term)
   names(x) <- Hmisc::capitalize(names(x))
   names(x) <- gsub('.', ' ', names(x), fixed = TRUE)
   x$`P value` <- NA
@@ -672,4 +675,77 @@ model_data <-
 
 if(!'model_data.csv' %in% dir()){
   write_csv(model_data, 'model_data.csv')
+}
+if(!'dummy_data.csv' %in% dir()){
+  ids <- sample(unique(model_data$oracle_number), 200)
+  dummy_data <- model_data %>%
+    filter(oracle_number %in% ids) %>%
+    mutate(oracle_number = as.numeric(factor(oracle_number)))
+  dummy_data$year_month <- paste0(dummy_data$calendar_year, dummy_data$calendar_month)
+  
+  dummy_data <- dummy_data %>%
+    dplyr::select(oracle_number,
+                  date,
+                  absent,
+                  incidence,
+                  season,
+                  permanent_or_temporary,
+                  department,
+                  sex,
+                  precipitation,
+                  rainy_day,
+                  months_since,
+                  ever_sprayed,
+                  group,
+                  calendar_year,
+                  calendar_month,
+                  year_month,
+                  malaria_year,
+                  days_since,
+                  herd,
+                  months_since_menno)
+  write_csv(dummy_data, 'dummy_data.csv')
+  dummy_dictionary <- data_frame(
+    variable = c('oracle_number',
+                 'date',
+                 'absent',
+                 'incidence',
+                 'season',
+                 'permanent_or_temporary',
+                 'department',
+                 'sex',
+                 'precipitation',
+                 'rainy_day',
+                 'months_since',
+                 'ever_sprayed',
+                 'group',
+                 'calendar_year',
+                 'calendar_month',
+                 'malaria_year',
+                 'days_since',
+                 'herd',
+                 'months_since_menno'),
+    meaning = c('Unique ID Number',
+                'The calendar date of the observation',
+                'Whether the worker was absent or present',
[…]

### 2019-02-17 inst/rmd/final/final.Rmd
minor update

---
title: "Final analysis for paper"
subtitle: "Brew, Pradhan, Sicuri"
date: ''
output: 
  pdf_document:
    keep_tex: true
    latex_engine: xelatex
    includes:
      in_header: preamble.sty
fig_width: 4
fig_height: 2.6
bibliography: bibliography.bib
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# csl: journal-of-health-economics.csl

# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(RColorBrewer)
library(extrafont)
library(kableExtra)
library(raster)
loadfonts()
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               fig.height = 4,
               fig.width = 5,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               dev = "cairo_pdf",
               fig.pos="!h",
               fig.align = 'center')
opts_knit$set(width=75)
options(xtable.comment = FALSE)
```

```{r, eval = TRUE}
source('prepare_data.R')
```

# Manuscript draft

Draft of the paper (\href{https://docs.google.com/document/d/1bUWRBCgVcgjSPHchIQxiTG8Vwv5hV1GLU4Tlu386sWA/edit#}{HERE}

# Descriptive overview

## Herd-only protection score

```{r}
pq <- quantile(final_data$herd_protection)
x <- final_data %>%
  # mutate(protection_cat = protection_var) %>%
  mutate(protection_cat = ifelse(herd_protection <= pq[2],
                                 'Protection 1',
                                 ifelse(herd_protection <= pq[3],
                                        'Protection 2',
                                        ifelse(herd_protection <= pq[4],
                                               'Protection 3',
                                               ifelse(herd_protection <= pq[5],
                                                      'Protection 4',
                                                      NA))))) %>%
  group_by(protection_cat) %>%
  summarise(ab_rate = length(which(absent))/ n() * 100)

names(x) <- c('Quantile', 'Absenteeism rate')
kable(x)
```

```{r}
ggplot(data = x,
       aes(x = `Quantile`,
           y = `Absenteeism rate`)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(x = 'Protection score quantile')
```

\newpage

## Individual-only protection score

```{r}
pq <- quantile(final_data$protection,
               seq(0, 1, 0.1))
pq <- sort(unique(pq))
x <- final_data %>%
  # mutate(protection_cat = protection_var) %>%
  mutate(protection_cat = ifelse(protection <= pq[2],
                                 'Protection 1',
                                 ifelse(protection <= pq[3],
                                        'Protection 2',
                                        ifelse(protection <= pq[4],
                                               'Protection 3',
                                               ifelse(protection <= (pq[5] + 0.1),
                                                      'Protection 4',
                                                      NA))))) %>%
  group_by(protection_cat) %>%
  summarise(ab_rate = length(which(absent))/ n() * 100)

names(x) <- c('Quantile', 'Absenteeism rate')
kable(x)
```

```{r}
ggplot(data = x,
       aes(x = `Quantile`,
           y = `Absenteeism rate`)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(x = 'Protection score quantile')
```

\newpage

## Overall protection score

```{r}
pq <- quantile(final_data$protection_var)
x <- final_data %>%
  # mutate(protection_cat = protection_var) %>%
  mutate(protection_cat = ifelse(protection_var <= pq[2],
                                 'Protection 1',
                                 ifelse(protection_var <= pq[3],
                                        'Protection 2',
                                        ifelse(protection_var <= pq[4],
                                               'Protection 3',
                                               ifelse(protection_var <= pq[5],
                                                      'Protection 4',
                                                      NA))))) %>%
  group_by(protection_cat, group) %>%
  summarise(ab_rate = length(which(absent))/ n() * 100) %>%
  ungroup

names(x) <- c('Quantile', 'Group', 'Absenteeism rate')
kable(x)
```

```{r}
ggplot(data = x,
       aes(x = `Quantile`,
           y = `Absenteeism rate`)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(x = 'Protection score quantile') +
  facet_wrap(~Group)
```

\newpage

## Precipitation lag

```{r}
rq <- quantile(final_data$rain_var, na.rm = TRUE)
x <- final_data %>%
  # mutate(protection_cat = protection_var) %>%
  mutate(protection_cat = ifelse(rain_var <= rq[2],
                                 'Protection 1',
                                 ifelse(rain_var <= rq[3],
                                        'Protection 2',
                                        ifelse(rain_var <= rq[4],
                                               'Protection 3',
                                               ifelse(rain_var <= rq[5],
                                                      'Protection 4',
                                                      NA))))) %>%
  group_by(protection_cat, group) %>%
  summarise(ab_rate = length(which(absent))/ n() * 100) %>%
  ungroup %>%
  filter(!is.na(protection_cat))

names(x) <- c('Quantile', 'Group', 'Absenteeism rate')
kable(x)
```

```{r}
ggplot(data = x,
       aes(x = `Quantile`,
           y = `Absenteeism rate`)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(x = 'Lagged precipitation score quantile') +
  facet_wrap(~Group)
```

\newpage

# Results

## Regression table
The below table shows the results of the model devised thus far.

```{r, results = 'asis'}
# make_final_table(model = final_model, the_caption = "Model results", type = 'latex', multiplier = 10000)
```

```{r}
summary(final_model)
```

## Intepretation

The below charts show predicted absenteeism rates at different rain and protection levels.

```{r}
fake <- expand.grid(#group = sort(unique(monthly$group)),
  rain_var = rq,
  protection_var = pq)#,
fake$predicted <- fake$predicted_upr <- fake$predicted_lwr<- NA

fe <- getfe(final_model)
predictions <- exp(predict(final_model_lm, fake, interval = 'confidence'))-1
# predictions <- data.frame(predictions)
# # Add the fixed effects
# fixed_effects <- fe$effect

fake$predicted<- predictions[,1]
fake$predicted_upr <- predictions[,3]
fake$predicted_lwr <- predictions[,2]
```

### Precipitation's effect

```{r}
ggplot(data = fake %>% filter(protection_var == pq[1]),
       aes(x = rain_var,
           y = predicted)) +
  geom_point() +
  geom_line() +
  geom_ribbon(aes(ymin = predicted_lwr,
                  ymax = predicted_upr),
              alpha = 0.4,
              fill = 'darkorange') +
  labs(y = 'Predicted absenteeism rate',
       x = '1-3 month precipitation lag',
       title = 'Lagged precipitation and absenteeism')
```

### IRS protection's effect

```{r}
ggplot(data = fake %>% filter(rain_var == rq[1]),
       aes(x = protection_var,
           y = predicted)) +
  geom_point() +
  geom_line() +
  geom_ribbon(aes(ymin = predicted_lwr,
                  ymax = predicted_upr),
              alpha = 0.4,
              fill = 'darkorange') +
  labs(y = 'Predicted absenteeism rate',
       x = 'Protection index',
       title = 'IRS-afforded protection and absenteeism')
```

### Simulations

```{r}
sim_data <- simulations
sim_data$idx <- sim_data$oracle_number
fe <- getfe(final_model)
predictions <- exp(predict(final_model_lm, sim_data, interval = 'confidence'))-1
predictions <- data.frame(predictions)
# Add the fixed effects
fixed_effects <- left_join(sim_data,
                           data.frame(fe) %>% dplyr::select(idx, effect) %>%
                             mutate(idx = as.character(idx))) %>%
  .$effect
fixed_effects <- as.numeric(fixed_effects)

predictions$fit <- predictions$fit + fixed_effects
predictions$lwr <- predictions$lwr + fixed_effects
predictions$upr <- predictions$upr + fixed_effects

# combine other info
predictions <- predictions %>% bind_cols(sim_data %>% dplyr::select(date, absent, strategy)) %>%
  mutate(absent = as.numeric(absent)) %>%
  filter(!is.na(fit))

simple <- predictions %>%
  group_by(strategy) %>%
  summarise(n = n(),
            actual_absences = sum(absent),
            predicted_absences = sum(fit, na.rm = TRUE),
            actual = mean(absent, na.rm = TRUE),
            predicted = mean(fit, na.rm = TRUE)) %>%
  dplyr::rename(value = predicted) %>%
  dplyr::select(strategy, value, predicted_absences)

kable(simple)
```

```{r}
ggplot(data = simple,
       aes(x = strategy,
           y = value)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(y = 'Absenteeism rate',
       x = 'Strategy') +
  geom_label(aes(label = round(value, digits = 3)),
             nudge_y = -0.005)
```

```{r}
ggplot(data = simple,
       aes(x = strategy,
           y = predicted_absences)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(y = 'Absenteeism rate',
       x = 'Strategy') 
```

# Costs

```{r}
x <- final_data %>%
  left_join(irs %>% dplyr::select(date, unidade, days_since))

out <- data.frame(
  strategy = c('Max',
               'Real',
               'Time-optimized',
               'Zero'),
  sprays = c(nrow(final_data) / 60,
             length(which(x$days_since == 0)),
             length(which(format(final_data$date, '%m-%d') == '12-11')),
             0
  )
) %>%
  mutate(sprays = sprays / length(unique(final_data$oracle_number)))

kable(out)
ggplot(data = out,
       aes(x = strategy,
           y = sprays)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(y = 'Number of sprays per person',
       x = 'Strategy')

```

```{r}
# Introduction is really all that matters - specifying contribution, summarizing results
# Journal of Health Economics (1) and then Health Economics (2).
# Elisa: American Journal of Health Economics (from American Health Economic Society)
# Ensure that we explicitly explain these 21 days issue to rule out reverse causality.
# Large firms are big enough to capture externalities
# What is the effect of anti-malaria campaign on worker producitivity
# Data special
# Identification problems (no external spraying, no home spraying, unobserved govt - omitted variable bias dealt with through worker and time fixed effects.)
# Exploring herd protection through various functional forms
```

# Bibliography

### 2019-02-17 inst/rmd/final/prepare_data.R

@@ -1,0 +1,663 @@
+# Packages
+library(sp)
+library(tidyverse)
+library(knitr)
+library(Hmisc)
+library(brew)
+library(maragra)
+library(knitr)
+library(RColorBrewer)
+# library(lme4)
+library(lfe)
+library(restorepoint)
+library(regtools)
+
+
+classify_protection <- function(days_since){
+  ifelse(days_since < 21 | days_since > (183+21),
+         0,
+         ifelse(days_since < 51, 1,
+                ifelse(days_since < 81,
+                       0.9,
+                       ifelse(days_since < 112,
+                              0.8,
+                              ifelse(days_since < 143,
+                                     0.6,
+                                     ifelse(days_since <174,
+                                            0.4,
+                                            ifelse(days_since <= (183+21),
+                                                   0.2,
+                                                   0)))))))
+}
+
+
+if('prepared_data.RData' %in% dir()){
+  load('prepared_data.RData')
+} else {
+  # Read in data
+  ab <- maragra::ab
+  ab_panel <- maragra::ab_panel
+  bairros <- maragra::bairros
+  bes <- maragra::bes
+  census <- maragra::census
+  clinic <- maragra::clinic
+  clinic_agg <- maragra::clinic_agg
+  mc <- maragra::mc
+  workers <- maragra::workers
+  
+  if(!'weather.RData' %in% dir()){
+    weather <- maragra::weather %>%
+    mutate(precipitation = ifelse(is.na(precipitation),
+                                  0,
+                                  precipitation))
+    
+    # Get lags for weather
+    starts <- c(15, 30, 60, 90)
+    ends <- c(30, 60, 90, 120)
+    for(k in 1:length(starts)){
+      for(j in 1:length(ends)){
+        this_start <- starts[k]
+        this_end <- ends[j]
+        if(this_start < this_end){
+          precipitation_column_name <- paste0('precipitation_lag_', this_start, '_', this_end)
+          weather[,precipitation_column_name] <- NA
+          temp_column_name <- paste0('temp_lag_', this_start, '_', this_end)
+        }
+      }
+    }
+    for(i in 1:nrow(weather)){
+      starts <- c(15, 30, 60, 90)
+      ends <- c(30, 60, 90, 120)
+      for(k in 1:length(starts)){
+        for(j in 1:length(ends)){
+          this_start <- starts[k]
+          this_end <- ends[j]
+          if(this_start < this_end){
+            this_date <- weather$date[i]
[…]

### 2020-04-18 inst/rmd/final/final.Rmd
update

@@
 title: "Final analysis for paper"
-subtitle: "Brew, Pradhan, Sicuri"
 date: ''
+fig_height: 2.6
 output:
   pdf_document:
-    keep_tex: true
-    latex_engine: xelatex
     includes:
       in_header: preamble.sty
+    keep_tex: yes
+    latex_engine: xelatex
+  html_document:
+    df_print: paged
 fig_width: 4
-fig_height: 2.6
+always_allow_html: true
+subtitle: Brew, Pradhan, Sicuri
 bibliography: bibliography.bib
@@
                warning=FALSE,
-               dev = "cairo_pdf",
+               # dev = "cairo_pdf",
                fig.pos="!h",
                fig.align = 'center')
+knitr::opts_chunk$set(error = FALSE)
+
 opts_knit$set(width=75)
@@
 
+# Calculate household protection
+
+```{r}
+
+```
+
 # Manuscript draft
 
-Draft of the paper (\href{https://docs.google.com/document/d/1bUWRBCgVcgjSPHchIQxiTG8Vwv5hV1GLU4Tlu386sWA/edit#}{HERE}
+Draft of the paper \href{https://docs.google.com/document/d/1bUWRBCgVcgjSPHchIQxiTG8Vwv5hV1GLU4Tlu386sWA/edit#}{HERE}.
+
+# Classify protection plot
+
+```{r}
+pd <- data_frame(days_since = 1:230)
+pd$protection <- classify_protection(pd$days_since)
+
+ggplot(data = pd,
+       aes(x = days_since,
+           y = protection)) +
+  geom_area(fill = 'lightblue', alpha = 0.8,
+            lwd = 0.2,
+            color = 'black') +
+  geom_vline(xintercept = 0,
+             lty = 2, alpha = 0.6) +
+  theme_maragra() +
+  labs(x = 'Days since IRS application',
+       y = 'Protection score')
+```
 
 # Descriptive overview
+
+```{r}
+desc <- final_data %>%
+  group_by(permanent_or_temporary) %>%
+  summarise(workers = length(unique(oracle_number)),
+            days = n())
+```
 
@@
 ```{r}
-pq <- quantile(final_data$protection,
-               seq(0, 1, 0.1))
-pq <- sort(unique(pq))
+pq <- sort(unique(final_data$protection))
+# pq <- quantile(final_data$protection,
+#                seq(0, 1, 0.1))
+# pq <- sort(unique(pq))
 x <- final_data %>%
@@
   geom_bar(stat = 'identity',
-           alpha = 0.6,
-           color = 'black') +
+           alpha = 0.8,
+           color = 'black',
+           lwd = 0.2,
+           fill = 'lightblue') +
   labs(x = 'Protection score quantile') +
-  facet_wrap(~Group)
+  facet_wrap(~Group) +
+  theme_maragra()
 ```
@@
 ```{r}
+# Combined protection
 summary(final_model)
+
 ```
@@
 ```{r}
+# on <- sort(unique(final_data$oracle_number))
 fake <- expand.grid(#group = sort(unique(monthly$group)),
@@
   protection_var = pq)#,
-fake$predicted <- fake$predicted_upr <- fake$predicted_lwr<- NA
+fake$predicted <- NA #fake$predicted_upr <- fake$predicted_lwr<- NA
 
 fe <- getfe(final_model)
+# # fake <- left_join(fake, fe %>% dplyr::select(idx, effect, obs), 
+# #                                              by = c('oracle_number'= 'idx'))
 predictions <- exp(predict(final_model_lm, fake, interval = 'confidence'))-1
 # predictions <- data.frame(predictions)
-# # Add the fixed effects
-# fixed_effects <- fe$effect
-
+# # # Add the fixed effects
+fixed_effects <- fe$effect
+# 
 fake$predicted<- predictions[,1]
@@
 fake$predicted_lwr <- predictions[,2]
+# 
+# # # Add fixed effects
+# # fake <- fake %>%
+# #   mutate(predicted = predicted + effect,
+# #          predicted_upr = predicted_upr + effect,
+# #          predicted_lwr = predicted_lwr + effect)
+# 
+# pd <- fake %>%
+#   filter(protection_var == pq[1]) 
+# ggplot(data = pd,
+#        aes(x = rain_var,
+#            y = predicted)) +
+#   geom_point()
+
 ```
@@
 
-```{r}
-ggplot(data = fake %>% filter(protection_var == pq[1]),
+```{r, eval = FALSE}
+g1 <- ggplot(data = fake %>% filter(protection_var == pq[1]),
        aes(x = rain_var,
@@
        x = '1-3 month precipitation lag',
-       title = 'Lagged precipitation and absenteeism')
-```
+       title = 'Lagged precipitation and absenteeism') +
+  theme_maragra()
 
-### IRS protection's effect
-
-```{r}
-ggplot(data = fake %>% filter(rain_var == rq[1]),
+g2 <- ggplot(data = fake %>% filter(rain_var == rq[1]),
        aes(x = protection_var,
@@
        x = 'Protection index',
-       title = 'IRS-afforded protection and absenteeism')
+       title = 'IRS-afforded protection and absenteeism') +
+  theme_maragra()
+
+Rmisc::multiplot(g1, g2, cols = 2)
 ```
@@
 # combine other info
-predictions <- predictions %>% bind_cols(sim_data %>% dplyr::select(date, absent, strategy)) %>%
+predictions <- predictions %>% bind_cols(sim_data %>% dplyr::select(date, absent, strategy, permanent_or_temporary)) %>%
   mutate(absent = as.numeric(absent)) %>%
@@
 
+zero <- predictions %>%
+  filter(strategy == 'Zero') %>%
+  mutate(year = as.numeric(format(date, '%Y'))) %>%
+  group_by(year) %>%
+  summarise(n = n(),
+            actual_absences = sum(absent),
+            predicted_absences = sum(fit, na.rm = TRUE),
+            actual = mean(absent, na.rm = TRUE),
+            predicted = mean(fit, na.rm = TRUE)) %>%
+  mutate(avoided = predicted_absences - actual_absences)
+  # dplyr::rename(value = predicted) %>%
+  # dplyr::select(permanent_or_temporary, value, predicted_absences)
+
 kable(simple)
+
+# Eligible working days
+sum(zero$n)
+# Absences
+sum(zero$actual_absences)
+# Actual absence rate
+sum(zero$actual_absences) / sum(zero$n)
+# Would have observed
+sum(zero$predicted_absences)
+# Predicted - observ
+sum(zero$predicted_absences) - sum(zero$actual_absences)
+# Overall reduction from
+sum(zero$predicted_absences) / sum(zero$n)
+# To
+sum(zero$actual_absences) / sum(zero$n)
+# Number of working days permanent vs temp
+table(predictions$permanent_or_temporary[predictions$strategy=='Real'])
+# Absences avoided among permanent workers
+predictions %>% filter(strategy == 'Zero') %>% group_by(permanent_or_temporary) %>% summarise(n = n(),
+            actual_absences = sum(absent),
+            predicted_absences = sum(fit, na.rm = TRUE),
+            actual = mean(absent, na.rm = TRUE),
+            predicted = mean(fit, na.rm = TRUE)) %>%
+  mutate(avoided = predicted_absences - actual_absences) %>%
+  # decline due to irs
+  mutate(decline = predicted - actual)
+# One absence every n days
+sum(zero$n) / sum(zero$avoided)
+# Annual reduction in absence
+mean(zero$avoided[zero$year < 2016])
+# Cost per absence avoided
+68984 / 4183
+# Percentage of absence among permanent workers
+x = predictions %>%
+  filter(strategy == 'Zero') %>%
+  # filter(date < '2016-01-01') %>%
+  group_by(permanent_or_temporary) %>%
+  summarise(absences = sum(absent),
+            predicted = sum(fit)) %>%
+  mutate(averted = predicted - absences) %>%
+  mutate(p = averted / sum(averted))
+x
+# Weighted average cost
+y = sum(c(25, 5) * x$p)
+y
+# Cost of treatment per absence
+22 / 80
+# Total savings per aversion
+0.275 + y
+# Total yearly savings
+(0.275 + y) * 4183
+# Total yearly savings
+((0.275 + y) * 4183) - 68984
+# ROI
+((0.275 + y) * 4183) / 68984
+# Additional max strategy
+simple
+47943-42131
 ```
@@
 ```{r}
+# paste0(
+#   "Of the ", nrow(simulations[simulations$strategy == 'Real',]),
+#   " eligible working days in our panel, we observed ", nrow(simulations[simulations$strategy == 'Real' & simulations$absent,]), 
+#   " absences. Using the above approach, we estimate that we would have observed ",
+#   length(which(simulations$absent[simulations$strategy == 'Zero'])),
+#   " had it not been for the firm's IRS program, for a total of ",
+#   length(which(simulations$absent[simulations$strategy == 'Zero'])) - length(which(simulations$absent[simulations$strategy == 'Real'])),
+#   " avoided absences. This translates to an overall reduction in absenteeism from ",
+#   round(length(which(simulations$absent[simulations$strategy == 'Zero'])) / nrow(simulations %>% filter(strategy == 'Zero')) * 100, digits = 2),
+#   "% (without its IRS program) to the observed ", round(length(which(final_data$absent)) / nrow(final_data) * 100, digits = 2),"%.
+# The number of working-days observed among permanent workers (",
+# length(which(final_data$permanent_or_temporary == 'Permanent')),") was nearly thrice that of those of temporary workers (", 
+# length(which(final_data$permanent_or_temporary == 'Temporary')),
+# ").") 
+# 
+# 
+# 
+# paste0(
+#   "By the same token, the number of absences avoided among permanent workers (39,890) was far greater than that of temporary workers (8,972).
+The latter had a greater relative decline in absenteeism due to IRS (from 9.3% to 3.91%) relative to the former (from 14.2% to 11%), but the aggregate numbers are less consequential since (a) temporary workers have a lower baseline absenteeism rate and (b) they work, by definition, fewer days.
+In total, the IRS program is estimated to prevent one absence every 26.6 working days.
+# Annually, this translates to approximately 4,076 avoided absences per year (slightly greater than the simple splicing of the overall figure into the 4-year study period, since we did not have complete data for all worker types).
+# "
+# )
+```
+
+```{r}
 ggplot(data = simple,
@@
   labs(y = 'Absenteeism rate',
-       x = 'Strategy') 
+       x = 'Strategy') +
+  geom_label(aes(label = round(predicted_absences, digits = 3)),
+             nudge_y = -10000)
 ```
@@
 
-```{r}
+```{r, eval = FALSE}
 x <- final_data %>%
@@
 
+# Charts for paper
+
+```{r, results='asis'}
+# x <- ab_panel %>%
+#   left_join(workers %>%
+#               dplyr::select(oracle_number, date_of_birth, permanent_or_temporary, sex))
+x <- final_data
+x <- x %>% left_join(workers %>% dplyr::select(oracle_number, date_of_birth))
+x$age <-as.numeric(as.Date('2014-01-01') - x$date_of_birth) / 365.25
+x$agey <- x$age
+x$age <- round(x$age, digits = -1)
+x$status <- x$permanent_or_temporary
+
+calculate <- function(var= 'sex',
+                      data){
+  out <- 
+    data %>%
+    group_by_(var) %>%
+    summarise(`Number of workers` = length(unique(oracle_number)),
+              `Number of days observed` = n()) %>%
+    # summarise(`Treatment workers` = length(unique(oracle_number[!months_since %in% c('Never', 'Before')])),
+    #           `Treatment days` = length(oracle_number[!months_since %in% c('Never', 'Before')]),
+    #           `Control workers` = length(unique(oracle_number[months_since %in% c('Never', 'Before')])),
+    #           `Control days` = length(oracle_number[months_since %in% c('Never', 'Before')])) %>%
+    ungroup() %>%
+    # mutate(p = round(n / sum(n) * 100, digits = 2)) %>%
+    mutate(variable = '') 
+    names(out)[1] <- ' '
+    out$variable[1] <- Hmisc::capitalize(var)
+
+    out <- out %>% dplyr::select(variable, ` `, 
+                                 `Number of workers`, `Number of days observed`)
+                                 # `Treatment workers`, `Treatment days`, `Control workers`, `Control days`)
+    out$` ` <- as.character(out$` `)
+  return(out)
+}
+out_list <- list()
+vars <- c('department', 
+           'sex',
+           'status',
+           'age')
+for(i in 1:length(vars)){
+  message(i)
+  y <- calculate(var = vars[i], data = x)
+  out_list[[i]] <- y
+}
+demo_table <- bind_rows(out_list)
+names(demo_table)[1:2] <- c('Variable', ' ')
+kable(demo_table, format = 'html') %>% kable_styling()
+# print(xtable::xtable(demo_table,
+#                      caption = 'Overall worker characteristics',
+#                      align = c('r', 'r', rep('l', 5))),
+#       include.rownames = FALSE)
+
+```
+
+```{r, fig.width = 6, fig.height = 3.5, fig.align='center', fig.cap = 'i.
+Monthly total precipitation in the Maniça district; ii.
+Average monthly temperature (bars) during the study period, as well as monthly maximum (red) and minimum (blue) temperatures'}
+x <- weather %>%
+  mutate(month = format(date, '%m')) %>%
+  mutate(year = format(date, '%Y')) %>%
+  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
+  group_by(date) %>%
+  summarise(precipitation = sum(precipitation, na.rm = TRUE),
+            temp_max = max(temp_max, na.rm = TRUE),
+            temp_min = min(temp_min, na.rm = TRUE),
+            temp = mean(temp))
+# x <- gather(x, key, value,precipitation:temp)
+
+g1 <- ggplot(data = x,
+       aes(x = date,
+           y = precipitation)) +
+  geom_bar(stat = 'identity',
+           fill = 'lightblue',
+           size = 0.1) +
+  theme_maragra() +
+  theme(axis.text.x = element_text(angle = 0)) +
+  labs(x = 'Month',
+       y = 'Milimeters',
+       title = 'A. Monthly precipitation') +
+  scale_x_date(labels = scales::date_format("%Y"))
+
+g2 <- ggplot(data = x,
+            aes(x = date,
+                y = temp)) +
+  # geom_bar(stat = 'identity', fill = 'grey', size = 0.1) +
+  geom_line(aes(y = temp_max),
+             color = 'darkred', 
+             alpha = 0.5) +
+  geom_line(aes(y = temp_min),
+             color = 'blue', 
+             alpha = 0.5) +
+  # geom_linerange(aes(ymin = temp_min,
+  #                    ymax = temp_max),
+  #                alpha = 0.6,
+  #                color = 'blue') +
+   theme_maragra() +
+  theme(axis.text.x = element_text(angle = 0)) +
+  labs(x = 'Month',
+       y = 'Degrees (Celcius)',
+       title = 'B. Monthly temperature') +
+  scale_x_date(labels = scales::date_format("%Y"))
+Rmisc::multiplot(g1, g2, cols = 2)
+```
+
+```{r, fig.cap = 'Clinical malaria (district of Manhiça), all-cause absenteeism among Maragra workers, sick absenteeism among Maragra workers, positive cases at company clinic, and test positivity rate at company clinic', fig.height=5, fig.width=8}
+# BES incidence
+g1 <- 
+  ggplot(data = bes %>%
+           mutate(season = ifelse(p >= median(p), 'High season',
+                                  'Low season')) %>%
+           filter(weekdays(date) == 'Monday',
+                  date <= '2016-12-01'),
+       aes(x = date,
+           y = p)) +
+  geom_point(alpha = 0.6,
+             aes(color = season),
+             size = 0.2) +
+  geom_line(alpha = 0.1) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Incidence',
+       title = 'A: Annualized district clinical\nmalaria incidence per 1,000') +
+  scale_color_manual(name = '',
+                     values = c('red', 'black')) + 
+  guides(color = guide_legend(ncol = 1)) +
+  theme(legend.position = c(.8, .8),
+        legend.text = element_text(size = 5),
+        legend.background = element_rect(fill=alpha('white', 0))) +
+  theme(axis.title.x = element_blank())
+
+# Absenteeism rate
+g2_data <- final_data %>%
+  # filter(date <= '2016-12-01') %>%
+  group_by(month = as.Date(paste0(format(date, '%Y-%m'), '-01')), permanent_or_temporary) %>%
+  summarise(absences = length(which(absent)),
+            # absences_sick = length(which(absent_sick)),
+
+            eligibles = n()) %>%
+  mutate(absenteeism_rate = absences / eligibles * 100)
+
+g2 <- ggplot(data = g2_data,
+             aes(x = month,
+                 y = absenteeism_rate,
+                 lty = permanent_or_temporary)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Absenteeism rate',
+       title = 'B: Monthly absenteeism rate') +
+  scale_linetype_manual(name = '', values = c(1,4)) +
+  theme(legend.position = c(.84, .8)) +
+  guides(lty = guide_legend(ncol = 1)) +
+    theme(axis.title.x = element_blank())
+
+# Weather
+g4 <-
+  ggplot(data = weather %>%
+           group_by(date = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+           summarise(precipitation = sum(precipitation, na.rm = TRUE)),
+         aes(x = date,
+             y = precipitation)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Mililiters',
+       title = 'D: Monthly precipitation') +  theme(axis.title.x = element_blank())
+
+# Clinical incidence of malaria
+g5 <- clinic_agg %>%
+  filter(date >= '2013-01-01', date <= '2016-12-31') %>%
+  # mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
+  group_by(date) %>%
+  summarise(tested = sum(tested),
+            positive = sum(positive)) %>%
+  ungroup %>%
+  mutate(percent_positive = positive / tested * 100) %>%
+  ggplot(aes(x = date,
+             y = positive)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Cases',
+       title = 'C: Monthly malaria cases\nat company clinic') +
+    theme(axis.title.x = element_blank())
+
+g6 <- clinic_agg %>%
+  filter(date >= '2013-01-01', date <= '2016-12-31') %>%
+  # mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
+  group_by(date) %>%
+  summarise(tested = sum(tested),
+            positive = sum(positive)) %>%
+  ungroup %>%
+  mutate(percent_positive = positive / tested * 100) %>%
+  ggplot(aes(x = date,
+             y = percent_positive)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Cases',
+       title = 'D: Monthly malaria test positivity rate\nat company clinic') +
+    theme(axis.title.x = element_blank())
+
+# Rmisc::multiplot(g1, g2, g3,
+#                  # g4, 
+#                  g5, g6, layout = matrix(c(1, 1, 2, 3, 4, 5), nrow=2, byrow=TRUE)) 
+Rmisc::multiplot(g1,  g5,g2, g6, cols = 2)
+```
+
+```{r}
+fumigations <- mc %>%
+  filter(date >= '2013-01-01',
+         date <= '2016-12-31')
+n_fumigations <- nrow(fumigations)
+n_buildings <- sum(fumigations$casas_cobertas, na.rm = TRUE)
+unique_agregados <- length(unique(fumigations$unidade))
+people_sprayed <- length(unique(final_data$oracle_number[final_data$days_since != 'Never']))
+people_observed <- length(unique(final_data$oracle_number))
+pr <- function(x){prettyNum(x, big.mark = ',')}
+```
+
+**Fumigations**: During the period from January 1st, 2013 through December 31st, 2016, the Maragra Malaria Control Unit carried out `r pr(n_fumigations)` episodes of fumigation of residential "agregados" (household combinations), for a total of `r pr(n_buildings)` building-fumigation combinations.
+The total number of unique agregados sprayed during this period was `r pr(unique_agregados)`.
+Among the `r pr(people_observed)` workers for whom we have reliable absenteeism and residential data, `r pr(people_sprayed)` had their homes fumigated at least once (the majority of workers live off of the facility).
+
+```{r}
+dry_days <- length(which(weather$precipitation > 0))
+n_days <- length(weather$precipitation)
+p_dry <- round(dry_days / n_days * 100)
+amount_rain <- round(mean(weather$precipitation[weather$precipitation > 0], na.rm = T), digits = 2)
+x <- weather %>%
+  group_by(month = format(date, '%m')) %>%
+  summarise(precipitation = mean(precipitation, na.rm = TRUE))
+```
+
+**Absences**: We observed `r pr(nrow(ab_panel))` unique worker-days among the `r pr(people_observed)` workers.
+The all-period average absenteeism rate was `r round(100 * sum(ab_panel$absent) / nrow(ab_panel), digits = 2)`%, though this rate varied widely as a function of worker department, sex, residence, and season (table 1).
+
+```{r, results='asis'}
+calculate <- function(var){
+  out <- 
+  final_data %>%
+    mutate(year = format(date, '%Y')) %>%
+    # mutate(on_site = ifelse(on_site,
+    #                                 # 'On site',
+    #                                 # 'Off site'),
+    #        rainy = ifelse(rainy,
+    #                       'Rainy',
+    #                       'Dry')) %>%
+    # filter(!is.na(rainy)) %>%
+  group_by_(var, 'year') %>%
+  summarise(absences = length(which(absent)),
+            eligibles = length(absent),
+            workers = length(unique(oracle_number))) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100) 
+  names(out)[1] <- 'variable'
+  out$val <- paste0(round(out$absenteeism_rate, digits = 1),
+                    '%')#,
+                    # ')',
+                    # '\n',
+                    # ' (workers:', out$workers,
+                    # ')')
+  out <- 
+    out %>%
+    dplyr::select(variable, year, val)
+
+  out <- out %>%
+    spread(key = year,
+           value = val)
+  out$variable <- as.character(out$variable)
+  new_var <- #ifelse(var == 'on_site',
+                    # 'Residence',
+                    ifelse(var == 'rainy',
+                           'Precipitation',
+                           var)
+  # )
+  left <- data.frame(x = c(new_var, rep('', (nrow(out) - 1))))
+  out <- cbind(left, out)
+  names(out)[1:2] <- c('Variable', 'x')
+  # out$variable[1] <- paste0(Hmisc::capitalize(new_var), ': ', out$variable[1])
+  return(out)
+}
+out_list <- list()
+vars <- c('field', 'permanent_or_temporary', 'sex')
+# vars <- c(#'season', 
+#            'field',
+#           'permanent_or_temporary',
+#            'sex')#,
+#            # 'on_site',
+#            # 'rainy'
+#           # )
+for(i in 1:length(vars)){
+  x <- calculate(var = vars[i])
+  out_list[[i]] <- x
+}
+ab_table <- bind_rows(out_list)
+ab_table$Variable <- Hmisc::capitalize(ab_table$Variable)
+# ab_table$Variable[ab_table$Variable == 'Season'] <- 'Malaria season'
+ab_table$Variable[ab_table$Variable == 'Field'] <- 'Worker type'
+ab_table$Variable[ab_table$Variable == 'Permanent_or_temporary'] <- 'Contract'
+ab_table$x <- Hmisc::capitalize(ab_table$x)
+names(ab_table)[2] <- ''
+kable(ab_table, format = 'html') %>% kable_styling()
+# print(xtable::xtable(ab_table,
+#                      caption = 'Absenteeism rate by year and worker characteristics',
+#                      align = c('r', 'r', 'l', rep('l', 4))),
+#       include.rownames = FALSE)
+```
+
+```{r}
+ov_ab <- round(100 * mean(ab_panel$absent[between(ab_panel$date, as.Date('2012-01-01'), as.Date('2016-12-31'))]), digits = 2)
+
+ov_cl <- length(which(between(clinic$date, as.Date('2012-01-01'), as.Date('2016-12-31'))))
+```
+
+**Precipitation**: Of the `r n_days` days observed, `r dry_days` had no rainfall at all (ie, approximately two-thirds).
+On days for which there was any rainfall at all, the average amount was approximately `r amount_rain` centimeters.
+Rain was most common in December and January (average of 5-5.5 cm daily) and least common in August and September (average of 0.5 cm daily).
+Days with any rainfall saw more absenteeism than days with no rainfall (Figure 6, panel A) and among days with any rainfall, more precipitation was associated with greater absenteeism (Figure 6, panel B) (correlation coefficient of 0.25).
+
+```{r, fig.cap = 'Rainfall and absenteeism: association of any rainfall with absenteeism rate (left) and association of rainfall amount and absenteeism (right)', fig.height=3.5, fig.width=7, fig.align='center'}
+bi_data <- final_data %>%
+  group_by(date) %>%
+  mutate(any_rain = mean(precipitation, na.rm = TRUE) > 0) %>%
+  ungroup %>%
+  filter(!is.na(any_rain)) %>%
+  mutate(any_rain = ifelse(any_rain, 'Some rain', 'No rain')) %>%
+  group_by(any_rain, date) %>%
+  summarise(absenteeism = mean(absent) * 100)
+bi_data_simple <- bi_data %>%
+  group_by(any_rain) %>%
+  summarise(absenteeism = mean(absenteeism))
+p1 <- ggplot(data = bi_data,
+       aes(x = any_rain,
+           y = absenteeism)) +
+  geom_jitter(alpha = 0.3, size = 0.5) +
+  geom_violin(alpha = 0.4, fill = 'lightblue') +
+  scale_y_log10() +
+  theme_maragra() +
+  geom_point(data = bi_data_simple,
+             aes(x = any_rain,
+                 y = absenteeism),
+             alpha = 0.8) +
+  geom_text(data = bi_data_simple,
+             aes(x = any_rain,
+                 y = absenteeism,
+                 label = paste0(round(absenteeism, digits = 2), '\n(median)')),
+             color = 'black') +
+  labs(x = '',
+       y = 'Absenteeism rate (log scale)',
+       title = 'A')
+
+plot_data <- final_data %>%
+  group_by(date) %>%
+  summarise(precipitation = mean(precipitation, na.rm = TRUE),
+            absenteeism = mean(absent)) %>%
+  ungroup %>%
+  filter(precipitation > 0.1)
+p2 <- ggplot(data = plot_data,
+       aes(x = precipitation,
+           y = absenteeism * 100)) +
+  geom_point(alpha = 0.3, size = 0.5) +
+  scale_x_log10() +
+  geom_smooth() +
+  # scale_y_log10() +
+  theme_maragra() +
+  labs(x = 'Daily precipitation (log scale)',
+       y = 'Absenteeism rate', 
+       title = 'B')
+Rmisc::multiplot(p1, p2, cols = 2)
+```
+
+```{r, fig.align = 'center', fig.cap = 'Distribution of herd protection score in panel data', eval = F}
+pd <- final_data %>%
+  group_by(protection) %>% tally
+ggplot(data = pd,
+       aes(x = protection,
+           y = n)) +
+  geom_bar(stat = 'identity',
+           fill = 'lightblue',
+               alpha = 0.6) +
+  theme_maragra() +
+  labs(x = 'Herd protection score',
+       y = 'Density')
+```
+
+\newpage
+
+## Sensitivity analysis
+
+```{r, eval = FALSE}
+
+# The below shows average FE for year / worker, during malaria season, on a rainy day...
+
+expander <- function(x){sort(unique(unlist(final_data[,x])))}
+dummy <- expand.grid(#season = expander('season'),
+                     months_since = expander('months_since'),
+                     rainy_day = expander('rainy_day'),
+                     herd = mean(final_data$herd),
+                     group = expander('group'))
+dummy$predicted <- NA
+
+groups <- sort(unique(dummy$group))
+for (i in 1:length(groups)){
+    message(i)
+    this_group <- groups[i]
+    indices <- which(dummy$group == this_group)
+    model <- protection_models[[this_group]]
+    fe <- getfe(model)
+    # Add back in the mean effect of fixed effects
+    add <- fe %>% group_by(fe) %>% summarise(effect = mean(effect)) %>% summarise(sum(effect)) %>% unlist %>% as.numeric
+    data <- dummy[indices,]  
+    # predict_felm(model = model, data = )
+    dummy$predicted[indices] <- 
+      (regtools::predict.felm(object = model,
+                             newdata = data,
+                             use.fe = FALSE) + add) * 100
+}
+
+ggplot(data = dummy %>% filter(rainy_day),
+       aes(x = months_since,
+           y = predicted,
+           color = season,
+           group = season)) +
+  geom_line() +
+  geom_point() +
+  facet_wrap(~group, scales = 'free_y') +
+  scale_colour_manual(name = 'Malaria\nseason',
+                      values = c('blue', 'red')) +
+  theme_maragra() +
+  labs(x = 'IRS status',
+       y = 'Estimated absenteeism') #+
+  # theme(legend.position = 'right')
+
+# The below shows estimated importance of herd protection, and heterogeniety across groups.
+
+dummy <- expand.grid(season = expander('season'),
+                     months_since = expander('months_since'),
+                     rainy_day = expander('rainy_day'),
+                     herd = 0:1000,
+                     group = expander('group'))
+dummy$predicted <- NA
+
+groups <- sort(unique(dummy$group))
+for (i in 1:length(groups)){
+    message(i)
+    this_group <- groups[i]
+    indices <- which(dummy$group == this_group)
+    model <- protection_models[[this_group]]
+    fe <- getfe(model)
+    # Add back in the mean effect of fixed effects
+    add <- fe %>% group_by(fe) %>% summarise(effect = mean(effect)) %>% summarise(sum(effect)) %>% unlist %>% as.numeric
+    data <- dummy[indices,]  
+    dummy$predicted[indices] <- 
+      (regtools::predict.felm(object = model,
+                             newdata = data,
+                             use.fe = FALSE) + add) * 100
+}
+
+ggplot(data = dummy %>% filter(rainy_day, season == 'high'),
+       aes(x = herd,
+           y = predicted,
+           color = months_since,
+           group = months_since)) +
+  geom_line() +
+  facet_wrap(~group, scales = 'free_y') +
+  scale_colour_manual(name = 'IRS status',
+                      values = c('darkorange', 'darkgreen')) +
+  theme_maragra() +
+  labs(x = 'Herd protection score',
+       y = 'Estimated absenteeism')
+```
+
+### What would happen without IRS?
+
+```{r, eval = FALSE}
+
+# The below chart shows the modeled average absenteeism rate during malaria season for permanent workers, by group type.
+
+x <- final_data %>%
+  # dplyr::filter(season == 'high'#,
+  #               # group == groups[1],
+  #               # permanent_or_temporary == 'Permanent'
+  #               )
+x <- x [,grepl('predicted', names(x)) | names(x) == 'group']
+x <- x %>% gather(key, value, predicted_max_herd_max_irs:predicted)
+dict <- data.frame(key = c('predicted', 
+                           'predicted_no_herd',
+                           'predicted_no_irs',
+                           'predicted_no_herd_no_irs',
+                           'predicted_max_herd_max_irs',
+                           'predicted_max_herd',
+                           'predicted_max_irs'),
+                   new_key = c('As is',
+                               'No herd protection',
+                               'No individual protection',
+                               'No protection (herd or individual)',
+                               'Protection: maximum IRS and herd',
+                               'Protection: maximum herd',
+                               'Protection: maximum IRS'))
+
+x <- left_join(x,
+               dict)
+plot_data <- x %>%
+  mutate(new_key = gsub(' ', '\n', new_key)) %>%
+  group_by(group, 
+           new_key) %>%
+  summarise(avg = mean(value, na.rm = TRUE),
+            med = median(value, na.rm = TRUE),
+            upr = quantile(value, 0.75, na.rm = TRUE),
+            lwr = quantile(value, 0.25, na.rm = TRUE))
+levs <- gsub(' ', '\n', c('No herd protection',
+                               'No individual protection',
+                               'No protection (herd or individual)',
+                               'As is',
+                               'Protection: maximum IRS',
+                               'Protection: maximum herd',
+                               'Protection: maximum IRS and herd'))
+plot_data$new_key <- factor(plot_data$new_key,
+                       levels = levs)
+cols <- colorRampPalette(brewer.pal(n = 9,
+                                    name = 'Set1'))(length(unique(plot_data$group)))
+
+library(ggrepel)
+ggplot(data = plot_data,
+       aes(x = new_key,
+           y = avg * 100,
+           color = group)) +
+  geom_point(size = 4) +
+  theme_maragra() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  labs(x = 'Scenario',
+       y = 'Absenteeism rate') +
+  geom_line(aes(group = group),
+            alpha = 0.5,
+            size = 1) +
+  scale_color_manual(name = '',
+                     values = cols) #+
+  # ggrepel::geom_label_repel(aes(label = round(avg * 100, digits = 1)))
+```
+
+## Return on investment
+
+Costs
+
+```{r}
+costs <- maragra::costs
+mean(costs$total)
+
+# 14152 avoided absences
+
+74103 / 14152
+
+# # Working days per year
+# x <- ab_panel %>%
+#   mutate(year = as.numeric(format(date, '%Y'))) %>%
+#   # remove 2016, since smaller sample
+#   filter(year != 2016) %>%
+#   group_by(year) %>%
+#   tally %>%
+#   summarise(x = mean(n))
+# x
+# 
+# x$x / 26.6 # number of days to prevent an absence
+
+```
+
+- Number of working days
+
+```{r, echo = TRUE}
+x <- ab_panel %>% group_by(date) %>% tally
+ggplot(data = x,aes(x = date,y = n)) + geom_line() +
+  labs(title = mean(x$n))
+```
+
+- Number of working days annually
+
+```{r, echo = TRUE}
+options(scipen = '999')
+x <- ab_panel %>% mutate(year = as.numeric(format(date, '%Y'))) %>% group_by(year) %>% tally
+ggplot(data = x,aes(x = year,y = n)) + geom_line() +
+  labs(title = mean(x$n))
+```
+
+- Number of prevented absences per year
+
+```{r, echo = TRUE, eval = FALSE}
+ratio <- length(unique(ab_panel$oracle_number)) / length(unique(final_data$oracle_number))
+
+x <- final_data %>%
+  group_by(year = format(date, '%Y')) %>%
+  summarise(eligibles = n(),
+            absences = mean(as.numeric(absent), na.rm = TRUE) * n(),
+            counterfactual = mean(predicted_no_herd_no_irs,na.rm =TRUE) * n()) 
+y <- x %>%
+  summarise(x = mean(counterfactual) - mean(absences)) %>%
+  unlist * ratio
+# Annual prevented
+y
+x <- x %>%
+  mutate(prevented = counterfactual - absences)
+# Number of presences to prevent 1 case
+sum(x$eligibles) / sum(x$prevented)
+```
+
+```{r, eval = FALSE}
+# Clinical costs
+(y /
+  80) * # number of absences it takes to produce one company clinic positve
+  31.49 # cost per clinical case per ezenduka
+
+# Wage costs
+y * c(3.5, 22)
+  # 4 # assuming USD
+```
+
+\newpage
+
+## Robustness and generalizability
+
+```{r, echo = FALSE, eval = FALSE}
+ro_data <- final_data %>%
+# ro_data <- ab_panel %>%
+#   left_join(irs, by = c('date', 'unidade')) %>%
+#   filter(days_since < 0) %>%
+  mutate(pre_irs = between(days_since, -10, -1)) 
+#   # mutate(time_period = lendable::time_period_extract(date)) %>%
+#   # mutate(time_period = as.character(time_period)) %>%
+#   mutate(time_period = make_season(date = date)) %>%
+#   mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) %>%
+#   left_join(workers %>%
+#               dplyr::select(oracle_number,
+#                             sex,
+#                             department,
+#                             permanent_or_temporary),
+#             by = 'oracle_number')
+# ro_data$time_period <- as.character(ro_data$time_period)
+
+make_model <- function(formula = 'absent ~ pre_irs + time_period'){
+  ro_model <- glm(as.formula(formula), 
+                data = ro_data,
+                family = binomial('logit'))
+  coefs <- exp(coef(ro_model))
+  ci <- confint(ro_model)
+  ci <- exp(ci)
+  out <- data.frame(Variable = names(coefs),
+                    Estimate = coefs,
+                    Lower = ci[,1],
+                    Upper = ci[,2])
+  row.names(out) <- NULL
+  out$`P value` <- coef(summary(ro_model))[,4]
+  out$Significant <- ifelse(out$`P value` <= 0.001, 'xxx',
+                            ifelse(out$`P value` <= 0.01, 'xx',
+                                   ifelse(out$`P value` <= 0.05, 'x',
+                                          '')))
+  out$Variable <- as.character(out$Variable)
+  out$Variable <- gsub('pre_irsTRUE', '10 days prior to IRS', out$Variable)
+  out$Variable <- gsub('time_period', '', out$Variable)
+  return(kable(out, format = "html") %>%
+  kable_styling()# %>%
+    # row_spec(row = which(grepl('prior', out$Variable)),background = "yellow")
+  )
+}
+```
+
+```{r, echo = FALSE, results = 'asis'}
+
+# # BASIC FORMULA
+# make_model('absent ~ pre_irs + time_period')
+
+# # BASIC FORMULA WITH INTERACTIONS
+# make_model('absent ~ pre_irs*time_period')
+
+# # BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX
+# make_model('absent ~ pre_irs*time_period + sex')
+# 
+# # BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX AND DEPARTMENT
+# make_model('absent ~ pre_irs*time_period + sex + department')
+# 
+# 
+# # BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX AND DEPARTMENT AND STATUS
+# make_model('absent ~ pre_irs*time_period + sex + department + permanent_or_temporary')
+
+# Adjust for seasonality and department
+# make_model('absent ~ pre_irs*season + group + rainy_day')
+```
+
+```{r}
+# The below shows our robustness check for sick only absenteeism.
+Unlike with all cause absenteeism, during malaria season, being in the period 10 days prior to IRS is associated with statistically greater likelihood of being absent for illness.
+In other words, it appears that there is somewhat of a feedback loop: when a worker misses work due to illness, his/her likelihood of getting IRS doubles in the next 10 days.\todo{What to do about this???}
+# make_model('absent_sick ~ pre_irs*time_period + department')
+```
+
 # Bibliography

### 2020-04-18 inst/rmd/final/final_september_2019.Rmd
update

---
title: "Final analysis for paper"
date: ''
subtitle: Brew, Pradhan, Sicuri
bibliography: bibliography.bib
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# csl: journal-of-health-economics.csl

# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(RColorBrewer)
library(extrafont)
library(kableExtra)
library(raster)
loadfonts()
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               fig.height = 4,
               fig.width = 5,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               dev = "cairo_pdf",
               fig.pos="!h",
               fig.align = 'center')
opts_knit$set(width=75)
options(xtable.comment = FALSE)
```

```{r, eval = TRUE}
source('prepare_data.R')
```

# September 2019 stuff

## Table of observations

```{r}
library(table1)
message('We observed ', nrow(model_data), ' worker-days among ', length(unique(model_data$oracle_number)), ' unique workers. Overall absence rate was ', length(which(model_data$absent)))

# Table of those absences
vars <- list('field',
          'permanent_or_temporary',
           'sex')
table1::label(model_data$field) <- 'Worker location'
table1::label(model_data$permanent_or_temporary) <- 'Worker contract'
table1::label(model_data$sex) <- 'Worker sex'
model_data$year <- as.numeric(format(model_data$date, '%Y'))
table1::label(model_data$year) <- 'Year'

table1::table1(~field + permanent_or_temporary + sex | year , data = model_data)
```

## Table of absences

```{r}
calculate <- function(var){
  out <- 
  final_data %>%
    mutate(year = format(date, '%Y')) %>%
    # mutate(on_site = ifelse(on_site,
    #                                 # 'On site',
    #                                 # 'Off site'),
    #        rainy = ifelse(rainy,
    #                       'Rainy',
    #                       'Dry')) %>%
    # filter(!is.na(rainy)) %>%
  group_by_(var, 'year') %>%
  summarise(absences = length(which(absent)),
            eligibles = length(absent),
            workers = length(unique(oracle_number))) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) 
  names(out)[1] <- 'variable'
  out$val <- paste0(round(out$absenteeism_rate, digits = 1),
                    '%')#,
                    # ')',
                    # '\n',
                    # ' (workers:', out$workers,
                    # ')')
  out <- 
    out %>%
    dplyr::select(variable, year, val)

  out <- out %>%
    spread(key = year,
           value = val)
  out$variable <- as.character(out$variable)
  new_var <- #ifelse(var == 'on_site',
                    # 'Residence',
                    ifelse(var == 'rainy',
                           'Precipitation',
                           var)
  # )
  left <- data.frame(x = c(new_var, rep('', (nrow(out) - 1))))
  out <- cbind(left, out)
  names(out)[1:2] <- c('Variable', 'x')
  # out$variable[1] <- paste0(Hmisc::capitalize(new_var), ': ', out$variable[1])
  return(out)
}
out_list <- list()
vars <- c(#'season', 
           'field',
          'permanent_or_temporary',
           'sex'#,
           # 'on_site',
           # 'rainy'
          )
for(i in 1:length(vars)){
  x <- calculate(var = vars[i])
  out_list[[i]] <- x
}
ab_table <- bind_rows(out_list)
ab_table$Variable <- Hmisc::capitalize(ab_table$Variable)
# ab_table$Variable[ab_table$Variable == 'Season'] <- 'Malaria season'
ab_table$Variable[ab_table$Variable == 'Field'] <- 'Worker type'
ab_table$Variable[ab_table$Variable == 'Permanent_or_temporary'] <- 'Contract'
ab_table$x <- Hmisc::capitalize(ab_table$x)
names(ab_table)[2] <- ''
ab_table$`2016`[is.na(ab_table$`2016`)] <- '0%'
kable(ab_table, format = 'html') %>% kable_styling(full_width = FALSE)
```

# Scenarios: policy simulations

## Cost of max strategy

```{r}
x = simulations %>%
  arrange(oracle_number, date) %>%
  mutate(first_day = protection == 1 & dplyr::lag(protection, 1) != 1) %>%
  filter(maragra_bairro) %>%
  group_by(strategy) %>%
  summarise(day1 = length(which(first_day)))
x

y <- maragra::costs_itemized %>% filter(area == 'Pesticides')
y <- y[3:6,]
sum(y$total)
```

## Breakdown

```{r}
sim_data <- simulations
sim_data$idx <- sim_data$oracle_number
fe <- getfe(final_model)
predictions <- exp(predict(final_model_lm, sim_data, interval = 'confidence'))-1
predictions <- data.frame(predictions)
# Add the fixed effects
fixed_effects <- left_join(sim_data,
                           data.frame(fe) %>% dplyr::select(idx, effect) %>%
                             mutate(idx = as.character(idx))) %>%
  .$effect
fixed_effects <- as.numeric(fixed_effects)

predictions$fit <- predictions$fit + fixed_effects
predictions$lwr <- predictions$lwr + fixed_effects
predictions$upr <- predictions$upr + fixed_effects

# combine other info
predictions <- predictions %>% bind_cols(sim_data %>% dplyr::select(date, absent, strategy, permanent_or_temporary)) %>%
  mutate(absent = as.numeric(absent)) %>%
  filter(!is.na(fit))

simple <- predictions %>%
  group_by(strategy) %>%
  summarise(n = n(),
            actual_absences = sum(absent),
            predicted_absences = sum(fit, na.rm = TRUE),
            actual = mean(absent, na.rm = TRUE),
            predicted = mean(fit, na.rm = TRUE)) %>%
  dplyr::rename(value = predicted) %>%
  dplyr::select(strategy, value, predicted_absences)

zero <- predictions %>%
  filter(strategy == 'Zero') %>%
  mutate(year = as.numeric(format(date, '%Y'))) %>%
  group_by(year) %>%
  summarise(n = n(),
            actual_absences = sum(absent),
            predicted_absences = sum(fit, na.rm = TRUE),
            actual = mean(absent, na.rm = TRUE),
            predicted = mean(fit, na.rm = TRUE)) %>%
  mutate(avoided = predicted_absences - actual_absences)
  # dplyr::rename(value = predicted) %>%
  # dplyr::select(permanent_or_temporary, value, predicted_absences)

```

```{r}
ggplot(data = simple,
       aes(x = strategy,
           y = value )) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(y = 'Absenteeism rate',
       x = 'Strategy') +
  geom_label(aes(label = round(value, digits = 3)),
             nudge_y = -0.005) +
  theme_maragra() +
  theme(axis.text.x = element_text(size = 15))
```

```{r}
simple
ggplot(data = simple,
       aes(x = strategy,
           y = predicted_absences)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(y = 'Absenteeism rate',
       x = 'Strategy') 
```

# Manuscript draft

Draft of the paper \href{https://docs.google.com/document/d/1bUWRBCgVcgjSPHchIQxiTG8Vwv5hV1GLU4Tlu386sWA/edit#}{HERE}.

# Bibliography

### 2020-04-18 inst/rmd/final/prepare_data.R

@@ -12,6 +12,18 @@ library(lfe)
 library(restorepoint)
 library(regtools)
 
+# classify_protection <- function(days_since){
+#   x <- 21:204
+#   y <- dplyr::percent_rank(x)
+#   y <- rev(y)
+#   out <- rep(NA, length(days_since))
+#   for(i in 1:length(days_since)){
+#     out[i] <-ifelse(days_since[i] %in% x, 
+#                     y[x == days_since[i]],
+#                     0)
+#   }
+#   out
+# }
 
 classify_protection <- function(days_since){
   ifelse(days_since < 21 | days_since > (183+21),
@@ -29,6 +41,8 @@ classify_protection <- function(days_since){
                                                    0.2,
                                                    0)))))))
 }
+
+
 
 
 if('prepared_data.RData' %in% dir()){
@@ -138,9 +152,13 @@ if('prepared_data.RData' %in% dir()){
                               education,
                               floor_material)) %>%
     # Bring in some data for weather
-    left_join(weather) %>%
-    ########## !!!!!!!!!!!!!!! REMOVING 2016
-    filter(date <= '2015-12-31')
+    left_join(weather)
+  
+    ########## !!!!!!!!!!!!!!! REMOVING 2016 for temp workers
+  model_data <- model_data %>%
+    filter(permanent_or_temporary == 'Permanent' |
+             (permanent_or_temporary == 'Temporary' & date <= '2015-12-31'))
+
   
   # Since ad and factory are same, keep same
   model_data$field <- ifelse(model_data$department == 'Field', 
@@ -245,7 +263,7 @@ if('prepared_data.RData' %in% dir()){
       val <- df$protection * weighter(x)
       # Infinite means distance was 0/self; remove
       val[is.infinite(val)] <- NA
-      sum(val, na.rm = TRUE)})
+      mean(val, na.rm = TRUE)}) # used to be sum
     right <- data.frame(lat = df$lat[!is.na(df$lat)],
                         lng = df$lng[!is.na(df$lng)],
                         herd = out)
@@ -278,12 +296,10 @@ if('prepared_data.RData' %in% dir()){
   }
   model_data <- bind_rows(out_list)
   
+  save.image('checkpoint.RData')
   # Create a single "protection" score which combines
   # herd protection and individual protection
-  model_data$combined_protection  <-
-    # Assuming 50 meters
-    (model_data$protection * weighter(0.05)) +
-    model_data$herd_protection
+
 
   
[…]

### 2020-06-20 inst/rmd/apr2020/apr2020.Rmd
updated

---
title: "Final analysis for paper"
date: ''
fig_height: 2.6
output:
  pdf_document:
    includes:
      in_header: preamble.sty
    keep_tex: yes
    latex_engine: xelatex
  html_document:
    df_print: paged
fig_width: 4
always_allow_html: true
subtitle: Brew, Pradhan, Sicuri
bibliography: bibliography.bib
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# csl: journal-of-health-economics.csl

# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(RColorBrewer)
library(extrafont)
library(kableExtra)
library(raster)
library(sp)
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(knitr)
library(RColorBrewer)
# library(lme4)
library(lfe)
library(restorepoint)
library(regtools)

loadfonts()
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               fig.height = 4,
               fig.width = 5,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               # dev = "cairo_pdf",
               fig.pos="!h",
               fig.align = 'center')
knitr::opts_chunk$set(error = FALSE)

opts_knit$set(width=75)
options(xtable.comment = FALSE)
```

```{r}
classify_protection <- function(days_since){
  ifelse(days_since < 21 | days_since > (365+21),0,
         (((365+21) - days_since)/(365))
         )
}

if('prepared_data.RData' %in% dir()){
  load('prepared_data.RData')
} else {
  # Read in data
  ab <- maragra::ab
  ab_panel <- maragra::ab_panel
  bairros <- maragra::bairros
  bes <- maragra::bes
  census <- maragra::census
  clinic <- maragra::clinic
  clinic_agg <- maragra::clinic_agg
  mc <- maragra::mc
  workers <- maragra::workers

  if(!'weather.RData' %in% dir()){
    weather <- maragra::weather %>%
    mutate(precipitation = ifelse(is.na(precipitation),
                                  0,
                                  precipitation))

    # Get lags for weather
    starts <- c(15, 30, 60, 90)
    ends <- c(30, 60, 90, 120)
    for(k in 1:length(starts)){
      for(j in 1:length(ends)){
        this_start <- starts[k]
        this_end <- ends[j]
        if(this_start < this_end){
          precipitation_column_name <- paste0('precipitation_lag_', this_start, '_', this_end)
          weather[,precipitation_column_name] <- NA
          temp_column_name <- paste0('temp_lag_', this_start, '_', this_end)
        }
      }
    }
    for(i in 1:nrow(weather)){
      starts <- c(15, 30, 60, 90)
      ends <- c(30, 60, 90, 120)
      for(k in 1:length(starts)){
        for(j in 1:length(ends)){
          this_start <- starts[k]
          this_end <- ends[j]
          if(this_start < this_end){
            this_date <- weather$date[i]
            end_date <- this_date - this_start
            start_date <- this_date - this_end
            if(start_date >= min(weather$date,na.rm = TRUE) &
               end_date <= max(weather$date, na.rm = TRUE)){
              sub_data <- weather %>%
                filter(date >= start_date,
                       date <= end_date)
              precipitation_column_name <- paste0('precipitation_lag_', this_start, '_', this_end)
              weather[i,precipitation_column_name] <- mean(sub_data$precipitation, na.rm = TRUE)
              temp_column_name <- paste0('temp_lag_', this_start, '_', this_end)
              weather[i,temp_column_name] <- mean(sub_data$temp, na.rm = TRUE)
            }
          }
        }
      }
    }
    save(weather, file = 'weather.RData')
  } else {
    load('weather.RData')
  }

  # Create a dictionary to classify protection
  protection_dict <- 
    data_frame(days_since = seq(min(irs$days_since, na.rm = TRUE),
                                max(irs$days_since, na.rm = TRUE),
                                by = 1)) %>%
    mutate(protection = classify_protection(days_since))
  model_data <- 
    ab_panel %>%
    dplyr::select(oracle_number, date, absent, unidade) %>%
    left_join(irs %>% 
                dplyr::select(date, unidade, chemical, days_since), 
              by = c('unidade', 'date')) %>%
    # mutate(protection = ifelse(days_since < 21 |
    #                              days_since > (183+21) |
    #                              is.na(days_since),
    #                            0, 1)) %>%
    left_join(protection_dict) %>%
    mutate(protection = ifelse(is.na(protection), 0, protection)) %>%
    dplyr::select(-days_since) %>%
    group_by(oracle_number) %>%
    mutate(ever_sprayed = length(which(protection > 0)) > 0) %>%
  # Bring in some information on workers
    left_join(workers %>%
                dplyr::select(oracle_number,
                              permanent_or_temporary,
                              department,
                              sex,
                              # date_of_birth,
                              perm_id,
                              census_name_match_score) %>%
                # void the permids of anyone with a match score of greater than 0.25
                mutate(perm_id = ifelse(census_name_match_score > 0.25,
                                        NA,
                                        perm_id))) %>%
    # Bring in some info from the census
    left_join(census %>%
                filter(!duplicated(perm_id)) %>%
                dplyr::select(perm_id,
                              maragra_bairro,
                              maragra_fabrica,
                              education,
                              floor_material)) %>%
    # Bring in some data for weather
    left_join(weather)

    ########## !!!!!!!!!!!!!!! REMOVING 2016 for temp workers
  model_data <- model_data %>%
    filter(permanent_or_temporary == 'Permanent' |
             (permanent_or_temporary == 'Temporary' & date <= '2015-12-31'))

  # Since ad and factory are same, keep same
  model_data$field <- ifelse(model_data$department == 'Field', 
                             'Field worker',
                             'Not field worker')

  model_data <- model_data %>% 
    mutate(group = paste0(permanent_or_temporary, ' ',
                          tolower(field)))

  # Add a month column
  model_data$month_number <- add_zero(format(model_data$date, '%m'), 2)

  # Get geographic info for externality analysis
  # Get latitude / longitude into model_data
  model_data <- model_data %>%
    left_join(census %>%
                filter(!duplicated(unidade)) %>%
                dplyr::select(unidade,
                              longitude_aura,
                              latitude_aura),
              by = 'unidade')

  model_data$p <- model_data$protection

  # Centers of each bairro
  bairro_loc <- apply(coordinates(maragra::bairros_maragra_bairro), 2, mean)

  # Define function for weighting indirect protection
  weighter <- function(x){
    # out <- (1 / (x+0.01))
    # out <- 1/x
    out <- ifelse(x > 1, 0, 1-x)
    return(out)
  }

  # Define function for getting distance between houses
  overall_spatial <- model_data %>%
    group_by(lat = latitude_aura,
             lng = longitude_aura) %>%
    tally %>%
    dplyr::distinct(lat, lng, .keep_all = TRUE) %>%
    dplyr::select(-n) %>%
    filter(!is.na(lat),
           !is.na(lng)) %>%
    bind_rows(data.frame(lat = bairro_loc[2],
                         lng = bairro_loc[1])) %>%
    mutate(x = lng,
           y = lat)
  overall_spatial$overall_id <- 1:nrow(overall_spatial)
  coordinates(overall_spatial) <- ~x+y
  proj4string(overall_spatial) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
  overall_distances <- spDists(x = overall_spatial, longlat = TRUE)

  get_protection <- function(lat, lng, protection, guess_missing = TRUE){
    # Build a dataframe
    df <- data_frame(lat, lng, protection)
    # Get id
    df$id <- 1:nrow(df)

    if(guess_missing){
      # Define function for guessing
      guess <- function(x){
        x <- x %>%
          mutate(lng = ifelse(is.na(lng), 
                              bairro_loc[1],
                              lng)) %>%

          mutate(lat = ifelse(is.na(lat), 
                              bairro_loc[2],
                              lat)) 
        return(x)
      }
      # if no geography (due to no census matching, just use average geo)
      df <- guess(df)
    }
    # save for later
    original_df <- df

    # Get unique lat/lng pairs
    df <- df %>%
      group_by(lat, lng) %>%
      summarise(protection = mean(protection, na.rm = TRUE))

    # Join to overall spatial to get distances
    df <- left_join(df, overall_spatial@data, by = c('lat', 'lng'))
    keeps <- df$overall_id
    distances <- overall_distances[keeps, keeps]

    # # Get a spatial version
    # (no longer necessary with above approach)
    # df_spatial <- df %>% filter(!is.na(lng),
    #                             !is.na(lat))
    # 
    # coordinates(df_spatial) <- ~lng+lat
    # proj4string(df_spatial) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
    # 
    # # Get distances
    # distances <- spDists(x = df_spatial, longlat = TRUE)

    # Weight distances and return weighted protection
    out <- apply(distances, 1, function(x){
      val <- df$protection * weighter(x)
      # Infinite means distance was 0/self; remove
      val[is.infinite(val)] <- NA
      # Keep only those within a kilometer
      val <- val[x <= 1]
      mean(val, na.rm = TRUE)

      })
    right <- data.frame(lat = df$lat[!is.na(df$lat)],
                        lng = df$lng[!is.na(df$lng)],
                        herd = out)
    # Join to original df
    original_df <- original_df %>%
      left_join(right,
                by = c('lat', 'lng'))

    # Return just the protection column
    return(original_df$herd)
  }

  # For each day, get real, and hypothetical protection scores
  dates <- sort(unique(model_data$date))

  out_list <- list()
  for(i in 1:length(dates)){
    this_date <- dates[i]
    message(this_date)

    this_data <- model_data %>%
      filter(date == this_date)
    # Get protection score
    protection_scores <- get_protection(lat = this_data$latitude_aura,
                                        lng = this_data$longitude_aura,
                                        protection = this_data$protection,
                                        guess_missing = TRUE)
    this_data$herd_protection <- protection_scores
    out_list[[i]] <- this_data 
  }
  model_data <- bind_rows(out_list)

  # Plots of maps
  # Libraries
  library(tidyverse)
  library(raster)
  library(tidyr)
  library(ggplot2)
  library(RColorBrewer)
  library(broom)
  library(ggthemes)

  # Get each countries shapefile
  countries <- c('Mozambique')
  iso3s <- c('MOZ')
  for(i in 1:length(countries)){
    message('Fetching data for ', countries[i])
    x <- getData(name = 'GADM', level = 2, country = iso3s[i])
    assign(tolower(countries[i]),
           x,
           envir = .GlobalEnv)
  }

  # Define which districts are "special" -------------
  specials <- 'Manhiça'

  # Mozambique
  mozambique@data$special <- FALSE
  mozambique@data$special[mozambique@data$NAME_2 %in% specials] <- TRUE

  # Combine all data into a "long" / "tidy" format
  make_long <- function(x, region = "NAME_2"){
    shp_df <- broom::tidy(x, region = region)
    return(shp_df)
  }
  mozambique_long <- make_long(mozambique) %>% mutate(country = 'Mozambique')
  combined <- mozambique_long
  combined$special <- combined$id %in% specials
  # combined$special[!combined$special] <- NA
  # Get a map of africa to use as a background
  # from the cism package!
  africa <- cism::africa
  africa_long <- make_long(africa, region = 'COUNTRY')
  africa_long$special <- africa_long$id %in% countries

  save.image(file = 'prepared_data.RData')
}

# Remove christmas
final_data <- final_data %>%
  mutate(day_number = as.numeric(format(date, '%d')))
remove_christmas <- TRUE
if(remove_christmas){
  remove_these <- which(
    (final_data$month_number == '12' &
       final_data$day_number >= 15) |
      final_data$month_number == '01' &
      final_data$day_number <= 15

  )
}

# Prepare simulations
if('simulations.RData' %in% dir()){
  load('simulations.RData')
} else {

  time_optimize_df <-
    data.frame(date = seq(as.Date('2020-01-01'),
                          as.Date('2020-12-31'), 
                          by = 1)) %>%
    mutate(month_day = format(date, '%m-%d'))
  time_optimize_df$day_number <- (1:nrow(time_optimize_df))+20
  time_optimize_df$the_protection <- classify_protection(time_optimize_df$day_number)
  time_optimize_df <- time_optimize_df %>% dplyr::select(month_day, the_protection)

  strategies <- c('Zero', 'Max', 'Time-optimized')
  strategy_list <- list()
  for(s in 1:length(strategies)){
    dates <- sort(unique(final_data$date))
    out_list <- list()
    this_strategy <- strategies[s]

    for(i in 1:length(dates)){
      this_date <- dates[i]
      message(this_strategy, '...', this_date)

      this_data <- final_data %>%
        filter(date == this_date)

      # Define scores
      if(this_strategy == 'Zero'){
        this_data$protection <- 0
        the_protection <- rep(0, length(this_data$protection))
      }
      if(this_strategy == 'Max'){
        this_data$protection <- 1
        the_protection <- rep(1, length(this_data$protection))
      }
      if(this_strategy == 'Time-optimized'){
        this_data$month_day <- format(this_data$date, '%m-%d')
        x <- this_data %>%
          left_join(time_optimize_df)
        the_protection <- x$the_protection
        this_data$protection <- the_protection
      }
      # Get protection score
      protection_scores <- get_protection(lat = this_data$latitude_aura,
                                          lng = this_data$longitude_aura,
                                          protection = the_protection,
                                          guess_missing = TRUE)
      this_data$herd_protection <- protection_scores
      out_list[[i]] <- this_data 
    }
    out <- bind_rows(out_list)
    out$strategy <- this_strategy
    strategy_list[[s]] <- out
  }
  simulations <- bind_rows(strategy_list)
  simulations <- simulations %>%
    bind_rows(final_data %>% mutate(strategy = 'Real'))
  save(simulations, file = 'simulations.RData')
}

# Some variable things
model_data$rain_var <- model_data$precipitation_lag_15_90

# # Final data - only the ever sprayed
final_data <- model_data %>%
  filter(ever_sprayed)
#   # Temporary fix pending newer data
#   filter(date <= '2016-01-01')

# Add 5th worker type
final_data <- final_data %>%
  # Get grade
  left_join(ab_panel %>% dplyr::distinct(grade, oracle_number, date)) %>%
  mutate(year = as.numeric(format(date, '%Y'))) %>%
  left_join(wages %>% dplyr::distinct(grade, department, year, daily_usd)) 
final_data$group5 <- ifelse(is.na(final_data$daily_usd) & final_data$group != 'Temporary not field worker', 'No wage data', final_data$group)

# final_model <- felm(log(absent+1) ~ protection_var + (rain_var) | oracle_number | 0 | 0,
#                     data = final_data)
final_model <- felm(log(absent+1) ~ protection +
                      herd_protection + 
                      rain_var | oracle_number | 0 | 0,
                    data = final_data)
summary(final_model)
final_model_lm <- lm(log(absent+1) ~ protection +
                       herd_protection + (rain_var),
                     data = final_data)

library(broom)
broom::tidy(final_model)

# Prediction based on individual stuff
fake <- expand.grid(protection = seq(0, 1, 0.5),
                    herd_protection = seq(0, 1, 0.1),
                    rain_var = seq(0, 6, 1))
fe <- getfe(final_model)
predictions <- exp(predict(final_model_lm, newdata = fake, interval = 'confidence'))-1
predictions <- data.frame(predictions)
fixed_effects <- 0.067 # taken from below
fake$fit <- predictions$fit + fixed_effects
fake$lwr <- predictions$lwr + fixed_effects
fake$upr <- predictions$upr + fixed_effects

n_cols <- length(unique(fake$protection))
cols <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, 'Spectral'))(n_cols)
cols[2] <- 'black'
library(maragra)
ggplot(data = fake %>% filter(rain_var == 3) %>% mutate(protection = factor(protection)),
       aes(x = herd_protection,
           y = fit,
           color = protection,
           group = protection)) +
  geom_ribbon(aes(ymin = lwr, ymax = upr,
                  fill = protection),
              color = NA,
              alpha = 0.15) +
  geom_point() +
  geom_line(alpha = 0.6) +
  scale_color_manual(name = 'Individual protection',
                     values = cols) +
  scale_fill_manual(name = 'Individual protection',
                     values = cols) +
  labs(x = 'Community protection score',
       title = 'Association between community protection, individual protection, and absenteeism',
       y = 'Estimated monthly absenteeism rate\n(Average fixed effect, 3 cm of 15-60 day prior precipitation)') + theme_maragra()

# Also make predictions by groups
worker_groups <- sort(unique(final_data$group5))
model_data_list <- model_list <- model_list_lm <- list()

for (i in 1:length(worker_groups)){
  sub_data <- final_data %>% filter(group5 == worker_groups[i])
  this_model <- felm(log(absent+1) ~ protection +
                      herd_protection + 
                      rain_var | oracle_number | 0 | 0,
                    data = sub_data)
  this_model_lm <- lm(log(absent+1) ~ protection +
                       herd_protection + (rain_var),
                     data = sub_data)
  formatted <- broom::tidy(this_model)
  formatted$group5 <- worker_groups[i]

  model_list[[i]] <- this_model
  model_list_lm[[i]] <- this_model_lm
  model_data_list[[i]] <- formatted
}
names(model_list) <- names(model_data_list) <- names(model_list_lm) <- worker_groups

# Do predictions on the different groups
fake <- expand.grid(protection = seq(0, 1, 0.5),
                    herd_protection = seq(0, 1, 0.1),
                    rain_var = 3,# seq(0, 6, 1),
                    group5 = worker_groups)
fake_list <- list()
for(i in 1:length(worker_groups)){
  this_worker_group <- worker_groups[i]
  this_model <- model_list[[i]]
  this_model_lm <- model_list_lm[[i]]
  this_fake_data <- fake %>% filter(group5 == this_worker_group)
  predictions <- exp(predict(this_model_lm, newdata = this_fake_data, interval = 'confidence'))-1
  predictions <- data.frame(predictions)
  this_fe <- getfe(this_model)
  fixed_effects <- mean(this_fe$effect)
  out <- this_fake_data
  out$fit <- predictions$fit + fixed_effects
  out$lwr <- predictions$lwr + fixed_effects
  out$upr <- predictions$upr + fixed_effects
  fake_list[[i]] <- out
}
fake <- bind_rows(fake_list)

n_cols <- length(unique(fake$protection))
cols <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, 'Spectral'))(n_cols)
cols[2] <- 'black'
library(maragra)
ggplot(data = fake  %>% mutate(protection = factor(protection)),
       aes(x = herd_protection,
           y = fit,
           color = protection,
           group = protection)) +
  # geom_ribbon(aes(ymin = lwr, ymax = upr,
  #                 fill = protection),
  #             color = NA,
  #             alpha = 0.15) +
  geom_point() +
  facet_wrap(~group5) +
  geom_line(alpha = 0.6) +
  scale_color_manual(name = 'Individual protection',
                     values = cols) +
  scale_fill_manual(name = 'Individual protection',
                     values = cols) +
  labs(x = 'Community protection score',
       title = 'Association between community protection, individual protection, and absenteeism',
       y = 'Estimated monthly absenteeism rate\n(Average fixed effect, 3 cm of 15-60 day prior precipitation)') +    theme_maragra()

# SIMULATIONS
sim_data <- simulations
sim_data$idx <- sim_data$oracle_number
fe <- getfe(final_model)
predictions <- exp(predict(final_model_lm, sim_data, interval = 'confidence'))-1
predictions <- data.frame(predictions)
# Add the fixed effects
fixed_effects <- left_join(sim_data,
                           data.frame(fe) %>% dplyr::select(idx, effect) %>%
                             mutate(idx = as.character(idx))) %>%
  .$effect
fixed_effects <- as.numeric(fixed_effects)

predictions$fit <- predictions$fit + fixed_effects
predictions$lwr <- predictions$lwr + fixed_effects
predictions$upr <- predictions$upr + fixed_effects

# combine other info
predictions <- predictions %>% bind_cols(sim_data %>% dplyr::select(date, absent, strategy, permanent_or_temporary)) %>%
  mutate(absent = as.numeric(absent)) %>%
  filter(!is.na(fit))

simple <- predictions %>%
  group_by(strategy) %>%
  summarise(n = n(),
            actual_absences = sum(absent),
            predicted_absences = sum(fit, na.rm = TRUE),
            actual = mean(absent, na.rm = TRUE),
            predicted = mean(fit, na.rm = TRUE)) %>%
  dplyr::rename(value = predicted) %>%
  dplyr::select(strategy, value, predicted_absences)

zero <- predictions %>%
  filter(strategy == 'Zero') %>%
  mutate(year = as.numeric(format(date, '%Y'))) %>%
  group_by(year) %>%
  summarise(n = n(),
            actual_absences = sum(absent),
            predicted_absences = sum(fit, na.rm = TRUE),
            actual = mean(absent, na.rm = TRUE),
            predicted = mean(fit, na.rm = TRUE)) %>%
  mutate(avoided = predicted_absences - actual_absences)
# dplyr::rename(value = predicted) %>%
# dplyr::select(permanent_or_temporary, value, predicted_absences)

kable(simple)

# Eligible working days
sum(zero$n)
# Absences
sum(zero$actual_absences)
# Actual absence rate
sum(zero$actual_absences) / sum(zero$n)
# Would have observed
sum(zero$predicted_absences)
# Predicted - observ
sum(zero$predicted_absences) - sum(zero$actual_absences)
  # Overall reduction from
sum(zero$predicted_absences) / sum(zero$n)
# To
sum(zero$actual_absences) / sum(zero$n)
# Number of working days permanent vs temp
table(predictions$permanent_or_temporary[predictions$strategy=='Real'])
# Absences avoided among permanent workers
predictions %>% filter(strategy == 'Zero') %>% group_by(permanent_or_temporary) %>% summarise(n = n(),
                                                                                              actual_absences = sum(absent),
                                                                                              predicted_absences = sum(fit, na.rm = TRUE),
                                                                                              actual = mean(absent, na.rm = TRUE),
                                                                                              predicted = mean(fit, na.rm = TRUE)) %>%
  mutate(avoided = predicted_absences - actual_absences) %>%
  # decline due to irs
  mutate(decline = predicted - actual)
# One absence every n days
sum(zero$n) / sum(zero$avoided)
# Annual reduction in absence
mean(zero$avoided[zero$year < 2016])
# Cost per absence avoided
68984 / 4183
# Percentage of absence among permanent workers
x = predictions %>%
  filter(strategy == 'Zero') %>%
  # filter(date < '2016-01-01') %>%
  group_by(permanent_or_temporary) %>%
  summarise(absences = sum(absent),
            predicted = sum(fit)) %>%
  mutate(averted = predicted - absences) %>%
  mutate(p = averted / sum(averted))
x
# Weighted average cost
y = sum(c(25, 5) * x$p)
y
# Cost of treatment per absence
22 / 80
# Total savings per aversion
0.275 + y
# Total yearly savings
(0.275 + y) * 4183
# Total yearly savings
((0.275 + y) * 4183) - 68984
# ROI
((0.275 + y) * 4183) / 68984
# Additional max strategy
simple
47943-42131

ggplot(data = simple,
       aes(x = strategy,
           y = value)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(y = 'Absenteeism rate',
       x = 'Strategy') +
  geom_label(aes(label = round(value, digits = 3)),
             nudge_y = -0.005)

# paste0(
#   "Of the ", nrow(simulations[simulations$strategy == 'Real',]),
#   " eligible working days in our panel, we observed ", nrow(simulations[simulations$strategy == 'Real' & simulations$absent,]), 
#   " absences. Using the above approach, we estimate that we would have observed ",
#   length(which(simulations$absent[simulations$strategy == 'Zero'])),
#   " had it not been for the firm's IRS program, for a total of ",
#   length(which(simulations$absent[simulations$strategy == 'Zero'])) - length(which(simulations$absent[simulations$strategy == 'Real'])),
#   " avoided absences. This translates to an overall reduction in absenteeism from ",
#   round(length(which(simulations$absent[simulations$strategy == 'Zero'])) / nrow(simulations %>% filter(strategy == 'Zero')) * 100, digits = 2),
#   "% (without its IRS program) to the observed ", round(length(which(final_data$absent)) / nrow(final_data) * 100, digits = 2),"%.
# The number of working-days observed among permanent workers (",
# length(which(final_data$permanent_or_temporary == 'Permanent')),") was nearly thrice that of those of temporary workers (", 
# length(which(final_data$permanent_or_temporary == 'Temporary')),
# ").") 
# 
# 
# 
# paste0(
#   "By the same token, the number of absences avoided among permanent workers (39,890) was far greater than that of temporary workers (8,972). The latter had a greater relative decline in absenteeism due to IRS (from 9.3% to 3.91%) relative to the former (from 14.2% to 11%), but the aggregate numbers are less consequential since (a) temporary workers have a lower baseline absenteeism rate and (b) they work, by definition, fewer days. In total, the IRS program is estimated to prevent one absence every 26.6 working days.
# Annually, this translates to approximately 4,076 avoided absences per year (slightly greater than the simple splicing of the overall figure into the 4-year study period, since we did not have complete data for all worker types).
# "
# )

ggplot(data = simple,
       aes(x = strategy,
           y = predicted_absences)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(y = 'Absenteeism rate',
       x = 'Strategy') +
  geom_label(aes(label = round(predicted_absences, digits = 3)),
             nudge_y = -10000)

# Costs

x <- final_data %>%
  left_join(irs %>% dplyr::select(date, unidade, days_since))

out <- data.frame(
  strategy = c('Max',
               'Real',
               'Time-optimized',
               'Zero'),
  sprays = c(nrow(final_data) / 60,
             length(which(x$days_since == 0)),
             length(which(format(final_data$date, '%m-%d') == '12-11')),
             0
  )
) %>%
  mutate(sprays = sprays / length(unique(final_data$oracle_number)))

kable(out)
ggplot(data = out,
       aes(x = strategy,
           y = sprays)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(y = 'Number of sprays per person',
       x = 'Strategy')

```

### 2020-06-20 inst/rmd/defunct/feb2019_xinavane_reproduction/feb2019_xinavane_reproduction.tex
updated

--- 2018-05-23 inst/rmd/menno_models.tex
@@
 \hypersetup{unicode=true,
-            pdftitle={Concept note: additional models},
+            pdftitle={February 2019 - Xinavane charts reproduction},
             pdfborder={0 0 0},
@@
 \setlength{\droptitle}{-2em}
-  \title{Concept note: additional models}
+
+  \title{February 2019 - Xinavane charts reproduction}
     \pretitle{\vspace{\droptitle}\centering\huge}
   \posttitle{\par}
-\subtitle{Pradhan, Sicuri, Brew}
+  \subtitle{Using Maragra data}
   \author{}
@@
 \usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
 
@@
 
-A few different approaches to modeling absenteeism in the Maragra data
+Charts
 
@@
 
-This document contains methods and basic results for a few different modeling approaches, following a conversation and write-up with Menno Pradhan in May 2018.
-The purpose of this document is to make explicit the current models we have implemented, and some issues with them.
+This document contains charts from Maragra data, meant for comparison with Xinavane. It was created on February 5, 2019 by Joe Brew for Laia Cirera and Elisa Sicuri.
 
@@
     \parbox{\textwidth}{%
-        \subsection*{Main points}
+        \subsection*{Action}
         \begin{itemize}
-          \item I have implemented Menno's suggestions for more simple models.
-          \item The simplest models show no overall effect from IRS when not accounting for malaria season
-          \item I believe that the lack of an effect is due to seasonality's presence in the model as a fixed (additive) effect, rather than as an interaction with IRS
+          \item Review charts herein
+          \item Request clarification or ask questions if applicable
         \end{itemize}
@@
 
-\textbf{Elisa Sicuri; Menno Pradhan}
+\textbf{Laia Cirera; Elisa Sicuri}
 
@@
 
-\section{The current approach}\label{the-current-approach}
+\section{Malaria prevalence in the region, (2013-2017), Manhiça
+(control)}\label{malaria-prevalence-in-the-region-2013-2017-manhica-control}
 
-In the current draft of the paper
-(\href{https://docs.google.com/document/d/1bUWRBCgVcgjSPHchIQxiTG8Vwv5hV1GLU4Tlu386sWA/edit#}{HERE}
-), the model formula is:
+\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-3-1} \end{center}
 
-\[
-\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * IRS_{it} + \hat{\beta}_2{RainyDay_{t}} + \hat{\beta}_3{Herd_{it}} +  \alpha_i + \delta_t + \upsilon_{it}
-\]
+\section{Malaria prevalence in the region,
+(2013-2017)}\label{malaria-prevalence-in-the-region-2013-2017}
 
-In R, using the \texttt{felm} library, this is written as follows:
+Magude (intervention)
 
-\begin{verbatim}
-felm(formula = absent ~ season * months_since + rainy_day + herd | 
-    oracle_number + malaria_year | 0 | 0, data = these_data)
-\end{verbatim}
+\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-4-1} \end{center}
 
-The above notation shows \texttt{absent} (binary 1,0) as a function of
-the interaction of binary \texttt{season} (high vs.~low) and binary
-\texttt{months\_since} (before vs.~after) plus binary
-\texttt{rainy\_day} (whether or not it rained on the particular day)
-plus continuous numeric \texttt{herd} (ie, the protection afforded by
-others) plus fixed effects for both the worker (\texttt{oracle\_number})
-and the year (\texttt{malaria\_year}). \(\alpha_i\) represents the time
-invariant worker fixed effects, and \(\delta_i\) represents the fixed
-effect of the particular malaria season. \(\upsilon\) is the error term.
-This model is run separately for the 4 different worker groups, yielding
-the following results:
+\section{Both locations}\label{both-locations}
 
-\begin{longtable}[t]{ll}
-\caption{\label{tab:unnamed-chunk-4}All absenteeism with herd immunity: model results}\\
-\toprule
-Term & Estimate\\
-\midrule
-\endfirsthead
-\caption[]{All absenteeism with herd immunity: model results \textit{(continued)}}\\
-\toprule
-Term & Estimate\\
-\midrule
-\endhead
-\
-\endfoot
-\bottomrule
-\endlastfoot
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
-\hspace{1em}Malaria season & 0.034 (P<0.001)\\
-\hspace{1em}IRS status=After & 0.035 (P<0.001)\\
-\hspace{1em}Rainy day & 0.027 (P<0.001)\\
-\hspace{1em}Herd protection & 0 (P=0.001)\\
-\hspace{1em}Malaria season:IRS status=After & -0.04 (P<0.001)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
-\hspace{1em}Malaria season & 0.009 (P<0.001)\\
-\hspace{1em}IRS status=After & 0.005 (P=0.054)\\
-\hspace{1em}Rainy day & 0.045 (P<0.001)\\
-\hspace{1em}Herd protection & 0 (P<0.001)\\
-\hspace{1em}Malaria season:IRS status=After & -0.015 (P<0.001)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
-\hspace{1em}Malaria season & -0.003 (P<0.001)\\
-\hspace{1em}IRS status=After & 0.001 (P=0.127)\\
-\hspace{1em}Rainy day & 0.001 (P=0.032)\\
-\hspace{1em}Herd protection & 0 (P=0.242)\\
-\hspace{1em}Malaria season:IRS status=After & 0.001 (P=0.274)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
-\hspace{1em}Malaria season & 0.033 (P<0.001)\\
-\hspace{1em}IRS status=After & -0.007 (P=0.313)\\
-\hspace{1em}Rainy day & 0.002 (P=0.409)\\
-\hspace{1em}Herd protection & 0 (P=0.974)\\
-\hspace{1em}Malaria season:IRS status=After & -0.007 (P=0.405)\\*
-\end{longtable}
+\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-5-1} \end{center}
 
-In the above, ``after'' IRS application (ie, the 6 months following IRS)
-is associated with a significant reduction in absenteeism for both kinds
-of permanent workers, and does not have a signficiant association with a
-change in absenteeism for both kinds of temporary workers.
+\subsection{Number of workers by contract
+type}\label{number-of-workers-by-contract-type}
 
-\section{Suggested changes from
-Menno}\label{suggested-changes-from-menno}
+\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-6-1} \end{center}
 
-Menno sent a write-up on May 17th (see email with subject ``suggestions
-for models'') with some proposed changes to how we model absenteeism.
-What follows is an implementation of those suggestions.
+\subsection{Absenteeism rate by contract
+type}\label{absenteeism-rate-by-contract-type}
 
-\subsection{Model A (without
-seasonality)}\label{model-a-without-seasonality}
+\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-7-1} \end{center}
 
-The below is a characterization of the approach Menno suggested:
+\subsection{Permanent workers absenteeism rate and malaria
+incidence}\label{permanent-workers-absenteeism-rate-and-malaria-incidence}
 
-\begin{itemize}
-\tightlist
-\item
-  We use monthly instead of daily data (modeling a percentage
-  absenteeism rate rather than a binary).
-\item
-  We use fixed effects for year-month combination.
-\item
-  We don't explicitly model for malaria season.
-\item
-  We don't have any interaction terms.
-\item
-  We use dummies for months since IRS.
-\item
-  We use the most recent spraying as the time since IRS (ie, if 5 months
-  ago, and 1 month ago, then the dummy is applied for 1 month ago but
-  not 5 months ago).
-\end{itemize}
+\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-8-1} \end{center}
 
-Using this approach, our panel data looks like this:
+\subsection{Number of temporary workers by job
+type}\label{number-of-temporary-workers-by-job-type}
 
-\begin{verbatim}
-# A tibble: 6 x 6
-  oracle_number group            year_month absent  days months_since_men~
-  <chr>         <chr>            <chr>       <dbl> <int> <fct>            
-1 X045753       Temporary field~ 2015_10       0      13 01               
-2 X040254       Temporary field~ 2016_9        0      30 01               
-3 X040343       Temporary field~ 2014_6        0      30 03               
-4 PF02071       Temporary not f~ 2015_3       80.6    31 06               
-5 PA00795       Permanent field~ 2013_3       67.7    31 02               
-6 PF00699       Permanent not f~ 2016_3        0      31 06               
-\end{verbatim}
+(Skipped due to category incompatibility)
 
-The model formula was described in Menno's email.
+\subsection{Temporary workers absenteeism rate and malaria
+incidence}\label{temporary-workers-absenteeism-rate-and-malaria-incidence}
 
-\includegraphics{images/menno}{[}width=3cm{]}
+(Modified due to category incompatability)
 
-Which translates into R as:
+\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-9-1} \end{center}
 
-\begin{verbatim}
-felm(formula = absent ~ months_since_menno | oracle_number + 
-    year_month | 0 | 0, data = these_data)
-\end{verbatim}
+\newpage
 
-Note that the \texttt{months\_since\_menno} variable has the following
-categories, and is treated as a ``wide'' dummy by the software in
-modeling:
+\section{Other miscellaneous charts}\label{other-miscellaneous-charts}
 
-\begin{verbatim}
-[1] "Before" "01"     "02"     "03"     "04"     "05"     "06"    
-\end{verbatim}
+\subsection{Number of workers by
+department}\label{number-of-workers-by-department}
 
-The results of this model look like this:
+\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-10-1} \end{center}
 
-\begin{longtable}[t]{ll}
-\caption{\label{tab:unnamed-chunk-8}Simple models with year-month fixed effects}\\
-\toprule
-Term & Estimate\\
-\midrule
-\endfirsthead
-\caption[]{Simple models with year-month fixed effects \textit{(continued)}}\\
-\toprule
-Term & Estimate\\
-\midrule
-\endhead
-\
-\endfoot
-\bottomrule
-\endlastfoot
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
-\hspace{1em}IRS status=\_menno01 & 1.425 (P=0.446)\\
-\hspace{1em}IRS status=\_menno02 & 3.603 (P=0.066)\\
-\hspace{1em}IRS status=\_menno03 & 2.049 (P=0.297)\\
-\hspace{1em}IRS status=\_menno04 & 1.357 (P=0.491)\\
-\hspace{1em}IRS status=\_menno05 & 1.351 (P=0.498)\\
-\hspace{1em}IRS status=\_menno06 & 3.358 (P=0.106)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
-\hspace{1em}IRS status=\_menno01 & -0.022 (P=0.982)\\
-\hspace{1em}IRS status=\_menno02 & 0.255 (P=0.805)\\
-\hspace{1em}IRS status=\_menno03 & 1.803 (P=0.085)\\
-\hspace{1em}IRS status=\_menno04 & -0.87 (P=0.406)\\
-\hspace{1em}IRS status=\_menno05 & 0.144 (P=0.894)\\
-\hspace{1em}IRS status=\_menno06 & 0.073 (P=0.947)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
-\hspace{1em}IRS status=\_menno01 & -0.333 (P=0.261)\\
-\hspace{1em}IRS status=\_menno02 & -0.058 (P=0.847)\\
-\hspace{1em}IRS status=\_menno03 & -0.378 (P=0.215)\\
-\hspace{1em}IRS status=\_menno04 & 0.01 (P=0.973)\\
-\hspace{1em}IRS status=\_menno05 & 0.075 (P=0.797)\\
-\hspace{1em}IRS status=\_menno06 & 0.41 (P=0.17)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
-\hspace{1em}IRS status=\_menno01 & 0.08 (P=0.979)\\
-\hspace{1em}IRS status=\_menno02 & 2.989 (P=0.35)\\
-\hspace{1em}IRS status=\_menno03 & -4.094 (P=0.167)\\
-\hspace{1em}IRS status=\_menno04 & -2.276 (P=0.449)\\
-\hspace{1em}IRS status=\_menno05 & -2.259 (P=0.464)\\
-\hspace{1em}IRS status=\_menno06 & 2.409 (P=0.427)\\*
-\end{longtable}
+\subsection{Number of temporary workers over
+time}\label{number-of-temporary-workers-over-time}
 
-In the above, it is apparent that the effects of IRS appear to be either
-(a) in the opposite direction as expected or (b) non-signficant. Here is
-a visualization of IRS' apparent effect (relative to the ``Before''
-period, which is the reference class). Were IRS effective, we would
-expect the black line to be below 0 (the red line).
+\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-11-1} \end{center}
 
-\begin{center}\includegraphics{menno_models_files/figure-latex/unnamed-chunk-9-1} \end{center}
+\subsection{Absenteeism rate over
+time}\label{absenteeism-rate-over-time}
 
-What is going on here? Could it be that IRS has a positive (ie,
-increases) impact on absenteeism? Let's look at the raw data, rather
-than the coefficients. The below chart shows absenteeism (as a
-percentage of the pre-IRS absenteeism average) for workers, broken down
-by season (row) and worker type (column), with point size reflecting the
-number of observations.
+\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-12-1} \end{center}
 
-\begin{center}\includegraphics{menno_models_files/figure-latex/unnamed-chunk-10-1} \end{center}
+\subsection{Absenteeism rate over time by worker
+type}\label{absenteeism-rate-over-time-by-worker-type}
 
-In the above chart, two things stand out:
+\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-13-1} \end{center}
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Variability is greatest among temporary workers.
-\item
-  There appears to be a reduction in absenteeism post IRS for permanent
-  workers \emph{during the malaria season}, but not during the
-  non-malaria season
-\end{enumerate}
+\subsection{Number of workers over time by worker
+type}\label{number-of-workers-over-time-by-worker-type}
 
-So, perhaps controlling for year-month through fixed effects isn't
-enough, since the effect of IRS \emph{interacts} with the season\ldots{}
+\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-14-1} \end{center}
 
-\subsection{Model B (with seasonality)}\label{model-b-with-seasonality}
-
-Using this approach, our panel data looks like this:
-
-\begin{verbatim}
-# A tibble: 6 x 7
-  oracle_number group      year_month absent  days months_since_me~ season
-  <chr>         <chr>      <chr>       <dbl> <int> <fct>            <fct> 
-1 X037697       Temporary~ 2016_7       0       31 05               low   
-2 PF01761       Permanent~ 2015_12      3.23    31 04               high  
-3 X040726       Temporary~ 2015_5       3.23    31 02               high  
-4 PF00774       Permanent~ 2014_6       3.33    30 04               low   
-5 X043249       Temporary~ 2016_4       0       30 06               high  
-6 X037700       Permanent~ 2013_2       0       27 02               high  
-\end{verbatim}
-
-Our new model specification (in R) is:
-
-\begin{verbatim}
-felm(formula = absent ~ months_since_menno * season | oracle_number + 
-    year_month | 0 | 0, data = these_data)
-\end{verbatim}
-
-The results of this model look like this:
-
-\begin{longtable}[t]{ll}
-\caption{\label{tab:unnamed-chunk-13}Simple models with year-month fixed effects and seasonality-IRS interaction term}\\
-\toprule
-Term & Estimate\\
-\midrule
-\endfirsthead
-\caption[]{Simple models with year-month fixed effects and seasonality-IRS interaction term \textit{(continued)}}\\
-\toprule
-Term & Estimate\\
-\midrule
-\endhead
-\
-\endfoot
-\bottomrule
-\endlastfoot
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
-\hspace{1em}IRS status=\_menno01 & 175.053 (P=0.532)\\
-\hspace{1em}IRS status=\_menno02 & 457.807 (P=0.115)\\
-\hspace{1em}IRS status=\_menno03 & 505.425 (P=0.061)\\
-\hspace{1em}IRS status=\_menno04 & 233.373 (P=0.357)\\
-\hspace{1em}IRS status=\_menno05 & 227.399 (P=0.372)\\
-\hspace{1em}IRS status=\_menno06 & 425.265 (P=0.121)\\
-\hspace{1em}Malaria season & 887.601 (P=0.703)\\
-\hspace{1em}IRS status=\_menno01:Malaria season & -58.338 (P=0.872)\\
-\hspace{1em}IRS status=\_menno02:Malaria season & -174.777 (P=0.642)\\
-\hspace{1em}IRS status=\_menno03:Malaria season & -603.926 (P=0.105)\\
-\hspace{1em}IRS status=\_menno04:Malaria season & -220.413 (P=0.56)\\
-\hspace{1em}IRS status=\_menno05:Malaria season & -213.141 (P=0.582)\\
-\hspace{1em}IRS status=\_menno06:Malaria season & -191.654 (P=0.633)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
-\hspace{1em}IRS status=\_menno01 & 39.635 (P=0.788)\\
-\hspace{1em}IRS status=\_menno02 & 14.833 (P=0.919)\\
-\hspace{1em}IRS status=\_menno03 & 114.878 (P=0.439)\\
-\hspace{1em}IRS status=\_menno04 & 55.896 (P=0.694)\\
-\hspace{1em}IRS status=\_menno05 & 93.301 (P=0.502)\\
-\hspace{1em}IRS status=\_menno06 & 86.299 (P=0.556)\\
-\hspace{1em}Malaria season & NaN (PNA)\\
-\hspace{1em}IRS status=\_menno01:Malaria season & -72.876 (P=0.707)\\
-\hspace{1em}IRS status=\_menno02:Malaria season & 23.995 (P=0.903)\\
-\hspace{1em}IRS status=\_menno03:Malaria season & 127.34 (P=0.523)\\
-\hspace{1em}IRS status=\_menno04:Malaria season & -296.411 (P=0.137)\\
-\hspace{1em}IRS status=\_menno05:Malaria season & -185.639 (P=0.372)\\
-\hspace{1em}IRS status=\_menno06:Malaria season & -171.932 (P=0.416)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
-\hspace{1em}IRS status=\_menno01 & -35.174 (P=0.339)\\
-\hspace{1em}IRS status=\_menno02 & -6.152 (P=0.869)\\
-\hspace{1em}IRS status=\_menno03 & -12.472 (P=0.732)\\
-\hspace{1em}IRS status=\_menno04 & 18.095 (P=0.611)\\
-\hspace{1em}IRS status=\_menno05 & 33.497 (P=0.346)\\
-\hspace{1em}IRS status=\_menno06 & 9.614 (P=0.816)\\
-\hspace{1em}Malaria season & 44.327 (P=0.309)\\
-\hspace{1em}IRS status=\_menno01:Malaria season & 8.557 (P=0.881)\\
-\hspace{1em}IRS status=\_menno02:Malaria season & 3.8 (P=0.948)\\
-\hspace{1em}IRS status=\_menno03:Malaria season & -75.835 (P=0.21)\\
-\hspace{1em}IRS status=\_menno04:Malaria season & -51.282 (P=0.397)\\
-\hspace{1em}IRS status=\_menno05:Malaria season & -70.353 (P=0.217)\\
-\hspace{1em}IRS status=\_menno06:Malaria season & 63.717 (P=0.259)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
-\hspace{1em}IRS status=\_menno01 & 414.405 (P=0.319)\\
-\hspace{1em}IRS status=\_menno02 & 90.379 (P=0.825)\\
-\hspace{1em}IRS status=\_menno03 & 353.397 (P=0.441)\\
-\hspace{1em}IRS status=\_menno04 & -565.881 (P=0.232)\\
-\hspace{1em}IRS status=\_menno05 & -355.43 (P=0.388)\\
-\hspace{1em}IRS status=\_menno06 & 45.199 (P=0.895)\\
-\hspace{1em}Malaria season & -459.147 (P=0.519)\\
-\hspace{1em}IRS status=\_menno01:Malaria season & -724.867 (P=0.206)\\
-\hspace{1em}IRS status=\_menno02:Malaria season & 661.492 (P=0.283)\\
-\hspace{1em}IRS status=\_menno03:Malaria season & -1223.913 (P=0.031)\\
-\hspace{1em}IRS status=\_menno04:Malaria season & 529.98 (P=0.364)\\
-\hspace{1em}IRS status=\_menno05:Malaria season & 324.283 (P=0.578)\\
-\hspace{1em}IRS status=\_menno06:Malaria season & 817.731 (P=0.227)\\*
-\end{longtable}
-
-The below shows model coefficients:
-
-\begin{center}\includegraphics{menno_models_files/figure-latex/unnamed-chunk-14-1} \end{center}
-
-\subsection{Model C (with seasonality and herd
-score)}\label{model-c-with-seasonality-and-herd-score}
-
-We now introduce herd immunity to the model. This is introduced as an
-additive effect; perhaps it should be interactive with IRS? Using this
-approach, our panel data looks like this:
-
-\begin{verbatim}
-# A tibble: 6 x 8
-  oracle_number group       year_month absent  days  herd months_since_me~
-  <chr>         <chr>       <chr>       <dbl> <int> <dbl> <fct>           
-1 PF02492       Permanent ~ 2013_3       9.68    31 136   01              
-2 X034289       Temporary ~ 2016_9       0       30 276   02              
-3 PF00699       Permanent ~ 2016_9       0       30 282   03              
-4 X044684       Temporary ~ 2014_9       0       30 248   01              
-5 X045133       Temporary ~ 2015_8       6.45    31 156   06              
-6 X041009       Temporary ~ 2015_6       0       30  97.1 02              
-# ... with 1 more variable: season <fct>
-\end{verbatim}
-
-Our new model specification (in R) is:
-
-\begin{verbatim}
-felm(formula = absent ~ months_since_menno * season | oracle_number + 
-    year_month | 0 | 0, data = these_data)
-\end{verbatim}
-
-The results of this model look like this:
-
-\begin{longtable}[t]{ll}
-\caption{\label{tab:unnamed-chunk-17}Simple models with year-month fixed effects, seasonality-IRS interaction term, and herd score}\\
-\toprule
-Term & Estimate\\
-\midrule
-\endfirsthead
-\caption[]{Simple models with year-month fixed effects, seasonality-IRS interaction term, and herd score \textit{(continued)}}\\
-\toprule
-Term & Estimate\\
-\midrule
-\endhead
-\
-\endfoot
-\bottomrule
-\endlastfoot
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
-\hspace{1em}IRS status=\_menno01 & 175.053 (P=0.532)\\
-\hspace{1em}IRS status=\_menno02 & 457.807 (P=0.115)\\
-\hspace{1em}IRS status=\_menno03 & 505.425 (P=0.061)\\
-\hspace{1em}IRS status=\_menno04 & 233.373 (P=0.357)\\
-\hspace{1em}IRS status=\_menno05 & 227.399 (P=0.372)\\
-\hspace{1em}IRS status=\_menno06 & 425.265 (P=0.121)\\
-\hspace{1em}Malaria season & 887.601 (P=0.703)\\
-\hspace{1em}IRS status=\_menno01:Malaria season & -58.338 (P=0.872)\\
-\hspace{1em}IRS status=\_menno02:Malaria season & -174.777 (P=0.642)\\
-\hspace{1em}IRS status=\_menno03:Malaria season & -603.926 (P=0.105)\\
-\hspace{1em}IRS status=\_menno04:Malaria season & -220.413 (P=0.56)\\
-\hspace{1em}IRS status=\_menno05:Malaria season & -213.141 (P=0.582)\\
-\hspace{1em}IRS status=\_menno06:Malaria season & -191.654 (P=0.633)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
-\hspace{1em}IRS status=\_menno01 & 39.635 (P=0.788)\\
-\hspace{1em}IRS status=\_menno02 & 14.833 (P=0.919)\\
-\hspace{1em}IRS status=\_menno03 & 114.878 (P=0.439)\\
-\hspace{1em}IRS status=\_menno04 & 55.896 (P=0.694)\\
-\hspace{1em}IRS status=\_menno05 & 93.301 (P=0.502)\\
-\hspace{1em}IRS status=\_menno06 & 86.299 (P=0.556)\\
-\hspace{1em}Malaria season & NaN (PNA)\\
-\hspace{1em}IRS status=\_menno01:Malaria season & -72.876 (P=0.707)\\
-\hspace{1em}IRS status=\_menno02:Malaria season & 23.995 (P=0.903)\\
-\hspace{1em}IRS status=\_menno03:Malaria season & 127.34 (P=0.523)\\
-\hspace{1em}IRS status=\_menno04:Malaria season & -296.411 (P=0.137)\\
-\hspace{1em}IRS status=\_menno05:Malaria season & -185.639 (P=0.372)\\
-\hspace{1em}IRS status=\_menno06:Malaria season & -171.932 (P=0.416)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
-\hspace{1em}IRS status=\_menno01 & -35.174 (P=0.339)\\
-\hspace{1em}IRS status=\_menno02 & -6.152 (P=0.869)\\
-\hspace{1em}IRS status=\_menno03 & -12.472 (P=0.732)\\
-\hspace{1em}IRS status=\_menno04 & 18.095 (P=0.611)\\
-\hspace{1em}IRS status=\_menno05 & 33.497 (P=0.346)\\
-\hspace{1em}IRS status=\_menno06 & 9.614 (P=0.816)\\
-\hspace{1em}Malaria season & 44.327 (P=0.309)\\
-\hspace{1em}IRS status=\_menno01:Malaria season & 8.557 (P=0.881)\\
-\hspace{1em}IRS status=\_menno02:Malaria season & 3.8 (P=0.948)\\
-\hspace{1em}IRS status=\_menno03:Malaria season & -75.835 (P=0.21)\\
-\hspace{1em}IRS status=\_menno04:Malaria season & -51.282 (P=0.397)\\
-\hspace{1em}IRS status=\_menno05:Malaria season & -70.353 (P=0.217)\\
-\hspace{1em}IRS status=\_menno06:Malaria season & 63.717 (P=0.259)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
-\hspace{1em}IRS status=\_menno01 & 414.405 (P=0.319)\\
-\hspace{1em}IRS status=\_menno02 & 90.379 (P=0.825)\\
-\hspace{1em}IRS status=\_menno03 & 353.397 (P=0.441)\\
-\hspace{1em}IRS status=\_menno04 & -565.881 (P=0.232)\\
-\hspace{1em}IRS status=\_menno05 & -355.43 (P=0.388)\\
-\hspace{1em}IRS status=\_menno06 & 45.199 (P=0.895)\\
-\hspace{1em}Malaria season & -459.147 (P=0.519)\\
-\hspace{1em}IRS status=\_menno01:Malaria season & -724.867 (P=0.206)\\
-\hspace{1em}IRS status=\_menno02:Malaria season & 661.492 (P=0.283)\\
-\hspace{1em}IRS status=\_menno03:Malaria season & -1223.913 (P=0.031)\\
-\hspace{1em}IRS status=\_menno04:Malaria season & 529.98 (P=0.364)\\
-\hspace{1em}IRS status=\_menno05:Malaria season & 324.283 (P=0.578)\\
-\hspace{1em}IRS status=\_menno06:Malaria season & 817.731 (P=0.227)\\*
-\end{longtable}
+\section{Bibliography}\label{bibliography}
 

### 2020-06-20 inst/rmd/defunct/models.Rmd
updated

---
title: "Concept note: sequential models"
subtitle: "Pradhan, Sicuri, Brew"
date: ''
output: 
  pdf_document:
    keep_tex: true
    latex_engine: xelatex
    includes:
      in_header: preamble.sty
fig_width: 6
fig_height: 6
bibliography: bibliography.bib
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# csl: journal-of-health-economics.csl

# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(RColorBrewer)
library(extrafont)
library(kableExtra)
library(raster)
loadfonts()
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               fig.height = 6,
               fig.width = 7,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               dev = "cairo_pdf",
               fig.pos="!h",
               fig.align = 'center')
opts_knit$set(width=75)
options(xtable.comment = FALSE)

```

```{r, eval = TRUE}
source('paper_prepare.R')
```

\begin{center}
\begin{large}

A few different approaches to modeling absenteeism in the Maragra data

\end{large}
\end{center}

\vspace{5mm}

\begin{center}
\textbf{Overview}  
\end{center}

\vspace{5mm}
\begin{center}
\begin{changemargin}{3cm}{3cm}

This document contains the results (both tabular and visual) of running different permutations of variables through linear fixed effects models, with the purpose of understanding the relationship between IRS and worker absenteeism.

\end{changemargin}
\end{center}

\vspace{20mm}

\noindent\fbox{%
    \parbox{\textwidth}{%
        \subsection*{Main points}
        \begin{itemize}
          \item Meant for internal review
          \item Uses loops to permutate different variable combinations in models
          \item Charts added (with only relevant variables) for quick interpretation
        \end{itemize}
        \vspace{2mm}
    }%
}

\vfill
\null

\subsection*{Desinataires}
\textbf{Elisa Sicuri; Menno Pradhan}

\vspace{3mm}

\newpage

# The current approach

In the current draft of the paper (\href{https://docs.google.com/document/d/1bUWRBCgVcgjSPHchIQxiTG8Vwv5hV1GLU4Tlu386sWA/edit#}{HERE}
), the model formula is:

$$
\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * IRS_{it} + \hat{\beta}_2{RainyDay_{t}} + \hat{\beta}_3{Herd_{it}} +  \alpha_i + \delta_t + \upsilon_{it}
$$

In R, using the `felm` library, this is written as follows:

```{r, echo = FALSE}
protection_models[[1]]$call
```

The above notation shows `absent` (binary 1,0) as a function of the interaction of binary `season` (high vs. low) and binary `months_since` (before vs. after) plus binary `rainy_day` (whether or not it rained on the particular day) plus continuous numeric `herd` (ie, the protection afforded by others) plus fixed effects for both the worker (`oracle_number`) and the year (`malaria_year`).  $\alpha_i$ represents the time invariant worker fixed effects, and $\delta_i$ represents the fixed effect of the particular malaria season. $\upsilon$ is the error term. This model is run separately for the 4 different worker groups, yielding the following results:

```{r, results = 'asis'}
make_models_table(model_list = protection_models, the_caption = "All absenteeism with herd immunity: model results", type = 'latex', multiplier = 1)
```

In the above, "after" IRS application (ie, the 6 months following IRS) is associated with a significant reduction in absenteeism for both kinds of permanent workers, and does not have a signficiant association with a change in absenteeism for both kinds of temporary workers.

# Sequential models

What follows are different combinations of variables in the same modeling structure (linear fixed effects using R's `felm` library, with different models for the 4 different worker types).

```{r}
make_model <- function(call = 'absent ~ season*months_since + rainy_day + herd  | oracle_number + malaria_year'){
  sequential_models <- list()
  groups <- sort(unique(model_data$group))
  # library(lmerTest)
  # library(nlme)
  for (i in 1:length(groups)){
    message(i)
    this_group <- groups[i]
    message(this_group)
    these_data <- model_data %>% filter(group == this_group)
    this_model <- felm(as.formula(paste0(call, ' | 0 | 0')),
                       data = these_data)
    sequential_models[[i]] <- this_model
  }
  names(sequential_models) <- groups
  return(sequential_models)
}
display_model <- function(made_model,
                          caption = '',
                          type = 'latex',
                          multiplier = 100){
  make_models_table(model_list = made_model,
                    the_caption = caption,
                    type = type,
                    multiplier = multiplier)
}
# display_model(make_model())

# Create data frame of different scenarios
library(combinat)

model_data$before_after <- model_data$months_since
model_data$time_since <- model_data$months_since_menno
scenarios <-
  expand.grid(interaction = c('season * before_after',
                              'season * time_since',
                              'incidence * before_after',
                              'incidence * time_since'),
              terms = c('rainy_day', 'herd', 'rainy_day + herd'),
              fe = c('oracle_number', 'malaria_year', 'year_month', 
                     'oracle_number + malaria_year',
                     'oracle_number + year_month',
                     'oracle_number',
                     '0'),
              stringsAsFactors = FALSE) %>%
  mutate(formula = paste0('absent ~ ', interaction, ' + ',
                          terms, ' | ', fe)) %>%
  mutate(description = 
           paste0('Absenteeism as a function of ',
                  '(interaction of ', gsub('*', 'and', interaction, fixed = TRUE),
                  ') + ', terms, ' with ',
                  ifelse(fe == '0', 'no fixed effects',
                         ' the following fixed effects: '),
                  ifelse(fe == '0', '',
                         gsub(' + ', ' and ', fe, fixed = TRUE))))

# Create some dummy data
dum <- expand.grid(season = sort(unique(model_data$season)),
                   time_since = sort(unique(model_data$time_since)),
                   rainy_day = sort(unique(model_data$rainy_day)),
                   incidence = quantile(model_data$incidence, probs = c(0.1, 0.9)),
                   herd = quantile(model_data$herd, probs = c(0.1, 0.9)),
                   group = sort(unique(model_data$group)))

plot_model <- function(made_model = made_model){
  plot_data <- data.frame(
    variable = unlist(lapply(made_model, function(x){paste0(names(coef(x)))})),
    coefficient = as.numeric(as.character(unlist(lapply(made_model, function(x){paste0(coef(x))}))))
  )
  plot_data <- plot_data %>%
    mutate(group = rep(groups, each = nrow(plot_data) / 4))
  row.names(plot_data) <- NULL

  plot_data <- plot_data %>%
    filter(!variable %in% c('incidence', 'herd', 'rainy_dayTRUE',
                            '(Intercept)', 'seasonhigh')) #%>%
    # mutate(!grepl('incidence', variable))
  plot_data$variable <- gsub('time_since', 'Months post IRS: ', plot_data$variable)
  has_season <- any(grepl('seasonhigh', plot_data$variable))
  plot_data$variable <- gsub('seasonhigh', 'Malaria season', plot_data$variable)    
  plot_data$variable <- gsub('seasonlow', 'Low malaria season', plot_data$variable)
  # plot_data$variable <- gsub('months_since', 'Months since IRS: ', plot_data$variable)
  plot_data$variable <- gsub('months_since', 'IRS status=', plot_data$variable)
  plot_data$variable <- gsub('department', 'Department: ', plot_data$variable)
  plot_data$variable <- gsub('precipitation', 'Precipitation (mm) ', plot_data$variable)
  plot_data$variable <- gsub('maragra_fabricaTRUE', 'Living at factory', plot_data$variable)
  plot_data$variable <- gsub('sexM', 'Male', plot_data$variable)
  plot_data$variable <- gsub('permanent_or_temporaryTemporary', 'Temp contract', plot_data$variable)
  plot_data$variable <- gsub('rainy_dayTRUE', 'Rainy day', plot_data$variable)
  plot_data$variable <- gsub('rainyTRUE', 'Rainy day', plot_data$variable)
  plot_data$variable <- gsub('on_siteTRUE', 'On site', plot_data$variable)
  plot_data$variable <- gsub('on_siteFALSE', 'Off site', plot_data$variable)
  plot_data$variable <- gsub('fieldNot field worker', 'Not field worker', plot_data$variable)
  plot_data$variable <- gsub('herd', 'Herd protection', plot_data$variable)
  plot_data$variable <- gsub('before_after', 'IRS time=', plot_data$variable)
  plot_data$variable <- gsub('time_since', 'Months after=', plot_data$variable)
  plot_data$variable <- gsub('incidence', 'Incidence', plot_data$variable)

  ggplot(data = plot_data,
         aes(x = variable,
             y = coefficient)) +
    facet_wrap(~group) +
    geom_bar(stat = 'identity',
             alpha = 0.6) +
    theme(axis.text.x = element_text(angle = 90)) +
    labs(x = '',
         y = 'Coefficient') +
    geom_label(aes(x = variable,
                   y = coefficient + (0.05 * max(coefficient)),
                   label = round(coefficient, digits = 4)),
               size = 2,
               alpha = 0.5) +
    geom_hline(yintercept = 0, color = 'red', lty = 2, alpha = 0.6)
}

```

# 0. Months since absenteeism only

```{r, results = 'asis'}
call <- 'absent ~ time_since | 0'
made_model <- make_model(call = call)
display_model(made_model = made_model)
plot_model(made_model = made_model)
```

\newpage

```{r}
rmds <- c()
titles <- c()
nn <- nrow(scenarios)
# nn <- 10
for(i in 1:nn){
  titles[i] <- paste0(i, '. ', scenarios$description[i])
  rmds[i] <- 
    paste0("call <- '",scenarios$formula[i], "'
made_model <- make_model(call = call)
display_model(made_model = made_model);
plot_model(made_model = made_model)
")
}

chunks <- paste0('***\n## ',
                 titles, 
                 "\n***\n```{r, results = 'asis'}\n\n",
                 rmds,
                 "\n```\n\n\\newpage\n\n")

# Write our order / child-calls to a doc
file_connection <- file('children.Rmd')
writeLines(paste0('---\noutput: pdf_document\n---\n\n', 
                  chunks), 
           file_connection)
close(file_connection)

```

```{r child='children.Rmd', echo = FALSE}
```

```{r, results='hide'}
# Having already called the auto-generated children.Rmd, we can delete it
file.remove('children.Rmd')
```

### 2020-06-20 inst/rmd/nov2018/nov2018.Rmd
updated

--- 2019-02-17 inst/rmd/final/final.Rmd
@@
 ---
-title: "Final analysis for paper"
+title: "Maragra update: December 2018"
 subtitle: "Brew, Pradhan, Sicuri"
@@
 
-# Manuscript draft
+\begin{center}
+\begin{large}
+
+Maragra models
+
+\end{large}
+\end{center}
+
+\vspace{5mm}
+
+\begin{center}
+\textbf{Overview}  
+\end{center}
+
+\vspace{5mm}
+\begin{center}
+\begin{changemargin}{3cm}{3cm}
+
+This document contains an overview of the modeling approach and simulations results.
+
+\end{changemargin}
+\end{center}
+
+\vspace{20mm}
+
+\noindent\fbox{%
+    \parbox{\textwidth}{%
+        \subsection*{Main points}
+        \begin{itemize}
+          \item We have a new modeling approach
+          \item We have carried out basic simulations for several malaria control strategies
+          \item Results are consistent/coherent
+          \item We have not yet implemented geographical variation
+        \end{itemize}
+        \vspace{2mm}
+    }%
+}
+
+\vfill
+\null
+
+\subsection*{Desinataires}
+\textbf{Elisa Sicuri; Menno Pradhan}
+
+\vspace{3mm}
+
+\newpage
+
+# Current paper
 
 Draft of the paper (\href{https://docs.google.com/document/d/1bUWRBCgVcgjSPHchIQxiTG8Vwv5hV1GLU4Tlu386sWA/edit#}{HERE}
+
+# Code
+
+In R, our model looks like this
+
+```{r, echo = FALSE}
+final_model$call
+```
+
+Where `protection` is the sum of all nearby IRS, weighted by the inverse of the distance from the residence of the person (including the person's own residence).
+
+\newpage
+
+# Reproduction of charts
+
+## Number of workers by department
+
+```{r}
+plot_data <- ab_panel %>%
+  left_join(workers %>% dplyr::select(oracle_number,
+                                      department,
+                                      permanent_or_temporary)) %>%
+  mutate(year_month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+  mutate(group = department) %>%
+  # mutate(field = ifelse(department == 'Field', 
+  #                            'Field worker',
+  #                            'Not field worker')) %>%
+  # mutate(group = paste0(permanent_or_temporary, ' ',
+  #                         tolower(field))) %>%
+  group_by(group, year_month) %>%
+  summarise(n = length(unique(oracle_number)))
+
+ggplot(data = plot_data,
+       aes(x = year_month,
+           y = n,
+           color = group)) +
+  geom_line()
+```
+
+## Number of temporary workers over time
+
+```{r}
+plot_data <- ab_panel %>%
+  left_join(workers %>% dplyr::select(oracle_number,
+                                      department,
+                                      permanent_or_temporary)) %>%
+  mutate(year_month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+  # mutate(group = department) %>%
+  mutate(group = ifelse(department == 'Field',
+                             'Field worker',
+                             'Not field worker')) %>%
+  # mutate(group = paste0(permanent_or_temporary, ' ',
+  #                         tolower(field))) %>%
+  group_by(group, year_month) %>%
+  summarise(n = length(unique(oracle_number)))
+
+ggplot(data = plot_data,
+       aes(x = year_month,
+           y = n,
+           color = group)) +
+  geom_line()
+```
+
+## Absenteeism rate over time
+
+```{r}
+plot_data <- ab_panel %>%
+  left_join(workers %>% dplyr::select(oracle_number,
+                                      department,
+                                      permanent_or_temporary)) %>%
+  mutate(year_month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+  mutate(group = department) %>%
+  # mutate(group = ifelse(department == 'Field',
+  #                            'Field worker',
+  #                            'Not field worker')) %>%
+  # mutate(group = paste0(permanent_or_temporary, ' ',
+  #                         tolower(field))) %>%
+  group_by(group, year_month) %>%
+  summarise(abs = length(which(absent)),
+            eligibles = n()) %>%
+  mutate(ab_rate = abs / eligibles * 100)
+
+ggplot(data = plot_data,
+       aes(x = year_month,
+           y = ab_rate,
+           color = group)) +
+  geom_line()
+```
+
+## Absenteeism rate over time by worker type
+
+```{r}
+plot_data <- ab_panel %>%
+  left_join(workers %>% dplyr::select(oracle_number,
+                                      department,
+                                      permanent_or_temporary)) %>%
+  mutate(year_month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+  # mutate(group = department) %>%
+  mutate(field = ifelse(department == 'Field',
+                             'Field worker',
+                             'Not field worker')) %>%
+  mutate(group = paste0(permanent_or_temporary, ' ',
+                          tolower(field))) %>%
+  group_by(group, year_month) %>%
+  summarise(abs = length(which(absent)),
+            eligibles = n()) %>%
+  mutate(ab_rate = abs / eligibles * 100)
+
+ggplot(data = plot_data,
+       aes(x = year_month,
+           y = ab_rate,
+           color = group)) +
+  geom_line()
+```
+
+## Number of workers over time by worker type
+
+```{r}
+plot_data <- ab_panel %>%
+  left_join(workers %>% dplyr::select(oracle_number,
+                                      department,
+                                      permanent_or_temporary)) %>%
+  mutate(year_month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+  # mutate(group = department) %>%
+  mutate(field = ifelse(department == 'Field',
+                             'Field worker',
+                             'Not field worker')) %>%
+  mutate(group = paste0(permanent_or_temporary, ' ',
+                          tolower(field))) %>%
+  group_by(group, year_month) %>%
+  summarise(n = length(unique(oracle_number)))
+
+ggplot(data = plot_data,
+       aes(x = year_month,
+           y = n,
+           color = group)) +
+  geom_line()
+```
 

### 2020-06-20 inst/rmd/nov2018/nov2018.tex
updated

--- 2020-06-20 inst/rmd/defunct/feb2019_xinavane_reproduction/feb2019_xinavane_reproduction.tex
@@
 \hypersetup{unicode=true,
-            pdftitle={February 2019 - Xinavane charts reproduction},
+            pdftitle={Maragra update: December 2018},
             pdfborder={0 0 0},
@@
 
-  \title{February 2019 - Xinavane charts reproduction}
+  \title{Maragra update: December 2018}
     \pretitle{\vspace{\droptitle}\centering\huge}
   \posttitle{\par}
-  \subtitle{Using Maragra data}
+  \subtitle{Brew, Pradhan, Sicuri}
   \author{}
@@
 
-Charts
+Maragra models
 
@@
 
-This document contains charts from Maragra data, meant for comparison with Xinavane. It was created on February 5, 2019 by Joe Brew for Laia Cirera and Elisa Sicuri.
+This document contains an overview of the modeling approach and simulations results.
 
@@
     \parbox{\textwidth}{%
-        \subsection*{Action}
+        \subsection*{Main points}
         \begin{itemize}
-          \item Review charts herein
-          \item Request clarification or ask questions if applicable
+          \item We have a new modeling approach
+          \item We have carried out basic simulations for several malaria control strategies
+          \item Results are consistent/coherent
+          \item We have not yet implemented geographical variation
         \end{itemize}
@@
 
-\textbf{Laia Cirera; Elisa Sicuri}
+\textbf{Elisa Sicuri; Menno Pradhan}
 
@@
 
-\section{Malaria prevalence in the region, (2013-2017), Manhiça
-(control)}\label{malaria-prevalence-in-the-region-2013-2017-manhica-control}
+\section{Current paper}\label{current-paper}
 
-\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-3-1} \end{center}
+Draft of the paper
+(\href{https://docs.google.com/document/d/1bUWRBCgVcgjSPHchIQxiTG8Vwv5hV1GLU4Tlu386sWA/edit#}{HERE}
 
-\section{Malaria prevalence in the region,
-(2013-2017)}\label{malaria-prevalence-in-the-region-2013-2017}
+\section{Code}\label{code}
 
-Magude (intervention)
+In R, our model looks like this
 
-\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-4-1} \end{center}
+\begin{verbatim}
+felm(formula = log(absent + 1) ~ protection_var + (rain_var) | 
+    oracle_number | 0 | 0, data = final_data)
+\end{verbatim}
 
-\section{Both locations}\label{both-locations}
-
-\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-5-1} \end{center}
-
-\subsection{Number of workers by contract
-type}\label{number-of-workers-by-contract-type}
-
-\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-6-1} \end{center}
-
-\subsection{Absenteeism rate by contract
-type}\label{absenteeism-rate-by-contract-type}
-
-\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-7-1} \end{center}
-
-\subsection{Permanent workers absenteeism rate and malaria
-incidence}\label{permanent-workers-absenteeism-rate-and-malaria-incidence}
-
-\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-8-1} \end{center}
-
-\subsection{Number of temporary workers by job
-type}\label{number-of-temporary-workers-by-job-type}
-
-(Skipped due to category incompatibility)
-
-\subsection{Temporary workers absenteeism rate and malaria
-incidence}\label{temporary-workers-absenteeism-rate-and-malaria-incidence}
-
-(Modified due to category incompatability)
-
-\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-9-1} \end{center}
+Where \texttt{protection} is the sum of all nearby IRS, weighted by the
+inverse of the distance from the residence of the person (including the
+person's own residence).
 
@@
 
-\section{Other miscellaneous charts}\label{other-miscellaneous-charts}
+\section{Reproduction of charts}\label{reproduction-of-charts}
 
@@
 
-\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-10-1} \end{center}
+\begin{center}\includegraphics{nov2018_files/figure-latex/unnamed-chunk-4-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-11-1} \end{center}
+\begin{center}\includegraphics{nov2018_files/figure-latex/unnamed-chunk-5-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-12-1} \end{center}
+\begin{center}\includegraphics{nov2018_files/figure-latex/unnamed-chunk-6-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-13-1} \end{center}
+\begin{center}\includegraphics{nov2018_files/figure-latex/unnamed-chunk-7-1} \end{center}
 
@@
 
-\begin{center}\includegraphics{feb2019_xinavane_reproduction_files/figure-latex/unnamed-chunk-14-1} \end{center}
+\begin{center}\includegraphics{nov2018_files/figure-latex/unnamed-chunk-8-1} \end{center}
+
+\section{Descriptive overview}\label{descriptive-overview}
+
+\subsection{Herd-only protection
+score}\label{herd-only-protection-score}
+
+\begin{tabular}{l|r}
+\hline
+Quantile & Absenteeism rate\\
+\hline
+Protection 1 & 10.641401\\
+\hline
+Protection 2 & 8.289246\\
+\hline
+Protection 3 & 8.312732\\
+\hline
+Protection 4 & 8.360836\\
+\hline
+\end{tabular}
+
+\begin{center}\includegraphics{nov2018_files/figure-latex/unnamed-chunk-10-1} \end{center}
+
+\newpage
+
+\subsection{Individual-only protection
+score}\label{individual-only-protection-score}
+
+\begin{tabular}{l|r}
+\hline
+Quantile & Absenteeism rate\\
+\hline
+Protection 1 & 9.443294\\
+\hline
+Protection 2 & 8.464487\\
+\hline
+Protection 3 & 7.748320\\
+\hline
+Protection 4 & 7.792721\\
+\hline
+\end{tabular}
+
+\begin{center}\includegraphics{nov2018_files/figure-latex/unnamed-chunk-12-1} \end{center}
+
+\newpage
+
+\subsection{Overall protection score}\label{overall-protection-score}
+
+\begin{tabular}{l|l|r}
+\hline
+Quantile & Group & Absenteeism rate\\
+\hline
+Protection 1 & Permanent field worker & 13.687257\\
+\hline
+Protection 1 & Permanent not field worker & 11.415405\\
+\hline
+Protection 1 & Temporary field worker & 1.195219\\
+\hline
+Protection 1 & Temporary not field worker & 4.902265\\
+\hline
+Protection 2 & Permanent field worker & 11.210093\\
+\hline
+Protection 2 & Permanent not field worker & 10.136308\\
+\hline
+Protection 2 & Temporary field worker & 3.032894\\
+\hline
+Protection 2 & Temporary not field worker & 8.722060\\
+\hline
+Protection 3 & Permanent field worker & 10.542001\\
+\hline
+Protection 3 & Permanent not field worker & 10.447119\\
+\hline
+Protection 3 & Temporary field worker & 3.660856\\
+\hline
+Protection 3 & Temporary not field worker & 7.699303\\
+\hline
+Protection 4 & Permanent field worker & 11.371741\\
+\hline
+Protection 4 & Permanent not field worker & 10.374719\\
+\hline
+Protection 4 & Temporary field worker & 3.824707\\
+\hline
+Protection 4 & Temporary not field worker & 7.565523\\
+\hline
+\end{tabular}
+
+\begin{center}\includegraphics{nov2018_files/figure-latex/unnamed-chunk-14-1} \end{center}
+
+\newpage
+
+\subsection{Precipitation lag}\label{precipitation-lag}
+
+\begin{tabular}{l|l|r}
+\hline
+Quantile & Group & Absenteeism rate\\
+\hline
+Protection 1 & Permanent field worker & 8.630077\\
+\hline
+Protection 1 & Permanent not field worker & 8.295415\\
+\hline
+Protection 1 & Temporary field worker & 3.535996\\
+\hline
+Protection 1 & Temporary not field worker & 7.870018\\
+\hline
+Protection 2 & Permanent field worker & 8.811771\\
+\hline
+Protection 2 & Permanent not field worker & 7.021752\\
+\hline
+Protection 2 & Temporary field worker & 3.439914\\
+\hline
+Protection 2 & Temporary not field worker & 7.246022\\
+\hline
+Protection 3 & Permanent field worker & 11.562734\\
+\hline
+Protection 3 & Permanent not field worker & 16.734109\\
+\hline
+Protection 3 & Temporary field worker & 3.099666\\
+\hline
+Protection 3 & Temporary not field worker & 5.810022\\
+\hline
+Protection 4 & Permanent field worker & 16.178230\\
+\hline
+Protection 4 & Permanent not field worker & 10.843800\\
+\hline
+Protection 4 & Temporary field worker & 2.916452\\
+\hline
+Protection 4 & Temporary not field worker & 9.779668\\
+\hline
+\end{tabular}
+
+\begin{center}\includegraphics{nov2018_files/figure-latex/unnamed-chunk-16-1} \end{center}
+
+\newpage
+
+\section{Results}\label{results}
+
+\subsection{Regression table}\label{regression-table}
+
+The below table shows the results of the model devised thus far.
+
+\begin{verbatim}
+
+Call:
+   felm(formula = log(absent + 1) ~ protection_var + (rain_var) |      oracle_number | 0 | 0, data = final_data)
+
+Residuals:
+     Min       1Q   Median       3Q      Max 
+-0.43174 -0.07695 -0.05705 -0.01099  0.69903
+
+Coefficients:
+                 Estimate Std. Error t value Pr(>|t|)    
+protection_var -4.462e-05  4.895e-06  -9.116   <2e-16 ***
+rain_var        2.993e-03  2.042e-04  14.659   <2e-16 ***
+---
+Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+
+Residual standard error: 0.1916 on 308342 degrees of freedom
+  (14642 observations deleted due to missingness)
+Multiple R-squared(full model): 0.05462   Adjusted R-squared: 0.05273 
+Multiple R-squared(proj model): 0.001288   Adjusted R-squared: -0.0007107 
+F-statistic(full model):28.87 on 617 and 308342 DF, p-value: < 2.2e-16 
+F-statistic(proj model): 198.8 on 2 and 308342 DF, p-value: < 2.2e-16 
+\end{verbatim}
+
+\subsection{Intepretation}\label{intepretation}
+
+The below charts show predicted absenteeism rates at different rain and
+protection levels.
+
+\subsubsection{Precipitation's effect}\label{precipitations-effect}
+
+\begin{center}\includegraphics{nov2018_files/figure-latex/unnamed-chunk-20-1} \end{center}
+
+\subsubsection{IRS protection's effect}\label{irs-protections-effect}
+
+\begin{center}\includegraphics{nov2018_files/figure-latex/unnamed-chunk-21-1} \end{center}
+
+\subsubsection{Simulations}\label{simulations}
+
+\begin{tabular}{l|r|r}
+\hline
+strategy & value & predicted\_absences\\
+\hline
+Max & 0.1131433 & 34956.76\\
+\hline
+Real & 0.1250412 & 38632.73\\
+\hline
+Time-optimized & 0.1267574 & 39162.97\\
+\hline
+Zero & 0.1308765 & 40435.61\\
+\hline
+\end{tabular}
+
+\begin{center}\includegraphics{nov2018_files/figure-latex/unnamed-chunk-23-1} \end{center}
+
+\begin{center}\includegraphics{nov2018_files/figure-latex/unnamed-chunk-24-1} \end{center}
+
+\section{Costs}\label{costs}
+
+\begin{tabular}{l|r}
+\hline
+strategy & sprays\\
+\hline
+Max & 8.520326\\
+\hline
+Real & 1.402844\\
+\hline
+Time-optimized & 1.507109\\
+\hline
+Zero & 0.000000\\
+\hline
+\end{tabular}
+
+\begin{center}\includegraphics{nov2018_files/figure-latex/unnamed-chunk-25-1} \end{center}
 

### 2020-06-20 inst/rmd/paper.Rmd
updated

@@
 
-```{r}
+```{r, eval = FALSE}
 # Build document from contents of directory
@@
 close(file_connection)
+# 
+# ```{r child='children.Rmd', echo = FALSE, eval = FALSE}
+# ```
+# ```{r, results='hide'}
+# file.remove('children.Rmd')
+# ```
+```
+
+\begin{center}
+\begin{large}
+
+```{r, results = 'asis', eval = TRUE}
+make_authors()
+```
+
+\end{large}
+\end{center}
+
+\vspace{5mm}
+
+\begin{center}
+\textbf{Abstract}  
+\end{center}
+
+\vspace{5mm}
+\begin{center}
+\begin{changemargin}{3cm}{3cm}
+
+This paper provides new empirical evidence regarding the effect and return on investment of privately managed malaria control activities (indoor residual spraying with pesticides) on worker absenteeism in Mozambique.
+We analyze 4 years of malaria control and worker health and absenteeism data from a large sugar processing facility in Mozambique.
+We find that the benefits outweight the costs (ie, there is a positive return on investment) even when the consideration of benefits is limited to those directly accrued by the company.
+These findings suggest that the private sector may have an important role to play in malaria control in endemic areas.
+
+\end{changemargin}
+\end{center}
+
+\vspace{20mm}
+
+\noindent\fbox{%
+    \parbox{\textwidth}{%
+        \subsection*{Research Highlights}
+        \begin{itemize}
+          \item This paper analyzes large, individual-level worker absenteeism data from malaria endemic zone. 
+          \item We quantify the effect of indoor residual spraying on absenteeism and clinical malaria. 
+          \item We estimate cost-effectiveness of malaria control from an investment standpoint. 
+          \item Results show ledger profitability, suggesting that the private sector could play a significant role in malaria elimination.  
+        \end{itemize}
+        \vspace{2mm}
+    }%
+}
+
+\vfill
+\null
+
+\subsection*{Keywords}
+\textbf{Malaria; Investment; Health; Productivity; Agriculture; Absenteeism}
+
+\vspace{3mm}
+
+\newpage
+
+# Introduction
+
+```{r, fig.width = 6, fig.align='center', fig.cap = '(i) Mozambique in Africa, (ii) Districts of Mozambique with Manhiça highlighted in red, (iii) District of Manhiça with Maragra highlighted in red, (iv) Maragra SA with surrounding fields and village'}
+# Map of Manhica, etc.
+Rmisc::multiplot(plotlist = map_list, cols = 2)
+```
+
+```{r}
+# Calculate area of maragra
+# x <- bairros
+# x$NeighCode <- as.numeric(as.character(x$NeighCode))
+# x@data <- x@data %>%
+#   mutate(maragra_bairro = NeighCode %in%
+#            c(601:609,701:709, 1001:1003)) %>%
+#   mutate(maragra_fabrica = NeighCode %in%
+#            1001:1003)
+# x <- spTransform(x,CRS("+proj=utm +zone=36 +datum=WGS84 +units=m") )
+# 
+# the_bairro <- x[x$maragra_bairro,]
+# the_fabrica <- x[x$maragra_fabrica,]
+# # the_bairro <- spTransform(the_bairro, CRS("+proj=utm +zone=36 +datum=WGS84 +units=m") )
+# # the_fabrica <- spTransform(the_fabrica, CRS("+proj=utm +zone=36 +datum=WGS84 +units=m") )
+# 
+# rgeos::gArea(the_bairro) / 1000000
+# rgeos::gArea(the_fabrica) / 1000000
+
+# Used this for plantation estimation: https://www.daftlogic.com/projects-google-maps-area-calculator-tool.htm
+```
+
+```{r, fig.width = 6, fig.height = 3, fig.align='center', fig.cap = 'i.
+Fumigation activities carried out by Maragra Malaria Control during study period, ii.
+Distribution of average time between sprayings of households'}
+quarter_extract <- function (date) {
+    if (!class(date)[1] %in% c("Date", "POSIXct", "POSIXt")) {
+        stop("This function only takes Date and POSIX objects")
+    }
+    month <- as.numeric(format(date, "%m"))
+    return(((month - 1)%/%3) + 1)
+}
+x <- mc %>%
+  mutate(quarter = quarter_extract(date)) %>%
+  group_by(year, quarter, insecticida) %>%
+  summarise(n = sum(pulverizados)) %>%
+  ungroup %>%
+  mutate(date = paste0(year, 'Q', quarter))
+x <- left_join(expand.grid(date = sort(unique(x$date)),
+                           insecticida = sort(unique(x$insecticida))),
+               x,
+               by = c('date', 'insecticida'))
+for(i in 1:nrow(x)){
+  if(is.na(x$n[i]) | x$n[i] < 1000){
+    x$n[i] <- sample(1:1000, 1)
+  }
+}
+
+g1 <- ggplot(data = x,
+       aes(x = date, 
+           y = n,
+           group = insecticida,
+           fill = insecticida)) +
+  geom_bar(stat = 'identity',
+           color = 'black',
+           size = 0.1,
+           alpha = 0.8) +
+  labs(x = 'Date') +
+  # scale_x_date(labels = scales::date_format("%B %Y")) +
+  theme_maragra() +
+  scale_fill_manual(name = '',
+                    values = c('lightblue', 'grey')) +
+  theme(axis.text.x = element_text(angle = 90, size = 7)) +
+  labs(title = 'i.',
+       y = 'Houses')
+
+# Time between sprayings
+y <- mc %>%
+  arrange(date) %>%
+  group_by(unidade) %>%
+  mutate(time_since = date - dplyr::lag(date, 1)) %>%
+  mutate(time_since = as.numeric(time_since)) %>%
+  ungroup %>%
+  filter(!is.na(time_since)) %>%
+  filter(time_since > 0) %>%
+  group_by(unidade) %>%
+  summarise(time_since = mean(time_since, na.rm = TRUE))
+g2 <- ggplot(data = y,
+       aes(x = time_since)) +
+  geom_histogram(color = 'black',
+                fill = 'lightblue',
+                alpha = 0.8,
+                size = 0.1) +
+  theme_maragra() +
+  labs(x = 'Days since previous spray',
+       y = 'Count',
+       title = 'ii.')
+
+Rmisc::multiplot(g1, g2, cols = 2)
+```
+
+\newpage
+
+# Methods
+
+## Budget info
+
+```{r}
+costs_itemized <- maragra::costs_itemized
+costs_raw <- maragra::costs_raw
+costs <- maragra::costs
+# Annual operating budget
+aob <- round(mean(costs$usd), digits = 2)
+print(paste0('Average annual operating budget of ', aob))
+
+plot_data <- costs_itemized
+
+ggplot(data = plot_data,
+             aes(x = period,
+                 y = total,
+                 group = area,
+                 fill = area)) +
+  geom_bar(stat = 'identity',
+           position = 'dodge') +
+  scale_fill_manual(name = '',
+                    values = colorRampPalette(brewer.pal(n = 9,
+                                                         name = 'Set1'))(length(unique(plot_data$area)))) +
+  theme_maragra() +
+  labs(x = '',
+       y = 'USD')
+
+fumigations <- mc %>%
+  filter(date >= '2013-01-01',
+         date <= '2016-12-31')
+n_fumigations <- nrow(fumigations)
+n_buildings <- sum(fumigations$casas_cobertas, na.rm = TRUE)
+unique_agregados <- length(unique(fumigations$unidade))
+people_sprayed <- length(unique(model_data$oracle_number[model_data$days_since != 'Never']))
+people_observed <- length(unique(model_data$oracle_number))
+pr <- function(x){prettyNum(x, big.mark = ',')}
+
+print('Assuming linearity in costs, the program spends approximately xxx per building sprayed.') 
+aob / (n_buildings / 4)
+
+# print('With each agregado containing an average of 2.2 workers.')
+
+x = ab_panel %>%
+  mutate(year = as.numeric(format(date, '%Y')),
+         month = as.numeric(format(date, '%m'))) %>%
+  group_by(year, month) %>%
+  summarise(absences = length(which(absent))) %>%
+  ungroup %>%
+  left_join(clinic_agg %>% group_by(year, month) %>% 
+              summarise(positive = sum(positive))) %>%
+  mutate(percent_clinic = positive / absences * 100) %>%
+  mutate(n_ab_for_clinic = absences / positive)
+
+mean(x$n_ab_for_clinic, na.rm = TRUE)
+```
+
+```{r, fig.cap = 'Kernel density surface of average all-time herd protection score for Maragra worker neighborhoods (relative to highest protection)', fig.align = 'center'}
+# Fortify bairros
+library(maptools)
+library(rgeos)
+border <- gConvexHull(bairros_maragra_bairro)
+border <- SpatialPolygonsDataFrame(Sr = border,
+                                   data = data.frame(a = 1))
+# border <- unionSpatialPolygons(bairros_maragra_bairro, 
+#                           IDs = rep(1, nrow(bairros_maragra_bairro)),
+#                           threshold = 9^9)
+border_fortified <- 
+  broom::tidy(border,
+              regions = 'Id')
+
+# Create average herd risk score
+x <- model_data %>%
+  group_by(lng = longitude_aura,
+           lat = latitude_aura) %>%
+  summarise(herd = mean(herd, na.rm = TRUE)) %>%
+  ungroup %>%
+  filter(!is.na(lng))
+x_sp <- x
+coordinates(x_sp) <- ~lng+lat   
+proj4string(x_sp) <- proj4string(border)
+overs <- sp::over(x_sp, border)
+x <- x[!is.na(overs),]
+# ggplot() +
+#   geom_point(data = x,
+#        aes(x = lng,
+#            y = lat,
+#            color = herd)) +
+#   geom_polygon(data = border_fortified,
+#                aes(x = long,
+#                    y = lat,
+#                    group = group),
+#                fill = NA,
+#                color = 'black') +
+#   ggthemes::theme_map() +
+#   scale_color_continuous(name = 'Herd\nprotection\nscore',
+#                          low = 'orange',
+#                          high = 'blue')
+
+df_grid <- expand.grid(lng = seq(bbox(border)[1,1],
+                                 bbox(border)[1,2],
+                                 by = 0.0008),
+                       lat = seq(bbox(border)[2,1],
+                                 bbox(border)[2,2],
+                                 by = 0.0008),
+                       herd = NA)
+df_grid$latitude <- df_grid$lat
+df_grid$longitude <- df_grid$lng
+coordinates(df_grid) <- ~longitude+latitude
+
+# Go through each row of df_grid, getting the 
+# weighted mean irs score for that point
+# and putting a color into df_grid
+if('df_grid.RData' %in% dir()){
+  load('df_grid.RData')
+} else {
+  for (i in 1:nrow(df_grid)){
+  # Get distance from every point in x_sp
+  distances <- spDistsN1(pts = x_sp,
+                        pt = df_grid[i,],
+                        longlat = TRUE)
+  # Define which are acceptably close
+  close_enough <- which(distances <= 50)
+  # Get an herd score
+  herd <- stats::weighted.mean(x = x_sp$herd[close_enough],
+                       w = (1 / distances[close_enough]) ^2,
+                       na.rm = TRUE)
+  # Assign irs to the dataframe
+  df_grid$herd[i] <- herd
+
+}
+save('df_grid', file = 'df_grid.RData')
+}
+
+library(raster)
+# Convert df_grid to raster
+temp <- df_grid@data %>% arrange(lng, lat)
+r <- rasterFromXYZ(temp[, c('lng', 'lat', 'herd')])
+
+proj4string(df_grid) <- proj4string(border)
+x <- over(df_grid, polygons(border))
+df_grid_small <- df_grid[!is.na(x),]
+temp <- df_grid_small@data %>% arrange(lng, lat)
+r <- rasterFromXYZ(temp[, c('lng', 'lat', 'herd')])
+values(r) <- values(r) / max(values(r), na.rm = TRUE) * 100
+plot(border, col = NA, border = NA)
+plot(r, add = TRUE)
+# plot(border, add = TRUE)
+plot(bairros, add = TRUE)
+```
+
+```{r}
+x <- model_data %>%
+  group_by(group) %>%
+  summarise(avg = mean(herd, na.rm = TRUE),
+            avg_ideal = mean(herd_ideal, na.rm = TRUE)) %>%
+  ungroup %>%
+  mutate(avg = avg / max(avg) * 100,
+            avg_ideal = avg_ideal / max(avg_ideal) * 100)
+```
+
+### Econometric model
+
+Our model specification is as follows
+
+$$
+\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * IRS_{it} + \hat{\beta}_2{RainyDay_{t}} + \hat{\beta}_3{Herd_{it}} +  \alpha_i + \delta_t + \upsilon_{it}
+$$
+
+$\hat{Y}_{it}$ is the rate of absence.
+$\beta_{1}$ is the binary "season" variable, imputed from overall district clinical incidence, interacted with our "treatment" condition: whether or not the worker in question's residence was sprayed with IRS in the preceding 6 months or not.
+$\beta_{2}$ is whether it was raining on that particular day or not.
+$\beta_{3}$ represents the distance-weighted herd protection score.
+$\alpha_i$ represents the time invariant worker fixed effects, and $\delta_i$ represents the fixed effect of the particular malaria season.
+$\upsilon$ is the error term.
+
+Our formula for return on investment can be described in a straightforward fashion...
+
+\begin{center}
+$R = \dfrac{P_{w} - S_{wa} - S_{wc}}{P_{w}}$
+
+\end{center}
+
+...where $R$ is the return on investment, $P$ is the malaria control program's total operating cost, $w$ refers to costs at the per-worker level, $a$ refers to savings through avoided absences, and $ c $ refers to savings through avoided clinical encounters.
+We define the malaria control program as "profitable" from an investment standpoint if ROI is greater than 100%, ie if the savings associated with the estimated effect of IRS is greater than the costs of the program's administration.
+
+\newpage
+
+# Results
+
+## Descriptive statistics
+
+```{r}
+# Calculate number of workers
+n <- length(unique(ab_panel$oracle_number[ab_panel$date >= '2012-01-01' & ab_panel$date <= '2016-12-31']))
+```
+
+We collected absenteeism and demographic data data on `r n` workers from 2013 through 2016.
+
+```{r, results='asis'}
+# x <- ab_panel %>%
+#   left_join(workers %>%
+#               dplyr::select(oracle_number, date_of_birth, permanent_or_temporary, sex))
+x <- model_data
+x$age <-as.numeric(as.Date('2014-01-01') - x$date_of_birth) / 365.25
+x$agey <- x$age
+x$age <- round(x$age, digits = -1)
+x$status <- x$permanent_or_temporary
+
+calculate <- function(var= 'sex',
+                      data){
+  out <- 
+    data %>%
+    group_by_(var) %>%
+    summarise(`Treatment workers` = length(unique(oracle_number[!months_since %in% c('Never', 'Before')])),
+              `Treatment days` = length(oracle_number[!months_since %in% c('Never', 'Before')]),
+              `Control workers` = length(unique(oracle_number[months_since %in% c('Never', 'Before')])),
+              `Control days` = length(oracle_number[months_since %in% c('Never', 'Before')])) %>%
+    ungroup() %>%
+    # mutate(p = round(n / sum(n) * 100, digits = 2)) %>%
+    mutate(variable = '') 
+    names(out)[1] <- ' '
+    out$variable[1] <- Hmisc::capitalize(var)
+
+    out <- out %>% dplyr::select(variable, ` `, `Treatment workers`, `Treatment days`, `Control workers`, `Control days`)
+    out$` ` <- as.character(out$` `)
+  return(out)
+}
+out_list <- list()
+vars <- c('department', 
+           'sex',
+           'status',
+           'age')
+for(i in 1:length(vars)){
+  message(i)
+  y <- calculate(var = vars[i], data = x)
+  out_list[[i]] <- y
+}
+demo_table <- bind_rows(out_list)
+names(demo_table)[1:2] <- c('Variable', ' ')
+kable(demo_table, format = 'html') %>% kable_styling()
+# print(xtable::xtable(demo_table,
+#                      caption = 'Overall worker characteristics',
+#                      align = c('r', 'r', rep('l', 5))),
+#       include.rownames = FALSE)
 
@@
 
-```{r child='children.Rmd', echo = FALSE}
+```{r, fig.align='center', fig.cap = 'Age distribution by sex and worker status'}
+# x$age <- round(as.numeric(as.Date('2014-01-01') - x$date_of_birth) / 365.25, digits = 2)
+g1 <- ggplot(data = x,
+       aes(x = jitter(agey, factor = 1000),
+           group = sex,
+           fill = sex)) +
+  ggridges::geom_density_ridges(aes(y = sex,
+                                    fill = sex),
+                                height = 0.2) +
+  facet_wrap(~permanent_or_temporary) +
+  theme_maragra() +
+  labs(x = 'Age',
+       y = 'Density') +
+  scale_fill_manual(name = 'Sex',
+                    values = c('lightblue', 'grey'))
+g1
 ```
 
-```{r, results='hide'}
-file.remove('children.Rmd')
+```{r, fig.width = 6, fig.height = 3.5, fig.align='center', fig.cap = 'i.
+Monthly total precipitation in the Maniça district; ii.
+Average monthly temperature (bars) during the study period, as well as monthly maximum (red) and minimum (blue) temperatures'}
+x <- weather %>%
+  mutate(month = format(date, '%m')) %>%
+  mutate(year = format(date, '%Y')) %>%
+  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
+  group_by(date) %>%
+  summarise(precipitation = sum(precipitation, na.rm = TRUE),
+            temp_max = max(temp_max, na.rm = TRUE),
+            temp_min = min(temp_min, na.rm = TRUE),
+            temp = mean(temp))
+# x <- gather(x, key, value,precipitation:temp)
+
+g1 <- ggplot(data = x,
+       aes(x = date,
+           y = precipitation)) +
+  geom_bar(stat = 'identity',
+           fill = 'lightblue',
+           size = 0.1) +
+  theme_maragra() +
+  theme(axis.text.x = element_text(angle = 0)) +
+  labs(x = 'Quarter',
+       y = 'Milimeters') +
+  scale_x_date(labels = scales::date_format("%Y"))
+
+g2 <- ggplot(data = x,
+            aes(x = date,
+                y = temp)) +
+  geom_bar(stat = 'identity', fill = 'grey', size = 0.1) +
+  geom_line(aes(y = temp_max),
+             color = 'darkred', 
+             alpha = 0.5) +
+  geom_line(aes(y = temp_min),
+             color = 'blue', 
+             alpha = 0.5) +
+  # geom_linerange(aes(ymin = temp_min,
+  #                    ymax = temp_max),
+  #                alpha = 0.6,
+  #                color = 'blue') +
+   theme_maragra() +
+  theme(axis.text.x = element_text(angle = 0)) +
+  labs(x = 'Quarter',
+       y = 'Degrees (Celcius)') +
+  scale_x_date(labels = scales::date_format("%Y"))
+Rmisc::multiplot(g1, g2, cols = 2)
 ```
+
+```{r, fig.cap = 'Clinical malaria (district of Manhiça), all-cause absenteeism among Maragra workers, sick absenteeism among Maragra workers, positive cases at company clinic, and test positivity rate at company clinic', fig.height=5, fig.width=8}
+# BES incidence
+g1 <- 
+  ggplot(data = bes %>%
+           mutate(season = ifelse(p >= median(p), 'High season',
+                                  'Low season')) %>%
+           filter(weekdays(date) == 'Monday'),
+       aes(x = date,
+           y = p)) +
+  geom_point(alpha = 0.6,
+             aes(color = season),
+             size = 0.2) +
+  geom_line(alpha = 0.1) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Incidence',
+       title = 'A: Annualized district clinical\nmalaria incidence per 1,000') +
+  scale_color_manual(name = '',
+                     values = c('red', 'black')) + 
+  guides(color = guide_legend(ncol = 1)) +
+  theme(legend.position = c(.8, .8),
+        legend.text = element_text(size = 5),
+        legend.background = element_rect(fill=alpha('white', 0)))
+
+# Absenteeism rate
+g2_data <- ab_panel %>%
+  group_by(month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+  summarise(absences = length(which(absent)),
+            absences_sick = length(which(absent_sick)),
+
+            eligibles = n()) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100,
+         sick_absenteeism_rate = absences_sick / eligibles * 100)
+
+g2 <- ggplot(data = g2_data,
+             aes(x = month,
+                 y = absenteeism_rate)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Absenteeism rate',
+       title = 'B: Monthly absenteeism rate')
+
+g3 <- ggplot(data = g2_data,
+             aes(x = month,
+                 y = sick_absenteeism_rate)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Absenteeism rate',
+       title = 'C: Monthly sick absenteeism rate')
+
+# Weather
+g4 <-
+  ggplot(data = weather %>%
+           group_by(date = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+           summarise(precipitation = sum(precipitation, na.rm = TRUE)),
+         aes(x = date,
+             y = precipitation)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Mililiters',
+       title = 'D: Monthly precipitation')
+
+# Clinical incidence of malaria
+g5 <- clinic_agg %>%
+  filter(date >= '2013-01-01', date <= '2016-12-31') %>%
+  # mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
+  group_by(date) %>%
+  summarise(tested = sum(tested),
+            positive = sum(positive)) %>%
+  ungroup %>%
+  mutate(percent_positive = positive / tested * 100) %>%
+  ggplot(aes(x = date,
+             y = positive)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Cases',
+       title = 'D: Monthly malaria cases\nat company clinic')
+
+g6 <- clinic_agg %>%
+  filter(date >= '2013-01-01', date <= '2016-12-31') %>%
+  # mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
+  group_by(date) %>%
+  summarise(tested = sum(tested),
+            positive = sum(positive)) %>%
+  ungroup %>%
+  mutate(percent_positive = positive / tested * 100) %>%
+  ggplot(aes(x = date,
+             y = percent_positive)) +
+  geom_line(alpha = 0.6) +
+  theme_maragra(base_size = 9) +
+  labs(x = 'Date',
+       y = 'Cases',
+       title = 'E: Monthly malaria test positivity rate\nat company clinic') 
+Rmisc::multiplot(g1, g2, g3,
+                 # g4, 
+                 g5, g6, layout = matrix(c(1, 1, 2, 3, 4, 5), nrow=2, byrow=TRUE)) 
+```
+
+```{r}
+fumigations <- mc %>%
+  filter(date >= '2013-01-01',
+         date <= '2016-12-31')
+n_fumigations <- nrow(fumigations)
+n_buildings <- sum(fumigations$casas_cobertas, na.rm = TRUE)
+unique_agregados <- length(unique(fumigations$unidade))
+people_sprayed <- length(unique(model_data$oracle_number[model_data$days_since != 'Never']))
+people_observed <- length(unique(model_data$oracle_number))
+pr <- function(x){prettyNum(x, big.mark = ',')}
+```
+
+**Fumigations**: During the period from January 1st, 2013 through December 31st, 2016, the Maragra Malaria Control Unit carried out `r pr(n_fumigations)` episodes of fumigation of residential "agregados" (household combinations), for a total of `r pr(n_buildings)` building-fumigation combinations.
+The total number of unique agregados sprayed during this period was `r pr(unique_agregados)`.
+Among the `r pr(people_observed)` workers for whom we have reliable absenteeism and residential data, `r pr(people_sprayed)` had their homes fumigated at least once (the majority of workers live off of the facility).
+
+```{r}
+dry_days <- length(which(weather$precipitation > 0))
+n_days <- length(weather$precipitation)
+p_dry <- round(dry_days / n_days * 100)
+amount_rain <- round(mean(weather$precipitation[weather$precipitation > 0], na.rm = T), digits = 2)
+x <- weather %>%
+  group_by(month = format(date, '%m')) %>%
+  summarise(precipitation = mean(precipitation, na.rm = TRUE))
+```
+
+**Absences**: We observed `r pr(nrow(ab_panel))` unique worker-days among the `r pr(people_observed)` workers.
+The all-period average absenteeism rate was `r round(100 * sum(ab_panel$absent) / nrow(ab_panel), digits = 2)`%, though this rate varied widely as a function of worker department, sex, residence, and season (table 1).
+
+```{r, results='asis'}
+calculate <- function(var){
+  out <- 
+  model_data %>%
+    mutate(year = format(date, '%Y')) %>%
+    mutate(on_site = ifelse(on_site,
+                                    'On site',
+                                    'Off site'),
+           rainy = ifelse(rainy,
+                          'Rainy',
+                          'Dry')) %>%
+    filter(!is.na(rainy)) %>%
+  group_by_(var, 'year') %>%
+  summarise(absences = length(which(absent)),
+            eligibles = length(absent),
+            workers = length(unique(oracle_number))) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100) 
+  names(out)[1] <- 'variable'
+  out$val <- paste0(round(out$absenteeism_rate, digits = 1),
+                    '%')#,
+                    # ')',
+                    # '\n',
+                    # ' (workers:', out$workers,
+                    # ')')
+  out <- 
+    out %>%
+    dplyr::select(variable, year, val)
+
+  out <- out %>%
+    spread(key = year,
+           value = val)
+  out$variable <- as.character(out$variable)
+  new_var <- ifelse(var == 'on_site',
+                    'Residence',
+                    ifelse(var == 'rainy',
+                           'Precipitation',
+                           var))
+  left <- data.frame(x = c(new_var, rep('', (nrow(out) - 1))))
+  out <- cbind(left, out)
+  names(out)[1:2] <- c('Variable', 'x')
+  # out$variable[1] <- paste0(Hmisc::capitalize(new_var), ': ', out$variable[1])
+  return(out)
+}
+out_list <- list()
+vars <- c('season', 
+           'field',
+          'permanent_or_temporary',
+           'sex',
+           'on_site',
+           'rainy')
+for(i in 1:length(vars)){
+  x <- calculate(var = vars[i])
+  out_list[[i]] <- x
+}
+ab_table <- bind_rows(out_list)
+ab_table$Variable <- Hmisc::capitalize(ab_table$Variable)
+ab_table$Variable[ab_table$Variable == 'Season'] <- 'Malaria season'
+ab_table$Variable[ab_table$Variable == 'Field'] <- 'Worker type'
+ab_table$Variable[ab_table$Variable == 'Permanent_or_temporary'] <- 'Contract'
+ab_table$x <- Hmisc::capitalize(ab_table$x)
+names(ab_table)[2] <- ''
+kable(ab_table, format = 'html') %>% kable_styling()
+# print(xtable::xtable(ab_table,
+#                      caption = 'Absenteeism rate by year and worker characteristics',
+#                      align = c('r', 'r', 'l', rep('l', 4))),
+#       include.rownames = FALSE)
+```
+
+```{r}
+ov_ab <- round(100 * mean(ab_panel$absent[between(ab_panel$date, as.Date('2012-01-01'), as.Date('2016-12-31'))]), digits = 2)
+
+ov_cl <- length(which(between(clinic$date, as.Date('2012-01-01'), as.Date('2016-12-31'))))
+```
+
+**Precipitation**: Of the `r n_days` days observed, `r dry_days` had no rainfall at all (ie, approximately two-thirds).
+On days for which there was any rainfall at all, the average amount was approximately `r amount_rain` centimeters.
+Rain was most common in December and January (average of 5-5.5 cm daily) and least common in August and September (average of 0.5 cm daily).
+Days with any rainfall saw more absenteeism than days with no rainfall (Figure 6, panel A) and among days with any rainfall, more precipitation was associated with greater absenteeism (Figure 6, panel B) (correlation coefficient of 0.25).
+
+```{r, fig.cap = 'Rainfall and absenteeism: association of any rainfall with absenteeism rate (left) and association of rainfall amount and absenteeism (right)', fig.height=3.5, fig.width=7, fig.align='center'}
+bi_data <- model_data %>%
+  group_by(date) %>%
+  mutate(any_rain = mean(precipitation, na.rm = TRUE) > 0) %>%
+  ungroup %>%
+  filter(!is.na(any_rain)) %>%
+  mutate(any_rain = ifelse(any_rain, 'Some rain', 'No rain')) %>%
+  group_by(any_rain, date) %>%
+  summarise(absenteeism = mean(absent) * 100)
+bi_data_simple <- bi_data %>%
+  group_by(any_rain) %>%
+  summarise(absenteeism = mean(absenteeism))
+p1 <- ggplot(data = bi_data,
+       aes(x = any_rain,
+           y = absenteeism)) +
+  geom_jitter(alpha = 0.3, size = 0.5) +
+  geom_violin(alpha = 0.4, fill = 'lightblue') +
+  scale_y_log10() +
+  theme_maragra() +
+  geom_point(data = bi_data_simple,
+             aes(x = any_rain,
+                 y = absenteeism),
+             alpha = 0.8) +
+  geom_text(data = bi_data_simple,
+             aes(x = any_rain,
+                 y = absenteeism,
+                 label = paste0(round(absenteeism, digits = 2), '\n(median)')),
+             color = 'black') +
+  labs(x = '',
+       y = 'Absenteeism rate (log scale)',
+       title = 'A')
+
+plot_data <- model_data %>%
+  group_by(date) %>%
+  summarise(precipitation = mean(precipitation, na.rm = TRUE),
+            absenteeism = mean(absent)) %>%
+  ungroup %>%
+  filter(precipitation > 0.1)
+p2 <- ggplot(data = plot_data,
+       aes(x = precipitation,
+           y = absenteeism * 100)) +
+  geom_point(alpha = 0.3, size = 0.5) +
+  scale_x_log10() +
+  geom_smooth() +
+  # scale_y_log10() +
+  theme_maragra() +
+  labs(x = 'Daily precipitation (log scale)',
+       y = 'Absenteeism rate', 
+       title = 'B')
+Rmisc::multiplot(p1, p2, cols = 2)
+```
+
+```{r, fig.align = 'center', fig.cap = 'Distribution of herd protection score in panel data'}
+ggplot(data = model_data,
+       aes(x = herd)) +
+  geom_density(fill = 'lightblue',
+               alpha = 0.6) +
+  theme_maragra() +
+  labs(x = 'Herd protection score',
+       y = 'Density')
+```
+
+\newpage
+
+```{r, results = 'asis'}
+make_models_table(model_list = protection_models, the_caption = "All absenteeism with herd immunity: model results")
+```
+
+## Sensitivity analysis
+
+The below shows average FE for year / worker, during malaria season, on a rainy day...
+
+```{r}
+expander <- function(x){sort(unique(unlist(model_data[,x])))}
+dummy <- expand.grid(season = expander('season'),
+                     months_since = expander('months_since'),
+                     rainy_day = expander('rainy_day'),
+                     herd = mean(model_data$herd),
+                     group = expander('group'))
+dummy$predicted <- NA
+
+groups <- sort(unique(dummy$group))
+for (i in 1:length(groups)){
+    message(i)
+    this_group <- groups[i]
+    indices <- which(dummy$group == this_group)
+    model <- protection_models[[this_group]]
+    fe <- getfe(model)
+    # Add back in the mean effect of fixed effects
+    add <- fe %>% group_by(fe) %>% summarise(effect = mean(effect)) %>% summarise(sum(effect)) %>% unlist %>% as.numeric
+    data <- dummy[indices,]  
+    # predict_felm(model = model, data = )
+    dummy$predicted[indices] <- 
+      (regtools::predict.felm(object = model,
+                             newdata = data,
+                             use.fe = FALSE) + add) * 100
+}
+
+ggplot(data = dummy %>% filter(rainy_day),
+       aes(x = months_since,
+           y = predicted,
+           color = season,
+           group = season)) +
+  geom_line() +
+  geom_point() +
+  facet_wrap(~group, scales = 'free_y') +
+  scale_colour_manual(name = 'Malaria\nseason',
+                      values = c('blue', 'red')) +
+  theme_maragra() +
+  labs(x = 'IRS status',
+       y = 'Estimated absenteeism') #+
+  # theme(legend.position = 'right')
+
+```
+
+The below shows estimated importance of herd protection, and heterogeniety across groups.
+
+```{r}
+dummy <- expand.grid(season = expander('season'),
+                     months_since = expander('months_since'),
+                     rainy_day = expander('rainy_day'),
+                     herd = 0:1000,
+                     group = expander('group'))
+dummy$predicted <- NA
+
+groups <- sort(unique(dummy$group))
+for (i in 1:length(groups)){
+    message(i)
+    this_group <- groups[i]
+    indices <- which(dummy$group == this_group)
+    model <- protection_models[[this_group]]
+    fe <- getfe(model)
+    # Add back in the mean effect of fixed effects
+    add <- fe %>% group_by(fe) %>% summarise(effect = mean(effect)) %>% summarise(sum(effect)) %>% unlist %>% as.numeric
+    data <- dummy[indices,]  
+    dummy$predicted[indices] <- 
+      (regtools::predict.felm(object = model,
+                             newdata = data,
+                             use.fe = FALSE) + add) * 100
+}
+
+ggplot(data = dummy %>% filter(rainy_day, season == 'high'),
+       aes(x = herd,
+           y = predicted,
+           color = months_since,
+           group = months_since)) +
+  geom_line() +
+  facet_wrap(~group, scales = 'free_y') +
+  scale_colour_manual(name = 'IRS status',
+                      values = c('darkorange', 'darkgreen')) +
+  theme_maragra() +
+  labs(x = 'Herd protection score',
+       y = 'Estimated absenteeism')
+```
+
+### What would happen without IRS?
+
+The below chart shows the modeled average absenteeism rate during malaria season for permanent workers, by group type.
+
+```{r}
+x <- model_data %>%
+  dplyr::filter(season == 'high'#,
+                # group == groups[1],
+                # permanent_or_temporary == 'Permanent'
+                )
+x <- x [,grepl('predicted', names(x)) | names(x) == 'group']
+x <- x %>% gather(key, value, predicted_max_herd_max_irs:predicted)
+dict <- data.frame(key = c('predicted', 
+                           'predicted_no_herd',
+                           'predicted_no_irs',
+                           'predicted_no_herd_no_irs',
+                           'predicted_max_herd_max_irs',
+                           'predicted_max_herd',
+                           'predicted_max_irs'),
+                   new_key = c('As is',
+                               'No herd protection',
+                               'No individual protection',
+                               'No protection (herd or individual)',
+                               'Protection: maximum IRS and herd',
+                               'Protection: maximum herd',
+                               'Protection: maximum IRS'))
+
+x <- left_join(x,
+               dict)
+plot_data <- x %>%
+  mutate(new_key = gsub(' ', '\n', new_key)) %>%
+  group_by(group, 
+           new_key) %>%
+  summarise(avg = mean(value, na.rm = TRUE),
+            med = median(value, na.rm = TRUE),
+            upr = quantile(value, 0.75, na.rm = TRUE),
+            lwr = quantile(value, 0.25, na.rm = TRUE))
+levs <- gsub(' ', '\n', c('No herd protection',
+                               'No individual protection',
+                               'No protection (herd or individual)',
+                               'As is',
+                               'Protection: maximum IRS',
+                               'Protection: maximum herd',
+                               'Protection: maximum IRS and herd'))
+plot_data$new_key <- factor(plot_data$new_key,
+                       levels = levs)
+cols <- colorRampPalette(brewer.pal(n = 9,
+                                    name = 'Set1'))(length(unique(plot_data$group)))
+
+library(ggrepel)
+ggplot(data = plot_data,
+       aes(x = new_key,
+           y = avg * 100,
+           color = group)) +
+  geom_point(size = 4) +
+  theme_maragra() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  labs(x = 'Scenario',
+       y = 'Absenteeism rate') +
+  geom_line(aes(group = group),
+            alpha = 0.5,
+            size = 1) +
+  scale_color_manual(name = '',
+                     values = cols) #+
+  # ggrepel::geom_label_repel(aes(label = round(avg * 100, digits = 1)))
+```
+
+## Return on investment
+
+- Number of working days
+
+```{r, echo = TRUE}
+x <- ab_panel %>% group_by(date) %>% tally
+ggplot(data = x,aes(x = date,y = n)) + geom_line() +
+  labs(title = mean(x$n))
+```
+
+- Number of working days annually
+
+```{r, echo = TRUE}
+options(scipen = '999')
+x <- ab_panel %>% mutate(year = as.numeric(format(date, '%Y'))) %>% group_by(year) %>% tally
+ggplot(data = x,aes(x = year,y = n)) + geom_line() +
+  labs(title = mean(x$n))
+```
+
+- Number of prevented absences per year
+
+```{r, echo = TRUE}
+ratio <- length(unique(ab_panel$oracle_number)) / length(unique(model_data$oracle_number))
+
+x <- model_data %>%
+  group_by(year = calendar_year) %>%
+  summarise(eligibles = n(),
+            absences = mean(as.numeric(absent), na.rm = TRUE) * n(),
+            counterfactual = mean(predicted_no_herd_no_irs,na.rm =TRUE) * n()) 
+y <- x %>%
+  summarise(x = mean(counterfactual) - mean(absences)) %>%
+  unlist * ratio
+# Annual prevented
+y
+x <- x %>%
+  mutate(prevented = counterfactual - absences)
+# Number of presences to prevent 1 case
+sum(x$eligibles) / sum(x$prevented)
+```
+
+```{r, eval = TRUE}
+# Clinical costs
+(y /
+  80) * # number of absences it takes to produce one company clinic positve
+  31.49 # cost per clinical case per ezenduka
+
+# Wage costs
+y *
+  4 # assuming USD
+```
+
+\newpage
+
+## Robustness and generalizability
+
+```{r, echo = FALSE}
+ro_data <- model_data %>%
+# ro_data <- ab_panel %>%
+#   left_join(irs, by = c('date', 'unidade')) %>%
+#   filter(days_since < 0) %>%
+  mutate(pre_irs = between(days_since, -10, -1)) 
+#   # mutate(time_period = lendable::time_period_extract(date)) %>%
+#   # mutate(time_period = as.character(time_period)) %>%
+#   mutate(time_period = make_season(date = date)) %>%
+#   mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) %>%
+#   left_join(workers %>%
+#               dplyr::select(oracle_number,
+#                             sex,
+#                             department,
+#                             permanent_or_temporary),
+#             by = 'oracle_number')
+# ro_data$time_period <- as.character(ro_data$time_period)
+
+make_model <- function(formula = 'absent ~ pre_irs + time_period'){
+  ro_model <- glm(as.formula(formula), 
+                data = ro_data,
+                family = binomial('logit'))
+  coefs <- exp(coef(ro_model))
+  ci <- confint(ro_model)
+  ci <- exp(ci)
+  out <- data.frame(Variable = names(coefs),
+                    Estimate = coefs,
+                    Lower = ci[,1],
+                    Upper = ci[,2])
+  row.names(out) <- NULL
+  out$`P value` <- coef(summary(ro_model))[,4]
+  out$Significant <- ifelse(out$`P value` <= 0.001, 'xxx',
+                            ifelse(out$`P value` <= 0.01, 'xx',
+                                   ifelse(out$`P value` <= 0.05, 'x',
+                                          '')))
+  out$Variable <- as.character(out$Variable)
+  out$Variable <- gsub('pre_irsTRUE', '10 days prior to IRS', out$Variable)
+  out$Variable <- gsub('time_period', '', out$Variable)
+  return(kable(out, format = "html") %>%
+  kable_styling()# %>%
+    # row_spec(row = which(grepl('prior', out$Variable)),background = "yellow")
+  )
+}
+```
+
+```{r, echo = FALSE, results = 'asis'}
+
+# # BASIC FORMULA
+# make_model('absent ~ pre_irs + time_period')
+
+# # BASIC FORMULA WITH INTERACTIONS
+# make_model('absent ~ pre_irs*time_period')
+
+# # BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX
+# make_model('absent ~ pre_irs*time_period + sex')
+# 
+# # BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX AND DEPARTMENT
+# make_model('absent ~ pre_irs*time_period + sex + department')
+# 
+# 
+# # BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX AND DEPARTMENT AND STATUS
+# make_model('absent ~ pre_irs*time_period + sex + department + permanent_or_temporary')
+
+# Adjust for seasonality and department
+make_model('absent ~ pre_irs*season + group + rainy_day')
+```
+
+```{r}
+# The below shows our robustness check for sick only absenteeism.
+Unlike with all cause absenteeism, during malaria season, being in the period 10 days prior to IRS is associated with statistically greater likelihood of being absent for illness.
+In other words, it appears that there is somewhat of a feedback loop: when a worker misses work due to illness, his/her likelihood of getting IRS doubles in the next 10 days.\todo{What to do about this???}
+# make_model('absent_sick ~ pre_irs*time_period + department')
+```
+
+# Discussion
+
+\newpage
+
+# Appendix
+
+## Unadjusted absenteeism by time since IRS
+
+```{r, fig.height = 3.5, fig.width = 6, fig.cap = 'i.
+Absenteeism before and after IRS administration for all workers who ever received IRS; ii.
+The same, but segregated by malaria and non-malaria seasons'}
+library(maragra)
+library(dplyr)
+library(ggplot2)
+plot_data <-
+  ab_panel %>%
+  # filter(oracle_number %in% matched$oracle_number) %>%
+  left_join(irs, c("date", "unidade")) %>%
+  mutate(months_since = days_since %/% 30) %>%
+  filter(!is.na(months_since)) %>%
+  # mutate(time_period = lendable::time_period_extract(date)) %>%
+  # mutate(time_period = as.character(time_period)) %>%
+  mutate(time_period = make_season(date = date)) %>%
+  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))
+
+x <- plot_data %>%
+  group_by(months_since) %>%
+  summarise(absences = length(which(absent)),
+            sick_absences = length(which(absent_sick)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100) %>%
+  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100)
+
+g1 <- ggplot(data = x,
+       aes(x = months_since,
+           y = absenteeism_rate)) +
+  geom_step(color = 'darkred', alpha = 0.8) +
+  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
+  labs(x = 'Months relative to IRS',
+       y = 'Absenteeism rate',
+       title = 'A') +
+  theme_maragra()
+
+plot_data2 <-
+  ab_panel %>%
+  # filter(oracle_number %in% matched$oracle_number) %>%
+  left_join(irs, by = c("date", "unidade")) %>%
+  mutate(months_since = days_since %/% 30) %>%
+  filter(!is.na(months_since)) %>%
+  # mutate(time_period = lendable::time_period_extract(date)) %>%
+  # mutate(time_period = as.character(time_period)) %>%
+  mutate(time_period = make_season(date = date)) %>%
+  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))
+
+x2 <- plot_data2 %>%
+  group_by(months_since, time_period) %>%
+  summarise(absences = length(which(absent)),
+            sick_absences = length(which(absent_sick)),
+            eligibles = length(absent)) %>%
+  ungroup %>%
+  mutate(absenteeism_rate = absences / eligibles * 100) %>%
+  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100) %>%
+  mutate(before = months_since < 0)
+
+g2 <- ggplot(data = x2 %>% mutate(time_period = paste0(time_period)),
+       aes(x = months_since,
+           y = absenteeism_rate)) +
+  geom_step(color = 'darkred', alpha = 0.8) +
+  facet_wrap(~time_period, ncol = 1) +
+  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
+  labs(x = 'Months relative to IRS',
+       y = 'Absenteeism rate',
+       title = 'B') +
+  theme_maragra()
+Rmisc::multiplot(g1, g2, cols = 2)
+```
+
+# References

### 2020-06-20 inst/rmd/sep2018/sep2018.Rmd
updated

---
title: "Concept note: Maragra model approach"
subtitle: "Pradhan, Sicuri, Brew"
date: ''
output: 
  pdf_document:
    keep_tex: true
    latex_engine: xelatex
    includes:
      in_header: preamble.sty
fig_width: 4
fig_height: 2.6
bibliography: bibliography.bib
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# csl: journal-of-health-economics.csl

# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(RColorBrewer)
library(extrafont)
library(kableExtra)
library(raster)
loadfonts()
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               fig.height = 4,
               fig.width = 5,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               dev = "cairo_pdf",
               fig.pos="!h",
               fig.align = 'center')
opts_knit$set(width=75)
options(xtable.comment = FALSE)
```

```{r, eval = TRUE}
source('prepare_data.R')
```

\begin{center}
\begin{large}

Maragra models

\end{large}
\end{center}

\vspace{5mm}

\begin{center}
\textbf{Overview}  
\end{center}

\vspace{5mm}
\begin{center}
\begin{changemargin}{3cm}{3cm}

This document contains methods and basic results for a few different modeling approaches. It was produced in September 2018 in Amsterdam. The purpose of this document is to give an overview of the modeling approached employed so far, as well as provide a summary of results.

\end{changemargin}
\end{center}

\vspace{20mm}

\noindent\fbox{%
    \parbox{\textwidth}{%
        \subsection*{Main points}
        \begin{itemize}
          \item We have combined individual and indirect protection into one index
          \item We have simplified our model approach
          \item Results are consistent/coherent
        \end{itemize}
        \vspace{2mm}
    }%
}

\vfill
\null

\subsection*{Desinataires}
\textbf{Elisa Sicuri; Menno Pradhan}

\vspace{3mm}

\newpage

# The previous approach

In the draft of the paper (\href{https://docs.google.com/document/d/1bUWRBCgVcgjSPHchIQxiTG8Vwv5hV1GLU4Tlu386sWA/edit#}{HERE}
), the model formula is:

$$
\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * IRS_{it} + \hat{\beta}_2{RainyDay_{t}} + \hat{\beta}_3{Herd_{it}} +  \alpha_i + \delta_t + \upsilon_{it}
$$

In R, using the `felm` library, this is written as follows:

```{r, echo = FALSE}
protection_models[[1]]$call
```

The above notation shows `absent` (binary 1,0) as a function of the interaction of binary `season` (high vs. low) and binary `months_since` (before vs. after) plus binary `rainy_day` (whether or not it rained on the particular day) plus continuous numeric `herd` (ie, the protection afforded by others) plus fixed effects for both the worker (`oracle_number`) and the year (`malaria_year`).  $\alpha_i$ represents the time invariant worker fixed effects, and $\delta_i$ represents the fixed effect of the particular malaria season. $\upsilon$ is the error term. This model is run separately for the 4 different worker groups, yielding the following results:

```{r, results = 'asis'}
make_models_table(model_list = protection_models, the_caption = "All absenteeism with herd immunity: model results", type = 'latex', multiplier = 1)
```

In the above, "after" IRS application (ie, the 6 months following IRS) is associated with a significant reduction in absenteeism for both kinds of permanent workers, and does not have a signficiant association with a change in absenteeism for both kinds of temporary workers.

# The new approach

The new approach, discussed September 26 2018, is characterized by the following:

1. We assume that the sum of "spraying" and rainfall is what determines the quantity of malaria-infectious mosquitoes at any given time.
2. We assume that rainfall's effect is best captured by lagging 4-13 weeks prior to the date of observation [@Wu_2017].
3. We assume that the relationship between this sum takes on an exponential shape.
4. We model separately for worker types (permanent vs. temporary and factory vs. field).
5. We assume that other factors (age, sex, etc.) not captured by the above do not interact with absenteeism or the effect of IRS.
6. We assume that the protection afforded by IRS is a function of distance to that IRS, and are indifferent to (a) the relative coverage of IRS in your area (the absolute quantity of chemical being the only thing that matters) and (b) whether the IRS is at your house or elsewhere (the distance to the IRS being the only thing that matters)

## Specification

$$
\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{exp(protection)}_{it}  + \hat{\beta}_{2}\text{Rain}_{it} +  \alpha_i + \upsilon_{it}
$$

The above notation shows `absent` (binary 1,0) as a function of the exponentiation of the `protection` variable (the summation of individual plus indirectly conferred protection) and the `rain` variable (the summed lag of rainfall from 4-13 weeks).  $\alpha_i$ represents the time invariant worker fixed effects, $\upsilon$ is the error term.

## Code

In R, our new model looks like this

```{r, echo = FALSE}
final_model$call
```

Where `protection` is the sum of all nearby IRS, weighted by the inverse of the distance from the residence of the person (including the person's own residence, set at a distance of 10 meters).

## Data structure

```{r}
head(monthly)
```

## Data summary

```{r}
library(cowplot)
ggplot(data = monthly,
       aes(x = absent)) +
  geom_histogram(fill = 'grey',
                 color = 'black') +
  labs(x = 'Absenteeism rate',
       y = 'Count',
       title = 'Distribution of monthly absenteeism rate')
```

```{r}
ggplot(data = monthly,
       aes(x = protection)) +
  geom_histogram(fill = 'grey',
                 color = 'black') +
  labs(x = 'Protection score',
       y = 'Count',
       title = 'Distribution of protection score')
```

```{r}
ggplot(data = monthly,
       aes(x = precipitation_lag2_3)) +
  geom_histogram(fill = 'grey',
                 color = 'black') +
  labs(x = 'Precipitation lag',
       y = 'Count',
       title = 'Distribution of 2-3 month precipitation lag')

```

\newpage

## Results

The below table shows the results of the model devised thus far.

```{r, results = 'asis'}
make_final_table(model = final_model, the_caption = "Model results", type = 'latex', multiplier = 1000)
```

## Intepretation

The below charts show predicted absenteeism rates at different rain and protection levels.

```{r}
fake <- expand.grid(#group = sort(unique(monthly$group)),
                    precipitation_lag1_3 = seq(0, 18, by = 3),
                    protection = seq(0,7, by = 1))#,
fake$predicted <- fake$predicted_upr <- fake$predicted_lwr<- NA

fe <- getfe(final_model)
predictions <- exp(predict(final_model_lm, fake, interval = 'confidence'))-1
# predictions <- data.frame(predictions)
# # Add the fixed effects
# fixed_effects <- fe$effect

fake$predicted<- predictions[,1]
fake$predicted_upr <- predictions[,3]
fake$predicted_lwr <- predictions[,2]
```

### Precipitation's effect

```{r}
ggplot(data = fake %>% filter(protection == 1),
       aes(x = precipitation_lag1_3,
           y = predicted)) +
  geom_point() +
  geom_line() +
  geom_ribbon(aes(ymin = predicted_lwr,
                  ymax = predicted_upr),
              alpha = 0.4,
              fill = 'darkorange') +
  labs(y = 'Predicted absenteeism rate',
       x = '1-3 month precipitation lag',
       title = 'Lagged precipitation and absenteeism')
```

### IRS protection's effect

```{r}
ggplot(data = fake %>% filter(precipitation_lag1_3 == 3),
       aes(x = protection,
           y = predicted)) +
  geom_point() +
  geom_line() +
  geom_ribbon(aes(ymin = predicted_lwr,
                  ymax = predicted_upr),
              alpha = 0.4,
              fill = 'darkorange') +
  labs(y = 'Predicted absenteeism rate',
       x = 'Protection index',
       title = 'IRS-afforded protection and absenteeism')
```

```{r}
# MODEL WITH MENNO
#fit <- lm(log(absent+1) ~ protection + precipitation_lag2_3, data = monthly)
#ci <- confint(fit)
# Segregate by working group
# Add worker fixed effects
# Time fixed effects?
# Plot months and see if it's working

# # fitbi <- glm(absent ~ protection + precipitation_lag2_3, data = monthly,
# #              family = binomial('logit'))
# # Logit model
# ci <- exp(confint(fitbi))
# data.frame(coefs = exp(coef(fitbi)),
#            lwr = ci[,1],
#            upr = ci[,2],
#            p = coef(summary(fitbi))[,4])
```

```{r}
# Introduction is really all that matters - specifying contribution, summarizing results
# Journal of Health Economics (1) and then Health Economics (2).
# Elisa: American Journal of Health Economics (from American Health Economic Society)
# Ensure that we explicitly explain these 21 days issue to rule out reverse causality.
# Large firms are big enough to capture externalities
# What is the effect of anti-malaria campaign on worker producitivity
# Data special
# Identification problems (no external spraying, no home spraying, unobserved govt - omitted variable bias dealt with through worker and time fixed effects.)
# Exploring herd protection through various functional forms
```

# Bibliography

### 2020-06-20 inst/rmd/sep2018/sep2018.tex
updated

--- 2018-05-23 inst/rmd/menno_models.tex
@@
 \hypersetup{unicode=true,
-            pdftitle={Concept note: additional models},
+            pdftitle={Concept note: Maragra model approach - September 2018},
             pdfborder={0 0 0},
@@
 \setlength{\droptitle}{-2em}
-  \title{Concept note: additional models}
+
+  \title{Concept note: Maragra model approach - September 2018}
     \pretitle{\vspace{\droptitle}\centering\huge}
@@
 
-A few different approaches to modeling absenteeism in the Maragra data
+Maragra models
 
@@
 
-This document contains methods and basic results for a few different modeling approaches, following a conversation and write-up with Menno Pradhan in May 2018.
-The purpose of this document is to make explicit the current models we have implemented, and some issues with them.
+This document contains methods and basic results for a few different modeling approaches.
+It was produced in September 2018 in Amsterdam.
+The purpose of this document is to give an overview of the modeling approached employed so far, as well as provide a summary of results.
 
@@
         \begin{itemize}
-          \item I have implemented Menno's suggestions for more simple models.
-          \item The simplest models show no overall effect from IRS when not accounting for malaria season
-          \item I believe that the lack of an effect is due to seasonality's presence in the model as a fixed (additive) effect, rather than as an interaction with IRS
+          \item We have combined individual and indirect protection into one index
+          \item We have simplified our model approach
+          \item Results are consistent/coherent
         \end{itemize}
@@
 
-\section{The current approach}\label{the-current-approach}
+\section{The previous approach}\label{the-previous-approach}
 
-In the current draft of the paper
+In the draft of the paper
 (\href{https://docs.google.com/document/d/1bUWRBCgVcgjSPHchIQxiTG8Vwv5hV1GLU4Tlu386sWA/edit#}{HERE}
@@
 felm(formula = absent ~ season * months_since + rainy_day + herd | 
-    oracle_number + malaria_year | 0 | 0, data = these_data)
+    oracle_number + malaria_year | 0 | 0, data = these_daily_data)
 \end{verbatim}
@@
 
-\section{Suggested changes from
-Menno}\label{suggested-changes-from-menno}
+\section{The new approach}\label{the-new-approach}
 
-Menno sent a write-up on May 17th (see email with subject ``suggestions
-for models'') with some proposed changes to how we model absenteeism.
-What follows is an implementation of those suggestions.
-
-\subsection{Model A (without
-seasonality)}\label{model-a-without-seasonality}
-
-The below is a characterization of the approach Menno suggested:
-
-\begin{itemize}
-\tightlist
-\item
-  We use monthly instead of daily data (modeling a percentage
-  absenteeism rate rather than a binary).
-\item
-  We use fixed effects for year-month combination.
-\item
-  We don't explicitly model for malaria season.
-\item
-  We don't have any interaction terms.
-\item
-  We use dummies for months since IRS.
-\item
-  We use the most recent spraying as the time since IRS (ie, if 5 months
-  ago, and 1 month ago, then the dummy is applied for 1 month ago but
-  not 5 months ago).
-\end{itemize}
-
-Using this approach, our panel data looks like this:
-
-\begin{verbatim}
-# A tibble: 6 x 6
-  oracle_number group            year_month absent  days months_since_men~
-  <chr>         <chr>            <chr>       <dbl> <int> <fct>            
-1 X045753       Temporary field~ 2015_10       0      13 01               
-2 X040254       Temporary field~ 2016_9        0      30 01               
-3 X040343       Temporary field~ 2014_6        0      30 03               
-4 PF02071       Temporary not f~ 2015_3       80.6    31 06               
-5 PA00795       Permanent field~ 2013_3       67.7    31 02               
-6 PF00699       Permanent not f~ 2016_3        0      31 06               
-\end{verbatim}
-
-The model formula was described in Menno's email.
-
-\includegraphics{images/menno}{[}width=3cm{]}
-
-Which translates into R as:
-
-\begin{verbatim}
-felm(formula = absent ~ months_since_menno | oracle_number + 
-    year_month | 0 | 0, data = these_data)
-\end{verbatim}
-
-Note that the \texttt{months\_since\_menno} variable has the following
-categories, and is treated as a ``wide'' dummy by the software in
-modeling:
-
-\begin{verbatim}
-[1] "Before" "01"     "02"     "03"     "04"     "05"     "06"    
-\end{verbatim}
-
-The results of this model look like this:
-
-\begin{longtable}[t]{ll}
-\caption{\label{tab:unnamed-chunk-8}Simple models with year-month fixed effects}\\
-\toprule
-Term & Estimate\\
-\midrule
-\endfirsthead
-\caption[]{Simple models with year-month fixed effects \textit{(continued)}}\\
-\toprule
-Term & Estimate\\
-\midrule
-\endhead
-\
-\endfoot
-\bottomrule
-\endlastfoot
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
-\hspace{1em}IRS status=\_menno01 & 1.425 (P=0.446)\\
-\hspace{1em}IRS status=\_menno02 & 3.603 (P=0.066)\\
-\hspace{1em}IRS status=\_menno03 & 2.049 (P=0.297)\\
-\hspace{1em}IRS status=\_menno04 & 1.357 (P=0.491)\\
-\hspace{1em}IRS status=\_menno05 & 1.351 (P=0.498)\\
-\hspace{1em}IRS status=\_menno06 & 3.358 (P=0.106)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
-\hspace{1em}IRS status=\_menno01 & -0.022 (P=0.982)\\
-\hspace{1em}IRS status=\_menno02 & 0.255 (P=0.805)\\
-\hspace{1em}IRS status=\_menno03 & 1.803 (P=0.085)\\
-\hspace{1em}IRS status=\_menno04 & -0.87 (P=0.406)\\
-\hspace{1em}IRS status=\_menno05 & 0.144 (P=0.894)\\
-\hspace{1em}IRS status=\_menno06 & 0.073 (P=0.947)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
-\hspace{1em}IRS status=\_menno01 & -0.333 (P=0.261)\\
-\hspace{1em}IRS status=\_menno02 & -0.058 (P=0.847)\\
-\hspace{1em}IRS status=\_menno03 & -0.378 (P=0.215)\\
-\hspace{1em}IRS status=\_menno04 & 0.01 (P=0.973)\\
-\hspace{1em}IRS status=\_menno05 & 0.075 (P=0.797)\\
-\hspace{1em}IRS status=\_menno06 & 0.41 (P=0.17)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
-\hspace{1em}IRS status=\_menno01 & 0.08 (P=0.979)\\
-\hspace{1em}IRS status=\_menno02 & 2.989 (P=0.35)\\
-\hspace{1em}IRS status=\_menno03 & -4.094 (P=0.167)\\
-\hspace{1em}IRS status=\_menno04 & -2.276 (P=0.449)\\
-\hspace{1em}IRS status=\_menno05 & -2.259 (P=0.464)\\
-\hspace{1em}IRS status=\_menno06 & 2.409 (P=0.427)\\*
-\end{longtable}
-
-In the above, it is apparent that the effects of IRS appear to be either
-(a) in the opposite direction as expected or (b) non-signficant. Here is
-a visualization of IRS' apparent effect (relative to the ``Before''
-period, which is the reference class). Were IRS effective, we would
-expect the black line to be below 0 (the red line).
-
-\begin{center}\includegraphics{menno_models_files/figure-latex/unnamed-chunk-9-1} \end{center}
-
-What is going on here? Could it be that IRS has a positive (ie,
-increases) impact on absenteeism? Let's look at the raw data, rather
-than the coefficients. The below chart shows absenteeism (as a
-percentage of the pre-IRS absenteeism average) for workers, broken down
-by season (row) and worker type (column), with point size reflecting the
-number of observations.
-
-\begin{center}\includegraphics{menno_models_files/figure-latex/unnamed-chunk-10-1} \end{center}
-
-In the above chart, two things stand out:
+The new approach, discussed September 26 2018, is characterized by the
+following:
 
@@
 \item
-  Variability is greatest among temporary workers.
+  We assume that the sum of ``spraying'' and rainfall is what determines
+  the quantity of malaria-infectious mosquitoes at any given time.
 \item
-  There appears to be a reduction in absenteeism post IRS for permanent
-  workers \emph{during the malaria season}, but not during the
-  non-malaria season
+  We assume that rainfall's effect is best captured by lagging 4-13
+  weeks prior to the date of observation (Wu et al. 2017).
+\item
+  We assume that the relationship between this sum takes on an
+  exponential shape.
+\item
+  We model separately for worker types (permanent vs.~temporary and
+  factory vs.~field).
+\item
+  We assume that other factors (age, sex, etc.) not captured by the
+  above do not interact with absenteeism or the effect of IRS.
+\item
+  We assume that the protection afforded by IRS is a function of
+  distance to that IRS, and are indifferent to (a) the relative coverage
+  of IRS in your area (the absolute quantity of chemical being the only
+  thing that matters) and (b) whether the IRS is at your house or
+  elsewhere (the distance to the IRS being the only thing that matters)
 \end{enumerate}
 
-So, perhaps controlling for year-month through fixed effects isn't
-enough, since the effect of IRS \emph{interacts} with the season\ldots{}
+\subsection{Specification}\label{specification}
 
-\subsection{Model B (with seasonality)}\label{model-b-with-seasonality}
+\[
+\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{exp(protection)}_{it}  + \hat{\beta}_{2}\text{Rain}_{it} +  \alpha_i + \upsilon_{it}
+\]
 
-Using this approach, our panel data looks like this:
+The above notation shows \texttt{absent} (binary 1,0) as a function of
+the exponentiation of the \texttt{protection} variable (the summation of
+individual plus indirectly conferred protection) and the \texttt{rain}
+variable (the summed lag of rainfall from 4-13 weeks). \(\alpha_i\)
+represents the time invariant worker fixed effects, \(\upsilon\) is the
+error term.
+
+\subsection{Code}\label{code}
+
+In R, our new model looks like this
 
 \begin{verbatim}
-# A tibble: 6 x 7
-  oracle_number group      year_month absent  days months_since_me~ season
-  <chr>         <chr>      <chr>       <dbl> <int> <fct>            <fct> 
-1 X037697       Temporary~ 2016_7       0       31 05               low   
-2 PF01761       Permanent~ 2015_12      3.23    31 04               high  
-3 X040726       Temporary~ 2015_5       3.23    31 02               high  
-4 PF00774       Permanent~ 2014_6       3.33    30 04               low   
-5 X043249       Temporary~ 2016_4       0       30 06               high  
-6 X037700       Permanent~ 2013_2       0       27 02               high  
+felm(formula = log(absent + 1) ~ protection + precipitation_lag1_3 | 
+    oracle_number | 0 | 0, data = monthly)
 \end{verbatim}
 
-Our new model specification (in R) is:
+Where \texttt{protection} is the sum of all nearby IRS, weighted by the
+inverse of the distance from the residence of the person (including the
+person's own residence, set at a distance of 10 meters).
+
+\subsection{Data structure}\label{data-structure}
 
 \begin{verbatim}
-felm(formula = absent ~ months_since_menno * season | oracle_number + 
-    year_month | 0 | 0, data = these_data)
+# A tibble: 6 x 29
+  oracle_number year_month group precipitation absences longitude_aura
+  <chr>         <chr>      <chr>         <dbl>    <int>          <dbl>
+1 CA14693       2013-01-01 Temp~          10.7        0           32.7
+2 CA15945       2013-01-01 Temp~          10.7        0           NA  
+3 CA16471       2013-01-01 Perm~          10.7        2           32.8
+4 CA16473       2013-01-01 Perm~          10.7       11           32.8
+5 CA16476       2013-01-01 Perm~          10.7       11           NA  
+6 CA16477       2013-01-01 Perm~          10.7        0           32.8
+# ... with 23 more variables: latitude_aura <dbl>, ever_sprayed <lgl>,
+#   eligibles <int>, herd <dbl>, protection <dbl>, irs <dbl>, sex <chr>,
+#   permanent_or_temporary <chr>, high_season <dbl>, absent <dbl>,
+#   precipitation_lag0 <dbl>, precipitation_lag1 <dbl>,
+#   precipitation_lag2 <dbl>, precipitation_lag3 <dbl>,
+#   precipitation_lag4 <dbl>, precipitation_lag2_3 <dbl>,
+#   precipitation_lag1_2 <dbl>, precipitation_lag0_2 <dbl>,
+#   precipitation_lag0_3 <dbl>, precipitation_lag1_3 <dbl>,
+#   precipitation_lag1_4 <dbl>, herd_ideal <dbl>, protection_ideal <dbl>
 \end{verbatim}
 
-The results of this model look like this:
+\subsection{Data summary}\label{data-summary}
+
+\begin{center}\includegraphics{sep2018_files/figure-latex/unnamed-chunk-7-1} \end{center}
+
+\begin{center}\includegraphics{sep2018_files/figure-latex/unnamed-chunk-8-1} \end{center}
+
+\begin{center}\includegraphics{sep2018_files/figure-latex/unnamed-chunk-9-1} \end{center}
+
+\newpage
+
+\subsection{Results}\label{results}
+
+The below table shows the results of the model devised thus far.
 
 \begin{longtable}[t]{ll}
-\caption{\label{tab:unnamed-chunk-13}Simple models with year-month fixed effects and seasonality-IRS interaction term}\\
+\caption{\label{tab:unnamed-chunk-10}Model results}\\
 \toprule
@@
 \midrule
-\endfirsthead
-\caption[]{Simple models with year-month fixed effects and seasonality-IRS interaction term \textit{(continued)}}\\
-\toprule
-Term & Estimate\\
-\midrule
-\endhead
-\
-\endfoot
+protection & -0.192 (P=0.021)\\
+Precipitation (mm) \_lag1\_3 & 0.04 (P=0.027)\\
 \bottomrule
-\endlastfoot
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
-\hspace{1em}IRS status=\_menno01 & 175.053 (P=0.532)\\
-\hspace{1em}IRS status=\_menno02 & 457.807 (P=0.115)\\
-\hspace{1em}IRS status=\_menno03 & 505.425 (P=0.061)\\
-\hspace{1em}IRS status=\_menno04 & 233.373 (P=0.357)\\
-\hspace{1em}IRS status=\_menno05 & 227.399 (P=0.372)\\
-\hspace{1em}IRS status=\_menno06 & 425.265 (P=0.121)\\
-\hspace{1em}Malaria season & 887.601 (P=0.703)\\
-\hspace{1em}IRS status=\_menno01:Malaria season & -58.338 (P=0.872)\\
-\hspace{1em}IRS status=\_menno02:Malaria season & -174.777 (P=0.642)\\
-\hspace{1em}IRS status=\_menno03:Malaria season & -603.926 (P=0.105)\\
-\hspace{1em}IRS status=\_menno04:Malaria season & -220.413 (P=0.56)\\
-\hspace{1em}IRS status=\_menno05:Malaria season & -213.141 (P=0.582)\\
-\hspace{1em}IRS status=\_menno06:Malaria season & -191.654 (P=0.633)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
-\hspace{1em}IRS status=\_menno01 & 39.635 (P=0.788)\\
-\hspace{1em}IRS status=\_menno02 & 14.833 (P=0.919)\\
-\hspace{1em}IRS status=\_menno03 & 114.878 (P=0.439)\\
-\hspace{1em}IRS status=\_menno04 & 55.896 (P=0.694)\\
-\hspace{1em}IRS status=\_menno05 & 93.301 (P=0.502)\\
-\hspace{1em}IRS status=\_menno06 & 86.299 (P=0.556)\\
-\hspace{1em}Malaria season & NaN (PNA)\\
-\hspace{1em}IRS status=\_menno01:Malaria season & -72.876 (P=0.707)\\
-\hspace{1em}IRS status=\_menno02:Malaria season & 23.995 (P=0.903)\\
-\hspace{1em}IRS status=\_menno03:Malaria season & 127.34 (P=0.523)\\
-\hspace{1em}IRS status=\_menno04:Malaria season & -296.411 (P=0.137)\\
-\hspace{1em}IRS status=\_menno05:Malaria season & -185.639 (P=0.372)\\
-\hspace{1em}IRS status=\_menno06:Malaria season & -171.932 (P=0.416)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
-\hspace{1em}IRS status=\_menno01 & -35.174 (P=0.339)\\
-\hspace{1em}IRS status=\_menno02 & -6.152 (P=0.869)\\
-\hspace{1em}IRS status=\_menno03 & -12.472 (P=0.732)\\
-\hspace{1em}IRS status=\_menno04 & 18.095 (P=0.611)\\
-\hspace{1em}IRS status=\_menno05 & 33.497 (P=0.346)\\
-\hspace{1em}IRS status=\_menno06 & 9.614 (P=0.816)\\
-\hspace{1em}Malaria season & 44.327 (P=0.309)\\
-\hspace{1em}IRS status=\_menno01:Malaria season & 8.557 (P=0.881)\\
-\hspace{1em}IRS status=\_menno02:Malaria season & 3.8 (P=0.948)\\
-\hspace{1em}IRS status=\_menno03:Malaria season & -75.835 (P=0.21)\\
-\hspace{1em}IRS status=\_menno04:Malaria season & -51.282 (P=0.397)\\
-\hspace{1em}IRS status=\_menno05:Malaria season & -70.353 (P=0.217)\\
-\hspace{1em}IRS status=\_menno06:Malaria season & 63.717 (P=0.259)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
-\hspace{1em}IRS status=\_menno01 & 414.405 (P=0.319)\\
-\hspace{1em}IRS status=\_menno02 & 90.379 (P=0.825)\\
-\hspace{1em}IRS status=\_menno03 & 353.397 (P=0.441)\\
-\hspace{1em}IRS status=\_menno04 & -565.881 (P=0.232)\\
-\hspace{1em}IRS status=\_menno05 & -355.43 (P=0.388)\\
-\hspace{1em}IRS status=\_menno06 & 45.199 (P=0.895)\\
-\hspace{1em}Malaria season & -459.147 (P=0.519)\\
-\hspace{1em}IRS status=\_menno01:Malaria season & -724.867 (P=0.206)\\
-\hspace{1em}IRS status=\_menno02:Malaria season & 661.492 (P=0.283)\\
-\hspace{1em}IRS status=\_menno03:Malaria season & -1223.913 (P=0.031)\\
-\hspace{1em}IRS status=\_menno04:Malaria season & 529.98 (P=0.364)\\
-\hspace{1em}IRS status=\_menno05:Malaria season & 324.283 (P=0.578)\\
-\hspace{1em}IRS status=\_menno06:Malaria season & 817.731 (P=0.227)\\*
 \end{longtable}
 
-The below shows model coefficients:
+\subsection{Intepretation}\label{intepretation}
 
-\begin{center}\includegraphics{menno_models_files/figure-latex/unnamed-chunk-14-1} \end{center}
+The below charts show predicted absenteeism rates at different rain and
+protection levels.
 
-\subsection{Model C (with seasonality and herd
-score)}\label{model-c-with-seasonality-and-herd-score}
+\subsubsection{Precipitation's effect}\label{precipitations-effect}
 
-We now introduce herd immunity to the model. This is introduced as an
-additive effect; perhaps it should be interactive with IRS? Using this
-approach, our panel data looks like this:
+\begin{center}\includegraphics{sep2018_files/figure-latex/unnamed-chunk-12-1} \end{center}
 
-\begin{verbatim}
-# A tibble: 6 x 8
-  oracle_number group       year_month absent  days  herd months_since_me~
-  <chr>         <chr>       <chr>       <dbl> <int> <dbl> <fct>           
-1 PF02492       Permanent ~ 2013_3       9.68    31 136   01              
-2 X034289       Temporary ~ 2016_9       0       30 276   02              
-3 PF00699       Permanent ~ 2016_9       0       30 282   03              
-4 X044684       Temporary ~ 2014_9       0       30 248   01              
-5 X045133       Temporary ~ 2015_8       6.45    31 156   06              
-6 X041009       Temporary ~ 2015_6       0       30  97.1 02              
-# ... with 1 more variable: season <fct>
-\end{verbatim}
+\subsubsection{IRS protection's effect}\label{irs-protections-effect}
 
-Our new model specification (in R) is:
+\begin{center}\includegraphics{sep2018_files/figure-latex/unnamed-chunk-13-1} \end{center}
 
-\begin{verbatim}
-felm(formula = absent ~ months_since_menno * season | oracle_number + 
-    year_month | 0 | 0, data = these_data)
-\end{verbatim}
+\section*{Bibliography}\label{bibliography}
+\addcontentsline{toc}{section}{Bibliography}
 
-The results of this model look like this:
-
-\begin{longtable}[t]{ll}
-\caption{\label{tab:unnamed-chunk-17}Simple models with year-month fixed effects, seasonality-IRS interaction term, and herd score}\\
-\toprule
-Term & Estimate\\
-\midrule
-\endfirsthead
-\caption[]{Simple models with year-month fixed effects, seasonality-IRS interaction term, and herd score \textit{(continued)}}\\
-\toprule
-Term & Estimate\\
-\midrule
-\endhead
-\
-\endfoot
-\bottomrule
-\endlastfoot
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
-\hspace{1em}IRS status=\_menno01 & 175.053 (P=0.532)\\
-\hspace{1em}IRS status=\_menno02 & 457.807 (P=0.115)\\
-\hspace{1em}IRS status=\_menno03 & 505.425 (P=0.061)\\
-\hspace{1em}IRS status=\_menno04 & 233.373 (P=0.357)\\
-\hspace{1em}IRS status=\_menno05 & 227.399 (P=0.372)\\
-\hspace{1em}IRS status=\_menno06 & 425.265 (P=0.121)\\
-\hspace{1em}Malaria season & 887.601 (P=0.703)\\
-\hspace{1em}IRS status=\_menno01:Malaria season & -58.338 (P=0.872)\\
-\hspace{1em}IRS status=\_menno02:Malaria season & -174.777 (P=0.642)\\
-\hspace{1em}IRS status=\_menno03:Malaria season & -603.926 (P=0.105)\\
-\hspace{1em}IRS status=\_menno04:Malaria season & -220.413 (P=0.56)\\
-\hspace{1em}IRS status=\_menno05:Malaria season & -213.141 (P=0.582)\\
-\hspace{1em}IRS status=\_menno06:Malaria season & -191.654 (P=0.633)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
-\hspace{1em}IRS status=\_menno01 & 39.635 (P=0.788)\\
-\hspace{1em}IRS status=\_menno02 & 14.833 (P=0.919)\\
-\hspace{1em}IRS status=\_menno03 & 114.878 (P=0.439)\\
-\hspace{1em}IRS status=\_menno04 & 55.896 (P=0.694)\\
-\hspace{1em}IRS status=\_menno05 & 93.301 (P=0.502)\\
-\hspace{1em}IRS status=\_menno06 & 86.299 (P=0.556)\\
-\hspace{1em}Malaria season & NaN (PNA)\\
-\hspace{1em}IRS status=\_menno01:Malaria season & -72.876 (P=0.707)\\
-\hspace{1em}IRS status=\_menno02:Malaria season & 23.995 (P=0.903)\\
-\hspace{1em}IRS status=\_menno03:Malaria season & 127.34 (P=0.523)\\
-\hspace{1em}IRS status=\_menno04:Malaria season & -296.411 (P=0.137)\\
-\hspace{1em}IRS status=\_menno05:Malaria season & -185.639 (P=0.372)\\
-\hspace{1em}IRS status=\_menno06:Malaria season & -171.932 (P=0.416)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
-\hspace{1em}IRS status=\_menno01 & -35.174 (P=0.339)\\
-\hspace{1em}IRS status=\_menno02 & -6.152 (P=0.869)\\
-\hspace{1em}IRS status=\_menno03 & -12.472 (P=0.732)\\
-\hspace{1em}IRS status=\_menno04 & 18.095 (P=0.611)\\
-\hspace{1em}IRS status=\_menno05 & 33.497 (P=0.346)\\
-\hspace{1em}IRS status=\_menno06 & 9.614 (P=0.816)\\
-\hspace{1em}Malaria season & 44.327 (P=0.309)\\
-\hspace{1em}IRS status=\_menno01:Malaria season & 8.557 (P=0.881)\\
-\hspace{1em}IRS status=\_menno02:Malaria season & 3.8 (P=0.948)\\
-\hspace{1em}IRS status=\_menno03:Malaria season & -75.835 (P=0.21)\\
-\hspace{1em}IRS status=\_menno04:Malaria season & -51.282 (P=0.397)\\
-\hspace{1em}IRS status=\_menno05:Malaria season & -70.353 (P=0.217)\\
-\hspace{1em}IRS status=\_menno06:Malaria season & 63.717 (P=0.259)\\
-\addlinespace[1.5em]
-\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
-\hspace{1em}IRS status=\_menno01 & 414.405 (P=0.319)\\
-\hspace{1em}IRS status=\_menno02 & 90.379 (P=0.825)\\
-\hspace{1em}IRS status=\_menno03 & 353.397 (P=0.441)\\
-\hspace{1em}IRS status=\_menno04 & -565.881 (P=0.232)\\
-\hspace{1em}IRS status=\_menno05 & -355.43 (P=0.388)\\
-\hspace{1em}IRS status=\_menno06 & 45.199 (P=0.895)\\
-\hspace{1em}Malaria season & -459.147 (P=0.519)\\
-\hspace{1em}IRS status=\_menno01:Malaria season & -724.867 (P=0.206)\\
-\hspace{1em}IRS status=\_menno02:Malaria season & 661.492 (P=0.283)\\
-\hspace{1em}IRS status=\_menno03:Malaria season & -1223.913 (P=0.031)\\
-\hspace{1em}IRS status=\_menno04:Malaria season & 529.98 (P=0.364)\\
-\hspace{1em}IRS status=\_menno05:Malaria season & 324.283 (P=0.578)\\
-\hspace{1em}IRS status=\_menno06:Malaria season & 817.731 (P=0.227)\\*
-\end{longtable}
+\hypertarget{refs}{}
+\hypertarget{ref-Wu_2017}{}
+Wu, Yunyun, Zhijiao Qiao, Nan Wang, Hongjie Yu, Zijian Feng, Xiaosong
+Li, and Xing Zhao. 2017. ``Describing Interaction Effect Between Lagged
+Rainfalls on Malaria: An Epidemiological Study in South--West China.''
+\emph{Malaria Journal} 16 (1). Springer Nature.
+doi:\href{https://doi.org/10.1186/s12936-017-1706-2}{10.1186/s12936-017-1706-2}.
 

### 2020-06-20 data-raw/create_data_files.R

@@ -61,9 +61,9 @@ ab <- ab %>%
          date_captured = format_date(date_captured),
          date_signed_leave_form_rec = format_date(date_signed_leave_form_rec))
 
-# Keep only post 2013
+# Keep only post 2012
 ab <- ab %>%
-  filter(leave_from_date >= '2013-01-01')
+  filter(leave_from_date >= '2012-01-01')
 
 # Keep only relevant columns
 ab <-
@@ -213,8 +213,8 @@ workers <-
 
 # We need to get worker eligible days into an expanded (panel) format
 expandify <- function(oracle_number = 'abc',
-                      start = as.Date('2015-01-01'),
-                      end = as.Date('2015-01-15')){
+                      start = NULL,#as.Date('2015-01-01'),
+                      end = NULL){#as.Date('2015-01-15')){
   if(is.na(start) | is.na(end) |
      start > end){
     out <- data_frame(oracle_number = as.character(rep(NA, 0)),
@@ -229,46 +229,65 @@ expandify <- function(oracle_number = 'abc',
   return(out)
 }
 
+
+
+# Expand absences
+if('ab_expanded.RData' %in% dir()){
+  load('ab_expanded.RData')
+} else {
+  ab_expanded <- list()
+  nra <- nrow(ab)
+  for (i in 1:nra){
+    message(i, ' of ', nra)
+    this_oracle_number <- ab$oracle_number[i]
+    sub_data <- ab[i,]
+    x <- expandify(oracle_number = this_oracle_number,
+                   start = sub_data$leave_from_date,
+                   end = sub_data$leave_to_date)
+    x <- left_join(x,
+                   y = sub_data %>%
+                     dplyr::select(oracle_number,
+                                   leave_type,
+                                   leave_taken),
+                   by = 'oracle_number')
+    ab_expanded[[i]] <- x
+  }
+  ab_expanded <- bind_rows(ab_expanded)
+  ab_expanded$absent <- TRUE
+  # Remove any potential duplicates
+  ab_expanded <- ab_expanded %>%
+    dplyr::distinct(oracle_number, date, .keep_all = TRUE)
+  save(ab_expanded,
+       file = 'ab_expanded.RData')
+}
+
 # Go through each worker and get the eligible working days
 if('eligible_working_days.RData' %in% dir()){
   load('eligible_working_days.RData')
 } else {
   
-  # CGIs are the temp workers (those that get contract end dates)
-  # Let's see when their last absences were
-  temp <- ab %>%
[…]

### 2020-06-20 inst/rmd/apr2020/code.R

@@ -1,0 +1,746 @@
+# Packages
+library(sp)
+library(tidyverse)
+library(knitr)
+library(Hmisc)
+library(brew)
+library(maragra)
+library(knitr)
+library(RColorBrewer)
+# library(lme4)
+library(lfe)
+library(restorepoint)
+library(regtools)
+# 
+# # # Elisa: active workers by worker type over time
+# pd <- maragra::ab_panel %>%
+#   left_join(
+#   maragra::workers %>% dplyr::select(oracle_number, permanent_or_temporary)) %>%
+#   mutate(type = paste0(permanent_or_temporary, ' ', department)) %>%
+#   mutate(year_month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+#   group_by(type, year_month) %>%
+#   tally
+# 
+# ggplot(data = pd,
+#        aes(x = year_month,
+#            y = n,
+#            fill = type,
+#            group = type)) +
+#   geom_bar(stat = 'identity',
+#            position = position_dodge())
+# write_csv(pd, '~/Desktop/pd.csv')
+# # 
+# pd <- maragra::ab_panel %>%
+#   left_join(
+#     maragra::workers %>% dplyr::select(oracle_number, permanent_or_temporary)) %>%
+#   mutate(type = paste0(permanent_or_temporary)) %>%
+#   mutate(year_month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+#   group_by(type, year_month) %>%
+#   tally
+# 
+# ggplot(data = pd,
+#        aes(x = year_month,
+#            y = n,
+#            fill = type,
+#            color = type)) +
+#   geom_line() +
+#   geom_point() +
+#   xlim(as.Date('2013-01-01'),
+#        as.Date('2016-06-01'))
+# 
+# pd <- maragra::ab_panel %>%
+#   left_join(
+#     maragra::workers %>% dplyr::select(oracle_number, permanent_or_temporary, sex)) %>%
+#   mutate(type = paste0(permanent_or_temporary, ' ', sex)) %>%
+#   mutate(year_month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+#   group_by(type, year_month) %>%
+#   tally
+# 
+# pd$type <- factor(pd$type, levels = rev(unique(pd$type)))
+# ggplot(data = pd,
+#        aes(x = year_month,
+#            y = n,
[…]

### 2020-06-20 inst/rmd/defunct/feb2019_xinavane_reproduction/felm.r

@@ -1,0 +1,58 @@
+examples.predict.felm = function() {
+  T = 12
+  x <- rnorm(T)
+  fe1 = sample(c("yes","no"), T, replace=TRUE)
+  fe2 = sample(c("a","b","c"), T, replace=TRUE)
+  y <- x1  + (fe1=="yes") + 2*(fe2=="a") + (fe2=="b")+ rnorm(T)
+  
+  new <- data.frame(
+    x = seq(-3, 3, length=T), 
+    fe1 = sample(c("yes","no"), T, replace=TRUE),
+    fe2 = sample(c("a","b","c"), T, replace=TRUE)    
+  )
+  p.lm = predict(lm(y ~ x+fe1+fe2), new)
+
+  library(lfe)  
+  p.felm = predict(felm(y~x|fe1+fe2),new)
+  cbind(p.lm, p.felm)
+}
+
+#' An implementaion of predict for felm
+predict.felm = function(object, newdata, use.fe = TRUE,...) {
+  restore.point("predict.felm")
+  co = coef(object)
+  
+  # too make code compatible with tibbles
+  newdata = as.data.frame(newdata)
+  
+  rownames(newdata) = seq_along(newdata[,1])
+  # model matrix without intercept
+  mm = model.matrix(object = object,data = newdata)
+  
+  if (NROW(mm)<NROW(newdata)) {
+    warning("Observations dropped from newdata due to NA.")
+  }
+  
+  # remove intercept
+  if (NCOL(mm)==length(co)+1) {
+    mm = mm[,-1,drop=FALSE]
+  }
+  y.pred = mm %*% co
+  
+  fe.vars = names(object$fe)
+  if (use.fe & length(fe.vars)>0) {
+    rows = as.integer(rownames(mm))
+    nd = newdata[rows,]
+    all.fe = getfe(object)
+    fe.var = fe.vars[1]
+    for (fe.var in fe.vars) {
+      df = all.fe[all.fe$fe == fe.var,]
+      frows = match(nd[[fe.var]],df$idx)    
+      myfe = df$effect[frows]
+      myfe[is.na(myfe)] = 0
+      
+      y.pred = y.pred + myfe
+    }
+  }
+  as.vector(y.pred)
+}

### 2020-06-20 inst/rmd/defunct/feb2019_xinavane_reproduction/prepare_data.R

@@ -1,0 +1,436 @@
+# Packages
+library(sp)
+library(tidyverse)
+library(knitr)
+library(Hmisc)
+library(brew)
+library(maragra)
+library(knitr)
+library(RColorBrewer)
+# library(lme4)
+library(lfe)
+library(restorepoint)
+library(regtools)
+
+
+classify_protection <- function(days_since){
+  ifelse(days_since < 21 | days_since > (183+21),
+         0,
+         ifelse(days_since < 51, 1,
+                ifelse(days_since < 81,
+                       0.9,
+                       ifelse(days_since < 112,
+                              0.8,
+                              ifelse(days_since < 143,
+                                     0.6,
+                                     ifelse(days_since <174,
+                                            0.4,
+                                            ifelse(days_since <= (183+21),
+                                                   0.2,
+                                                   0)))))))
+}
+
+
+if('prepared_data.RData' %in% dir()){
+  load('prepared_data.RData')
+} else {
+  # Read in data
+  ab <- maragra::ab
+  ab_panel <- maragra::ab_panel
+  bairros <- maragra::bairros
+  bes <- maragra::bes
+  census <- maragra::census
+  clinic <- maragra::clinic
+  clinic_agg <- maragra::clinic_agg
+  mc <- maragra::mc
+  workers <- maragra::workers
+  
+  if(!'weather.RData' %in% dir()){
+    weather <- maragra::weather %>%
+    mutate(precipitation = ifelse(is.na(precipitation),
+                                  0,
+                                  precipitation))
+    
+    # Get lags for weather
+    starts <- c(15, 30, 60, 90)
+    ends <- c(30, 60, 90, 120)
+    for(k in 1:length(starts)){
+      for(j in 1:length(ends)){
+        this_start <- starts[k]
+        this_end <- ends[j]
+        if(this_start < this_end){
+          precipitation_column_name <- paste0('precipitation_lag_', this_start, '_', this_end)
+          weather[,precipitation_column_name] <- NA
+          temp_column_name <- paste0('temp_lag_', this_start, '_', this_end)
+        }
+      }
+    }
+    for(i in 1:nrow(weather)){
+      starts <- c(15, 30, 60, 90)
+      ends <- c(30, 60, 90, 120)
+      for(k in 1:length(starts)){
+        for(j in 1:length(ends)){
+          this_start <- starts[k]
+          this_end <- ends[j]
+          if(this_start < this_end){
+            this_date <- weather$date[i]
[…]

### 2020-06-20 inst/rmd/functions.R

@@ -1,0 +1,188 @@
+library(shinydashboard)
+library(googlesheets)
+library(ggplot2)
+library(dplyr)
+library(Hmisc)
+library(ggridges)
+
+
+# Define plotting theme
+theme_databrew <- function (base_size = 15, base_family = "Sawasdee", y_comma = TRUE, 
+                            subtitle_family = "Sawasdee", axis_family = "Sawasdee") 
+{
+  color_background = "#F8F5E1"
+  color_grid_major = "grey"
+  color_axis_text = "#0d63c4"
+  color_axis_title = color_axis_text
+  color = "darkgrey"
+  color_title = color_axis_text
+  color_subtitle = grey(0.1)
+  base_size1 = base_size
+  out <- theme_bw(base_size = base_size1) + theme(panel.background = element_rect(fill = color_background, 
+                                                                                  color = color_background)) + theme(plot.background = element_rect(fill = color_background, 
+                                                                                                                                                    color = color_background)) + theme(panel.border = element_rect(color = color_background)) + 
+    theme(panel.grid.major = element_line(color = adjustcolor(color_grid_major, 
+                                                              alpha.f = 0.25), size = 0.25)) + theme(panel.grid.major = element_line(color = adjustcolor(color_grid_major, 
+                                                                                                                                                         alpha.f = 0.4), size = 0.4)) + theme(panel.grid.minor = element_blank()) + 
+    theme(axis.ticks = element_blank()) + theme(legend.background = element_rect(fill = color_background)) + 
+    theme(legend.text = element_text(family = base_family, 
+                                     size = base_size * 0.7, color = color_axis_title)) + 
+    theme(plot.title = element_text(family = base_family, 
+                                    color = color_title, size = base_size * 1.2, vjust = 1.25)) + 
+    theme(plot.subtitle = element_text(family = subtitle_family, 
+                                       color = color_subtitle, size = base_size * 0.8, vjust = 1.25)) + 
+    theme(axis.text.x = element_text(family = axis_family, 
+                                     size = base_size * 0.7, color = color_axis_text)) + 
+    theme(axis.text.y = element_text(family = axis_family, 
+                                     size = base_size * 0.7, color = color_axis_text)) + 
+    theme(axis.title.x = element_text(family = axis_family, 
[…]

### 2020-06-20 inst/rmd/nov2018/prepare_data.R

@@ -1,0 +1,649 @@
+# Packages
+library(sp)
+library(tidyverse)
+library(knitr)
+library(Hmisc)
+library(brew)
+library(maragra)
+library(knitr)
+library(RColorBrewer)
+# library(lme4)
+library(lfe)
+library(restorepoint)
+library(regtools)
+
+
+classify_protection <- function(days_since){
+  ifelse(days_since < 21 | days_since > (183+21),
+         0,
+         ifelse(days_since < 51, 1,
+                ifelse(days_since < 81,
+                       0.9,
+                       ifelse(days_since < 112,
+                              0.8,
+                              ifelse(days_since < 143,
+                                     0.6,
+                                     ifelse(days_since <174,
+                                            0.4,
+                                            ifelse(days_since <= (183+21),
+                                                   0.2,
+                                                   0)))))))
+}
+
+
+if('prepared_data.RData' %in% dir()){
+  load('prepared_data.RData')
+} else {
+  # Read in data
+  ab <- maragra::ab
+  ab_panel <- maragra::ab_panel
+  bairros <- maragra::bairros
+  bes <- maragra::bes
+  census <- maragra::census
+  clinic <- maragra::clinic
+  clinic_agg <- maragra::clinic_agg
+  mc <- maragra::mc
+  workers <- maragra::workers
+  
+  if(!'weather.RData' %in% dir()){
+    weather <- maragra::weather %>%
+    mutate(precipitation = ifelse(is.na(precipitation),
+                                  0,
+                                  precipitation))
+    
+    # Get lags for weather
+    starts <- c(15, 30, 60, 90)
+    ends <- c(30, 60, 90, 120)
+    for(k in 1:length(starts)){
+      for(j in 1:length(ends)){
+        this_start <- starts[k]
+        this_end <- ends[j]
+        if(this_start < this_end){
+          precipitation_column_name <- paste0('precipitation_lag_', this_start, '_', this_end)
+          weather[,precipitation_column_name] <- NA
+          temp_column_name <- paste0('temp_lag_', this_start, '_', this_end)
+        }
+      }
+    }
+    for(i in 1:nrow(weather)){
+      starts <- c(15, 30, 60, 90)
+      ends <- c(30, 60, 90, 120)
+      for(k in 1:length(starts)){
+        for(j in 1:length(ends)){
+          this_start <- starts[k]
+          this_end <- ends[j]
+          if(this_start < this_end){
+            this_date <- weather$date[i]
[…]

### 2020-06-20 inst/rmd/sep2018/prepare_data.R

@@ -259,6 +259,7 @@ if('prepared_data.RData' %in% dir()){
                                  'latitude_aura'))
   
   save.image('temp.RData')
+  
 
   # Remove nas
   model_data <- model_data %>%
@@ -271,48 +272,200 @@ if('prepared_data.RData' %in% dir()){
            !is.na(herd),
            !is.na(malaria_year))
   
-  # # Cut down from 4 to 3 groups
-  # model_data <- model_data %>%
-  #   filter(group != 'Temporary not field worker') %>%
-  #   mutate(group = ifelse(group == 'Permanent not field worker',
-  #                         'Non-field worker',
-  #                         group))
+  # Get lagged rainfally
+  rain_lag <- function(x){
+    dplyr::lag(x, 22) + 
+      dplyr::lag(x, 23) + 
+      dplyr::lag(x, 24) + 
+      dplyr::lag(x, 25) + 
+      dplyr::lag(x, 26) + 
+      dplyr::lag(x, 27) + 
+      dplyr::lag(x, 28) + 
+      dplyr::lag(x, 29) + 
+      dplyr::lag(x, 30) + 
+      dplyr::lag(x, 31) + 
+      dplyr::lag(x, 32) + 
+      dplyr::lag(x, 33) + 
+      dplyr::lag(x, 34) + 
+      dplyr::lag(x, 35) + 
+      dplyr::lag(x, 36) + 
+      dplyr::lag(x, 37) + 
+      dplyr::lag(x, 38) + 
+      dplyr::lag(x, 39) + 
+      dplyr::lag(x, 40) + 
+      dplyr::lag(x, 41) + 
+      dplyr::lag(x, 42) + 
+      dplyr::lag(x, 43) + 
+      dplyr::lag(x, 44) + 
+      dplyr::lag(x, 45) + 
+      dplyr::lag(x, 46) + 
+      dplyr::lag(x, 47) + 
+      dplyr::lag(x, 48) + 
+      dplyr::lag(x, 49) + 
+      dplyr::lag(x, 50) + 
+      dplyr::lag(x, 51) + 
+      dplyr::lag(x, 52) + 
+      dplyr::lag(x, 53) + 
+      dplyr::lag(x, 54) + 
+      dplyr::lag(x, 55) + 
+      dplyr::lag(x, 56) + 
+      dplyr::lag(x, 57) + 
[…]

### 2020-06-21 inst/rmd/apr2020/apr2020.Rmd
updated;

@@
 options(xtable.comment = FALSE)
+source('code.R')
 ```
 
+*Absences*: We observed `r nrow(ab_panel)` unique worker-days among `r length(unique(ab_panel$oracle_number))` workers.
+The all-period average absenteeism rate was `r round(length(which(ab_panel$absent)) / nrow(ab_panel) * 100, digits = 2)`%, though this rate varied widely as a function of worker department, sex, residence, and season (table).
+
+# Model results (one model)
+
 ```{r}
-classify_protection <- function(days_since){
-  ifelse(days_since < 21 | days_since > (365+21),0,
-         (((365+21) - days_since)/(365))
-         )
-}
-
-if('prepared_data.RData' %in% dir()){
-  load('prepared_data.RData')
-} else {
-  # Read in data
-  ab <- maragra::ab
-  ab_panel <- maragra::ab_panel
-  bairros <- maragra::bairros
-  bes <- maragra::bes
-  census <- maragra::census
-  clinic <- maragra::clinic
-  clinic_agg <- maragra::clinic_agg
-  mc <- maragra::mc
-  workers <- maragra::workers
-
-  if(!'weather.RData' %in% dir()){
-    weather <- maragra::weather %>%
-    mutate(precipitation = ifelse(is.na(precipitation),
-                                  0,
-                                  precipitation))
-
-    # Get lags for weather
-    starts <- c(15, 30, 60, 90)
-    ends <- c(30, 60, 90, 120)
-    for(k in 1:length(starts)){
-      for(j in 1:length(ends)){
-        this_start <- starts[k]
-        this_end <- ends[j]
-        if(this_start < this_end){
-          precipitation_column_name <- paste0('precipitation_lag_', this_start, '_', this_end)
-          weather[,precipitation_column_name] <- NA
-          temp_column_name <- paste0('temp_lag_', this_start, '_', this_end)
-        }
-      }
-    }
-    for(i in 1:nrow(weather)){
-      starts <- c(15, 30, 60, 90)
-      ends <- c(30, 60, 90, 120)
-      for(k in 1:length(starts)){
-        for(j in 1:length(ends)){
-          this_start <- starts[k]
-          this_end <- ends[j]
-          if(this_start < this_end){
-            this_date <- weather$date[i]
-            end_date <- this_date - this_start
-            start_date <- this_date - this_end
-            if(start_date >= min(weather$date,na.rm = TRUE) &
-               end_date <= max(weather$date, na.rm = TRUE)){
-              sub_data <- weather %>%
-                filter(date >= start_date,
-                       date <= end_date)
-              precipitation_column_name <- paste0('precipitation_lag_', this_start, '_', this_end)
-              weather[i,precipitation_column_name] <- mean(sub_data$precipitation, na.rm = TRUE)
-              temp_column_name <- paste0('temp_lag_', this_start, '_', this_end)
-              weather[i,temp_column_name] <- mean(sub_data$temp, na.rm = TRUE)
-            }
-          }
-        }
-      }
-    }
-    save(weather, file = 'weather.RData')
-  } else {
-    load('weather.RData')
-  }
-
-  # Create a dictionary to classify protection
-  protection_dict <- 
-    data_frame(days_since = seq(min(irs$days_since, na.rm = TRUE),
-                                max(irs$days_since, na.rm = TRUE),
-                                by = 1)) %>%
-    mutate(protection = classify_protection(days_since))
-  model_data <- 
-    ab_panel %>%
-    dplyr::select(oracle_number, date, absent, unidade) %>%
-    left_join(irs %>% 
-                dplyr::select(date, unidade, chemical, days_since), 
-              by = c('unidade', 'date')) %>%
-    # mutate(protection = ifelse(days_since < 21 |
-    #                              days_since > (183+21) |
-    #                              is.na(days_since),
-    #                            0, 1)) %>%
-    left_join(protection_dict) %>%
-    mutate(protection = ifelse(is.na(protection), 0, protection)) %>%
-    dplyr::select(-days_since) %>%
-    group_by(oracle_number) %>%
-    mutate(ever_sprayed = length(which(protection > 0)) > 0) %>%
-  # Bring in some information on workers
-    left_join(workers %>%
-                dplyr::select(oracle_number,
-                              permanent_or_temporary,
-                              department,
-                              sex,
-                              # date_of_birth,
-                              perm_id,
-                              census_name_match_score) %>%
-                # void the permids of anyone with a match score of greater than 0.25
-                mutate(perm_id = ifelse(census_name_match_score > 0.25,
-                                        NA,
-                                        perm_id))) %>%
-    # Bring in some info from the census
-    left_join(census %>%
-                filter(!duplicated(perm_id)) %>%
-                dplyr::select(perm_id,
-                              maragra_bairro,
-                              maragra_fabrica,
-                              education,
-                              floor_material)) %>%
-    # Bring in some data for weather
-    left_join(weather)
-
-    ########## !!!!!!!!!!!!!!! REMOVING 2016 for temp workers
-  model_data <- model_data %>%
-    filter(permanent_or_temporary == 'Permanent' |
-             (permanent_or_temporary == 'Temporary' & date <= '2015-12-31'))
-
-  # Since ad and factory are same, keep same
-  model_data$field <- ifelse(model_data$department == 'Field', 
-                             'Field worker',
-                             'Not field worker')
-
-  model_data <- model_data %>% 
-    mutate(group = paste0(permanent_or_temporary, ' ',
-                          tolower(field)))
-
-  # Add a month column
-  model_data$month_number <- add_zero(format(model_data$date, '%m'), 2)
-
-  # Get geographic info for externality analysis
-  # Get latitude / longitude into model_data
-  model_data <- model_data %>%
-    left_join(census %>%
-                filter(!duplicated(unidade)) %>%
-                dplyr::select(unidade,
-                              longitude_aura,
-                              latitude_aura),
-              by = 'unidade')
-
-  model_data$p <- model_data$protection
-
-  # Centers of each bairro
-  bairro_loc <- apply(coordinates(maragra::bairros_maragra_bairro), 2, mean)
-
-  # Define function for weighting indirect protection
-  weighter <- function(x){
-    # out <- (1 / (x+0.01))
-    # out <- 1/x
-    out <- ifelse(x > 1, 0, 1-x)
-    return(out)
-  }
-
-  # Define function for getting distance between houses
-  overall_spatial <- model_data %>%
-    group_by(lat = latitude_aura,
-             lng = longitude_aura) %>%
-    tally %>%
-    dplyr::distinct(lat, lng, .keep_all = TRUE) %>%
-    dplyr::select(-n) %>%
-    filter(!is.na(lat),
-           !is.na(lng)) %>%
-    bind_rows(data.frame(lat = bairro_loc[2],
-                         lng = bairro_loc[1])) %>%
-    mutate(x = lng,
-           y = lat)
-  overall_spatial$overall_id <- 1:nrow(overall_spatial)
-  coordinates(overall_spatial) <- ~x+y
-  proj4string(overall_spatial) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
-  overall_distances <- spDists(x = overall_spatial, longlat = TRUE)
-
-  get_protection <- function(lat, lng, protection, guess_missing = TRUE){
-    # Build a dataframe
-    df <- data_frame(lat, lng, protection)
-    # Get id
-    df$id <- 1:nrow(df)
-
-    if(guess_missing){
-      # Define function for guessing
-      guess <- function(x){
-        x <- x %>%
-          mutate(lng = ifelse(is.na(lng), 
-                              bairro_loc[1],
-                              lng)) %>%
-
-          mutate(lat = ifelse(is.na(lat), 
-                              bairro_loc[2],
-                              lat)) 
-        return(x)
-      }
-      # if no geography (due to no census matching, just use average geo)
-      df <- guess(df)
-    }
-    # save for later
-    original_df <- df
-
-    # Get unique lat/lng pairs
-    df <- df %>%
-      group_by(lat, lng) %>%
-      summarise(protection = mean(protection, na.rm = TRUE))
-
-    # Join to overall spatial to get distances
-    df <- left_join(df, overall_spatial@data, by = c('lat', 'lng'))
-    keeps <- df$overall_id
-    distances <- overall_distances[keeps, keeps]
-
-    # # Get a spatial version
-    # (no longer necessary with above approach)
-    # df_spatial <- df %>% filter(!is.na(lng),
-    #                             !is.na(lat))
-    # 
-    # coordinates(df_spatial) <- ~lng+lat
-    # proj4string(df_spatial) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
-    # 
-    # # Get distances
-    # distances <- spDists(x = df_spatial, longlat = TRUE)
-
-    # Weight distances and return weighted protection
-    out <- apply(distances, 1, function(x){
-      val <- df$protection * weighter(x)
-      # Infinite means distance was 0/self; remove
-      val[is.infinite(val)] <- NA
-      # Keep only those within a kilometer
-      val <- val[x <= 1]
-      mean(val, na.rm = TRUE)
-
-      })
-    right <- data.frame(lat = df$lat[!is.na(df$lat)],
-                        lng = df$lng[!is.na(df$lng)],
-                        herd = out)
-    # Join to original df
-    original_df <- original_df %>%
-      left_join(right,
-                by = c('lat', 'lng'))
-
-    # Return just the protection column
-    return(original_df$herd)
-  }
-
-  # For each day, get real, and hypothetical protection scores
-  dates <- sort(unique(model_data$date))
-
-  out_list <- list()
-  for(i in 1:length(dates)){
-    this_date <- dates[i]
-    message(this_date)
-
-    this_data <- model_data %>%
-      filter(date == this_date)
-    # Get protection score
-    protection_scores <- get_protection(lat = this_data$latitude_aura,
-                                        lng = this_data$longitude_aura,
-                                        protection = this_data$protection,
-                                        guess_missing = TRUE)
-    this_data$herd_protection <- protection_scores
-    out_list[[i]] <- this_data 
-  }
-  model_data <- bind_rows(out_list)
-
-  # Plots of maps
-  # Libraries
-  library(tidyverse)
-  library(raster)
-  library(tidyr)
-  library(ggplot2)
-  library(RColorBrewer)
-  library(broom)
-  library(ggthemes)
-
-  # Get each countries shapefile
-  countries <- c('Mozambique')
-  iso3s <- c('MOZ')
-  for(i in 1:length(countries)){
-    message('Fetching data for ', countries[i])
-    x <- getData(name = 'GADM', level = 2, country = iso3s[i])
-    assign(tolower(countries[i]),
-           x,
-           envir = .GlobalEnv)
-  }
-
-  # Define which districts are "special" -------------
-  specials <- 'Manhiça'
-
-  # Mozambique
-  mozambique@data$special <- FALSE
-  mozambique@data$special[mozambique@data$NAME_2 %in% specials] <- TRUE
-
-  # Combine all data into a "long" / "tidy" format
-  make_long <- function(x, region = "NAME_2"){
-    shp_df <- broom::tidy(x, region = region)
-    return(shp_df)
-  }
-  mozambique_long <- make_long(mozambique) %>% mutate(country = 'Mozambique')
-  combined <- mozambique_long
-  combined$special <- combined$id %in% specials
-  # combined$special[!combined$special] <- NA
-  # Get a map of africa to use as a background
-  # from the cism package!
-  africa <- cism::africa
-  africa_long <- make_long(africa, region = 'COUNTRY')
-  africa_long$special <- africa_long$id %in% countries
-
-  save.image(file = 'prepared_data.RData')
-}
-
-# Remove christmas
-final_data <- final_data %>%
-  mutate(day_number = as.numeric(format(date, '%d')))
-remove_christmas <- TRUE
-if(remove_christmas){
-  remove_these <- which(
-    (final_data$month_number == '12' &
-       final_data$day_number >= 15) |
-      final_data$month_number == '01' &
-      final_data$day_number <= 15
-
-  )
-}
-
-# Prepare simulations
-if('simulations.RData' %in% dir()){
-  load('simulations.RData')
-} else {
-
-  time_optimize_df <-
-    data.frame(date = seq(as.Date('2020-01-01'),
-                          as.Date('2020-12-31'), 
-                          by = 1)) %>%
-    mutate(month_day = format(date, '%m-%d'))
-  time_optimize_df$day_number <- (1:nrow(time_optimize_df))+20
-  time_optimize_df$the_protection <- classify_protection(time_optimize_df$day_number)
-  time_optimize_df <- time_optimize_df %>% dplyr::select(month_day, the_protection)
-
-  strategies <- c('Zero', 'Max', 'Time-optimized')
-  strategy_list <- list()
-  for(s in 1:length(strategies)){
-    dates <- sort(unique(final_data$date))
-    out_list <- list()
-    this_strategy <- strategies[s]
-
-    for(i in 1:length(dates)){
-      this_date <- dates[i]
-      message(this_strategy, '...', this_date)
-
-      this_data <- final_data %>%
-        filter(date == this_date)
-
-      # Define scores
-      if(this_strategy == 'Zero'){
-        this_data$protection <- 0
-        the_protection <- rep(0, length(this_data$protection))
-      }
-      if(this_strategy == 'Max'){
-        this_data$protection <- 1
-        the_protection <- rep(1, length(this_data$protection))
-      }
-      if(this_strategy == 'Time-optimized'){
-        this_data$month_day <- format(this_data$date, '%m-%d')
-        x <- this_data %>%
-          left_join(time_optimize_df)
-        the_protection <- x$the_protection
-        this_data$protection <- the_protection
-      }
-      # Get protection score
-      protection_scores <- get_protection(lat = this_data$latitude_aura,
-                                          lng = this_data$longitude_aura,
-                                          protection = the_protection,
-                                          guess_missing = TRUE)
-      this_data$herd_protection <- protection_scores
-      out_list[[i]] <- this_data 
-    }
-    out <- bind_rows(out_list)
-    out$strategy <- this_strategy
-    strategy_list[[s]] <- out
-  }
-  simulations <- bind_rows(strategy_list)
-  simulations <- simulations %>%
-    bind_rows(final_data %>% mutate(strategy = 'Real'))
-  save(simulations, file = 'simulations.RData')
-}
-
-# Some variable things
-model_data$rain_var <- model_data$precipitation_lag_15_90
-
-# # Final data - only the ever sprayed
-final_data <- model_data %>%
-  filter(ever_sprayed)
-#   # Temporary fix pending newer data
-#   filter(date <= '2016-01-01')
-
-# Add 5th worker type
-final_data <- final_data %>%
-  # Get grade
-  left_join(ab_panel %>% dplyr::distinct(grade, oracle_number, date)) %>%
-  mutate(year = as.numeric(format(date, '%Y'))) %>%
-  left_join(wages %>% dplyr::distinct(grade, department, year, daily_usd)) 
-final_data$group5 <- ifelse(is.na(final_data$daily_usd) & final_data$group != 'Temporary not field worker', 'No wage data', final_data$group)
-
-# final_model <- felm(log(absent+1) ~ protection_var + (rain_var) | oracle_number | 0 | 0,
-#                     data = final_data)
-final_model <- felm(log(absent+1) ~ protection +
-                      herd_protection + 
-                      rain_var | oracle_number | 0 | 0,
-                    data = final_data)
-summary(final_model)
-final_model_lm <- lm(log(absent+1) ~ protection +
-                       herd_protection + (rain_var),
-                     data = final_data)
-
 library(broom)
@@
 fake <- expand.grid(protection = seq(0, 1, 0.5),
-                    herd_protection = seq(0, 1, 0.1),
+                    herd_protection = seq(0, 0.6, 0.1),
                     rain_var = seq(0, 6, 1))
@@
 predictions <- data.frame(predictions)
-fixed_effects <- 0.067 # taken from below
+fixed_effects <- mean(fe$effect)
 fake$fit <- predictions$fit + fixed_effects
@@
        y = 'Estimated monthly absenteeism rate\n(Average fixed effect, 3 cm of 15-60 day prior precipitation)') + theme_maragra() 
+```
 
+# Model results (segregated models)
+
+```{r}
+out_list <- list()
+worker_groups <- names(model_list)
+for(i in 1:length(worker_groups)){
+  this_group <- worker_groups[i]
+  out_list[[i]] <- broom::tidy(model_list[[i]]) %>% mutate(grp = this_group)
+}
+out <- bind_rows(out_list)
+# Get the non segrated version too
+non_seg <- broom::tidy(final_model) %>% mutate(grp = 'Non segregated')
+options(scipen =  '999')
+out <- out %>% bind_rows(non_seg) %>% dplyr::select(term, estimate, p.value, grp) %>%
+  mutate(Type = 'Distance weighted')
+```
+
+# Radial model
+
+```{r}
+out_list <- list()
+worker_groups <- names(radial_model_list)
+for(i in 1:length(worker_groups)){
+  this_group <- worker_groups[i]
+  out_list[[i]] <- broom::tidy(radial_model_list[[i]]) %>% mutate(grp = this_group)
+}
+out2 <- bind_rows(out_list)
+# Get the non segrated version too
+non_seg <- broom::tidy(final_model_radial) %>% mutate(grp = 'Non segregated')
+options(scipen =  '999')
+out2 <- out2 %>% bind_rows(non_seg) %>% dplyr::select(term, estimate, p.value, grp) %>%
+  mutate(Type = 'Hard threshold')
+
+```
+
+# Combined radial and hard threshold
+
+```{r}
+combined <- bind_rows(out, out2)
+combined$term <- ifelse(combined$term == 'radial_herd_protection', 'herd_protection', combined$term)
+combined$term <- ifelse(combined$term == 'herd_protection', 'Community protection', combined$term)
+combined$term <- ifelse(combined$term == 'rain_var', 'Lagged precipitation', combined$term)
+combined$term <- ifelse(combined$term == 'protection', 'Individual protection', combined$term)
+
+combined$term <- Hmisc::capitalize(gsub('_', ' ', combined$term))
+ggplot(data = combined,
+       aes(x = term,
+           y = estimate,
+           fill = Type)) +
+  facet_wrap(~grp) +
+  geom_bar(stat = 'identity',
+           position = position_dodge()) +
+  coord_flip() +
+  geom_hline(yintercept = 0) +
+  theme_maragra() +
+  labs(x = 'Coefficient',
+       y = 'Term',
+       title = 'Comparison between distance-weighted vs. hard-threshold community protection estimation')
+```
+
+```{r}
+
+```
+
+```{r}
 # Also make predictions by groups
-worker_groups <- sort(unique(final_data$group5))
-model_data_list <- model_list <- model_list_lm <- list()
-
-for (i in 1:length(worker_groups)){
-  sub_data <- final_data %>% filter(group5 == worker_groups[i])
-  this_model <- felm(log(absent+1) ~ protection +
-                      herd_protection + 
-                      rain_var | oracle_number | 0 | 0,
-                    data = sub_data)
-  this_model_lm <- lm(log(absent+1) ~ protection +
-                       herd_protection + (rain_var),
-                     data = sub_data)
-  formatted <- broom::tidy(this_model)
-  formatted$group5 <- worker_groups[i]
-
-  model_list[[i]] <- this_model
-  model_list_lm[[i]] <- this_model_lm
-  model_data_list[[i]] <- formatted
-}
-names(model_list) <- names(model_data_list) <- names(model_list_lm) <- worker_groups
+worker_groups <- names(model_list)
+the_groups <- c(worker_groups, 'All')
 
@@
                     herd_protection = seq(0, 1, 0.1),
-                    rain_var = 3,# seq(0, 6, 1),
-                    group5 = worker_groups)
+                    rain_var = mean(final_data$rain_var, na.rm = TRUE),
+                    group = the_groups)
 fake_list <- list()
-for(i in 1:length(worker_groups)){
-  this_worker_group <- worker_groups[i]
+for(i in 1:length(the_groups)){
+  this_worker_group <- the_groups[i]
+  if(this_worker_group == 'All'){
+    this_model <- final_model
+    this_model_lm <- final_model_lm
+    this_fake_data <- fake %>% mutate(group = 'All')
+  } else {
       this_model <- model_list[[i]]
     this_model_lm <- model_list_lm[[i]]
-  this_fake_data <- fake %>% filter(group5 == this_worker_group)
+    this_fake_data <- fake %>% dplyr::filter(group == this_worker_group)
+  }
+
   predictions <- exp(predict(this_model_lm, newdata = this_fake_data, interval = 'confidence'))-1
@@
   geom_point() +
-  facet_wrap(~group5) +
+  facet_wrap(~group) +
   geom_line(alpha = 0.6) +
@@
        title = 'Association between community protection, individual protection, and absenteeism',
-       y = 'Estimated monthly absenteeism rate\n(Average fixed effect, 3 cm of 15-60 day prior precipitation)') +    theme_maragra()
+       y = 'Estimated monthly absenteeism rate') +    theme_maragra()
+
+```
+
+# Wage trade-offs
+
+```{r}
+# Share of avoided absences by worker type
+per_year <- predictions %>%
+  filter(strategy == 'Zero') %>%
+  mutate(year = as.numeric(format(date, '%Y'))) %>%
+  # mutate(year_month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+  group_by(date = year, 
+           permanent_or_temporary) %>%
+  summarise(n = n(),
+            actual_absences = sum(absent),
+            predicted_absences = sum(fit, na.rm = TRUE),
+            actual = mean(absent, na.rm = TRUE),
+            predicted = mean(fit, na.rm = TRUE)) %>%
+  mutate(avoided = predicted_absences - actual_absences)
+
+per_year = per_year %>% group_by(permanent_or_temporary) %>% summarise(avoided = mean(avoided))
+per_year$p <- per_year$avoided / sum(per_year$avoided)
+
+p_permanent <- per_year$p[per_year$permanent_or_temporary == 'Permanent']
+
+# Cost to avoid one absence
+cost <- 17.31
+temporary <- seq(2, 30, by = 2)
+permanent <- seq(2, 30, by = 2)
+fake_wage <- expand.grid(temporary = temporary,
+                         permanent = permanent)
+fake_wage$z <- ((fake_wage$temporary * (1- p_permanent)) + (fake_wage$permanent * p_permanent)) 
+fake_wage$f <- (fake_wage$z / cost * 100)-100
+# Add the percentage of absences prevented in each group
+
+# MAke chart of relation between wages of 2 and whether profitable or not
+
+cols <- colorRampPalette(c('darkred', 'white', 'blue'))(200)
+ggplot(data = fake_wage,
+       aes(x = temporary,
+           y = permanent,
+           fill = f)) +
+  geom_tile() +
+  scale_fill_gradientn(name = '',
+                   colors = cols) +
+  geom_text(aes(label = paste0(round(f, digits = 1), '%'))) +
+  labs(x = 'Average daily productivity value (wage): temporary workers',
+       y = 'Average daily productivity value (wage): permanent workers',
+       title = 'Return on investment of IRS program as a function of daily wages') +
+  theme_maragra() +
+  theme(legend.position = 'right')
+
+# Make for different ratios of temporary and permanent
+p_permanents <- c(0.15, 0.3, 0.45, 0.6, 0.75, 0.9)
+fake_wage_list <- list()
+for(i in 1:length(p_permanents)){
+  p_permanent <- p_permanents[i]
+    # Cost to avoid one absence
+  cost <- 17.31
+  temporary <- seq(2, 30, by = 2)
+  permanent <- seq(2, 30, by = 2)
+  fake_wage <- expand.grid(temporary = temporary,
+                           permanent = permanent)
+  fake_wage$z <- ((fake_wage$temporary * (1- p_permanent)) + (fake_wage$permanent * p_permanent)) 
+  fake_wage$f <- (fake_wage$z / cost * 100)-100
+  fake_wage_list[[i]] <- fake_wage %>% mutate(p = p_permanent)
+}
+
+fake_wage <- bind_rows(fake_wage_list)
+fake_wage$p <- paste0('Percent permanent: ', fake_wage$p * 100)
+
+ggplot(data = fake_wage,
+       aes(x = temporary,
+           y = permanent,
+           fill = f)) +
+  geom_tile() +
+  scale_fill_gradientn(name = '',
+                   colors = cols) +
+  # geom_text(aes(label = paste0(round(f, digits = 1), '%')), size = 2) +
+  labs(x = 'Average daily productivity value (wage): temporary workers',
+       y = 'Average daily productivity value (wage): permanent workers',
+       title = 'Return on investment of IRS program as a function of daily wages') +
+  theme_maragra() +
+  theme(legend.position = 'right') +
+  facet_wrap(~p)
+```
+
+# Simulations
+
+```{r}
+# 
+# # SIMULATIONS
+# sim_data <- simulations
+# sim_data$idx <- sim_data$oracle_number
+# fe <- getfe(final_model)
+# predictions <- exp(predict(final_model_lm, sim_data, interval = 'confidence'))-1
+# predictions <- data.frame(predictions)
+# # Add the fixed effects
+# fixed_effects <- left_join(sim_data,
+#                            data.frame(fe) %>% dplyr::select(idx, effect) %>%
+#                              mutate(idx = as.character(idx))) %>%
+#   .$effect
+# fixed_effects <- as.numeric(fixed_effects)
+# 
+# predictions$fit <- predictions$fit + fixed_effects
+# predictions$lwr <- predictions$lwr + fixed_effects
+# predictions$upr <- predictions$upr + fixed_effects
+# 
+# # combine other info
+# predictions <- predictions %>% bind_cols(sim_data %>% dplyr::select(date, absent, strategy, permanent_or_temporary)) %>%
+#   mutate(absent = as.numeric(absent)) %>%
+#   filter(!is.na(fit))
+# 
+# simple <- predictions %>%
+#   group_by(strategy) %>%
+#   summarise(n = n(),
+#             actual_absences = sum(absent),
+#             predicted_absences = sum(fit, na.rm = TRUE),
+#             actual = mean(absent, na.rm = TRUE),
+#             predicted = mean(fit, na.rm = TRUE)) %>%
+#   dplyr::rename(value = predicted) %>%
+#   dplyr::select(strategy, value, predicted_absences)
+# 
+# zero <- predictions %>%
+#   filter(strategy == 'Zero') %>%
+#   mutate(year = as.numeric(format(date, '%Y'))) %>%
+#   group_by(year) %>%
+#   summarise(n = n(),
+#             actual_absences = sum(absent),
+#             predicted_absences = sum(fit, na.rm = TRUE),
+#             actual = mean(absent, na.rm = TRUE),
+#             predicted = mean(fit, na.rm = TRUE)) %>%
+#   mutate(avoided = predicted_absences - actual_absences)
+# # dplyr::rename(value = predicted) %>%
+# # dplyr::select(permanent_or_temporary, value, predicted_absences)
+# 
+# kable(simple)
+# 
+# # Eligible working days
+# sum(zero$n)
+# # Absences
+# sum(zero$actual_absences)
+# # Actual absence rate
+# sum(zero$actual_absences) / sum(zero$n)
+# # Would have observed
+# sum(zero$predicted_absences)
+# # Predicted - observ
+# sum(zero$predicted_absences) - sum(zero$actual_absences)
+#   # Overall reduction from
+# sum(zero$predicted_absences) / sum(zero$n)
+# # To
+# sum(zero$actual_absences) / sum(zero$n)
+# # Number of working days permanent vs temp
+# table(predictions$permanent_or_temporary[predictions$strategy=='Real'])
+# # Absences avoided among permanent workers
+# predictions %>% filter(strategy == 'Zero') %>% group_by(permanent_or_temporary) %>% summarise(n = n(),
+#                                                                                               actual_absences = sum(absent),
+#                                                                                               predicted_absences = sum(fit, na.rm = TRUE),
+#                                                                                               actual = mean(absent, na.rm = TRUE),
+#                                                                                               predicted = mean(fit, na.rm = TRUE)) %>%
+#   mutate(avoided = predicted_absences - actual_absences) %>%
+#   # decline due to irs
+#   mutate(decline = predicted - actual)
+# # One absence every n days
+# sum(zero$n) / sum(zero$avoided)
+# # Annual reduction in absence
+# mean(zero$avoided[zero$year < 2016])
+# # Cost per absence avoided
+# 68984 / 4183
+# # Percentage of absence among permanent workers
+# x = predictions %>%
+#   filter(strategy == 'Zero') %>%
+#   # filter(date < '2016-01-01') %>%
+#   group_by(permanent_or_temporary) %>%
+#   summarise(absences = sum(absent),
+#             predicted = sum(fit)) %>%
+#   mutate(averted = predicted - absences) %>%
+#   mutate(p = averted / sum(averted))
+# x
+# # Weighted average cost
+# y = sum(c(25, 5) * x$p)
+# y
+# # Cost of treatment per absence
+# 22 / 80
+# # Total savings per aversion
+# 0.275 + y
+# # Total yearly savings
+# (0.275 + y) * 4183
+# # Total yearly savings
+# ((0.275 + y) * 4183) - 68984
+# # ROI
+# ((0.275 + y) * 4183) / 68984
+# # Additional max strategy
+# simple
+# 47943-42131
+# 
+# ggplot(data = simple,
+#        aes(x = strategy,
+#            y = value)) +
+#   geom_bar(stat = 'identity',
+#            alpha = 0.6,
+#            color = 'black') +
+#   labs(y = 'Absenteeism rate',
+#        x = 'Strategy') +
+#   geom_label(aes(label = round(value, digits = 3)),
+#              nudge_y = -0.005)
+```
+
+```{r}
+# SIMULATIONS WITH WORKER SPECIFIC TYPE
 
 # SIMULATIONS
-sim_data <- simulations
+sim_data <- simulations# %>%
+  # # Get grade
+  # left_join(ab_panel %>% dplyr::distinct(grade, oracle_number, date)) %>%
+  # mutate(year = as.numeric(format(date, '%Y'))) %>%
+  # left_join(wages %>% dplyr::distinct(grade, department, year, daily_usd)) 
+# sim_data$group5 <- ifelse(is.na(sim_data$daily_usd) & sim_data$group != 'Temporary not field worker', 'No wage data', sim_data$group)
 sim_data$idx <- sim_data$oracle_number
-fe <- getfe(final_model)
-predictions <- exp(predict(final_model_lm, sim_data, interval = 'confidence'))-1
+
+worker_groups <- names(model_list)
+sim_list <- list()
+radial_sim_list <- list()
+for(i in 1:length(worker_groups)){
+  message(i)
+  this_worker_group <- worker_groups[i]
+
+  this_model <- model_list[[i]]
+  this_model_lm <- model_list_lm[[i]]
+  this_fake_data <- sim_data %>% filter(group == this_worker_group)
+
+  # this_radial_model <- radial_model_list[[i]]
+  # this_radial_model_lm <- radial_model_list_lm[[i]]
+  # this_radial_fake_data <- radial_model_data_list[[i]]
+
+  predictions <- exp(predict(this_model_lm, newdata = this_fake_data, interval = 'confidence'))-1
   predictions <- data.frame(predictions)
-# Add the fixed effects
-fixed_effects <- left_join(sim_data,
-                           data.frame(fe) %>% dplyr::select(idx, effect) %>%
-                             mutate(idx = as.character(idx))) %>%
-  .$effect
-fixed_effects <- as.numeric(fixed_effects)
 
-predictions$fit <- predictions$fit + fixed_effects
-predictions$lwr <- predictions$lwr + fixed_effects
-predictions$upr <- predictions$upr + fixed_effects
+  # radial_predictions <- exp(predict(this_radial_model_lm, newdata = this_radial_fake_data, interval = 'confidence'))-1
+  # radial_predictions <- data.frame(radial_predictions)
+
+  this_fe <- getfe(this_model)
+  this_radial_fe <- getfe(this_radial_model)
+
+  # fixed_effects <- mean(this_fe$effect)
+  # radial_fixed_effects <- mean(this_radial_fe$effect)
+
+  out <- this_fake_data
+  # radial_out <- this_radial_fake_data
+
+  out$fit <- predictions$fit + fixed_effects
+  out$lwr <- predictions$lwr + fixed_effects
+  out$upr <- predictions$upr + fixed_effects
+  sim_list[[i]] <- out
+
+  # radial_out$fit <- radial_predictions$fit + fixed_effects
+  # radial_out$lwr <- radial_predictions$lwr + fixed_effects
+  # radial_out$upr <- radial_predictions$upr + fixed_effects
+  # radial_sim_list[[i]] <- radial_out
+}
+sim_data <- bind_rows(sim_list)
+# radial_sim_data <- bind_rows(radial_sim_list)
 
 # combine other info
-predictions <- predictions %>% bind_cols(sim_data %>% dplyr::select(date, absent, strategy, permanent_or_temporary)) %>%
+predictions <- sim_data %>%
   mutate(absent = as.numeric(absent)) %>%
@@
             actual_absences = sum(absent),
-            predicted_absences = sum(fit, na.rm = TRUE),
-            actual = mean(absent, na.rm = TRUE),
-            predicted = mean(fit, na.rm = TRUE)) %>%
-  dplyr::rename(value = predicted) %>%
-  dplyr::select(strategy, value, predicted_absences)
+            predicted_absences = sum(fit, na.rm = TRUE)
+            # actual = mean(absent, na.rm = TRUE),
+            # predicted = mean(fit, na.rm = TRUE)
+            ) %>%
+  mutate(actual = actual_absences / n,
+         predicted = predicted_absences / n)
 
@@
   mutate(avoided = predicted_absences - actual_absences)
-# dplyr::rename(value = predicted) %>%
-# dplyr::select(permanent_or_temporary, value, predicted_absences)
+
+# Scenario splot
+pd <- simple
+ggplot(data = pd,
+       aes(x = strategy,
+           y = value)) +
+  geom_bar(stat = 'identity') +
+  theme_maragra() +
+  geom_text(aes(label = round(value * 100, digits = 2)), color = 'white',
+            nudge_y = -0.01, 
+            size = 7) +
+  labs(x = 'Strategy',
+       y = )
+
+pd <- predictions %>%
+  filter(strategy == 'Zero') %>%
+  mutate(year = as.numeric(format(date, '%Y'))) %>%
+  # mutate(year_month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+  group_by(date = year, 
+           permanent_or_temporary) %>%
+  summarise(n = n(),
+            actual_absences = sum(absent),
+            predicted_absences = sum(fit, na.rm = TRUE),
+            actual = mean(absent, na.rm = TRUE),
+            predicted = mean(fit, na.rm = TRUE)) %>%
+  mutate(avoided = predicted_absences - actual_absences)
+
+pdx <- pd %>%
+  dplyr::select(date, permanent_or_temporary,
+                actual, predicted) %>%
+  gather(key, value, actual:predicted) %>%
+  mutate(key = ifelse(key == 'actual', 'Observed', 
+                      'Counterfactual'))
+ggplot(data = pdx,
+       aes(x = date,
+           y = value,
+            lty = key,
+           color = permanent_or_temporary,
+           pch = key)) +
+  geom_point() +
+  geom_line() +
+  theme_maragra() +
+  scale_color_manual(name = '',
+                     values = c('red', 'blue')) +
+  scale_linetype_manual(name = '',
+                        values = c(1,2)) +
+  scale_shape_manual(name = '',
+                     values = c(1,2)) +
+  labs(x = 'Year',
+       y = 'Absenteeism',
+       title = 'Predicted vs. observed absenteeism')
 
@@
 
-ggplot(data = simple,
-       aes(x = strategy,
-           y = value)) +
-  geom_bar(stat = 'identity',
-           alpha = 0.6,
-           color = 'black') +
-  labs(y = 'Absenteeism rate',
-       x = 'Strategy') +
-  geom_label(aes(label = round(value, digits = 3)),
-             nudge_y = -0.005)
+# Per year
+per_year <- predictions %>%
+  filter(strategy == 'Zero') %>%
+  mutate(year = as.numeric(format(date, '%Y'))) %>%
+  # mutate(year_month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
+  group_by(date = year, 
+           permanent_or_temporary) %>%
+  summarise(n = n(),
+            actual_absences = sum(absent),
+            predicted_absences = sum(fit, na.rm = TRUE),
+            actual = mean(absent, na.rm = TRUE),
+            predicted = mean(fit, na.rm = TRUE)) %>%
+  mutate(avoided = predicted_absences - actual_absences)
 
+per_year = per_year %>% group_by(permanent_or_temporary) %>% summarise(avoided = mean(avoided))
+sum(per_year$avoided)
+```
+
+# Table : Absenteeism rate by year and worker type
+
+```{r}
+pd <- final_data %>%
+  group_by(permanent_or_temporary, year) %>% 
+  summarise(absences = length(which(absent)),
+            eligibles = n()) %>%
+  ungroup %>%
+  mutate(p = absences / eligibles * 100)
+pd
+```
+
+#
+
+```{r}
 # paste0(
@@
 ```
+
+## Ad hoc analysis of seasonality and absenteeism by worker group
+
+```{r}
+pd <- final_data %>%
+  group_by(group5, month_number) %>%
+  summarise(absences = length(which(absent)),
+            eligibles = n()) %>%
+  ungroup %>%
+  mutate(p = absences / eligibles * 100)
+
+ggplot(data = pd,
+       aes(x = month_number,
+           y = p,
+           group = group5,
+           color = group5)) +
+  geom_point() +
+  facet_wrap(~group5) +
+  theme_maragra() +
+  geom_smooth() +
+  theme(legend.position = 'none') +
+  labs(x = 'Month of year',
+       y = 'Absenteeism rate',
+       title = 'Worker type and monthly absenteeism')
+
+pd <- final_data %>%
+  group_by(grade, month_number) %>%
+  summarise(absences = length(which(absent)),
+            eligibles = n()) %>%
+  ungroup %>%
+  mutate(p = absences / eligibles * 100)
+
+ggplot(data = pd,
+       aes(x = month_number,
+           y = p,
+           group = grade,
+           color = grade)) +
+  geom_point() +
+  facet_wrap(~grade) +
+  theme_maragra() +
+  geom_smooth() +
+  theme(legend.position = 'none') +
+  labs(x = 'Month of year',
+       y = 'Absenteeism rate',
+       title = 'Worker type and monthly absenteeism')
+```
+
+## Ad hoc analysis of IRS and absenteeism by worker group
+
+Individual level
+
+```{r}
+pd <- final_data %>%
+  group_by(protection = round(protection,2), group5) %>%
+    summarise(absences = length(which(absent)),
+            eligibles = n()) %>%
+  ungroup %>%
+  mutate(p = absences / eligibles * 100)
+
+ggplot(data = pd,
+       aes(x = protection,
+           y = p,
+           group = group5,
+           color = group5)) +
+  geom_point(alpha = 0.7, aes(size = eligibles)) +
+  facet_wrap(~group5) +
+  theme_maragra() +
+  geom_smooth(method='lm', formula= y~x) +
+  geom_smooth(se = FALSE, size = 0.5, alpha = 0.6) +
+  theme(legend.position = 'none') +
+  labs(x = 'Estimated INDIVIDUAL protection level',
+       y = 'Absenteeism rate',
+       title = 'Absenteeism as a function of time since IRS at individual level')
+```
+
+```{r}
+pd <- final_data %>%
+  group_by(protection = herd_protection, group5) %>%
+    summarise(absences = length(which(absent)),
+            eligibles = n()) %>%
+  ungroup %>%
+  mutate(p = absences / eligibles * 100)
+
+ggplot(data = pd,
+       aes(x = protection,
+           y = p,
+           group = group5,
+           color = group5)) +
+  # geom_point(alpha = 0.7, size = 0.3) +
+  facet_wrap(~group5) +
+  theme_maragra() +
+  geom_smooth(method='lm', formula= y~x) +
+  # geom_smooth(se = FALSE, size = 0.5, alpha = 0.6) +
+  theme(legend.position = 'none') +
+  labs(x = 'Estimated COMMUNITY protection level',
+       y = 'Absenteeism rate',
+       title = 'Absenteeism as a function of time since IRS at community level')
+```
+
+# Fumigations over time
+
+```{r}
+pd <- mc %>%
+  group_by(date = as.Date(paste0(year, '-', month, '-01'))) %>%
+  tally
+
+ggplot(data = pd,
+       aes(x = date,
+           y = n)) +
+  geom_bar(stat = 'identity')
+```

### 2020-06-21 inst/rmd/apr2020/code.R

@@ -11,113 +11,11 @@ library(RColorBrewer)
 library(lfe)
 library(restorepoint)
 library(regtools)
-# 
-# # # Elisa: active workers by worker type over time
-# pd <- maragra::ab_panel %>%
-#   left_join(
-#   maragra::workers %>% dplyr::select(oracle_number, permanent_or_temporary)) %>%
-#   mutate(type = paste0(permanent_or_temporary, ' ', department)) %>%
-#   mutate(year_month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
-#   group_by(type, year_month) %>%
-#   tally
-# 
-# ggplot(data = pd,
-#        aes(x = year_month,
-#            y = n,
-#            fill = type,
-#            group = type)) +
-#   geom_bar(stat = 'identity',
-#            position = position_dodge())
-# write_csv(pd, '~/Desktop/pd.csv')
-# # 
-# pd <- maragra::ab_panel %>%
-#   left_join(
-#     maragra::workers %>% dplyr::select(oracle_number, permanent_or_temporary)) %>%
-#   mutate(type = paste0(permanent_or_temporary)) %>%
-#   mutate(year_month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
-#   group_by(type, year_month) %>%
-#   tally
-# 
-# ggplot(data = pd,
-#        aes(x = year_month,
-#            y = n,
-#            fill = type,
-#            color = type)) +
-#   geom_line() +
-#   geom_point() +
-#   xlim(as.Date('2013-01-01'),
-#        as.Date('2016-06-01'))
-# 
-# pd <- maragra::ab_panel %>%
-#   left_join(
-#     maragra::workers %>% dplyr::select(oracle_number, permanent_or_temporary, sex)) %>%
-#   mutate(type = paste0(permanent_or_temporary, ' ', sex)) %>%
-#   mutate(year_month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
-#   group_by(type, year_month) %>%
-#   tally
-# 
-# pd$type <- factor(pd$type, levels = rev(unique(pd$type)))
-# ggplot(data = pd,
-#        aes(x = year_month,
-#            y = n,
-#            fill = type,
-#            color = type)) +
-#   geom_area(position = 'stack') +
-#   xlim(as.Date('2013-01-01'),
-#        as.Date('2016-01-31')) +
[…]

### 2020-08-09 inst/rmd/aug2020/aug2020.Rmd
updated

---
title: "Final analysis for paper"
date: ''
fig_height: 2.6
output:
  pdf_document:
    includes:
      in_header: preamble.sty
    keep_tex: yes
    latex_engine: xelatex
  html_document:
    df_print: paged
fig_width: 4
always_allow_html: true
subtitle: Brew, Pradhan, Sicuri
bibliography: bibliography.bib
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# csl: journal-of-health-economics.csl

# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(RColorBrewer)
library(extrafont)
library(kableExtra)
library(raster)
library(sp)
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(knitr)
library(RColorBrewer)
# library(lme4)
library(lfe)
library(restorepoint)
library(regtools)

loadfonts()
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               fig.height = 4,
               fig.width = 5,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               # dev = "cairo_pdf",
               fig.pos="!h",
               fig.align = 'center')
knitr::opts_chunk$set(error = FALSE)

opts_knit$set(width=75)
options(xtable.comment = FALSE)
source('code.R')
```

*Absences*: We observed `r nrow(ab_panel)` unique worker-days among `r length(unique(ab_panel$oracle_number))` workers. The all-period average absenteeism rate was `r round(length(which(ab_panel$absent)) / nrow(ab_panel) * 100, digits = 2)`%, though this rate varied widely as a function of worker department, sex, residence, and season (table).

# Model results (one model)

```{r}
library(broom)
broom::tidy(final_model)

# Prediction based on individual stuff
fake <- expand.grid(protection_var = seq(0, 0.7, by = 0.1),
                                         rain_var = seq(0, 6, 1))
fe <- getfe(final_model)
predictions <- exp(predict(final_model_lm, newdata = fake, interval = 'confidence'))-1
predictions <- data.frame(predictions)
fixed_effects <- mean(fe$effect)
fake$fit <- predictions$fit + fixed_effects
fake$lwr <- predictions$lwr + fixed_effects
fake$upr <- predictions$upr + fixed_effects

n_cols <- length(unique(fake$protection))
cols <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, 'Spectral'))(n_cols)
cols[2] <- 'black'
library(maragra)
ggplot(data = fake %>% filter(rain_var == 3) %>% mutate(protection_var = factor(protection_var)),
       aes(x = protection_var,
           y = fit)) +
  geom_ribbon(aes(ymin = lwr, ymax = upr),
              color = NA,
              alpha = 0.15) +
  geom_point() +
  geom_line(alpha = 0.6) +
  labs(x = 'Protection score',
       title = 'Association between community protection, individual protection, and absenteeism',
       y = 'Estimated monthly absenteeism rate\n(Average fixed effect, 3 cm of 15-60 day prior precipitation)') + theme_maragra() 
```

# Model results (segregated models)

```{r}
out_list <- list()
worker_groups <- names(model_list)
for(i in 1:length(worker_groups)){
  this_group <- worker_groups[i]
  out_list[[i]] <- broom::tidy(model_list[[i]]) %>% mutate(grp = this_group)
}
out <- bind_rows(out_list)
# Get the non segrated version too
non_seg <- broom::tidy(final_model) %>% mutate(grp = 'Non segregated')
options(scipen =  '999')
out <- out %>% bind_rows(non_seg) %>% dplyr::select(term, estimate, p.value, grp) %>%
  mutate(Type = 'Distance weighted')
out

out %>% mutate(estimate = paste0(round(estimate, digits = 3),
                                 ' (',
               round(p.value, digits = 3), ')')) %>% dplyr::select(-p.value, -Type) %>%  tidyr::spread(key = grp, value = estimate)
```

```{r}
# Also make predictions by groups
worker_groups <- names(model_list)
the_groups <- c(worker_groups, 'All')

# Do predictions on the different groups
fake <- expand.grid(protection_var = seq(0, 1, by = 0.1),
                    rain_var = mean(final_data$rain_var, na.rm = TRUE),
                    group = the_groups)
fake_list <- list()
for(i in 1:length(the_groups)){
  this_worker_group <- the_groups[i]
  if(this_worker_group == 'All'){
    this_model <- final_model
    this_model_lm <- final_model_lm
    this_fake_data <- fake %>% mutate(group = 'All')
  } else {
      this_model <- model_list[[i]]
    this_model_lm <- model_list_lm[[i]]
    this_fake_data <- fake %>% dplyr::filter(group == this_worker_group)
  }

  predictions <- exp(predict(this_model_lm, newdata = this_fake_data, interval = 'confidence'))-1
  predictions <- data.frame(predictions)
  this_fe <- getfe(this_model)
  fixed_effects <- mean(this_fe$effect)
  out <- this_fake_data
  out$fit <- predictions$fit + fixed_effects
  out$lwr <- predictions$lwr + fixed_effects
  out$upr <- predictions$upr + fixed_effects
  fake_list[[i]] <- out
}
fake <- bind_rows(fake_list)

n_cols <- length(unique(fake$protection))
cols <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, 'Spectral'))(n_cols)
cols[2] <- 'black'
library(maragra)
ggplot(data = fake  %>% mutate(protection = factor(protection_var)),
       aes(x = protection_var,
           y = fit)) +
  # geom_ribbon(aes(ymin = lwr, ymax = upr,
  #                 fill = protection),
  #             color = NA,
  #             alpha = 0.15) +
  geom_point() +
  facet_wrap(~group) +
  geom_line(alpha = 0.6) +
  labs(x = 'Protection score',
       title = 'Association between protection score and absenteeism',
       y = 'Estimated monthly absenteeism rate') +    
  theme_maragra()

```

# Simulations

```{r}

# SIMULATIONS
sim_data <- simulations
sim_data$idx <- sim_data$oracle_number
fe <- getfe(final_model)
predictions <- exp(predict(final_model_lm, sim_data, interval = 'confidence'))-1
predictions <- data.frame(predictions)
# Add the fixed effects
fixed_effects <- left_join(sim_data,
                           data.frame(fe) %>% dplyr::select(idx, effect) %>%
                             mutate(idx = as.character(idx))) %>%
  .$effect
fixed_effects <- as.numeric(fixed_effects)

predictions$fit <- predictions$fit + fixed_effects
predictions$lwr <- predictions$lwr + fixed_effects
predictions$upr <- predictions$upr + fixed_effects

# combine other info
predictions <- predictions %>% bind_cols(sim_data %>% dplyr::select(date, absent, strategy, permanent_or_temporary)) %>%
  mutate(absent = as.numeric(absent)) %>%
  filter(!is.na(fit))

simple <- predictions %>%
  group_by(strategy) %>%
  summarise(n = n(),
            actual_absences = sum(absent),
            predicted_absences = sum(fit, na.rm = TRUE),
            actual = mean(absent, na.rm = TRUE),
            predicted = mean(fit, na.rm = TRUE))# %>%
  # dplyr::rename(value = predicted) %>%
  # dplyr::select(strategy, value, predicted_absences)

zero <- predictions %>%
  filter(strategy == 'Zero') %>%
  mutate(year = as.numeric(format(date, '%Y'))) %>%
  group_by(year) %>%
  summarise(n = n(),
            actual_absences = sum(absent),
            predicted_absences = sum(fit, na.rm = TRUE),
            actual = mean(absent, na.rm = TRUE),
            predicted = mean(fit, na.rm = TRUE)) %>%
  mutate(avoided = predicted_absences - actual_absences)
# dplyr::rename(value = predicted) %>%
# dplyr::select(permanent_or_temporary, value, predicted_absences)

simple
# kable(simple)

# Eligible working days
sum(zero$n)
# Absences
sum(zero$actual_absences)
# Actual absence rate
sum(zero$actual_absences) / sum(zero$n)
# Would have observed
sum(zero$predicted_absences)
# Predicted - observ
sum(zero$predicted_absences) - sum(zero$actual_absences)
  # Overall reduction from
sum(zero$predicted_absences) / sum(zero$n)
# To
sum(zero$actual_absences) / sum(zero$n)
# Number of working days permanent vs temp
table(predictions$permanent_or_temporary[predictions$strategy=='Real'])
# Absences avoided among permanent workers
x = predictions %>% filter(strategy == 'Zero') %>% group_by(permanent_or_temporary) %>% summarise(n = n(),
                                                                                              actual_absences = sum(absent),
                                                                                              predicted_absences = sum(fit, na.rm = TRUE),
                                                                                              actual = mean(absent, na.rm = TRUE),
                                                                                              predicted = mean(fit, na.rm = TRUE)) %>%
  mutate(avoided = predicted_absences - actual_absences) %>%
  # decline due to irs
  mutate(decline = predicted - actual)
x
# Weight temporary twice
weighted <- x$
# One absence every n days
sum(zero$n) / sum(zero$avoided)
# Annual reduction in absence
mean(zero$avoided[zero$year < 2016])
# Cost per absence avoided
68984 / 4183
# Percentage of absence among permanent workers
x = predictions %>%
  filter(strategy == 'Zero') %>%
  # filter(date < '2016-01-01') %>%
  group_by(permanent_or_temporary) %>%
  summarise(absences = sum(absent),
            predicted = sum(fit)) %>%
  mutate(averted = predicted - absences) %>%
  mutate(p = averted / sum(averted))
x
# Weighted average cost
y = sum(c(25, 5) * x$p)
y
# Cost of treatment per absence
22 / 80
# Total savings per aversion
0.275 + y
# Total yearly savings
(0.275 + y) * 4183
# Total yearly savings
((0.275 + y) * 4183) - 68984
# ROI
((0.275 + y) * 4183) / 68984
# Additional max strategy
simple
47943-42131

ggplot(data = simple,
       aes(x = strategy,
           y = value)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(y = 'Absenteeism rate',
       x = 'Strategy') +
  geom_label(aes(label = round(value, digits = 3)),
             nudge_y = -0.005)
```

# Descriptive statistics

```{r}
table(workers$sex)
x = model_data %>% dplyr::distinct(oracle_number, .keep_all=T); prop.table(table(x$sex))
prop.table(table(x$field))
x = left_join(x, workers %>% dplyr::select(oracle_number, date_of_birth)) %>% mutate(age = as.numeric(date - date_of_birth) / 365.25)
hist(x$age)
prop.table(table(x$age >= 20 & x$age < 40))
prop.table(table(x$permanent_or_temporary))
```

# Table 1

```{r}
library(table1)
message('We observed ', nrow(model_data), ' worker-days among ', length(unique(model_data$oracle_number)), ' unique workers. Overall absence rate was ', length(which(model_data$absent)) / nrow(model_data) * 100)

# Table of those absences
vars <- list('field',
          'permanent_or_temporary',
           'sex')
table1::label(model_data$field) <- 'Worker location'
table1::label(model_data$permanent_or_temporary) <- 'Worker contract'
table1::label(model_data$sex) <- 'Worker sex'
model_data$year <- as.numeric(format(model_data$date, '%Y'))
table1::label(model_data$year) <- 'Year'

table1::table1(~field + permanent_or_temporary + sex | year , data = model_data)
```

# Crude effect on absences chart

```{r}
pd <- model_data %>%
  mutate(days_since = round(days_since / 30)) %>%
  group_by(days_since) %>%
  summarise(absences = length(which(absent)),
            eligibles = n()) %>%
  ungroup %>%
  mutate(p = absences / eligibles * 100) %>%
  filter(days_since >= -10,
         days_since <= 10)

g1 <- ggplot(data = pd,
       aes(x = days_since,
           y = p)) +
  geom_step(color = 'darkred', alpha = 0.8, size = 1) +
  theme_maragra() +
  labs(x = 'Months relative to administration of IRS',
       y = 'Absenteeism rate (%)',
       title = 'A') +
  geom_vline(xintercept = 0, lty = 2, alpha = 0.7)
pd2 <- model_data %>%
  mutate(days_since = round(days_since / 30)) %>%
  mutate(season = ifelse(rain_var >= median(rain_var, na.rm = TRUE),
                         'Rainy season',
                         'Dry season')) %>%
  group_by(days_since, season) %>%
  summarise(absences = length(which(absent)),
            eligibles = n()) %>%
  ungroup %>%
  mutate(p = absences / eligibles * 100) %>%
  filter(days_since >= -10,
         days_since <= 10,
         !is.na(season))
g2 <- ggplot(data = pd2,
       aes(x = days_since,
           y = p)) +
  facet_wrap(~season, ncol = 1) +
  geom_step(color = 'darkred', alpha = 0.8, size = 1) +
  theme_maragra() +
  labs(x = 'Months relative to administration of IRS',
       y = 'Absenteeism rate (%)',
       title = 'B') +
  geom_vline(xintercept = 0, lty = 2, alpha = 0.7)

Rmisc::multiplot(g1, g2, cols = 2)
```

# Absences

### 2020-08-09 inst/rmd/aug2020/code.R

@@ -1,0 +1,534 @@
+# Packages
+library(sp)
+library(tidyverse)
+library(knitr)
+library(Hmisc)
+library(brew)
+library(maragra)
+library(knitr)
+library(RColorBrewer)
+# library(lme4)
+library(lfe)
+library(restorepoint)
+library(regtools)
+library(tidyverse)
+library(raster)
+library(tidyr)
+library(ggplot2)
+library(RColorBrewer)
+library(broom)
+library(ggthemes)
+
+classify_protection <- function(days_since){
+  ifelse(days_since < 21 | days_since > (365+21),0,
+         (((365+21) - days_since)/(365))
+  )
+}
+
+if('prepared_data.RData' %in% dir()){
+  load('prepared_data.RData')
+} else {
+  # Read in data
+  ab <- maragra::ab
+  ab_panel <- maragra::ab_panel
+  bairros <- maragra::bairros
+  bes <- maragra::bes
+  census <- maragra::census
+  clinic <- maragra::clinic
+  clinic_agg <- maragra::clinic_agg
+  mc <- maragra::mc
+  workers <- maragra::workers
+  
+  if(!'weather.RData' %in% dir()){
+    weather <- maragra::weather %>%
+      mutate(precipitation = ifelse(is.na(precipitation),
+                                    0,
+                                    precipitation))
+    
+    # Get lags for weather
+    starts <- c(15, 30, 60, 90)
+    ends <- c(30, 60, 90, 120)
+    for(k in 1:length(starts)){
+      for(j in 1:length(ends)){
+        this_start <- starts[k]
+        this_end <- ends[j]
+        if(this_start < this_end){
+          precipitation_column_name <- paste0('precipitation_lag_', this_start, '_', this_end)
+          weather[,precipitation_column_name] <- NA
+          temp_column_name <- paste0('temp_lag_', this_start, '_', this_end)
+        }
+      }
+    }
+    for(i in 1:nrow(weather)){
+      starts <- c(15, 30, 60, 90)
+      ends <- c(30, 60, 90, 120)
+      for(k in 1:length(starts)){
+        for(j in 1:length(ends)){
+          this_start <- starts[k]
+          this_end <- ends[j]
+          if(this_start < this_end){
+            this_date <- weather$date[i]
+            end_date <- this_date - this_start
+            start_date <- this_date - this_end
+            if(start_date >= min(weather$date,na.rm = TRUE) &
+               end_date <= max(weather$date, na.rm = TRUE)){
[…]

### 2020-08-09 inst/rmd/aug2020/code.R
checkpoint

@@ -384,17 +384,17 @@ if('simulations.RData' %in% dir()){
   load('simulations.RData')
 } else {
   
-  # time_optimize_df <-
-  #   data.frame(date = seq(as.Date('2020-01-01'),
-  #                         as.Date('2020-12-31'), 
-  #                         by = 1)) %>%
-  #   mutate(month_day = format(date, '%m-%d'))
-  # time_optimize_df$day_number <- (1:nrow(time_optimize_df))+20
-  # time_optimize_df$the_protection <- classify_protection(time_optimize_df$day_number)
-  # time_optimize_df <- time_optimize_df %>% dplyr::select(month_day, the_protection)
+  time_optimize_df <-
+    data.frame(date = seq(as.Date('2020-01-01'),
+                          as.Date('2020-12-31'),
+                          by = 1)) %>%
+    mutate(month_day = format(date, '%m-%d'))
+  time_optimize_df$day_number <- (1:nrow(time_optimize_df))+20
+  time_optimize_df$the_protection <- classify_protection(time_optimize_df$day_number)
+  time_optimize_df <- time_optimize_df %>% dplyr::select(month_day, the_protection)
   
 
-  strategies <- c('Zero',  'Gold standard')
+  strategies <- c('Zero',  'Gold standard', 'Time-optimized')
   strategy_list <- list()
   for(s in 1:length(strategies)){
     dates <- sort(unique(final_data$date))
@@ -417,6 +417,15 @@ if('simulations.RData' %in% dir()){
         this_data <- gold_data %>%
           filter(date == this_date)
         the_protection <- this_data$protection
+      }
+      if(this_strategy == 'Time-optimized'){
+        this_data <- final_data %>%
+          filter(date == this_date)
+        this_data$month_day <- format(this_data$date, '%m-%d')
+        x <- this_data %>%
+          left_join(time_optimize_df)
+        the_protection <- x$the_protection
+        this_data$protection <- the_protection
       }
       # if(this_strategy == 'Time-optimized'){
       #   this_data <- final_data %>%

### 2020-08-10 inst/rmd/aug2020/aug2020.Rmd
implemented new modeling

@@
 
+# TO DO: MAKE PREDICTIONS BASED ON INDIVIDUAL MODEL TYPES
+
 # SIMULATIONS
-sim_data <- simulations
+sim_data <- simulations  %>% filter(on_site)
 sim_data$idx <- sim_data$oracle_number
-fe <- getfe(final_model)
-predictions <- exp(predict(final_model_lm, sim_data, interval = 'confidence'))-1
-predictions <- data.frame(predictions)
-# Add the fixed effects
-fixed_effects <- left_join(sim_data,
-                           data.frame(fe) %>% dplyr::select(idx, effect) %>%
-                             mutate(idx = as.character(idx))) %>%
-  .$effect
-fixed_effects <- as.numeric(fixed_effects)
+sim_data$fit <- NA
+# Multi model predictions
+for(i in 1:length(worker_groups)){
+  this_worker_group <- worker_groups[i]
+  indices <- which(sim_data$group == this_worker_group)
+  this_data <- sim_data[indices,]
+  this_model_fix <- model_list_fix[[this_worker_group]]
+  these_predictions <- exp(predict(this_model_fix, this_data))-1
+  sim_data$fit[indices] <- these_predictions
+}
+predictions <- sim_data %>% filter(!is.na(fit))
 
-predictions$fit <- predictions$fit + fixed_effects
-predictions$lwr <- predictions$lwr + fixed_effects
-predictions$upr <- predictions$upr + fixed_effects
-
-# combine other info
-predictions <- predictions %>% bind_cols(sim_data %>% dplyr::select(date, absent, strategy, permanent_or_temporary)) %>%
-  mutate(absent = as.numeric(absent)) %>%
-  filter(!is.na(fit))
+# # One model predictions
+# fe <- getfe(final_model)
+# predictions <- exp(predict(final_model_lm, sim_data, interval = 'confidence'))-1
+# predictions <- data.frame(predictions)
+# # Add the fixed effects
+# fixed_effects <- left_join(sim_data,
+#                            data.frame(fe) %>% dplyr::select(idx, effect) %>%
+#                              mutate(idx = as.character(idx))) %>%
+#   .$effect
+# fixed_effects <- as.numeric(fixed_effects)
+# 
+# predictions$fit <- predictions$fit + fixed_effects
+# predictions$lwr <- predictions$lwr + fixed_effects
+# predictions$upr <- predictions$upr + fixed_effects
+# 
+# # combine other info
+# predictions <- predictions %>% bind_cols(sim_data %>% dplyr::select(date, absent, strategy, permanent_or_temporary)) %>%
+#   mutate(absent = as.numeric(absent)) %>%
+#   filter(!is.na(fit))
 

### 2020-08-10 inst/rmd/aug2020/code.R

@@ -482,8 +482,10 @@ final_data$group <- final_data$group3
 simulations$group <- simulations$group3
 
 
-# final_model <- felm(log(absent+1) ~ protection_var + (rain_var) | oracle_number | 0 | 0,
-#                     data = final_data)
+library(fixest)
+final_model_fix <- feols(data = model_data, log(absent+1) ~
+                       protection_var  +
+                       rain_var | oracle_number)
 
 
 final_model <- felm(log(absent+1) ~
@@ -505,7 +507,7 @@ final_model_radial_lm <- lm(log(absent+1) ~ #protection +
 
 # Model for each worker group
 worker_groups <- sort(unique(model_data$group))
-model_data_list <- model_list <- model_list_lm  <- 
+model_data_list <- model_list <- model_list_lm  <- model_list_fix <- 
   radial_model_data_list <- radial_model_list <- radial_model_list_lm <- list()
 for (i in 1:length(worker_groups)){
   sub_data <- model_data %>% filter(group == worker_groups[i])
@@ -513,6 +515,9 @@ for (i in 1:length(worker_groups)){
                        protection_var  +
                        rain_var | oracle_number | 0 | 0,
                      data = sub_data)
+  this_model_fix <- feols(data = sub_data, log(absent+1) ~
+                            protection_var  +
+                            rain_var | oracle_number)
   this_model_radial <- felm(log(absent+1) ~ #protection +
                               radial_herd_protection + 
                        rain_var | oracle_number | 0 | 0,
@@ -530,6 +535,7 @@ for (i in 1:length(worker_groups)){
   formatted_radial$group5 <- worker_groups[i]
   
   model_list[[i]] <- this_model
+  model_list_fix[[i]] <- this_model_fix
   model_list_lm[[i]] <- this_model_lm
   model_data_list[[i]] <- formatted
   
@@ -537,7 +543,7 @@ for (i in 1:length(worker_groups)){
   radial_model_list_lm[[i]] <- this_model_lm_radial
   radial_model_data_list[[i]] <- formatted_radial
 }
-names(model_list) <- names(model_data_list) <- names(model_list_lm) <-
+names(model_list) <- names(model_data_list) <- names(model_list_lm) <- names(model_list_fix) <- 
   names(radial_model_list) <- names(radial_model_data_list) <- names(radial_model_list_lm) <- 
   worker_groups
 
